### 2020-04-09 rR.tex
started writing

% Created by tikzDevice version 0.12.3 on 2020-04-09 18:48:49
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,433.62);
\begin{scope}
\path[clip] (  0.00,216.81) rectangle (289.08,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,216.81) rectangle (289.08,433.62);
\end{scope}
\begin{scope}
\path[clip] ( 27.31,247.50) rectangle (283.58,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31,247.50) rectangle (283.58,410.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31,247.68) --
	( 30.56,248.50) --
	( 33.80,249.34) --
	( 37.04,250.20) --
	( 40.29,251.09) --
	( 43.53,251.99) --
	( 46.78,252.92) --
	( 50.02,253.86) --
	( 53.26,254.83) --
	( 56.51,255.81) --
	( 59.75,256.82) --
	( 63.00,257.85) --
	( 66.24,258.90) --
	( 69.48,259.97) --
	( 72.73,261.06) --
	( 75.97,262.17) --
	( 79.22,263.30) --
	( 82.46,264.46) --
	( 85.70,265.64) --
	( 88.95,266.84) --
	( 92.19,268.05) --
	( 95.43,269.29) --
	( 98.68,270.54) --
	(101.92,271.81) --
	(105.17,273.11) --
	(108.41,274.42) --
	(111.65,275.76) --
	(114.90,277.13) --
	(118.14,278.52) --
	(121.39,279.93) --
	(124.63,281.37) --
	(127.87,282.82) --
	(131.12,284.29) --
	(134.36,285.79) --
	(137.61,287.30) --
	(140.85,288.84) --
	(144.09,290.41) --
	(147.34,292.01) --
	(150.58,293.64) --
	(153.82,295.30) --
	(157.07,296.99) --
	(160.31,298.71) --
	(163.56,300.44) --
	(166.80,302.19) --
	(170.04,303.97) --
	(173.29,305.78) --
	(176.53,307.62) --
	(179.78,309.49) --
	(183.02,311.40) --
	(186.26,313.35) --
	(189.51,315.33) --
	(192.75,317.36) --
	(196.00,319.41) --
	(199.24,321.50) --
	(202.48,323.61) --
	(205.73,325.76) --
	(208.97,327.95) --
	(212.21,330.17) --
	(215.46,332.42) --
	(218.70,334.72) --
	(221.95,337.06) --
	(225.19,339.44) --
	(228.43,341.86) --
	(231.68,344.31) --
	(234.92,346.79) --
	(238.17,349.31) --
	(241.41,351.87) --
	(244.65,354.47) --
	(247.90,357.09) --
	(251.14,359.76) --
	(254.39,362.46) --
	(257.63,365.21) --
	(260.87,367.98) --
	(264.12,370.80) --
	(267.36,373.65) --
	(270.60,376.54) --
	(273.85,379.47) --
	(277.09,382.43) --
	(280.34,385.43) --
	(283.58,388.46);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 27.31,247.65) --
	( 30.56,248.48) --
	( 33.80,249.33) --
	( 37.04,250.20) --
	( 40.29,251.09) --
	( 43.53,252.00) --
	( 46.78,252.94) --
	( 50.02,253.89) --
	( 53.26,254.86) --
	( 56.51,255.85) --
	( 59.75,256.87) --
	( 63.00,257.90) --
	( 66.24,258.95) --
	( 69.48,260.03) --
	( 72.73,261.12) --
	( 75.97,262.24) --
	( 79.22,263.38) --
	( 82.46,264.54) --
	( 85.70,265.72) --
	( 88.95,266.92) --
	( 92.19,268.14) --
	( 95.43,269.38) --
	( 98.68,270.63) --
	(101.92,271.91) --
	(105.17,273.21) --
	(108.41,274.52) --
	(111.65,275.86) --
	(114.90,277.23) --
	(118.14,278.62) --
	(121.39,280.03) --
	(124.63,281.47) --
	(127.87,282.92) --
	(131.12,284.38) --
	(134.36,285.87) --
	(137.61,287.38) --
	(140.85,288.91) --
	(144.09,290.47) --
	(147.34,292.06) --
	(150.58,293.68) --
	(153.82,295.34) --
	(157.07,297.03) --
	(160.31,298.74) --
	(163.56,300.47) --
	(166.80,302.22) --
	(170.04,304.00) --
	(173.29,305.80) --
	(176.53,307.64) --
	(179.78,309.51) --
	(183.02,311.41) --
	(186.26,313.36) --
	(189.51,315.34) --
	(192.75,317.37) --
	(196.00,319.42) --
	(199.24,321.50) --
	(202.48,323.61) --
	(205.73,325.76) --
	(208.97,327.94) --
	(212.21,330.16) --
	(215.46,332.42) --
	(218.70,334.72) --
	(221.95,337.06) --
	(225.19,339.44) --
	(228.43,341.86) --
	(231.68,344.32) --
	(234.92,346.81) --
	(238.17,349.33) --
	(241.41,351.90) --
	(244.65,354.49) --
	(247.90,357.13) --
	(251.14,359.80) --
	(254.39,362.51) --
	(257.63,365.26) --
	(260.87,368.04) --
	(264.12,370.87) --
	(267.36,373.73) --
	(270.60,376.62) --
	(273.85,379.56) --
	(277.09,382.53) --
	(280.34,385.54) --
	(283.58,388.58);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 27.31,247.64) --
	( 30.56,248.47) --
	( 33.80,249.33) --
	( 37.04,250.20) --
	( 40.29,251.09) --
	( 43.53,252.00) --
	( 46.78,252.94) --
	( 50.02,253.89) --
	( 53.26,254.86) --
	( 56.51,255.86) --
	( 59.75,256.87) --
	( 63.00,257.91) --
	( 66.24,258.96) --
	( 69.48,260.04) --
	( 72.73,261.13) --
	( 75.97,262.25) --
	( 79.22,263.39) --
	( 82.46,264.55) --
	( 85.70,265.73) --
	( 88.95,266.93) --
	( 92.19,268.15) --
	( 95.43,269.39) --
	( 98.68,270.65) --
	(101.92,271.92) --
	(105.17,273.22) --
	(108.41,274.54) --
	(111.65,275.88) --
	(114.90,277.24) --
	(118.14,278.63) --
	(121.39,280.04) --
	(124.63,281.48) --
	(127.87,282.92) --
	(131.12,284.39) --
	(134.36,285.87) --
	(137.61,287.38) --
	(140.85,288.91) --
	(144.09,290.47) --
	(147.34,292.05) --
	(150.58,293.67) --
	(153.82,295.33) --
	(157.07,297.02) --
	(160.31,298.73) --
	(163.56,300.45) --
	(166.80,302.20) --
	(170.04,303.97) --
	(173.29,305.78) --
	(176.53,307.61) --
	(179.78,309.47) --
	(183.02,311.38) --
	(186.26,313.32) --
	(189.51,315.31) --
	(192.75,317.33) --
	(196.00,319.38) --
	(199.24,321.47) --
	(202.48,323.58) --
	(205.73,325.73) --
	(208.97,327.92) --
	(212.21,330.14) --
	(215.46,332.40) --
	(218.70,334.70) --
	(221.95,337.05) --
	(225.19,339.44) --
	(228.43,341.86) --
	(231.68,344.33) --
	(234.92,346.82) --
	(238.17,349.36) --
	(241.41,351.93) --
	(244.65,354.54) --
	(247.90,357.18) --
	(251.14,359.87) --
	(254.39,362.59) --
	(257.63,365.35) --
	(260.87,368.14) --
	(264.12,370.98) --
	(267.36,373.85) --
	(270.60,376.76) --
	(273.85,379.71) --
	(277.09,382.70) --
	(280.34,385.72) --
	(283.58,388.79);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31,247.64) --
	( 30.56,248.47) --
	( 33.80,249.32) --
	( 37.04,250.19) --
	( 40.29,251.09) --
	( 43.53,252.00) --
	( 46.78,252.94) --
	( 50.02,253.89) --
	( 53.26,254.86) --
	( 56.51,255.86) --
	( 59.75,256.87) --
	( 63.00,257.91) --
	( 66.24,258.97) --
	( 69.48,260.04) --
	( 72.73,261.14) --
	( 75.97,262.26) --
	( 79.22,263.40) --
	( 82.46,264.56) --
	( 85.70,265.74) --
	( 88.95,266.94) --
	( 92.19,268.16) --
	( 95.43,269.40) --
	( 98.68,270.66) --
	(101.92,271.93) --
	(105.17,273.23) --
	(108.41,274.55) --
	(111.65,275.89) --
	(114.90,277.25) --
	(118.14,278.64) --
	(121.39,280.05) --
	(124.63,281.48) --
	(127.87,282.93) --
	(131.12,284.39) --
	(134.36,285.88) --
	(137.61,287.38) --
	(140.85,288.91) --
	(144.09,290.46) --
	(147.34,292.05) --
	(150.58,293.67) --
	(153.82,295.32) --
	(157.07,297.01) --
	(160.31,298.72) --
	(163.56,300.44) --
	(166.80,302.19) --
	(170.04,303.96) --
	(173.29,305.76) --
	(176.53,307.59) --
	(179.78,309.45) --
	(183.02,311.35) --
	(186.26,313.30) --
	(189.51,315.28) --
	(192.75,317.31) --
	(196.00,319.36) --
	(199.24,321.45) --
	(202.48,323.57) --
	(205.73,325.72) --
	(208.97,327.91) --
	(212.21,330.14) --
	(215.46,332.41) --
	(218.70,334.71) --
	(221.95,337.07) --
	(225.19,339.46) --
	(228.43,341.90) --
	(231.68,344.37) --
	(234.92,346.88) --
	(238.17,349.42) --
	(241.41,352.01) --
	(244.65,354.63) --
	(247.90,357.28) --
	(251.14,359.98) --
	(254.39,362.71) --
	(257.63,365.49) --
	(260.87,368.30) --
	(264.12,371.15) --
	(267.36,374.05) --
	(270.60,376.98) --
	(273.85,379.94) --
	(277.09,382.95) --
	(280.34,385.99) --
	(283.58,389.07);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 27.31,251.38) --
	( 30.68,245.55) --
	( 23.95,245.55) --
	cycle;

\path[fill=fillColor] ( 35.86,253.72) --
	( 39.22,247.89) --
	( 32.49,247.89) --
	cycle;

\path[fill=fillColor] ( 44.40,256.16) --
	( 47.76,250.34) --
	( 41.03,250.34) --
	cycle;

\path[fill=fillColor] ( 52.94,258.73) --
	( 56.30,252.90) --
	( 49.58,252.90) --
	cycle;

\path[fill=fillColor] ( 61.48,261.41) --
	( 64.85,255.58) --
	( 58.12,255.58) --
	cycle;

\path[fill=fillColor] ( 70.02,264.21) --
	( 73.39,258.38) --
	( 66.66,258.38) --
	cycle;

\path[fill=fillColor] ( 78.57,267.14) --
	( 81.93,261.31) --
	( 75.20,261.31) --
	cycle;

\path[fill=fillColor] ( 87.11,270.19) --
	( 90.47,264.37) --
	( 83.75,264.37) --
	cycle;

\path[fill=fillColor] ( 95.65,273.40) --
	( 99.01,267.58) --
	( 92.29,267.58) --
	cycle;

\path[fill=fillColor] (104.19,276.77) --
	(107.56,270.94) --
	(100.83,270.94) --
	cycle;

\path[fill=fillColor] (112.74,280.26) --
	(116.10,274.43) --
	(109.37,274.43) --
	cycle;

\path[fill=fillColor] (121.28,283.89) --
	(124.64,278.06) --
	(117.91,278.06) --
	cycle;

\path[fill=fillColor] (129.82,287.67) --
	(133.18,281.85) --
	(126.46,281.85) --
	cycle;

\path[fill=fillColor] (138.36,291.63) --
	(141.73,285.80) --
	(135.00,285.80) --
	cycle;

\path[fill=fillColor] (146.90,295.78) --
	(150.27,289.96) --
	(143.54,289.96) --
	cycle;

\path[fill=fillColor] (155.45,300.06) --
	(158.81,294.24) --
	(152.08,294.24) --
	cycle;

\path[fill=fillColor] (163.99,304.54) --
	(167.35,298.72) --
	(160.63,298.72) --
	cycle;

\path[fill=fillColor] (172.53,309.20) --
	(175.89,303.38) --
	(169.17,303.38) --
	cycle;

\path[fill=fillColor] (181.07,314.20) --
	(184.44,308.38) --
	(177.71,308.38) --
	cycle;

\path[fill=fillColor] (189.62,319.33) --
	(192.98,313.51) --
	(186.25,313.51) --
	cycle;

\path[fill=fillColor] (198.16,324.68) --
	(201.52,318.86) --
	(194.79,318.86) --
	cycle;

\path[fill=fillColor] (206.70,330.22) --
	(210.06,324.39) --
	(203.34,324.39) --
	cycle;

\path[fill=fillColor] (215.24,336.05) --
	(218.61,330.22) --
	(211.88,330.22) --
	cycle;

\path[fill=fillColor] (223.78,342.17) --
	(227.15,336.34) --
	(220.42,336.34) --
	cycle;

\path[fill=fillColor] (232.33,348.47) --
	(235.69,342.65) --
	(228.96,342.65) --
	cycle;

\path[fill=fillColor] (240.87,355.10) --
	(244.23,349.27) --
	(237.51,349.27) --
	cycle;

\path[fill=fillColor] (249.41,361.97) --
	(252.77,356.14) --
	(246.05,356.14) --
	cycle;

\path[fill=fillColor] (257.95,369.32) --
	(261.32,363.50) --
	(254.59,363.50) --
	cycle;

\path[fill=fillColor] (266.50,376.85) --
	(269.86,371.02) --
	(263.13,371.02) --
	cycle;

\path[fill=fillColor] (275.04,384.53) --
	(278.40,378.70) --
	(271.67,378.70) --
	cycle;

\path[fill=fillColor] (283.58,392.80) --
	(286.94,386.97) --
	(280.22,386.97) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 24.82,245.00) --
	( 29.81,245.00) --
	( 29.81,249.99) --
	( 24.82,249.99) --
	cycle;

\path[fill=fillColor] ( 33.36,247.34) --
	( 38.35,247.34) --
	( 38.35,252.33) --
	( 33.36,252.33) --
	cycle;

\path[fill=fillColor] ( 41.90,249.78) --
	( 46.90,249.78) --
	( 46.90,254.78) --
	( 41.90,254.78) --
	cycle;

\path[fill=fillColor] ( 50.44,252.35) --
	( 55.44,252.35) --
	( 55.44,257.34) --
	( 50.44,257.34) --
	cycle;

\path[fill=fillColor] ( 58.98,255.03) --
	( 63.98,255.03) --
	( 63.98,260.02) --
	( 58.98,260.02) --
	cycle;

\path[fill=fillColor] ( 67.53,257.83) --
	( 72.52,257.83) --
	( 72.52,262.83) --
	( 67.53,262.83) --
	cycle;

\path[fill=fillColor] ( 76.07,260.76) --
	( 81.06,260.76) --
	( 81.06,265.75) --
	( 76.07,265.75) --
	cycle;

\path[fill=fillColor] ( 84.61,263.82) --
	( 89.61,263.82) --
	( 89.61,268.81) --
	( 84.61,268.81) --
	cycle;

\path[fill=fillColor] ( 93.15,267.03) --
	( 98.15,267.03) --
	( 98.15,272.03) --
	( 93.15,272.03) --
	cycle;

\path[fill=fillColor] (101.70,270.41) --
	(106.69,270.41) --
	(106.69,275.40) --
	(101.70,275.40) --
	cycle;

\path[fill=fillColor] (110.24,273.89) --
	(115.23,273.89) --
	(115.23,278.89) --
	(110.24,278.89) --
	cycle;

\path[fill=fillColor] (118.78,277.52) --
	(123.78,277.52) --
	(123.78,282.51) --
	(118.78,282.51) --
	cycle;

\path[fill=fillColor] (127.32,281.30) --
	(132.32,281.30) --
	(132.32,286.30) --
	(127.32,286.30) --
	cycle;

\path[fill=fillColor] (135.86,285.26) --
	(140.86,285.26) --
	(140.86,290.25) --
	(135.86,290.25) --
	cycle;

\path[fill=fillColor] (144.41,289.41) --
	(149.40,289.41) --
	(149.40,294.41) --
	(144.41,294.41) --
	cycle;

\path[fill=fillColor] (152.95,293.67) --
	(157.94,293.67) --
	(157.94,298.67) --
	(152.95,298.67) --
	cycle;

\path[fill=fillColor] (161.49,298.13) --
	(166.49,298.13) --
	(166.49,303.13) --
	(161.49,303.13) --
	cycle;

\path[fill=fillColor] (170.03,302.76) --
	(175.03,302.76) --
	(175.03,307.76) --
	(170.03,307.76) --
	cycle;

\path[fill=fillColor] (178.58,307.80) --
	(183.57,307.80) --
	(183.57,312.80) --
	(178.58,312.80) --
	cycle;

\path[fill=fillColor] (187.12,312.93) --
	(192.11,312.93) --
	(192.11,317.93) --
	(187.12,317.93) --
	cycle;

\path[fill=fillColor] (195.66,318.29) --
	(200.66,318.29) --
	(200.66,323.28) --
	(195.66,323.28) --
	cycle;

\path[fill=fillColor] (204.20,323.80) --
	(209.20,323.80) --
	(209.20,328.79) --
	(204.20,328.79) --
	cycle;

\path[fill=fillColor] (212.74,329.65) --
	(217.74,329.65) --
	(217.74,334.64) --
	(212.74,334.64) --
	cycle;

\path[fill=fillColor] (221.29,335.78) --
	(226.28,335.78) --
	(226.28,340.78) --
	(221.29,340.78) --
	cycle;

\path[fill=fillColor] (229.83,342.05) --
	(234.82,342.05) --
	(234.82,347.05) --
	(229.83,347.05) --
	cycle;

\path[fill=fillColor] (238.37,348.70) --
	(243.37,348.70) --
	(243.37,353.69) --
	(238.37,353.69) --
	cycle;

\path[fill=fillColor] (246.91,355.61) --
	(251.91,355.61) --
	(251.91,360.60) --
	(246.91,360.60) --
	cycle;

\path[fill=fillColor] (255.46,363.10) --
	(260.45,363.10) --
	(260.45,368.10) --
	(255.46,368.10) --
	cycle;

\path[fill=fillColor] (264.00,370.65) --
	(268.99,370.65) --
	(268.99,375.65) --
	(264.00,375.65) --
	cycle;

\path[fill=fillColor] (272.54,378.27) --
	(277.54,378.27) --
	(277.54,383.27) --
	(272.54,383.27) --
	cycle;

\path[fill=fillColor] (281.08,386.62) --
	(286.08,386.62) --
	(286.08,391.61) --
	(281.08,391.61) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 23.78,247.50) -- ( 30.85,247.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.31,243.96) -- ( 27.31,251.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.32,249.83) -- ( 39.39,249.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.86,246.30) -- ( 35.86,253.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87,252.27) -- ( 47.93,252.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.40,248.74) -- ( 44.40,255.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.41,254.84) -- ( 56.47,254.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.94,251.31) -- ( 52.94,258.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.95,257.52) -- ( 65.01,257.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.48,253.99) -- ( 61.48,261.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.49,260.33) -- ( 73.56,260.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.02,256.79) -- ( 70.02,263.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.03,263.25) -- ( 82.10,263.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.57,259.72) -- ( 78.57,266.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.58,266.32) -- ( 90.64,266.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.11,262.78) -- ( 87.11,269.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.12,269.54) -- ( 99.18,269.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65,266.00) -- ( 95.65,273.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.66,272.92) -- (107.73,272.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.19,269.39) -- (104.19,276.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.20,276.41) -- (116.27,276.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.74,272.87) -- (112.74,279.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.75,280.03) -- (124.81,280.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.28,276.49) -- (121.28,283.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.29,283.81) -- (133.35,283.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.82,280.27) -- (129.82,287.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.83,287.76) -- (141.89,287.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.36,284.23) -- (138.36,291.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.37,291.93) -- (150.44,291.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.90,288.39) -- (146.90,295.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.91,296.17) -- (158.98,296.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (155.45,292.64) -- (155.45,299.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.46,300.60) -- (167.52,300.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.99,297.07) -- (163.99,304.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.00,305.21) -- (176.06,305.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.53,301.68) -- (172.53,308.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.54,310.29) -- (184.61,310.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.07,306.76) -- (181.07,313.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.08,315.42) -- (193.15,315.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.62,311.89) -- (189.62,318.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.63,320.79) -- (201.69,320.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.16,317.26) -- (198.16,324.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.17,326.28) -- (210.23,326.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.70,322.74) -- (206.70,329.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.71,332.15) -- (218.77,332.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.24,328.62) -- (215.24,335.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.25,338.30) -- (227.32,338.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.78,334.76) -- (223.78,341.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.79,344.53) -- (235.86,344.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.33,341.00) -- (232.33,348.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.34,351.21) -- (244.40,351.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.87,347.68) -- (240.87,354.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.88,358.17) -- (252.94,358.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.41,354.64) -- (249.41,361.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.42,365.83) -- (261.49,365.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.95,362.30) -- (257.95,369.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.96,373.41) -- (270.03,373.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (266.50,369.87) -- (266.50,376.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.51,380.97) -- (278.57,380.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.04,377.44) -- (275.04,384.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.05,389.39) -- (287.11,389.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.58,385.85) -- (283.58,392.92);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 27.31,247.50) circle (  2.50);

\path[fill=fillColor] ( 35.86,249.83) circle (  2.50);

\path[fill=fillColor] ( 44.40,252.27) circle (  2.50);

\path[fill=fillColor] ( 52.94,254.82) circle (  2.50);

\path[fill=fillColor] ( 61.48,257.51) circle (  2.50);

\path[fill=fillColor] ( 70.02,260.28) circle (  2.50);

\path[fill=fillColor] ( 78.57,263.20) circle (  2.50);

\path[fill=fillColor] ( 87.11,266.27) circle (  2.50);

\path[fill=fillColor] ( 95.65,269.43) circle (  2.50);

\path[fill=fillColor] (104.19,272.79) circle (  2.50);

\path[fill=fillColor] (112.74,276.24) circle (  2.50);

\path[fill=fillColor] (121.28,279.86) circle (  2.50);

\path[fill=fillColor] (129.82,283.67) circle (  2.50);

\path[fill=fillColor] (138.36,287.69) circle (  2.50);

\path[fill=fillColor] (146.90,291.76) circle (  2.50);

\path[fill=fillColor] (155.45,296.17) circle (  2.50);

\path[fill=fillColor] (163.99,300.71) circle (  2.50);

\path[fill=fillColor] (172.53,305.39) circle (  2.50);

\path[fill=fillColor] (181.07,310.31) circle (  2.50);

\path[fill=fillColor] (189.62,315.41) circle (  2.50);

\path[fill=fillColor] (198.16,320.80) circle (  2.50);

\path[fill=fillColor] (206.70,326.25) circle (  2.50);

\path[fill=fillColor] (215.24,332.23) circle (  2.50);

\path[fill=fillColor] (223.78,338.13) circle (  2.50);

\path[fill=fillColor] (232.33,344.44) circle (  2.50);

\path[fill=fillColor] (240.87,351.38) circle (  2.50);

\path[fill=fillColor] (249.41,358.30) circle (  2.50);

\path[fill=fillColor] (257.95,365.26) circle (  2.50);

\path[fill=fillColor] (266.50,372.85) circle (  2.50);

\path[fill=fillColor] (275.04,380.60) circle (  2.50);

\path[fill=fillColor] (283.58,388.73) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31,247.50) rectangle (283.58,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,247.50) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,294.20) --
	( 27.31,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,340.91) --
	( 27.31,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,387.61) --
	( 27.31,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,244.75) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.74,244.75) --
	(112.74,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.16,244.75) --
	(198.16,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.74,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.16,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (155.45,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,329.23) {Reproduction number $\mathcal R$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 39.98,332.94) rectangle (168.41,407.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,381.80) rectangle ( 59.93,396.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 46.93,389.03) -- ( 58.49,389.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 52.71,389.03) circle (  2.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,367.35) rectangle ( 59.93,381.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 46.93,374.57) -- ( 58.49,374.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 52.71,378.46) --
	( 56.07,372.63) --
	( 49.34,372.63) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,352.89) rectangle ( 59.93,367.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 46.93,360.12) -- ( 58.49,360.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 50.21,357.62) --
	( 55.20,357.62) --
	( 55.20,362.62) --
	( 50.21,362.62) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,338.44) rectangle ( 59.93,352.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 46.93,345.67) -- ( 58.49,345.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.17,345.67) -- ( 56.24,345.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.71,342.13) -- ( 52.71,349.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,386.00) {Generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,371.54) {Serial interval ($\rho=0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,357.09) {Serial interval ($\rho=0$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,342.63) {Serial interval ($\rho=-0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,419.03) {A};
\end{scope}
\begin{scope}
\path[clip] (289.08,216.81) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,216.81) rectangle (578.16,433.62);
\end{scope}
\begin{scope}
\path[clip] (316.39,247.50) rectangle (572.66,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (316.39,247.50) rectangle (572.66,410.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (316.39,247.68) --
	(319.64,248.50) --
	(322.88,249.34) --
	(326.12,250.20) --
	(329.37,251.09) --
	(332.61,251.99) --
	(335.86,252.92) --
	(339.10,253.86) --
	(342.34,254.83) --
	(345.59,255.81) --
	(348.83,256.82) --
	(352.08,257.85) --
	(355.32,258.90) --
	(358.56,259.97) --
	(361.81,261.06) --
	(365.05,262.17) --
	(368.30,263.30) --
	(371.54,264.46) --
	(374.78,265.64) --
	(378.03,266.84) --
	(381.27,268.05) --
	(384.51,269.29) --
	(387.76,270.54) --
	(391.00,271.81) --
	(394.25,273.10) --
	(397.49,274.42) --
	(400.73,275.76) --
	(403.98,277.12) --
	(407.22,278.51) --
	(410.47,279.93) --
	(413.71,281.36) --
	(416.95,282.81) --
	(420.20,284.29) --
	(423.44,285.78) --
	(426.69,287.29) --
	(429.93,288.83) --
	(433.17,290.40) --
	(436.42,292.00) --
	(439.66,293.63) --
	(442.90,295.29) --
	(446.15,296.98) --
	(449.39,298.70) --
	(452.64,300.44) --
	(455.88,302.19) --
	(459.12,303.98) --
	(462.37,305.79) --
	(465.61,307.63) --
	(468.86,309.51) --
	(472.10,311.42) --
	(475.34,313.37) --
	(478.59,315.36) --
	(481.83,317.38) --
	(485.08,319.43) --
	(488.32,321.51) --
	(491.56,323.63) --
	(494.81,325.77) --
	(498.05,327.95) --
	(501.29,330.17) --
	(504.54,332.42) --
	(507.78,334.72) --
	(511.03,337.06) --
	(514.27,339.44) --
	(517.51,341.85) --
	(520.76,344.31) --
	(524.00,346.79) --
	(527.25,349.31) --
	(530.49,351.87) --
	(533.73,354.47) --
	(536.98,357.10) --
	(540.22,359.77) --
	(543.47,362.47) --
	(546.71,365.21) --
	(549.95,367.99) --
	(553.20,370.81) --
	(556.44,373.67) --
	(559.68,376.56) --
	(562.93,379.49) --
	(566.17,382.45) --
	(569.42,385.45) --
	(572.66,388.49);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (319.64,248.28) --
	(322.88,249.19) --
	(326.12,250.12) --
	(329.37,251.04) --
	(332.61,251.98) --
	(335.86,252.92) --
	(339.10,253.86) --
	(342.34,254.82) --
	(345.59,255.77) --
	(348.83,256.74) --
	(352.08,257.71) --
	(355.32,258.69) --
	(358.56,259.67) --
	(361.81,260.66) --
	(365.05,261.65) --
	(368.30,262.65) --
	(371.54,263.65) --
	(374.78,264.66) --
	(378.03,265.68) --
	(381.27,266.70) --
	(384.51,267.73) --
	(387.76,268.77) --
	(391.00,269.81) --
	(394.25,270.86) --
	(397.49,271.92) --
	(400.73,272.98) --
	(403.98,274.04) --
	(407.22,275.11) --
	(410.47,276.18) --
	(413.71,277.25) --
	(416.95,278.34) --
	(420.20,279.43) --
	(423.44,280.54) --
	(426.69,281.64) --
	(429.93,282.75) --
	(433.17,283.86) --
	(436.42,284.97) --
	(439.66,286.07) --
	(442.90,287.16) --
	(446.15,288.24) --
	(449.39,289.34) --
	(452.64,290.44) --
	(455.88,291.56) --
	(459.12,292.67) --
	(462.37,293.78) --
	(465.61,294.89) --
	(468.86,295.99) --
	(472.10,297.07) --
	(475.34,298.14) --
	(478.59,299.19) --
	(481.83,300.23) --
	(485.08,301.27) --
	(488.32,302.30) --
	(491.56,303.33) --
	(494.81,304.35) --
	(498.05,305.36) --
	(501.29,306.36) --
	(504.54,307.33) --
	(507.78,308.29) --
	(511.03,309.23) --
	(514.27,310.14) --
	(517.51,311.04) --
	(520.76,311.93) --
	(524.00,312.80) --
	(527.25,313.67) --
	(530.49,314.51) --
	(533.73,315.35) --
	(536.98,316.17) --
	(540.22,316.98) --
	(543.47,317.77) --
	(546.71,318.55) --
	(549.95,319.31) --
	(553.20,320.05) --
	(556.44,320.78) --
	(559.68,321.50) --
	(562.93,322.20) --
	(566.17,322.89) --
	(569.42,323.56) --
	(572.66,324.22);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (319.64,248.18) --
	(322.88,249.13) --
	(326.12,250.07) --
	(329.37,251.02) --
	(332.61,251.97) --
	(335.86,252.91) --
	(339.10,253.86) --
	(342.34,254.80) --
	(345.59,255.75) --
	(348.83,256.69) --
	(352.08,257.64) --
	(355.32,258.59) --
	(358.56,259.53) --
	(361.81,260.48) --
	(365.05,261.42) --
	(368.30,262.36) --
	(371.54,263.30) --
	(374.78,264.24) --
	(378.03,265.18) --
	(381.27,266.12) --
	(384.51,267.07) --
	(387.76,268.02) --
	(391.00,268.97) --
	(394.25,269.92) --
	(397.49,270.86) --
	(400.73,271.81) --
	(403.98,272.75) --
	(407.22,273.68) --
	(410.47,274.60) --
	(413.71,275.53) --
	(416.95,276.46) --
	(420.20,277.40) --
	(423.44,278.34) --
	(426.69,279.28) --
	(429.93,280.21) --
	(433.17,281.13) --
	(436.42,282.03) --
	(439.66,282.92) --
	(442.90,283.77) --
	(446.15,284.61) --
	(449.39,285.45) --
	(452.64,286.29) --
	(455.88,287.13) --
	(459.12,287.97) --
	(462.37,288.79) --
	(465.61,289.59) --
	(468.86,290.36) --
	(472.10,291.11) --
	(475.34,291.82) --
	(478.59,292.49) --
	(481.83,293.12) --
	(485.08,293.74) --
	(488.32,294.35) --
	(491.56,294.93) --
	(494.81,295.49) --
	(498.05,296.02) --
	(501.29,296.52) --
	(504.54,296.98) --
	(507.78,297.40) --
	(511.03,297.78) --
	(514.27,298.10) --
	(517.51,298.40) --
	(520.76,298.67) --
	(524.00,298.90) --
	(527.25,299.11) --
	(530.49,299.28) --
	(533.73,299.42) --
	(536.98,299.53) --
	(540.22,299.60) --
	(543.47,299.64) --
	(546.71,299.65) --
	(549.95,299.62) --
	(553.20,299.56) --
	(556.44,299.46) --
	(559.68,299.32) --
	(562.93,299.16) --
	(566.17,298.96) --
	(569.42,298.72) --
	(572.66,298.46);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (319.64,248.11) --
	(322.88,249.08) --
	(326.12,250.04) --
	(329.37,251.00) --
	(332.61,251.95) --
	(335.86,252.90) --
	(339.10,253.84) --
	(342.34,254.78) --
	(345.59,255.71) --
	(348.83,256.63) --
	(352.08,257.56) --
	(355.32,258.47) --
	(358.56,259.38) --
	(361.81,260.28) --
	(365.05,261.18) --
	(368.30,262.07) --
	(371.54,262.95) --
	(374.78,263.83) --
	(378.03,264.69) --
	(381.27,265.56) --
	(384.51,266.43) --
	(387.76,267.30) --
	(391.00,268.16) --
	(394.25,269.02) --
	(397.49,269.87) --
	(400.73,270.71) --
	(403.98,271.54) --
	(407.22,272.35) --
	(410.47,273.15) --
	(413.71,273.94) --
	(416.95,274.74) --
	(420.20,275.54) --
	(423.44,276.33) --
	(426.69,277.12) --
	(429.93,277.89) --
	(433.17,278.64) --
	(436.42,279.36) --
	(439.66,280.06) --
	(442.90,280.71) --
	(446.15,281.34) --
	(449.39,281.96) --
	(452.64,282.57) --
	(455.88,283.18) --
	(459.12,283.78) --
	(462.37,284.35) --
	(465.61,284.89) --
	(468.86,285.40) --
	(472.10,285.86) --
	(475.34,286.28) --
	(478.59,286.64) --
	(481.83,286.95) --
	(485.08,287.25) --
	(488.32,287.51) --
	(491.56,287.75) --
	(494.81,287.95) --
	(498.05,288.11) --
	(501.29,288.23) --
	(504.54,288.30) --
	(507.78,288.32) --
	(511.03,288.28) --
	(514.27,288.19) --
	(517.51,288.05) --
	(520.76,287.88) --
	(524.00,287.67) --
	(527.25,287.41) --
	(530.49,287.11) --
	(533.73,286.77) --
	(536.98,286.39) --
	(540.22,285.96) --
	(543.47,285.49) --
	(546.71,284.98) --
	(549.95,284.41) --
	(553.20,283.81) --
	(556.44,283.16) --
	(559.68,282.46) --
	(562.93,281.72) --
	(566.17,280.94) --
	(569.42,280.11) --
	(572.66,279.25);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] (316.39,251.38) --
	(319.76,245.55) --
	(313.03,245.55) --
	cycle;

\path[fill=fillColor] (324.94,253.70) --
	(328.30,247.88) --
	(321.57,247.88) --
	cycle;

\path[fill=fillColor] (333.48,256.09) --
	(336.84,250.27) --
	(330.11,250.27) --
	cycle;

\path[fill=fillColor] (342.02,258.54) --
	(345.38,252.71) --
	(338.66,252.71) --
	cycle;

\path[fill=fillColor] (350.56,261.05) --
	(353.93,255.22) --
	(347.20,255.22) --
	cycle;

\path[fill=fillColor] (359.10,263.61) --
	(362.47,257.79) --
	(355.74,257.79) --
	cycle;

\path[fill=fillColor] (367.65,266.23) --
	(371.01,260.40) --
	(364.28,260.40) --
	cycle;

\path[fill=fillColor] (376.19,268.90) --
	(379.55,263.07) --
	(372.83,263.07) --
	cycle;

\path[fill=fillColor] (384.73,271.62) --
	(388.09,265.79) --
	(381.37,265.79) --
	cycle;

\path[fill=fillColor] (393.27,274.38) --
	(396.64,268.55) --
	(389.91,268.55) --
	cycle;

\path[fill=fillColor] (401.82,277.18) --
	(405.18,271.35) --
	(398.45,271.35) --
	cycle;

\path[fill=fillColor] (410.36,280.01) --
	(413.72,274.19) --
	(406.99,274.19) --
	cycle;

\path[fill=fillColor] (418.90,282.88) --
	(422.26,277.05) --
	(415.54,277.05) --
	cycle;

\path[fill=fillColor] (427.44,285.76) --
	(430.81,279.94) --
	(424.08,279.94) --
	cycle;

\path[fill=fillColor] (435.98,288.66) --
	(439.35,282.84) --
	(432.62,282.84) --
	cycle;

\path[fill=fillColor] (444.53,291.57) --
	(447.89,285.74) --
	(441.16,285.74) --
	cycle;

\path[fill=fillColor] (453.07,294.48) --
	(456.43,288.65) --
	(449.71,288.65) --
	cycle;

\path[fill=fillColor] (461.61,297.37) --
	(464.97,291.55) --
	(458.25,291.55) --
	cycle;

\path[fill=fillColor] (470.15,300.25) --
	(473.52,294.42) --
	(466.79,294.42) --
	cycle;

\path[fill=fillColor] (478.70,303.09) --
	(482.06,297.26) --
	(475.33,297.26) --
	cycle;

\path[fill=fillColor] (487.24,305.88) --
	(490.60,300.06) --
	(483.87,300.06) --
	cycle;

\path[fill=fillColor] (495.78,308.62) --
	(499.14,302.79) --
	(492.42,302.79) --
	cycle;

\path[fill=fillColor] (504.32,311.28) --
	(507.69,305.46) --
	(500.96,305.46) --
	cycle;

\path[fill=fillColor] (512.86,313.85) --
	(516.23,308.03) --
	(509.50,308.03) --
	cycle;

\path[fill=fillColor] (521.41,316.31) --
	(524.77,310.49) --
	(518.04,310.49) --
	cycle;

\path[fill=fillColor] (529.95,318.64) --
	(533.31,312.82) --
	(526.59,312.82) --
	cycle;

\path[fill=fillColor] (538.49,320.82) --
	(541.85,314.99) --
	(535.13,314.99) --
	cycle;

\path[fill=fillColor] (547.03,322.82) --
	(550.40,317.00) --
	(543.67,317.00) --
	cycle;

\path[fill=fillColor] (555.58,324.63) --
	(558.94,318.80) --
	(552.21,318.80) --
	cycle;

\path[fill=fillColor] (564.12,326.21) --
	(567.48,320.38) --
	(560.75,320.38) --
	cycle;

\path[fill=fillColor] (572.66,327.53) --
	(576.02,321.70) --
	(569.30,321.70) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] (313.90,245.00) --
	(318.89,245.00) --
	(318.89,249.99) --
	(313.90,249.99) --
	cycle;

\path[fill=fillColor] (322.44,247.31) --
	(327.43,247.31) --
	(327.43,252.31) --
	(322.44,252.31) --
	cycle;

\path[fill=fillColor] (330.98,249.68) --
	(335.98,249.68) --
	(335.98,254.67) --
	(330.98,254.67) --
	cycle;

\path[fill=fillColor] (339.52,252.08) --
	(344.52,252.08) --
	(344.52,257.07) --
	(339.52,257.07) --
	cycle;

\path[fill=fillColor] (348.06,254.52) --
	(353.06,254.52) --
	(353.06,259.51) --
	(348.06,259.51) --
	cycle;

\path[fill=fillColor] (356.61,256.99) --
	(361.60,256.99) --
	(361.60,261.98) --
	(356.61,261.98) --
	cycle;

\path[fill=fillColor] (365.15,259.48) --
	(370.14,259.48) --
	(370.14,264.47) --
	(365.15,264.47) --
	cycle;

\path[fill=fillColor] (373.69,261.99) --
	(378.69,261.99) --
	(378.69,266.99) --
	(373.69,266.99) --
	cycle;

\path[fill=fillColor] (382.23,264.51) --
	(387.23,264.51) --
	(387.23,269.51) --
	(382.23,269.51) --
	cycle;

\path[fill=fillColor] (390.78,267.04) --
	(395.77,267.04) --
	(395.77,272.03) --
	(390.78,272.03) --
	cycle;

\path[fill=fillColor] (399.32,269.56) --
	(404.31,269.56) --
	(404.31,274.55) --
	(399.32,274.55) --
	cycle;

\path[fill=fillColor] (407.86,272.06) --
	(412.86,272.06) --
	(412.86,277.05) --
	(407.86,277.05) --
	cycle;

\path[fill=fillColor] (416.40,274.53) --
	(421.40,274.53) --
	(421.40,279.53) --
	(416.40,279.53) --
	cycle;

\path[fill=fillColor] (424.94,276.97) --
	(429.94,276.97) --
	(429.94,281.96) --
	(424.94,281.96) --
	cycle;

\path[fill=fillColor] (433.49,279.35) --
	(438.48,279.35) --
	(438.48,284.35) --
	(433.49,284.35) --
	cycle;

\path[fill=fillColor] (442.03,281.67) --
	(447.02,281.67) --
	(447.02,286.67) --
	(442.03,286.67) --
	cycle;

\path[fill=fillColor] (450.57,283.91) --
	(455.57,283.91) --
	(455.57,288.91) --
	(450.57,288.91) --
	cycle;

\path[fill=fillColor] (459.11,286.06) --
	(464.11,286.06) --
	(464.11,291.05) --
	(459.11,291.05) --
	cycle;

\path[fill=fillColor] (467.66,288.09) --
	(472.65,288.09) --
	(472.65,293.08) --
	(467.66,293.08) --
	cycle;

\path[fill=fillColor] (476.20,289.98) --
	(481.19,289.98) --
	(481.19,294.98) --
	(476.20,294.98) --
	cycle;

\path[fill=fillColor] (484.74,291.73) --
	(489.74,291.73) --
	(489.74,296.72) --
	(484.74,296.72) --
	cycle;

\path[fill=fillColor] (493.28,293.30) --
	(498.28,293.30) --
	(498.28,298.29) --
	(493.28,298.29) --
	cycle;

\path[fill=fillColor] (501.82,294.66) --
	(506.82,294.66) --
	(506.82,299.66) --
	(501.82,299.66) --
	cycle;

\path[fill=fillColor] (510.37,295.81) --
	(515.36,295.81) --
	(515.36,300.80) --
	(510.37,300.80) --
	cycle;

\path[fill=fillColor] (518.91,296.70) --
	(523.90,296.70) --
	(523.90,301.70) --
	(518.91,301.70) --
	cycle;

\path[fill=fillColor] (527.45,297.31) --
	(532.45,297.31) --
	(532.45,302.31) --
	(527.45,302.31) --
	cycle;

\path[fill=fillColor] (535.99,297.62) --
	(540.99,297.62) --
	(540.99,302.61) --
	(535.99,302.61) --
	cycle;

\path[fill=fillColor] (544.54,297.58) --
	(549.53,297.58) --
	(549.53,302.58) --
	(544.54,302.58) --
	cycle;

\path[fill=fillColor] (553.08,297.18) --
	(558.07,297.18) --
	(558.07,302.18) --
	(553.08,302.18) --
	cycle;

\path[fill=fillColor] (561.62,296.38) --
	(566.62,296.38) --
	(566.62,301.38) --
	(561.62,301.38) --
	cycle;

\path[fill=fillColor] (570.16,295.16) --
	(575.16,295.16) --
	(575.16,300.16) --
	(570.16,300.16) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.86,247.50) -- (319.93,247.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.39,243.96) -- (316.39,251.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.40,249.80) -- (328.47,249.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (324.94,246.27) -- (324.94,253.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.95,252.14) -- (337.01,252.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.48,248.61) -- (333.48,255.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.49,254.50) -- (345.55,254.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.02,250.97) -- (342.02,258.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (347.03,256.87) -- (354.09,256.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.56,253.34) -- (350.56,260.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (355.57,259.25) -- (362.64,259.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.10,255.72) -- (359.10,262.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.11,261.63) -- (371.18,261.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.65,258.10) -- (367.65,265.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.66,263.99) -- (379.72,263.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.19,260.46) -- (376.19,267.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.20,266.33) -- (388.26,266.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.73,262.80) -- (384.73,269.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.74,268.64) -- (396.81,268.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.27,265.11) -- (393.27,272.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.28,270.90) -- (405.35,270.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.82,267.37) -- (401.82,274.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.83,273.10) -- (413.89,273.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.36,269.57) -- (410.36,276.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (415.37,275.23) -- (422.43,275.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.90,271.70) -- (418.90,278.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.91,277.27) -- (430.97,277.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.44,273.73) -- (427.44,280.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (432.45,279.20) -- (439.52,279.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.98,275.67) -- (435.98,282.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (440.99,281.01) -- (448.06,281.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.53,277.48) -- (444.53,284.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (449.54,282.68) -- (456.60,282.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.07,279.14) -- (453.07,286.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.08,284.18) -- (465.14,284.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.61,280.65) -- (461.61,287.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (466.62,285.51) -- (473.69,285.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.15,281.98) -- (470.15,289.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.16,286.63) -- (482.23,286.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.70,283.10) -- (478.70,290.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.71,287.53) -- (490.77,287.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.24,283.99) -- (487.24,291.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (492.25,288.17) -- (499.31,288.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.78,284.64) -- (495.78,291.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.79,288.54) -- (507.85,288.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (504.32,285.00) -- (504.32,292.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (509.33,288.60) -- (516.40,288.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (512.86,285.07) -- (512.86,292.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (517.87,288.33) -- (524.94,288.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (521.41,284.80) -- (521.41,291.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.42,287.71) -- (533.48,287.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (529.95,284.18) -- (529.95,291.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (534.96,286.71) -- (542.02,286.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (538.49,283.17) -- (538.49,290.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (543.50,285.30) -- (550.57,285.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.03,281.77) -- (547.03,288.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.04,283.48) -- (559.11,283.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (555.58,279.95) -- (555.58,287.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.59,281.22) -- (567.65,281.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.12,277.69) -- (564.12,284.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13,278.53) -- (576.19,278.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66,275.00) -- (572.66,282.06);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (316.39,247.50) circle (  2.50);

\path[fill=fillColor] (324.94,249.83) circle (  2.50);

\path[fill=fillColor] (333.48,252.27) circle (  2.50);

\path[fill=fillColor] (342.02,254.83) circle (  2.50);

\path[fill=fillColor] (350.56,257.49) circle (  2.50);

\path[fill=fillColor] (359.10,260.29) circle (  2.50);

\path[fill=fillColor] (367.65,263.21) circle (  2.50);

\path[fill=fillColor] (376.19,266.26) circle (  2.50);

\path[fill=fillColor] (384.73,269.43) circle (  2.50);

\path[fill=fillColor] (393.27,272.82) circle (  2.50);

\path[fill=fillColor] (401.82,276.23) circle (  2.50);

\path[fill=fillColor] (410.36,279.89) circle (  2.50);

\path[fill=fillColor] (418.90,283.67) circle (  2.50);

\path[fill=fillColor] (427.44,287.63) circle (  2.50);

\path[fill=fillColor] (435.98,291.78) circle (  2.50);

\path[fill=fillColor] (444.53,296.18) circle (  2.50);

\path[fill=fillColor] (453.07,300.57) circle (  2.50);

\path[fill=fillColor] (461.61,305.43) circle (  2.50);

\path[fill=fillColor] (470.15,310.38) circle (  2.50);

\path[fill=fillColor] (478.70,315.48) circle (  2.50);

\path[fill=fillColor] (487.24,320.76) circle (  2.50);

\path[fill=fillColor] (495.78,326.28) circle (  2.50);

\path[fill=fillColor] (504.32,332.33) circle (  2.50);

\path[fill=fillColor] (512.86,338.07) circle (  2.50);

\path[fill=fillColor] (521.41,344.75) circle (  2.50);

\path[fill=fillColor] (529.95,350.97) circle (  2.50);

\path[fill=fillColor] (538.49,358.06) circle (  2.50);

\path[fill=fillColor] (547.03,365.33) circle (  2.50);

\path[fill=fillColor] (555.58,372.97) circle (  2.50);

\path[fill=fillColor] (564.12,380.73) circle (  2.50);

\path[fill=fillColor] (572.66,388.73) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.39,247.50) rectangle (572.66,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,247.50) --
	(316.39,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,294.20) --
	(316.39,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,340.91) --
	(316.39,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,387.61) --
	(316.39,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.39,244.75) --
	(316.39,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.82,244.75) --
	(401.82,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.24,244.75) --
	(487.24,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (316.39,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (401.82,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (487.24,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (444.53,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,329.23) {Reproduction number $\mathcal R$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (316.39,419.03) {B};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 27.31, 30.69) rectangle (283.58,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31, 30.69) rectangle (283.58,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31, 47.92) --
	( 30.56, 47.94) --
	( 33.80, 47.96) --
	( 37.04, 47.98) --
	( 40.29, 48.00) --
	( 43.53, 48.02) --
	( 46.78, 48.03) --
	( 50.02, 48.05) --
	( 53.26, 48.06) --
	( 56.51, 48.07) --
	( 59.75, 48.09) --
	( 63.00, 48.09) --
	( 66.24, 48.10) --
	( 69.48, 48.11) --
	( 72.73, 48.12) --
	( 75.97, 48.12) --
	( 79.22, 48.13) --
	( 82.46, 48.13) --
	( 85.70, 48.14) --
	( 88.95, 48.14) --
	( 92.19, 48.14) --
	( 95.43, 48.14) --
	( 98.68, 48.14) --
	(101.92, 48.13) --
	(105.17, 48.12) --
	(108.41, 48.12) --
	(111.65, 48.11) --
	(114.90, 48.10) --
	(118.14, 48.10) --
	(121.39, 48.10) --
	(124.63, 48.11) --
	(127.87, 48.12) --
	(131.12, 48.13) --
	(134.36, 48.14) --
	(137.61, 48.16) --
	(140.85, 48.17) --
	(144.09, 48.19) --
	(147.34, 48.20) --
	(150.58, 48.21) --
	(153.82, 48.22) --
	(157.07, 48.22) --
	(160.31, 48.22) --
	(163.56, 48.22) --
	(166.80, 48.22) --
	(170.04, 48.22) --
	(173.29, 48.22) --
	(176.53, 48.22) --
	(179.78, 48.21) --
	(183.02, 48.21) --
	(186.26, 48.21) --
	(189.51, 48.20) --
	(192.75, 48.20) --
	(196.00, 48.20) --
	(199.24, 48.20) --
	(202.48, 48.19) --
	(205.73, 48.19) --
	(208.97, 48.19) --
	(212.21, 48.18) --
	(215.46, 48.18) --
	(218.70, 48.18) --
	(221.95, 48.18) --
	(225.19, 48.18) --
	(228.43, 48.18) --
	(231.68, 48.17) --
	(234.92, 48.17) --
	(238.17, 48.17) --
	(241.41, 48.17) --
	(244.65, 48.17) --
	(247.90, 48.17) --
	(251.14, 48.16) --
	(254.39, 48.16) --
	(257.63, 48.16) --
	(260.87, 48.16) --
	(264.12, 48.16) --
	(267.36, 48.16) --
	(270.60, 48.16) --
	(273.85, 48.17) --
	(277.09, 48.17) --
	(280.34, 48.17) --
	(283.58, 48.17);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 27.31, 49.28) --
	( 30.56, 50.77) --
	( 33.80, 52.24) --
	( 37.04, 53.70) --
	( 40.29, 55.14) --
	( 43.53, 56.57) --
	( 46.78, 57.98) --
	( 50.02, 59.38) --
	( 53.26, 60.76) --
	( 56.51, 62.12) --
	( 59.75, 63.47) --
	( 63.00, 64.80) --
	( 66.24, 66.12) --
	( 69.48, 67.42) --
	( 72.73, 68.70) --
	( 75.97, 69.97) --
	( 79.22, 71.22) --
	( 82.46, 72.46) --
	( 85.70, 73.69) --
	( 88.95, 74.89) --
	( 92.19, 76.08) --
	( 95.43, 77.25) --
	( 98.68, 78.41) --
	(101.92, 79.55) --
	(105.17, 80.67) --
	(108.41, 81.77) --
	(111.65, 82.86) --
	(114.90, 83.94) --
	(118.14, 85.01) --
	(121.39, 86.06) --
	(124.63, 87.10) --
	(127.87, 88.12) --
	(131.12, 89.11) --
	(134.36, 90.09) --
	(137.61, 91.06) --
	(140.85, 92.01) --
	(144.09, 92.95) --
	(147.34, 93.89) --
	(150.58, 94.82) --
	(153.82, 95.75) --
	(157.07, 96.69) --
	(160.31, 97.61) --
	(163.56, 98.51) --
	(166.80, 99.41) --
	(170.04,100.30) --
	(173.29,101.18) --
	(176.53,102.05) --
	(179.78,102.92) --
	(183.02,103.78) --
	(186.26,104.64) --
	(189.51,105.49) --
	(192.75,106.34) --
	(196.00,107.19) --
	(199.24,108.02) --
	(202.48,108.85) --
	(205.73,109.67) --
	(208.97,110.49) --
	(212.21,111.30) --
	(215.46,112.10) --
	(218.70,112.90) --
	(221.95,113.69) --
	(225.19,114.48) --
	(228.43,115.26) --
	(231.68,116.03) --
	(234.92,116.79) --
	(238.17,117.55) --
	(241.41,118.30) --
	(244.65,119.05) --
	(247.90,119.79) --
	(251.14,120.52) --
	(254.39,121.25) --
	(257.63,121.97) --
	(260.87,122.69) --
	(264.12,123.39) --
	(267.36,124.10) --
	(270.60,124.79) --
	(273.85,125.48) --
	(277.09,126.16) --
	(280.34,126.84) --
	(283.58,127.51);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 27.31, 49.13) --
	( 30.56, 51.18) --
	( 33.80, 53.21) --
	( 37.04, 55.22) --
	( 40.29, 57.21) --
	( 43.53, 59.18) --
	( 46.78, 61.13) --
	( 50.02, 63.06) --
	( 53.26, 64.97) --
	( 56.51, 66.86) --
	( 59.75, 68.72) --
	( 63.00, 70.57) --
	( 66.24, 72.40) --
	( 69.48, 74.20) --
	( 72.73, 75.99) --
	( 75.97, 77.75) --
	( 79.22, 79.50) --
	( 82.46, 81.23) --
	( 85.70, 82.94) --
	( 88.95, 84.63) --
	( 92.19, 86.29) --
	( 95.43, 87.93) --
	( 98.68, 89.55) --
	(101.92, 91.15) --
	(105.17, 92.73) --
	(108.41, 94.28) --
	(111.65, 95.82) --
	(114.90, 97.35) --
	(118.14, 98.86) --
	(121.39,100.35) --
	(124.63,101.82) --
	(127.87,103.26) --
	(131.12,104.67) --
	(134.36,106.06) --
	(137.61,107.43) --
	(140.85,108.79) --
	(144.09,110.14) --
	(147.34,111.48) --
	(150.58,112.81) --
	(153.82,114.15) --
	(157.07,115.48) --
	(160.31,116.80) --
	(163.56,118.11) --
	(166.80,119.40) --
	(170.04,120.68) --
	(173.29,121.95) --
	(176.53,123.21) --
	(179.78,124.46) --
	(183.02,125.71) --
	(186.26,126.95) --
	(189.51,128.18) --
	(192.75,129.41) --
	(196.00,130.63) --
	(199.24,131.84) --
	(202.48,133.05) --
	(205.73,134.24) --
	(208.97,135.43) --
	(212.21,136.61) --
	(215.46,137.78) --
	(218.70,138.94) --
	(221.95,140.09) --
	(225.19,141.23) --
	(228.43,142.37) --
	(231.68,143.50) --
	(234.92,144.62) --
	(238.17,145.73) --
	(241.41,146.83) --
	(244.65,147.92) --
	(247.90,149.01) --
	(251.14,150.08) --
	(254.39,151.15) --
	(257.63,152.21) --
	(260.87,153.26) --
	(264.12,154.30) --
	(267.36,155.33) --
	(270.60,156.36) --
	(273.85,157.38) --
	(277.09,158.39) --
	(280.34,159.39) --
	(283.58,160.38);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 48.86) --
	( 30.56, 51.44) --
	( 33.80, 54.00) --
	( 37.04, 56.54) --
	( 40.29, 59.06) --
	( 43.53, 61.55) --
	( 46.78, 64.02) --
	( 50.02, 66.46) --
	( 53.26, 68.88) --
	( 56.51, 71.28) --
	( 59.75, 73.65) --
	( 63.00, 76.00) --
	( 66.24, 78.32) --
	( 69.48, 80.62) --
	( 72.73, 82.90) --
	( 75.97, 85.15) --
	( 79.22, 87.38) --
	( 82.46, 89.59) --
	( 85.70, 91.78) --
	( 88.95, 93.94) --
	( 92.19, 96.08) --
	( 95.43, 98.19) --
	( 98.68,100.27) --
	(101.92,102.33) --
	(105.17,104.37) --
	(108.41,106.38) --
	(111.65,108.37) --
	(114.90,110.34) --
	(118.14,112.29) --
	(121.39,114.23) --
	(124.63,116.14) --
	(127.87,118.01) --
	(131.12,119.85) --
	(134.36,121.67) --
	(137.61,123.46) --
	(140.85,125.24) --
	(144.09,127.00) --
	(147.34,128.75) --
	(150.58,130.50) --
	(153.82,132.25) --
	(157.07,134.01) --
	(160.31,135.74) --
	(163.56,137.46) --
	(166.80,139.17) --
	(170.04,140.86) --
	(173.29,142.54) --
	(176.53,144.21) --
	(179.78,145.87) --
	(183.02,147.53) --
	(186.26,149.17) --
	(189.51,150.81) --
	(192.75,152.44) --
	(196.00,154.07) --
	(199.24,155.68) --
	(202.48,157.28) --
	(205.73,158.87) --
	(208.97,160.45) --
	(212.21,162.03) --
	(215.46,163.59) --
	(218.70,165.14) --
	(221.95,166.68) --
	(225.19,168.21) --
	(228.43,169.74) --
	(231.68,171.25) --
	(234.92,172.75) --
	(238.17,174.24) --
	(241.41,175.72) --
	(244.65,177.20) --
	(247.90,178.66) --
	(251.14,180.11) --
	(254.39,181.55) --
	(257.63,182.98) --
	(260.87,184.41) --
	(264.12,185.82) --
	(267.36,187.22) --
	(270.60,188.61) --
	(273.85,189.99) --
	(277.09,191.36) --
	(280.34,192.73) --
	(283.58,194.08);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 27.31, 53.19) --
	( 30.68, 47.36) --
	( 23.95, 47.36) --
	cycle;

\path[fill=fillColor] ( 35.86, 57.06) --
	( 39.22, 51.23) --
	( 32.49, 51.23) --
	cycle;

\path[fill=fillColor] ( 44.40, 60.55) --
	( 47.76, 54.73) --
	( 41.03, 54.73) --
	cycle;

\path[fill=fillColor] ( 52.94, 64.55) --
	( 56.30, 58.73) --
	( 49.58, 58.73) --
	cycle;

\path[fill=fillColor] ( 61.48, 68.20) --
	( 64.85, 62.37) --
	( 58.12, 62.37) --
	cycle;

\path[fill=fillColor] ( 70.02, 71.66) --
	( 73.39, 65.83) --
	( 66.66, 65.83) --
	cycle;

\path[fill=fillColor] ( 78.57, 74.89) --
	( 81.93, 69.06) --
	( 75.20, 69.06) --
	cycle;

\path[fill=fillColor] ( 87.11, 78.06) --
	( 90.47, 72.23) --
	( 83.75, 72.23) --
	cycle;

\path[fill=fillColor] ( 95.65, 81.32) --
	( 99.01, 75.49) --
	( 92.29, 75.49) --
	cycle;

\path[fill=fillColor] (104.19, 84.53) --
	(107.56, 78.70) --
	(100.83, 78.70) --
	cycle;

\path[fill=fillColor] (112.74, 87.29) --
	(116.10, 81.47) --
	(109.37, 81.47) --
	cycle;

\path[fill=fillColor] (121.28, 89.96) --
	(124.64, 84.14) --
	(117.91, 84.14) --
	cycle;

\path[fill=fillColor] (129.82, 92.46) --
	(133.18, 86.64) --
	(126.46, 86.64) --
	cycle;

\path[fill=fillColor] (138.36, 95.06) --
	(141.73, 89.23) --
	(135.00, 89.23) --
	cycle;

\path[fill=fillColor] (146.90, 97.72) --
	(150.27, 91.89) --
	(143.54, 91.89) --
	cycle;

\path[fill=fillColor] (155.45,100.07) --
	(158.81, 94.25) --
	(152.08, 94.25) --
	cycle;

\path[fill=fillColor] (163.99,102.38) --
	(167.35, 96.55) --
	(160.63, 96.55) --
	cycle;

\path[fill=fillColor] (172.53,104.65) --
	(175.89, 98.82) --
	(169.17, 98.82) --
	cycle;

\path[fill=fillColor] (181.07,107.23) --
	(184.44,101.41) --
	(177.71,101.41) --
	cycle;

\path[fill=fillColor] (189.62,109.45) --
	(192.98,103.63) --
	(186.25,103.63) --
	cycle;

\path[fill=fillColor] (198.16,111.78) --
	(201.52,105.95) --
	(194.79,105.95) --
	cycle;

\path[fill=fillColor] (206.70,113.72) --
	(210.06,107.89) --
	(203.34,107.89) --
	cycle;

\path[fill=fillColor] (215.24,115.86) --
	(218.61,110.04) --
	(211.88,110.04) --
	cycle;

\path[fill=fillColor] (223.78,118.03) --
	(227.15,112.20) --
	(220.42,112.20) --
	cycle;

\path[fill=fillColor] (232.33,120.05) --
	(235.69,114.22) --
	(228.96,114.22) --
	cycle;

\path[fill=fillColor] (240.87,122.05) --
	(244.23,116.22) --
	(237.51,116.22) --
	cycle;

\path[fill=fillColor] (249.41,124.03) --
	(252.77,118.21) --
	(246.05,118.21) --
	cycle;

\path[fill=fillColor] (257.95,125.86) --
	(261.32,120.04) --
	(254.59,120.04) --
	cycle;

\path[fill=fillColor] (266.50,127.84) --
	(269.86,122.01) --
	(263.13,122.01) --
	cycle;

\path[fill=fillColor] (275.04,129.56) --
	(278.40,123.73) --
	(271.67,123.73) --
	cycle;

\path[fill=fillColor] (283.58,131.45) --
	(286.94,125.62) --
	(280.22,125.62) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 24.82, 46.71) --
	( 29.81, 46.71) --
	( 29.81, 51.71) --
	( 24.82, 51.71) --
	cycle;

\path[fill=fillColor] ( 33.36, 51.99) --
	( 38.35, 51.99) --
	( 38.35, 56.98) --
	( 33.36, 56.98) --
	cycle;

\path[fill=fillColor] ( 41.90, 56.82) --
	( 46.90, 56.82) --
	( 46.90, 61.82) --
	( 41.90, 61.82) --
	cycle;

\path[fill=fillColor] ( 50.44, 62.34) --
	( 55.44, 62.34) --
	( 55.44, 67.33) --
	( 50.44, 67.33) --
	cycle;

\path[fill=fillColor] ( 58.98, 67.33) --
	( 63.98, 67.33) --
	( 63.98, 72.32) --
	( 58.98, 72.32) --
	cycle;

\path[fill=fillColor] ( 67.53, 72.15) --
	( 72.52, 72.15) --
	( 72.52, 77.15) --
	( 67.53, 77.15) --
	cycle;

\path[fill=fillColor] ( 76.07, 76.68) --
	( 81.06, 76.68) --
	( 81.06, 81.67) --
	( 76.07, 81.67) --
	cycle;

\path[fill=fillColor] ( 84.61, 81.16) --
	( 89.61, 81.16) --
	( 89.61, 86.16) --
	( 84.61, 86.16) --
	cycle;

\path[fill=fillColor] ( 93.15, 85.71) --
	( 98.15, 85.71) --
	( 98.15, 90.70) --
	( 93.15, 90.70) --
	cycle;

\path[fill=fillColor] (101.70, 90.24) --
	(106.69, 90.24) --
	(106.69, 95.23) --
	(101.70, 95.23) --
	cycle;

\path[fill=fillColor] (110.24, 94.02) --
	(115.23, 94.02) --
	(115.23, 99.01) --
	(110.24, 99.01) --
	cycle;

\path[fill=fillColor] (118.78, 97.87) --
	(123.78, 97.87) --
	(123.78,102.86) --
	(118.78,102.86) --
	cycle;

\path[fill=fillColor] (127.32,101.41) --
	(132.32,101.41) --
	(132.32,106.41) --
	(127.32,106.41) --
	cycle;

\path[fill=fillColor] (135.86,105.12) --
	(140.86,105.12) --
	(140.86,110.12) --
	(135.86,110.12) --
	cycle;

\path[fill=fillColor] (144.41,108.90) --
	(149.40,108.90) --
	(149.40,113.89) --
	(144.41,113.89) --
	cycle;

\path[fill=fillColor] (152.95,112.30) --
	(157.94,112.30) --
	(157.94,117.30) --
	(152.95,117.30) --
	cycle;

\path[fill=fillColor] (161.49,115.57) --
	(166.49,115.57) --
	(166.49,120.56) --
	(161.49,120.56) --
	cycle;

\path[fill=fillColor] (170.03,118.83) --
	(175.03,118.83) --
	(175.03,123.82) --
	(170.03,123.82) --
	cycle;

\path[fill=fillColor] (178.58,122.59) --
	(183.57,122.59) --
	(183.57,127.59) --
	(178.58,127.59) --
	cycle;

\path[fill=fillColor] (187.12,125.78) --
	(192.11,125.78) --
	(192.11,130.78) --
	(187.12,130.78) --
	cycle;

\path[fill=fillColor] (195.66,129.19) --
	(200.66,129.19) --
	(200.66,134.19) --
	(195.66,134.19) --
	cycle;

\path[fill=fillColor] (204.20,131.97) --
	(209.20,131.97) --
	(209.20,136.97) --
	(204.20,136.97) --
	cycle;

\path[fill=fillColor] (212.74,135.14) --
	(217.74,135.14) --
	(217.74,140.13) --
	(212.74,140.13) --
	cycle;

\path[fill=fillColor] (221.29,138.27) --
	(226.28,138.27) --
	(226.28,143.26) --
	(221.29,143.26) --
	cycle;

\path[fill=fillColor] (229.83,141.17) --
	(234.82,141.17) --
	(234.82,146.17) --
	(229.83,146.17) --
	cycle;

\path[fill=fillColor] (238.37,144.08) --
	(243.37,144.08) --
	(243.37,149.07) --
	(238.37,149.07) --
	cycle;

\path[fill=fillColor] (246.91,147.11) --
	(251.91,147.11) --
	(251.91,152.11) --
	(246.91,152.11) --
	cycle;

\path[fill=fillColor] (255.46,149.72) --
	(260.45,149.72) --
	(260.45,154.72) --
	(255.46,154.72) --
	cycle;

\path[fill=fillColor] (264.00,152.62) --
	(268.99,152.62) --
	(268.99,157.62) --
	(264.00,157.62) --
	cycle;

\path[fill=fillColor] (272.54,155.18) --
	(277.54,155.18) --
	(277.54,160.17) --
	(272.54,160.17) --
	cycle;

\path[fill=fillColor] (281.08,157.94) --
	(286.08,157.94) --
	(286.08,162.94) --
	(281.08,162.94) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 23.78, 49.02) -- ( 30.85, 49.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.31, 45.48) -- ( 27.31, 52.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.32, 55.61) -- ( 39.39, 55.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.86, 52.08) -- ( 35.86, 59.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87, 61.73) -- ( 47.93, 61.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.40, 58.20) -- ( 44.40, 65.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.41, 68.71) -- ( 56.47, 68.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.94, 65.18) -- ( 52.94, 72.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.95, 74.96) -- ( 65.01, 74.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.48, 71.43) -- ( 61.48, 78.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.49, 81.15) -- ( 73.56, 81.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.02, 77.61) -- ( 70.02, 84.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.03, 86.95) -- ( 82.10, 86.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.57, 83.42) -- ( 78.57, 90.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.58, 92.74) -- ( 90.64, 92.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.11, 89.21) -- ( 87.11, 96.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.12, 98.56) -- ( 99.18, 98.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65, 95.03) -- ( 95.65,102.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.66,104.43) -- (107.73,104.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.19,100.89) -- (104.19,107.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.20,109.18) -- (116.27,109.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.74,105.64) -- (112.74,112.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.75,114.24) -- (124.81,114.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.28,110.71) -- (121.28,117.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.29,118.85) -- (133.35,118.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.82,115.32) -- (129.82,122.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.83,123.69) -- (141.89,123.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.36,120.16) -- (138.36,127.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.37,128.65) -- (150.44,128.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.90,125.12) -- (146.90,132.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.91,133.17) -- (158.98,133.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (155.45,129.63) -- (155.45,136.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.46,137.39) -- (167.52,137.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.99,133.86) -- (163.99,140.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.00,141.70) -- (176.06,141.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.53,138.17) -- (172.53,145.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.54,146.71) -- (184.61,146.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.07,143.18) -- (181.07,150.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.08,150.92) -- (193.15,150.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.62,147.39) -- (189.62,154.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.63,155.51) -- (201.69,155.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.16,151.98) -- (198.16,159.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.17,159.19) -- (210.23,159.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.70,155.65) -- (206.70,162.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.71,163.43) -- (218.77,163.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.24,159.90) -- (215.24,166.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.25,167.61) -- (227.32,167.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.78,164.08) -- (223.78,171.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.79,171.46) -- (235.86,171.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.33,167.93) -- (232.33,174.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.34,175.31) -- (244.40,175.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.87,171.78) -- (240.87,178.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.88,179.58) -- (252.94,179.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.41,176.04) -- (249.41,183.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.42,182.99) -- (261.49,182.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.95,179.46) -- (257.95,186.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.96,186.92) -- (270.03,186.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (266.50,183.38) -- (266.50,190.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.51,190.40) -- (278.57,190.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.04,186.87) -- (275.04,193.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.05,194.15) -- (287.11,194.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.58,190.62) -- (283.58,197.68);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 27.31, 47.39) circle (  2.50);

\path[fill=fillColor] ( 35.86, 48.28) circle (  2.50);

\path[fill=fillColor] ( 44.40, 48.31) circle (  2.50);

\path[fill=fillColor] ( 52.94, 47.63) circle (  2.50);

\path[fill=fillColor] ( 61.48, 48.69) circle (  2.50);

\path[fill=fillColor] ( 70.02, 48.09) circle (  2.50);

\path[fill=fillColor] ( 78.57, 48.15) circle (  2.50);

\path[fill=fillColor] ( 87.11, 48.59) circle (  2.50);

\path[fill=fillColor] ( 95.65, 48.01) circle (  2.50);

\path[fill=fillColor] (104.19, 48.33) circle (  2.50);

\path[fill=fillColor] (112.74, 47.97) circle (  2.50);

\path[fill=fillColor] (121.28, 47.68) circle (  2.50);

\path[fill=fillColor] (129.82, 47.78) circle (  2.50);

\path[fill=fillColor] (138.36, 48.47) circle (  2.50);

\path[fill=fillColor] (146.90, 47.94) circle (  2.50);

\path[fill=fillColor] (155.45, 48.48) circle (  2.50);

\path[fill=fillColor] (163.99, 48.56) circle (  2.50);

\path[fill=fillColor] (172.53, 48.33) circle (  2.50);

\path[fill=fillColor] (181.07, 48.25) circle (  2.50);

\path[fill=fillColor] (189.62, 48.13) circle (  2.50);

\path[fill=fillColor] (198.16, 48.23) circle (  2.50);

\path[fill=fillColor] (206.70, 47.82) circle (  2.50);

\path[fill=fillColor] (215.24, 48.29) circle (  2.50);

\path[fill=fillColor] (223.78, 47.84) circle (  2.50);

\path[fill=fillColor] (232.33, 47.85) circle (  2.50);

\path[fill=fillColor] (240.87, 48.44) circle (  2.50);

\path[fill=fillColor] (249.41, 48.50) circle (  2.50);

\path[fill=fillColor] (257.95, 48.19) circle (  2.50);

\path[fill=fillColor] (266.50, 48.39) circle (  2.50);

\path[fill=fillColor] (275.04, 48.09) circle (  2.50);

\path[fill=fillColor] (283.58, 47.98) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 30.69) rectangle (283.58,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 55.97) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,150.33) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 59.00) --
	( 27.31, 59.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,153.36) --
	( 27.31,153.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 27.94) --
	( 27.31, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.74, 27.94) --
	(112.74, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.16, 27.94) --
	(198.16, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.74, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.16, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (155.45,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,112.42) {Mean generation/serial interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,202.22) {C};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,216.81);
\end{scope}
\begin{scope}
\path[clip] (323.24, 30.69) rectangle (572.66,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (323.24, 30.69) rectangle (572.66,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (323.24, 49.95) --
	(326.39, 49.96) --
	(329.55, 49.96) --
	(332.71, 49.96) --
	(335.87, 49.97) --
	(339.02, 49.97) --
	(342.18, 49.98) --
	(345.34, 49.98) --
	(348.49, 49.98) --
	(351.65, 49.98) --
	(354.81, 49.99) --
	(357.97, 49.99) --
	(361.12, 49.99) --
	(364.28, 49.99) --
	(367.44, 49.99) --
	(370.60, 50.00) --
	(373.75, 50.00) --
	(376.91, 50.00) --
	(380.07, 50.00) --
	(383.22, 50.00) --
	(386.38, 50.00) --
	(389.54, 50.00) --
	(392.70, 50.00) --
	(395.85, 50.00) --
	(399.01, 50.00) --
	(402.17, 50.00) --
	(405.32, 50.00) --
	(408.48, 49.99) --
	(411.64, 49.99) --
	(414.80, 49.99) --
	(417.95, 49.99) --
	(421.11, 49.98) --
	(424.27, 49.98) --
	(427.43, 49.97) --
	(430.58, 49.97) --
	(433.74, 49.96) --
	(436.90, 49.95) --
	(440.05, 49.95) --
	(443.21, 49.94) --
	(446.37, 49.94) --
	(449.53, 49.94) --
	(452.68, 49.94) --
	(455.84, 49.94) --
	(459.00, 49.95) --
	(462.16, 49.95) --
	(465.31, 49.95) --
	(468.47, 49.95) --
	(471.63, 49.96) --
	(474.78, 49.96) --
	(477.94, 49.96) --
	(481.10, 49.97) --
	(484.26, 49.97) --
	(487.41, 49.97) --
	(490.57, 49.98) --
	(493.73, 49.98) --
	(496.89, 49.98) --
	(500.04, 49.99) --
	(503.20, 49.99) --
	(506.36, 49.99) --
	(509.51, 50.00) --
	(512.67, 50.00) --
	(515.83, 50.00) --
	(518.99, 50.00) --
	(522.14, 50.00) --
	(525.30, 50.00) --
	(528.46, 50.00) --
	(531.62, 50.00) --
	(534.77, 49.99) --
	(537.93, 49.99) --
	(541.09, 49.99) --
	(544.24, 49.99) --
	(547.40, 49.99) --
	(550.56, 49.98) --
	(553.72, 49.98) --
	(556.87, 49.97) --
	(560.03, 49.97) --
	(563.19, 49.96) --
	(566.35, 49.96) --
	(569.50, 49.95) --
	(572.66, 49.94);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (323.24,140.06) --
	(326.39,138.91) --
	(329.55,137.78) --
	(332.71,136.67) --
	(335.87,135.58) --
	(339.02,134.51) --
	(342.18,133.46) --
	(345.34,132.43) --
	(348.49,131.43) --
	(351.65,130.44) --
	(354.81,129.47) --
	(357.97,128.52) --
	(361.12,127.60) --
	(364.28,126.69) --
	(367.44,125.80) --
	(370.60,124.93) --
	(373.75,124.08) --
	(376.91,123.25) --
	(380.07,122.43) --
	(383.22,121.64) --
	(386.38,120.87) --
	(389.54,120.12) --
	(392.70,119.40) --
	(395.85,118.70) --
	(399.01,118.02) --
	(402.17,117.36) --
	(405.32,116.71) --
	(408.48,116.08) --
	(411.64,115.46) --
	(414.80,114.85) --
	(417.95,114.27) --
	(421.11,113.70) --
	(424.27,113.17) --
	(427.43,112.65) --
	(430.58,112.15) --
	(433.74,111.66) --
	(436.90,111.18) --
	(440.05,110.70) --
	(443.21,110.23) --
	(446.37,109.76) --
	(449.53,109.28) --
	(452.68,108.82) --
	(455.84,108.37) --
	(459.00,107.93) --
	(462.16,107.50) --
	(465.31,107.08) --
	(468.47,106.66) --
	(471.63,106.25) --
	(474.78,105.85) --
	(477.94,105.44) --
	(481.10,105.04) --
	(484.26,104.64) --
	(487.41,104.25) --
	(490.57,103.86) --
	(493.73,103.48) --
	(496.89,103.10) --
	(500.04,102.73) --
	(503.20,102.36) --
	(506.36,102.00) --
	(509.51,101.65) --
	(512.67,101.30) --
	(515.83,100.95) --
	(518.99,100.61) --
	(522.14,100.27) --
	(525.30, 99.94) --
	(528.46, 99.61) --
	(531.62, 99.29) --
	(534.77, 98.98) --
	(537.93, 98.67) --
	(541.09, 98.37) --
	(544.24, 98.07) --
	(547.40, 97.77) --
	(550.56, 97.48) --
	(553.72, 97.20) --
	(556.87, 96.92) --
	(560.03, 96.65) --
	(563.19, 96.38) --
	(566.35, 96.12) --
	(569.50, 95.86) --
	(572.66, 95.61);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (323.24,172.74) --
	(326.39,170.89) --
	(329.55,169.07) --
	(332.71,167.28) --
	(335.87,165.53) --
	(339.02,163.81) --
	(342.18,162.12) --
	(345.34,160.46) --
	(348.49,158.84) --
	(351.65,157.25) --
	(354.81,155.70) --
	(357.97,154.18) --
	(361.12,152.70) --
	(364.28,151.25) --
	(367.44,149.83) --
	(370.60,148.44) --
	(373.75,147.08) --
	(376.91,145.75) --
	(380.07,144.46) --
	(383.22,143.19) --
	(386.38,141.97) --
	(389.54,140.79) --
	(392.70,139.65) --
	(395.85,138.55) --
	(399.01,137.48) --
	(402.17,136.44) --
	(405.32,135.42) --
	(408.48,134.43) --
	(411.64,133.46) --
	(414.80,132.50) --
	(417.95,131.58) --
	(421.11,130.71) --
	(424.27,129.87) --
	(427.43,129.06) --
	(430.58,128.28) --
	(433.74,127.53) --
	(436.90,126.78) --
	(440.05,126.05) --
	(443.21,125.32) --
	(446.37,124.59) --
	(449.53,123.86) --
	(452.68,123.14) --
	(455.84,122.45) --
	(459.00,121.78) --
	(462.16,121.12) --
	(465.31,120.48) --
	(468.47,119.84) --
	(471.63,119.22) --
	(474.78,118.60) --
	(477.94,117.99) --
	(481.10,117.38) --
	(484.26,116.77) --
	(487.41,116.17) --
	(490.57,115.58) --
	(493.73,115.01) --
	(496.89,114.44) --
	(500.04,113.88) --
	(503.20,113.33) --
	(506.36,112.79) --
	(509.51,112.25) --
	(512.67,111.72) --
	(515.83,111.20) --
	(518.99,110.69) --
	(522.14,110.19) --
	(525.30,109.69) --
	(528.46,109.21) --
	(531.62,108.73) --
	(534.77,108.26) --
	(537.93,107.80) --
	(541.09,107.35) --
	(544.24,106.91) --
	(547.40,106.48) --
	(550.56,106.05) --
	(553.72,105.63) --
	(556.87,105.22) --
	(560.03,104.82) --
	(563.19,104.43) --
	(566.35,104.05) --
	(569.50,103.67) --
	(572.66,103.31);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (339.86,189.79) --
	(342.81,187.67) --
	(345.76,185.60) --
	(348.70,183.56) --
	(351.65,181.56) --
	(354.60,179.59) --
	(357.55,177.67) --
	(360.49,175.78) --
	(363.44,173.93) --
	(366.39,172.12) --
	(369.33,170.35) --
	(372.28,168.62) --
	(375.23,166.93) --
	(378.17,165.28) --
	(381.12,163.66) --
	(384.07,162.09) --
	(387.01,160.54) --
	(389.96,159.03) --
	(392.91,157.56) --
	(395.85,156.12) --
	(398.80,154.71) --
	(401.75,153.36) --
	(404.69,152.05) --
	(407.64,150.80) --
	(410.59,149.59) --
	(413.53,148.42) --
	(416.48,147.28) --
	(419.43,146.18) --
	(422.37,145.09) --
	(425.32,144.03) --
	(428.27,142.98) --
	(431.21,141.94) --
	(434.16,140.93) --
	(437.11,139.96) --
	(440.05,139.04) --
	(443.00,138.14) --
	(445.95,137.27) --
	(448.90,136.41) --
	(451.84,135.56) --
	(454.79,134.70) --
	(457.74,133.83) --
	(460.68,132.98) --
	(463.63,132.14) --
	(466.58,131.32) --
	(469.52,130.52) --
	(472.47,129.73) --
	(475.42,128.95) --
	(478.36,128.19) --
	(481.31,127.43) --
	(484.26,126.69) --
	(487.20,125.95) --
	(490.15,125.22) --
	(493.10,124.50) --
	(496.04,123.80) --
	(498.99,123.11) --
	(501.94,122.43) --
	(504.88,121.76) --
	(507.83,121.11) --
	(510.78,120.46) --
	(513.72,119.81) --
	(516.67,119.18) --
	(519.62,118.55) --
	(522.56,117.94) --
	(525.51,117.34) --
	(528.46,116.75) --
	(531.41,116.16) --
	(534.35,115.59) --
	(537.30,115.03) --
	(540.25,114.48) --
	(543.19,113.94) --
	(546.14,113.41) --
	(549.09,112.89) --
	(552.03,112.37) --
	(554.98,111.87) --
	(557.93,111.38) --
	(560.87,110.90) --
	(563.82,110.43) --
	(566.77,109.97) --
	(569.71,109.51) --
	(572.66,109.07);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] (323.24,144.25) --
	(326.60,138.42) --
	(319.87,138.42) --
	cycle;

\path[fill=fillColor] (331.55,140.97) --
	(334.91,135.14) --
	(328.19,135.14) --
	cycle;

\path[fill=fillColor] (339.86,138.16) --
	(343.23,132.34) --
	(336.50,132.34) --
	cycle;

\path[fill=fillColor] (348.18,135.26) --
	(351.54,129.43) --
	(344.81,129.43) --
	cycle;

\path[fill=fillColor] (356.49,132.48) --
	(359.86,126.65) --
	(353.13,126.65) --
	cycle;

\path[fill=fillColor] (364.81,130.21) --
	(368.17,124.38) --
	(361.44,124.38) --
	cycle;

\path[fill=fillColor] (373.12,128.06) --
	(376.48,122.24) --
	(369.76,122.24) --
	cycle;

\path[fill=fillColor] (381.44,125.99) --
	(384.80,120.16) --
	(378.07,120.16) --
	cycle;

\path[fill=fillColor] (389.75,123.88) --
	(393.11,118.06) --
	(386.39,118.06) --
	cycle;

\path[fill=fillColor] (398.06,121.85) --
	(401.43,116.03) --
	(394.70,116.03) --
	cycle;

\path[fill=fillColor] (406.38,120.12) --
	(409.74,114.29) --
	(403.01,114.29) --
	cycle;

\path[fill=fillColor] (414.69,118.70) --
	(418.06,112.87) --
	(411.33,112.87) --
	cycle;

\path[fill=fillColor] (423.01,117.31) --
	(426.37,111.48) --
	(419.64,111.48) --
	cycle;

\path[fill=fillColor] (431.32,116.01) --
	(434.68,110.19) --
	(427.96,110.19) --
	cycle;

\path[fill=fillColor] (439.63,114.62) --
	(443.00,108.79) --
	(436.27,108.79) --
	cycle;

\path[fill=fillColor] (447.95,113.52) --
	(451.31,107.69) --
	(444.58,107.69) --
	cycle;

\path[fill=fillColor] (456.26,112.31) --
	(459.63,106.48) --
	(452.90,106.48) --
	cycle;

\path[fill=fillColor] (464.58,111.32) --
	(467.94,105.49) --
	(461.21,105.49) --
	cycle;

\path[fill=fillColor] (472.89,109.94) --
	(476.25,104.11) --
	(469.53,104.11) --
	cycle;

\path[fill=fillColor] (481.20,108.82) --
	(484.57,102.99) --
	(477.84,102.99) --
	cycle;

\path[fill=fillColor] (489.52,107.83) --
	(492.88,102.00) --
	(486.15,102.00) --
	cycle;

\path[fill=fillColor] (497.83,106.85) --
	(501.20,101.02) --
	(494.47,101.02) --
	cycle;

\path[fill=fillColor] (506.15,105.93) --
	(509.51,100.10) --
	(502.78,100.10) --
	cycle;

\path[fill=fillColor] (514.46,104.96) --
	(517.82, 99.14) --
	(511.10, 99.14) --
	cycle;

\path[fill=fillColor] (522.78,104.12) --
	(526.14, 98.29) --
	(519.41, 98.29) --
	cycle;

\path[fill=fillColor] (531.09,103.27) --
	(534.45, 97.45) --
	(527.73, 97.45) --
	cycle;

\path[fill=fillColor] (539.40,102.56) --
	(542.77, 96.73) --
	(536.04, 96.73) --
	cycle;

\path[fill=fillColor] (547.72,101.64) --
	(551.08, 95.82) --
	(544.35, 95.82) --
	cycle;

\path[fill=fillColor] (556.03,100.83) --
	(559.40, 95.01) --
	(552.67, 95.01) --
	cycle;

\path[fill=fillColor] (564.35,100.20) --
	(567.71, 94.37) --
	(560.98, 94.37) --
	cycle;

\path[fill=fillColor] (572.66, 99.42) --
	(576.02, 93.59) --
	(569.30, 93.59) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] (320.74,170.78) --
	(325.73,170.78) --
	(325.73,175.77) --
	(320.74,175.77) --
	cycle;

\path[fill=fillColor] (329.05,165.46) --
	(334.05,165.46) --
	(334.05,170.45) --
	(329.05,170.45) --
	cycle;

\path[fill=fillColor] (337.37,160.96) --
	(342.36,160.96) --
	(342.36,165.95) --
	(337.37,165.95) --
	cycle;

\path[fill=fillColor] (345.68,156.26) --
	(350.68,156.26) --
	(350.68,161.25) --
	(345.68,161.25) --
	cycle;

\path[fill=fillColor] (353.99,151.74) --
	(358.99,151.74) --
	(358.99,156.74) --
	(353.99,156.74) --
	cycle;

\path[fill=fillColor] (362.31,148.12) --
	(367.30,148.12) --
	(367.30,153.11) --
	(362.31,153.11) --
	cycle;

\path[fill=fillColor] (370.62,144.68) --
	(375.62,144.68) --
	(375.62,149.67) --
	(370.62,149.67) --
	cycle;

\path[fill=fillColor] (378.94,141.38) --
	(383.93,141.38) --
	(383.93,146.38) --
	(378.94,146.38) --
	cycle;

\path[fill=fillColor] (387.25,138.08) --
	(392.25,138.08) --
	(392.25,143.07) --
	(387.25,143.07) --
	cycle;

\path[fill=fillColor] (395.57,134.92) --
	(400.56,134.92) --
	(400.56,139.92) --
	(395.57,139.92) --
	cycle;

\path[fill=fillColor] (403.88,132.16) --
	(408.88,132.16) --
	(408.88,137.16) --
	(403.88,137.16) --
	cycle;

\path[fill=fillColor] (412.19,129.98) --
	(417.19,129.98) --
	(417.19,134.97) --
	(412.19,134.97) --
	cycle;

\path[fill=fillColor] (420.51,127.83) --
	(425.50,127.83) --
	(425.50,132.82) --
	(420.51,132.82) --
	cycle;

\path[fill=fillColor] (428.82,125.76) --
	(433.82,125.76) --
	(433.82,130.76) --
	(428.82,130.76) --
	cycle;

\path[fill=fillColor] (437.14,123.59) --
	(442.13,123.59) --
	(442.13,128.58) --
	(437.14,128.58) --
	cycle;

\path[fill=fillColor] (445.45,121.88) --
	(450.45,121.88) --
	(450.45,126.88) --
	(445.45,126.88) --
	cycle;

\path[fill=fillColor] (453.76,120.04) --
	(458.76,120.04) --
	(458.76,125.03) --
	(453.76,125.03) --
	cycle;

\path[fill=fillColor] (462.08,118.49) --
	(467.07,118.49) --
	(467.07,123.48) --
	(462.08,123.48) --
	cycle;

\path[fill=fillColor] (470.39,116.43) --
	(475.39,116.43) --
	(475.39,121.43) --
	(470.39,121.43) --
	cycle;

\path[fill=fillColor] (478.71,114.73) --
	(483.70,114.73) --
	(483.70,119.72) --
	(478.71,119.72) --
	cycle;

\path[fill=fillColor] (487.02,113.23) --
	(492.02,113.23) --
	(492.02,118.23) --
	(487.02,118.23) --
	cycle;

\path[fill=fillColor] (495.34,111.77) --
	(500.33,111.77) --
	(500.33,116.77) --
	(495.34,116.77) --
	cycle;

\path[fill=fillColor] (503.65,110.35) --
	(508.64,110.35) --
	(508.64,115.34) --
	(503.65,115.34) --
	cycle;

\path[fill=fillColor] (511.96,108.90) --
	(516.96,108.90) --
	(516.96,113.90) --
	(511.96,113.90) --
	cycle;

\path[fill=fillColor] (520.28,107.65) --
	(525.27,107.65) --
	(525.27,112.64) --
	(520.28,112.64) --
	cycle;

\path[fill=fillColor] (528.59,106.36) --
	(533.59,106.36) --
	(533.59,111.36) --
	(528.59,111.36) --
	cycle;

\path[fill=fillColor] (536.91,105.34) --
	(541.90,105.34) --
	(541.90,110.33) --
	(536.91,110.33) --
	cycle;

\path[fill=fillColor] (545.22,104.01) --
	(550.22,104.01) --
	(550.22,109.01) --
	(545.22,109.01) --
	cycle;

\path[fill=fillColor] (553.53,102.76) --
	(558.53,102.76) --
	(558.53,107.75) --
	(553.53,107.75) --
	cycle;

\path[fill=fillColor] (561.85,101.83) --
	(566.84,101.83) --
	(566.84,106.83) --
	(561.85,106.83) --
	cycle;

\path[fill=fillColor] (570.16,100.67) --
	(575.16,100.67) --
	(575.16,105.66) --
	(570.16,105.66) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (336.33,190.59) -- (343.40,190.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (339.86,187.06) -- (339.86,194.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (344.65,184.03) -- (351.71,184.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (348.18,180.49) -- (348.18,187.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (352.96,177.69) -- (360.02,177.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.49,174.16) -- (356.49,181.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.27,172.70) -- (368.34,172.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.81,169.17) -- (364.81,176.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.59,167.92) -- (376.65,167.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (373.12,164.38) -- (373.12,171.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.90,163.35) -- (384.97,163.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.44,159.82) -- (381.44,166.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.22,158.83) -- (393.28,158.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.75,155.30) -- (389.75,162.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (394.53,154.55) -- (401.60,154.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.06,151.01) -- (398.06,158.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (402.85,150.76) -- (409.91,150.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.38,147.23) -- (406.38,154.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (411.16,147.81) -- (418.22,147.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (414.69,144.28) -- (414.69,151.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.47,144.92) -- (426.54,144.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.01,141.39) -- (423.01,148.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.79,142.06) -- (434.85,142.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.32,138.53) -- (431.32,145.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.10,139.16) -- (443.17,139.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (439.63,135.62) -- (439.63,142.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.42,136.86) -- (451.48,136.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (447.95,133.33) -- (447.95,140.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.73,134.36) -- (459.79,134.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.26,130.83) -- (456.26,137.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.04,132.27) -- (468.11,132.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.58,128.74) -- (464.58,135.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.36,129.57) -- (476.42,129.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (472.89,126.04) -- (472.89,133.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.67,127.31) -- (484.74,127.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.20,123.78) -- (481.20,130.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.99,125.32) -- (493.05,125.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (489.52,121.79) -- (489.52,128.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.30,123.43) -- (501.36,123.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (497.83,119.89) -- (497.83,126.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.61,121.47) -- (509.68,121.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (506.15,117.93) -- (506.15,125.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.93,119.60) -- (517.99,119.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (514.46,116.06) -- (514.46,123.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.24,117.95) -- (526.31,117.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.78,114.42) -- (522.78,121.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (527.56,116.25) -- (534.62,116.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (531.09,112.72) -- (531.09,119.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.87,114.97) -- (542.94,114.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (539.40,111.44) -- (539.40,118.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.19,113.27) -- (551.25,113.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.72,109.73) -- (547.72,116.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.50,111.59) -- (559.56,111.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (556.03,108.05) -- (556.03,115.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.81,110.40) -- (567.88,110.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.35,106.87) -- (564.35,113.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13,108.89) -- (576.19,108.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66,105.36) -- (572.66,112.42);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (323.24, 49.94) circle (  2.50);

\path[fill=fillColor] (331.55, 50.02) circle (  2.50);

\path[fill=fillColor] (339.86, 50.01) circle (  2.50);

\path[fill=fillColor] (348.18, 49.81) circle (  2.50);

\path[fill=fillColor] (356.49, 49.97) circle (  2.50);

\path[fill=fillColor] (364.81, 50.10) circle (  2.50);

\path[fill=fillColor] (373.12, 49.98) circle (  2.50);

\path[fill=fillColor] (381.44, 50.04) circle (  2.50);

\path[fill=fillColor] (389.75, 50.05) circle (  2.50);

\path[fill=fillColor] (398.06, 49.95) circle (  2.50);

\path[fill=fillColor] (406.38, 50.17) circle (  2.50);

\path[fill=fillColor] (414.69, 49.86) circle (  2.50);

\path[fill=fillColor] (423.01, 49.74) circle (  2.50);

\path[fill=fillColor] (431.32, 49.98) circle (  2.50);

\path[fill=fillColor] (439.63, 50.13) circle (  2.50);

\path[fill=fillColor] (447.95, 50.04) circle (  2.50);

\path[fill=fillColor] (456.26, 49.94) circle (  2.50);

\path[fill=fillColor] (464.58, 49.95) circle (  2.50);

\path[fill=fillColor] (472.89, 49.84) circle (  2.50);

\path[fill=fillColor] (481.20, 49.92) circle (  2.50);

\path[fill=fillColor] (489.52, 49.86) circle (  2.50);

\path[fill=fillColor] (497.83, 50.08) circle (  2.50);

\path[fill=fillColor] (506.15, 49.96) circle (  2.50);

\path[fill=fillColor] (514.46, 50.12) circle (  2.50);

\path[fill=fillColor] (522.78, 50.25) circle (  2.50);

\path[fill=fillColor] (531.09, 49.77) circle (  2.50);

\path[fill=fillColor] (539.40, 49.90) circle (  2.50);

\path[fill=fillColor] (547.72, 50.13) circle (  2.50);

\path[fill=fillColor] (556.03, 50.12) circle (  2.50);

\path[fill=fillColor] (564.35, 49.92) circle (  2.50);

\path[fill=fillColor] (572.66, 49.83) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.24, 30.69) rectangle (572.66,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29, 27.66) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29, 78.74) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,129.82) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,180.91) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49, 30.69) --
	(323.24, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49, 81.77) --
	(323.24, 81.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,132.85) --
	(323.24,132.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,183.94) --
	(323.24,183.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.24, 27.94) --
	(323.24, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.38, 27.94) --
	(406.38, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (489.52, 27.94) --
	(489.52, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (323.24, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (406.38, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.52, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (447.95,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,112.42) {Squared coefficient of variation};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (323.24,202.22) {D};
\end{scope}
\end{tikzpicture}

\end{document}