### 2020-04-14 serial_analysis.tex
i like this new figure

% Created by tikzDevice version 0.12.3 on 2020-04-12 17:51:37
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 34.64, 30.69) rectangle (283.58,266.42);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 34.64, 30.69) rectangle (283.58,266.42);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.97,138.18) -- ( 41.97,138.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.97,138.18) -- ( 41.97,138.18);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 39.22,138.18) --
	( 39.22,138.18) --
	( 44.71,138.18) --
	( 44.71,138.18) --
	( 39.22,138.18) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 39.22,138.18) -- ( 44.71,138.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 56.61,226.33) -- ( 56.61,234.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 56.61,209.04) -- ( 56.61,186.58);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 53.86,226.33) --
	( 53.86,209.04) --
	( 59.36,209.04) --
	( 59.36,226.33) --
	( 53.86,226.33) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 53.86,217.69) -- ( 59.36,217.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 63.93,196.95) -- ( 63.93,228.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 63.93,158.92) -- ( 63.93,152.01);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 61.19,196.95) --
	( 61.19,158.92) --
	( 66.68,158.92) --
	( 66.68,196.95) --
	( 61.19,196.95) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 61.19,165.84) -- ( 66.68,165.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.25,200.40) -- ( 71.25,207.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 71.25,186.58) -- ( 71.25,179.66);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 68.51,200.40) --
	( 68.51,186.58) --
	( 74.00,186.58) --
	( 74.00,200.40) --
	( 68.51,200.40) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 68.51,193.49) -- ( 74.00,193.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.57,221.14) -- ( 78.57,234.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.57,172.75) -- ( 78.57,138.18);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 75.83,221.14) --
	( 75.83,172.75) --
	( 81.32,172.75) --
	( 81.32,221.14) --
	( 75.83,221.14) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 75.83,207.32) -- ( 81.32,207.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 85.90,221.14) -- ( 85.90,221.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 85.90,221.14) -- ( 85.90,221.14);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 83.15,221.14) --
	( 83.15,221.14) --
	( 88.64,221.14) --
	( 88.64,221.14) --
	( 83.15,221.14) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 83.15,221.14) -- ( 88.64,221.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 93.22,238.42) -- ( 93.22,255.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 93.22,203.86) -- ( 93.22,186.58);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 90.47,238.42) --
	( 90.47,203.86) --
	( 95.96,203.86) --
	( 95.96,238.42) --
	( 90.47,238.42) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 90.47,221.14) -- ( 95.96,221.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.54,200.40) -- (100.54,207.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.54,186.58) -- (100.54,179.66);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 97.79,200.40) --
	( 97.79,186.58) --
	(103.29,186.58) --
	(103.29,200.40) --
	( 97.79,200.40) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 97.79,193.49) -- (103.29,193.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (107.86,214.23) -- (107.86,214.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (107.86,214.23) -- (107.86,214.23);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (105.12,214.23) --
	(105.12,214.23) --
	(110.61,214.23) --
	(110.61,214.23) --
	(105.12,214.23) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (105.12,214.23) -- (110.61,214.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (115.18,198.67) -- (115.18,214.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (115.18,164.11) -- (115.18,158.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (112.44,198.67) --
	(112.44,164.11) --
	(117.93,164.11) --
	(117.93,198.67) --
	(112.44,198.67) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (112.44,179.66) -- (117.93,179.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (122.50,186.58) -- (122.50,234.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (122.50,138.18) -- (122.50,103.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (119.76,186.58) --
	(119.76,138.18) --
	(125.25,138.18) --
	(125.25,186.58) --
	(119.76,186.58) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (119.76,158.92) -- (125.25,158.92);
\definecolor{fillColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (129.83,221.14) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.83,171.02) -- (129.83,193.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.83,139.91) -- (129.83,124.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (127.08,171.02) --
	(127.08,139.91) --
	(132.57,139.91) --
	(132.57,171.02) --
	(127.08,171.02) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (127.08,155.47) -- (132.57,155.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.15,183.12) -- (137.15,200.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.15,148.55) -- (137.15,117.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (134.40,183.12) --
	(134.40,148.55) --
	(139.89,148.55) --
	(139.89,183.12) --
	(134.40,183.12) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (134.40,165.84) -- (139.89,165.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (144.47,172.75) -- (144.47,214.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (144.47,131.27) -- (144.47,110.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (141.72,172.75) --
	(141.72,131.27) --
	(147.21,131.27) --
	(147.21,172.75) --
	(141.72,172.75) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (141.72,152.01) -- (147.21,152.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.79,179.66) -- (151.79,221.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.79,138.18) -- (151.79,117.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (149.05,179.66) --
	(149.05,138.18) --
	(154.54,138.18) --
	(154.54,179.66) --
	(149.05,179.66) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (149.05,158.92) -- (154.54,158.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.11,184.85) -- (159.11,228.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.11,126.09) -- (159.11, 55.23);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (156.37,184.85) --
	(156.37,126.09) --
	(161.86,126.09) --
	(161.86,184.85) --
	(156.37,184.85) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (156.37,155.47) -- (161.86,155.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.43,174.48) -- (166.43,200.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.43,129.54) -- (166.43,110.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (163.69,174.48) --
	(163.69,129.54) --
	(169.18,129.54) --
	(169.18,174.48) --
	(163.69,174.48) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (163.69,152.01) -- (169.18,152.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (173.76,171.02) -- (173.76,193.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (173.76,133.00) -- (173.76,117.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (171.01,171.02) --
	(171.01,133.00) --
	(176.50,133.00) --
	(176.50,171.02) --
	(171.01,171.02) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (171.01,152.01) -- (176.50,152.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (181.08,152.01) -- (181.08,179.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (181.08,110.53) -- (181.08, 82.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (178.33,152.01) --
	(178.33,110.53) --
	(183.82,110.53) --
	(183.82,152.01) --
	(178.33,152.01) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (178.33,131.27) -- (183.82,131.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (188.40,165.84) -- (188.40,193.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (188.40,110.53) -- (188.40, 75.97);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (185.65,165.84) --
	(185.65,110.53) --
	(191.14,110.53) --
	(191.14,165.84) --
	(185.65,165.84) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (185.65,138.18) -- (191.14,138.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.72,160.65) -- (195.72,200.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.72,108.80) -- (195.72, 82.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (192.97,160.65) --
	(192.97,108.80) --
	(198.47,108.80) --
	(198.47,160.65) --
	(192.97,160.65) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (192.97,134.73) -- (198.47,134.73);
\definecolor{fillColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (203.04, 48.31) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.04,155.47) -- (203.04,179.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.04,120.90) -- (203.04,103.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (200.30,155.47) --
	(200.30,120.90) --
	(205.79,120.90) --
	(205.79,155.47) --
	(200.30,155.47) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (200.30,138.18) -- (205.79,138.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.36,143.37) -- (210.36,172.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.36,107.08) -- (210.36, 62.14);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (207.62,143.37) --
	(207.62,107.08) --
	(213.11,107.08) --
	(213.11,143.37) --
	(207.62,143.37) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (207.62,127.81) -- (213.11,127.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.69,131.27) -- (217.69,145.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.69, 96.71) -- (217.69, 75.97);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (214.94,131.27) --
	(214.94, 96.71) --
	(220.43, 96.71) --
	(220.43,131.27) --
	(214.94,131.27) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (214.94,117.45) -- (220.43,117.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.01,134.73) -- (225.01,145.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.01,113.99) -- (225.01, 82.88);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (222.26,134.73) --
	(222.26,113.99) --
	(227.75,113.99) --
	(227.75,134.73) --
	(222.26,134.73) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (222.26,124.36) -- (227.75,124.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (232.33,134.73) -- (232.33,145.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (232.33,113.99) -- (232.33, 96.71);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (229.58,134.73) --
	(229.58,113.99) --
	(235.07,113.99) --
	(235.07,134.73) --
	(229.58,134.73) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (229.58,124.36) -- (235.07,124.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (239.65,126.09) -- (239.65,131.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (239.65,115.72) -- (239.65,110.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (236.90,126.09) --
	(236.90,115.72) --
	(242.40,115.72) --
	(242.40,126.09) --
	(236.90,126.09) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (236.90,120.90) -- (242.40,120.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (246.97,110.53) -- (246.97,110.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (246.97,110.53) -- (246.97,110.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (244.23,110.53) --
	(244.23,110.53) --
	(249.72,110.53) --
	(249.72,110.53) --
	(244.23,110.53) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (244.23,110.53) -- (249.72,110.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.62, 89.79) -- (261.62, 89.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.62, 89.79) -- (261.62, 89.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (258.87, 89.79) --
	(258.87, 89.79) --
	(264.36, 89.79) --
	(264.36, 89.79) --
	(258.87, 89.79) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (258.87, 89.79) -- (264.36, 89.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.26, 93.25) -- (276.26,110.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.26, 58.68) -- (276.26, 41.40);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (273.51, 93.25) --
	(273.51, 58.68) --
	(279.00, 58.68) --
	(279.00, 93.25) --
	(273.51, 93.25) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (273.51, 75.97) -- (279.00, 75.97);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.64, 30.69) rectangle (283.58,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69, 45.28) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69,114.41) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69,183.55) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69,252.68) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89, 48.31) --
	( 34.64, 48.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89,117.45) --
	( 34.64,117.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89,186.58) --
	( 34.64,186.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89,255.71) --
	( 34.64,255.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 63.93, 27.94) --
	( 63.93, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.15, 27.94) --
	(137.15, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.36, 27.94) --
	(210.36, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (283.58, 27.94) --
	(283.58, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 63.93, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (137.15, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (210.36, 19.68) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (283.58, 19.68) {30};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (159.11,  7.64) {Cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,148.55) {Forward serial intervals (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 34.64,274.49) {A};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,289.08);
\end{scope}
\begin{scope}
\path[clip] (316.39, 30.69) rectangle (572.66,266.42);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (316.39, 30.69) rectangle (572.66,266.42);
\definecolor{drawColor}{RGB}{240,228,66}
\definecolor{fillColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (346.54,255.71) --
	(354.08,234.51) --
	(361.62,227.85) --
	(369.15,224.34) --
	(376.69,232.95) --
	(384.23,225.18) --
	(391.77,220.08) --
	(399.30,212.60) --
	(406.84,137.83) --
	(414.38,119.60) --
	(421.91,116.21) --
	(429.45,104.34) --
	(436.99,101.52) --
	(444.53, 91.89) --
	(452.06, 87.88) --
	(459.60, 85.77) --
	(467.14, 79.62) --
	(474.68, 76.53) --
	(482.21, 72.74) --
	(489.75, 71.32) --
	(497.29, 69.30) --
	(504.82, 68.00) --
	(512.36, 66.17) --
	(519.90, 65.86) --
	(527.44, 65.06) --
	(534.97, 64.84) --
	(542.51, 64.84) --
	(550.05, 64.82) --
	(557.59, 64.82) --
	(565.12, 64.08) --
	(565.12, 56.77) --
	(557.59, 57.72) --
	(550.05, 57.72) --
	(542.51, 58.18) --
	(534.97, 58.18) --
	(527.44, 58.40) --
	(519.90, 58.16) --
	(512.36, 58.79) --
	(504.82, 60.03) --
	(497.29, 60.94) --
	(489.75, 62.46) --
	(482.21, 65.05) --
	(474.68, 66.78) --
	(467.14, 70.56) --
	(459.60, 73.93) --
	(452.06, 74.95) --
	(444.53, 77.75) --
	(436.99, 84.76) --
	(429.45, 84.17) --
	(421.91, 90.54) --
	(414.38, 89.60) --
	(406.84, 93.27) --
	(399.30,129.71) --
	(391.77,137.14) --
	(384.23,136.34) --
	(376.69,134.94) --
	(369.15,135.04) --
	(361.62,120.43) --
	(354.08,124.23) --
	(346.54,119.49) --
	cycle;
\definecolor{drawColor}{RGB}{0,114,178}
\definecolor{fillColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (346.54,167.99) --
	(354.08,158.37) --
	(361.62,155.61) --
	(369.15,153.41) --
	(376.69,158.49) --
	(384.23,154.03) --
	(391.77,151.93) --
	(399.30,148.30) --
	(406.84,107.71) --
	(414.38, 96.60) --
	(421.91, 94.03) --
	(429.45, 86.26) --
	(436.99, 84.52) --
	(444.53, 77.77) --
	(452.06, 75.45) --
	(459.60, 73.98) --
	(467.14, 69.74) --
	(474.68, 67.73) --
	(482.21, 65.35) --
	(489.75, 64.41) --
	(497.29, 62.76) --
	(504.82, 61.94) --
	(512.36, 60.56) --
	(519.90, 60.51) --
	(527.44, 60.01) --
	(534.97, 59.88) --
	(542.51, 59.88) --
	(550.05, 59.77) --
	(557.59, 59.77) --
	(565.12, 59.40) --
	(565.12, 54.49) --
	(557.59, 54.89) --
	(550.05, 54.89) --
	(542.51, 55.23) --
	(534.97, 55.23) --
	(527.44, 55.32) --
	(519.90, 55.43) --
	(512.36, 55.76) --
	(504.82, 56.48) --
	(497.29, 57.32) --
	(489.75, 58.34) --
	(482.21, 59.85) --
	(474.68, 61.35) --
	(467.14, 63.63) --
	(459.60, 66.52) --
	(452.06, 66.97) --
	(444.53, 68.74) --
	(436.99, 73.15) --
	(429.45, 73.04) --
	(421.91, 77.40) --
	(414.38, 77.19) --
	(406.84, 79.82) --
	(399.30,101.35) --
	(391.77,108.09) --
	(384.23,107.05) --
	(376.69,105.61) --
	(369.15,105.29) --
	(361.62, 96.62) --
	(354.08, 98.80) --
	(346.54, 95.78) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (316.39, 41.40) -- (572.66, 41.40);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (346.54,170.94) --
	(354.08,170.79) --
	(361.62,163.30) --
	(369.15,167.34) --
	(376.69,172.68) --
	(384.23,172.43) --
	(391.77,174.86) --
	(399.30,167.73) --
	(406.84,112.65) --
	(414.38,105.04) --
	(421.91,102.84) --
	(429.45, 93.38) --
	(436.99, 91.84) --
	(444.53, 83.99) --
	(452.06, 80.95) --
	(459.60, 80.00) --
	(467.14, 75.18) --
	(474.68, 71.77) --
	(482.21, 68.56) --
	(489.75, 66.30) --
	(497.29, 65.11) --
	(504.82, 63.71) --
	(512.36, 62.37) --
	(519.90, 61.91) --
	(527.44, 61.69) --
	(534.97, 61.52) --
	(542.51, 61.52) --
	(550.05, 61.33) --
	(557.59, 61.33) --
	(565.12, 60.65);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (346.54,126.39) --
	(354.08,125.79) --
	(361.62,122.09) --
	(369.15,124.37) --
	(376.69,127.48) --
	(384.23,126.99) --
	(391.77,128.28) --
	(399.30,123.96) --
	(406.84, 92.40) --
	(414.38, 87.06) --
	(421.91, 85.30) --
	(429.45, 79.23) --
	(436.99, 77.99) --
	(444.53, 73.04) --
	(452.06, 70.90) --
	(459.60, 70.16) --
	(467.14, 66.88) --
	(474.68, 64.57) --
	(482.21, 62.37) --
	(489.75, 60.94) --
	(497.29, 60.13) --
	(504.82, 59.14) --
	(512.36, 58.16) --
	(519.90, 57.82) --
	(527.44, 57.64) --
	(534.97, 57.52) --
	(542.51, 57.52) --
	(550.05, 57.39) --
	(557.59, 57.39) --
	(565.12, 56.99);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.39, 30.69) rectangle (572.66,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44, 38.37) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44, 90.38) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,142.40) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,194.41) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,246.42) {5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64, 41.40) --
	(316.39, 41.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64, 93.41) --
	(316.39, 93.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,145.43) --
	(316.39,145.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,197.44) --
	(316.39,197.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,249.45) --
	(316.39,249.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (346.54, 27.94) --
	(346.54, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (421.91, 27.94) --
	(421.91, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (497.29, 27.94) --
	(497.29, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (572.66, 27.94) --
	(572.66, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (346.54, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (421.91, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (497.29, 19.68) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (572.66, 19.68) {30};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (444.53,  7.64) {Cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,148.55) {Reproduction number $\mathcal R$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (450.83,208.36) rectangle (566.36,253.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (456.33,228.31) rectangle (470.78,242.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{240,228,66}
\definecolor{fillColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor,fill opacity=0.50] (457.04,229.02) rectangle (470.07,242.05);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (457.77,235.54) -- (469.33,235.54);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (456.33,213.86) rectangle (470.78,228.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,114,178}
\definecolor{fillColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line cap=rect,fill=fillColor,fill opacity=0.50] (457.04,214.57) rectangle (470.07,227.60);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (457.77,221.09) -- (469.33,221.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (476.28,232.51) {6 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (476.28,218.05) {8 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (316.39,274.49) {B};
\end{scope}
\end{tikzpicture}

\end{document}