### 2020-04-14 serial_guide.tex

% Created by tikzDevice version 0.12.3 on 2020-04-14 02:04:29
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,433.62);
\begin{scope}
\path[clip] (  0.00,216.81) rectangle (722.70,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,216.81) rectangle (722.70,433.62);
\end{scope}
\begin{scope}
\path[clip] (  8.25,225.06) rectangle (717.20,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (  8.25,225.06) rectangle (717.20,410.96);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04,275.76) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.16,275.76) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (443.29,275.76) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (604.41,275.76) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04,332.09) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.16,388.43) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (443.29,360.26) circle (  3.57);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (282.16,257.88) {$\ell$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (121.04,257.88) {$\ell-x_0$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (121.04,243.79) {Infector};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (282.16,243.79) {Primary cohort};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,243.79) {Infectee};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,257.88) {$\ell-x_0+\sigma$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (604.41,257.88) {$\ell-x_0+\sigma+x_1$};

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 40.48,275.76) -- (684.98,275.76);

\path[draw=drawColor,line width= 1.1pt,line join=round] (672.46,268.53) --
	(684.98,275.76) --
	(672.46,282.99);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (121.04,275.76) -- (121.04,332.09);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (282.16,275.76) -- (282.16,388.43);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (443.29,275.76) -- (443.29,360.26);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (604.41,275.76) -- (604.41,388.43);

\path[draw=drawColor,line width= 1.1pt,line join=round] (282.16,303.93) -- (121.04,303.93);

\path[draw=drawColor,line width= 1.1pt,line join=round] (127.30,307.54) --
	(121.04,303.93) --
	(127.30,300.31);

\path[draw=drawColor,line width= 1.1pt,line join=round] (121.04,332.09) -- (443.29,332.09);

\path[draw=drawColor,line width= 1.1pt,line join=round] (437.03,328.48) --
	(443.29,332.09) --
	(437.03,335.71);

\path[draw=drawColor,line width= 1.1pt,line join=round] (443.29,360.26) -- (604.41,360.26);

\path[draw=drawColor,line width= 1.1pt,line join=round] (598.15,356.65) --
	(604.41,360.26) --
	(598.15,363.88);

\path[draw=drawColor,line width= 1.1pt,line join=round] (282.16,388.43) -- (604.41,388.43);

\path[draw=drawColor,line width= 1.1pt,line join=round] (598.15,384.82) --
	(604.41,388.43) --
	(598.15,392.04);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,398.71) {Forward serial interval};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (201.60,314.21) {Backward incubation period};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (201.60,342.38) {Forward generation interval};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (523.85,370.54) {Forward incubation period};
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round,fill=fillColor] (282.16,275.76) circle (  5.71);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round,fill=fillColor] (604.41,275.76) circle (  5.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  8.25,419.03) {A};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,216.81);
\end{scope}
\begin{scope}
\path[clip] (  8.25,  8.25) rectangle (717.20,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (  8.25,  8.25) rectangle (717.20,194.15);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04, 58.95) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.16, 58.95) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (443.29, 58.95) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (604.41, 58.95) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04,143.45) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (604.41,171.62) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (443.29,115.28) circle (  3.57);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (604.41, 41.07) {$\ell$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29, 41.07) {$\ell-x_1$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (121.04, 41.07) {$\ell-x_1-\sigma$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (282.16, 41.07) {$\ell-x_1-\sigma+x_0$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (121.04, 26.98) {Infector};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29, 26.98) {Infectee};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (604.41, 26.98) {Secondary cohort};

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 40.48, 58.95) -- (684.98, 58.95);

\path[draw=drawColor,line width= 1.1pt,line join=round] (672.46, 51.72) --
	(684.98, 58.95) --
	(672.46, 66.18);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (121.04, 58.95) -- (121.04,143.45);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (282.16, 58.95) -- (282.16,171.62);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (443.29, 58.95) -- (443.29,115.28);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (604.41, 58.95) -- (604.41,171.62);

\path[draw=drawColor,line width= 1.1pt,line join=round] (604.41, 87.12) -- (443.29, 87.12);

\path[draw=drawColor,line width= 1.1pt,line join=round] (449.55, 90.73) --
	(443.29, 87.12) --
	(449.55, 83.50);

\path[draw=drawColor,line width= 1.1pt,line join=round] (443.29,115.28) -- (121.04,115.28);

\path[draw=drawColor,line width= 1.1pt,line join=round] (127.30,118.90) --
	(121.04,115.28) --
	(127.30,111.67);

\path[draw=drawColor,line width= 1.1pt,line join=round] (121.04,143.45) -- (282.16,143.45);

\path[draw=drawColor,line width= 1.1pt,line join=round] (275.90,139.84) --
	(282.16,143.45) --
	(275.90,147.07);

\path[draw=drawColor,line width= 1.1pt,line join=round] (604.41,171.62) -- (282.16,171.62);

\path[draw=drawColor,line width= 1.1pt,line join=round] (288.42,175.23) --
	(282.16,171.62) --
	(288.42,168.01);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,181.90) {Backward serial interval};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (523.85, 97.40) {Backward incubation period};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (362.72,125.57) {Backward generation interval};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (201.60,153.73) {Forward incubation period};
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round,fill=fillColor] (604.41, 58.95) circle (  5.71);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round,fill=fillColor] (282.16, 58.95) circle (  5.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  8.25,202.22) {B};
\end{scope}
\end{tikzpicture}

\end{document}