### 2020-04-14 rR.tex
there's one thing that I'm missing...huh...

% Created by tikzDevice version 0.12.3 on 2020-04-14 11:43:25
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,433.62);
\begin{scope}
\path[clip] (  0.00,216.81) rectangle (289.08,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,216.81) rectangle (289.08,433.62);
\end{scope}
\begin{scope}
\path[clip] ( 27.31,247.50) rectangle (283.58,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31,247.50) rectangle (283.58,410.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31,247.65) --
	( 30.56,248.50) --
	( 33.80,249.37) --
	( 37.04,250.25) --
	( 40.29,251.16) --
	( 43.53,252.09) --
	( 46.78,253.03) --
	( 50.02,254.00) --
	( 53.26,254.98) --
	( 56.51,255.98) --
	( 59.75,257.00) --
	( 63.00,258.04) --
	( 66.24,259.10) --
	( 69.48,260.18) --
	( 72.73,261.28) --
	( 75.97,262.40) --
	( 79.22,263.54) --
	( 82.46,264.70) --
	( 85.70,265.88) --
	( 88.95,267.07) --
	( 92.19,268.29) --
	( 95.43,269.52) --
	( 98.68,270.77) --
	(101.92,272.04) --
	(105.17,273.33) --
	(108.41,274.63) --
	(111.65,275.96) --
	(114.90,277.31) --
	(118.14,278.69) --
	(121.39,280.08) --
	(124.63,281.50) --
	(127.87,282.94) --
	(131.12,284.39) --
	(134.36,285.86) --
	(137.61,287.35) --
	(140.85,288.87) --
	(144.09,290.41) --
	(147.34,291.97) --
	(150.58,293.56) --
	(153.82,295.18) --
	(157.07,296.83) --
	(160.31,298.49) --
	(163.56,300.18) --
	(166.80,301.88) --
	(170.04,303.60) --
	(173.29,305.35) --
	(176.53,307.13) --
	(179.78,308.93) --
	(183.02,310.77) --
	(186.26,312.64) --
	(189.51,314.55) --
	(192.75,316.48) --
	(196.00,318.44) --
	(199.24,320.43) --
	(202.48,322.45) --
	(205.73,324.49) --
	(208.97,326.57) --
	(212.21,328.67) --
	(215.46,330.81) --
	(218.70,332.98) --
	(221.95,335.19) --
	(225.19,337.43) --
	(228.43,339.70) --
	(231.68,342.00) --
	(234.92,344.33) --
	(238.17,346.69) --
	(241.41,349.09) --
	(244.65,351.51) --
	(247.90,353.96) --
	(251.14,356.45) --
	(254.39,358.97) --
	(257.63,361.51) --
	(260.87,364.10) --
	(264.12,366.71) --
	(267.36,369.35) --
	(270.60,372.03) --
	(273.85,374.73) --
	(277.09,377.47) --
	(280.34,380.24) --
	(283.58,383.04);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 27.31,247.62) --
	( 30.56,248.48) --
	( 33.80,249.35) --
	( 37.04,250.25) --
	( 40.29,251.16) --
	( 43.53,252.10) --
	( 46.78,253.05) --
	( 50.02,254.02) --
	( 53.26,255.01) --
	( 56.51,256.02) --
	( 59.75,257.05) --
	( 63.00,258.09) --
	( 66.24,259.16) --
	( 69.48,260.24) --
	( 72.73,261.35) --
	( 75.97,262.47) --
	( 79.22,263.61) --
	( 82.46,264.77) --
	( 85.70,265.96) --
	( 88.95,267.16) --
	( 92.19,268.37) --
	( 95.43,269.61) --
	( 98.68,270.86) --
	(101.92,272.13) --
	(105.17,273.42) --
	(108.41,274.73) --
	(111.65,276.06) --
	(114.90,277.41) --
	(118.14,278.78) --
	(121.39,280.18) --
	(124.63,281.60) --
	(127.87,283.02) --
	(131.12,284.47) --
	(134.36,285.93) --
	(137.61,287.42) --
	(140.85,288.93) --
	(144.09,290.46) --
	(147.34,292.01) --
	(150.58,293.60) --
	(153.82,295.21) --
	(157.07,296.86) --
	(160.31,298.52) --
	(163.56,300.20) --
	(166.80,301.90) --
	(170.04,303.62) --
	(173.29,305.37) --
	(176.53,307.15) --
	(179.78,308.95) --
	(183.02,310.78) --
	(186.26,312.65) --
	(189.51,314.56) --
	(192.75,316.49) --
	(196.00,318.45) --
	(199.24,320.44) --
	(202.48,322.45) --
	(205.73,324.50) --
	(208.97,326.57) --
	(212.21,328.67) --
	(215.46,330.81) --
	(218.70,332.98) --
	(221.95,335.19) --
	(225.19,337.43) --
	(228.43,339.70) --
	(231.68,342.01) --
	(234.92,344.34) --
	(238.17,346.70) --
	(241.41,349.10) --
	(244.65,351.53) --
	(247.90,353.98) --
	(251.14,356.47) --
	(254.39,358.99) --
	(257.63,361.55) --
	(260.87,364.13) --
	(264.12,366.75) --
	(267.36,369.40) --
	(270.60,372.08) --
	(273.85,374.79) --
	(277.09,377.53) --
	(280.34,380.31) --
	(283.58,383.11);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 27.31,247.61) --
	( 30.56,248.47) --
	( 33.80,249.35) --
	( 37.04,250.24) --
	( 40.29,251.16) --
	( 43.53,252.10) --
	( 46.78,253.05) --
	( 50.02,254.02) --
	( 53.26,255.02) --
	( 56.51,256.03) --
	( 59.75,257.05) --
	( 63.00,258.10) --
	( 66.24,259.17) --
	( 69.48,260.25) --
	( 72.73,261.36) --
	( 75.97,262.48) --
	( 79.22,263.63) --
	( 82.46,264.79) --
	( 85.70,265.97) --
	( 88.95,267.17) --
	( 92.19,268.39) --
	( 95.43,269.63) --
	( 98.68,270.88) --
	(101.92,272.15) --
	(105.17,273.44) --
	(108.41,274.75) --
	(111.65,276.08) --
	(114.90,277.43) --
	(118.14,278.80) --
	(121.39,280.20) --
	(124.63,281.61) --
	(127.87,283.04) --
	(131.12,284.48) --
	(134.36,285.94) --
	(137.61,287.42) --
	(140.85,288.92) --
	(144.09,290.45) --
	(147.34,292.01) --
	(150.58,293.59) --
	(153.82,295.20) --
	(157.07,296.84) --
	(160.31,298.50) --
	(163.56,300.18) --
	(166.80,301.87) --
	(170.04,303.59) --
	(173.29,305.33) --
	(176.53,307.10) --
	(179.78,308.90) --
	(183.02,310.74) --
	(186.26,312.60) --
	(189.51,314.51) --
	(192.75,316.44) --
	(196.00,318.40) --
	(199.24,320.39) --
	(202.48,322.41) --
	(205.73,324.45) --
	(208.97,326.53) --
	(212.21,328.64) --
	(215.46,330.78) --
	(218.70,332.95) --
	(221.95,335.17) --
	(225.19,337.42) --
	(228.43,339.70) --
	(231.68,342.01) --
	(234.92,344.35) --
	(238.17,346.73) --
	(241.41,349.13) --
	(244.65,351.57) --
	(247.90,354.04) --
	(251.14,356.54) --
	(254.39,359.07) --
	(257.63,361.64) --
	(260.87,364.24) --
	(264.12,366.87) --
	(267.36,369.54) --
	(270.60,372.24) --
	(273.85,374.97) --
	(277.09,377.73) --
	(280.34,380.52) --
	(283.58,383.35);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31,247.60) --
	( 30.56,248.46) --
	( 33.80,249.34) --
	( 37.04,250.24) --
	( 40.29,251.16) --
	( 43.53,252.09) --
	( 46.78,253.05) --
	( 50.02,254.02) --
	( 53.26,255.02) --
	( 56.51,256.03) --
	( 59.75,257.06) --
	( 63.00,258.11) --
	( 66.24,259.17) --
	( 69.48,260.26) --
	( 72.73,261.37) --
	( 75.97,262.49) --
	( 79.22,263.64) --
	( 82.46,264.80) --
	( 85.70,265.98) --
	( 88.95,267.18) --
	( 92.19,268.40) --
	( 95.43,269.64) --
	( 98.68,270.89) --
	(101.92,272.16) --
	(105.17,273.45) --
	(108.41,274.76) --
	(111.65,276.09) --
	(114.90,277.44) --
	(118.14,278.81) --
	(121.39,280.21) --
	(124.63,281.62) --
	(127.87,283.04) --
	(131.12,284.48) --
	(134.36,285.94) --
	(137.61,287.42) --
	(140.85,288.92) --
	(144.09,290.45) --
	(147.34,292.00) --
	(150.58,293.58) --
	(153.82,295.19) --
	(157.07,296.83) --
	(160.31,298.48) --
	(163.56,300.16) --
	(166.80,301.85) --
	(170.04,303.57) --
	(173.29,305.31) --
	(176.53,307.07) --
	(179.78,308.87) --
	(183.02,310.70) --
	(186.26,312.57) --
	(189.51,314.47) --
	(192.75,316.41) --
	(196.00,318.37) --
	(199.24,320.36) --
	(202.48,322.38) --
	(205.73,324.43) --
	(208.97,326.51) --
	(212.21,328.63) --
	(215.46,330.77) --
	(218.70,332.96) --
	(221.95,335.18) --
	(225.19,337.44) --
	(228.43,339.73) --
	(231.68,342.05) --
	(234.92,344.41) --
	(238.17,346.80) --
	(241.41,349.22) --
	(244.65,351.67) --
	(247.90,354.15) --
	(251.14,356.67) --
	(254.39,359.22) --
	(257.63,361.81) --
	(260.87,364.43) --
	(264.12,367.08) --
	(267.36,369.77) --
	(270.60,372.48) --
	(273.85,375.24) --
	(277.09,378.02) --
	(280.34,380.84) --
	(283.58,383.69);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 27.31,251.38) --
	( 30.68,245.55) --
	( 23.95,245.55) --
	cycle;

\path[fill=fillColor] ( 35.86,253.77) --
	( 39.22,247.94) --
	( 32.49,247.94) --
	cycle;

\path[fill=fillColor] ( 44.40,256.26) --
	( 47.76,250.43) --
	( 41.03,250.43) --
	cycle;

\path[fill=fillColor] ( 52.94,258.86) --
	( 56.30,253.03) --
	( 49.58,253.03) --
	cycle;

\path[fill=fillColor] ( 61.48,261.57) --
	( 64.85,255.75) --
	( 58.12,255.75) --
	cycle;

\path[fill=fillColor] ( 70.02,264.40) --
	( 73.39,258.58) --
	( 66.66,258.58) --
	cycle;

\path[fill=fillColor] ( 78.57,267.35) --
	( 81.93,261.52) --
	( 75.20,261.52) --
	cycle;

\path[fill=fillColor] ( 87.11,270.41) --
	( 90.47,264.59) --
	( 83.75,264.59) --
	cycle;

\path[fill=fillColor] ( 95.65,273.62) --
	( 99.01,267.79) --
	( 92.29,267.79) --
	cycle;

\path[fill=fillColor] (104.19,276.97) --
	(107.56,271.14) --
	(100.83,271.14) --
	cycle;

\path[fill=fillColor] (112.74,280.44) --
	(116.10,274.61) --
	(109.37,274.61) --
	cycle;

\path[fill=fillColor] (121.28,284.03) --
	(124.64,278.21) --
	(117.91,278.21) --
	cycle;

\path[fill=fillColor] (129.82,287.77) --
	(133.18,281.94) --
	(126.46,281.94) --
	cycle;

\path[fill=fillColor] (138.36,291.66) --
	(141.73,285.83) --
	(135.00,285.83) --
	cycle;

\path[fill=fillColor] (146.90,295.73) --
	(150.27,289.90) --
	(143.54,289.90) --
	cycle;

\path[fill=fillColor] (155.45,299.91) --
	(158.81,294.09) --
	(152.08,294.09) --
	cycle;

\path[fill=fillColor] (163.99,304.28) --
	(167.35,298.45) --
	(160.63,298.45) --
	cycle;

\path[fill=fillColor] (172.53,308.80) --
	(175.89,302.97) --
	(169.17,302.97) --
	cycle;

\path[fill=fillColor] (181.07,313.61) --
	(184.44,307.78) --
	(177.71,307.78) --
	cycle;

\path[fill=fillColor] (189.62,318.54) --
	(192.98,312.71) --
	(186.25,312.71) --
	cycle;

\path[fill=fillColor] (198.16,323.65) --
	(201.52,317.83) --
	(194.79,317.83) --
	cycle;

\path[fill=fillColor] (206.70,328.94) --
	(210.06,323.11) --
	(203.34,323.11) --
	cycle;

\path[fill=fillColor] (215.24,334.46) --
	(218.61,328.63) --
	(211.88,328.63) --
	cycle;

\path[fill=fillColor] (223.78,340.25) --
	(227.15,334.42) --
	(220.42,334.42) --
	cycle;

\path[fill=fillColor] (232.33,346.20) --
	(235.69,340.38) --
	(228.96,340.38) --
	cycle;

\path[fill=fillColor] (240.87,352.40) --
	(244.23,346.58) --
	(237.51,346.58) --
	cycle;

\path[fill=fillColor] (249.41,358.80) --
	(252.77,352.98) --
	(246.05,352.98) --
	cycle;

\path[fill=fillColor] (257.95,365.62) --
	(261.32,359.79) --
	(254.59,359.79) --
	cycle;

\path[fill=fillColor] (266.50,372.58) --
	(269.86,366.75) --
	(263.13,366.75) --
	cycle;

\path[fill=fillColor] (275.04,379.67) --
	(278.40,373.84) --
	(271.67,373.84) --
	cycle;

\path[fill=fillColor] (283.58,387.25) --
	(286.94,381.43) --
	(280.22,381.43) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 24.82,245.00) --
	( 29.81,245.00) --
	( 29.81,249.99) --
	( 24.82,249.99) --
	cycle;

\path[fill=fillColor] ( 33.36,247.39) --
	( 38.35,247.39) --
	( 38.35,252.38) --
	( 33.36,252.38) --
	cycle;

\path[fill=fillColor] ( 41.90,249.87) --
	( 46.90,249.87) --
	( 46.90,254.87) --
	( 41.90,254.87) --
	cycle;

\path[fill=fillColor] ( 50.44,252.48) --
	( 55.44,252.48) --
	( 55.44,257.47) --
	( 50.44,257.47) --
	cycle;

\path[fill=fillColor] ( 58.98,255.19) --
	( 63.98,255.19) --
	( 63.98,260.19) --
	( 58.98,260.19) --
	cycle;

\path[fill=fillColor] ( 67.53,258.02) --
	( 72.52,258.02) --
	( 72.52,263.02) --
	( 67.53,263.02) --
	cycle;

\path[fill=fillColor] ( 76.07,260.97) --
	( 81.06,260.97) --
	( 81.06,265.96) --
	( 76.07,265.96) --
	cycle;

\path[fill=fillColor] ( 84.61,264.04) --
	( 89.61,264.04) --
	( 89.61,269.03) --
	( 84.61,269.03) --
	cycle;

\path[fill=fillColor] ( 93.15,267.25) --
	( 98.15,267.25) --
	( 98.15,272.25) --
	( 93.15,272.25) --
	cycle;

\path[fill=fillColor] (101.70,270.62) --
	(106.69,270.62) --
	(106.69,275.61) --
	(101.70,275.61) --
	cycle;

\path[fill=fillColor] (110.24,274.08) --
	(115.23,274.08) --
	(115.23,279.08) --
	(110.24,279.08) --
	cycle;

\path[fill=fillColor] (118.78,277.67) --
	(123.78,277.67) --
	(123.78,282.67) --
	(118.78,282.67) --
	cycle;

\path[fill=fillColor] (127.32,281.40) --
	(132.32,281.40) --
	(132.32,286.40) --
	(127.32,286.40) --
	cycle;

\path[fill=fillColor] (135.86,285.29) --
	(140.86,285.29) --
	(140.86,290.28) --
	(135.86,290.28) --
	cycle;

\path[fill=fillColor] (144.41,289.36) --
	(149.40,289.36) --
	(149.40,294.36) --
	(144.41,294.36) --
	cycle;

\path[fill=fillColor] (152.95,293.52) --
	(157.94,293.52) --
	(157.94,298.51) --
	(152.95,298.51) --
	cycle;

\path[fill=fillColor] (161.49,297.85) --
	(166.49,297.85) --
	(166.49,302.85) --
	(161.49,302.85) --
	cycle;

\path[fill=fillColor] (170.03,302.34) --
	(175.03,302.34) --
	(175.03,307.33) --
	(170.03,307.33) --
	cycle;

\path[fill=fillColor] (178.58,307.20) --
	(183.57,307.20) --
	(183.57,312.19) --
	(178.58,312.19) --
	cycle;

\path[fill=fillColor] (187.12,312.13) --
	(192.11,312.13) --
	(192.11,317.12) --
	(187.12,317.12) --
	cycle;

\path[fill=fillColor] (195.66,317.25) --
	(200.66,317.25) --
	(200.66,322.24) --
	(195.66,322.24) --
	cycle;

\path[fill=fillColor] (204.20,322.50) --
	(209.20,322.50) --
	(209.20,327.49) --
	(204.20,327.49) --
	cycle;

\path[fill=fillColor] (212.74,328.04) --
	(217.74,328.04) --
	(217.74,333.04) --
	(212.74,333.04) --
	cycle;

\path[fill=fillColor] (221.29,333.85) --
	(226.28,333.85) --
	(226.28,338.85) --
	(221.29,338.85) --
	cycle;

\path[fill=fillColor] (229.83,339.77) --
	(234.82,339.77) --
	(234.82,344.76) --
	(229.83,344.76) --
	cycle;

\path[fill=fillColor] (238.37,345.99) --
	(243.37,345.99) --
	(243.37,350.98) --
	(238.37,350.98) --
	cycle;

\path[fill=fillColor] (246.91,352.43) --
	(251.91,352.43) --
	(251.91,357.43) --
	(246.91,357.43) --
	cycle;

\path[fill=fillColor] (255.46,359.42) --
	(260.45,359.42) --
	(260.45,364.42) --
	(255.46,364.42) --
	cycle;

\path[fill=fillColor] (264.00,366.41) --
	(268.99,366.41) --
	(268.99,371.40) --
	(264.00,371.40) --
	cycle;

\path[fill=fillColor] (272.54,373.42) --
	(277.54,373.42) --
	(277.54,378.41) --
	(272.54,378.41) --
	cycle;

\path[fill=fillColor] (281.08,381.10) --
	(286.08,381.10) --
	(286.08,386.09) --
	(281.08,386.09) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 23.78,247.50) -- ( 30.85,247.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.31,243.96) -- ( 27.31,251.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.32,249.88) -- ( 39.39,249.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.86,246.35) -- ( 35.86,253.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87,252.37) -- ( 47.93,252.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.40,248.83) -- ( 44.40,255.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.41,254.97) -- ( 56.47,254.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.94,251.44) -- ( 52.94,258.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.95,257.69) -- ( 65.01,257.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.48,254.15) -- ( 61.48,261.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.49,260.52) -- ( 73.56,260.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.02,256.98) -- ( 70.02,264.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.03,263.46) -- ( 82.10,263.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.57,259.93) -- ( 78.57,266.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.58,266.54) -- ( 90.64,266.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.11,263.01) -- ( 87.11,270.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.12,269.76) -- ( 99.18,269.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65,266.23) -- ( 95.65,273.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.66,273.14) -- (107.73,273.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.19,269.61) -- (104.19,276.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.20,276.60) -- (116.27,276.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.74,273.07) -- (112.74,280.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.75,280.18) -- (124.81,280.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.28,276.65) -- (121.28,283.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.29,283.91) -- (133.35,283.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.82,280.38) -- (129.82,287.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.83,287.80) -- (141.89,287.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.36,284.26) -- (138.36,291.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.37,291.88) -- (150.44,291.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.90,288.34) -- (146.90,295.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.91,296.01) -- (158.98,296.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (155.45,292.47) -- (155.45,299.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.46,300.31) -- (167.52,300.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.99,296.78) -- (163.99,303.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.00,304.77) -- (176.06,304.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.53,301.23) -- (172.53,308.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.54,309.68) -- (184.61,309.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.07,306.15) -- (181.07,313.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.08,314.61) -- (193.15,314.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.62,311.08) -- (189.62,318.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.63,319.74) -- (201.69,319.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.16,316.21) -- (198.16,323.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.17,324.96) -- (210.23,324.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.70,321.43) -- (206.70,328.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.71,330.54) -- (218.77,330.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.24,327.01) -- (215.24,334.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.25,336.37) -- (227.32,336.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.78,332.83) -- (223.78,339.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.79,342.23) -- (235.86,342.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.33,338.70) -- (232.33,345.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.34,348.49) -- (244.40,348.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.87,344.96) -- (240.87,352.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.88,355.00) -- (252.94,355.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.41,351.47) -- (249.41,358.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.42,362.19) -- (261.49,362.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.95,358.66) -- (257.95,365.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.96,369.22) -- (270.03,369.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (266.50,365.69) -- (266.50,372.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.51,376.15) -- (278.57,376.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.04,372.62) -- (275.04,379.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.05,383.92) -- (287.11,383.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.58,380.39) -- (283.58,387.45);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 27.31,247.50) circle (  2.50);

\path[fill=fillColor] ( 35.86,249.88) circle (  2.50);

\path[fill=fillColor] ( 44.40,252.37) circle (  2.50);

\path[fill=fillColor] ( 52.94,254.94) circle (  2.50);

\path[fill=fillColor] ( 61.48,257.68) circle (  2.50);

\path[fill=fillColor] ( 70.02,260.47) circle (  2.50);

\path[fill=fillColor] ( 78.57,263.41) circle (  2.50);

\path[fill=fillColor] ( 87.11,266.50) circle (  2.50);

\path[fill=fillColor] ( 95.65,269.65) circle (  2.50);

\path[fill=fillColor] (104.19,273.00) circle (  2.50);

\path[fill=fillColor] (112.74,276.42) circle (  2.50);

\path[fill=fillColor] (121.28,279.99) circle (  2.50);

\path[fill=fillColor] (129.82,283.76) circle (  2.50);

\path[fill=fillColor] (138.36,287.74) circle (  2.50);

\path[fill=fillColor] (146.90,291.70) circle (  2.50);

\path[fill=fillColor] (155.45,296.03) circle (  2.50);

\path[fill=fillColor] (163.99,300.46) circle (  2.50);

\path[fill=fillColor] (172.53,304.98) circle (  2.50);

\path[fill=fillColor] (181.07,309.73) circle (  2.50);

\path[fill=fillColor] (189.62,314.62) circle (  2.50);

\path[fill=fillColor] (198.16,319.79) circle (  2.50);

\path[fill=fillColor] (206.70,324.93) circle (  2.50);

\path[fill=fillColor] (215.24,330.66) circle (  2.50);

\path[fill=fillColor] (223.78,336.18) circle (  2.50);

\path[fill=fillColor] (232.33,342.09) circle (  2.50);

\path[fill=fillColor] (240.87,348.75) circle (  2.50);

\path[fill=fillColor] (249.41,355.20) circle (  2.50);

\path[fill=fillColor] (257.95,361.58) circle (  2.50);

\path[fill=fillColor] (266.50,368.62) circle (  2.50);

\path[fill=fillColor] (275.04,375.76) circle (  2.50);

\path[fill=fillColor] (283.58,383.22) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31,247.50) rectangle (283.58,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,247.50) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,294.20) --
	( 27.31,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,340.91) --
	( 27.31,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,387.61) --
	( 27.31,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,244.75) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.74,244.75) --
	(112.74,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.16,244.75) --
	(198.16,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.74,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.16,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (155.45,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,329.23) {Reproduction number $\mathcal R$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 39.98,332.94) rectangle (168.41,407.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,381.80) rectangle ( 59.93,396.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 46.93,389.03) -- ( 58.49,389.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 52.71,389.03) circle (  2.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,367.35) rectangle ( 59.93,381.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 46.93,374.57) -- ( 58.49,374.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 52.71,378.46) --
	( 56.07,372.63) --
	( 49.34,372.63) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,352.89) rectangle ( 59.93,367.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 46.93,360.12) -- ( 58.49,360.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 50.21,357.62) --
	( 55.20,357.62) --
	( 55.20,362.62) --
	( 50.21,362.62) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,338.44) rectangle ( 59.93,352.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 46.93,345.67) -- ( 58.49,345.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.17,345.67) -- ( 56.24,345.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.71,342.13) -- ( 52.71,349.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,386.00) {Generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,371.54) {Serial interval ($\rho=0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,357.09) {Serial interval ($\rho=0$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,342.63) {Serial interval ($\rho=-0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,419.03) {A};
\end{scope}
\begin{scope}
\path[clip] (289.08,216.81) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,216.81) rectangle (578.16,433.62);
\end{scope}
\begin{scope}
\path[clip] (316.39,247.50) rectangle (572.66,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (316.39,247.50) rectangle (572.66,410.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (316.39,247.65) --
	(319.64,248.49) --
	(322.88,249.36) --
	(326.12,250.25) --
	(329.37,251.16) --
	(332.61,252.08) --
	(335.86,253.03) --
	(339.10,254.00) --
	(342.34,254.98) --
	(345.59,255.98) --
	(348.83,257.00) --
	(352.08,258.04) --
	(355.32,259.10) --
	(358.56,260.18) --
	(361.81,261.28) --
	(365.05,262.40) --
	(368.30,263.54) --
	(371.54,264.70) --
	(374.78,265.88) --
	(378.03,267.07) --
	(381.27,268.29) --
	(384.51,269.52) --
	(387.76,270.77) --
	(391.00,272.04) --
	(394.25,273.32) --
	(397.49,274.63) --
	(400.73,275.96) --
	(403.98,277.31) --
	(407.22,278.68) --
	(410.47,280.08) --
	(413.71,281.49) --
	(416.95,282.93) --
	(420.20,284.38) --
	(423.44,285.85) --
	(426.69,287.34) --
	(429.93,288.85) --
	(433.17,290.39) --
	(436.42,291.96) --
	(439.66,293.55) --
	(442.90,295.17) --
	(446.15,296.82) --
	(449.39,298.48) --
	(452.64,300.17) --
	(455.88,301.88) --
	(459.12,303.61) --
	(462.37,305.37) --
	(465.61,307.15) --
	(468.86,308.96) --
	(472.10,310.80) --
	(475.34,312.67) --
	(478.59,314.58) --
	(481.83,316.51) --
	(485.08,318.47) --
	(488.32,320.46) --
	(491.56,322.47) --
	(494.81,324.51) --
	(498.05,326.58) --
	(501.29,328.68) --
	(504.54,330.81) --
	(507.78,332.98) --
	(511.03,335.19) --
	(514.27,337.43) --
	(517.51,339.70) --
	(520.76,342.00) --
	(524.00,344.33) --
	(527.25,346.69) --
	(530.49,349.09) --
	(533.73,351.51) --
	(536.98,353.97) --
	(540.22,356.45) --
	(543.47,358.97) --
	(546.71,361.52) --
	(549.95,364.10) --
	(553.20,366.72) --
	(556.44,369.36) --
	(559.68,372.04) --
	(562.93,374.75) --
	(566.17,377.49) --
	(569.42,380.26) --
	(572.66,383.06);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (319.64,248.30) --
	(322.88,249.24) --
	(326.12,250.18) --
	(329.37,251.12) --
	(332.61,252.08) --
	(335.86,253.04) --
	(339.10,254.00) --
	(342.34,254.98) --
	(345.59,255.96) --
	(348.83,256.94) --
	(352.08,257.94) --
	(355.32,258.94) --
	(358.56,259.94) --
	(361.81,260.95) --
	(365.05,261.97) --
	(368.30,263.00) --
	(371.54,264.03) --
	(374.78,265.06) --
	(378.03,266.11) --
	(381.27,267.16) --
	(384.51,268.22) --
	(387.76,269.28) --
	(391.00,270.36) --
	(394.25,271.44) --
	(397.49,272.53) --
	(400.73,273.62) --
	(403.98,274.72) --
	(407.22,275.82) --
	(410.47,276.92) --
	(413.71,278.03) --
	(416.95,279.15) --
	(420.20,280.28) --
	(423.44,281.42) --
	(426.69,282.56) --
	(429.93,283.71) --
	(433.17,284.85) --
	(436.42,286.00) --
	(439.66,287.14) --
	(442.90,288.28) --
	(446.15,289.41) --
	(449.39,290.55) --
	(452.64,291.70) --
	(455.88,292.86) --
	(459.12,294.02) --
	(462.37,295.18) --
	(465.61,296.34) --
	(468.86,297.49) --
	(472.10,298.64) --
	(475.34,299.77) --
	(478.59,300.88) --
	(481.83,301.99) --
	(485.08,303.10) --
	(488.32,304.20) --
	(491.56,305.31) --
	(494.81,306.40) --
	(498.05,307.49) --
	(501.29,308.57) --
	(504.54,309.63) --
	(507.78,310.68) --
	(511.03,311.72) --
	(514.27,312.73) --
	(517.51,313.74) --
	(520.76,314.73) --
	(524.00,315.72) --
	(527.25,316.70) --
	(530.49,317.67) --
	(533.73,318.63) --
	(536.98,319.58) --
	(540.22,320.52) --
	(543.47,321.45) --
	(546.71,322.36) --
	(549.95,323.27) --
	(553.20,324.16) --
	(556.44,325.04) --
	(559.68,325.91) --
	(562.93,326.77) --
	(566.17,327.62) --
	(569.42,328.46) --
	(572.66,329.29);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (319.64,248.18) --
	(322.88,249.16) --
	(326.12,250.13) --
	(329.37,251.09) --
	(332.61,252.06) --
	(335.86,253.02) --
	(339.10,253.99) --
	(342.34,254.95) --
	(345.59,255.91) --
	(348.83,256.87) --
	(352.08,257.83) --
	(355.32,258.78) --
	(358.56,259.74) --
	(361.81,260.69) --
	(365.05,261.64) --
	(368.30,262.58) --
	(371.54,263.53) --
	(374.78,264.47) --
	(378.03,265.40) --
	(381.27,266.35) --
	(384.51,267.29) --
	(387.76,268.24) --
	(391.00,269.18) --
	(394.25,270.13) --
	(397.49,271.07) --
	(400.73,272.00) --
	(403.98,272.93) --
	(407.22,273.85) --
	(410.47,274.75) --
	(413.71,275.66) --
	(416.95,276.58) --
	(420.20,277.50) --
	(423.44,278.42) --
	(426.69,279.33) --
	(429.93,280.24) --
	(433.17,281.13) --
	(436.42,282.00) --
	(439.66,282.86) --
	(442.90,283.68) --
	(446.15,284.48) --
	(449.39,285.28) --
	(452.64,286.08) --
	(455.88,286.88) --
	(459.12,287.67) --
	(462.37,288.45) --
	(465.61,289.20) --
	(468.86,289.93) --
	(472.10,290.62) --
	(475.34,291.28) --
	(478.59,291.89) --
	(481.83,292.48) --
	(485.08,293.04) --
	(488.32,293.59) --
	(491.56,294.11) --
	(494.81,294.61) --
	(498.05,295.08) --
	(501.29,295.51) --
	(504.54,295.91) --
	(507.78,296.26) --
	(511.03,296.58) --
	(514.27,296.84) --
	(517.51,297.07) --
	(520.76,297.27) --
	(524.00,297.44) --
	(527.25,297.58) --
	(530.49,297.68) --
	(533.73,297.76) --
	(536.98,297.79) --
	(540.22,297.80) --
	(543.47,297.77) --
	(546.71,297.70) --
	(549.95,297.60) --
	(553.20,297.46) --
	(556.44,297.29) --
	(559.68,297.08) --
	(562.93,296.84) --
	(566.17,296.56) --
	(569.42,296.25) --
	(572.66,295.91);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (319.64,248.10) --
	(322.88,249.10) --
	(326.12,250.09) --
	(329.37,251.07) --
	(332.61,252.04) --
	(335.86,253.00) --
	(339.10,253.95) --
	(342.34,254.90) --
	(345.59,255.84) --
	(348.83,256.77) --
	(352.08,257.69) --
	(355.32,258.61) --
	(358.56,259.51) --
	(361.81,260.41) --
	(365.05,261.29) --
	(368.30,262.17) --
	(371.54,263.03) --
	(374.78,263.89) --
	(378.03,264.73) --
	(381.27,265.58) --
	(384.51,266.42) --
	(387.76,267.26) --
	(391.00,268.09) --
	(394.25,268.91) --
	(397.49,269.72) --
	(400.73,270.52) --
	(403.98,271.30) --
	(407.22,272.06) --
	(410.47,272.80) --
	(413.71,273.54) --
	(416.95,274.28) --
	(420.20,275.02) --
	(423.44,275.74) --
	(426.69,276.46) --
	(429.93,277.16) --
	(433.17,277.83) --
	(436.42,278.48) --
	(439.66,279.09) --
	(442.90,279.66) --
	(446.15,280.19) --
	(449.39,280.71) --
	(452.64,281.23) --
	(455.88,281.73) --
	(459.12,282.22) --
	(462.37,282.68) --
	(465.61,283.10) --
	(468.86,283.49) --
	(472.10,283.84) --
	(475.34,284.13) --
	(478.59,284.37) --
	(481.83,284.56) --
	(485.08,284.73) --
	(488.32,284.86) --
	(491.56,284.96) --
	(494.81,285.02) --
	(498.05,285.04) --
	(501.29,285.02) --
	(504.54,284.95) --
	(507.78,284.83) --
	(511.03,284.65) --
	(514.27,284.42) --
	(517.51,284.15) --
	(520.76,283.83) --
	(524.00,283.47) --
	(527.25,283.07) --
	(530.49,282.63) --
	(533.73,282.14) --
	(536.98,281.61) --
	(540.22,281.03) --
	(543.47,280.41) --
	(546.71,279.74) --
	(549.95,279.03) --
	(553.20,278.27) --
	(556.44,277.46) --
	(559.68,276.61) --
	(562.93,275.72) --
	(566.17,274.78) --
	(569.42,273.79) --
	(572.66,272.77);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] (316.39,251.38) --
	(319.76,245.55) --
	(313.03,245.55) --
	cycle;

\path[fill=fillColor] (324.94,253.76) --
	(328.30,247.93) --
	(321.57,247.93) --
	cycle;

\path[fill=fillColor] (333.48,256.20) --
	(336.84,250.37) --
	(330.11,250.37) --
	cycle;

\path[fill=fillColor] (342.02,258.70) --
	(345.38,252.88) --
	(338.66,252.88) --
	cycle;

\path[fill=fillColor] (350.56,261.27) --
	(353.93,255.45) --
	(347.20,255.45) --
	cycle;

\path[fill=fillColor] (359.10,263.90) --
	(362.47,258.07) --
	(355.74,258.07) --
	cycle;

\path[fill=fillColor] (367.65,266.58) --
	(371.01,260.76) --
	(364.28,260.76) --
	cycle;

\path[fill=fillColor] (376.19,269.32) --
	(379.55,263.50) --
	(372.83,263.50) --
	cycle;

\path[fill=fillColor] (384.73,272.11) --
	(388.09,266.29) --
	(381.37,266.29) --
	cycle;

\path[fill=fillColor] (393.27,274.95) --
	(396.64,269.13) --
	(389.91,269.13) --
	cycle;

\path[fill=fillColor] (401.82,277.83) --
	(405.18,272.01) --
	(398.45,272.01) --
	cycle;

\path[fill=fillColor] (410.36,280.75) --
	(413.72,274.93) --
	(406.99,274.93) --
	cycle;

\path[fill=fillColor] (418.90,283.71) --
	(422.26,277.88) --
	(415.54,277.88) --
	cycle;

\path[fill=fillColor] (427.44,286.69) --
	(430.81,280.87) --
	(424.08,280.87) --
	cycle;

\path[fill=fillColor] (435.98,289.70) --
	(439.35,283.87) --
	(432.62,283.87) --
	cycle;

\path[fill=fillColor] (444.53,292.72) --
	(447.89,286.89) --
	(441.16,286.89) --
	cycle;

\path[fill=fillColor] (453.07,295.74) --
	(456.43,289.92) --
	(449.71,289.92) --
	cycle;

\path[fill=fillColor] (461.61,298.77) --
	(464.97,292.94) --
	(458.25,292.94) --
	cycle;

\path[fill=fillColor] (470.15,301.79) --
	(473.52,295.96) --
	(466.79,295.96) --
	cycle;

\path[fill=fillColor] (478.70,304.79) --
	(482.06,298.96) --
	(475.33,298.96) --
	cycle;

\path[fill=fillColor] (487.24,307.75) --
	(490.60,301.93) --
	(483.87,301.93) --
	cycle;

\path[fill=fillColor] (495.78,310.68) --
	(499.14,304.85) --
	(492.42,304.85) --
	cycle;

\path[fill=fillColor] (504.32,313.55) --
	(507.69,307.73) --
	(500.96,307.73) --
	cycle;

\path[fill=fillColor] (512.86,316.36) --
	(516.23,310.53) --
	(509.50,310.53) --
	cycle;

\path[fill=fillColor] (521.41,319.08) --
	(524.77,313.25) --
	(518.04,313.25) --
	cycle;

\path[fill=fillColor] (529.95,321.70) --
	(533.31,315.88) --
	(526.59,315.88) --
	cycle;

\path[fill=fillColor] (538.49,324.21) --
	(541.85,318.39) --
	(535.13,318.39) --
	cycle;

\path[fill=fillColor] (547.03,326.59) --
	(550.40,320.76) --
	(543.67,320.76) --
	cycle;

\path[fill=fillColor] (555.58,328.81) --
	(558.94,322.98) --
	(552.21,322.98) --
	cycle;

\path[fill=fillColor] (564.12,330.86) --
	(567.48,325.03) --
	(560.75,325.03) --
	cycle;

\path[fill=fillColor] (572.66,332.71) --
	(576.02,326.88) --
	(569.30,326.88) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] (313.90,245.00) --
	(318.89,245.00) --
	(318.89,249.99) --
	(313.90,249.99) --
	cycle;

\path[fill=fillColor] (322.44,247.36) --
	(327.43,247.36) --
	(327.43,252.36) --
	(322.44,252.36) --
	cycle;

\path[fill=fillColor] (330.98,249.77) --
	(335.98,249.77) --
	(335.98,254.76) --
	(330.98,254.76) --
	cycle;

\path[fill=fillColor] (339.52,252.21) --
	(344.52,252.21) --
	(344.52,257.20) --
	(339.52,257.20) --
	cycle;

\path[fill=fillColor] (348.06,254.68) --
	(353.06,254.68) --
	(353.06,259.68) --
	(348.06,259.68) --
	cycle;

\path[fill=fillColor] (356.61,257.18) --
	(361.60,257.18) --
	(361.60,262.17) --
	(356.61,262.17) --
	cycle;

\path[fill=fillColor] (365.15,259.69) --
	(370.14,259.69) --
	(370.14,264.68) --
	(365.15,264.68) --
	cycle;

\path[fill=fillColor] (373.69,262.21) --
	(378.69,262.21) --
	(378.69,267.20) --
	(373.69,267.20) --
	cycle;

\path[fill=fillColor] (382.23,264.73) --
	(387.23,264.73) --
	(387.23,269.73) --
	(382.23,269.73) --
	cycle;

\path[fill=fillColor] (390.78,267.25) --
	(395.77,267.25) --
	(395.77,272.24) --
	(390.78,272.24) --
	cycle;

\path[fill=fillColor] (399.32,269.74) --
	(404.31,269.74) --
	(404.31,274.74) --
	(399.32,274.74) --
	cycle;

\path[fill=fillColor] (407.86,272.21) --
	(412.86,272.21) --
	(412.86,277.21) --
	(407.86,277.21) --
	cycle;

\path[fill=fillColor] (416.40,274.64) --
	(421.40,274.64) --
	(421.40,279.63) --
	(416.40,279.63) --
	cycle;

\path[fill=fillColor] (424.94,277.02) --
	(429.94,277.02) --
	(429.94,282.01) --
	(424.94,282.01) --
	cycle;

\path[fill=fillColor] (433.49,279.33) --
	(438.48,279.33) --
	(438.48,284.33) --
	(433.49,284.33) --
	cycle;

\path[fill=fillColor] (442.03,281.56) --
	(447.02,281.56) --
	(447.02,286.56) --
	(442.03,286.56) --
	cycle;

\path[fill=fillColor] (450.57,283.70) --
	(455.57,283.70) --
	(455.57,288.70) --
	(450.57,288.70) --
	cycle;

\path[fill=fillColor] (459.11,285.74) --
	(464.11,285.74) --
	(464.11,290.73) --
	(459.11,290.73) --
	cycle;

\path[fill=fillColor] (467.66,287.64) --
	(472.65,287.64) --
	(472.65,292.63) --
	(467.66,292.63) --
	cycle;

\path[fill=fillColor] (476.20,289.40) --
	(481.19,289.40) --
	(481.19,294.39) --
	(476.20,294.39) --
	cycle;

\path[fill=fillColor] (484.74,290.99) --
	(489.74,290.99) --
	(489.74,295.98) --
	(484.74,295.98) --
	cycle;

\path[fill=fillColor] (493.28,292.39) --
	(498.28,292.39) --
	(498.28,297.39) --
	(493.28,297.39) --
	cycle;

\path[fill=fillColor] (501.82,293.59) --
	(506.82,293.59) --
	(506.82,298.58) --
	(501.82,298.58) --
	cycle;

\path[fill=fillColor] (510.37,294.55) --
	(515.36,294.55) --
	(515.36,299.55) --
	(510.37,299.55) --
	cycle;

\path[fill=fillColor] (518.91,295.26) --
	(523.90,295.26) --
	(523.90,300.25) --
	(518.91,300.25) --
	cycle;

\path[fill=fillColor] (527.45,295.68) --
	(532.45,295.68) --
	(532.45,300.68) --
	(527.45,300.68) --
	cycle;

\path[fill=fillColor] (535.99,295.80) --
	(540.99,295.80) --
	(540.99,300.79) --
	(535.99,300.79) --
	cycle;

\path[fill=fillColor] (544.54,295.58) --
	(549.53,295.58) --
	(549.53,300.58) --
	(544.54,300.58) --
	cycle;

\path[fill=fillColor] (553.08,295.01) --
	(558.07,295.01) --
	(558.07,300.00) --
	(553.08,300.00) --
	cycle;

\path[fill=fillColor] (561.62,294.05) --
	(566.62,294.05) --
	(566.62,299.05) --
	(561.62,299.05) --
	cycle;

\path[fill=fillColor] (570.16,292.69) --
	(575.16,292.69) --
	(575.16,297.69) --
	(570.16,297.69) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.86,247.50) -- (319.93,247.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.39,243.96) -- (316.39,251.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.40,249.85) -- (328.47,249.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (324.94,246.32) -- (324.94,253.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.95,252.22) -- (337.01,252.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.48,248.69) -- (333.48,255.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.49,254.60) -- (345.55,254.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.02,251.07) -- (342.02,258.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (347.03,256.98) -- (354.09,256.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.56,253.45) -- (350.56,260.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (355.57,259.36) -- (362.64,259.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.10,255.82) -- (359.10,262.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.11,261.71) -- (371.18,261.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.65,258.18) -- (367.65,265.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.66,264.03) -- (379.72,264.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.19,260.50) -- (376.19,267.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.20,266.31) -- (388.26,266.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.73,262.78) -- (384.73,269.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.74,268.54) -- (396.81,268.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.27,265.01) -- (393.27,272.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.28,270.69) -- (405.35,270.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.82,267.16) -- (401.82,274.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.83,272.77) -- (413.89,272.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.36,269.23) -- (410.36,276.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (415.37,274.74) -- (422.43,274.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.90,271.21) -- (418.90,278.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.91,276.60) -- (430.97,276.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.44,273.07) -- (427.44,280.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (432.45,278.33) -- (439.52,278.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.98,274.80) -- (435.98,281.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (440.99,279.91) -- (448.06,279.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.53,276.38) -- (444.53,283.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (449.54,281.32) -- (456.60,281.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.07,277.79) -- (453.07,284.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.08,282.55) -- (465.14,282.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.61,279.02) -- (461.61,286.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (466.62,283.58) -- (473.69,283.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.15,280.04) -- (470.15,287.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.16,284.37) -- (482.23,284.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.70,280.84) -- (478.70,287.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.71,284.92) -- (490.77,284.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.24,281.39) -- (487.24,288.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (492.25,285.20) -- (499.31,285.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.78,281.66) -- (495.78,288.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.79,285.18) -- (507.85,285.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (504.32,281.65) -- (504.32,288.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (509.33,284.85) -- (516.40,284.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (512.86,281.32) -- (512.86,288.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (517.87,284.18) -- (524.94,284.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (521.41,280.65) -- (521.41,287.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.42,283.15) -- (533.48,283.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (529.95,279.62) -- (529.95,286.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (534.96,281.75) -- (542.02,281.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (538.49,278.22) -- (538.49,285.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (543.50,279.96) -- (550.57,279.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.03,276.43) -- (547.03,283.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.04,277.78) -- (559.11,277.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (555.58,274.24) -- (555.58,281.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.59,275.19) -- (567.65,275.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.12,271.66) -- (564.12,278.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13,272.21) -- (576.19,272.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66,268.68) -- (572.66,275.74);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (316.39,247.50) circle (  2.50);

\path[fill=fillColor] (324.94,249.88) circle (  2.50);

\path[fill=fillColor] (333.48,252.36) circle (  2.50);

\path[fill=fillColor] (342.02,254.96) circle (  2.50);

\path[fill=fillColor] (350.56,257.65) circle (  2.50);

\path[fill=fillColor] (359.10,260.48) circle (  2.50);

\path[fill=fillColor] (367.65,263.42) circle (  2.50);

\path[fill=fillColor] (376.19,266.48) circle (  2.50);

\path[fill=fillColor] (384.73,269.65) circle (  2.50);

\path[fill=fillColor] (393.27,273.04) circle (  2.50);

\path[fill=fillColor] (401.82,276.42) circle (  2.50);

\path[fill=fillColor] (410.36,280.04) circle (  2.50);

\path[fill=fillColor] (418.90,283.76) circle (  2.50);

\path[fill=fillColor] (427.44,287.65) circle (  2.50);

\path[fill=fillColor] (435.98,291.72) circle (  2.50);

\path[fill=fillColor] (444.53,296.05) circle (  2.50);

\path[fill=fillColor] (453.07,300.28) circle (  2.50);

\path[fill=fillColor] (461.61,305.03) circle (  2.50);

\path[fill=fillColor] (470.15,309.81) circle (  2.50);

\path[fill=fillColor] (478.70,314.71) circle (  2.50);

\path[fill=fillColor] (487.24,319.74) circle (  2.50);

\path[fill=fillColor] (495.78,324.97) circle (  2.50);

\path[fill=fillColor] (504.32,330.80) circle (  2.50);

\path[fill=fillColor] (512.86,336.10) circle (  2.50);

\path[fill=fillColor] (521.41,342.50) circle (  2.50);

\path[fill=fillColor] (529.95,348.19) circle (  2.50);

\path[fill=fillColor] (538.49,354.89) circle (  2.50);

\path[fill=fillColor] (547.03,361.68) circle (  2.50);

\path[fill=fillColor] (555.58,368.79) circle (  2.50);

\path[fill=fillColor] (564.12,375.90) circle (  2.50);

\path[fill=fillColor] (572.66,383.19) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.39,247.50) rectangle (572.66,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,247.50) --
	(316.39,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,294.20) --
	(316.39,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,340.91) --
	(316.39,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,387.61) --
	(316.39,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.39,244.75) --
	(316.39,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.82,244.75) --
	(401.82,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.24,244.75) --
	(487.24,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (316.39,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (401.82,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (487.24,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (444.53,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,329.23) {Reproduction number $\mathcal R$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (316.39,419.03) {B};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 27.31, 30.69) rectangle (283.58,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31, 30.69) rectangle (283.58,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31, 54.52) --
	( 30.56, 54.54) --
	( 33.80, 54.57) --
	( 37.04, 54.60) --
	( 40.29, 54.62) --
	( 43.53, 54.64) --
	( 46.78, 54.67) --
	( 50.02, 54.69) --
	( 53.26, 54.70) --
	( 56.51, 54.72) --
	( 59.75, 54.73) --
	( 63.00, 54.75) --
	( 66.24, 54.76) --
	( 69.48, 54.77) --
	( 72.73, 54.77) --
	( 75.97, 54.78) --
	( 79.22, 54.79) --
	( 82.46, 54.80) --
	( 85.70, 54.80) --
	( 88.95, 54.81) --
	( 92.19, 54.81) --
	( 95.43, 54.81) --
	( 98.68, 54.80) --
	(101.92, 54.79) --
	(105.17, 54.78) --
	(108.41, 54.77) --
	(111.65, 54.77) --
	(114.90, 54.76) --
	(118.14, 54.76) --
	(121.39, 54.76) --
	(124.63, 54.76) --
	(127.87, 54.77) --
	(131.12, 54.79) --
	(134.36, 54.80) --
	(137.61, 54.82) --
	(140.85, 54.83) --
	(144.09, 54.85) --
	(147.34, 54.87) --
	(150.58, 54.88) --
	(153.82, 54.89) --
	(157.07, 54.89) --
	(160.31, 54.89) --
	(163.56, 54.89) --
	(166.80, 54.89) --
	(170.04, 54.89) --
	(173.29, 54.89) --
	(176.53, 54.89) --
	(179.78, 54.88) --
	(183.02, 54.88) --
	(186.26, 54.88) --
	(189.51, 54.88) --
	(192.75, 54.87) --
	(196.00, 54.87) --
	(199.24, 54.87) --
	(202.48, 54.87) --
	(205.73, 54.86) --
	(208.97, 54.86) --
	(212.21, 54.86) --
	(215.46, 54.86) --
	(218.70, 54.86) --
	(221.95, 54.85) --
	(225.19, 54.85) --
	(228.43, 54.85) --
	(231.68, 54.85) --
	(234.92, 54.85) --
	(238.17, 54.84) --
	(241.41, 54.84) --
	(244.65, 54.84) --
	(247.90, 54.84) --
	(251.14, 54.83) --
	(254.39, 54.83) --
	(257.63, 54.83) --
	(260.87, 54.83) --
	(264.12, 54.83) --
	(267.36, 54.83) --
	(270.60, 54.83) --
	(273.85, 54.83) --
	(277.09, 54.82) --
	(280.34, 54.82) --
	(283.58, 54.82);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 27.31, 55.83) --
	( 30.56, 56.93) --
	( 33.80, 58.02) --
	( 37.04, 59.09) --
	( 40.29, 60.16) --
	( 43.53, 61.21) --
	( 46.78, 62.25) --
	( 50.02, 63.27) --
	( 53.26, 64.29) --
	( 56.51, 65.29) --
	( 59.75, 66.28) --
	( 63.00, 67.26) --
	( 66.24, 68.22) --
	( 69.48, 69.18) --
	( 72.73, 70.12) --
	( 75.97, 71.05) --
	( 79.22, 71.97) --
	( 82.46, 72.88) --
	( 85.70, 73.77) --
	( 88.95, 74.66) --
	( 92.19, 75.53) --
	( 95.43, 76.38) --
	( 98.68, 77.22) --
	(101.92, 78.05) --
	(105.17, 78.87) --
	(108.41, 79.68) --
	(111.65, 80.48) --
	(114.90, 81.26) --
	(118.14, 82.04) --
	(121.39, 82.81) --
	(124.63, 83.56) --
	(127.87, 84.30) --
	(131.12, 85.03) --
	(134.36, 85.74) --
	(137.61, 86.44) --
	(140.85, 87.13) --
	(144.09, 87.82) --
	(147.34, 88.50) --
	(150.58, 89.18) --
	(153.82, 89.85) --
	(157.07, 90.53) --
	(160.31, 91.19) --
	(163.56, 91.85) --
	(166.80, 92.50) --
	(170.04, 93.14) --
	(173.29, 93.78) --
	(176.53, 94.41) --
	(179.78, 95.03) --
	(183.02, 95.66) --
	(186.26, 96.27) --
	(189.51, 96.89) --
	(192.75, 97.50) --
	(196.00, 98.11) --
	(199.24, 98.72) --
	(202.48, 99.31) --
	(205.73, 99.91) --
	(208.97,100.50) --
	(212.21,101.08) --
	(215.46,101.66) --
	(218.70,102.23) --
	(221.95,102.80) --
	(225.19,103.37) --
	(228.43,103.93) --
	(231.68,104.49) --
	(234.92,105.04) --
	(238.17,105.59) --
	(241.41,106.13) --
	(244.65,106.66) --
	(247.90,107.20) --
	(251.14,107.72) --
	(254.39,108.25) --
	(257.63,108.77) --
	(260.87,109.28) --
	(264.12,109.79) --
	(267.36,110.29) --
	(270.60,110.79) --
	(273.85,111.29) --
	(277.09,111.78) --
	(280.34,112.27) --
	(283.58,112.75);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 27.31, 55.69) --
	( 30.56, 57.48) --
	( 33.80, 59.25) --
	( 37.04, 61.00) --
	( 40.29, 62.74) --
	( 43.53, 64.46) --
	( 46.78, 66.16) --
	( 50.02, 67.84) --
	( 53.26, 69.50) --
	( 56.51, 71.15) --
	( 59.75, 72.78) --
	( 63.00, 74.39) --
	( 66.24, 75.98) --
	( 69.48, 77.56) --
	( 72.73, 79.11) --
	( 75.97, 80.65) --
	( 79.22, 82.18) --
	( 82.46, 83.68) --
	( 85.70, 85.17) --
	( 88.95, 86.65) --
	( 92.19, 88.10) --
	( 95.43, 89.53) --
	( 98.68, 90.94) --
	(101.92, 92.34) --
	(105.17, 93.71) --
	(108.41, 95.07) --
	(111.65, 96.42) --
	(114.90, 97.74) --
	(118.14, 99.06) --
	(121.39,100.36) --
	(124.63,101.64) --
	(127.87,102.90) --
	(131.12,104.13) --
	(134.36,105.34) --
	(137.61,106.54) --
	(140.85,107.73) --
	(144.09,108.90) --
	(147.34,110.07) --
	(150.58,111.23) --
	(153.82,112.39) --
	(157.07,113.56) --
	(160.31,114.71) --
	(163.56,115.85) --
	(166.80,116.97) --
	(170.04,118.09) --
	(173.29,119.20) --
	(176.53,120.30) --
	(179.78,121.39) --
	(183.02,122.48) --
	(186.26,123.56) --
	(189.51,124.64) --
	(192.75,125.71) --
	(196.00,126.77) --
	(199.24,127.83) --
	(202.48,128.88) --
	(205.73,129.92) --
	(208.97,130.96) --
	(212.21,131.98) --
	(215.46,133.00) --
	(218.70,134.02) --
	(221.95,135.02) --
	(225.19,136.02) --
	(228.43,137.01) --
	(231.68,137.99) --
	(234.92,138.97) --
	(238.17,139.94) --
	(241.41,140.90) --
	(244.65,141.85) --
	(247.90,142.80) --
	(251.14,143.74) --
	(254.39,144.67) --
	(257.63,145.59) --
	(260.87,146.51) --
	(264.12,147.42) --
	(267.36,148.32) --
	(270.60,149.21) --
	(273.85,150.10) --
	(277.09,150.98) --
	(280.34,151.85) --
	(283.58,152.72);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 55.36) --
	( 30.56, 57.79) --
	( 33.80, 60.21) --
	( 37.04, 62.60) --
	( 40.29, 64.97) --
	( 43.53, 67.32) --
	( 46.78, 69.65) --
	( 50.02, 71.96) --
	( 53.26, 74.25) --
	( 56.51, 76.51) --
	( 59.75, 78.75) --
	( 63.00, 80.97) --
	( 66.24, 83.17) --
	( 69.48, 85.34) --
	( 72.73, 87.50) --
	( 75.97, 89.63) --
	( 79.22, 91.75) --
	( 82.46, 93.84) --
	( 85.70, 95.91) --
	( 88.95, 97.97) --
	( 92.19, 99.99) --
	( 95.43,102.00) --
	( 98.68,103.98) --
	(101.92,105.93) --
	(105.17,107.87) --
	(108.41,109.78) --
	(111.65,111.68) --
	(114.90,113.56) --
	(118.14,115.41) --
	(121.39,117.26) --
	(124.63,119.08) --
	(127.87,120.86) --
	(131.12,122.62) --
	(134.36,124.35) --
	(137.61,126.07) --
	(140.85,127.76) --
	(144.09,129.45) --
	(147.34,131.13) --
	(150.58,132.80) --
	(153.82,134.48) --
	(157.07,136.16) --
	(160.31,137.82) --
	(163.56,139.47) --
	(166.80,141.11) --
	(170.04,142.73) --
	(173.29,144.35) --
	(176.53,145.95) --
	(179.78,147.55) --
	(183.02,149.14) --
	(186.26,150.72) --
	(189.51,152.30) --
	(192.75,153.87) --
	(196.00,155.43) --
	(199.24,156.98) --
	(202.48,158.52) --
	(205.73,160.06) --
	(208.97,161.58) --
	(212.21,163.09) --
	(215.46,164.60) --
	(218.70,166.10) --
	(221.95,167.58) --
	(225.19,169.06) --
	(228.43,170.53) --
	(231.68,171.99) --
	(234.92,173.44) --
	(238.17,174.88) --
	(241.41,176.31) --
	(244.65,177.74) --
	(247.90,179.15) --
	(251.14,180.55) --
	(254.39,181.95) --
	(257.63,183.34) --
	(260.87,184.71) --
	(264.12,186.08) --
	(267.36,187.44) --
	(270.60,188.79) --
	(273.85,190.13) --
	(277.09,191.46) --
	(280.34,192.78) --
	(283.58,194.09);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 27.31, 59.72) --
	( 30.68, 53.90) --
	( 23.95, 53.90) --
	cycle;

\path[fill=fillColor] ( 35.86, 62.59) --
	( 39.22, 56.77) --
	( 32.49, 56.77) --
	cycle;

\path[fill=fillColor] ( 44.40, 65.16) --
	( 47.76, 59.33) --
	( 41.03, 59.33) --
	cycle;

\path[fill=fillColor] ( 52.94, 68.11) --
	( 56.30, 62.29) --
	( 49.58, 62.29) --
	cycle;

\path[fill=fillColor] ( 61.48, 70.80) --
	( 64.85, 64.98) --
	( 58.12, 64.98) --
	cycle;

\path[fill=fillColor] ( 70.02, 73.35) --
	( 73.39, 67.53) --
	( 66.66, 67.53) --
	cycle;

\path[fill=fillColor] ( 78.57, 75.71) --
	( 81.93, 69.89) --
	( 75.20, 69.89) --
	cycle;

\path[fill=fillColor] ( 87.11, 77.99) --
	( 90.47, 72.17) --
	( 83.75, 72.17) --
	cycle;

\path[fill=fillColor] ( 95.65, 80.38) --
	( 99.01, 74.56) --
	( 92.29, 74.56) --
	cycle;

\path[fill=fillColor] (104.19, 82.73) --
	(107.56, 76.90) --
	(100.83, 76.90) --
	cycle;

\path[fill=fillColor] (112.74, 84.79) --
	(116.10, 78.96) --
	(109.37, 78.96) --
	cycle;

\path[fill=fillColor] (121.28, 86.71) --
	(124.64, 80.88) --
	(117.91, 80.88) --
	cycle;

\path[fill=fillColor] (129.82, 88.52) --
	(133.18, 82.69) --
	(126.46, 82.69) --
	cycle;

\path[fill=fillColor] (138.36, 90.39) --
	(141.73, 84.57) --
	(135.00, 84.57) --
	cycle;

\path[fill=fillColor] (146.90, 92.35) --
	(150.27, 86.52) --
	(143.54, 86.52) --
	cycle;

\path[fill=fillColor] (155.45, 94.05) --
	(158.81, 88.22) --
	(152.08, 88.22) --
	cycle;

\path[fill=fillColor] (163.99, 95.73) --
	(167.35, 89.90) --
	(160.63, 89.90) --
	cycle;

\path[fill=fillColor] (172.53, 97.38) --
	(175.89, 91.55) --
	(169.17, 91.55) --
	cycle;

\path[fill=fillColor] (181.07, 99.22) --
	(184.44, 93.40) --
	(177.71, 93.40) --
	cycle;

\path[fill=fillColor] (189.62,100.84) --
	(192.98, 95.02) --
	(186.25, 95.02) --
	cycle;

\path[fill=fillColor] (198.16,102.49) --
	(201.52, 96.67) --
	(194.79, 96.67) --
	cycle;

\path[fill=fillColor] (206.70,103.90) --
	(210.06, 98.08) --
	(203.34, 98.08) --
	cycle;

\path[fill=fillColor] (215.24,105.43) --
	(218.61, 99.60) --
	(211.88, 99.60) --
	cycle;

\path[fill=fillColor] (223.78,107.01) --
	(227.15,101.18) --
	(220.42,101.18) --
	cycle;

\path[fill=fillColor] (232.33,108.48) --
	(235.69,102.65) --
	(228.96,102.65) --
	cycle;

\path[fill=fillColor] (240.87,109.92) --
	(244.23,104.10) --
	(237.51,104.10) --
	cycle;

\path[fill=fillColor] (249.41,111.32) --
	(252.77,105.49) --
	(246.05,105.49) --
	cycle;

\path[fill=fillColor] (257.95,112.65) --
	(261.32,106.82) --
	(254.59,106.82) --
	cycle;

\path[fill=fillColor] (266.50,114.09) --
	(269.86,108.26) --
	(263.13,108.26) --
	cycle;

\path[fill=fillColor] (275.04,115.30) --
	(278.40,109.48) --
	(271.67,109.48) --
	cycle;

\path[fill=fillColor] (283.58,116.69) --
	(286.94,110.86) --
	(280.22,110.86) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 24.82, 53.26) --
	( 29.81, 53.26) --
	( 29.81, 58.26) --
	( 24.82, 58.26) --
	cycle;

\path[fill=fillColor] ( 33.36, 57.87) --
	( 38.35, 57.87) --
	( 38.35, 62.86) --
	( 33.36, 62.86) --
	cycle;

\path[fill=fillColor] ( 41.90, 62.08) --
	( 46.90, 62.08) --
	( 46.90, 67.07) --
	( 41.90, 67.07) --
	cycle;

\path[fill=fillColor] ( 50.44, 66.89) --
	( 55.44, 66.89) --
	( 55.44, 71.88) --
	( 50.44, 71.88) --
	cycle;

\path[fill=fillColor] ( 58.98, 71.24) --
	( 63.98, 71.24) --
	( 63.98, 76.24) --
	( 58.98, 76.24) --
	cycle;

\path[fill=fillColor] ( 67.53, 75.45) --
	( 72.52, 75.45) --
	( 72.52, 80.44) --
	( 67.53, 80.44) --
	cycle;

\path[fill=fillColor] ( 76.07, 79.40) --
	( 81.06, 79.40) --
	( 81.06, 84.39) --
	( 76.07, 84.39) --
	cycle;

\path[fill=fillColor] ( 84.61, 83.31) --
	( 89.61, 83.31) --
	( 89.61, 88.30) --
	( 84.61, 88.30) --
	cycle;

\path[fill=fillColor] ( 93.15, 87.27) --
	( 98.15, 87.27) --
	( 98.15, 92.27) --
	( 93.15, 92.27) --
	cycle;

\path[fill=fillColor] (101.70, 91.22) --
	(106.69, 91.22) --
	(106.69, 96.22) --
	(101.70, 96.22) --
	cycle;

\path[fill=fillColor] (110.24, 94.52) --
	(115.23, 94.52) --
	(115.23, 99.52) --
	(110.24, 99.52) --
	cycle;

\path[fill=fillColor] (118.78, 97.88) --
	(123.78, 97.88) --
	(123.78,102.87) --
	(118.78,102.87) --
	cycle;

\path[fill=fillColor] (127.32,100.97) --
	(132.32,100.97) --
	(132.32,105.96) --
	(127.32,105.96) --
	cycle;

\path[fill=fillColor] (135.86,104.20) --
	(140.86,104.20) --
	(140.86,109.20) --
	(135.86,109.20) --
	cycle;

\path[fill=fillColor] (144.41,107.50) --
	(149.40,107.50) --
	(149.40,112.49) --
	(144.41,112.49) --
	cycle;

\path[fill=fillColor] (152.95,110.47) --
	(157.94,110.47) --
	(157.94,115.46) --
	(152.95,115.46) --
	cycle;

\path[fill=fillColor] (161.49,113.31) --
	(166.49,113.31) --
	(166.49,118.31) --
	(161.49,118.31) --
	cycle;

\path[fill=fillColor] (170.03,116.16) --
	(175.03,116.16) --
	(175.03,121.15) --
	(170.03,121.15) --
	cycle;

\path[fill=fillColor] (178.58,119.44) --
	(183.57,119.44) --
	(183.57,124.44) --
	(178.58,124.44) --
	cycle;

\path[fill=fillColor] (187.12,122.22) --
	(192.11,122.22) --
	(192.11,127.22) --
	(187.12,127.22) --
	cycle;

\path[fill=fillColor] (195.66,125.20) --
	(200.66,125.20) --
	(200.66,130.19) --
	(195.66,130.19) --
	cycle;

\path[fill=fillColor] (204.20,127.62) --
	(209.20,127.62) --
	(209.20,132.62) --
	(204.20,132.62) --
	cycle;

\path[fill=fillColor] (212.74,130.39) --
	(217.74,130.39) --
	(217.74,135.38) --
	(212.74,135.38) --
	cycle;

\path[fill=fillColor] (221.29,133.11) --
	(226.28,133.11) --
	(226.28,138.11) --
	(221.29,138.11) --
	cycle;

\path[fill=fillColor] (229.83,135.65) --
	(234.82,135.65) --
	(234.82,140.64) --
	(229.83,140.64) --
	cycle;

\path[fill=fillColor] (238.37,138.18) --
	(243.37,138.18) --
	(243.37,143.18) --
	(238.37,143.18) --
	cycle;

\path[fill=fillColor] (246.91,140.83) --
	(251.91,140.83) --
	(251.91,145.82) --
	(246.91,145.82) --
	cycle;

\path[fill=fillColor] (255.46,143.10) --
	(260.45,143.10) --
	(260.45,148.10) --
	(255.46,148.10) --
	cycle;

\path[fill=fillColor] (264.00,145.63) --
	(268.99,145.63) --
	(268.99,150.63) --
	(264.00,150.63) --
	cycle;

\path[fill=fillColor] (272.54,147.86) --
	(277.54,147.86) --
	(277.54,152.86) --
	(272.54,152.86) --
	cycle;

\path[fill=fillColor] (281.08,150.27) --
	(286.08,150.27) --
	(286.08,155.27) --
	(281.08,155.27) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 23.78, 55.53) -- ( 30.85, 55.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.31, 52.00) -- ( 27.31, 59.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.32, 61.72) -- ( 39.39, 61.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.86, 58.18) -- ( 35.86, 65.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87, 67.50) -- ( 47.93, 67.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.40, 63.97) -- ( 44.40, 71.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.41, 74.08) -- ( 56.47, 74.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.94, 70.55) -- ( 52.94, 77.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.95, 79.95) -- ( 65.01, 79.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.48, 76.42) -- ( 61.48, 83.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.49, 85.82) -- ( 73.56, 85.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.02, 82.28) -- ( 70.02, 89.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.03, 91.32) -- ( 82.10, 91.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.57, 87.79) -- ( 78.57, 94.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.58, 96.85) -- ( 90.64, 96.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.11, 93.32) -- ( 87.11,100.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.12,102.37) -- ( 99.18,102.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65, 98.84) -- ( 95.65,105.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.66,107.95) -- (107.73,107.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.19,104.42) -- (104.19,111.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.20,112.40) -- (116.27,112.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.74,108.86) -- (112.74,115.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.75,117.27) -- (124.81,117.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.28,113.74) -- (121.28,120.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.29,121.67) -- (133.35,121.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.82,118.13) -- (129.82,125.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.83,126.30) -- (141.89,126.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.36,122.76) -- (138.36,129.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.37,131.03) -- (150.44,131.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.90,127.50) -- (146.90,134.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.91,135.37) -- (158.98,135.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (155.45,131.84) -- (155.45,138.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.46,139.39) -- (167.52,139.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.99,135.86) -- (163.99,142.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.00,143.52) -- (176.06,143.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.53,139.99) -- (172.53,147.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.54,148.36) -- (184.61,148.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.07,144.83) -- (181.07,151.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.08,152.40) -- (193.15,152.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.62,148.86) -- (189.62,155.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.63,156.84) -- (201.69,156.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.16,153.31) -- (198.16,160.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.17,160.37) -- (210.23,160.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.70,156.84) -- (206.70,163.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.71,164.46) -- (218.77,164.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.24,160.93) -- (215.24,168.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.25,168.49) -- (227.32,168.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.78,164.96) -- (223.78,172.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.79,172.17) -- (235.86,172.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.33,168.64) -- (232.33,175.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.34,175.87) -- (244.40,175.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.87,172.34) -- (240.87,179.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.88,180.09) -- (252.94,180.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.41,176.56) -- (249.41,183.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.42,183.35) -- (261.49,183.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.95,179.81) -- (257.95,186.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.96,187.13) -- (270.03,187.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (266.50,183.60) -- (266.50,190.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.51,190.53) -- (278.57,190.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.04,187.00) -- (275.04,194.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.05,194.15) -- (287.11,194.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.58,190.62) -- (283.58,197.68);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 27.31, 53.86) circle (  2.50);

\path[fill=fillColor] ( 35.86, 54.98) circle (  2.50);

\path[fill=fillColor] ( 44.40, 55.01) circle (  2.50);

\path[fill=fillColor] ( 52.94, 54.11) circle (  2.50);

\path[fill=fillColor] ( 61.48, 55.47) circle (  2.50);

\path[fill=fillColor] ( 70.02, 54.78) circle (  2.50);

\path[fill=fillColor] ( 78.57, 54.81) circle (  2.50);

\path[fill=fillColor] ( 87.11, 55.37) circle (  2.50);

\path[fill=fillColor] ( 95.65, 54.67) circle (  2.50);

\path[fill=fillColor] (104.19, 55.03) circle (  2.50);

\path[fill=fillColor] (112.74, 54.66) circle (  2.50);

\path[fill=fillColor] (121.28, 54.19) circle (  2.50);

\path[fill=fillColor] (129.82, 54.27) circle (  2.50);

\path[fill=fillColor] (138.36, 55.20) circle (  2.50);

\path[fill=fillColor] (146.90, 54.61) circle (  2.50);

\path[fill=fillColor] (155.45, 55.25) circle (  2.50);

\path[fill=fillColor] (163.99, 55.30) circle (  2.50);

\path[fill=fillColor] (172.53, 55.03) circle (  2.50);

\path[fill=fillColor] (181.07, 54.88) circle (  2.50);

\path[fill=fillColor] (189.62, 54.77) circle (  2.50);

\path[fill=fillColor] (198.16, 54.88) circle (  2.50);

\path[fill=fillColor] (206.70, 54.44) circle (  2.50);

\path[fill=fillColor] (215.24, 54.97) circle (  2.50);

\path[fill=fillColor] (223.78, 54.48) circle (  2.50);

\path[fill=fillColor] (232.33, 54.54) circle (  2.50);

\path[fill=fillColor] (240.87, 55.09) circle (  2.50);

\path[fill=fillColor] (249.41, 55.20) circle (  2.50);

\path[fill=fillColor] (257.95, 54.91) circle (  2.50);

\path[fill=fillColor] (266.50, 55.15) circle (  2.50);

\path[fill=fillColor] (275.04, 54.72) circle (  2.50);

\path[fill=fillColor] (283.58, 54.55) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 30.69) rectangle (283.58,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 52.35) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,134.65) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 55.38) --
	( 27.31, 55.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,137.68) --
	( 27.31,137.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 27.94) --
	( 27.31, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.74, 27.94) --
	(112.74, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.16, 27.94) --
	(198.16, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.74, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.16, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (155.45,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,112.42) {Mean generation/serial interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,202.22) {C};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,216.81);
\end{scope}
\begin{scope}
\path[clip] (323.24, 30.69) rectangle (572.66,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (323.24, 30.69) rectangle (572.66,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (323.24, 68.09) --
	(326.39, 68.10) --
	(329.55, 68.10) --
	(332.71, 68.11) --
	(335.87, 68.12) --
	(339.02, 68.12) --
	(342.18, 68.13) --
	(345.34, 68.14) --
	(348.49, 68.14) --
	(351.65, 68.14) --
	(354.81, 68.15) --
	(357.97, 68.15) --
	(361.12, 68.15) --
	(364.28, 68.16) --
	(367.44, 68.16) --
	(370.60, 68.16) --
	(373.75, 68.16) --
	(376.91, 68.16) --
	(380.07, 68.16) --
	(383.22, 68.16) --
	(386.38, 68.16) --
	(389.54, 68.16) --
	(392.70, 68.16) --
	(395.85, 68.16) --
	(399.01, 68.16) --
	(402.17, 68.16) --
	(405.32, 68.15) --
	(408.48, 68.15) --
	(411.64, 68.15) --
	(414.80, 68.14) --
	(417.95, 68.13) --
	(421.11, 68.12) --
	(424.27, 68.11) --
	(427.43, 68.09) --
	(430.58, 68.08) --
	(433.74, 68.06) --
	(436.90, 68.04) --
	(440.05, 68.03) --
	(443.21, 68.02) --
	(446.37, 68.01) --
	(449.53, 68.01) --
	(452.68, 68.01) --
	(455.84, 68.01) --
	(459.00, 68.02) --
	(462.16, 68.02) --
	(465.31, 68.03) --
	(468.47, 68.04) --
	(471.63, 68.04) --
	(474.78, 68.05) --
	(477.94, 68.06) --
	(481.10, 68.06) --
	(484.26, 68.07) --
	(487.41, 68.08) --
	(490.57, 68.09) --
	(493.73, 68.10) --
	(496.89, 68.11) --
	(500.04, 68.12) --
	(503.20, 68.13) --
	(506.36, 68.13) --
	(509.51, 68.14) --
	(512.67, 68.14) --
	(515.83, 68.15) --
	(518.99, 68.15) --
	(522.14, 68.15) --
	(525.30, 68.15) --
	(528.46, 68.15) --
	(531.62, 68.14) --
	(534.77, 68.14) --
	(537.93, 68.14) --
	(541.09, 68.14) --
	(544.24, 68.13) --
	(547.40, 68.12) --
	(550.56, 68.12) --
	(553.72, 68.11) --
	(556.87, 68.10) --
	(560.03, 68.08) --
	(563.19, 68.07) --
	(566.35, 68.06) --
	(569.50, 68.04) --
	(572.66, 68.03);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (323.24,141.29) --
	(326.39,140.30) --
	(329.55,139.33) --
	(332.71,138.37) --
	(335.87,137.43) --
	(339.02,136.50) --
	(342.18,135.60) --
	(345.34,134.71) --
	(348.49,133.83) --
	(351.65,132.97) --
	(354.81,132.13) --
	(357.97,131.31) --
	(361.12,130.51) --
	(364.28,129.71) --
	(367.44,128.94) --
	(370.60,128.18) --
	(373.75,127.44) --
	(376.91,126.71) --
	(380.07,126.00) --
	(383.22,125.30) --
	(386.38,124.63) --
	(389.54,123.97) --
	(392.70,123.33) --
	(395.85,122.72) --
	(399.01,122.12) --
	(402.17,121.53) --
	(405.32,120.96) --
	(408.48,120.39) --
	(411.64,119.84) --
	(414.80,119.30) --
	(417.95,118.78) --
	(421.11,118.27) --
	(424.27,117.79) --
	(427.43,117.32) --
	(430.58,116.87) --
	(433.74,116.43) --
	(436.90,115.99) --
	(440.05,115.56) --
	(443.21,115.13) --
	(446.37,114.70) --
	(449.53,114.27) --
	(452.68,113.85) --
	(455.84,113.44) --
	(459.00,113.04) --
	(462.16,112.65) --
	(465.31,112.27) --
	(468.47,111.89) --
	(471.63,111.52) --
	(474.78,111.14) --
	(477.94,110.78) --
	(481.10,110.41) --
	(484.26,110.04) --
	(487.41,109.68) --
	(490.57,109.32) --
	(493.73,108.97) --
	(496.89,108.62) --
	(500.04,108.28) --
	(503.20,107.94) --
	(506.36,107.61) --
	(509.51,107.28) --
	(512.67,106.96) --
	(515.83,106.63) --
	(518.99,106.32) --
	(522.14,106.01) --
	(525.30,105.70) --
	(528.46,105.39) --
	(531.62,105.09) --
	(534.77,104.80) --
	(537.93,104.51) --
	(541.09,104.23) --
	(544.24,103.94) --
	(547.40,103.67) --
	(550.56,103.40) --
	(553.72,103.13) --
	(556.87,102.86) --
	(560.03,102.60) --
	(563.19,102.35) --
	(566.35,102.10) --
	(569.50,101.85) --
	(572.66,101.61);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (323.24,185.67) --
	(326.39,183.76) --
	(329.55,181.89) --
	(332.71,180.05) --
	(335.87,178.24) --
	(339.02,176.47) --
	(342.18,174.73) --
	(345.34,173.02) --
	(348.49,171.35) --
	(351.65,169.71) --
	(354.81,168.11) --
	(357.97,166.55) --
	(361.12,165.01) --
	(364.28,163.51) --
	(367.44,162.05) --
	(370.60,160.61) --
	(373.75,159.21) --
	(376.91,157.84) --
	(380.07,156.49) --
	(383.22,155.18) --
	(386.38,153.92) --
	(389.54,152.70) --
	(392.70,151.52) --
	(395.85,150.37) --
	(399.01,149.26) --
	(402.17,148.17) --
	(405.32,147.12) --
	(408.48,146.09) --
	(411.64,145.08) --
	(414.80,144.08) --
	(417.95,143.12) --
	(421.11,142.21) --
	(424.27,141.33) --
	(427.43,140.48) --
	(430.58,139.67) --
	(433.74,138.87) --
	(436.90,138.09) --
	(440.05,137.32) --
	(443.21,136.56) --
	(446.37,135.79) --
	(449.53,135.02) --
	(452.68,134.27) --
	(455.84,133.54) --
	(459.00,132.83) --
	(462.16,132.14) --
	(465.31,131.46) --
	(468.47,130.80) --
	(471.63,130.14) --
	(474.78,129.49) --
	(477.94,128.85) --
	(481.10,128.20) --
	(484.26,127.56) --
	(487.41,126.93) --
	(490.57,126.31) --
	(493.73,125.70) --
	(496.89,125.11) --
	(500.04,124.52) --
	(503.20,123.94) --
	(506.36,123.36) --
	(509.51,122.80) --
	(512.67,122.24) --
	(515.83,121.69) --
	(518.99,121.15) --
	(522.14,120.61) --
	(525.30,120.09) --
	(528.46,119.58) --
	(531.62,119.07) --
	(534.77,118.57) --
	(537.93,118.08) --
	(541.09,117.60) --
	(544.24,117.13) --
	(547.40,116.67) --
	(550.56,116.22) --
	(553.72,115.77) --
	(556.87,115.33) --
	(560.03,114.91) --
	(563.19,114.49) --
	(566.35,114.07) --
	(569.50,113.67) --
	(572.66,113.28);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (364.81,191.74) --
	(367.44,190.07) --
	(370.07,188.42) --
	(372.70,186.81) --
	(375.33,185.22) --
	(377.96,183.67) --
	(380.59,182.13) --
	(383.22,180.63) --
	(385.86,179.15) --
	(388.49,177.70) --
	(391.12,176.27) --
	(393.75,174.87) --
	(396.38,173.50) --
	(399.01,172.16) --
	(401.64,170.85) --
	(404.27,169.56) --
	(406.90,168.29) --
	(409.53,167.05) --
	(412.17,165.83) --
	(414.80,164.63) --
	(417.43,163.47) --
	(420.06,162.34) --
	(422.69,161.26) --
	(425.32,160.20) --
	(427.95,159.18) --
	(430.58,158.17) --
	(433.21,157.19) --
	(435.85,156.21) --
	(438.48,155.25) --
	(441.11,154.29) --
	(443.74,153.36) --
	(446.37,152.46) --
	(449.00,151.59) --
	(451.63,150.73) --
	(454.26,149.89) --
	(456.89,149.05) --
	(459.52,148.23) --
	(462.16,147.40) --
	(464.79,146.58) --
	(467.42,145.76) --
	(470.05,144.95) --
	(472.68,144.16) --
	(475.31,143.38) --
	(477.94,142.60) --
	(480.57,141.84) --
	(483.20,141.09) --
	(485.84,140.34) --
	(488.47,139.61) --
	(491.10,138.89) --
	(493.73,138.18) --
	(496.36,137.48) --
	(498.99,136.79) --
	(501.62,136.11) --
	(504.25,135.44) --
	(506.88,134.78) --
	(509.51,134.14) --
	(512.15,133.50) --
	(514.78,132.87) --
	(517.41,132.25) --
	(520.04,131.64) --
	(522.67,131.03) --
	(525.30,130.43) --
	(527.93,129.84) --
	(530.56,129.25) --
	(533.19,128.68) --
	(535.83,128.11) --
	(538.46,127.56) --
	(541.09,127.01) --
	(543.72,126.47) --
	(546.35,125.94) --
	(548.98,125.42) --
	(551.61,124.90) --
	(554.24,124.39) --
	(556.87,123.90) --
	(559.50,123.41) --
	(562.14,122.93) --
	(564.77,122.45) --
	(567.40,121.99) --
	(570.03,121.53) --
	(572.66,121.08);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] (323.24,145.42) --
	(326.60,139.59) --
	(319.87,139.59) --
	cycle;

\path[fill=fillColor] (331.55,142.62) --
	(334.91,136.79) --
	(328.19,136.79) --
	cycle;

\path[fill=fillColor] (339.86,140.20) --
	(343.23,134.37) --
	(336.50,134.37) --
	cycle;

\path[fill=fillColor] (348.18,137.66) --
	(351.54,131.83) --
	(344.81,131.83) --
	cycle;

\path[fill=fillColor] (356.49,135.29) --
	(359.86,129.47) --
	(353.13,129.47) --
	cycle;

\path[fill=fillColor] (364.81,133.27) --
	(368.17,127.44) --
	(361.44,127.44) --
	cycle;

\path[fill=fillColor] (373.12,131.42) --
	(376.48,125.59) --
	(369.76,125.59) --
	cycle;

\path[fill=fillColor] (381.44,129.62) --
	(384.80,123.79) --
	(378.07,123.79) --
	cycle;

\path[fill=fillColor] (389.75,127.76) --
	(393.11,121.93) --
	(386.39,121.93) --
	cycle;

\path[fill=fillColor] (398.06,125.95) --
	(401.43,120.13) --
	(394.70,120.13) --
	cycle;

\path[fill=fillColor] (406.38,124.42) --
	(409.74,118.60) --
	(403.01,118.60) --
	cycle;

\path[fill=fillColor] (414.69,123.14) --
	(418.06,117.32) --
	(411.33,117.32) --
	cycle;

\path[fill=fillColor] (423.01,121.89) --
	(426.37,116.06) --
	(419.64,116.06) --
	cycle;

\path[fill=fillColor] (431.32,120.75) --
	(434.68,114.92) --
	(427.96,114.92) --
	cycle;

\path[fill=fillColor] (439.63,119.48) --
	(443.00,113.66) --
	(436.27,113.66) --
	cycle;

\path[fill=fillColor] (447.95,118.47) --
	(451.31,112.64) --
	(444.58,112.64) --
	cycle;

\path[fill=fillColor] (456.26,117.35) --
	(459.63,111.52) --
	(452.90,111.52) --
	cycle;

\path[fill=fillColor] (464.58,116.47) --
	(467.94,110.64) --
	(461.21,110.64) --
	cycle;

\path[fill=fillColor] (472.89,115.21) --
	(476.25,109.39) --
	(469.53,109.39) --
	cycle;

\path[fill=fillColor] (481.20,114.19) --
	(484.57,108.37) --
	(477.84,108.37) --
	cycle;

\path[fill=fillColor] (489.52,113.29) --
	(492.88,107.46) --
	(486.15,107.46) --
	cycle;

\path[fill=fillColor] (497.83,112.38) --
	(501.20,106.55) --
	(494.47,106.55) --
	cycle;

\path[fill=fillColor] (506.15,111.57) --
	(509.51,105.74) --
	(502.78,105.74) --
	cycle;

\path[fill=fillColor] (514.46,110.65) --
	(517.82,104.82) --
	(511.10,104.82) --
	cycle;

\path[fill=fillColor] (522.78,109.83) --
	(526.14,104.01) --
	(519.41,104.01) --
	cycle;

\path[fill=fillColor] (531.09,109.08) --
	(534.45,103.25) --
	(527.73,103.25) --
	cycle;

\path[fill=fillColor] (539.40,108.38) --
	(542.77,102.55) --
	(536.04,102.55) --
	cycle;

\path[fill=fillColor] (547.72,107.52) --
	(551.08,101.70) --
	(544.35,101.70) --
	cycle;

\path[fill=fillColor] (556.03,106.78) --
	(559.40,100.96) --
	(552.67,100.96) --
	cycle;

\path[fill=fillColor] (564.35,106.17) --
	(567.71,100.34) --
	(560.98,100.34) --
	cycle;

\path[fill=fillColor] (572.66,105.44) --
	(576.02, 99.61) --
	(569.30, 99.61) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] (320.74,183.71) --
	(325.73,183.71) --
	(325.73,188.71) --
	(320.74,188.71) --
	cycle;

\path[fill=fillColor] (329.05,178.25) --
	(334.05,178.25) --
	(334.05,183.24) --
	(329.05,183.24) --
	cycle;

\path[fill=fillColor] (337.37,173.63) --
	(342.36,173.63) --
	(342.36,178.62) --
	(337.37,178.62) --
	cycle;

\path[fill=fillColor] (345.68,168.76) --
	(350.68,168.76) --
	(350.68,173.76) --
	(345.68,173.76) --
	cycle;

\path[fill=fillColor] (353.99,164.12) --
	(358.99,164.12) --
	(358.99,169.12) --
	(353.99,169.12) --
	cycle;

\path[fill=fillColor] (362.31,160.33) --
	(367.30,160.33) --
	(367.30,165.33) --
	(362.31,165.33) --
	cycle;

\path[fill=fillColor] (370.62,156.78) --
	(375.62,156.78) --
	(375.62,161.77) --
	(370.62,161.77) --
	cycle;

\path[fill=fillColor] (378.94,153.43) --
	(383.93,153.43) --
	(383.93,158.43) --
	(378.94,158.43) --
	cycle;

\path[fill=fillColor] (387.25,149.99) --
	(392.25,149.99) --
	(392.25,154.99) --
	(387.25,154.99) --
	cycle;

\path[fill=fillColor] (395.57,146.71) --
	(400.56,146.71) --
	(400.56,151.71) --
	(395.57,151.71) --
	cycle;

\path[fill=fillColor] (403.88,143.82) --
	(408.88,143.82) --
	(408.88,148.82) --
	(403.88,148.82) --
	cycle;

\path[fill=fillColor] (412.19,141.55) --
	(417.19,141.55) --
	(417.19,146.55) --
	(412.19,146.55) --
	cycle;

\path[fill=fillColor] (420.51,139.32) --
	(425.50,139.32) --
	(425.50,144.31) --
	(420.51,144.31) --
	cycle;

\path[fill=fillColor] (428.82,137.17) --
	(433.82,137.17) --
	(433.82,142.17) --
	(428.82,142.17) --
	cycle;

\path[fill=fillColor] (437.14,134.85) --
	(442.13,134.85) --
	(442.13,139.85) --
	(437.14,139.85) --
	cycle;

\path[fill=fillColor] (445.45,133.06) --
	(450.45,133.06) --
	(450.45,138.05) --
	(445.45,138.05) --
	cycle;

\path[fill=fillColor] (453.76,131.12) --
	(458.76,131.12) --
	(458.76,136.11) --
	(453.76,136.11) --
	cycle;

\path[fill=fillColor] (462.08,129.46) --
	(467.07,129.46) --
	(467.07,134.45) --
	(462.08,134.45) --
	cycle;

\path[fill=fillColor] (470.39,127.34) --
	(475.39,127.34) --
	(475.39,132.34) --
	(470.39,132.34) --
	cycle;

\path[fill=fillColor] (478.71,125.54) --
	(483.70,125.54) --
	(483.70,130.54) --
	(478.71,130.54) --
	cycle;

\path[fill=fillColor] (487.02,123.98) --
	(492.02,123.98) --
	(492.02,128.97) --
	(487.02,128.97) --
	cycle;

\path[fill=fillColor] (495.34,122.45) --
	(500.33,122.45) --
	(500.33,127.44) --
	(495.34,127.44) --
	cycle;

\path[fill=fillColor] (503.65,120.96) --
	(508.64,120.96) --
	(508.64,125.95) --
	(503.65,125.95) --
	cycle;

\path[fill=fillColor] (511.96,119.41) --
	(516.96,119.41) --
	(516.96,124.40) --
	(511.96,124.40) --
	cycle;

\path[fill=fillColor] (520.28,118.05) --
	(525.27,118.05) --
	(525.27,123.05) --
	(520.28,123.05) --
	cycle;

\path[fill=fillColor] (528.59,116.70) --
	(533.59,116.70) --
	(533.59,121.70) --
	(528.59,121.70) --
	cycle;

\path[fill=fillColor] (536.91,115.61) --
	(541.90,115.61) --
	(541.90,120.60) --
	(536.91,120.60) --
	cycle;

\path[fill=fillColor] (545.22,114.22) --
	(550.22,114.22) --
	(550.22,119.22) --
	(545.22,119.22) --
	cycle;

\path[fill=fillColor] (553.53,112.86) --
	(558.53,112.86) --
	(558.53,117.86) --
	(553.53,117.86) --
	cycle;

\path[fill=fillColor] (561.85,111.89) --
	(566.84,111.89) --
	(566.84,116.88) --
	(561.85,116.88) --
	cycle;

\path[fill=fillColor] (570.16,110.62) --
	(575.16,110.62) --
	(575.16,115.62) --
	(570.16,115.62) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.27,192.11) -- (368.34,192.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.81,188.58) -- (364.81,195.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.59,186.73) -- (376.65,186.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (373.12,183.20) -- (373.12,190.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.90,181.78) -- (384.97,181.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.44,178.25) -- (381.44,185.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.22,176.74) -- (393.28,176.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.75,173.21) -- (389.75,180.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (394.53,171.94) -- (401.60,171.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.06,168.41) -- (398.06,175.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (402.85,167.61) -- (409.91,167.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.38,164.08) -- (406.38,171.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (411.16,164.39) -- (418.22,164.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (414.69,160.86) -- (414.69,167.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.47,161.21) -- (426.54,161.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.01,157.67) -- (423.01,164.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.79,157.99) -- (434.85,157.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.32,154.46) -- (431.32,161.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.10,154.70) -- (443.17,154.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (439.63,151.16) -- (439.63,158.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.42,152.12) -- (451.48,152.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (447.95,148.59) -- (447.95,155.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.73,149.36) -- (459.79,149.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.26,145.83) -- (456.26,152.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.04,146.94) -- (468.11,146.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.58,143.40) -- (464.58,150.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.36,144.00) -- (476.42,144.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (472.89,140.47) -- (472.89,147.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.67,141.47) -- (484.74,141.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.20,137.94) -- (481.20,145.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.99,139.29) -- (493.05,139.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (489.52,135.76) -- (489.52,142.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.30,137.22) -- (501.36,137.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (497.83,133.69) -- (497.83,140.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.61,134.96) -- (509.68,134.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (506.15,131.43) -- (506.15,138.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.93,132.89) -- (517.99,132.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (514.46,129.36) -- (514.46,136.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.24,131.00) -- (526.31,131.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.78,127.46) -- (522.78,134.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (527.56,129.08) -- (534.62,129.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (531.09,125.55) -- (531.09,132.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.87,127.75) -- (542.94,127.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (539.40,124.22) -- (539.40,131.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.19,125.86) -- (551.25,125.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.72,122.33) -- (547.72,129.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.50,123.89) -- (559.56,123.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (556.03,120.36) -- (556.03,127.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.81,122.63) -- (567.88,122.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.35,119.09) -- (564.35,126.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13,120.90) -- (576.19,120.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66,117.37) -- (572.66,124.44);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (323.24, 68.06) circle (  2.50);

\path[fill=fillColor] (331.55, 68.24) circle (  2.50);

\path[fill=fillColor] (339.86, 68.21) circle (  2.50);

\path[fill=fillColor] (348.18, 67.81) circle (  2.50);

\path[fill=fillColor] (356.49, 68.07) circle (  2.50);

\path[fill=fillColor] (364.81, 68.35) circle (  2.50);

\path[fill=fillColor] (373.12, 68.10) circle (  2.50);

\path[fill=fillColor] (381.44, 68.30) circle (  2.50);

\path[fill=fillColor] (389.75, 68.27) circle (  2.50);

\path[fill=fillColor] (398.06, 68.02) circle (  2.50);

\path[fill=fillColor] (406.38, 68.54) circle (  2.50);

\path[fill=fillColor] (414.69, 67.86) circle (  2.50);

\path[fill=fillColor] (423.01, 67.64) circle (  2.50);

\path[fill=fillColor] (431.32, 68.15) circle (  2.50);

\path[fill=fillColor] (439.63, 68.39) circle (  2.50);

\path[fill=fillColor] (447.95, 68.19) circle (  2.50);

\path[fill=fillColor] (456.26, 68.00) circle (  2.50);

\path[fill=fillColor] (464.58, 68.03) circle (  2.50);

\path[fill=fillColor] (472.89, 67.80) circle (  2.50);

\path[fill=fillColor] (481.20, 67.94) circle (  2.50);

\path[fill=fillColor] (489.52, 67.80) circle (  2.50);

\path[fill=fillColor] (497.83, 68.37) circle (  2.50);

\path[fill=fillColor] (506.15, 68.09) circle (  2.50);

\path[fill=fillColor] (514.46, 68.40) circle (  2.50);

\path[fill=fillColor] (522.78, 68.75) circle (  2.50);

\path[fill=fillColor] (531.09, 67.59) circle (  2.50);

\path[fill=fillColor] (539.40, 68.00) circle (  2.50);

\path[fill=fillColor] (547.72, 68.44) circle (  2.50);

\path[fill=fillColor] (556.03, 68.43) circle (  2.50);

\path[fill=fillColor] (564.35, 67.96) circle (  2.50);

\path[fill=fillColor] (572.66, 67.78) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.24, 30.69) rectangle (572.66,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29, 27.66) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29, 78.74) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,129.82) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,180.91) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49, 30.69) --
	(323.24, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49, 81.77) --
	(323.24, 81.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,132.85) --
	(323.24,132.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,183.94) --
	(323.24,183.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.24, 27.94) --
	(323.24, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.38, 27.94) --
	(406.38, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (489.52, 27.94) --
	(489.52, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (323.24, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (406.38, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.52, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (447.95,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,112.42) {Squared coefficient of variation};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (323.24,202.22) {D};
\end{scope}
\end{tikzpicture}

\end{document}