### 2020-04-15 rR.tex
almost done with the results section

% Created by tikzDevice version 0.12.3 on 2020-04-14 17:43:43
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,433.62);
\begin{scope}
\path[clip] (  0.00,216.81) rectangle (289.08,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,216.81) rectangle (289.08,433.62);
\end{scope}
\begin{scope}
\path[clip] ( 27.31,247.50) rectangle (283.58,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31,247.50) rectangle (283.58,410.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31,247.65) --
	( 30.56,248.50) --
	( 33.80,249.37) --
	( 37.04,250.25) --
	( 40.29,251.16) --
	( 43.53,252.09) --
	( 46.78,253.03) --
	( 50.02,254.00) --
	( 53.26,254.98) --
	( 56.51,255.98) --
	( 59.75,257.00) --
	( 63.00,258.04) --
	( 66.24,259.10) --
	( 69.48,260.18) --
	( 72.73,261.28) --
	( 75.97,262.40) --
	( 79.22,263.54) --
	( 82.46,264.70) --
	( 85.70,265.88) --
	( 88.95,267.07) --
	( 92.19,268.29) --
	( 95.43,269.52) --
	( 98.68,270.77) --
	(101.92,272.04) --
	(105.17,273.33) --
	(108.41,274.63) --
	(111.65,275.96) --
	(114.90,277.31) --
	(118.14,278.69) --
	(121.39,280.08) --
	(124.63,281.50) --
	(127.87,282.94) --
	(131.12,284.39) --
	(134.36,285.86) --
	(137.61,287.35) --
	(140.85,288.87) --
	(144.09,290.41) --
	(147.34,291.97) --
	(150.58,293.56) --
	(153.82,295.18) --
	(157.07,296.83) --
	(160.31,298.49) --
	(163.56,300.18) --
	(166.80,301.88) --
	(170.04,303.60) --
	(173.29,305.35) --
	(176.53,307.13) --
	(179.78,308.93) --
	(183.02,310.77) --
	(186.26,312.64) --
	(189.51,314.55) --
	(192.75,316.48) --
	(196.00,318.44) --
	(199.24,320.43) --
	(202.48,322.45) --
	(205.73,324.49) --
	(208.97,326.57) --
	(212.21,328.67) --
	(215.46,330.81) --
	(218.70,332.98) --
	(221.95,335.19) --
	(225.19,337.43) --
	(228.43,339.70) --
	(231.68,342.00) --
	(234.92,344.33) --
	(238.17,346.69) --
	(241.41,349.09) --
	(244.65,351.51) --
	(247.90,353.96) --
	(251.14,356.45) --
	(254.39,358.97) --
	(257.63,361.51) --
	(260.87,364.10) --
	(264.12,366.71) --
	(267.36,369.35) --
	(270.60,372.03) --
	(273.85,374.73) --
	(277.09,377.47) --
	(280.34,380.24) --
	(283.58,383.04);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 27.31,247.62) --
	( 30.56,248.48) --
	( 33.80,249.35) --
	( 37.04,250.25) --
	( 40.29,251.16) --
	( 43.53,252.10) --
	( 46.78,253.05) --
	( 50.02,254.02) --
	( 53.26,255.01) --
	( 56.51,256.02) --
	( 59.75,257.05) --
	( 63.00,258.09) --
	( 66.24,259.16) --
	( 69.48,260.24) --
	( 72.73,261.35) --
	( 75.97,262.47) --
	( 79.22,263.61) --
	( 82.46,264.77) --
	( 85.70,265.96) --
	( 88.95,267.16) --
	( 92.19,268.38) --
	( 95.43,269.61) --
	( 98.68,270.86) --
	(101.92,272.13) --
	(105.17,273.42) --
	(108.41,274.73) --
	(111.65,276.06) --
	(114.90,277.41) --
	(118.14,278.78) --
	(121.39,280.18) --
	(124.63,281.60) --
	(127.87,283.02) --
	(131.12,284.47) --
	(134.36,285.93) --
	(137.61,287.42) --
	(140.85,288.92) --
	(144.09,290.45) --
	(147.34,292.01) --
	(150.58,293.59) --
	(153.82,295.21) --
	(157.07,296.85) --
	(160.31,298.51) --
	(163.56,300.20) --
	(166.80,301.90) --
	(170.04,303.62) --
	(173.29,305.37) --
	(176.53,307.14) --
	(179.78,308.95) --
	(183.02,310.78) --
	(186.26,312.65) --
	(189.51,314.55) --
	(192.75,316.49) --
	(196.00,318.45) --
	(199.24,320.44) --
	(202.48,322.45) --
	(205.73,324.49) --
	(208.97,326.57) --
	(212.21,328.67) --
	(215.46,330.81) --
	(218.70,332.98) --
	(221.95,335.19) --
	(225.19,337.43) --
	(228.43,339.70) --
	(231.68,342.01) --
	(234.92,344.34) --
	(238.17,346.71) --
	(241.41,349.11) --
	(244.65,351.53) --
	(247.90,353.99) --
	(251.14,356.48) --
	(254.39,359.01) --
	(257.63,361.56) --
	(260.87,364.15) --
	(264.12,366.77) --
	(267.36,369.42) --
	(270.60,372.10) --
	(273.85,374.82) --
	(277.09,377.56) --
	(280.34,380.34) --
	(283.58,383.15);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 27.31,247.61) --
	( 30.56,248.47) --
	( 33.80,249.35) --
	( 37.04,250.24) --
	( 40.29,251.16) --
	( 43.53,252.10) --
	( 46.78,253.05) --
	( 50.02,254.02) --
	( 53.26,255.02) --
	( 56.51,256.03) --
	( 59.75,257.06) --
	( 63.00,258.10) --
	( 66.24,259.17) --
	( 69.48,260.26) --
	( 72.73,261.36) --
	( 75.97,262.48) --
	( 79.22,263.63) --
	( 82.46,264.79) --
	( 85.70,265.97) --
	( 88.95,267.17) --
	( 92.19,268.39) --
	( 95.43,269.63) --
	( 98.68,270.88) --
	(101.92,272.15) --
	(105.17,273.44) --
	(108.41,274.75) --
	(111.65,276.08) --
	(114.90,277.43) --
	(118.14,278.80) --
	(121.39,280.20) --
	(124.63,281.61) --
	(127.87,283.04) --
	(131.12,284.48) --
	(134.36,285.94) --
	(137.61,287.42) --
	(140.85,288.92) --
	(144.09,290.45) --
	(147.34,292.00) --
	(150.58,293.58) --
	(153.82,295.19) --
	(157.07,296.84) --
	(160.31,298.49) --
	(163.56,300.17) --
	(166.80,301.87) --
	(170.04,303.59) --
	(173.29,305.33) --
	(176.53,307.10) --
	(179.78,308.90) --
	(183.02,310.73) --
	(186.26,312.60) --
	(189.51,314.50) --
	(192.75,316.44) --
	(196.00,318.40) --
	(199.24,320.39) --
	(202.48,322.40) --
	(205.73,324.45) --
	(208.97,326.53) --
	(212.21,328.63) --
	(215.46,330.78) --
	(218.70,332.95) --
	(221.95,335.17) --
	(225.19,337.42) --
	(228.43,339.70) --
	(231.68,342.01) --
	(234.92,344.36) --
	(238.17,346.74) --
	(241.41,349.14) --
	(244.65,351.58) --
	(247.90,354.06) --
	(251.14,356.56) --
	(254.39,359.10) --
	(257.63,361.67) --
	(260.87,364.27) --
	(264.12,366.91) --
	(267.36,369.58) --
	(270.60,372.28) --
	(273.85,375.01) --
	(277.09,377.78) --
	(280.34,380.58) --
	(283.58,383.41);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31,247.60) --
	( 30.56,248.46) --
	( 33.80,249.34) --
	( 37.04,250.24) --
	( 40.29,251.16) --
	( 43.53,252.09) --
	( 46.78,253.05) --
	( 50.02,254.02) --
	( 53.26,255.02) --
	( 56.51,256.03) --
	( 59.75,257.06) --
	( 63.00,258.11) --
	( 66.24,259.18) --
	( 69.48,260.26) --
	( 72.73,261.37) --
	( 75.97,262.49) --
	( 79.22,263.64) --
	( 82.46,264.80) --
	( 85.70,265.98) --
	( 88.95,267.19) --
	( 92.19,268.40) --
	( 95.43,269.64) --
	( 98.68,270.89) --
	(101.92,272.16) --
	(105.17,273.45) --
	(108.41,274.76) --
	(111.65,276.09) --
	(114.90,277.44) --
	(118.14,278.81) --
	(121.39,280.21) --
	(124.63,281.62) --
	(127.87,283.04) --
	(131.12,284.48) --
	(134.36,285.94) --
	(137.61,287.42) --
	(140.85,288.92) --
	(144.09,290.44) --
	(147.34,291.99) --
	(150.58,293.57) --
	(153.82,295.18) --
	(157.07,296.82) --
	(160.31,298.48) --
	(163.56,300.15) --
	(166.80,301.85) --
	(170.04,303.56) --
	(173.29,305.30) --
	(176.53,307.07) --
	(179.78,308.87) --
	(183.02,310.70) --
	(186.26,312.56) --
	(189.51,314.47) --
	(192.75,316.41) --
	(196.00,318.37) --
	(199.24,320.36) --
	(202.48,322.38) --
	(205.73,324.43) --
	(208.97,326.51) --
	(212.21,328.63) --
	(215.46,330.77) --
	(218.70,332.96) --
	(221.95,335.18) --
	(225.19,337.45) --
	(228.43,339.74) --
	(231.68,342.07) --
	(234.92,344.42) --
	(238.17,346.82) --
	(241.41,349.24) --
	(244.65,351.70) --
	(247.90,354.18) --
	(251.14,356.71) --
	(254.39,359.26) --
	(257.63,361.85) --
	(260.87,364.48) --
	(264.12,367.14) --
	(267.36,369.83) --
	(270.60,372.55) --
	(273.85,375.31) --
	(277.09,378.11) --
	(280.34,380.93) --
	(283.58,383.79);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 27.31,251.38) --
	( 30.68,245.55) --
	( 23.95,245.55) --
	cycle;

\path[fill=fillColor] ( 35.86,253.77) --
	( 39.22,247.94) --
	( 32.49,247.94) --
	cycle;

\path[fill=fillColor] ( 44.40,256.26) --
	( 47.76,250.43) --
	( 41.03,250.43) --
	cycle;

\path[fill=fillColor] ( 52.94,258.86) --
	( 56.30,253.03) --
	( 49.58,253.03) --
	cycle;

\path[fill=fillColor] ( 61.48,261.57) --
	( 64.85,255.75) --
	( 58.12,255.75) --
	cycle;

\path[fill=fillColor] ( 70.02,264.40) --
	( 73.39,258.57) --
	( 66.66,258.57) --
	cycle;

\path[fill=fillColor] ( 78.57,267.34) --
	( 81.93,261.52) --
	( 75.20,261.52) --
	cycle;

\path[fill=fillColor] ( 87.11,270.42) --
	( 90.47,264.59) --
	( 83.75,264.59) --
	cycle;

\path[fill=fillColor] ( 95.65,273.63) --
	( 99.01,267.80) --
	( 92.29,267.80) --
	cycle;

\path[fill=fillColor] (104.19,276.97) --
	(107.56,271.14) --
	(100.83,271.14) --
	cycle;

\path[fill=fillColor] (112.74,280.44) --
	(116.10,274.61) --
	(109.37,274.61) --
	cycle;

\path[fill=fillColor] (121.28,284.03) --
	(124.64,278.20) --
	(117.91,278.20) --
	cycle;

\path[fill=fillColor] (129.82,287.78) --
	(133.18,281.95) --
	(126.46,281.95) --
	cycle;

\path[fill=fillColor] (138.36,291.67) --
	(141.73,285.84) --
	(135.00,285.84) --
	cycle;

\path[fill=fillColor] (146.90,295.72) --
	(150.27,289.89) --
	(143.54,289.89) --
	cycle;

\path[fill=fillColor] (155.45,299.91) --
	(158.81,294.09) --
	(152.08,294.09) --
	cycle;

\path[fill=fillColor] (163.99,304.26) --
	(167.35,298.43) --
	(160.63,298.43) --
	cycle;

\path[fill=fillColor] (172.53,308.81) --
	(175.89,302.98) --
	(169.17,302.98) --
	cycle;

\path[fill=fillColor] (181.07,313.60) --
	(184.44,307.77) --
	(177.71,307.77) --
	cycle;

\path[fill=fillColor] (189.62,318.53) --
	(192.98,312.70) --
	(186.25,312.70) --
	cycle;

\path[fill=fillColor] (198.16,323.63) --
	(201.52,317.81) --
	(194.79,317.81) --
	cycle;

\path[fill=fillColor] (206.70,328.95) --
	(210.06,323.12) --
	(203.34,323.12) --
	cycle;

\path[fill=fillColor] (215.24,334.48) --
	(218.61,328.65) --
	(211.88,328.65) --
	cycle;

\path[fill=fillColor] (223.78,340.26) --
	(227.15,334.44) --
	(220.42,334.44) --
	cycle;

\path[fill=fillColor] (232.33,346.20) --
	(235.69,340.37) --
	(228.96,340.37) --
	cycle;

\path[fill=fillColor] (240.87,352.35) --
	(244.23,346.52) --
	(237.51,346.52) --
	cycle;

\path[fill=fillColor] (249.41,358.83) --
	(252.77,353.01) --
	(246.05,353.01) --
	cycle;

\path[fill=fillColor] (257.95,365.61) --
	(261.32,359.79) --
	(254.59,359.79) --
	cycle;

\path[fill=fillColor] (266.50,372.59) --
	(269.86,366.77) --
	(263.13,366.77) --
	cycle;

\path[fill=fillColor] (275.04,379.74) --
	(278.40,373.92) --
	(271.67,373.92) --
	cycle;

\path[fill=fillColor] (283.58,387.27) --
	(286.94,381.45) --
	(280.22,381.45) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 24.82,245.00) --
	( 29.81,245.00) --
	( 29.81,249.99) --
	( 24.82,249.99) --
	cycle;

\path[fill=fillColor] ( 33.36,247.39) --
	( 38.35,247.39) --
	( 38.35,252.38) --
	( 33.36,252.38) --
	cycle;

\path[fill=fillColor] ( 41.90,249.87) --
	( 46.90,249.87) --
	( 46.90,254.87) --
	( 41.90,254.87) --
	cycle;

\path[fill=fillColor] ( 50.44,252.48) --
	( 55.44,252.48) --
	( 55.44,257.47) --
	( 50.44,257.47) --
	cycle;

\path[fill=fillColor] ( 58.98,255.19) --
	( 63.98,255.19) --
	( 63.98,260.19) --
	( 58.98,260.19) --
	cycle;

\path[fill=fillColor] ( 67.53,258.02) --
	( 72.52,258.02) --
	( 72.52,263.01) --
	( 67.53,263.01) --
	cycle;

\path[fill=fillColor] ( 76.07,260.97) --
	( 81.06,260.97) --
	( 81.06,265.96) --
	( 76.07,265.96) --
	cycle;

\path[fill=fillColor] ( 84.61,264.05) --
	( 89.61,264.05) --
	( 89.61,269.04) --
	( 84.61,269.04) --
	cycle;

\path[fill=fillColor] ( 93.15,267.26) --
	( 98.15,267.26) --
	( 98.15,272.26) --
	( 93.15,272.26) --
	cycle;

\path[fill=fillColor] (101.70,270.61) --
	(106.69,270.61) --
	(106.69,275.61) --
	(101.70,275.61) --
	cycle;

\path[fill=fillColor] (110.24,274.08) --
	(115.23,274.08) --
	(115.23,279.07) --
	(110.24,279.07) --
	cycle;

\path[fill=fillColor] (118.78,277.67) --
	(123.78,277.67) --
	(123.78,282.67) --
	(118.78,282.67) --
	cycle;

\path[fill=fillColor] (127.32,281.41) --
	(132.32,281.41) --
	(132.32,286.41) --
	(127.32,286.41) --
	cycle;

\path[fill=fillColor] (135.86,285.30) --
	(140.86,285.30) --
	(140.86,290.30) --
	(135.86,290.30) --
	cycle;

\path[fill=fillColor] (144.41,289.34) --
	(149.40,289.34) --
	(149.40,294.33) --
	(144.41,294.33) --
	cycle;

\path[fill=fillColor] (152.95,293.52) --
	(157.94,293.52) --
	(157.94,298.51) --
	(152.95,298.51) --
	cycle;

\path[fill=fillColor] (161.49,297.83) --
	(166.49,297.83) --
	(166.49,302.82) --
	(161.49,302.82) --
	cycle;

\path[fill=fillColor] (170.03,302.36) --
	(175.03,302.36) --
	(175.03,307.36) --
	(170.03,307.36) --
	cycle;

\path[fill=fillColor] (178.58,307.19) --
	(183.57,307.19) --
	(183.57,312.18) --
	(178.58,312.18) --
	cycle;

\path[fill=fillColor] (187.12,312.12) --
	(192.11,312.12) --
	(192.11,317.12) --
	(187.12,317.12) --
	cycle;

\path[fill=fillColor] (195.66,317.22) --
	(200.66,317.22) --
	(200.66,322.21) --
	(195.66,322.21) --
	cycle;

\path[fill=fillColor] (204.20,322.52) --
	(209.20,322.52) --
	(209.20,327.51) --
	(204.20,327.51) --
	cycle;

\path[fill=fillColor] (212.74,328.07) --
	(217.74,328.07) --
	(217.74,333.06) --
	(212.74,333.06) --
	cycle;

\path[fill=fillColor] (221.29,333.88) --
	(226.28,333.88) --
	(226.28,338.88) --
	(221.29,338.88) --
	cycle;

\path[fill=fillColor] (229.83,339.76) --
	(234.82,339.76) --
	(234.82,344.75) --
	(229.83,344.75) --
	cycle;

\path[fill=fillColor] (238.37,345.90) --
	(243.37,345.90) --
	(243.37,350.89) --
	(238.37,350.89) --
	cycle;

\path[fill=fillColor] (246.91,352.51) --
	(251.91,352.51) --
	(251.91,357.51) --
	(246.91,357.51) --
	cycle;

\path[fill=fillColor] (255.46,359.41) --
	(260.45,359.41) --
	(260.45,364.40) --
	(255.46,364.40) --
	cycle;

\path[fill=fillColor] (264.00,366.42) --
	(268.99,366.42) --
	(268.99,371.41) --
	(264.00,371.41) --
	cycle;

\path[fill=fillColor] (272.54,373.53) --
	(277.54,373.53) --
	(277.54,378.53) --
	(272.54,378.53) --
	cycle;

\path[fill=fillColor] (281.08,381.15) --
	(286.08,381.15) --
	(286.08,386.15) --
	(281.08,386.15) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 23.78,247.50) -- ( 30.85,247.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.31,243.96) -- ( 27.31,251.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.32,249.88) -- ( 39.39,249.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.86,246.35) -- ( 35.86,253.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87,252.37) -- ( 47.93,252.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.40,248.83) -- ( 44.40,255.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.41,254.97) -- ( 56.47,254.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.94,251.44) -- ( 52.94,258.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.95,257.69) -- ( 65.01,257.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.48,254.15) -- ( 61.48,261.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.49,260.51) -- ( 73.56,260.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.02,256.98) -- ( 70.02,264.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.03,263.46) -- ( 82.10,263.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.57,259.93) -- ( 78.57,266.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.58,266.55) -- ( 90.64,266.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.11,263.02) -- ( 87.11,270.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.12,269.78) -- ( 99.18,269.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65,266.24) -- ( 95.65,273.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.66,273.13) -- (107.73,273.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.19,269.60) -- (104.19,276.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.20,276.59) -- (116.27,276.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.74,273.06) -- (112.74,280.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.75,280.19) -- (124.81,280.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.28,276.66) -- (121.28,283.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.29,283.92) -- (133.35,283.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.82,280.39) -- (129.82,287.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.83,287.81) -- (141.89,287.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.36,284.28) -- (138.36,291.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.37,291.84) -- (150.44,291.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.90,288.31) -- (146.90,295.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.91,296.01) -- (158.98,296.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (155.45,292.47) -- (155.45,299.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.46,300.29) -- (167.52,300.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.99,296.75) -- (163.99,303.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.00,304.81) -- (176.06,304.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.53,301.27) -- (172.53,308.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.54,309.66) -- (184.61,309.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.07,306.13) -- (181.07,313.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.08,314.61) -- (193.15,314.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.62,311.07) -- (189.62,318.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.63,319.70) -- (201.69,319.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.16,316.17) -- (198.16,323.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.17,324.99) -- (210.23,324.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.70,321.46) -- (206.70,328.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.71,330.57) -- (218.77,330.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.24,327.04) -- (215.24,334.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.25,336.41) -- (227.32,336.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.78,332.88) -- (223.78,339.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.79,342.22) -- (235.86,342.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.33,338.69) -- (232.33,345.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.34,348.37) -- (244.40,348.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.87,344.84) -- (240.87,351.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.88,355.17) -- (252.94,355.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.41,351.64) -- (249.41,358.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.42,362.17) -- (261.49,362.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.95,358.64) -- (257.95,365.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.96,369.21) -- (270.03,369.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (266.50,365.68) -- (266.50,372.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.51,376.31) -- (278.57,376.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.04,372.78) -- (275.04,379.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.05,384.02) -- (287.11,384.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.58,380.49) -- (283.58,387.55);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 27.31,247.50) circle (  2.50);

\path[fill=fillColor] ( 35.86,249.88) circle (  2.50);

\path[fill=fillColor] ( 44.40,252.37) circle (  2.50);

\path[fill=fillColor] ( 52.94,254.94) circle (  2.50);

\path[fill=fillColor] ( 61.48,257.68) circle (  2.50);

\path[fill=fillColor] ( 70.02,260.47) circle (  2.50);

\path[fill=fillColor] ( 78.57,263.41) circle (  2.50);

\path[fill=fillColor] ( 87.11,266.50) circle (  2.50);

\path[fill=fillColor] ( 95.65,269.65) circle (  2.50);

\path[fill=fillColor] (104.19,273.00) circle (  2.50);

\path[fill=fillColor] (112.74,276.42) circle (  2.50);

\path[fill=fillColor] (121.28,279.99) circle (  2.50);

\path[fill=fillColor] (129.82,283.76) circle (  2.50);

\path[fill=fillColor] (138.36,287.74) circle (  2.50);

\path[fill=fillColor] (146.90,291.70) circle (  2.50);

\path[fill=fillColor] (155.45,296.03) circle (  2.50);

\path[fill=fillColor] (163.99,300.46) circle (  2.50);

\path[fill=fillColor] (172.53,304.98) circle (  2.50);

\path[fill=fillColor] (181.07,309.73) circle (  2.50);

\path[fill=fillColor] (189.62,314.62) circle (  2.50);

\path[fill=fillColor] (198.16,319.79) circle (  2.50);

\path[fill=fillColor] (206.70,324.93) circle (  2.50);

\path[fill=fillColor] (215.24,330.66) circle (  2.50);

\path[fill=fillColor] (223.78,336.18) circle (  2.50);

\path[fill=fillColor] (232.33,342.09) circle (  2.50);

\path[fill=fillColor] (240.87,348.75) circle (  2.50);

\path[fill=fillColor] (249.41,355.20) circle (  2.50);

\path[fill=fillColor] (257.95,361.58) circle (  2.50);

\path[fill=fillColor] (266.50,368.62) circle (  2.50);

\path[fill=fillColor] (275.04,375.76) circle (  2.50);

\path[fill=fillColor] (283.58,383.22) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31,247.50) rectangle (283.58,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,247.50) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,294.20) --
	( 27.31,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,340.91) --
	( 27.31,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,387.61) --
	( 27.31,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,244.75) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.74,244.75) --
	(112.74,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.16,244.75) --
	(198.16,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.74,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.16,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (155.45,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,329.23) {Reproduction number $\mathcal R$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 39.98,332.94) rectangle (168.41,407.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,381.80) rectangle ( 59.93,396.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 46.93,389.03) -- ( 58.49,389.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 52.71,389.03) circle (  2.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,367.35) rectangle ( 59.93,381.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 46.93,374.57) -- ( 58.49,374.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 52.71,378.46) --
	( 56.07,372.63) --
	( 49.34,372.63) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,352.89) rectangle ( 59.93,367.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 46.93,360.12) -- ( 58.49,360.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 50.21,357.62) --
	( 55.20,357.62) --
	( 55.20,362.62) --
	( 50.21,362.62) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 45.48,338.44) rectangle ( 59.93,352.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 46.93,345.67) -- ( 58.49,345.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.17,345.67) -- ( 56.24,345.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.71,342.13) -- ( 52.71,349.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,386.00) {Generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,371.54) {Serial interval ($\rho=0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,357.09) {Serial interval ($\rho=0$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 65.43,342.63) {Serial interval ($\rho=-0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,419.03) {A};
\end{scope}
\begin{scope}
\path[clip] (289.08,216.81) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,216.81) rectangle (578.16,433.62);
\end{scope}
\begin{scope}
\path[clip] (316.39,247.50) rectangle (572.66,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (316.39,247.50) rectangle (572.66,410.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (316.39,247.65) --
	(319.64,248.49) --
	(322.88,249.36) --
	(326.12,250.25) --
	(329.37,251.16) --
	(332.61,252.08) --
	(335.86,253.03) --
	(339.10,254.00) --
	(342.34,254.98) --
	(345.59,255.98) --
	(348.83,257.00) --
	(352.08,258.04) --
	(355.32,259.10) --
	(358.56,260.18) --
	(361.81,261.28) --
	(365.05,262.40) --
	(368.30,263.54) --
	(371.54,264.70) --
	(374.78,265.88) --
	(378.03,267.07) --
	(381.27,268.29) --
	(384.51,269.52) --
	(387.76,270.77) --
	(391.00,272.04) --
	(394.25,273.32) --
	(397.49,274.63) --
	(400.73,275.96) --
	(403.98,277.31) --
	(407.22,278.68) --
	(410.47,280.08) --
	(413.71,281.49) --
	(416.95,282.93) --
	(420.20,284.38) --
	(423.44,285.85) --
	(426.69,287.34) --
	(429.93,288.85) --
	(433.17,290.39) --
	(436.42,291.96) --
	(439.66,293.55) --
	(442.90,295.17) --
	(446.15,296.82) --
	(449.39,298.48) --
	(452.64,300.17) --
	(455.88,301.88) --
	(459.12,303.61) --
	(462.37,305.37) --
	(465.61,307.15) --
	(468.86,308.96) --
	(472.10,310.80) --
	(475.34,312.67) --
	(478.59,314.58) --
	(481.83,316.51) --
	(485.08,318.47) --
	(488.32,320.46) --
	(491.56,322.47) --
	(494.81,324.51) --
	(498.05,326.58) --
	(501.29,328.68) --
	(504.54,330.81) --
	(507.78,332.98) --
	(511.03,335.19) --
	(514.27,337.43) --
	(517.51,339.70) --
	(520.76,342.00) --
	(524.00,344.33) --
	(527.25,346.69) --
	(530.49,349.09) --
	(533.73,351.51) --
	(536.98,353.97) --
	(540.22,356.45) --
	(543.47,358.97) --
	(546.71,361.52) --
	(549.95,364.10) --
	(553.20,366.72) --
	(556.44,369.36) --
	(559.68,372.04) --
	(562.93,374.75) --
	(566.17,377.49) --
	(569.42,380.26) --
	(572.66,383.06);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (319.64,248.30) --
	(322.88,249.24) --
	(326.12,250.18) --
	(329.37,251.12) --
	(332.61,252.08) --
	(335.86,253.04) --
	(339.10,254.00) --
	(342.34,254.98) --
	(345.59,255.96) --
	(348.83,256.94) --
	(352.08,257.93) --
	(355.32,258.93) --
	(358.56,259.94) --
	(361.81,260.95) --
	(365.05,261.97) --
	(368.30,262.99) --
	(371.54,264.02) --
	(374.78,265.05) --
	(378.03,266.09) --
	(381.27,267.14) --
	(384.51,268.20) --
	(387.76,269.26) --
	(391.00,270.34) --
	(394.25,271.42) --
	(397.49,272.50) --
	(400.73,273.59) --
	(403.98,274.68) --
	(407.22,275.78) --
	(410.47,276.87) --
	(413.71,277.98) --
	(416.95,279.10) --
	(420.20,280.22) --
	(423.44,281.36) --
	(426.69,282.50) --
	(429.93,283.64) --
	(433.17,284.78) --
	(436.42,285.92) --
	(439.66,287.06) --
	(442.90,288.19) --
	(446.15,289.31) --
	(449.39,290.45) --
	(452.64,291.59) --
	(455.88,292.74) --
	(459.12,293.89) --
	(462.37,295.05) --
	(465.61,296.19) --
	(468.86,297.34) --
	(472.10,298.47) --
	(475.34,299.59) --
	(478.59,300.69) --
	(481.83,301.79) --
	(485.08,302.89) --
	(488.32,303.98) --
	(491.56,305.07) --
	(494.81,306.15) --
	(498.05,307.23) --
	(501.29,308.29) --
	(504.54,309.34) --
	(507.78,310.37) --
	(511.03,311.39) --
	(514.27,312.39) --
	(517.51,313.37) --
	(520.76,314.35) --
	(524.00,315.32) --
	(527.25,316.28) --
	(530.49,317.22) --
	(533.73,318.16) --
	(536.98,319.09) --
	(540.22,320.00) --
	(543.47,320.91) --
	(546.71,321.80) --
	(549.95,322.67) --
	(553.20,323.54) --
	(556.44,324.40) --
	(559.68,325.24) --
	(562.93,326.07) --
	(566.17,326.89) --
	(569.42,327.70) --
	(572.66,328.50);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (319.64,248.18) --
	(322.88,249.15) --
	(326.12,250.12) --
	(329.37,251.09) --
	(332.61,252.06) --
	(335.86,253.02) --
	(339.10,253.99) --
	(342.34,254.95) --
	(345.59,255.91) --
	(348.83,256.87) --
	(352.08,257.82) --
	(355.32,258.78) --
	(358.56,259.73) --
	(361.81,260.68) --
	(365.05,261.63) --
	(368.30,262.58) --
	(371.54,263.52) --
	(374.78,264.45) --
	(378.03,265.39) --
	(381.27,266.33) --
	(384.51,267.27) --
	(387.76,268.21) --
	(391.00,269.15) --
	(394.25,270.09) --
	(397.49,271.03) --
	(400.73,271.96) --
	(403.98,272.88) --
	(407.22,273.79) --
	(410.47,274.70) --
	(413.71,275.60) --
	(416.95,276.51) --
	(420.20,277.43) --
	(423.44,278.34) --
	(426.69,279.25) --
	(429.93,280.15) --
	(433.17,281.03) --
	(436.42,281.90) --
	(439.66,282.74) --
	(442.90,283.56) --
	(446.15,284.35) --
	(449.39,285.14) --
	(452.64,285.93) --
	(455.88,286.72) --
	(459.12,287.50) --
	(462.37,288.26) --
	(465.61,289.01) --
	(468.86,289.72) --
	(472.10,290.40) --
	(475.34,291.04) --
	(478.59,291.64) --
	(481.83,292.21) --
	(485.08,292.76) --
	(488.32,293.29) --
	(491.56,293.79) --
	(494.81,294.27) --
	(498.05,294.72) --
	(501.29,295.13) --
	(504.54,295.51) --
	(507.78,295.84) --
	(511.03,296.12) --
	(514.27,296.36) --
	(517.51,296.57) --
	(520.76,296.74) --
	(524.00,296.88) --
	(527.25,296.99) --
	(530.49,297.06) --
	(533.73,297.10) --
	(536.98,297.11) --
	(540.22,297.08) --
	(543.47,297.02) --
	(546.71,296.91) --
	(549.95,296.77) --
	(553.20,296.60) --
	(556.44,296.39) --
	(559.68,296.14) --
	(562.93,295.86) --
	(566.17,295.54) --
	(569.42,295.19) --
	(572.66,294.80);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (319.64,248.10) --
	(322.88,249.10) --
	(326.12,250.09) --
	(329.37,251.07) --
	(332.61,252.04) --
	(335.86,253.00) --
	(339.10,253.95) --
	(342.34,254.90) --
	(345.59,255.84) --
	(348.83,256.77) --
	(352.08,257.69) --
	(355.32,258.60) --
	(358.56,259.51) --
	(361.81,260.40) --
	(365.05,261.29) --
	(368.30,262.16) --
	(371.54,263.02) --
	(374.78,263.87) --
	(378.03,264.71) --
	(381.27,265.55) --
	(384.51,266.39) --
	(387.76,267.23) --
	(391.00,268.05) --
	(394.25,268.87) --
	(397.49,269.68) --
	(400.73,270.47) --
	(403.98,271.25) --
	(407.22,272.00) --
	(410.47,272.74) --
	(413.71,273.47) --
	(416.95,274.20) --
	(420.20,274.93) --
	(423.44,275.65) --
	(426.69,276.36) --
	(429.93,277.05) --
	(433.17,277.72) --
	(436.42,278.35) --
	(439.66,278.95) --
	(442.90,279.51) --
	(446.15,280.03) --
	(449.39,280.54) --
	(452.64,281.05) --
	(455.88,281.54) --
	(459.12,282.01) --
	(462.37,282.46) --
	(465.61,282.87) --
	(468.86,283.25) --
	(472.10,283.58) --
	(475.34,283.86) --
	(478.59,284.08) --
	(481.83,284.25) --
	(485.08,284.40) --
	(488.32,284.51) --
	(491.56,284.59) --
	(494.81,284.63) --
	(498.05,284.63) --
	(501.29,284.58) --
	(504.54,284.48) --
	(507.78,284.34) --
	(511.03,284.13) --
	(514.27,283.88) --
	(517.51,283.57) --
	(520.76,283.23) --
	(524.00,282.84) --
	(527.25,282.40) --
	(530.49,281.93) --
	(533.73,281.41) --
	(536.98,280.84) --
	(540.22,280.23) --
	(543.47,279.57) --
	(546.71,278.87) --
	(549.95,278.12) --
	(553.20,277.32) --
	(556.44,276.47) --
	(559.68,275.58) --
	(562.93,274.65) --
	(566.17,273.66) --
	(569.42,272.64) --
	(572.66,271.57);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] (316.39,251.38) --
	(319.76,245.55) --
	(313.03,245.55) --
	cycle;

\path[fill=fillColor] (324.94,253.76) --
	(328.30,247.93) --
	(321.57,247.93) --
	cycle;

\path[fill=fillColor] (333.48,256.20) --
	(336.84,250.37) --
	(330.11,250.37) --
	cycle;

\path[fill=fillColor] (342.02,258.70) --
	(345.38,252.88) --
	(338.66,252.88) --
	cycle;

\path[fill=fillColor] (350.56,261.27) --
	(353.93,255.44) --
	(347.20,255.44) --
	cycle;

\path[fill=fillColor] (359.10,263.89) --
	(362.47,258.07) --
	(355.74,258.07) --
	cycle;

\path[fill=fillColor] (367.65,266.57) --
	(371.01,260.75) --
	(364.28,260.75) --
	cycle;

\path[fill=fillColor] (376.19,269.31) --
	(379.55,263.48) --
	(372.83,263.48) --
	cycle;

\path[fill=fillColor] (384.73,272.09) --
	(388.09,266.27) --
	(381.37,266.27) --
	cycle;

\path[fill=fillColor] (393.27,274.93) --
	(396.64,269.10) --
	(389.91,269.10) --
	cycle;

\path[fill=fillColor] (401.82,277.80) --
	(405.18,271.97) --
	(398.45,271.97) --
	cycle;

\path[fill=fillColor] (410.36,280.71) --
	(413.72,274.89) --
	(406.99,274.89) --
	cycle;

\path[fill=fillColor] (418.90,283.66) --
	(422.26,277.83) --
	(415.54,277.83) --
	cycle;

\path[fill=fillColor] (427.44,286.63) --
	(430.81,280.80) --
	(424.08,280.80) --
	cycle;

\path[fill=fillColor] (435.98,289.62) --
	(439.35,283.79) --
	(432.62,283.79) --
	cycle;

\path[fill=fillColor] (444.53,292.62) --
	(447.89,286.79) --
	(441.16,286.79) --
	cycle;

\path[fill=fillColor] (453.07,295.63) --
	(456.43,289.80) --
	(449.71,289.80) --
	cycle;

\path[fill=fillColor] (461.61,298.63) --
	(464.97,292.81) --
	(458.25,292.81) --
	cycle;

\path[fill=fillColor] (470.15,301.63) --
	(473.52,295.80) --
	(466.79,295.80) --
	cycle;

\path[fill=fillColor] (478.70,304.60) --
	(482.06,298.77) --
	(475.33,298.77) --
	cycle;

\path[fill=fillColor] (487.24,307.54) --
	(490.60,301.71) --
	(483.87,301.71) --
	cycle;

\path[fill=fillColor] (495.78,310.43) --
	(499.14,304.60) --
	(492.42,304.60) --
	cycle;

\path[fill=fillColor] (504.32,313.26) --
	(507.69,307.44) --
	(500.96,307.44) --
	cycle;

\path[fill=fillColor] (512.86,316.03) --
	(516.23,310.20) --
	(509.50,310.20) --
	cycle;

\path[fill=fillColor] (521.41,318.70) --
	(524.77,312.87) --
	(518.04,312.87) --
	cycle;

\path[fill=fillColor] (529.95,321.27) --
	(533.31,315.44) --
	(526.59,315.44) --
	cycle;

\path[fill=fillColor] (538.49,323.72) --
	(541.85,317.89) --
	(535.13,317.89) --
	cycle;

\path[fill=fillColor] (547.03,326.03) --
	(550.40,320.20) --
	(543.67,320.20) --
	cycle;

\path[fill=fillColor] (555.58,328.18) --
	(558.94,322.35) --
	(552.21,322.35) --
	cycle;

\path[fill=fillColor] (564.12,330.14) --
	(567.48,324.32) --
	(560.75,324.32) --
	cycle;

\path[fill=fillColor] (572.66,331.91) --
	(576.02,326.08) --
	(569.30,326.08) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] (313.90,245.00) --
	(318.89,245.00) --
	(318.89,249.99) --
	(313.90,249.99) --
	cycle;

\path[fill=fillColor] (322.44,247.36) --
	(327.43,247.36) --
	(327.43,252.36) --
	(322.44,252.36) --
	cycle;

\path[fill=fillColor] (330.98,249.77) --
	(335.98,249.77) --
	(335.98,254.76) --
	(330.98,254.76) --
	cycle;

\path[fill=fillColor] (339.52,252.21) --
	(344.52,252.21) --
	(344.52,257.20) --
	(339.52,257.20) --
	cycle;

\path[fill=fillColor] (348.06,254.68) --
	(353.06,254.68) --
	(353.06,259.67) --
	(348.06,259.67) --
	cycle;

\path[fill=fillColor] (356.61,257.17) --
	(361.60,257.17) --
	(361.60,262.16) --
	(356.61,262.16) --
	cycle;

\path[fill=fillColor] (365.15,259.67) --
	(370.14,259.67) --
	(370.14,264.67) --
	(365.15,264.67) --
	cycle;

\path[fill=fillColor] (373.69,262.19) --
	(378.69,262.19) --
	(378.69,267.19) --
	(373.69,267.19) --
	cycle;

\path[fill=fillColor] (382.23,264.71) --
	(387.23,264.71) --
	(387.23,269.70) --
	(382.23,269.70) --
	cycle;

\path[fill=fillColor] (390.78,267.21) --
	(395.77,267.21) --
	(395.77,272.21) --
	(390.78,272.21) --
	cycle;

\path[fill=fillColor] (399.32,269.70) --
	(404.31,269.70) --
	(404.31,274.69) --
	(399.32,274.69) --
	cycle;

\path[fill=fillColor] (407.86,272.15) --
	(412.86,272.15) --
	(412.86,277.15) --
	(407.86,277.15) --
	cycle;

\path[fill=fillColor] (416.40,274.57) --
	(421.40,274.57) --
	(421.40,279.56) --
	(416.40,279.56) --
	cycle;

\path[fill=fillColor] (424.94,276.93) --
	(429.94,276.93) --
	(429.94,281.93) --
	(424.94,281.93) --
	cycle;

\path[fill=fillColor] (433.49,279.22) --
	(438.48,279.22) --
	(438.48,284.22) --
	(433.49,284.22) --
	cycle;

\path[fill=fillColor] (442.03,281.44) --
	(447.02,281.44) --
	(447.02,286.43) --
	(442.03,286.43) --
	cycle;

\path[fill=fillColor] (450.57,283.55) --
	(455.57,283.55) --
	(455.57,288.55) --
	(450.57,288.55) --
	cycle;

\path[fill=fillColor] (459.11,285.55) --
	(464.11,285.55) --
	(464.11,290.55) --
	(459.11,290.55) --
	cycle;

\path[fill=fillColor] (467.66,287.42) --
	(472.65,287.42) --
	(472.65,292.42) --
	(467.66,292.42) --
	cycle;

\path[fill=fillColor] (476.20,289.14) --
	(481.19,289.14) --
	(481.19,294.14) --
	(476.20,294.14) --
	cycle;

\path[fill=fillColor] (484.74,290.69) --
	(489.74,290.69) --
	(489.74,295.69) --
	(484.74,295.69) --
	cycle;

\path[fill=fillColor] (493.28,292.05) --
	(498.28,292.05) --
	(498.28,297.04) --
	(493.28,297.04) --
	cycle;

\path[fill=fillColor] (501.82,293.19) --
	(506.82,293.19) --
	(506.82,298.19) --
	(501.82,298.19) --
	cycle;

\path[fill=fillColor] (510.37,294.09) --
	(515.36,294.09) --
	(515.36,299.09) --
	(510.37,299.09) --
	cycle;

\path[fill=fillColor] (518.91,294.73) --
	(523.90,294.73) --
	(523.90,299.73) --
	(518.91,299.73) --
	cycle;

\path[fill=fillColor] (527.45,295.08) --
	(532.45,295.08) --
	(532.45,300.07) --
	(527.45,300.07) --
	cycle;

\path[fill=fillColor] (535.99,295.11) --
	(540.99,295.11) --
	(540.99,300.11) --
	(535.99,300.11) --
	cycle;

\path[fill=fillColor] (544.54,294.80) --
	(549.53,294.80) --
	(549.53,299.79) --
	(544.54,299.79) --
	cycle;

\path[fill=fillColor] (553.08,294.12) --
	(558.07,294.12) --
	(558.07,299.12) --
	(553.08,299.12) --
	cycle;

\path[fill=fillColor] (561.62,293.05) --
	(566.62,293.05) --
	(566.62,298.05) --
	(561.62,298.05) --
	cycle;

\path[fill=fillColor] (570.16,291.57) --
	(575.16,291.57) --
	(575.16,296.57) --
	(570.16,296.57) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.86,247.50) -- (319.93,247.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.39,243.96) -- (316.39,251.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.40,249.85) -- (328.47,249.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (324.94,246.32) -- (324.94,253.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.95,252.22) -- (337.01,252.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.48,248.69) -- (333.48,255.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.49,254.60) -- (345.55,254.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.02,251.07) -- (342.02,258.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (347.03,256.98) -- (354.09,256.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.56,253.45) -- (350.56,260.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (355.57,259.35) -- (362.64,259.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.10,255.81) -- (359.10,262.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.11,261.70) -- (371.18,261.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.65,258.16) -- (367.65,265.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.66,264.01) -- (379.72,264.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.19,260.48) -- (376.19,267.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.20,266.28) -- (388.26,266.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.73,262.75) -- (384.73,269.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.74,268.50) -- (396.81,268.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.27,264.97) -- (393.27,272.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.28,270.64) -- (405.35,270.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.82,267.11) -- (401.82,274.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.83,272.70) -- (413.89,272.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.36,269.17) -- (410.36,276.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (415.37,274.66) -- (422.43,274.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.90,271.13) -- (418.90,278.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.91,276.50) -- (430.97,276.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.44,272.97) -- (427.44,280.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (432.45,278.21) -- (439.52,278.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.98,274.67) -- (435.98,281.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (440.99,279.76) -- (448.06,279.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.53,276.23) -- (444.53,283.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (449.54,281.15) -- (456.60,281.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.07,277.61) -- (453.07,284.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.08,282.34) -- (465.14,282.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.61,278.81) -- (461.61,285.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (466.62,283.32) -- (473.69,283.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.15,279.79) -- (470.15,286.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.16,284.08) -- (482.23,284.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.70,280.55) -- (478.70,287.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.71,284.58) -- (490.77,284.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.24,281.04) -- (487.24,288.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (492.25,284.80) -- (499.31,284.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.78,281.27) -- (495.78,288.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.79,284.72) -- (507.85,284.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (504.32,281.19) -- (504.32,288.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (509.33,284.32) -- (516.40,284.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (512.86,280.79) -- (512.86,287.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (517.87,283.58) -- (524.94,283.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (521.41,280.04) -- (521.41,287.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.42,282.47) -- (533.48,282.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (529.95,278.94) -- (529.95,286.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (534.96,280.98) -- (542.02,280.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (538.49,277.44) -- (538.49,284.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (543.50,279.09) -- (550.57,279.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.03,275.56) -- (547.03,282.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.04,276.80) -- (559.11,276.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (555.58,273.26) -- (555.58,280.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.59,274.10) -- (567.65,274.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.12,270.57) -- (564.12,277.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13,271.01) -- (576.19,271.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66,267.47) -- (572.66,274.54);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (316.39,247.50) circle (  2.50);

\path[fill=fillColor] (324.94,249.88) circle (  2.50);

\path[fill=fillColor] (333.48,252.36) circle (  2.50);

\path[fill=fillColor] (342.02,254.96) circle (  2.50);

\path[fill=fillColor] (350.56,257.65) circle (  2.50);

\path[fill=fillColor] (359.10,260.48) circle (  2.50);

\path[fill=fillColor] (367.65,263.42) circle (  2.50);

\path[fill=fillColor] (376.19,266.48) circle (  2.50);

\path[fill=fillColor] (384.73,269.65) circle (  2.50);

\path[fill=fillColor] (393.27,273.04) circle (  2.50);

\path[fill=fillColor] (401.82,276.42) circle (  2.50);

\path[fill=fillColor] (410.36,280.04) circle (  2.50);

\path[fill=fillColor] (418.90,283.76) circle (  2.50);

\path[fill=fillColor] (427.44,287.65) circle (  2.50);

\path[fill=fillColor] (435.98,291.72) circle (  2.50);

\path[fill=fillColor] (444.53,296.05) circle (  2.50);

\path[fill=fillColor] (453.07,300.28) circle (  2.50);

\path[fill=fillColor] (461.61,305.03) circle (  2.50);

\path[fill=fillColor] (470.15,309.81) circle (  2.50);

\path[fill=fillColor] (478.70,314.71) circle (  2.50);

\path[fill=fillColor] (487.24,319.74) circle (  2.50);

\path[fill=fillColor] (495.78,324.97) circle (  2.50);

\path[fill=fillColor] (504.32,330.80) circle (  2.50);

\path[fill=fillColor] (512.86,336.10) circle (  2.50);

\path[fill=fillColor] (521.41,342.50) circle (  2.50);

\path[fill=fillColor] (529.95,348.19) circle (  2.50);

\path[fill=fillColor] (538.49,354.89) circle (  2.50);

\path[fill=fillColor] (547.03,361.68) circle (  2.50);

\path[fill=fillColor] (555.58,368.79) circle (  2.50);

\path[fill=fillColor] (564.12,375.90) circle (  2.50);

\path[fill=fillColor] (572.66,383.19) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.39,247.50) rectangle (572.66,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,247.50) --
	(316.39,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,294.20) --
	(316.39,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,340.91) --
	(316.39,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,387.61) --
	(316.39,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.39,244.75) --
	(316.39,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.82,244.75) --
	(401.82,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.24,244.75) --
	(487.24,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (316.39,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (401.82,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (487.24,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (444.53,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,329.23) {Naive reproduction number ${\mathcal R}_{\textrm{\tiny naive}}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (316.39,419.03) {B};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 27.31, 30.69) rectangle (283.58,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31, 30.69) rectangle (283.58,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31, 54.39) --
	( 30.56, 54.42) --
	( 33.80, 54.45) --
	( 37.04, 54.47) --
	( 40.29, 54.50) --
	( 43.53, 54.52) --
	( 46.78, 54.54) --
	( 50.02, 54.56) --
	( 53.26, 54.58) --
	( 56.51, 54.59) --
	( 59.75, 54.61) --
	( 63.00, 54.62) --
	( 66.24, 54.63) --
	( 69.48, 54.64) --
	( 72.73, 54.65) --
	( 75.97, 54.66) --
	( 79.22, 54.66) --
	( 82.46, 54.67) --
	( 85.70, 54.68) --
	( 88.95, 54.68) --
	( 92.19, 54.68) --
	( 95.43, 54.68) --
	( 98.68, 54.67) --
	(101.92, 54.67) --
	(105.17, 54.66) --
	(108.41, 54.65) --
	(111.65, 54.64) --
	(114.90, 54.63) --
	(118.14, 54.63) --
	(121.39, 54.63) --
	(124.63, 54.64) --
	(127.87, 54.65) --
	(131.12, 54.66) --
	(134.36, 54.67) --
	(137.61, 54.69) --
	(140.85, 54.71) --
	(144.09, 54.72) --
	(147.34, 54.74) --
	(150.58, 54.75) --
	(153.82, 54.76) --
	(157.07, 54.76) --
	(160.31, 54.76) --
	(163.56, 54.76) --
	(166.80, 54.76) --
	(170.04, 54.76) --
	(173.29, 54.76) --
	(176.53, 54.76) --
	(179.78, 54.76) --
	(183.02, 54.75) --
	(186.26, 54.75) --
	(189.51, 54.75) --
	(192.75, 54.75) --
	(196.00, 54.75) --
	(199.24, 54.74) --
	(202.48, 54.74) --
	(205.73, 54.74) --
	(208.97, 54.74) --
	(212.21, 54.73) --
	(215.46, 54.73) --
	(218.70, 54.73) --
	(221.95, 54.73) --
	(225.19, 54.73) --
	(228.43, 54.72) --
	(231.68, 54.72) --
	(234.92, 54.72) --
	(238.17, 54.72) --
	(241.41, 54.71) --
	(244.65, 54.71) --
	(247.90, 54.71) --
	(251.14, 54.71) --
	(254.39, 54.70) --
	(257.63, 54.70) --
	(260.87, 54.70) --
	(264.12, 54.70) --
	(267.36, 54.70) --
	(270.60, 54.70) --
	(273.85, 54.70) --
	(277.09, 54.70) --
	(280.34, 54.70) --
	(283.58, 54.70);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 27.31, 55.70) --
	( 30.56, 56.81) --
	( 33.80, 57.90) --
	( 37.04, 58.98) --
	( 40.29, 60.06) --
	( 43.53, 61.12) --
	( 46.78, 62.16) --
	( 50.02, 63.20) --
	( 53.26, 64.22) --
	( 56.51, 65.23) --
	( 59.75, 66.23) --
	( 63.00, 67.21) --
	( 66.24, 68.18) --
	( 69.48, 69.14) --
	( 72.73, 70.09) --
	( 75.97, 71.03) --
	( 79.22, 71.95) --
	( 82.46, 72.86) --
	( 85.70, 73.77) --
	( 88.95, 74.66) --
	( 92.19, 75.53) --
	( 95.43, 76.39) --
	( 98.68, 77.24) --
	(101.92, 78.07) --
	(105.17, 78.90) --
	(108.41, 79.71) --
	(111.65, 80.51) --
	(114.90, 81.30) --
	(118.14, 82.08) --
	(121.39, 82.85) --
	(124.63, 83.61) --
	(127.87, 84.35) --
	(131.12, 85.08) --
	(134.36, 85.79) --
	(137.61, 86.49) --
	(140.85, 87.19) --
	(144.09, 87.87) --
	(147.34, 88.55) --
	(150.58, 89.23) --
	(153.82, 89.91) --
	(157.07, 90.59) --
	(160.31, 91.26) --
	(163.56, 91.92) --
	(166.80, 92.57) --
	(170.04, 93.22) --
	(173.29, 93.86) --
	(176.53, 94.49) --
	(179.78, 95.12) --
	(183.02, 95.75) --
	(186.26, 96.37) --
	(189.51, 96.99) --
	(192.75, 97.61) --
	(196.00, 98.22) --
	(199.24, 98.82) --
	(202.48, 99.42) --
	(205.73,100.02) --
	(208.97,100.61) --
	(212.21,101.20) --
	(215.46,101.78) --
	(218.70,102.35) --
	(221.95,102.93) --
	(225.19,103.50) --
	(228.43,104.06) --
	(231.68,104.62) --
	(234.92,105.17) --
	(238.17,105.72) --
	(241.41,106.27) --
	(244.65,106.81) --
	(247.90,107.34) --
	(251.14,107.87) --
	(254.39,108.40) --
	(257.63,108.92) --
	(260.87,109.43) --
	(264.12,109.95) --
	(267.36,110.45) --
	(270.60,110.95) --
	(273.85,111.45) --
	(277.09,111.95) --
	(280.34,112.43) --
	(283.58,112.92);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 27.31, 55.54) --
	( 30.56, 57.34) --
	( 33.80, 59.12) --
	( 37.04, 60.88) --
	( 40.29, 62.63) --
	( 43.53, 64.35) --
	( 46.78, 66.06) --
	( 50.02, 67.75) --
	( 53.26, 69.43) --
	( 56.51, 71.08) --
	( 59.75, 72.71) --
	( 63.00, 74.33) --
	( 66.24, 75.93) --
	( 69.48, 77.51) --
	( 72.73, 79.07) --
	( 75.97, 80.62) --
	( 79.22, 82.15) --
	( 82.46, 83.66) --
	( 85.70, 85.16) --
	( 88.95, 86.63) --
	( 92.19, 88.09) --
	( 95.43, 89.53) --
	( 98.68, 90.94) --
	(101.92, 92.34) --
	(105.17, 93.72) --
	(108.41, 95.08) --
	(111.65, 96.43) --
	(114.90, 97.76) --
	(118.14, 99.08) --
	(121.39,100.38) --
	(124.63,101.67) --
	(127.87,102.92) --
	(131.12,104.16) --
	(134.36,105.37) --
	(137.61,106.57) --
	(140.85,107.76) --
	(144.09,108.93) --
	(147.34,110.10) --
	(150.58,111.27) --
	(153.82,112.43) --
	(157.07,113.59) --
	(160.31,114.75) --
	(163.56,115.89) --
	(166.80,117.02) --
	(170.04,118.14) --
	(173.29,119.26) --
	(176.53,120.36) --
	(179.78,121.46) --
	(183.02,122.55) --
	(186.26,123.63) --
	(189.51,124.71) --
	(192.75,125.78) --
	(196.00,126.85) --
	(199.24,127.91) --
	(202.48,128.96) --
	(205.73,130.00) --
	(208.97,131.03) --
	(212.21,132.06) --
	(215.46,133.08) --
	(218.70,134.09) --
	(221.95,135.10) --
	(225.19,136.09) --
	(228.43,137.09) --
	(231.68,138.07) --
	(234.92,139.04) --
	(238.17,140.01) --
	(241.41,140.97) --
	(244.65,141.92) --
	(247.90,142.87) --
	(251.14,143.81) --
	(254.39,144.74) --
	(257.63,145.66) --
	(260.87,146.58) --
	(264.12,147.48) --
	(267.36,148.38) --
	(270.60,149.28) --
	(273.85,150.16) --
	(277.09,151.04) --
	(280.34,151.91) --
	(283.58,152.77);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 55.20) --
	( 30.56, 57.64) --
	( 33.80, 60.07) --
	( 37.04, 62.47) --
	( 40.29, 64.85) --
	( 43.53, 67.21) --
	( 46.78, 69.55) --
	( 50.02, 71.86) --
	( 53.26, 74.15) --
	( 56.51, 76.43) --
	( 59.75, 78.67) --
	( 63.00, 80.90) --
	( 66.24, 83.10) --
	( 69.48, 85.29) --
	( 72.73, 87.45) --
	( 75.97, 89.59) --
	( 79.22, 91.71) --
	( 82.46, 93.81) --
	( 85.70, 95.89) --
	( 88.95, 97.94) --
	( 92.19, 99.98) --
	( 95.43,101.98) --
	( 98.68,103.97) --
	(101.92,105.93) --
	(105.17,107.87) --
	(108.41,109.79) --
	(111.65,111.69) --
	(114.90,113.57) --
	(118.14,115.43) --
	(121.39,117.28) --
	(124.63,119.10) --
	(127.87,120.88) --
	(131.12,122.64) --
	(134.36,124.37) --
	(137.61,126.09) --
	(140.85,127.78) --
	(144.09,129.47) --
	(147.34,131.15) --
	(150.58,132.82) --
	(153.82,134.50) --
	(157.07,136.18) --
	(160.31,137.85) --
	(163.56,139.50) --
	(166.80,141.14) --
	(170.04,142.77) --
	(173.29,144.39) --
	(176.53,146.00) --
	(179.78,147.60) --
	(183.02,149.19) --
	(186.26,150.78) --
	(189.51,152.36) --
	(192.75,153.93) --
	(196.00,155.49) --
	(199.24,157.04) --
	(202.48,158.58) --
	(205.73,160.11) --
	(208.97,161.63) --
	(212.21,163.15) --
	(215.46,164.65) --
	(218.70,166.14) --
	(221.95,167.63) --
	(225.19,169.10) --
	(228.43,170.57) --
	(231.68,172.02) --
	(234.92,173.47) --
	(238.17,174.91) --
	(241.41,176.33) --
	(244.65,177.75) --
	(247.90,179.16) --
	(251.14,180.56) --
	(254.39,181.95) --
	(257.63,183.33) --
	(260.87,184.70) --
	(264.12,186.06) --
	(267.36,187.41) --
	(270.60,188.75) --
	(273.85,190.08) --
	(277.09,191.40) --
	(280.34,192.72) --
	(283.58,194.02);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 27.31, 59.60) --
	( 30.68, 53.77) --
	( 23.95, 53.77) --
	cycle;

\path[fill=fillColor] ( 35.86, 62.47) --
	( 39.22, 56.65) --
	( 32.49, 56.65) --
	cycle;

\path[fill=fillColor] ( 44.40, 65.07) --
	( 47.76, 59.25) --
	( 41.03, 59.25) --
	cycle;

\path[fill=fillColor] ( 52.94, 68.02) --
	( 56.30, 62.19) --
	( 49.58, 62.19) --
	cycle;

\path[fill=fillColor] ( 61.48, 70.77) --
	( 64.85, 64.95) --
	( 58.12, 64.95) --
	cycle;

\path[fill=fillColor] ( 70.02, 73.26) --
	( 73.39, 67.44) --
	( 66.66, 67.44) --
	cycle;

\path[fill=fillColor] ( 78.57, 75.65) --
	( 81.93, 69.83) --
	( 75.20, 69.83) --
	cycle;

\path[fill=fillColor] ( 87.11, 78.08) --
	( 90.47, 72.25) --
	( 83.75, 72.25) --
	cycle;

\path[fill=fillColor] ( 95.65, 80.50) --
	( 99.01, 74.67) --
	( 92.29, 74.67) --
	cycle;

\path[fill=fillColor] (104.19, 82.71) --
	(107.56, 76.88) --
	(100.83, 76.88) --
	cycle;

\path[fill=fillColor] (112.74, 84.81) --
	(116.10, 78.99) --
	(109.37, 78.99) --
	cycle;

\path[fill=fillColor] (121.28, 86.66) --
	(124.64, 80.84) --
	(117.91, 80.84) --
	cycle;

\path[fill=fillColor] (129.82, 88.60) --
	(133.18, 82.78) --
	(126.46, 82.78) --
	cycle;

\path[fill=fillColor] (138.36, 90.48) --
	(141.73, 84.65) --
	(135.00, 84.65) --
	cycle;

\path[fill=fillColor] (146.90, 92.35) --
	(150.27, 86.53) --
	(143.54, 86.53) --
	cycle;

\path[fill=fillColor] (155.45, 94.13) --
	(158.81, 88.30) --
	(152.08, 88.30) --
	cycle;

\path[fill=fillColor] (163.99, 95.79) --
	(167.35, 89.97) --
	(160.63, 89.97) --
	cycle;

\path[fill=fillColor] (172.53, 97.50) --
	(175.89, 91.67) --
	(169.17, 91.67) --
	cycle;

\path[fill=fillColor] (181.07, 99.31) --
	(184.44, 93.48) --
	(177.71, 93.48) --
	cycle;

\path[fill=fillColor] (189.62,100.91) --
	(192.98, 95.09) --
	(186.25, 95.09) --
	cycle;

\path[fill=fillColor] (198.16,102.55) --
	(201.52, 96.72) --
	(194.79, 96.72) --
	cycle;

\path[fill=fillColor] (206.70,104.08) --
	(210.06, 98.25) --
	(203.34, 98.25) --
	cycle;

\path[fill=fillColor] (215.24,105.58) --
	(218.61, 99.75) --
	(211.88, 99.75) --
	cycle;

\path[fill=fillColor] (223.78,107.13) --
	(227.15,101.30) --
	(220.42,101.30) --
	cycle;

\path[fill=fillColor] (232.33,108.61) --
	(235.69,102.78) --
	(228.96,102.78) --
	cycle;

\path[fill=fillColor] (240.87,109.99) --
	(244.23,104.16) --
	(237.51,104.16) --
	cycle;

\path[fill=fillColor] (249.41,111.47) --
	(252.77,105.64) --
	(246.05,105.64) --
	cycle;

\path[fill=fillColor] (257.95,112.80) --
	(261.32,106.97) --
	(254.59,106.97) --
	cycle;

\path[fill=fillColor] (266.50,114.25) --
	(269.86,108.43) --
	(263.13,108.43) --
	cycle;

\path[fill=fillColor] (275.04,115.51) --
	(278.40,109.68) --
	(271.67,109.68) --
	cycle;

\path[fill=fillColor] (283.58,116.83) --
	(286.94,111.01) --
	(280.22,111.01) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 24.82, 53.14) --
	( 29.81, 53.14) --
	( 29.81, 58.14) --
	( 24.82, 58.14) --
	cycle;

\path[fill=fillColor] ( 33.36, 57.75) --
	( 38.35, 57.75) --
	( 38.35, 62.75) --
	( 33.36, 62.75) --
	cycle;

\path[fill=fillColor] ( 41.90, 61.97) --
	( 46.90, 61.97) --
	( 46.90, 66.96) --
	( 41.90, 66.96) --
	cycle;

\path[fill=fillColor] ( 50.44, 66.76) --
	( 55.44, 66.76) --
	( 55.44, 71.75) --
	( 50.44, 71.75) --
	cycle;

\path[fill=fillColor] ( 58.98, 71.20) --
	( 63.98, 71.20) --
	( 63.98, 76.19) --
	( 58.98, 76.19) --
	cycle;

\path[fill=fillColor] ( 67.53, 75.33) --
	( 72.52, 75.33) --
	( 72.52, 80.32) --
	( 67.53, 80.32) --
	cycle;

\path[fill=fillColor] ( 76.07, 79.35) --
	( 81.06, 79.35) --
	( 81.06, 84.34) --
	( 76.07, 84.34) --
	cycle;

\path[fill=fillColor] ( 84.61, 83.44) --
	( 89.61, 83.44) --
	( 89.61, 88.43) --
	( 84.61, 88.43) --
	cycle;

\path[fill=fillColor] ( 93.15, 87.44) --
	( 98.15, 87.44) --
	( 98.15, 92.44) --
	( 93.15, 92.44) --
	cycle;

\path[fill=fillColor] (101.70, 91.12) --
	(106.69, 91.12) --
	(106.69, 96.12) --
	(101.70, 96.12) --
	cycle;

\path[fill=fillColor] (110.24, 94.51) --
	(115.23, 94.51) --
	(115.23, 99.50) --
	(110.24, 99.50) --
	cycle;

\path[fill=fillColor] (118.78, 97.80) --
	(123.78, 97.80) --
	(123.78,102.79) --
	(118.78,102.79) --
	cycle;

\path[fill=fillColor] (127.32,101.03) --
	(132.32,101.03) --
	(132.32,106.02) --
	(127.32,106.02) --
	cycle;

\path[fill=fillColor] (135.86,104.29) --
	(140.86,104.29) --
	(140.86,109.28) --
	(135.86,109.28) --
	cycle;

\path[fill=fillColor] (144.41,107.42) --
	(149.40,107.42) --
	(149.40,112.41) --
	(144.41,112.41) --
	cycle;

\path[fill=fillColor] (152.95,110.55) --
	(157.94,110.55) --
	(157.94,115.55) --
	(152.95,115.55) --
	cycle;

\path[fill=fillColor] (161.49,113.41) --
	(166.49,113.41) --
	(166.49,118.40) --
	(161.49,118.40) --
	cycle;

\path[fill=fillColor] (170.03,116.27) --
	(175.03,116.27) --
	(175.03,121.26) --
	(170.03,121.26) --
	cycle;

\path[fill=fillColor] (178.58,119.47) --
	(183.57,119.47) --
	(183.57,124.47) --
	(178.58,124.47) --
	cycle;

\path[fill=fillColor] (187.12,122.28) --
	(192.11,122.28) --
	(192.11,127.28) --
	(187.12,127.28) --
	cycle;

\path[fill=fillColor] (195.66,125.17) --
	(200.66,125.17) --
	(200.66,130.16) --
	(195.66,130.16) --
	cycle;

\path[fill=fillColor] (204.20,127.80) --
	(209.20,127.80) --
	(209.20,132.80) --
	(204.20,132.80) --
	cycle;

\path[fill=fillColor] (212.74,130.49) --
	(217.74,130.49) --
	(217.74,135.48) --
	(212.74,135.48) --
	cycle;

\path[fill=fillColor] (221.29,133.19) --
	(226.28,133.19) --
	(226.28,138.18) --
	(221.29,138.18) --
	cycle;

\path[fill=fillColor] (229.83,135.74) --
	(234.82,135.74) --
	(234.82,140.73) --
	(229.83,140.73) --
	cycle;

\path[fill=fillColor] (238.37,138.17) --
	(243.37,138.17) --
	(243.37,143.16) --
	(238.37,143.16) --
	cycle;

\path[fill=fillColor] (246.91,140.85) --
	(251.91,140.85) --
	(251.91,145.85) --
	(246.91,145.85) --
	cycle;

\path[fill=fillColor] (255.46,143.24) --
	(260.45,143.24) --
	(260.45,148.23) --
	(255.46,148.23) --
	cycle;

\path[fill=fillColor] (264.00,145.71) --
	(268.99,145.71) --
	(268.99,150.71) --
	(264.00,150.71) --
	cycle;

\path[fill=fillColor] (272.54,147.90) --
	(277.54,147.90) --
	(277.54,152.89) --
	(272.54,152.89) --
	cycle;

\path[fill=fillColor] (281.08,150.34) --
	(286.08,150.34) --
	(286.08,155.34) --
	(281.08,155.34) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 23.78, 55.41) -- ( 30.85, 55.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.31, 51.87) -- ( 27.31, 58.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.32, 61.62) -- ( 39.39, 61.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.86, 58.09) -- ( 35.86, 65.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87, 67.34) -- ( 47.93, 67.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.40, 63.81) -- ( 44.40, 70.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.41, 73.88) -- ( 56.47, 73.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.94, 70.35) -- ( 52.94, 77.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.95, 79.89) -- ( 65.01, 79.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.48, 76.36) -- ( 61.48, 83.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.49, 85.65) -- ( 73.56, 85.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.02, 82.12) -- ( 70.02, 89.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.03, 91.28) -- ( 82.10, 91.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.57, 87.75) -- ( 78.57, 94.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.58, 97.06) -- ( 90.64, 97.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.11, 93.53) -- ( 87.11,100.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.12,102.60) -- ( 99.18,102.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65, 99.07) -- ( 95.65,106.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.66,107.76) -- (107.73,107.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.19,104.23) -- (104.19,111.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.20,112.34) -- (116.27,112.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.74,108.81) -- (112.74,115.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.75,117.19) -- (124.81,117.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.28,113.66) -- (121.28,120.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.29,121.71) -- (133.35,121.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.82,118.17) -- (129.82,125.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.83,126.41) -- (141.89,126.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.36,122.88) -- (138.36,129.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.37,130.85) -- (150.44,130.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.90,127.32) -- (146.90,134.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.91,135.49) -- (158.98,135.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (155.45,131.96) -- (155.45,139.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.46,139.55) -- (167.52,139.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.99,136.02) -- (163.99,143.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.00,143.62) -- (176.06,143.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.53,140.09) -- (172.53,147.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.54,148.33) -- (184.61,148.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.07,144.80) -- (181.07,151.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.08,152.45) -- (193.15,152.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.62,148.92) -- (189.62,155.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.63,156.73) -- (201.69,156.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.16,153.19) -- (198.16,160.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.17,160.57) -- (210.23,160.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.70,157.04) -- (206.70,164.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.71,164.54) -- (218.77,164.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.24,161.01) -- (215.24,168.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.25,168.54) -- (227.32,168.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.78,165.01) -- (223.78,172.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.79,172.25) -- (235.86,172.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.33,168.71) -- (232.33,175.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.34,175.81) -- (244.40,175.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.87,172.28) -- (240.87,179.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.88,179.95) -- (252.94,179.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.41,176.42) -- (249.41,183.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.42,183.53) -- (261.49,183.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.95,179.99) -- (257.95,187.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.96,187.15) -- (270.03,187.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (266.50,183.62) -- (266.50,190.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.51,190.32) -- (278.57,190.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.04,186.78) -- (275.04,193.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.05,194.15) -- (287.11,194.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.58,190.62) -- (283.58,197.68);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 27.31, 53.74) circle (  2.50);

\path[fill=fillColor] ( 35.86, 54.86) circle (  2.50);

\path[fill=fillColor] ( 44.40, 54.88) circle (  2.50);

\path[fill=fillColor] ( 52.94, 53.99) circle (  2.50);

\path[fill=fillColor] ( 61.48, 55.34) circle (  2.50);

\path[fill=fillColor] ( 70.02, 54.65) circle (  2.50);

\path[fill=fillColor] ( 78.57, 54.68) circle (  2.50);

\path[fill=fillColor] ( 87.11, 55.24) circle (  2.50);

\path[fill=fillColor] ( 95.65, 54.54) circle (  2.50);

\path[fill=fillColor] (104.19, 54.90) circle (  2.50);

\path[fill=fillColor] (112.74, 54.54) circle (  2.50);

\path[fill=fillColor] (121.28, 54.06) circle (  2.50);

\path[fill=fillColor] (129.82, 54.15) circle (  2.50);

\path[fill=fillColor] (138.36, 55.07) circle (  2.50);

\path[fill=fillColor] (146.90, 54.49) circle (  2.50);

\path[fill=fillColor] (155.45, 55.12) circle (  2.50);

\path[fill=fillColor] (163.99, 55.17) circle (  2.50);

\path[fill=fillColor] (172.53, 54.90) circle (  2.50);

\path[fill=fillColor] (181.07, 54.75) circle (  2.50);

\path[fill=fillColor] (189.62, 54.64) circle (  2.50);

\path[fill=fillColor] (198.16, 54.75) circle (  2.50);

\path[fill=fillColor] (206.70, 54.31) circle (  2.50);

\path[fill=fillColor] (215.24, 54.85) circle (  2.50);

\path[fill=fillColor] (223.78, 54.36) circle (  2.50);

\path[fill=fillColor] (232.33, 54.41) circle (  2.50);

\path[fill=fillColor] (240.87, 54.96) circle (  2.50);

\path[fill=fillColor] (249.41, 55.07) circle (  2.50);

\path[fill=fillColor] (257.95, 54.79) circle (  2.50);

\path[fill=fillColor] (266.50, 55.03) circle (  2.50);

\path[fill=fillColor] (275.04, 54.59) circle (  2.50);

\path[fill=fillColor] (283.58, 54.42) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 30.69) rectangle (283.58,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 52.22) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,134.09) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 55.25) --
	( 27.31, 55.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,137.12) --
	( 27.31,137.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 27.94) --
	( 27.31, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.74, 27.94) --
	(112.74, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.16, 27.94) --
	(198.16, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.74, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.16, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (155.45,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,112.42) {Mean generation/serial interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,202.22) {C};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,216.81);
\end{scope}
\begin{scope}
\path[clip] (323.24, 30.69) rectangle (572.66,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (323.24, 30.69) rectangle (572.66,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (323.24, 68.09) --
	(326.39, 68.10) --
	(329.55, 68.10) --
	(332.71, 68.11) --
	(335.87, 68.12) --
	(339.02, 68.12) --
	(342.18, 68.13) --
	(345.34, 68.14) --
	(348.49, 68.14) --
	(351.65, 68.14) --
	(354.81, 68.15) --
	(357.97, 68.15) --
	(361.12, 68.15) --
	(364.28, 68.16) --
	(367.44, 68.16) --
	(370.60, 68.16) --
	(373.75, 68.16) --
	(376.91, 68.16) --
	(380.07, 68.16) --
	(383.22, 68.16) --
	(386.38, 68.16) --
	(389.54, 68.16) --
	(392.70, 68.16) --
	(395.85, 68.16) --
	(399.01, 68.16) --
	(402.17, 68.16) --
	(405.32, 68.15) --
	(408.48, 68.15) --
	(411.64, 68.15) --
	(414.80, 68.14) --
	(417.95, 68.13) --
	(421.11, 68.12) --
	(424.27, 68.11) --
	(427.43, 68.09) --
	(430.58, 68.08) --
	(433.74, 68.06) --
	(436.90, 68.04) --
	(440.05, 68.03) --
	(443.21, 68.02) --
	(446.37, 68.01) --
	(449.53, 68.01) --
	(452.68, 68.01) --
	(455.84, 68.01) --
	(459.00, 68.02) --
	(462.16, 68.02) --
	(465.31, 68.03) --
	(468.47, 68.04) --
	(471.63, 68.04) --
	(474.78, 68.05) --
	(477.94, 68.06) --
	(481.10, 68.06) --
	(484.26, 68.07) --
	(487.41, 68.08) --
	(490.57, 68.09) --
	(493.73, 68.10) --
	(496.89, 68.11) --
	(500.04, 68.12) --
	(503.20, 68.13) --
	(506.36, 68.13) --
	(509.51, 68.14) --
	(512.67, 68.14) --
	(515.83, 68.15) --
	(518.99, 68.15) --
	(522.14, 68.15) --
	(525.30, 68.15) --
	(528.46, 68.15) --
	(531.62, 68.14) --
	(534.77, 68.14) --
	(537.93, 68.14) --
	(541.09, 68.14) --
	(544.24, 68.13) --
	(547.40, 68.12) --
	(550.56, 68.12) --
	(553.72, 68.11) --
	(556.87, 68.10) --
	(560.03, 68.08) --
	(563.19, 68.07) --
	(566.35, 68.06) --
	(569.50, 68.04) --
	(572.66, 68.03);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (323.24,142.26) --
	(326.39,141.25) --
	(329.55,140.26) --
	(332.71,139.28) --
	(335.87,138.32) --
	(339.02,137.38) --
	(342.18,136.45) --
	(345.34,135.55) --
	(348.49,134.65) --
	(351.65,133.78) --
	(354.81,132.92) --
	(357.97,132.09) --
	(361.12,131.26) --
	(364.28,130.46) --
	(367.44,129.67) --
	(370.60,128.90) --
	(373.75,128.14) --
	(376.91,127.40) --
	(380.07,126.68) --
	(383.22,125.97) --
	(386.38,125.28) --
	(389.54,124.61) --
	(392.70,123.97) --
	(395.85,123.34) --
	(399.01,122.73) --
	(402.17,122.13) --
	(405.32,121.55) --
	(408.48,120.98) --
	(411.64,120.42) --
	(414.80,119.87) --
	(417.95,119.34) --
	(421.11,118.83) --
	(424.27,118.34) --
	(427.43,117.87) --
	(430.58,117.41) --
	(433.74,116.96) --
	(436.90,116.52) --
	(440.05,116.09) --
	(443.21,115.65) --
	(446.37,115.22) --
	(449.53,114.78) --
	(452.68,114.36) --
	(455.84,113.94) --
	(459.00,113.54) --
	(462.16,113.14) --
	(465.31,112.75) --
	(468.47,112.37) --
	(471.63,111.99) --
	(474.78,111.61) --
	(477.94,111.24) --
	(481.10,110.87) --
	(484.26,110.49) --
	(487.41,110.13) --
	(490.57,109.77) --
	(493.73,109.41) --
	(496.89,109.06) --
	(500.04,108.71) --
	(503.20,108.37) --
	(506.36,108.03) --
	(509.51,107.70) --
	(512.67,107.37) --
	(515.83,107.04) --
	(518.99,106.72) --
	(522.14,106.40) --
	(525.30,106.09) --
	(528.46,105.78) --
	(531.62,105.48) --
	(534.77,105.18) --
	(537.93,104.88) --
	(541.09,104.59) --
	(544.24,104.31) --
	(547.40,104.03) --
	(550.56,103.75) --
	(553.72,103.48) --
	(556.87,103.21) --
	(560.03,102.94) --
	(563.19,102.69) --
	(566.35,102.43) --
	(569.50,102.18) --
	(572.66,101.93);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (323.24,186.85) --
	(326.39,184.91) --
	(329.55,183.00) --
	(332.71,181.13) --
	(335.87,179.29) --
	(339.02,177.48) --
	(342.18,175.71) --
	(345.34,173.98) --
	(348.49,172.28) --
	(351.65,170.62) --
	(354.81,168.99) --
	(357.97,167.40) --
	(361.12,165.84) --
	(364.28,164.32) --
	(367.44,162.83) --
	(370.60,161.37) --
	(373.75,159.95) --
	(376.91,158.56) --
	(380.07,157.20) --
	(383.22,155.87) --
	(386.38,154.59) --
	(389.54,153.35) --
	(392.70,152.15) --
	(395.85,150.99) --
	(399.01,149.87) --
	(402.17,148.77) --
	(405.32,147.71) --
	(408.48,146.66) --
	(411.64,145.64) --
	(414.80,144.64) --
	(417.95,143.67) --
	(421.11,142.74) --
	(424.27,141.86) --
	(427.43,141.01) --
	(430.58,140.19) --
	(433.74,139.39) --
	(436.90,138.60) --
	(440.05,137.83) --
	(443.21,137.06) --
	(446.37,136.28) --
	(449.53,135.51) --
	(452.68,134.75) --
	(455.84,134.01) --
	(459.00,133.30) --
	(462.16,132.60) --
	(465.31,131.92) --
	(468.47,131.24) --
	(471.63,130.58) --
	(474.78,129.92) --
	(477.94,129.27) --
	(481.10,128.62) --
	(484.26,127.97) --
	(487.41,127.33) --
	(490.57,126.70) --
	(493.73,126.09) --
	(496.89,125.48) --
	(500.04,124.89) --
	(503.20,124.30) --
	(506.36,123.72) --
	(509.51,123.14) --
	(512.67,122.58) --
	(515.83,122.02) --
	(518.99,121.47) --
	(522.14,120.93) --
	(525.30,120.40) --
	(528.46,119.87) --
	(531.62,119.36) --
	(534.77,118.86) --
	(537.93,118.36) --
	(541.09,117.87) --
	(544.24,117.39) --
	(547.40,116.92) --
	(550.56,116.46) --
	(553.72,116.00) --
	(556.87,115.56) --
	(560.03,115.12) --
	(563.19,114.69) --
	(566.35,114.27) --
	(569.50,113.86) --
	(572.66,113.46);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (364.81,192.52) --
	(367.44,190.83) --
	(370.07,189.17) --
	(372.70,187.54) --
	(375.33,185.94) --
	(377.96,184.36) --
	(380.59,182.82) --
	(383.22,181.30) --
	(385.86,179.80) --
	(388.49,178.34) --
	(391.12,176.90) --
	(393.75,175.49) --
	(396.38,174.11) --
	(399.01,172.76) --
	(401.64,171.43) --
	(404.27,170.13) --
	(406.90,168.86) --
	(409.53,167.61) --
	(412.17,166.38) --
	(414.80,165.18) --
	(417.43,164.00) --
	(420.06,162.88) --
	(422.69,161.79) --
	(425.32,160.73) --
	(427.95,159.70) --
	(430.58,158.69) --
	(433.21,157.70) --
	(435.85,156.73) --
	(438.48,155.76) --
	(441.11,154.80) --
	(443.74,153.86) --
	(446.37,152.96) --
	(449.00,152.08) --
	(451.63,151.21) --
	(454.26,150.37) --
	(456.89,149.53) --
	(459.52,148.70) --
	(462.16,147.87) --
	(464.79,147.03) --
	(467.42,146.20) --
	(470.05,145.39) --
	(472.68,144.58) --
	(475.31,143.79) --
	(477.94,143.00) --
	(480.57,142.23) --
	(483.20,141.46) --
	(485.84,140.71) --
	(488.47,139.96) --
	(491.10,139.22) --
	(493.73,138.50) --
	(496.36,137.79) --
	(498.99,137.08) --
	(501.62,136.40) --
	(504.25,135.72) --
	(506.88,135.06) --
	(509.51,134.41) --
	(512.15,133.76) --
	(514.78,133.13) --
	(517.41,132.50) --
	(520.04,131.88) --
	(522.67,131.26) --
	(525.30,130.65) --
	(527.93,130.05) --
	(530.56,129.46) --
	(533.19,128.87) --
	(535.83,128.30) --
	(538.46,127.73) --
	(541.09,127.17) --
	(543.72,126.62) --
	(546.35,126.08) --
	(548.98,125.55) --
	(551.61,125.02) --
	(554.24,124.51) --
	(556.87,124.00) --
	(559.50,123.49) --
	(562.14,123.00) --
	(564.77,122.52) --
	(567.40,122.04) --
	(570.03,121.57) --
	(572.66,121.12);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] (323.24,146.40) --
	(326.60,140.57) --
	(319.87,140.57) --
	cycle;

\path[fill=fillColor] (331.55,143.56) --
	(334.91,137.74) --
	(328.19,137.74) --
	cycle;

\path[fill=fillColor] (339.86,141.07) --
	(343.23,135.25) --
	(336.50,135.25) --
	cycle;

\path[fill=fillColor] (348.18,138.44) --
	(351.54,132.61) --
	(344.81,132.61) --
	cycle;

\path[fill=fillColor] (356.49,136.06) --
	(359.86,130.23) --
	(353.13,130.23) --
	cycle;

\path[fill=fillColor] (364.81,134.02) --
	(368.17,128.19) --
	(361.44,128.19) --
	cycle;

\path[fill=fillColor] (373.12,132.12) --
	(376.48,126.29) --
	(369.76,126.29) --
	cycle;

\path[fill=fillColor] (381.44,130.29) --
	(384.80,124.46) --
	(378.07,124.46) --
	cycle;

\path[fill=fillColor] (389.75,128.36) --
	(393.11,122.53) --
	(386.39,122.53) --
	cycle;

\path[fill=fillColor] (398.06,126.62) --
	(401.43,120.79) --
	(394.70,120.79) --
	cycle;

\path[fill=fillColor] (406.38,125.06) --
	(409.74,119.23) --
	(403.01,119.23) --
	cycle;

\path[fill=fillColor] (414.69,123.72) --
	(418.06,117.89) --
	(411.33,117.89) --
	cycle;

\path[fill=fillColor] (423.01,122.40) --
	(426.37,116.58) --
	(419.64,116.58) --
	cycle;

\path[fill=fillColor] (431.32,121.22) --
	(434.68,115.39) --
	(427.96,115.39) --
	cycle;

\path[fill=fillColor] (439.63,120.06) --
	(443.00,114.23) --
	(436.27,114.23) --
	cycle;

\path[fill=fillColor] (447.95,118.96) --
	(451.31,113.14) --
	(444.58,113.14) --
	cycle;

\path[fill=fillColor] (456.26,117.95) --
	(459.63,112.12) --
	(452.90,112.12) --
	cycle;

\path[fill=fillColor] (464.58,116.87) --
	(467.94,111.05) --
	(461.21,111.05) --
	cycle;

\path[fill=fillColor] (472.89,115.69) --
	(476.25,109.87) --
	(469.53,109.87) --
	cycle;

\path[fill=fillColor] (481.20,114.67) --
	(484.57,108.84) --
	(477.84,108.84) --
	cycle;

\path[fill=fillColor] (489.52,113.75) --
	(492.88,107.92) --
	(486.15,107.92) --
	cycle;

\path[fill=fillColor] (497.83,112.81) --
	(501.20,106.98) --
	(494.47,106.98) --
	cycle;

\path[fill=fillColor] (506.15,111.96) --
	(509.51,106.13) --
	(502.78,106.13) --
	cycle;

\path[fill=fillColor] (514.46,111.02) --
	(517.82,105.19) --
	(511.10,105.19) --
	cycle;

\path[fill=fillColor] (522.78,110.24) --
	(526.14,104.41) --
	(519.41,104.41) --
	cycle;

\path[fill=fillColor] (531.09,109.50) --
	(534.45,103.67) --
	(527.73,103.67) --
	cycle;

\path[fill=fillColor] (539.40,108.72) --
	(542.77,102.89) --
	(536.04,102.89) --
	cycle;

\path[fill=fillColor] (547.72,107.92) --
	(551.08,102.09) --
	(544.35,102.09) --
	cycle;

\path[fill=fillColor] (556.03,107.13) --
	(559.40,101.31) --
	(552.67,101.31) --
	cycle;

\path[fill=fillColor] (564.35,106.48) --
	(567.71,100.65) --
	(560.98,100.65) --
	cycle;

\path[fill=fillColor] (572.66,105.76) --
	(576.02, 99.93) --
	(569.30, 99.93) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] (320.74,184.90) --
	(325.73,184.90) --
	(325.73,189.89) --
	(320.74,189.89) --
	cycle;

\path[fill=fillColor] (329.05,179.38) --
	(334.05,179.38) --
	(334.05,184.38) --
	(329.05,184.38) --
	cycle;

\path[fill=fillColor] (337.37,174.63) --
	(342.36,174.63) --
	(342.36,179.62) --
	(337.37,179.62) --
	cycle;

\path[fill=fillColor] (345.68,169.62) --
	(350.68,169.62) --
	(350.68,174.62) --
	(345.68,174.62) --
	cycle;

\path[fill=fillColor] (353.99,164.98) --
	(358.99,164.98) --
	(358.99,169.97) --
	(353.99,169.97) --
	cycle;

\path[fill=fillColor] (362.31,161.16) --
	(367.30,161.16) --
	(367.30,166.16) --
	(362.31,166.16) --
	cycle;

\path[fill=fillColor] (370.62,157.55) --
	(375.62,157.55) --
	(375.62,162.54) --
	(370.62,162.54) --
	cycle;

\path[fill=fillColor] (378.94,154.05) --
	(383.93,154.05) --
	(383.93,159.04) --
	(378.94,159.04) --
	cycle;

\path[fill=fillColor] (387.25,150.58) --
	(392.25,150.58) --
	(392.25,155.57) --
	(387.25,155.57) --
	cycle;

\path[fill=fillColor] (395.57,147.37) --
	(400.56,147.37) --
	(400.56,152.36) --
	(395.57,152.36) --
	cycle;

\path[fill=fillColor] (403.88,144.46) --
	(408.88,144.46) --
	(408.88,149.46) --
	(403.88,149.46) --
	cycle;

\path[fill=fillColor] (412.19,142.16) --
	(417.19,142.16) --
	(417.19,147.15) --
	(412.19,147.15) --
	cycle;

\path[fill=fillColor] (420.51,139.77) --
	(425.50,139.77) --
	(425.50,144.76) --
	(420.51,144.76) --
	cycle;

\path[fill=fillColor] (428.82,137.57) --
	(433.82,137.57) --
	(433.82,142.56) --
	(428.82,142.56) --
	cycle;

\path[fill=fillColor] (437.14,135.43) --
	(442.13,135.43) --
	(442.13,140.43) --
	(437.14,140.43) --
	cycle;

\path[fill=fillColor] (445.45,133.56) --
	(450.45,133.56) --
	(450.45,138.56) --
	(445.45,138.56) --
	cycle;

\path[fill=fillColor] (453.76,131.74) --
	(458.76,131.74) --
	(458.76,136.74) --
	(453.76,136.74) --
	cycle;

\path[fill=fillColor] (462.08,129.78) --
	(467.07,129.78) --
	(467.07,134.77) --
	(462.08,134.77) --
	cycle;

\path[fill=fillColor] (470.39,127.78) --
	(475.39,127.78) --
	(475.39,132.78) --
	(470.39,132.78) --
	cycle;

\path[fill=fillColor] (478.71,126.01) --
	(483.70,126.01) --
	(483.70,131.01) --
	(478.71,131.01) --
	cycle;

\path[fill=fillColor] (487.02,124.40) --
	(492.02,124.40) --
	(492.02,129.39) --
	(487.02,129.39) --
	cycle;

\path[fill=fillColor] (495.34,122.77) --
	(500.33,122.77) --
	(500.33,127.77) --
	(495.34,127.77) --
	cycle;

\path[fill=fillColor] (503.65,121.27) --
	(508.64,121.27) --
	(508.64,126.27) --
	(503.65,126.27) --
	cycle;

\path[fill=fillColor] (511.96,119.69) --
	(516.96,119.69) --
	(516.96,124.69) --
	(511.96,124.69) --
	cycle;

\path[fill=fillColor] (520.28,118.37) --
	(525.27,118.37) --
	(525.27,123.37) --
	(520.28,123.37) --
	cycle;

\path[fill=fillColor] (528.59,117.08) --
	(533.59,117.08) --
	(533.59,122.07) --
	(528.59,122.07) --
	cycle;

\path[fill=fillColor] (536.91,115.80) --
	(541.90,115.80) --
	(541.90,120.79) --
	(536.91,120.79) --
	cycle;

\path[fill=fillColor] (545.22,114.56) --
	(550.22,114.56) --
	(550.22,119.56) --
	(545.22,119.56) --
	cycle;

\path[fill=fillColor] (553.53,113.11) --
	(558.53,113.11) --
	(558.53,118.11) --
	(553.53,118.11) --
	cycle;

\path[fill=fillColor] (561.85,112.00) --
	(566.84,112.00) --
	(566.84,116.99) --
	(561.85,116.99) --
	cycle;

\path[fill=fillColor] (570.16,110.83) --
	(575.16,110.83) --
	(575.16,115.83) --
	(570.16,115.83) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.27,192.97) -- (368.34,192.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.81,189.44) -- (364.81,196.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.59,187.54) -- (376.65,187.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (373.12,184.00) -- (373.12,191.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.90,182.28) -- (384.97,182.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.44,178.74) -- (381.44,185.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.22,177.28) -- (393.28,177.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.75,173.74) -- (389.75,180.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (394.53,172.56) -- (401.60,172.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.06,169.03) -- (398.06,176.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (402.85,168.23) -- (409.91,168.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.38,164.70) -- (406.38,171.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (411.16,165.07) -- (418.22,165.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (414.69,161.54) -- (414.69,168.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.47,161.58) -- (426.54,161.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.01,158.05) -- (423.01,165.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.79,158.33) -- (434.85,158.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.32,154.80) -- (431.32,161.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.10,155.26) -- (443.17,155.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (439.63,151.73) -- (439.63,158.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.42,152.71) -- (451.48,152.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (447.95,149.18) -- (447.95,156.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.73,150.06) -- (459.79,150.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.26,146.53) -- (456.26,153.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.04,147.16) -- (468.11,147.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.58,143.63) -- (464.58,150.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.36,144.40) -- (476.42,144.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (472.89,140.87) -- (472.89,147.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.67,141.94) -- (484.74,141.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.20,138.40) -- (481.20,145.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.99,139.68) -- (493.05,139.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (489.52,136.15) -- (489.52,143.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.30,137.43) -- (501.36,137.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (497.83,133.89) -- (497.83,140.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.61,135.23) -- (509.68,135.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (506.15,131.70) -- (506.15,138.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.93,133.11) -- (517.99,133.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (514.46,129.58) -- (514.46,136.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.24,131.25) -- (526.31,131.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.78,127.72) -- (522.78,134.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (527.56,129.44) -- (534.62,129.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (531.09,125.91) -- (531.09,132.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.87,127.77) -- (542.94,127.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (539.40,124.24) -- (539.40,131.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.19,126.17) -- (551.25,126.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.72,122.64) -- (547.72,129.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.50,124.08) -- (559.56,124.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (556.03,120.55) -- (556.03,127.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.81,122.46) -- (567.88,122.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.35,118.92) -- (564.35,125.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13,121.01) -- (576.19,121.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66,117.48) -- (572.66,124.54);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (323.24, 68.06) circle (  2.50);

\path[fill=fillColor] (331.55, 68.24) circle (  2.50);

\path[fill=fillColor] (339.86, 68.21) circle (  2.50);

\path[fill=fillColor] (348.18, 67.81) circle (  2.50);

\path[fill=fillColor] (356.49, 68.07) circle (  2.50);

\path[fill=fillColor] (364.81, 68.35) circle (  2.50);

\path[fill=fillColor] (373.12, 68.10) circle (  2.50);

\path[fill=fillColor] (381.44, 68.30) circle (  2.50);

\path[fill=fillColor] (389.75, 68.27) circle (  2.50);

\path[fill=fillColor] (398.06, 68.02) circle (  2.50);

\path[fill=fillColor] (406.38, 68.54) circle (  2.50);

\path[fill=fillColor] (414.69, 67.86) circle (  2.50);

\path[fill=fillColor] (423.01, 67.64) circle (  2.50);

\path[fill=fillColor] (431.32, 68.15) circle (  2.50);

\path[fill=fillColor] (439.63, 68.39) circle (  2.50);

\path[fill=fillColor] (447.95, 68.19) circle (  2.50);

\path[fill=fillColor] (456.26, 68.00) circle (  2.50);

\path[fill=fillColor] (464.58, 68.03) circle (  2.50);

\path[fill=fillColor] (472.89, 67.80) circle (  2.50);

\path[fill=fillColor] (481.20, 67.94) circle (  2.50);

\path[fill=fillColor] (489.52, 67.80) circle (  2.50);

\path[fill=fillColor] (497.83, 68.37) circle (  2.50);

\path[fill=fillColor] (506.15, 68.09) circle (  2.50);

\path[fill=fillColor] (514.46, 68.40) circle (  2.50);

\path[fill=fillColor] (522.78, 68.75) circle (  2.50);

\path[fill=fillColor] (531.09, 67.59) circle (  2.50);

\path[fill=fillColor] (539.40, 68.00) circle (  2.50);

\path[fill=fillColor] (547.72, 68.44) circle (  2.50);

\path[fill=fillColor] (556.03, 68.43) circle (  2.50);

\path[fill=fillColor] (564.35, 67.96) circle (  2.50);

\path[fill=fillColor] (572.66, 67.78) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.24, 30.69) rectangle (572.66,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29, 27.66) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29, 78.74) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,129.82) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,180.91) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49, 30.69) --
	(323.24, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49, 81.77) --
	(323.24, 81.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,132.85) --
	(323.24,132.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,183.94) --
	(323.24,183.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.24, 27.94) --
	(323.24, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.38, 27.94) --
	(406.38, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (489.52, 27.94) --
	(489.52, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (323.24, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (406.38, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.52, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (447.95,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,112.42) {Squared coefficient of variation};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (323.24,202.22) {D};
\end{scope}
\end{tikzpicture}

\end{document}