### 2020-04-20 serial_analysis.tex
stuff

% Created by tikzDevice version 0.12.3 on 2020-04-20 00:29:27
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (650.43,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (216.81,216.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 34.64, 30.69) rectangle (211.31,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (100.89,194.15) --
	(106.42,189.20) --
	(111.94,184.25) --
	(117.46,179.29) --
	(122.98,174.34) --
	(128.50,169.39) --
	(134.02,164.43) --
	(139.54,159.48) --
	(145.06,154.52) --
	(150.58,149.57) --
	(156.10,144.62) --
	(161.62,139.66) --
	(167.14,134.71) --
	(172.66,129.76) --
	(178.19,124.80) --
	(183.71,119.85) --
	(189.23,114.90) --
	(194.75,109.94) --
	(200.27,104.99) --
	(205.79,100.04) --
	(211.31, 95.08);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 34.64,104.99) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 45.69,161.34) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 51.21,136.36) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 56.73,144.62) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 62.25,144.62) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 67.77,164.43) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.29,164.43) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 78.81,144.62) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 84.33,159.48) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 89.85,137.19) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 95.37,120.06) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (100.89,120.16) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (106.42,124.02) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (111.94,113.45) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (117.46,121.02) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (122.98,112.94) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (128.50,109.78) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (134.02,113.37) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (139.54,101.81) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (145.06,105.30) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (150.58, 99.53) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (156.10, 97.67) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (161.62, 97.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (167.14, 92.46) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (172.66, 89.55) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (178.19, 95.98) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.71, 94.09) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (189.23, 85.17) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (200.27, 70.31) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (211.31, 60.41) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.64,104.99) --
	( 34.64,104.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.64,104.99) --
	( 34.64,104.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.64,104.99) --
	( 34.64,104.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 45.69,139.66) --
	( 45.69,139.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 45.69,139.66) --
	( 45.69,174.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 45.69,174.34) --
	( 45.69,174.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 51.21,114.90) --
	( 51.21,114.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 51.21,114.90) --
	( 51.21,169.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 51.21,169.39) --
	( 51.21,169.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 56.73,134.71) --
	( 56.73,134.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 56.73,134.71) --
	( 56.73,154.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 56.73,154.52) --
	( 56.73,154.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.25,104.99) --
	( 62.25,104.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.25,104.99) --
	( 62.25,174.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.25,174.34) --
	( 62.25,174.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.77,164.43) --
	( 67.77,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.77,164.43) --
	( 67.77,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.77,164.43) --
	( 67.77,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.29,139.66) --
	( 73.29,139.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.29,139.66) --
	( 73.29,189.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.29,189.20) --
	( 73.29,189.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.81,134.71) --
	( 78.81,134.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.81,134.71) --
	( 78.81,154.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.81,154.52) --
	( 78.81,154.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.33,159.48) --
	( 84.33,159.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.33,159.48) --
	( 84.33,159.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.33,159.48) --
	( 84.33,159.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.85,119.85) --
	( 89.85,119.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.85,119.85) --
	( 89.85,159.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.85,159.48) --
	( 89.85,159.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.37, 80.22) --
	( 95.37, 80.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.37, 80.22) --
	( 95.37,174.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.37,174.34) --
	( 95.37,174.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.89, 95.08) --
	(100.89, 95.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.89, 95.08) --
	(100.89,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.89,164.43) --
	(100.89,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.42, 90.13) --
	(106.42, 90.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.42, 90.13) --
	(106.42,149.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.42,149.57) --
	(106.42,149.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.94, 85.17) --
	(111.94, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.94, 85.17) --
	(111.94,159.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.94,159.48) --
	(111.94,159.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (117.46, 90.13) --
	(117.46, 90.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (117.46, 90.13) --
	(117.46,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (117.46,164.43) --
	(117.46,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (122.98, 45.55) --
	(122.98, 45.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (122.98, 45.55) --
	(122.98,169.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (122.98,169.39) --
	(122.98,169.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (128.50, 85.17) --
	(128.50, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (128.50, 85.17) --
	(128.50,149.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (128.50,149.57) --
	(128.50,149.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.02, 90.13) --
	(134.02, 90.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.02, 90.13) --
	(134.02,144.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.02,144.62) --
	(134.02,144.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.54, 65.36) --
	(139.54, 65.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.54, 65.36) --
	(139.54,134.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.54,134.71) --
	(139.54,134.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.06, 60.41) --
	(145.06, 60.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.06, 60.41) --
	(145.06,144.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.06,144.62) --
	(145.06,144.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.58, 65.36) --
	(150.58, 65.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.58, 65.36) --
	(150.58,149.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.58,149.57) --
	(150.58,149.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.10, 40.59) --
	(156.10, 40.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.10, 40.59) --
	(156.10,134.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.10,134.71) --
	(156.10,134.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.62, 50.50) --
	(161.62, 50.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.62, 50.50) --
	(161.62,129.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.62,129.76) --
	(161.62,129.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.14, 60.41) --
	(167.14, 60.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.14, 60.41) --
	(167.14,109.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.14,109.94) --
	(167.14,109.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.66, 65.36) --
	(172.66, 65.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.66, 65.36) --
	(172.66,109.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.66,109.94) --
	(172.66,109.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.19, 75.27) --
	(178.19, 75.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.19, 75.27) --
	(178.19,109.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.19,109.94) --
	(178.19,109.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.71, 85.17) --
	(183.71, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.71, 85.17) --
	(183.71,100.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.71,100.04) --
	(183.71,100.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (189.23, 85.17) --
	(189.23, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (189.23, 85.17) --
	(189.23, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (189.23, 85.17) --
	(189.23, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.27, 70.31) --
	(200.27, 70.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.27, 70.31) --
	(200.27, 70.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.27, 70.31) --
	(200.27, 70.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.31, 35.64) --
	(211.31, 35.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.31, 35.64) --
	(211.31, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.31, 85.17) --
	(211.31, 85.17);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor,fill opacity=0.20] ( 34.64,171.01) --
	( 36.88,169.05) --
	( 39.12,167.14) --
	( 41.35,165.27) --
	( 43.59,163.46) --
	( 45.83,161.68) --
	( 48.06,159.96) --
	( 50.30,158.29) --
	( 52.54,156.67) --
	( 54.77,155.10) --
	( 57.01,153.59) --
	( 59.24,152.12) --
	( 61.48,150.70) --
	( 63.72,149.32) --
	( 65.95,147.97) --
	( 68.19,146.66) --
	( 70.43,145.37) --
	( 72.66,144.09) --
	( 74.90,142.81) --
	( 77.13,141.54) --
	( 79.37,140.25) --
	( 81.61,138.94) --
	( 83.84,137.61) --
	( 86.08,136.26) --
	( 88.32,134.88) --
	( 90.55,133.47) --
	( 92.79,132.04) --
	( 95.02,130.59) --
	( 97.26,129.14) --
	( 99.50,127.74) --
	(101.73,126.42) --
	(103.97,125.19) --
	(106.21,124.07) --
	(108.44,123.07) --
	(110.68,122.19) --
	(112.91,121.41) --
	(115.15,120.75) --
	(117.39,120.02) --
	(119.62,119.00) --
	(121.86,117.91) --
	(124.10,116.89) --
	(126.33,115.70) --
	(128.57,114.48) --
	(130.80,113.34) --
	(133.04,112.27) --
	(135.28,111.11) --
	(137.51,109.84) --
	(139.75,108.69) --
	(141.99,107.59) --
	(144.22,106.58) --
	(146.46,105.54) --
	(148.69,104.57) --
	(150.93,103.63) --
	(153.17,102.62) --
	(155.40,101.65) --
	(157.64,100.75) --
	(159.88, 99.93) --
	(162.11, 99.20) --
	(164.35, 98.48) --
	(166.58, 97.75) --
	(168.82, 97.00) --
	(171.06, 96.23) --
	(173.29, 95.46) --
	(175.53, 94.69) --
	(177.77, 93.95) --
	(180.00, 93.23) --
	(182.24, 92.56) --
	(184.47, 91.94) --
	(186.71, 91.37) --
	(188.95, 90.85) --
	(191.18, 90.39) --
	(193.42, 89.99) --
	(195.66, 89.64) --
	(197.89, 89.34) --
	(200.13, 89.10) --
	(202.36, 88.91) --
	(204.60, 88.77) --
	(206.84, 88.68) --
	(209.07, 88.64) --
	(211.31, 88.65) --
	(211.31, 46.93) --
	(209.07, 50.41) --
	(206.84, 53.75) --
	(204.60, 56.97) --
	(202.36, 60.04) --
	(200.13, 62.98) --
	(197.89, 65.79) --
	(195.66, 68.45) --
	(193.42, 70.97) --
	(191.18, 73.35) --
	(188.95, 75.58) --
	(186.71, 77.67) --
	(184.47, 79.62) --
	(182.24, 81.42) --
	(180.00, 83.08) --
	(177.77, 84.60) --
	(175.53, 86.01) --
	(173.29, 87.29) --
	(171.06, 88.48) --
	(168.82, 89.58) --
	(166.58, 90.59) --
	(164.35, 91.53) --
	(162.11, 92.39) --
	(159.88, 93.23) --
	(157.64, 94.16) --
	(155.40, 95.15) --
	(153.17, 96.16) --
	(150.93, 97.18) --
	(148.69, 98.28) --
	(146.46, 99.36) --
	(144.22,100.40) --
	(141.99,101.43) --
	(139.75,102.42) --
	(137.51,103.58) --
	(135.28,104.76) --
	(133.04,105.87) --
	(130.80,107.14) --
	(128.57,108.40) --
	(126.33,109.53) --
	(124.10,110.53) --
	(121.86,111.53) --
	(119.62,112.53) --
	(117.39,113.40) --
	(115.15,114.13) --
	(112.91,114.80) --
	(110.68,115.58) --
	(108.44,116.44) --
	(106.21,117.32) --
	(103.97,118.24) --
	(101.73,119.19) --
	( 99.50,120.18) --
	( 97.26,121.23) --
	( 95.02,122.36) --
	( 92.79,123.51) --
	( 90.55,124.65) --
	( 88.32,125.78) --
	( 86.08,126.91) --
	( 83.84,128.03) --
	( 81.61,129.15) --
	( 79.37,130.26) --
	( 77.13,131.35) --
	( 74.90,132.43) --
	( 72.66,133.48) --
	( 70.43,134.49) --
	( 68.19,135.47) --
	( 65.95,136.39) --
	( 63.72,137.26) --
	( 61.48,138.06) --
	( 59.24,138.79) --
	( 57.01,139.45) --
	( 54.77,140.04) --
	( 52.54,140.55) --
	( 50.30,140.98) --
	( 48.06,141.33) --
	( 45.83,141.61) --
	( 43.59,141.82) --
	( 41.35,141.95) --
	( 39.12,142.01) --
	( 36.88,142.01) --
	( 34.64,141.94) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 34.64,156.47) --
	( 36.88,155.53) --
	( 39.12,154.58) --
	( 41.35,153.61) --
	( 43.59,152.64) --
	( 45.83,151.65) --
	( 48.06,150.65) --
	( 50.30,149.64) --
	( 52.54,148.61) --
	( 54.77,147.57) --
	( 57.01,146.52) --
	( 59.24,145.46) --
	( 61.48,144.38) --
	( 63.72,143.29) --
	( 65.95,142.18) --
	( 68.19,141.06) --
	( 70.43,139.93) --
	( 72.66,138.78) --
	( 74.90,137.62) --
	( 77.13,136.44) --
	( 79.37,135.25) --
	( 81.61,134.04) --
	( 83.84,132.82) --
	( 86.08,131.59) --
	( 88.32,130.33) --
	( 90.55,129.06) --
	( 92.79,127.78) --
	( 95.02,126.48) --
	( 97.26,125.19) --
	( 99.50,123.96) --
	(101.73,122.80) --
	(103.97,121.71) --
	(106.21,120.70) --
	(108.44,119.75) --
	(110.68,118.89) --
	(112.91,118.11) --
	(115.15,117.44) --
	(117.39,116.71) --
	(119.62,115.77) --
	(121.86,114.72) --
	(124.10,113.71) --
	(126.33,112.61) --
	(128.57,111.44) --
	(130.80,110.24) --
	(133.04,109.07) --
	(135.28,107.93) --
	(137.51,106.71) --
	(139.75,105.56) --
	(141.99,104.51) --
	(144.22,103.49) --
	(146.46,102.45) --
	(148.69,101.42) --
	(150.93,100.41) --
	(153.17, 99.39) --
	(155.40, 98.40) --
	(157.64, 97.46) --
	(159.88, 96.58) --
	(162.11, 95.79) --
	(164.35, 95.01) --
	(166.58, 94.17) --
	(168.82, 93.29) --
	(171.06, 92.35) --
	(173.29, 91.38) --
	(175.53, 90.35) --
	(177.77, 89.28) --
	(180.00, 88.16) --
	(182.24, 86.99) --
	(184.47, 85.78) --
	(186.71, 84.52) --
	(188.95, 83.22) --
	(191.18, 81.87) --
	(193.42, 80.48) --
	(195.66, 79.04) --
	(197.89, 77.57) --
	(200.13, 76.04) --
	(202.36, 74.48) --
	(204.60, 72.87) --
	(206.84, 71.22) --
	(209.07, 69.53) --
	(211.31, 67.79);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.64, 30.69) rectangle (211.31,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69, 37.56) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69, 87.10) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69,136.63) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69,186.17) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89, 40.59) --
	( 34.64, 40.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89, 90.13) --
	( 34.64, 90.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89,139.66) --
	( 34.64,139.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89,189.20) --
	( 34.64,189.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 51.21, 27.94) --
	( 51.21, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.42, 27.94) --
	(106.42, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.62, 27.94) --
	(161.62, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 51.21, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (106.42, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (161.62, 19.68) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (122.98,  7.64) {Primary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,112.42) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 34.64,202.22) {A. Forward serial interval};
\end{scope}
\begin{scope}
\path[clip] (216.81,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (216.81,  0.00) rectangle (433.62,216.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.45, 30.69) rectangle (428.12,194.15);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (251.45,105.23) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (275.01, 47.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (280.90,110.02) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (286.79, 81.26) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (292.68,124.40) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (298.57, 90.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (304.45, 83.31) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (310.34, 98.52) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (316.23, 87.25) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (322.12, 95.32) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (328.01,102.53) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (333.90,105.54) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (339.79,106.72) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (345.68,105.89) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (351.57,100.32) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (357.45,107.92) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (363.34,112.15) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (369.23,113.22) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (375.12,111.54) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (381.01,112.98) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (386.90,115.62) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (392.79,121.14) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (398.68,124.15) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (404.56,122.56) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (410.45,139.38) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (416.34,133.19) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (422.23,167.55) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (428.12,148.37) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.45,105.23) --
	(251.45,105.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.45,105.23) --
	(251.45,105.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.45,105.23) --
	(251.45,105.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.01, 47.70) --
	(275.01, 47.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.01, 47.70) --
	(275.01, 47.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.01, 47.70) --
	(275.01, 47.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.90,105.23) --
	(280.90,105.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.90,105.23) --
	(280.90,114.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.90,114.82) --
	(280.90,114.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (286.79, 81.26) --
	(286.79, 81.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (286.79, 81.26) --
	(286.79, 81.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (286.79, 81.26) --
	(286.79, 81.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.68,124.40) --
	(292.68,124.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.68,124.40) --
	(292.68,124.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.68,124.40) --
	(292.68,124.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (298.57, 90.85) --
	(298.57, 90.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (298.57, 90.85) --
	(298.57, 90.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (298.57, 90.85) --
	(298.57, 90.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (304.45, 42.91) --
	(304.45, 42.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (304.45, 42.91) --
	(304.45,138.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (304.45,138.78) --
	(304.45,138.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (310.34, 86.05) --
	(310.34, 86.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (310.34, 86.05) --
	(310.34,133.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (310.34,133.99) --
	(310.34,133.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.23, 62.08) --
	(316.23, 62.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.23, 62.08) --
	(316.23,105.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.23,105.23) --
	(316.23,105.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (322.12, 52.50) --
	(322.12, 52.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (322.12, 52.50) --
	(322.12,153.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (322.12,153.17) --
	(322.12,153.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (328.01, 66.88) --
	(328.01, 66.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (328.01, 66.88) --
	(328.01,157.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (328.01,157.96) --
	(328.01,157.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90, 71.67) --
	(333.90, 71.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90, 71.67) --
	(333.90,162.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90,162.75) --
	(333.90,162.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.79, 62.08) --
	(339.79, 62.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.79, 62.08) --
	(339.79,167.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.79,167.55) --
	(339.79,167.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.68, 71.67) --
	(345.68, 71.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.68, 71.67) --
	(345.68,172.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.68,172.34) --
	(345.68,172.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.57, 66.88) --
	(351.57, 66.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.57, 66.88) --
	(351.57,133.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.57,133.99) --
	(351.57,133.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (357.45, 38.12) --
	(357.45, 38.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (357.45, 38.12) --
	(357.45,162.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (357.45,162.75) --
	(357.45,162.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (363.34, 90.85) --
	(363.34, 90.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (363.34, 90.85) --
	(363.34,172.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (363.34,172.34) --
	(363.34,172.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.23, 76.47) --
	(369.23, 76.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.23, 76.47) --
	(369.23,157.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.23,157.96) --
	(369.23,157.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.12, 86.05) --
	(375.12, 86.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.12, 86.05) --
	(375.12,157.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.12,157.96) --
	(375.12,157.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (381.01, 86.05) --
	(381.01, 86.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (381.01, 86.05) --
	(381.01,148.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (381.01,148.37) --
	(381.01,148.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (386.90, 71.67) --
	(386.90, 71.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (386.90, 71.67) --
	(386.90,162.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (386.90,162.75) --
	(386.90,162.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (392.79, 86.05) --
	(392.79, 86.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (392.79, 86.05) --
	(392.79,186.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (392.79,186.72) --
	(392.79,186.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.68, 95.64) --
	(398.68, 95.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.68, 95.64) --
	(398.68,172.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.68,172.34) --
	(398.68,172.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.56,100.43) --
	(404.56,100.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.56,100.43) --
	(404.56,148.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.56,148.37) --
	(404.56,148.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.45,110.02) --
	(410.45,110.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.45,110.02) --
	(410.45,162.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.45,162.75) --
	(410.45,162.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.34, 86.05) --
	(416.34, 86.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.34, 86.05) --
	(416.34,162.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.34,162.75) --
	(416.34,162.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (422.23,167.55) --
	(422.23,167.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (422.23,167.55) --
	(422.23,167.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (422.23,167.55) --
	(422.23,167.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (428.12,148.37) --
	(428.12,148.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (428.12,148.37) --
	(428.12,148.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (428.12,148.37) --
	(428.12,148.37);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor,fill opacity=0.20] (251.45,121.15) --
	(253.69,119.25) --
	(255.93,117.44) --
	(258.16,115.70) --
	(260.40,114.05) --
	(262.64,112.48) --
	(264.87,111.00) --
	(267.11,109.60) --
	(269.35,108.28) --
	(271.58,107.06) --
	(273.82,105.92) --
	(276.05,104.87) --
	(278.29,103.92) --
	(280.53,103.05) --
	(282.76,102.27) --
	(285.00,101.59) --
	(287.24,101.00) --
	(289.47,100.50) --
	(291.71,100.09) --
	(293.94, 99.78) --
	(296.18, 99.55) --
	(298.42, 99.42) --
	(300.65, 99.37) --
	(302.89, 99.40) --
	(305.13, 99.50) --
	(307.36, 99.68) --
	(309.60, 99.93) --
	(311.83,100.23) --
	(314.07,100.59) --
	(316.31,100.99) --
	(318.54,101.45) --
	(320.78,101.95) --
	(323.02,102.50) --
	(325.25,103.10) --
	(327.49,103.76) --
	(329.72,104.45) --
	(331.96,105.11) --
	(334.20,105.76) --
	(336.43,106.39) --
	(338.67,107.01) --
	(340.91,107.56) --
	(343.14,107.90) --
	(345.38,108.14) --
	(347.61,108.42) --
	(349.85,108.79) --
	(352.09,109.25) --
	(354.32,109.74) --
	(356.56,110.39) --
	(358.80,111.15) --
	(361.03,112.32) --
	(363.27,113.45) --
	(365.50,113.91) --
	(367.74,114.29) --
	(369.98,114.93) --
	(372.21,115.63) --
	(374.45,116.42) --
	(376.69,117.13) --
	(378.92,117.75) --
	(381.16,118.39) --
	(383.39,119.13) --
	(385.63,120.01) --
	(387.87,121.05) --
	(390.10,122.20) --
	(392.34,123.45) --
	(394.58,124.82) --
	(396.81,126.34) --
	(399.05,128.02) --
	(401.28,129.88) --
	(403.52,131.95) --
	(405.76,134.21) --
	(407.99,136.68) --
	(410.23,139.36) --
	(412.47,142.24) --
	(414.70,145.32) --
	(416.94,148.60) --
	(419.17,152.06) --
	(421.41,155.72) --
	(423.65,159.56) --
	(425.88,163.59) --
	(428.12,167.81) --
	(428.12,134.79) --
	(425.88,133.76) --
	(423.65,132.73) --
	(421.41,131.69) --
	(419.17,130.64) --
	(416.94,129.57) --
	(414.70,128.50) --
	(412.47,127.40) --
	(410.23,126.29) --
	(407.99,125.14) --
	(405.76,123.97) --
	(403.52,122.77) --
	(401.28,121.53) --
	(399.05,120.28) --
	(396.81,119.01) --
	(394.58,117.75) --
	(392.34,116.53) --
	(390.10,115.35) --
	(387.87,114.24) --
	(385.63,113.23) --
	(383.39,112.40) --
	(381.16,111.67) --
	(378.92,110.96) --
	(376.69,110.24) --
	(374.45,109.51) --
	(372.21,108.90) --
	(369.98,108.23) --
	(367.74,107.58) --
	(365.50,106.98) --
	(363.27,106.24) --
	(361.03,105.28) --
	(358.80,104.09) --
	(356.56,103.31) --
	(354.32,102.90) --
	(352.09,102.47) --
	(349.85,102.09) --
	(347.61,101.83) --
	(345.38,101.55) --
	(343.14,101.17) --
	(340.91,100.66) --
	(338.67,100.06) --
	(336.43, 99.41) --
	(334.20, 98.70) --
	(331.96, 97.92) --
	(329.72, 97.09) --
	(327.49, 96.21) --
	(325.25, 95.32) --
	(323.02, 94.44) --
	(320.78, 93.58) --
	(318.54, 92.75) --
	(316.31, 91.93) --
	(314.07, 91.14) --
	(311.83, 90.37) --
	(309.60, 89.60) --
	(307.36, 88.85) --
	(305.13, 88.09) --
	(302.89, 87.33) --
	(300.65, 86.56) --
	(298.42, 85.77) --
	(296.18, 84.96) --
	(293.94, 84.13) --
	(291.71, 83.27) --
	(289.47, 82.38) --
	(287.24, 81.46) --
	(285.00, 80.51) --
	(282.76, 79.53) --
	(280.53, 78.51) --
	(278.29, 77.47) --
	(276.05, 76.39) --
	(273.82, 75.29) --
	(271.58, 74.15) --
	(269.35, 72.98) --
	(267.11, 71.79) --
	(264.87, 70.56) --
	(262.64, 69.31) --
	(260.40, 68.03) --
	(258.16, 66.72) --
	(255.93, 65.38) --
	(253.69, 64.01) --
	(251.45, 62.62) --
	cycle;

\path[draw=drawColor,line width= 1.1pt,line join=round] (251.45, 91.89) --
	(253.69, 91.63) --
	(255.93, 91.41) --
	(258.16, 91.21) --
	(260.40, 91.04) --
	(262.64, 90.89) --
	(264.87, 90.78) --
	(267.11, 90.69) --
	(269.35, 90.63) --
	(271.58, 90.60) --
	(273.82, 90.60) --
	(276.05, 90.63) --
	(278.29, 90.69) --
	(280.53, 90.78) --
	(282.76, 90.90) --
	(285.00, 91.05) --
	(287.24, 91.23) --
	(289.47, 91.44) --
	(291.71, 91.68) --
	(293.94, 91.95) --
	(296.18, 92.26) --
	(298.42, 92.60) --
	(300.65, 92.96) --
	(302.89, 93.37) --
	(305.13, 93.80) --
	(307.36, 94.27) --
	(309.60, 94.76) --
	(311.83, 95.30) --
	(314.07, 95.86) --
	(316.31, 96.46) --
	(318.54, 97.10) --
	(320.78, 97.77) --
	(323.02, 98.47) --
	(325.25, 99.21) --
	(327.49, 99.98) --
	(329.72,100.77) --
	(331.96,101.52) --
	(334.20,102.23) --
	(336.43,102.90) --
	(338.67,103.53) --
	(340.91,104.11) --
	(343.14,104.53) --
	(345.38,104.85) --
	(347.61,105.13) --
	(349.85,105.44) --
	(352.09,105.86) --
	(354.32,106.32) --
	(356.56,106.85) --
	(358.80,107.62) --
	(361.03,108.80) --
	(363.27,109.84) --
	(365.50,110.44) --
	(367.74,110.94) --
	(369.98,111.58) --
	(372.21,112.27) --
	(374.45,112.97) --
	(376.69,113.68) --
	(378.92,114.35) --
	(381.16,115.03) --
	(383.39,115.76) --
	(385.63,116.62) --
	(387.87,117.64) --
	(390.10,118.77) --
	(392.34,119.99) --
	(394.58,121.29) --
	(396.81,122.67) --
	(399.05,124.15) --
	(401.28,125.71) --
	(403.52,127.36) --
	(405.76,129.09) --
	(407.99,130.91) --
	(410.23,132.82) --
	(412.47,134.82) --
	(414.70,136.91) --
	(416.94,139.09) --
	(419.17,141.35) --
	(421.41,143.70) --
	(423.65,146.14) --
	(425.88,148.68) --
	(428.12,151.30);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.45, 30.69) rectangle (428.12,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (246.50, 39.88) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (246.50, 87.82) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (246.50,135.75) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (246.50,183.69) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (248.70, 42.91) --
	(251.45, 42.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (248.70, 90.85) --
	(251.45, 90.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (248.70,138.78) --
	(251.45,138.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (248.70,186.72) --
	(251.45,186.72);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.45, 27.94) --
	(251.45, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (310.34, 27.94) --
	(310.34, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.23, 27.94) --
	(369.23, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (428.12, 27.94) --
	(428.12, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (251.45, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (310.34, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (369.23, 19.68) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (428.12, 19.68) {30};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (339.79,  7.64) {Secondary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (229.89,112.42) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (251.45,202.22) {B. Backward serial interval};
\end{scope}
\begin{scope}
\path[clip] (433.62,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (433.62,  0.00) rectangle (650.43,216.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (460.93, 30.69) rectangle (644.93,194.15);
\definecolor{drawColor}{RGB}{240,228,66}
\definecolor{fillColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (478.18,186.72) --
	(483.93,172.02) --
	(489.68,167.41) --
	(495.43,164.97) --
	(501.18,170.94) --
	(506.93,165.55) --
	(512.68,162.02) --
	(518.43,156.83) --
	(524.18,104.98) --
	(529.93, 92.34) --
	(535.68, 89.99) --
	(541.43, 81.76) --
	(547.18, 79.80) --
	(552.93, 73.13) --
	(558.68, 70.34) --
	(564.43, 68.88) --
	(570.18, 64.62) --
	(575.93, 62.47) --
	(581.68, 59.85) --
	(587.43, 58.86) --
	(593.18, 57.46) --
	(598.93, 56.56) --
	(604.68, 55.29) --
	(610.43, 55.07) --
	(616.18, 54.52) --
	(621.93, 54.37) --
	(627.68, 54.37) --
	(633.43, 54.35) --
	(639.18, 54.35) --
	(644.93, 53.84) --
	(644.93, 48.78) --
	(639.18, 49.43) --
	(633.43, 49.43) --
	(627.68, 49.75) --
	(621.93, 49.75) --
	(616.18, 49.90) --
	(610.43, 49.74) --
	(604.68, 50.17) --
	(598.93, 51.03) --
	(593.18, 51.66) --
	(587.43, 52.72) --
	(581.68, 54.51) --
	(575.93, 55.72) --
	(570.18, 58.33) --
	(564.43, 60.67) --
	(558.68, 61.38) --
	(552.93, 63.32) --
	(547.18, 68.18) --
	(541.43, 67.77) --
	(535.68, 72.19) --
	(529.93, 71.54) --
	(524.18, 74.09) --
	(518.43, 99.35) --
	(512.68,104.50) --
	(506.93,103.95) --
	(501.18,102.98) --
	(495.43,103.05) --
	(489.68, 92.92) --
	(483.93, 95.55) --
	(478.18, 92.27) --
	cycle;
\definecolor{drawColor}{RGB}{0,114,178}
\definecolor{fillColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (478.18,125.90) --
	(483.93,119.23) --
	(489.68,117.31) --
	(495.43,115.79) --
	(501.18,119.31) --
	(506.93,116.22) --
	(512.68,114.76) --
	(518.43,112.24) --
	(524.18, 84.09) --
	(529.93, 76.40) --
	(535.68, 74.61) --
	(541.43, 69.22) --
	(547.18, 68.02) --
	(552.93, 63.34) --
	(558.68, 61.72) --
	(564.43, 60.71) --
	(570.18, 57.76) --
	(575.93, 56.38) --
	(581.68, 54.72) --
	(587.43, 54.07) --
	(593.18, 52.92) --
	(598.93, 52.36) --
	(604.68, 51.40) --
	(610.43, 51.36) --
	(616.18, 51.02) --
	(621.93, 50.93) --
	(627.68, 50.93) --
	(633.43, 50.85) --
	(639.18, 50.85) --
	(644.93, 50.60) --
	(644.93, 47.19) --
	(639.18, 47.47) --
	(633.43, 47.47) --
	(627.68, 47.71) --
	(621.93, 47.71) --
	(616.18, 47.77) --
	(610.43, 47.85) --
	(604.68, 48.08) --
	(598.93, 48.58) --
	(593.18, 49.16) --
	(587.43, 49.86) --
	(581.68, 50.91) --
	(575.93, 51.95) --
	(570.18, 53.53) --
	(564.43, 55.54) --
	(558.68, 55.84) --
	(552.93, 57.08) --
	(547.18, 60.14) --
	(541.43, 60.06) --
	(535.68, 63.08) --
	(529.93, 62.94) --
	(524.18, 64.75) --
	(518.43, 79.69) --
	(512.68, 84.36) --
	(506.93, 83.64) --
	(501.18, 82.64) --
	(495.43, 82.42) --
	(489.68, 76.41) --
	(483.93, 77.92) --
	(478.18, 75.82) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (460.93, 38.12) -- (644.93, 38.12);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.18,127.94) --
	(483.93,127.84) --
	(489.68,122.64) --
	(495.43,125.45) --
	(501.18,129.15) --
	(506.93,128.97) --
	(512.68,130.66) --
	(518.43,125.72) --
	(524.18, 87.52) --
	(529.93, 82.25) --
	(535.68, 80.72) --
	(541.43, 74.16) --
	(547.18, 73.09) --
	(552.93, 67.65) --
	(558.68, 65.54) --
	(564.43, 64.88) --
	(570.18, 61.54) --
	(575.93, 59.17) --
	(581.68, 56.95) --
	(587.43, 55.38) --
	(593.18, 54.56) --
	(598.93, 53.59) --
	(604.68, 52.65) --
	(610.43, 52.34) --
	(616.18, 52.19) --
	(621.93, 52.07) --
	(627.68, 52.07) --
	(633.43, 51.94) --
	(639.18, 51.94) --
	(644.93, 51.46);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (478.18, 97.05) --
	(483.93, 96.63) --
	(489.68, 94.07) --
	(495.43, 95.65) --
	(501.18, 97.81) --
	(506.93, 97.47) --
	(512.68, 98.36) --
	(518.43, 95.36) --
	(524.18, 73.48) --
	(529.93, 69.78) --
	(535.68, 68.56) --
	(541.43, 64.35) --
	(547.18, 63.49) --
	(552.93, 60.05) --
	(558.68, 58.57) --
	(564.43, 58.06) --
	(570.18, 55.78) --
	(575.93, 54.18) --
	(581.68, 52.66) --
	(587.43, 51.66) --
	(593.18, 51.10) --
	(598.93, 50.42) --
	(604.68, 49.74) --
	(610.43, 49.50) --
	(616.18, 49.38) --
	(621.93, 49.29) --
	(627.68, 49.29) --
	(633.43, 49.20) --
	(639.18, 49.20) --
	(644.93, 48.92);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (460.93, 30.69) rectangle (644.93,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.98, 35.09) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.98, 71.15) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.98,107.22) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.98,143.29) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.98,179.35) {5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.18, 38.12) --
	(460.93, 38.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.18, 74.18) --
	(460.93, 74.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.18,110.25) --
	(460.93,110.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.18,146.32) --
	(460.93,146.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.18,182.38) --
	(460.93,182.38);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.18, 27.94) --
	(478.18, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (535.68, 27.94) --
	(535.68, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.18, 27.94) --
	(593.18, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (478.18, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (535.68, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (593.18, 19.68) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (552.93,  7.64) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (446.70,112.42) {Reproduction number $\mathcal R$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (528.28,146.93) rectangle (643.82,192.34);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (533.78,166.88) rectangle (548.24,181.34);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{240,228,66}
\definecolor{fillColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor,fill opacity=0.50] (534.49,167.59) rectangle (547.53,180.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (535.23,174.11) -- (546.79,174.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (533.78,152.43) rectangle (548.24,166.88);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,114,178}
\definecolor{fillColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line cap=rect,fill=fillColor,fill opacity=0.50] (534.49,153.14) rectangle (547.53,166.17);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (535.23,159.66) -- (546.79,159.66);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (553.74,171.08) {6 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (553.74,156.63) {8 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (460.93,202.22) {C. Cohort-averaged estimates};
\end{scope}
\end{tikzpicture}

\end{document}