### 2020-04-29 rR.tex
new fig

% Created by tikzDevice version 0.12.3 on 2020-04-29 03:55:17
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,433.62);
\begin{scope}
\path[clip] (  0.00,216.81) rectangle (289.08,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,216.81) rectangle (289.08,433.62);
\end{scope}
\begin{scope}
\path[clip] ( 27.31,247.50) rectangle (283.58,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31,247.50) rectangle (283.58,410.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 52.94,254.97) --
	( 55.54,255.77) --
	( 58.13,256.58) --
	( 60.73,257.40) --
	( 63.32,258.23) --
	( 65.92,259.08) --
	( 68.51,259.94) --
	( 71.11,260.81) --
	( 73.70,261.69) --
	( 76.30,262.59) --
	( 78.89,263.49) --
	( 81.49,264.42) --
	( 84.08,265.35) --
	( 86.68,266.30) --
	( 89.27,267.26) --
	( 91.87,268.23) --
	( 94.46,269.21) --
	( 97.06,270.20) --
	( 99.65,271.21) --
	(102.25,272.23) --
	(104.84,273.26) --
	(107.44,274.30) --
	(110.03,275.36) --
	(112.63,276.43) --
	(115.22,277.51) --
	(117.82,278.61) --
	(120.41,279.72) --
	(123.01,280.84) --
	(125.60,281.98) --
	(128.20,283.13) --
	(130.79,284.29) --
	(133.39,285.46) --
	(135.98,286.65) --
	(138.58,287.85) --
	(141.17,289.07) --
	(143.77,290.31) --
	(146.36,291.55) --
	(148.96,292.82) --
	(151.55,294.10) --
	(154.15,295.40) --
	(156.74,296.71) --
	(159.34,298.04) --
	(161.93,299.39) --
	(164.53,300.75) --
	(167.12,302.12) --
	(169.72,303.51) --
	(172.31,304.92) --
	(174.91,306.34) --
	(177.51,307.78) --
	(180.10,309.24) --
	(182.70,310.71) --
	(185.29,312.21) --
	(187.89,313.72) --
	(190.48,315.26) --
	(193.08,316.81) --
	(195.67,318.38) --
	(198.27,319.97) --
	(200.86,321.57) --
	(203.46,323.20) --
	(206.05,324.84) --
	(208.65,326.51) --
	(211.24,328.19) --
	(213.84,329.89) --
	(216.43,331.62) --
	(219.03,333.36) --
	(221.62,335.12) --
	(224.22,336.91) --
	(226.81,338.72) --
	(229.41,340.54) --
	(232.00,342.39) --
	(234.60,344.25) --
	(237.19,346.14) --
	(239.79,348.04) --
	(242.38,349.97) --
	(244.98,351.91) --
	(247.57,353.87) --
	(250.17,355.85) --
	(252.76,357.85) --
	(255.36,359.88) --
	(257.95,361.92);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 27.31,247.52) --
	( 30.34,248.34) --
	( 33.37,249.16) --
	( 36.40,250.01) --
	( 39.42,250.87) --
	( 42.45,251.74) --
	( 45.48,252.64) --
	( 48.51,253.54) --
	( 51.53,254.47) --
	( 54.56,255.41) --
	( 57.59,256.36) --
	( 60.62,257.34) --
	( 63.64,258.32) --
	( 66.67,259.33) --
	( 69.70,260.35) --
	( 72.73,261.38) --
	( 75.76,262.43) --
	( 78.78,263.50) --
	( 81.81,264.58) --
	( 84.84,265.68) --
	( 87.87,266.80) --
	( 90.89,267.93) --
	( 93.92,269.08) --
	( 96.95,270.25) --
	( 99.98,271.42) --
	(103.00,272.61) --
	(106.03,273.81) --
	(109.06,275.03) --
	(112.09,276.26) --
	(115.11,277.51) --
	(118.14,278.77) --
	(121.17,280.06) --
	(124.20,281.37) --
	(127.22,282.71) --
	(130.25,284.06) --
	(133.28,285.44) --
	(136.31,286.83) --
	(139.34,288.23) --
	(142.36,289.64) --
	(145.39,291.08) --
	(148.42,292.53) --
	(151.45,294.01) --
	(154.47,295.51) --
	(157.50,297.03) --
	(160.53,298.59) --
	(163.56,300.17) --
	(166.58,301.77) --
	(169.61,303.39) --
	(172.64,305.02) --
	(175.67,306.68) --
	(178.69,308.35) --
	(181.72,310.04) --
	(184.75,311.76) --
	(187.78,313.51) --
	(190.80,315.28) --
	(193.83,317.09) --
	(196.86,318.92) --
	(199.89,320.80) --
	(202.92,322.69) --
	(205.94,324.62) --
	(208.97,326.56) --
	(212.00,328.54) --
	(215.03,330.54) --
	(218.05,332.56) --
	(221.08,334.62) --
	(224.11,336.70) --
	(227.14,338.80) --
	(230.16,340.94) --
	(233.19,343.10) --
	(236.22,345.29) --
	(239.25,347.50) --
	(242.27,349.75) --
	(245.30,352.02) --
	(248.33,354.31) --
	(251.36,356.63) --
	(254.39,358.98) --
	(257.41,361.36) --
	(260.44,363.76) --
	(263.47,366.19) --
	(266.50,368.65);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 35.86,249.91) --
	( 38.88,250.76) --
	( 41.91,251.62) --
	( 44.94,252.51) --
	( 47.97,253.41) --
	( 50.99,254.32) --
	( 54.02,255.26) --
	( 57.05,256.21) --
	( 60.08,257.17) --
	( 63.10,258.15) --
	( 66.13,259.15) --
	( 69.16,260.17) --
	( 72.19,261.20) --
	( 75.21,262.25) --
	( 78.24,263.31) --
	( 81.27,264.39) --
	( 84.30,265.48) --
	( 87.33,266.60) --
	( 90.35,267.72) --
	( 93.38,268.87) --
	( 96.41,270.03) --
	( 99.44,271.21) --
	(102.46,272.41) --
	(105.49,273.62) --
	(108.52,274.84) --
	(111.55,276.08) --
	(114.57,277.32) --
	(117.60,278.59) --
	(120.63,279.86) --
	(123.66,281.16) --
	(126.68,282.48) --
	(129.71,283.81) --
	(132.74,285.17) --
	(135.77,286.56) --
	(138.79,287.97) --
	(141.82,289.39) --
	(144.85,290.83) --
	(147.88,292.29) --
	(150.91,293.76) --
	(153.93,295.24) --
	(156.96,296.75) --
	(159.99,298.28) --
	(163.02,299.84) --
	(166.04,301.42) --
	(169.07,303.03) --
	(172.10,304.67) --
	(175.13,306.33) --
	(178.15,308.01) --
	(181.18,309.70) --
	(184.21,311.42) --
	(187.24,313.15) --
	(190.26,314.90) --
	(193.29,316.68) --
	(196.32,318.49) --
	(199.35,320.33) --
	(202.37,322.21) --
	(205.40,324.12) --
	(208.43,326.06) --
	(211.46,328.03) --
	(214.49,330.03) --
	(217.51,332.06) --
	(220.54,334.11) --
	(223.57,336.19) --
	(226.60,338.30) --
	(229.62,340.44) --
	(232.65,342.61) --
	(235.68,344.80) --
	(238.71,347.02) --
	(241.73,349.28) --
	(244.76,351.56) --
	(247.79,353.87) --
	(250.82,356.21) --
	(253.84,358.58) --
	(256.87,360.97) --
	(259.90,363.40) --
	(262.93,365.85) --
	(265.95,368.33) --
	(268.98,370.84) --
	(272.01,373.38) --
	(275.04,375.95);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 44.40,252.39) --
	( 47.43,253.28) --
	( 50.45,254.18) --
	( 53.48,255.11) --
	( 56.51,256.05) --
	( 59.54,257.00) --
	( 62.56,257.98) --
	( 65.59,258.97) --
	( 68.62,259.97) --
	( 71.65,261.00) --
	( 74.67,262.04) --
	( 77.70,263.10) --
	( 80.73,264.17) --
	( 83.76,265.26) --
	( 86.78,266.37) --
	( 89.81,267.49) --
	( 92.84,268.63) --
	( 95.87,269.79) --
	( 98.89,270.97) --
	(101.92,272.16) --
	(104.95,273.37) --
	(107.98,274.60) --
	(111.01,275.84) --
	(114.03,277.11) --
	(117.06,278.38) --
	(120.09,279.67) --
	(123.12,280.97) --
	(126.14,282.28) --
	(129.17,283.61) --
	(132.20,284.96) --
	(135.23,286.33) --
	(138.25,287.72) --
	(141.28,289.14) --
	(144.31,290.58) --
	(147.34,292.05) --
	(150.36,293.53) --
	(153.39,295.03) --
	(156.42,296.54) --
	(159.45,298.06) --
	(162.48,299.61) --
	(165.50,301.18) --
	(168.53,302.77) --
	(171.56,304.38) --
	(174.59,306.03) --
	(177.61,307.70) --
	(180.64,309.41) --
	(183.67,311.14) --
	(186.70,312.89) --
	(189.72,314.66) --
	(192.75,316.45) --
	(195.78,318.25) --
	(198.81,320.09) --
	(201.83,321.94) --
	(204.86,323.83) --
	(207.89,325.76) --
	(210.92,327.71) --
	(213.94,329.71) --
	(216.97,331.74) --
	(220.00,333.80) --
	(223.03,335.89) --
	(226.06,338.01) --
	(229.08,340.16) --
	(232.11,342.33) --
	(235.14,344.54) --
	(238.17,346.77) --
	(241.19,349.04) --
	(244.22,351.34) --
	(247.25,353.67) --
	(250.28,356.02) --
	(253.30,358.41) --
	(256.33,360.83) --
	(259.36,363.28) --
	(262.39,365.76) --
	(265.41,368.27) --
	(268.44,370.81) --
	(271.47,373.38) --
	(274.50,375.98) --
	(277.52,378.61) --
	(280.55,381.26) --
	(283.58,383.95);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 27.31,251.38) --
	( 30.68,245.55) --
	( 23.95,245.55) --
	cycle;

\path[fill=fillColor] ( 61.48,261.57) --
	( 64.85,255.75) --
	( 58.12,255.75) --
	cycle;

\path[fill=fillColor] ( 95.65,273.63) --
	( 99.01,267.80) --
	( 92.29,267.80) --
	cycle;

\path[fill=fillColor] (129.82,287.78) --
	(133.18,281.95) --
	(126.46,281.95) --
	cycle;

\path[fill=fillColor] (163.99,304.26) --
	(167.35,298.43) --
	(160.63,298.43) --
	cycle;

\path[fill=fillColor] (198.16,323.63) --
	(201.52,317.81) --
	(194.79,317.81) --
	cycle;

\path[fill=fillColor] (232.33,346.20) --
	(235.69,340.37) --
	(228.96,340.37) --
	cycle;

\path[fill=fillColor] (266.50,372.59) --
	(269.86,366.77) --
	(263.13,366.77) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 33.36,247.39) --
	( 38.35,247.39) --
	( 38.35,252.38) --
	( 33.36,252.38) --
	cycle;

\path[fill=fillColor] ( 67.53,258.02) --
	( 72.52,258.02) --
	( 72.52,263.01) --
	( 67.53,263.01) --
	cycle;

\path[fill=fillColor] (101.70,270.61) --
	(106.69,270.61) --
	(106.69,275.61) --
	(101.70,275.61) --
	cycle;

\path[fill=fillColor] (135.86,285.30) --
	(140.86,285.30) --
	(140.86,290.30) --
	(135.86,290.30) --
	cycle;

\path[fill=fillColor] (170.03,302.36) --
	(175.03,302.36) --
	(175.03,307.36) --
	(170.03,307.36) --
	cycle;

\path[fill=fillColor] (204.20,322.52) --
	(209.20,322.52) --
	(209.20,327.51) --
	(204.20,327.51) --
	cycle;

\path[fill=fillColor] (238.37,345.90) --
	(243.37,345.90) --
	(243.37,350.89) --
	(238.37,350.89) --
	cycle;

\path[fill=fillColor] (272.54,373.53) --
	(277.54,373.53) --
	(277.54,378.53) --
	(272.54,378.53) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87,252.37) -- ( 47.93,252.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.40,248.83) -- ( 44.40,255.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.03,263.46) -- ( 82.10,263.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.57,259.93) -- ( 78.57,266.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.20,276.59) -- (116.27,276.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.74,273.06) -- (112.74,280.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.37,291.84) -- (150.44,291.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.90,288.31) -- (146.90,295.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.54,309.66) -- (184.61,309.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.07,306.13) -- (181.07,313.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.71,330.57) -- (218.77,330.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.24,327.04) -- (215.24,334.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.88,355.17) -- (252.94,355.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.41,351.64) -- (249.41,358.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.05,384.02) -- (287.11,384.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.58,380.49) -- (283.58,387.55);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 52.94,254.96) circle (  2.50);

\path[fill=fillColor] ( 87.11,266.50) circle (  2.50);

\path[fill=fillColor] (121.28,280.09) circle (  2.50);

\path[fill=fillColor] (155.45,296.05) circle (  2.50);

\path[fill=fillColor] (189.62,314.74) circle (  2.50);

\path[fill=fillColor] (223.78,336.55) circle (  2.50);

\path[fill=fillColor] (257.95,361.93) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31,247.50) rectangle (283.58,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,247.50) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,294.20) --
	( 27.31,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,340.91) --
	( 27.31,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,387.61) --
	( 27.31,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,244.75) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.74,244.75) --
	(112.74,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.16,244.75) --
	(198.16,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.74,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.16,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (155.45,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,329.23) {Reproduction number $\mathcal R$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.06,332.94) rectangle (192.70,407.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 36.56,381.80) rectangle ( 51.02,396.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 38.01,389.03) -- ( 49.57,389.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 43.79,389.03) circle (  2.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 36.56,367.35) rectangle ( 51.02,381.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 38.01,374.57) -- ( 49.57,374.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 43.79,378.46) --
	( 47.15,372.63) --
	( 40.43,372.63) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 36.56,352.89) rectangle ( 51.02,367.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 38.01,360.12) -- ( 49.57,360.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 41.29,357.62) --
	( 46.29,357.62) --
	( 46.29,362.62) --
	( 41.29,362.62) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 36.56,338.44) rectangle ( 51.02,352.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 38.01,345.67) -- ( 49.57,345.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.26,345.67) -- ( 47.32,345.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.79,342.13) -- ( 43.79,349.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 56.52,386.00) {Intrinsic generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 56.52,371.54) {Forward serial interval ($\rho=0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 56.52,357.09) {Forward serial interval ($\rho=0$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 56.52,342.63) {Forward serial interval ($\rho=-0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,419.03) {A};
\end{scope}
\begin{scope}
\path[clip] (289.08,216.81) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,216.81) rectangle (578.16,433.62);
\end{scope}
\begin{scope}
\path[clip] (316.39,247.50) rectangle (572.66,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (316.39,247.50) rectangle (572.66,410.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (316.39,247.64) --
	(319.64,248.49) --
	(322.88,249.36) --
	(326.12,250.25) --
	(329.37,251.16) --
	(332.61,252.09) --
	(335.86,253.03) --
	(339.10,254.00) --
	(342.34,254.99) --
	(345.59,255.99) --
	(348.83,257.01) --
	(352.08,258.06) --
	(355.32,259.12) --
	(358.56,260.20) --
	(361.81,261.30) --
	(365.05,262.42) --
	(368.30,263.56) --
	(371.54,264.72) --
	(374.78,265.90) --
	(378.03,267.10) --
	(381.27,268.32) --
	(384.51,269.55) --
	(387.76,270.81) --
	(391.00,272.08) --
	(394.25,273.37) --
	(397.49,274.68) --
	(400.73,276.01) --
	(403.98,277.36) --
	(407.22,278.74) --
	(410.47,280.13) --
	(413.71,281.55) --
	(416.95,282.99) --
	(420.20,284.44) --
	(423.44,285.91) --
	(426.69,287.40) --
	(429.93,288.91) --
	(433.17,290.45) --
	(436.42,292.02) --
	(439.66,293.61) --
	(442.90,295.23) --
	(446.15,296.88) --
	(449.39,298.55) --
	(452.64,300.24) --
	(455.88,301.95) --
	(459.12,303.68) --
	(462.37,305.44) --
	(465.61,307.23) --
	(468.86,309.04) --
	(472.10,310.89) --
	(475.34,312.76) --
	(478.59,314.68) --
	(481.83,316.62) --
	(485.08,318.59) --
	(488.32,320.58) --
	(491.56,322.60) --
	(494.81,324.66) --
	(498.05,326.74) --
	(501.29,328.85) --
	(504.54,330.99) --
	(507.78,333.17) --
	(511.03,335.39) --
	(514.27,337.63) --
	(517.51,339.91) --
	(520.76,342.22) --
	(524.00,344.56) --
	(527.25,346.93) --
	(530.49,349.33) --
	(533.73,351.76) --
	(536.98,354.22) --
	(540.22,356.72) --
	(543.47,359.24) --
	(546.71,361.80) --
	(549.95,364.38) --
	(553.20,367.00) --
	(556.44,369.66) --
	(559.68,372.34) --
	(562.93,375.05) --
	(566.17,377.80) --
	(569.42,380.57) --
	(572.66,383.38);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (319.64,248.30) --
	(322.88,249.24) --
	(326.12,250.18) --
	(329.37,251.12) --
	(332.61,252.08) --
	(335.86,253.04) --
	(339.10,254.00) --
	(342.34,254.98) --
	(345.59,255.96) --
	(348.83,256.94) --
	(352.08,257.93) --
	(355.32,258.93) --
	(358.56,259.94) --
	(361.81,260.95) --
	(365.05,261.97) --
	(368.30,262.99) --
	(371.54,264.02) --
	(374.78,265.05) --
	(378.03,266.09) --
	(381.27,267.14) --
	(384.51,268.20) --
	(387.76,269.26) --
	(391.00,270.34) --
	(394.25,271.42) --
	(397.49,272.50) --
	(400.73,273.59) --
	(403.98,274.68) --
	(407.22,275.78) --
	(410.47,276.87) --
	(413.71,277.98) --
	(416.95,279.10) --
	(420.20,280.22) --
	(423.44,281.36) --
	(426.69,282.50) --
	(429.93,283.64) --
	(433.17,284.78) --
	(436.42,285.92) --
	(439.66,287.06) --
	(442.90,288.19) --
	(446.15,289.31) --
	(449.39,290.45) --
	(452.64,291.59) --
	(455.88,292.74) --
	(459.12,293.89) --
	(462.37,295.05) --
	(465.61,296.19) --
	(468.86,297.34) --
	(472.10,298.47) --
	(475.34,299.59) --
	(478.59,300.69) --
	(481.83,301.79) --
	(485.08,302.89) --
	(488.32,303.98) --
	(491.56,305.07) --
	(494.81,306.15) --
	(498.05,307.23) --
	(501.29,308.29) --
	(504.54,309.34) --
	(507.78,310.37) --
	(511.03,311.39) --
	(514.27,312.39) --
	(517.51,313.37) --
	(520.76,314.35) --
	(524.00,315.32) --
	(527.25,316.28) --
	(530.49,317.22) --
	(533.73,318.16) --
	(536.98,319.09) --
	(540.22,320.00) --
	(543.47,320.91) --
	(546.71,321.80) --
	(549.95,322.67) --
	(553.20,323.54) --
	(556.44,324.40) --
	(559.68,325.24) --
	(562.93,326.07) --
	(566.17,326.89) --
	(569.42,327.70) --
	(572.66,328.50);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (319.64,248.18) --
	(322.88,249.15) --
	(326.12,250.12) --
	(329.37,251.09) --
	(332.61,252.06) --
	(335.86,253.02) --
	(339.10,253.99) --
	(342.34,254.95) --
	(345.59,255.91) --
	(348.83,256.87) --
	(352.08,257.82) --
	(355.32,258.78) --
	(358.56,259.73) --
	(361.81,260.68) --
	(365.05,261.63) --
	(368.30,262.58) --
	(371.54,263.52) --
	(374.78,264.45) --
	(378.03,265.39) --
	(381.27,266.33) --
	(384.51,267.27) --
	(387.76,268.21) --
	(391.00,269.15) --
	(394.25,270.09) --
	(397.49,271.03) --
	(400.73,271.96) --
	(403.98,272.88) --
	(407.22,273.79) --
	(410.47,274.70) --
	(413.71,275.60) --
	(416.95,276.51) --
	(420.20,277.43) --
	(423.44,278.34) --
	(426.69,279.25) --
	(429.93,280.15) --
	(433.17,281.03) --
	(436.42,281.90) --
	(439.66,282.74) --
	(442.90,283.56) --
	(446.15,284.35) --
	(449.39,285.14) --
	(452.64,285.93) --
	(455.88,286.72) --
	(459.12,287.50) --
	(462.37,288.26) --
	(465.61,289.01) --
	(468.86,289.72) --
	(472.10,290.40) --
	(475.34,291.04) --
	(478.59,291.64) --
	(481.83,292.21) --
	(485.08,292.76) --
	(488.32,293.29) --
	(491.56,293.79) --
	(494.81,294.27) --
	(498.05,294.72) --
	(501.29,295.13) --
	(504.54,295.51) --
	(507.78,295.84) --
	(511.03,296.12) --
	(514.27,296.36) --
	(517.51,296.57) --
	(520.76,296.74) --
	(524.00,296.88) --
	(527.25,296.99) --
	(530.49,297.06) --
	(533.73,297.10) --
	(536.98,297.11) --
	(540.22,297.08) --
	(543.47,297.02) --
	(546.71,296.91) --
	(549.95,296.77) --
	(553.20,296.60) --
	(556.44,296.39) --
	(559.68,296.14) --
	(562.93,295.86) --
	(566.17,295.54) --
	(569.42,295.19) --
	(572.66,294.80);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (319.64,248.10) --
	(322.88,249.10) --
	(326.12,250.09) --
	(329.37,251.07) --
	(332.61,252.04) --
	(335.86,253.00) --
	(339.10,253.95) --
	(342.34,254.90) --
	(345.59,255.84) --
	(348.83,256.77) --
	(352.08,257.69) --
	(355.32,258.60) --
	(358.56,259.51) --
	(361.81,260.40) --
	(365.05,261.29) --
	(368.30,262.16) --
	(371.54,263.02) --
	(374.78,263.87) --
	(378.03,264.71) --
	(381.27,265.55) --
	(384.51,266.39) --
	(387.76,267.23) --
	(391.00,268.05) --
	(394.25,268.87) --
	(397.49,269.68) --
	(400.73,270.47) --
	(403.98,271.25) --
	(407.22,272.00) --
	(410.47,272.74) --
	(413.71,273.47) --
	(416.95,274.20) --
	(420.20,274.93) --
	(423.44,275.65) --
	(426.69,276.36) --
	(429.93,277.05) --
	(433.17,277.72) --
	(436.42,278.35) --
	(439.66,278.95) --
	(442.90,279.51) --
	(446.15,280.03) --
	(449.39,280.54) --
	(452.64,281.05) --
	(455.88,281.54) --
	(459.12,282.01) --
	(462.37,282.46) --
	(465.61,282.87) --
	(468.86,283.25) --
	(472.10,283.58) --
	(475.34,283.86) --
	(478.59,284.08) --
	(481.83,284.25) --
	(485.08,284.40) --
	(488.32,284.51) --
	(491.56,284.59) --
	(494.81,284.63) --
	(498.05,284.63) --
	(501.29,284.58) --
	(504.54,284.48) --
	(507.78,284.34) --
	(511.03,284.13) --
	(514.27,283.88) --
	(517.51,283.57) --
	(520.76,283.23) --
	(524.00,282.84) --
	(527.25,282.40) --
	(530.49,281.93) --
	(533.73,281.41) --
	(536.98,280.84) --
	(540.22,280.23) --
	(543.47,279.57) --
	(546.71,278.87) --
	(549.95,278.12) --
	(553.20,277.32) --
	(556.44,276.47) --
	(559.68,275.58) --
	(562.93,274.65) --
	(566.17,273.66) --
	(569.42,272.64) --
	(572.66,271.57);
\definecolor{fillColor}{RGB}{0,158,115}

\path[fill=fillColor] (316.39,251.38) --
	(319.76,245.55) --
	(313.03,245.55) --
	cycle;

\path[fill=fillColor] (324.94,253.76) --
	(328.30,247.93) --
	(321.57,247.93) --
	cycle;

\path[fill=fillColor] (333.48,256.20) --
	(336.84,250.37) --
	(330.11,250.37) --
	cycle;

\path[fill=fillColor] (342.02,258.70) --
	(345.38,252.88) --
	(338.66,252.88) --
	cycle;

\path[fill=fillColor] (350.56,261.27) --
	(353.93,255.44) --
	(347.20,255.44) --
	cycle;

\path[fill=fillColor] (359.10,263.89) --
	(362.47,258.07) --
	(355.74,258.07) --
	cycle;

\path[fill=fillColor] (367.65,266.57) --
	(371.01,260.75) --
	(364.28,260.75) --
	cycle;

\path[fill=fillColor] (376.19,269.31) --
	(379.55,263.48) --
	(372.83,263.48) --
	cycle;

\path[fill=fillColor] (384.73,272.09) --
	(388.09,266.27) --
	(381.37,266.27) --
	cycle;

\path[fill=fillColor] (393.27,274.93) --
	(396.64,269.10) --
	(389.91,269.10) --
	cycle;

\path[fill=fillColor] (401.82,277.80) --
	(405.18,271.97) --
	(398.45,271.97) --
	cycle;

\path[fill=fillColor] (410.36,280.71) --
	(413.72,274.89) --
	(406.99,274.89) --
	cycle;

\path[fill=fillColor] (418.90,283.66) --
	(422.26,277.83) --
	(415.54,277.83) --
	cycle;

\path[fill=fillColor] (427.44,286.63) --
	(430.81,280.80) --
	(424.08,280.80) --
	cycle;

\path[fill=fillColor] (435.98,289.62) --
	(439.35,283.79) --
	(432.62,283.79) --
	cycle;

\path[fill=fillColor] (444.53,292.62) --
	(447.89,286.79) --
	(441.16,286.79) --
	cycle;

\path[fill=fillColor] (453.07,295.63) --
	(456.43,289.80) --
	(449.71,289.80) --
	cycle;

\path[fill=fillColor] (461.61,298.63) --
	(464.97,292.81) --
	(458.25,292.81) --
	cycle;

\path[fill=fillColor] (470.15,301.63) --
	(473.52,295.80) --
	(466.79,295.80) --
	cycle;

\path[fill=fillColor] (478.70,304.60) --
	(482.06,298.77) --
	(475.33,298.77) --
	cycle;

\path[fill=fillColor] (487.24,307.54) --
	(490.60,301.71) --
	(483.87,301.71) --
	cycle;

\path[fill=fillColor] (495.78,310.43) --
	(499.14,304.60) --
	(492.42,304.60) --
	cycle;

\path[fill=fillColor] (504.32,313.26) --
	(507.69,307.44) --
	(500.96,307.44) --
	cycle;

\path[fill=fillColor] (512.86,316.03) --
	(516.23,310.20) --
	(509.50,310.20) --
	cycle;

\path[fill=fillColor] (521.41,318.70) --
	(524.77,312.87) --
	(518.04,312.87) --
	cycle;

\path[fill=fillColor] (529.95,321.27) --
	(533.31,315.44) --
	(526.59,315.44) --
	cycle;

\path[fill=fillColor] (538.49,323.72) --
	(541.85,317.89) --
	(535.13,317.89) --
	cycle;

\path[fill=fillColor] (547.03,326.03) --
	(550.40,320.20) --
	(543.67,320.20) --
	cycle;

\path[fill=fillColor] (555.58,328.18) --
	(558.94,322.35) --
	(552.21,322.35) --
	cycle;

\path[fill=fillColor] (564.12,330.14) --
	(567.48,324.32) --
	(560.75,324.32) --
	cycle;

\path[fill=fillColor] (572.66,331.91) --
	(576.02,326.08) --
	(569.30,326.08) --
	cycle;
\definecolor{fillColor}{RGB}{240,228,66}

\path[fill=fillColor] (313.90,245.00) --
	(318.89,245.00) --
	(318.89,249.99) --
	(313.90,249.99) --
	cycle;

\path[fill=fillColor] (322.44,247.36) --
	(327.43,247.36) --
	(327.43,252.36) --
	(322.44,252.36) --
	cycle;

\path[fill=fillColor] (330.98,249.77) --
	(335.98,249.77) --
	(335.98,254.76) --
	(330.98,254.76) --
	cycle;

\path[fill=fillColor] (339.52,252.21) --
	(344.52,252.21) --
	(344.52,257.20) --
	(339.52,257.20) --
	cycle;

\path[fill=fillColor] (348.06,254.68) --
	(353.06,254.68) --
	(353.06,259.67) --
	(348.06,259.67) --
	cycle;

\path[fill=fillColor] (356.61,257.17) --
	(361.60,257.17) --
	(361.60,262.16) --
	(356.61,262.16) --
	cycle;

\path[fill=fillColor] (365.15,259.67) --
	(370.14,259.67) --
	(370.14,264.67) --
	(365.15,264.67) --
	cycle;

\path[fill=fillColor] (373.69,262.19) --
	(378.69,262.19) --
	(378.69,267.19) --
	(373.69,267.19) --
	cycle;

\path[fill=fillColor] (382.23,264.71) --
	(387.23,264.71) --
	(387.23,269.70) --
	(382.23,269.70) --
	cycle;

\path[fill=fillColor] (390.78,267.21) --
	(395.77,267.21) --
	(395.77,272.21) --
	(390.78,272.21) --
	cycle;

\path[fill=fillColor] (399.32,269.70) --
	(404.31,269.70) --
	(404.31,274.69) --
	(399.32,274.69) --
	cycle;

\path[fill=fillColor] (407.86,272.15) --
	(412.86,272.15) --
	(412.86,277.15) --
	(407.86,277.15) --
	cycle;

\path[fill=fillColor] (416.40,274.57) --
	(421.40,274.57) --
	(421.40,279.56) --
	(416.40,279.56) --
	cycle;

\path[fill=fillColor] (424.94,276.93) --
	(429.94,276.93) --
	(429.94,281.93) --
	(424.94,281.93) --
	cycle;

\path[fill=fillColor] (433.49,279.22) --
	(438.48,279.22) --
	(438.48,284.22) --
	(433.49,284.22) --
	cycle;

\path[fill=fillColor] (442.03,281.44) --
	(447.02,281.44) --
	(447.02,286.43) --
	(442.03,286.43) --
	cycle;

\path[fill=fillColor] (450.57,283.55) --
	(455.57,283.55) --
	(455.57,288.55) --
	(450.57,288.55) --
	cycle;

\path[fill=fillColor] (459.11,285.55) --
	(464.11,285.55) --
	(464.11,290.55) --
	(459.11,290.55) --
	cycle;

\path[fill=fillColor] (467.66,287.42) --
	(472.65,287.42) --
	(472.65,292.42) --
	(467.66,292.42) --
	cycle;

\path[fill=fillColor] (476.20,289.14) --
	(481.19,289.14) --
	(481.19,294.14) --
	(476.20,294.14) --
	cycle;

\path[fill=fillColor] (484.74,290.69) --
	(489.74,290.69) --
	(489.74,295.69) --
	(484.74,295.69) --
	cycle;

\path[fill=fillColor] (493.28,292.05) --
	(498.28,292.05) --
	(498.28,297.04) --
	(493.28,297.04) --
	cycle;

\path[fill=fillColor] (501.82,293.19) --
	(506.82,293.19) --
	(506.82,298.19) --
	(501.82,298.19) --
	cycle;

\path[fill=fillColor] (510.37,294.09) --
	(515.36,294.09) --
	(515.36,299.09) --
	(510.37,299.09) --
	cycle;

\path[fill=fillColor] (518.91,294.73) --
	(523.90,294.73) --
	(523.90,299.73) --
	(518.91,299.73) --
	cycle;

\path[fill=fillColor] (527.45,295.08) --
	(532.45,295.08) --
	(532.45,300.07) --
	(527.45,300.07) --
	cycle;

\path[fill=fillColor] (535.99,295.11) --
	(540.99,295.11) --
	(540.99,300.11) --
	(535.99,300.11) --
	cycle;

\path[fill=fillColor] (544.54,294.80) --
	(549.53,294.80) --
	(549.53,299.79) --
	(544.54,299.79) --
	cycle;

\path[fill=fillColor] (553.08,294.12) --
	(558.07,294.12) --
	(558.07,299.12) --
	(553.08,299.12) --
	cycle;

\path[fill=fillColor] (561.62,293.05) --
	(566.62,293.05) --
	(566.62,298.05) --
	(561.62,298.05) --
	cycle;

\path[fill=fillColor] (570.16,291.57) --
	(575.16,291.57) --
	(575.16,296.57) --
	(570.16,296.57) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.86,247.50) -- (319.93,247.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.39,243.96) -- (316.39,251.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.40,249.85) -- (328.47,249.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (324.94,246.32) -- (324.94,253.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.95,252.22) -- (337.01,252.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.48,248.69) -- (333.48,255.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.49,254.60) -- (345.55,254.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.02,251.07) -- (342.02,258.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (347.03,256.98) -- (354.09,256.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.56,253.45) -- (350.56,260.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (355.57,259.35) -- (362.64,259.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.10,255.81) -- (359.10,262.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.11,261.70) -- (371.18,261.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.65,258.16) -- (367.65,265.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.66,264.01) -- (379.72,264.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.19,260.48) -- (376.19,267.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.20,266.28) -- (388.26,266.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.73,262.75) -- (384.73,269.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.74,268.50) -- (396.81,268.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.27,264.97) -- (393.27,272.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.28,270.64) -- (405.35,270.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.82,267.11) -- (401.82,274.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.83,272.70) -- (413.89,272.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.36,269.17) -- (410.36,276.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (415.37,274.66) -- (422.43,274.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.90,271.13) -- (418.90,278.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.91,276.50) -- (430.97,276.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.44,272.97) -- (427.44,280.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (432.45,278.21) -- (439.52,278.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.98,274.67) -- (435.98,281.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (440.99,279.76) -- (448.06,279.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.53,276.23) -- (444.53,283.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (449.54,281.15) -- (456.60,281.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.07,277.61) -- (453.07,284.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.08,282.34) -- (465.14,282.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.61,278.81) -- (461.61,285.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (466.62,283.32) -- (473.69,283.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.15,279.79) -- (470.15,286.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.16,284.08) -- (482.23,284.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.70,280.55) -- (478.70,287.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.71,284.58) -- (490.77,284.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.24,281.04) -- (487.24,288.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (492.25,284.80) -- (499.31,284.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.78,281.27) -- (495.78,288.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.79,284.72) -- (507.85,284.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (504.32,281.19) -- (504.32,288.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (509.33,284.32) -- (516.40,284.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (512.86,280.79) -- (512.86,287.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (517.87,283.58) -- (524.94,283.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (521.41,280.04) -- (521.41,287.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.42,282.47) -- (533.48,282.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (529.95,278.94) -- (529.95,286.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (534.96,280.98) -- (542.02,280.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (538.49,277.44) -- (538.49,284.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (543.50,279.09) -- (550.57,279.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.03,275.56) -- (547.03,282.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.04,276.80) -- (559.11,276.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (555.58,273.26) -- (555.58,280.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.59,274.10) -- (567.65,274.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.12,270.57) -- (564.12,277.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13,271.01) -- (576.19,271.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66,267.47) -- (572.66,274.54);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (316.39,247.50) circle (  2.50);

\path[fill=fillColor] (324.94,249.88) circle (  2.50);

\path[fill=fillColor] (333.48,252.37) circle (  2.50);

\path[fill=fillColor] (342.02,254.96) circle (  2.50);

\path[fill=fillColor] (350.56,257.67) circle (  2.50);

\path[fill=fillColor] (359.10,260.49) circle (  2.50);

\path[fill=fillColor] (367.65,263.43) circle (  2.50);

\path[fill=fillColor] (376.19,266.50) circle (  2.50);

\path[fill=fillColor] (384.73,269.69) circle (  2.50);

\path[fill=fillColor] (393.27,273.02) circle (  2.50);

\path[fill=fillColor] (401.82,276.48) circle (  2.50);

\path[fill=fillColor] (410.36,280.09) circle (  2.50);

\path[fill=fillColor] (418.90,283.85) circle (  2.50);

\path[fill=fillColor] (427.44,287.76) circle (  2.50);

\path[fill=fillColor] (435.98,291.82) circle (  2.50);

\path[fill=fillColor] (444.53,296.05) circle (  2.50);

\path[fill=fillColor] (453.07,300.46) circle (  2.50);

\path[fill=fillColor] (461.61,305.04) circle (  2.50);

\path[fill=fillColor] (470.15,309.80) circle (  2.50);

\path[fill=fillColor] (478.70,314.74) circle (  2.50);

\path[fill=fillColor] (487.24,319.89) circle (  2.50);

\path[fill=fillColor] (495.78,325.23) circle (  2.50);

\path[fill=fillColor] (504.32,330.79) circle (  2.50);

\path[fill=fillColor] (512.86,336.55) circle (  2.50);

\path[fill=fillColor] (521.41,342.55) circle (  2.50);

\path[fill=fillColor] (529.95,348.77) circle (  2.50);

\path[fill=fillColor] (538.49,355.23) circle (  2.50);

\path[fill=fillColor] (547.03,361.93) circle (  2.50);

\path[fill=fillColor] (555.58,368.89) circle (  2.50);

\path[fill=fillColor] (564.12,376.12) circle (  2.50);

\path[fill=fillColor] (572.66,383.61) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.39,247.50) rectangle (572.66,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,247.50) --
	(316.39,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,294.20) --
	(316.39,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,340.91) --
	(316.39,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,387.61) --
	(316.39,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.39,244.75) --
	(316.39,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.82,244.75) --
	(401.82,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.24,244.75) --
	(487.24,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (316.39,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (401.82,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (487.24,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (444.53,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,329.23) {Reproduction number ${\mathcal R}_{\textrm{\tiny intrinsic}}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (320.07,332.94) rectangle (481.85,407.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (325.57,381.80) rectangle (340.02,396.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (327.01,389.03) -- (338.58,389.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (332.80,389.03) circle (  2.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (325.57,367.35) rectangle (340.02,381.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (327.01,374.57) -- (338.58,374.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{0,158,115}

\path[fill=fillColor] (332.80,378.46) --
	(336.16,372.63) --
	(329.43,372.63) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (325.57,352.89) rectangle (340.02,367.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (327.01,360.12) -- (338.58,360.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{240,228,66}

\path[fill=fillColor] (330.30,357.62) --
	(335.29,357.62) --
	(335.29,362.62) --
	(330.30,362.62) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (325.57,338.44) rectangle (340.02,352.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (327.01,345.67) -- (338.58,345.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.26,345.67) -- (336.33,345.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (332.80,342.13) -- (332.80,349.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (345.52,386.00) {Intrinsic generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (345.52,371.54) {Intrinsic serial interval ($\rho=0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (345.52,357.09) {Intrinsic serial interval ($\rho=0$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (345.52,342.63) {Intrinsic serial interval ($\rho=-0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (316.39,419.03) {B};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 27.31, 30.69) rectangle (283.58,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31, 30.69) rectangle (283.58,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31, 54.98) --
	( 30.56, 54.98) --
	( 33.80, 54.98) --
	( 37.04, 54.98) --
	( 40.29, 54.98) --
	( 43.53, 54.98) --
	( 46.78, 54.98) --
	( 50.02, 54.98) --
	( 53.26, 54.98) --
	( 56.51, 54.98) --
	( 59.75, 54.98) --
	( 63.00, 54.98) --
	( 66.24, 54.98) --
	( 69.48, 54.98) --
	( 72.73, 54.98) --
	( 75.97, 54.98) --
	( 79.22, 54.98) --
	( 82.46, 54.98) --
	( 85.70, 54.98) --
	( 88.95, 54.98) --
	( 92.19, 54.98) --
	( 95.43, 54.98) --
	( 98.68, 54.98) --
	(101.92, 54.98) --
	(105.17, 54.98) --
	(108.41, 54.98) --
	(111.65, 54.98) --
	(114.90, 54.98) --
	(118.14, 54.98) --
	(121.39, 54.98) --
	(124.63, 54.98) --
	(127.87, 54.98) --
	(131.12, 54.98) --
	(134.36, 54.98) --
	(137.61, 54.98) --
	(140.85, 54.98) --
	(144.09, 54.98) --
	(147.34, 54.98) --
	(150.58, 54.98) --
	(153.82, 54.98) --
	(157.07, 54.98) --
	(160.31, 54.98) --
	(163.56, 54.98) --
	(166.80, 54.98) --
	(170.04, 54.98) --
	(173.29, 54.98) --
	(176.53, 54.98) --
	(179.78, 54.98) --
	(183.02, 54.98) --
	(186.26, 54.98) --
	(189.51, 54.98) --
	(192.75, 54.98) --
	(196.00, 54.98) --
	(199.24, 54.98) --
	(202.48, 54.98) --
	(205.73, 54.98) --
	(208.97, 54.98) --
	(212.21, 54.98) --
	(215.46, 54.98) --
	(218.70, 54.98) --
	(221.95, 54.98) --
	(225.19, 54.98) --
	(228.43, 54.98) --
	(231.68, 54.98) --
	(234.92, 54.98) --
	(238.17, 54.98) --
	(241.41, 54.98) --
	(244.65, 54.98) --
	(247.90, 54.98) --
	(251.14, 54.98) --
	(254.39, 54.98) --
	(257.63, 54.98) --
	(260.87, 54.98) --
	(264.12, 54.98) --
	(267.36, 54.98) --
	(270.60, 54.98) --
	(273.85, 54.98) --
	(277.09, 54.98) --
	(280.34, 54.98) --
	(283.58, 54.98);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 27.31, 55.70) --
	( 30.56, 56.81) --
	( 33.80, 57.90) --
	( 37.04, 58.98) --
	( 40.29, 60.06) --
	( 43.53, 61.12) --
	( 46.78, 62.16) --
	( 50.02, 63.20) --
	( 53.26, 64.22) --
	( 56.51, 65.23) --
	( 59.75, 66.23) --
	( 63.00, 67.21) --
	( 66.24, 68.18) --
	( 69.48, 69.14) --
	( 72.73, 70.09) --
	( 75.97, 71.03) --
	( 79.22, 71.95) --
	( 82.46, 72.86) --
	( 85.70, 73.77) --
	( 88.95, 74.66) --
	( 92.19, 75.53) --
	( 95.43, 76.39) --
	( 98.68, 77.24) --
	(101.92, 78.07) --
	(105.17, 78.90) --
	(108.41, 79.71) --
	(111.65, 80.51) --
	(114.90, 81.30) --
	(118.14, 82.08) --
	(121.39, 82.85) --
	(124.63, 83.61) --
	(127.87, 84.35) --
	(131.12, 85.08) --
	(134.36, 85.79) --
	(137.61, 86.49) --
	(140.85, 87.19) --
	(144.09, 87.87) --
	(147.34, 88.55) --
	(150.58, 89.23) --
	(153.82, 89.91) --
	(157.07, 90.59) --
	(160.31, 91.26) --
	(163.56, 91.92) --
	(166.80, 92.57) --
	(170.04, 93.22) --
	(173.29, 93.86) --
	(176.53, 94.49) --
	(179.78, 95.12) --
	(183.02, 95.75) --
	(186.26, 96.37) --
	(189.51, 96.99) --
	(192.75, 97.61) --
	(196.00, 98.22) --
	(199.24, 98.82) --
	(202.48, 99.42) --
	(205.73,100.02) --
	(208.97,100.61) --
	(212.21,101.20) --
	(215.46,101.78) --
	(218.70,102.35) --
	(221.95,102.93) --
	(225.19,103.50) --
	(228.43,104.06) --
	(231.68,104.62) --
	(234.92,105.17) --
	(238.17,105.72) --
	(241.41,106.27) --
	(244.65,106.81) --
	(247.90,107.34) --
	(251.14,107.87) --
	(254.39,108.40) --
	(257.63,108.92) --
	(260.87,109.43) --
	(264.12,109.95) --
	(267.36,110.45) --
	(270.60,110.95) --
	(273.85,111.45) --
	(277.09,111.95) --
	(280.34,112.43) --
	(283.58,112.92);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 27.31, 55.54) --
	( 30.56, 57.34) --
	( 33.80, 59.12) --
	( 37.04, 60.88) --
	( 40.29, 62.63) --
	( 43.53, 64.35) --
	( 46.78, 66.06) --
	( 50.02, 67.75) --
	( 53.26, 69.43) --
	( 56.51, 71.08) --
	( 59.75, 72.71) --
	( 63.00, 74.33) --
	( 66.24, 75.93) --
	( 69.48, 77.51) --
	( 72.73, 79.07) --
	( 75.97, 80.62) --
	( 79.22, 82.15) --
	( 82.46, 83.66) --
	( 85.70, 85.16) --
	( 88.95, 86.63) --
	( 92.19, 88.09) --
	( 95.43, 89.53) --
	( 98.68, 90.94) --
	(101.92, 92.34) --
	(105.17, 93.72) --
	(108.41, 95.08) --
	(111.65, 96.43) --
	(114.90, 97.76) --
	(118.14, 99.08) --
	(121.39,100.38) --
	(124.63,101.67) --
	(127.87,102.92) --
	(131.12,104.16) --
	(134.36,105.37) --
	(137.61,106.57) --
	(140.85,107.76) --
	(144.09,108.93) --
	(147.34,110.10) --
	(150.58,111.27) --
	(153.82,112.43) --
	(157.07,113.59) --
	(160.31,114.75) --
	(163.56,115.89) --
	(166.80,117.02) --
	(170.04,118.14) --
	(173.29,119.26) --
	(176.53,120.36) --
	(179.78,121.46) --
	(183.02,122.55) --
	(186.26,123.63) --
	(189.51,124.71) --
	(192.75,125.78) --
	(196.00,126.85) --
	(199.24,127.91) --
	(202.48,128.96) --
	(205.73,130.00) --
	(208.97,131.03) --
	(212.21,132.06) --
	(215.46,133.08) --
	(218.70,134.09) --
	(221.95,135.10) --
	(225.19,136.09) --
	(228.43,137.09) --
	(231.68,138.07) --
	(234.92,139.04) --
	(238.17,140.01) --
	(241.41,140.97) --
	(244.65,141.92) --
	(247.90,142.87) --
	(251.14,143.81) --
	(254.39,144.74) --
	(257.63,145.66) --
	(260.87,146.58) --
	(264.12,147.48) --
	(267.36,148.38) --
	(270.60,149.28) --
	(273.85,150.16) --
	(277.09,151.04) --
	(280.34,151.91) --
	(283.58,152.77);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 55.20) --
	( 30.56, 57.64) --
	( 33.80, 60.07) --
	( 37.04, 62.47) --
	( 40.29, 64.85) --
	( 43.53, 67.21) --
	( 46.78, 69.55) --
	( 50.02, 71.86) --
	( 53.26, 74.15) --
	( 56.51, 76.43) --
	( 59.75, 78.67) --
	( 63.00, 80.90) --
	( 66.24, 83.10) --
	( 69.48, 85.29) --
	( 72.73, 87.45) --
	( 75.97, 89.59) --
	( 79.22, 91.71) --
	( 82.46, 93.81) --
	( 85.70, 95.89) --
	( 88.95, 97.94) --
	( 92.19, 99.98) --
	( 95.43,101.98) --
	( 98.68,103.97) --
	(101.92,105.93) --
	(105.17,107.87) --
	(108.41,109.79) --
	(111.65,111.69) --
	(114.90,113.57) --
	(118.14,115.43) --
	(121.39,117.28) --
	(124.63,119.10) --
	(127.87,120.88) --
	(131.12,122.64) --
	(134.36,124.37) --
	(137.61,126.09) --
	(140.85,127.78) --
	(144.09,129.47) --
	(147.34,131.15) --
	(150.58,132.82) --
	(153.82,134.50) --
	(157.07,136.18) --
	(160.31,137.85) --
	(163.56,139.50) --
	(166.80,141.14) --
	(170.04,142.77) --
	(173.29,144.39) --
	(176.53,146.00) --
	(179.78,147.60) --
	(183.02,149.19) --
	(186.26,150.78) --
	(189.51,152.36) --
	(192.75,153.93) --
	(196.00,155.49) --
	(199.24,157.04) --
	(202.48,158.58) --
	(205.73,160.11) --
	(208.97,161.63) --
	(212.21,163.15) --
	(215.46,164.65) --
	(218.70,166.14) --
	(221.95,167.63) --
	(225.19,169.10) --
	(228.43,170.57) --
	(231.68,172.02) --
	(234.92,173.47) --
	(238.17,174.91) --
	(241.41,176.33) --
	(244.65,177.75) --
	(247.90,179.16) --
	(251.14,180.56) --
	(254.39,181.95) --
	(257.63,183.33) --
	(260.87,184.70) --
	(264.12,186.06) --
	(267.36,187.41) --
	(270.60,188.75) --
	(273.85,190.08) --
	(277.09,191.40) --
	(280.34,192.72) --
	(283.58,194.02);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 27.31, 59.60) --
	( 30.68, 53.77) --
	( 23.95, 53.77) --
	cycle;

\path[fill=fillColor] ( 35.86, 62.47) --
	( 39.22, 56.65) --
	( 32.49, 56.65) --
	cycle;

\path[fill=fillColor] ( 44.40, 65.07) --
	( 47.76, 59.25) --
	( 41.03, 59.25) --
	cycle;

\path[fill=fillColor] ( 52.94, 68.02) --
	( 56.30, 62.19) --
	( 49.58, 62.19) --
	cycle;

\path[fill=fillColor] ( 61.48, 70.77) --
	( 64.85, 64.95) --
	( 58.12, 64.95) --
	cycle;

\path[fill=fillColor] ( 70.02, 73.26) --
	( 73.39, 67.44) --
	( 66.66, 67.44) --
	cycle;

\path[fill=fillColor] ( 78.57, 75.65) --
	( 81.93, 69.83) --
	( 75.20, 69.83) --
	cycle;

\path[fill=fillColor] ( 87.11, 78.08) --
	( 90.47, 72.25) --
	( 83.75, 72.25) --
	cycle;

\path[fill=fillColor] ( 95.65, 80.50) --
	( 99.01, 74.67) --
	( 92.29, 74.67) --
	cycle;

\path[fill=fillColor] (104.19, 82.71) --
	(107.56, 76.88) --
	(100.83, 76.88) --
	cycle;

\path[fill=fillColor] (112.74, 84.81) --
	(116.10, 78.99) --
	(109.37, 78.99) --
	cycle;

\path[fill=fillColor] (121.28, 86.66) --
	(124.64, 80.84) --
	(117.91, 80.84) --
	cycle;

\path[fill=fillColor] (129.82, 88.60) --
	(133.18, 82.78) --
	(126.46, 82.78) --
	cycle;

\path[fill=fillColor] (138.36, 90.48) --
	(141.73, 84.65) --
	(135.00, 84.65) --
	cycle;

\path[fill=fillColor] (146.90, 92.35) --
	(150.27, 86.53) --
	(143.54, 86.53) --
	cycle;

\path[fill=fillColor] (155.45, 94.13) --
	(158.81, 88.30) --
	(152.08, 88.30) --
	cycle;

\path[fill=fillColor] (163.99, 95.79) --
	(167.35, 89.97) --
	(160.63, 89.97) --
	cycle;

\path[fill=fillColor] (172.53, 97.50) --
	(175.89, 91.67) --
	(169.17, 91.67) --
	cycle;

\path[fill=fillColor] (181.07, 99.31) --
	(184.44, 93.48) --
	(177.71, 93.48) --
	cycle;

\path[fill=fillColor] (189.62,100.91) --
	(192.98, 95.09) --
	(186.25, 95.09) --
	cycle;

\path[fill=fillColor] (198.16,102.55) --
	(201.52, 96.72) --
	(194.79, 96.72) --
	cycle;

\path[fill=fillColor] (206.70,104.08) --
	(210.06, 98.25) --
	(203.34, 98.25) --
	cycle;

\path[fill=fillColor] (215.24,105.58) --
	(218.61, 99.75) --
	(211.88, 99.75) --
	cycle;

\path[fill=fillColor] (223.78,107.13) --
	(227.15,101.30) --
	(220.42,101.30) --
	cycle;

\path[fill=fillColor] (232.33,108.61) --
	(235.69,102.78) --
	(228.96,102.78) --
	cycle;

\path[fill=fillColor] (240.87,109.99) --
	(244.23,104.16) --
	(237.51,104.16) --
	cycle;

\path[fill=fillColor] (249.41,111.47) --
	(252.77,105.64) --
	(246.05,105.64) --
	cycle;

\path[fill=fillColor] (257.95,112.80) --
	(261.32,106.97) --
	(254.59,106.97) --
	cycle;

\path[fill=fillColor] (266.50,114.25) --
	(269.86,108.43) --
	(263.13,108.43) --
	cycle;

\path[fill=fillColor] (275.04,115.51) --
	(278.40,109.68) --
	(271.67,109.68) --
	cycle;

\path[fill=fillColor] (283.58,116.83) --
	(286.94,111.01) --
	(280.22,111.01) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 24.82, 53.14) --
	( 29.81, 53.14) --
	( 29.81, 58.14) --
	( 24.82, 58.14) --
	cycle;

\path[fill=fillColor] ( 33.36, 57.75) --
	( 38.35, 57.75) --
	( 38.35, 62.75) --
	( 33.36, 62.75) --
	cycle;

\path[fill=fillColor] ( 41.90, 61.97) --
	( 46.90, 61.97) --
	( 46.90, 66.96) --
	( 41.90, 66.96) --
	cycle;

\path[fill=fillColor] ( 50.44, 66.76) --
	( 55.44, 66.76) --
	( 55.44, 71.75) --
	( 50.44, 71.75) --
	cycle;

\path[fill=fillColor] ( 58.98, 71.20) --
	( 63.98, 71.20) --
	( 63.98, 76.19) --
	( 58.98, 76.19) --
	cycle;

\path[fill=fillColor] ( 67.53, 75.33) --
	( 72.52, 75.33) --
	( 72.52, 80.32) --
	( 67.53, 80.32) --
	cycle;

\path[fill=fillColor] ( 76.07, 79.35) --
	( 81.06, 79.35) --
	( 81.06, 84.34) --
	( 76.07, 84.34) --
	cycle;

\path[fill=fillColor] ( 84.61, 83.44) --
	( 89.61, 83.44) --
	( 89.61, 88.43) --
	( 84.61, 88.43) --
	cycle;

\path[fill=fillColor] ( 93.15, 87.44) --
	( 98.15, 87.44) --
	( 98.15, 92.44) --
	( 93.15, 92.44) --
	cycle;

\path[fill=fillColor] (101.70, 91.12) --
	(106.69, 91.12) --
	(106.69, 96.12) --
	(101.70, 96.12) --
	cycle;

\path[fill=fillColor] (110.24, 94.51) --
	(115.23, 94.51) --
	(115.23, 99.50) --
	(110.24, 99.50) --
	cycle;

\path[fill=fillColor] (118.78, 97.80) --
	(123.78, 97.80) --
	(123.78,102.79) --
	(118.78,102.79) --
	cycle;

\path[fill=fillColor] (127.32,101.03) --
	(132.32,101.03) --
	(132.32,106.02) --
	(127.32,106.02) --
	cycle;

\path[fill=fillColor] (135.86,104.29) --
	(140.86,104.29) --
	(140.86,109.28) --
	(135.86,109.28) --
	cycle;

\path[fill=fillColor] (144.41,107.42) --
	(149.40,107.42) --
	(149.40,112.41) --
	(144.41,112.41) --
	cycle;

\path[fill=fillColor] (152.95,110.55) --
	(157.94,110.55) --
	(157.94,115.55) --
	(152.95,115.55) --
	cycle;

\path[fill=fillColor] (161.49,113.41) --
	(166.49,113.41) --
	(166.49,118.40) --
	(161.49,118.40) --
	cycle;

\path[fill=fillColor] (170.03,116.27) --
	(175.03,116.27) --
	(175.03,121.26) --
	(170.03,121.26) --
	cycle;

\path[fill=fillColor] (178.58,119.47) --
	(183.57,119.47) --
	(183.57,124.47) --
	(178.58,124.47) --
	cycle;

\path[fill=fillColor] (187.12,122.28) --
	(192.11,122.28) --
	(192.11,127.28) --
	(187.12,127.28) --
	cycle;

\path[fill=fillColor] (195.66,125.17) --
	(200.66,125.17) --
	(200.66,130.16) --
	(195.66,130.16) --
	cycle;

\path[fill=fillColor] (204.20,127.80) --
	(209.20,127.80) --
	(209.20,132.80) --
	(204.20,132.80) --
	cycle;

\path[fill=fillColor] (212.74,130.49) --
	(217.74,130.49) --
	(217.74,135.48) --
	(212.74,135.48) --
	cycle;

\path[fill=fillColor] (221.29,133.19) --
	(226.28,133.19) --
	(226.28,138.18) --
	(221.29,138.18) --
	cycle;

\path[fill=fillColor] (229.83,135.74) --
	(234.82,135.74) --
	(234.82,140.73) --
	(229.83,140.73) --
	cycle;

\path[fill=fillColor] (238.37,138.17) --
	(243.37,138.17) --
	(243.37,143.16) --
	(238.37,143.16) --
	cycle;

\path[fill=fillColor] (246.91,140.85) --
	(251.91,140.85) --
	(251.91,145.85) --
	(246.91,145.85) --
	cycle;

\path[fill=fillColor] (255.46,143.24) --
	(260.45,143.24) --
	(260.45,148.23) --
	(255.46,148.23) --
	cycle;

\path[fill=fillColor] (264.00,145.71) --
	(268.99,145.71) --
	(268.99,150.71) --
	(264.00,150.71) --
	cycle;

\path[fill=fillColor] (272.54,147.90) --
	(277.54,147.90) --
	(277.54,152.89) --
	(272.54,152.89) --
	cycle;

\path[fill=fillColor] (281.08,150.34) --
	(286.08,150.34) --
	(286.08,155.34) --
	(281.08,155.34) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 23.78, 55.41) -- ( 30.85, 55.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.31, 51.87) -- ( 27.31, 58.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.32, 61.62) -- ( 39.39, 61.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.86, 58.09) -- ( 35.86, 65.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87, 67.34) -- ( 47.93, 67.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.40, 63.81) -- ( 44.40, 70.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.41, 73.88) -- ( 56.47, 73.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.94, 70.35) -- ( 52.94, 77.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.95, 79.89) -- ( 65.01, 79.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.48, 76.36) -- ( 61.48, 83.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.49, 85.65) -- ( 73.56, 85.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.02, 82.12) -- ( 70.02, 89.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.03, 91.28) -- ( 82.10, 91.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.57, 87.75) -- ( 78.57, 94.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.58, 97.06) -- ( 90.64, 97.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.11, 93.53) -- ( 87.11,100.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.12,102.60) -- ( 99.18,102.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65, 99.07) -- ( 95.65,106.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.66,107.76) -- (107.73,107.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.19,104.23) -- (104.19,111.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.20,112.34) -- (116.27,112.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.74,108.81) -- (112.74,115.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.75,117.19) -- (124.81,117.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.28,113.66) -- (121.28,120.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.29,121.71) -- (133.35,121.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.82,118.17) -- (129.82,125.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.83,126.41) -- (141.89,126.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.36,122.88) -- (138.36,129.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.37,130.85) -- (150.44,130.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.90,127.32) -- (146.90,134.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.91,135.49) -- (158.98,135.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (155.45,131.96) -- (155.45,139.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.46,139.55) -- (167.52,139.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.99,136.02) -- (163.99,143.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.00,143.62) -- (176.06,143.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.53,140.09) -- (172.53,147.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.54,148.33) -- (184.61,148.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.07,144.80) -- (181.07,151.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.08,152.45) -- (193.15,152.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.62,148.92) -- (189.62,155.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.63,156.73) -- (201.69,156.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.16,153.19) -- (198.16,160.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.17,160.57) -- (210.23,160.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.70,157.04) -- (206.70,164.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.71,164.54) -- (218.77,164.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.24,161.01) -- (215.24,168.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.25,168.54) -- (227.32,168.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.78,165.01) -- (223.78,172.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.79,172.25) -- (235.86,172.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.33,168.71) -- (232.33,175.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.34,175.81) -- (244.40,175.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.87,172.28) -- (240.87,179.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.88,179.95) -- (252.94,179.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.41,176.42) -- (249.41,183.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.42,183.53) -- (261.49,183.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.95,179.99) -- (257.95,187.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.96,187.15) -- (270.03,187.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (266.50,183.62) -- (266.50,190.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.51,190.32) -- (278.57,190.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.04,186.78) -- (275.04,193.85);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.05,194.15) -- (287.11,194.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.58,190.62) -- (283.58,197.68);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 27.31, 54.98) circle (  2.50);

\path[fill=fillColor] ( 35.86, 54.98) circle (  2.50);

\path[fill=fillColor] ( 44.40, 54.98) circle (  2.50);

\path[fill=fillColor] ( 52.94, 54.98) circle (  2.50);

\path[fill=fillColor] ( 61.48, 54.98) circle (  2.50);

\path[fill=fillColor] ( 70.02, 54.98) circle (  2.50);

\path[fill=fillColor] ( 78.57, 54.98) circle (  2.50);

\path[fill=fillColor] ( 87.11, 54.98) circle (  2.50);

\path[fill=fillColor] ( 95.65, 54.98) circle (  2.50);

\path[fill=fillColor] (104.19, 54.98) circle (  2.50);

\path[fill=fillColor] (112.74, 54.98) circle (  2.50);

\path[fill=fillColor] (121.28, 54.98) circle (  2.50);

\path[fill=fillColor] (129.82, 54.98) circle (  2.50);

\path[fill=fillColor] (138.36, 54.98) circle (  2.50);

\path[fill=fillColor] (146.90, 54.98) circle (  2.50);

\path[fill=fillColor] (155.45, 54.98) circle (  2.50);

\path[fill=fillColor] (163.99, 54.98) circle (  2.50);

\path[fill=fillColor] (172.53, 54.98) circle (  2.50);

\path[fill=fillColor] (181.07, 54.98) circle (  2.50);

\path[fill=fillColor] (189.62, 54.98) circle (  2.50);

\path[fill=fillColor] (198.16, 54.98) circle (  2.50);

\path[fill=fillColor] (206.70, 54.98) circle (  2.50);

\path[fill=fillColor] (215.24, 54.98) circle (  2.50);

\path[fill=fillColor] (223.78, 54.98) circle (  2.50);

\path[fill=fillColor] (232.33, 54.98) circle (  2.50);

\path[fill=fillColor] (240.87, 54.98) circle (  2.50);

\path[fill=fillColor] (249.41, 54.98) circle (  2.50);

\path[fill=fillColor] (257.95, 54.98) circle (  2.50);

\path[fill=fillColor] (266.50, 54.98) circle (  2.50);

\path[fill=fillColor] (275.04, 54.98) circle (  2.50);

\path[fill=fillColor] (283.58, 54.98) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 30.69) rectangle (283.58,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 52.22) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,134.09) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 55.25) --
	( 27.31, 55.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,137.12) --
	( 27.31,137.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 27.94) --
	( 27.31, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.74, 27.94) --
	(112.74, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.16, 27.94) --
	(198.16, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.74, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.16, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (155.45,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,112.42) {Mean interval (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,202.22) {C};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,216.81);
\end{scope}
\begin{scope}
\path[clip] (323.24, 30.69) rectangle (572.66,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (323.24, 30.69) rectangle (572.66,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (323.24, 58.94) --
	(326.39, 58.94) --
	(329.55, 58.94) --
	(332.71, 58.94) --
	(335.87, 58.94) --
	(339.02, 58.94) --
	(342.18, 58.94) --
	(345.34, 58.94) --
	(348.49, 58.94) --
	(351.65, 58.94) --
	(354.81, 58.94) --
	(357.97, 58.94) --
	(361.12, 58.94) --
	(364.28, 58.94) --
	(367.44, 58.94) --
	(370.60, 58.94) --
	(373.75, 58.94) --
	(376.91, 58.94) --
	(380.07, 58.94) --
	(383.22, 58.94) --
	(386.38, 58.94) --
	(389.54, 58.94) --
	(392.70, 58.94) --
	(395.85, 58.94) --
	(399.01, 58.94) --
	(402.17, 58.94) --
	(405.32, 58.94) --
	(408.48, 58.94) --
	(411.64, 58.94) --
	(414.80, 58.94) --
	(417.95, 58.94) --
	(421.11, 58.94) --
	(424.27, 58.94) --
	(427.43, 58.94) --
	(430.58, 58.94) --
	(433.74, 58.94) --
	(436.90, 58.94) --
	(440.05, 58.94) --
	(443.21, 58.94) --
	(446.37, 58.94) --
	(449.53, 58.94) --
	(452.68, 58.94) --
	(455.84, 58.94) --
	(459.00, 58.94) --
	(462.16, 58.94) --
	(465.31, 58.94) --
	(468.47, 58.94) --
	(471.63, 58.94) --
	(474.78, 58.94) --
	(477.94, 58.94) --
	(481.10, 58.94) --
	(484.26, 58.94) --
	(487.41, 58.94) --
	(490.57, 58.94) --
	(493.73, 58.94) --
	(496.89, 58.94) --
	(500.04, 58.94) --
	(503.20, 58.94) --
	(506.36, 58.94) --
	(509.51, 58.94) --
	(512.67, 58.94) --
	(515.83, 58.94) --
	(518.99, 58.94) --
	(522.14, 58.94) --
	(525.30, 58.94) --
	(528.46, 58.94) --
	(531.62, 58.94) --
	(534.77, 58.94) --
	(537.93, 58.94) --
	(541.09, 58.94) --
	(544.24, 58.94) --
	(547.40, 58.94) --
	(550.56, 58.94) --
	(553.72, 58.94) --
	(556.87, 58.94) --
	(560.03, 58.94) --
	(563.19, 58.94) --
	(566.35, 58.94) --
	(569.50, 58.94) --
	(572.66, 58.94);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (323.24,115.70) --
	(326.39,114.93) --
	(329.55,114.17) --
	(332.71,113.43) --
	(335.87,112.70) --
	(339.02,111.98) --
	(342.18,111.27) --
	(345.34,110.58) --
	(348.49,109.90) --
	(351.65,109.23) --
	(354.81,108.58) --
	(357.97,107.94) --
	(361.12,107.32) --
	(364.28,106.70) --
	(367.44,106.10) --
	(370.60,105.51) --
	(373.75,104.94) --
	(376.91,104.37) --
	(380.07,103.82) --
	(383.22,103.28) --
	(386.38,102.76) --
	(389.54,102.25) --
	(392.70,101.76) --
	(395.85,101.28) --
	(399.01,100.81) --
	(402.17,100.36) --
	(405.32, 99.92) --
	(408.48, 99.48) --
	(411.64, 99.06) --
	(414.80, 98.64) --
	(417.95, 98.23) --
	(421.11, 97.85) --
	(424.27, 97.47) --
	(427.43, 97.11) --
	(430.58, 96.76) --
	(433.74, 96.42) --
	(436.90, 96.09) --
	(440.05, 95.75) --
	(443.21, 95.42) --
	(446.37, 95.09) --
	(449.53, 94.76) --
	(452.68, 94.43) --
	(455.84, 94.12) --
	(459.00, 93.81) --
	(462.16, 93.51) --
	(465.31, 93.21) --
	(468.47, 92.92) --
	(471.63, 92.63) --
	(474.78, 92.34) --
	(477.94, 92.06) --
	(481.10, 91.77) --
	(484.26, 91.49) --
	(487.41, 91.21) --
	(490.57, 90.94) --
	(493.73, 90.67) --
	(496.89, 90.40) --
	(500.04, 90.13) --
	(503.20, 89.87) --
	(506.36, 89.61) --
	(509.51, 89.36) --
	(512.67, 89.11) --
	(515.83, 88.86) --
	(518.99, 88.61) --
	(522.14, 88.37) --
	(525.30, 88.13) --
	(528.46, 87.90) --
	(531.62, 87.67) --
	(534.77, 87.44) --
	(537.93, 87.22) --
	(541.09, 87.00) --
	(544.24, 86.78) --
	(547.40, 86.56) --
	(550.56, 86.35) --
	(553.72, 86.14) --
	(556.87, 85.94) --
	(560.03, 85.74) --
	(563.19, 85.54) --
	(566.35, 85.35) --
	(569.50, 85.16) --
	(572.66, 84.97);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (323.24,149.67) --
	(326.39,148.19) --
	(329.55,146.73) --
	(332.71,145.31) --
	(335.87,143.91) --
	(339.02,142.53) --
	(342.18,141.18) --
	(345.34,139.86) --
	(348.49,138.57) --
	(351.65,137.30) --
	(354.81,136.06) --
	(357.97,134.85) --
	(361.12,133.66) --
	(364.28,132.50) --
	(367.44,131.37) --
	(370.60,130.26) --
	(373.75,129.17) --
	(376.91,128.11) --
	(380.07,127.07) --
	(383.22,126.06) --
	(386.38,125.09) --
	(389.54,124.14) --
	(392.70,123.23) --
	(395.85,122.35) --
	(399.01,121.49) --
	(402.17,120.66) --
	(405.32,119.84) --
	(408.48,119.05) --
	(411.64,118.27) --
	(414.80,117.51) --
	(417.95,116.77) --
	(421.11,116.06) --
	(424.27,115.39) --
	(427.43,114.74) --
	(430.58,114.12) --
	(433.74,113.51) --
	(436.90,112.91) --
	(440.05,112.32) --
	(443.21,111.73) --
	(446.37,111.14) --
	(449.53,110.55) --
	(452.68,109.97) --
	(455.84,109.41) --
	(459.00,108.87) --
	(462.16,108.33) --
	(465.31,107.81) --
	(468.47,107.30) --
	(471.63,106.80) --
	(474.78,106.29) --
	(477.94,105.80) --
	(481.10,105.30) --
	(484.26,104.81) --
	(487.41,104.32) --
	(490.57,103.84) --
	(493.73,103.37) --
	(496.89,102.91) --
	(500.04,102.46) --
	(503.20,102.01) --
	(506.36,101.57) --
	(509.51,101.13) --
	(512.67,100.70) --
	(515.83,100.27) --
	(518.99, 99.86) --
	(522.14, 99.44) --
	(525.30, 99.04) --
	(528.46, 98.64) --
	(531.62, 98.25) --
	(534.77, 97.86) --
	(537.93, 97.48) --
	(541.09, 97.11) --
	(544.24, 96.75) --
	(547.40, 96.39) --
	(550.56, 96.04) --
	(553.72, 95.69) --
	(556.87, 95.35) --
	(560.03, 95.02) --
	(563.19, 94.69) --
	(566.35, 94.37) --
	(569.50, 94.06) --
	(572.66, 93.75);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (323.24,180.94) --
	(326.39,178.70) --
	(329.55,176.51) --
	(332.71,174.35) --
	(335.87,172.23) --
	(339.02,170.16) --
	(342.18,168.13) --
	(345.34,166.14) --
	(348.49,164.20) --
	(351.65,162.29) --
	(354.81,160.43) --
	(357.97,158.62) --
	(361.12,156.84) --
	(364.28,155.11) --
	(367.44,153.41) --
	(370.60,151.76) --
	(373.75,150.14) --
	(376.91,148.56) --
	(380.07,147.02) --
	(383.22,145.52) --
	(386.38,144.08) --
	(389.54,142.68) --
	(392.70,141.34) --
	(395.85,140.05) --
	(399.01,138.79) --
	(402.17,137.57) --
	(405.32,136.38) --
	(408.48,135.23) --
	(411.64,134.09) --
	(414.80,132.98) --
	(417.95,131.90) --
	(421.11,130.89) --
	(424.27,129.91) --
	(427.43,128.98) --
	(430.58,128.09) --
	(433.74,127.21) --
	(436.90,126.36) --
	(440.05,125.52) --
	(443.21,124.68) --
	(446.37,123.84) --
	(449.53,122.99) --
	(452.68,122.16) --
	(455.84,121.37) --
	(459.00,120.59) --
	(462.16,119.84) --
	(465.31,119.10) --
	(468.47,118.37) --
	(471.63,117.66) --
	(474.78,116.95) --
	(477.94,116.25) --
	(481.10,115.55) --
	(484.26,114.85) --
	(487.41,114.17) --
	(490.57,113.50) --
	(493.73,112.84) --
	(496.89,112.20) --
	(500.04,111.56) --
	(503.20,110.94) --
	(506.36,110.32) --
	(509.51,109.72) --
	(512.67,109.12) --
	(515.83,108.53) --
	(518.99,107.95) --
	(522.14,107.38) --
	(525.30,106.82) --
	(528.46,106.27) --
	(531.62,105.73) --
	(534.77,105.20) --
	(537.93,104.68) --
	(541.09,104.17) --
	(544.24,103.68) --
	(547.40,103.19) --
	(550.56,102.71) --
	(553.72,102.24) --
	(556.87,101.78) --
	(560.03,101.33) --
	(563.19,100.89) --
	(566.35,100.46) --
	(569.50,100.05) --
	(572.66, 99.64);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] (323.24,119.77) --
	(326.60,113.94) --
	(319.87,113.94) --
	cycle;

\path[fill=fillColor] (331.55,117.61) --
	(334.91,111.79) --
	(328.19,111.79) --
	cycle;

\path[fill=fillColor] (339.86,115.71) --
	(343.23,109.89) --
	(336.50,109.89) --
	cycle;

\path[fill=fillColor] (348.18,113.71) --
	(351.54,107.88) --
	(344.81,107.88) --
	cycle;

\path[fill=fillColor] (356.49,111.89) --
	(359.86,106.07) --
	(353.13,106.07) --
	cycle;

\path[fill=fillColor] (364.81,110.34) --
	(368.17,104.52) --
	(361.44,104.52) --
	cycle;

\path[fill=fillColor] (373.12,108.89) --
	(376.48,103.06) --
	(369.76,103.06) --
	cycle;

\path[fill=fillColor] (381.44,107.50) --
	(384.80,101.67) --
	(378.07,101.67) --
	cycle;

\path[fill=fillColor] (389.75,106.03) --
	(393.11,100.20) --
	(386.39,100.20) --
	cycle;

\path[fill=fillColor] (398.06,104.70) --
	(401.43, 98.88) --
	(394.70, 98.88) --
	cycle;

\path[fill=fillColor] (406.38,103.51) --
	(409.74, 97.69) --
	(403.01, 97.69) --
	cycle;

\path[fill=fillColor] (414.69,102.49) --
	(418.06, 96.67) --
	(411.33, 96.67) --
	cycle;

\path[fill=fillColor] (423.01,101.49) --
	(426.37, 95.66) --
	(419.64, 95.66) --
	cycle;

\path[fill=fillColor] (431.32,100.59) --
	(434.68, 94.76) --
	(427.96, 94.76) --
	cycle;

\path[fill=fillColor] (439.63, 99.70) --
	(443.00, 93.88) --
	(436.27, 93.88) --
	cycle;

\path[fill=fillColor] (447.95, 98.87) --
	(451.31, 93.04) --
	(444.58, 93.04) --
	cycle;

\path[fill=fillColor] (456.26, 98.10) --
	(459.63, 92.27) --
	(452.90, 92.27) --
	cycle;

\path[fill=fillColor] (464.58, 97.28) --
	(467.94, 91.45) --
	(461.21, 91.45) --
	cycle;

\path[fill=fillColor] (472.89, 96.38) --
	(476.25, 90.55) --
	(469.53, 90.55) --
	cycle;

\path[fill=fillColor] (481.20, 95.60) --
	(484.57, 89.77) --
	(477.84, 89.77) --
	cycle;

\path[fill=fillColor] (489.52, 94.90) --
	(492.88, 89.07) --
	(486.15, 89.07) --
	cycle;

\path[fill=fillColor] (497.83, 94.18) --
	(501.20, 88.35) --
	(494.47, 88.35) --
	cycle;

\path[fill=fillColor] (506.15, 93.53) --
	(509.51, 87.70) --
	(502.78, 87.70) --
	cycle;

\path[fill=fillColor] (514.46, 92.81) --
	(517.82, 86.99) --
	(511.10, 86.99) --
	cycle;

\path[fill=fillColor] (522.78, 92.22) --
	(526.14, 86.39) --
	(519.41, 86.39) --
	cycle;

\path[fill=fillColor] (531.09, 91.66) --
	(534.45, 85.83) --
	(527.73, 85.83) --
	cycle;

\path[fill=fillColor] (539.40, 91.06) --
	(542.77, 85.24) --
	(536.04, 85.24) --
	cycle;

\path[fill=fillColor] (547.72, 90.45) --
	(551.08, 84.63) --
	(544.35, 84.63) --
	cycle;

\path[fill=fillColor] (556.03, 89.86) --
	(559.40, 84.03) --
	(552.67, 84.03) --
	cycle;

\path[fill=fillColor] (564.35, 89.36) --
	(567.71, 83.53) --
	(560.98, 83.53) --
	cycle;

\path[fill=fillColor] (572.66, 88.81) --
	(576.02, 82.98) --
	(569.30, 82.98) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] (320.74,147.59) --
	(325.73,147.59) --
	(325.73,152.58) --
	(320.74,152.58) --
	cycle;

\path[fill=fillColor] (329.05,143.39) --
	(334.05,143.39) --
	(334.05,148.38) --
	(329.05,148.38) --
	cycle;

\path[fill=fillColor] (337.37,139.76) --
	(342.36,139.76) --
	(342.36,144.75) --
	(337.37,144.75) --
	cycle;

\path[fill=fillColor] (345.68,135.95) --
	(350.68,135.95) --
	(350.68,140.94) --
	(345.68,140.94) --
	cycle;

\path[fill=fillColor] (353.99,132.41) --
	(358.99,132.41) --
	(358.99,137.40) --
	(353.99,137.40) --
	cycle;

\path[fill=fillColor] (362.31,129.50) --
	(367.30,129.50) --
	(367.30,134.50) --
	(362.31,134.50) --
	cycle;

\path[fill=fillColor] (370.62,126.75) --
	(375.62,126.75) --
	(375.62,131.74) --
	(370.62,131.74) --
	cycle;

\path[fill=fillColor] (378.94,124.08) --
	(383.93,124.08) --
	(383.93,129.08) --
	(378.94,129.08) --
	cycle;

\path[fill=fillColor] (387.25,121.44) --
	(392.25,121.44) --
	(392.25,126.43) --
	(387.25,126.43) --
	cycle;

\path[fill=fillColor] (395.57,118.99) --
	(400.56,118.99) --
	(400.56,123.99) --
	(395.57,123.99) --
	cycle;

\path[fill=fillColor] (403.88,116.78) --
	(408.88,116.78) --
	(408.88,121.77) --
	(403.88,121.77) --
	cycle;

\path[fill=fillColor] (412.19,115.02) --
	(417.19,115.02) --
	(417.19,120.02) --
	(412.19,120.02) --
	cycle;

\path[fill=fillColor] (420.51,113.20) --
	(425.50,113.20) --
	(425.50,118.19) --
	(420.51,118.19) --
	cycle;

\path[fill=fillColor] (428.82,111.53) --
	(433.82,111.53) --
	(433.82,116.52) --
	(428.82,116.52) --
	cycle;

\path[fill=fillColor] (437.14,109.90) --
	(442.13,109.90) --
	(442.13,114.89) --
	(437.14,114.89) --
	cycle;

\path[fill=fillColor] (445.45,108.47) --
	(450.45,108.47) --
	(450.45,113.47) --
	(445.45,113.47) --
	cycle;

\path[fill=fillColor] (453.76,107.09) --
	(458.76,107.09) --
	(458.76,112.08) --
	(453.76,112.08) --
	cycle;

\path[fill=fillColor] (462.08,105.59) --
	(467.07,105.59) --
	(467.07,110.58) --
	(462.08,110.58) --
	cycle;

\path[fill=fillColor] (470.39,104.07) --
	(475.39,104.07) --
	(475.39,109.07) --
	(470.39,109.07) --
	cycle;

\path[fill=fillColor] (478.71,102.72) --
	(483.70,102.72) --
	(483.70,107.72) --
	(478.71,107.72) --
	cycle;

\path[fill=fillColor] (487.02,101.49) --
	(492.02,101.49) --
	(492.02,106.49) --
	(487.02,106.49) --
	cycle;

\path[fill=fillColor] (495.34,100.25) --
	(500.33,100.25) --
	(500.33,105.25) --
	(495.34,105.25) --
	cycle;

\path[fill=fillColor] (503.65, 99.11) --
	(508.64, 99.11) --
	(508.64,104.10) --
	(503.65,104.10) --
	cycle;

\path[fill=fillColor] (511.96, 97.90) --
	(516.96, 97.90) --
	(516.96,102.90) --
	(511.96,102.90) --
	cycle;

\path[fill=fillColor] (520.28, 96.90) --
	(525.27, 96.90) --
	(525.27,101.90) --
	(520.28,101.90) --
	cycle;

\path[fill=fillColor] (528.59, 95.91) --
	(533.59, 95.91) --
	(533.59,100.91) --
	(528.59,100.91) --
	cycle;

\path[fill=fillColor] (536.91, 94.94) --
	(541.90, 94.94) --
	(541.90, 99.93) --
	(536.91, 99.93) --
	cycle;

\path[fill=fillColor] (545.22, 94.00) --
	(550.22, 94.00) --
	(550.22, 98.99) --
	(545.22, 98.99) --
	cycle;

\path[fill=fillColor] (553.53, 92.89) --
	(558.53, 92.89) --
	(558.53, 97.89) --
	(553.53, 97.89) --
	cycle;

\path[fill=fillColor] (561.85, 92.04) --
	(566.84, 92.04) --
	(566.84, 97.04) --
	(561.85, 97.04) --
	cycle;

\path[fill=fillColor] (570.16, 91.16) --
	(575.16, 91.16) --
	(575.16, 96.15) --
	(570.16, 96.15) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (319.70,181.63) -- (326.77,181.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.24,178.10) -- (323.24,185.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.02,175.24) -- (335.08,175.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (331.55,171.71) -- (331.55,178.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (336.33,169.74) -- (343.40,169.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (339.86,166.20) -- (339.86,173.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (344.65,163.98) -- (351.71,163.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (348.18,160.45) -- (348.18,167.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (352.96,158.64) -- (360.02,158.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.49,155.11) -- (356.49,162.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.27,154.33) -- (368.34,154.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.81,150.80) -- (364.81,157.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.59,150.19) -- (376.65,150.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (373.12,146.66) -- (373.12,153.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.90,146.18) -- (384.97,146.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.44,142.65) -- (381.44,149.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.22,142.37) -- (393.28,142.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.75,138.84) -- (389.75,145.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (394.53,138.78) -- (401.60,138.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.06,135.25) -- (398.06,142.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (402.85,135.48) -- (409.91,135.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.38,131.95) -- (406.38,139.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (411.16,133.07) -- (418.22,133.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (414.69,129.54) -- (414.69,136.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.47,130.41) -- (426.54,130.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.01,126.88) -- (423.01,133.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.79,127.94) -- (434.85,127.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.32,124.40) -- (431.32,131.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.10,125.60) -- (443.17,125.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (439.63,122.07) -- (439.63,129.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.42,123.66) -- (451.48,123.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (447.95,120.13) -- (447.95,127.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.73,121.64) -- (459.79,121.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.26,118.11) -- (456.26,125.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.04,119.43) -- (468.11,119.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.58,115.90) -- (464.58,122.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.36,117.33) -- (476.42,117.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (472.89,113.80) -- (472.89,120.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.67,115.45) -- (484.74,115.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.20,111.92) -- (481.20,118.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.99,113.73) -- (493.05,113.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (489.52,110.20) -- (489.52,117.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.30,112.01) -- (501.36,112.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (497.83,108.48) -- (497.83,115.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.61,110.34) -- (509.68,110.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (506.15,106.81) -- (506.15,113.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.93,108.72) -- (517.99,108.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (514.46,105.19) -- (514.46,112.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.24,107.30) -- (526.31,107.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.78,103.77) -- (522.78,110.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (527.56,105.93) -- (534.62,105.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (531.09,102.40) -- (531.09,109.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.87,104.65) -- (542.94,104.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (539.40,101.12) -- (539.40,108.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.19,103.44) -- (551.25,103.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.72, 99.91) -- (547.72,106.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.50,101.84) -- (559.56,101.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (556.03, 98.31) -- (556.03,105.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.81,100.61) -- (567.88,100.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.35, 97.07) -- (564.35,104.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13, 99.50) -- (576.19, 99.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66, 95.97) -- (572.66,103.04);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (323.24, 58.94) circle (  2.50);

\path[fill=fillColor] (331.55, 58.94) circle (  2.50);

\path[fill=fillColor] (339.86, 58.94) circle (  2.50);

\path[fill=fillColor] (348.18, 58.94) circle (  2.50);

\path[fill=fillColor] (356.49, 58.94) circle (  2.50);

\path[fill=fillColor] (364.81, 58.94) circle (  2.50);

\path[fill=fillColor] (373.12, 58.94) circle (  2.50);

\path[fill=fillColor] (381.44, 58.94) circle (  2.50);

\path[fill=fillColor] (389.75, 58.94) circle (  2.50);

\path[fill=fillColor] (398.06, 58.94) circle (  2.50);

\path[fill=fillColor] (406.38, 58.94) circle (  2.50);

\path[fill=fillColor] (414.69, 58.94) circle (  2.50);

\path[fill=fillColor] (423.01, 58.94) circle (  2.50);

\path[fill=fillColor] (431.32, 58.94) circle (  2.50);

\path[fill=fillColor] (439.63, 58.94) circle (  2.50);

\path[fill=fillColor] (447.95, 58.94) circle (  2.50);

\path[fill=fillColor] (456.26, 58.94) circle (  2.50);

\path[fill=fillColor] (464.58, 58.94) circle (  2.50);

\path[fill=fillColor] (472.89, 58.94) circle (  2.50);

\path[fill=fillColor] (481.20, 58.94) circle (  2.50);

\path[fill=fillColor] (489.52, 58.94) circle (  2.50);

\path[fill=fillColor] (497.83, 58.94) circle (  2.50);

\path[fill=fillColor] (506.15, 58.94) circle (  2.50);

\path[fill=fillColor] (514.46, 58.94) circle (  2.50);

\path[fill=fillColor] (522.78, 58.94) circle (  2.50);

\path[fill=fillColor] (531.09, 58.94) circle (  2.50);

\path[fill=fillColor] (539.40, 58.94) circle (  2.50);

\path[fill=fillColor] (547.72, 58.94) circle (  2.50);

\path[fill=fillColor] (556.03, 58.94) circle (  2.50);

\path[fill=fillColor] (564.35, 58.94) circle (  2.50);

\path[fill=fillColor] (572.66, 58.94) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.24, 30.69) rectangle (572.66,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29, 27.66) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29, 66.58) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,105.50) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,144.42) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,183.34) {0.8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49, 30.69) --
	(323.24, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49, 69.61) --
	(323.24, 69.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,108.53) --
	(323.24,108.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,147.45) --
	(323.24,147.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,186.37) --
	(323.24,186.37);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.24, 27.94) --
	(323.24, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.38, 27.94) --
	(406.38, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (489.52, 27.94) --
	(489.52, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (323.24, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (406.38, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.52, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (447.95,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,112.42) {Squared coefficient of variation};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (323.24,202.22) {D};
\end{scope}
\end{tikzpicture}

\end{document}