### 2020-04-29 observedrRb.tex

% Created by tikzDevice version 0.12.3 on 2020-04-29 11:23:36
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (476.98,238.49);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (238.49,238.49);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (238.49,238.49);
\end{scope}
\begin{scope}
\path[clip] ( 49.68, 44.63) rectangle (230.49,230.49);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 49.68, 44.63) rectangle (230.49,230.49);
\definecolor{drawColor}{RGB}{204,121,167}
\definecolor{fillColor}{RGB}{204,121,167}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] ( 72.28,208.07) --
	( 76.80,164.32) --
	( 81.32,167.89) --
	( 85.84,158.27) --
	( 90.36,161.18) --
	( 94.88,155.47) --
	( 99.40,155.84) --
	(103.92,152.83) --
	(108.44,147.52) --
	(112.96,144.06) --
	(117.49,141.75) --
	(122.01,135.58) --
	(126.53,129.99) --
	(131.05,127.94) --
	(135.57,121.76) --
	(140.09,114.68) --
	(144.61,107.88) --
	(149.13,100.74) --
	(153.65, 92.87) --
	(158.17, 84.99) --
	(162.69, 79.17) --
	(167.21, 73.72) --
	(171.73, 70.27) --
	(176.25, 67.93) --
	(180.77, 66.45) --
	(185.29, 65.22) --
	(189.81, 64.69) --
	(194.33, 64.39) --
	(198.85, 63.99) --
	(203.37, 63.64) --
	(207.89, 63.40) --
	(212.41, 63.31) --
	(216.93, 63.30) --
	(221.45, 63.27) --
	(225.97, 63.25) --
	(230.49, 63.23) --
	(230.49, 59.98) --
	(225.97, 59.99) --
	(221.45, 59.99) --
	(216.93, 59.99) --
	(212.41, 60.01) --
	(207.89, 60.06) --
	(203.37, 60.28) --
	(198.85, 60.37) --
	(194.33, 60.61) --
	(189.81, 60.97) --
	(185.29, 61.22) --
	(180.77, 61.94) --
	(176.25, 63.28) --
	(171.73, 64.96) --
	(167.21, 67.32) --
	(162.69, 70.74) --
	(158.17, 75.21) --
	(153.65, 80.74) --
	(149.13, 86.75) --
	(144.61, 93.83) --
	(140.09,101.11) --
	(135.57,107.97) --
	(131.05,114.84) --
	(126.53,119.84) --
	(122.01,121.79) --
	(117.49,123.07) --
	(112.96,127.78) --
	(108.44,125.50) --
	(103.92,127.66) --
	( 99.40,118.76) --
	( 94.88,120.68) --
	( 90.36,112.09) --
	( 85.84,111.94) --
	( 81.32,109.08) --
	( 76.80,127.52) --
	( 72.28,125.38) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 72.28,166.88) --
	( 76.80,144.70) --
	( 81.32,138.94) --
	( 85.84,139.23) --
	( 90.36,139.42) --
	( 94.88,138.52) --
	( 99.40,138.58) --
	(103.92,136.90) --
	(108.44,135.32) --
	(112.96,134.23) --
	(117.49,130.80) --
	(122.01,129.38) --
	(126.53,125.23) --
	(131.05,120.57) --
	(135.57,114.75) --
	(140.09,107.84) --
	(144.61,100.77) --
	(149.13, 93.23) --
	(153.65, 86.29) --
	(158.17, 80.06) --
	(162.69, 75.18) --
	(167.21, 71.07) --
	(171.73, 67.95) --
	(176.25, 65.68) --
	(180.77, 64.26) --
	(185.29, 63.27) --
	(189.81, 62.65) --
	(194.33, 62.22) --
	(198.85, 61.93) --
	(203.37, 61.76) --
	(207.89, 61.63) --
	(212.41, 61.56) --
	(216.93, 61.51) --
	(221.45, 61.49) --
	(225.97, 61.48) --
	(230.49, 61.47);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 49.68,140.72) -- (230.49,140.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.68, 80.10) -- (230.49, 80.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (225.97,143.66) {Initial forward};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (225.97, 83.04) {Intrinsic};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] ( 49.68, 44.63) rectangle (230.49,230.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 40.23) {4.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 74.64) {5.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,109.06) {5.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,143.48) {6.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,177.90) {6.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,212.32) {7.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 44.63) --
	( 49.68, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 79.05) --
	( 49.68, 79.05);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,113.47) --
	( 49.68,113.47);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,147.89) --
	( 49.68,147.89);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,182.31) --
	( 49.68,182.31);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,216.72) --
	( 49.68,216.72);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 40.63) --
	( 49.68, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 94.88, 40.63) --
	( 94.88, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (140.09, 40.63) --
	(140.09, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (185.29, 40.63) --
	(185.29, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (230.49, 40.63) --
	(230.49, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 49.68, 28.62) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 94.88, 28.62) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (140.09, 28.62) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (185.29, 28.62) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (230.49, 28.62) {80};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (140.09, 11.11) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at ( 19.02,137.56) {Mean serial interval (days)};
\end{scope}
\begin{scope}
\path[clip] (238.49,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (238.49,  0.00) rectangle (476.98,238.49);
\end{scope}
\begin{scope}
\path[clip] (278.22, 44.63) rectangle (468.98,230.49);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (278.22, 44.63) rectangle (468.98,230.49);
\definecolor{drawColor}{RGB}{204,121,167}
\definecolor{fillColor}{RGB}{204,121,167}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (302.06,220.91) --
	(306.83,186.06) --
	(311.60,184.29) --
	(316.37,180.56) --
	(321.14,171.77) --
	(325.91,171.49) --
	(330.68,168.64) --
	(335.45,166.60) --
	(340.22,160.21) --
	(344.99,158.84) --
	(349.76,159.16) --
	(354.52,154.78) --
	(359.29,150.97) --
	(364.06,149.69) --
	(368.83,146.44) --
	(373.60,141.11) --
	(378.37,135.50) --
	(383.14,131.30) --
	(387.91,126.00) --
	(392.68,119.60) --
	(397.45,115.57) --
	(402.22,111.47) --
	(406.98,107.65) --
	(411.75,106.09) --
	(416.52,104.82) --
	(421.29,103.30) --
	(426.06,102.79) --
	(430.83,102.24) --
	(435.60,101.92) --
	(440.37,101.04) --
	(445.14,100.74) --
	(449.91,100.71) --
	(454.67,100.71) --
	(459.44,100.71) --
	(464.21,100.71) --
	(468.98,100.70) --
	(468.98, 97.00) --
	(464.21, 97.00) --
	(459.44, 97.00) --
	(454.67, 97.00) --
	(449.91, 97.02) --
	(445.14, 97.05) --
	(440.37, 97.67) --
	(435.60, 97.80) --
	(430.83, 98.05) --
	(426.06, 98.33) --
	(421.29, 98.61) --
	(416.52, 99.34) --
	(411.75,100.93) --
	(406.98,102.92) --
	(402.22,105.37) --
	(397.45,108.32) --
	(392.68,112.30) --
	(387.91,116.70) --
	(383.14,122.38) --
	(378.37,127.38) --
	(373.60,131.82) --
	(368.83,137.05) --
	(364.06,140.82) --
	(359.29,143.36) --
	(354.52,143.56) --
	(349.76,142.51) --
	(344.99,144.89) --
	(340.22,141.25) --
	(335.45,145.85) --
	(330.68,140.31) --
	(325.91,142.20) --
	(321.14,135.02) --
	(316.37,142.00) --
	(311.60,128.88) --
	(306.83,145.02) --
	(302.06,169.68) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.06,188.49) --
	(306.83,168.18) --
	(311.60,158.33) --
	(316.37,159.72) --
	(321.14,157.07) --
	(325.91,156.54) --
	(330.68,156.67) --
	(335.45,155.15) --
	(340.22,152.92) --
	(344.99,153.04) --
	(349.76,150.77) --
	(354.52,150.34) --
	(359.29,147.58) --
	(364.06,144.69) --
	(368.83,140.95) --
	(373.60,135.85) --
	(378.37,131.20) --
	(383.14,125.97) --
	(387.91,120.91) --
	(392.68,116.05) --
	(397.45,112.24) --
	(402.22,108.72) --
	(406.98,105.97) --
	(411.75,103.93) --
	(416.52,102.50) --
	(421.29,101.44) --
	(426.06,100.72) --
	(430.83,100.16) --
	(435.60, 99.70) --
	(440.37, 99.45) --
	(445.14, 99.21) --
	(449.91, 99.01) --
	(454.67, 98.73) --
	(459.44, 98.68) --
	(464.21, 98.67) --
	(468.98, 98.65);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (278.22,156.15) -- (468.98,156.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] (278.22,106.40) -- (468.98,106.40);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (464.21,159.09) {Initial forward};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (464.21,109.34) {Intrinsic};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] (278.22, 44.63) rectangle (468.98,230.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (271.02,105.28) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (271.02,198.20) {3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (274.22,109.68) --
	(278.22,109.68);

\path[draw=drawColor,line width= 0.8pt,line join=round] (274.22,202.61) --
	(278.22,202.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (278.22, 40.63) --
	(278.22, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (325.91, 40.63) --
	(325.91, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (373.60, 40.63) --
	(373.60, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (421.29, 40.63) --
	(421.29, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (468.98, 40.63) --
	(468.98, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (278.22, 28.62) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (325.91, 28.62) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (373.60, 28.62) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (421.29, 28.62) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (468.98, 28.62) {80};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (373.60, 11.11) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,238.49);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (257.51,137.56) {Reproduction number $\mathcal R$};
\end{scope}
\end{tikzpicture}

\end{document}