### 2020-04-30 observedrRa.tex
still making so many changes!!

% Created by tikzDevice version 0.12.3 on 2020-04-30 02:30:25
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (476.98,252.94);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (238.49,252.94);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (238.49,252.94);
\end{scope}
\begin{scope}
\path[clip] ( 39.73, 44.63) rectangle (230.49,244.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 39.73, 44.63) rectangle (230.49,244.94);
\definecolor{drawColor}{RGB}{213,94,0}
\definecolor{fillColor}{RGB}{213,94,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] ( 63.57,151.60) --
	( 68.34,146.65) --
	( 73.11,147.98) --
	( 77.88,158.41) --
	( 82.65,162.59) --
	( 87.42,151.69) --
	( 92.19,146.81) --
	( 96.96,140.33) --
	(101.73,135.45) --
	(106.50,136.85) --
	(111.26,135.88) --
	(116.03,135.19) --
	(120.80,136.14) --
	(125.57,136.07) --
	(130.34,138.08) --
	(135.11,138.96) --
	(139.88,141.42) --
	(144.65,144.70) --
	(149.42,148.23) --
	(154.19,152.08) --
	(158.95,155.95) --
	(163.72,159.65) --
	(168.49,163.18) --
	(173.26,165.92) --
	(178.03,167.97) --
	(182.80,169.68) --
	(187.57,170.79) --
	(192.34,171.85) --
	(197.11,172.27) --
	(201.88,172.82) --
	(206.65,173.27) --
	(211.41,173.45) --
	(216.18,173.54) --
	(220.95,173.58) --
	(225.72,173.59) --
	(230.49,173.60) --
	(230.49,171.46) --
	(225.72,171.42) --
	(220.95,171.39) --
	(216.18,171.28) --
	(211.41,171.15) --
	(206.65,170.99) --
	(201.88,170.55) --
	(197.11,169.95) --
	(192.34,169.00) --
	(187.57,167.59) --
	(182.80,166.10) --
	(178.03,163.30) --
	(173.26,160.34) --
	(168.49,156.48) --
	(163.72,152.40) --
	(158.95,148.33) --
	(154.19,144.23) --
	(149.42,141.58) --
	(144.65,138.62) --
	(139.88,135.15) --
	(135.11,132.35) --
	(130.34,130.38) --
	(125.57,127.70) --
	(120.80,124.26) --
	(116.03,122.94) --
	(111.26,118.36) --
	(106.50,120.00) --
	(101.73,118.03) --
	( 96.96,109.89) --
	( 92.19, 99.52) --
	( 87.42, 92.47) --
	( 82.65, 86.76) --
	( 77.88, 96.72) --
	( 73.11, 68.74) --
	( 68.34, 72.52) --
	( 63.57, 58.70) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 63.57, 97.58) --
	( 68.34,110.27) --
	( 73.11,113.96) --
	( 77.88,121.82) --
	( 82.65,121.18) --
	( 87.42,122.95) --
	( 92.19,125.71) --
	( 96.96,125.15) --
	(101.73,125.10) --
	(106.50,127.96) --
	(111.26,129.03) --
	(116.03,130.35) --
	(120.80,131.44) --
	(125.57,132.42) --
	(130.34,134.04) --
	(135.11,135.68) --
	(139.88,138.14) --
	(144.65,140.86) --
	(149.42,144.17) --
	(154.19,147.86) --
	(158.95,152.06) --
	(163.72,155.97) --
	(168.49,159.57) --
	(173.26,162.86) --
	(178.03,165.60) --
	(182.80,167.72) --
	(187.57,169.22) --
	(192.34,170.33) --
	(197.11,171.05) --
	(201.88,171.54) --
	(206.65,171.88) --
	(211.41,172.09) --
	(216.18,172.23) --
	(220.95,172.32) --
	(225.72,172.37) --
	(230.49,172.40);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 39.73,226.12) -- (230.49,226.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.73,185.10) -- (230.49,185.10);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (225.72,229.06) {Initial forward};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (225.72,188.04) {Intrinsic};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] ( 39.73, 44.63) rectangle (230.49,244.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 32.53, 40.23) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 32.53, 86.81) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 32.53,133.39) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 32.53,179.98) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 32.53,226.56) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 35.73, 44.63) --
	( 39.73, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 35.73, 91.22) --
	( 39.73, 91.22);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 35.73,137.80) --
	( 39.73,137.80);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 35.73,184.39) --
	( 39.73,184.39);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 35.73,230.97) --
	( 39.73,230.97);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 39.73, 40.63) --
	( 39.73, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 87.42, 40.63) --
	( 87.42, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (135.11, 40.63) --
	(135.11, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (182.80, 40.63) --
	(182.80, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (230.49, 40.63) --
	(230.49, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 39.73, 28.62) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 87.42, 28.62) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (135.11, 28.62) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (182.80, 28.62) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (230.49, 28.62) {80};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (135.11, 11.11) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at ( 19.02,144.79) {Mean serial interval (days)};
\end{scope}
\begin{scope}
\path[clip] (238.49,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (238.49,  0.00) rectangle (476.98,252.94);
\end{scope}
\begin{scope}
\path[clip] (278.22, 44.63) rectangle (468.98,244.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (278.22, 44.63) rectangle (468.98,244.94);
\definecolor{drawColor}{RGB}{213,94,0}
\definecolor{fillColor}{RGB}{213,94,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (302.06,190.17) --
	(306.83,179.27) --
	(311.60,176.02) --
	(316.37,177.60) --
	(321.14,178.34) --
	(325.91,168.63) --
	(330.68,162.88) --
	(335.45,153.39) --
	(340.22,145.37) --
	(344.99,145.09) --
	(349.76,145.62) --
	(354.52,145.58) --
	(359.29,144.70) --
	(364.06,145.09) --
	(368.83,145.28) --
	(373.60,145.60) --
	(378.37,146.15) --
	(383.14,147.24) --
	(387.91,148.81) --
	(392.68,149.88) --
	(397.45,152.35) --
	(402.22,153.62) --
	(406.98,154.03) --
	(411.75,155.27) --
	(416.52,156.10) --
	(421.29,156.14) --
	(426.06,156.59) --
	(430.83,156.61) --
	(435.60,156.74) --
	(440.37,156.10) --
	(445.14,156.09) --
	(449.91,156.15) --
	(454.67,156.22) --
	(459.44,156.25) --
	(464.21,156.27) --
	(468.98,156.27) --
	(468.98,151.85) --
	(464.21,151.83) --
	(459.44,151.81) --
	(454.67,151.77) --
	(449.91,151.73) --
	(445.14,151.68) --
	(440.37,152.28) --
	(435.60,152.14) --
	(430.83,151.96) --
	(426.06,151.33) --
	(421.29,151.07) --
	(416.52,150.15) --
	(411.75,149.07) --
	(406.98,148.09) --
	(402.22,146.76) --
	(397.45,145.42) --
	(392.68,143.05) --
	(387.91,143.24) --
	(383.14,142.23) --
	(378.37,139.89) --
	(373.60,137.14) --
	(368.83,138.45) --
	(364.06,136.58) --
	(359.29,132.83) --
	(354.52,132.09) --
	(349.76,129.32) --
	(344.99,129.61) --
	(340.22,127.18) --
	(335.45,126.40) --
	(330.68,117.34) --
	(325.91,116.24) --
	(321.14,109.30) --
	(316.37,110.12) --
	(311.60, 90.94) --
	(306.83,106.56) --
	(302.06,103.82) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.06,139.40) --
	(306.83,138.55) --
	(311.60,132.98) --
	(316.37,139.00) --
	(321.14,137.66) --
	(325.91,138.12) --
	(330.68,139.37) --
	(335.45,138.52) --
	(340.22,136.81) --
	(344.99,138.73) --
	(349.76,139.00) --
	(354.52,140.58) --
	(359.29,140.79) --
	(364.06,141.24) --
	(368.83,142.12) --
	(373.60,142.20) --
	(378.37,143.52) --
	(383.14,144.65) --
	(387.91,145.87) --
	(392.68,146.99) --
	(397.45,148.80) --
	(402.22,150.01) --
	(406.98,151.12) --
	(411.75,152.16) --
	(416.52,153.01) --
	(421.29,153.61) --
	(426.06,153.97) --
	(430.83,154.14) --
	(435.60,154.15) --
	(440.37,154.19) --
	(445.14,154.14) --
	(449.91,154.03) --
	(454.67,153.78) --
	(459.44,153.77) --
	(464.21,153.79) --
	(468.98,153.80);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (278.22,222.69) -- (468.98,222.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (278.22,163.11) -- (468.98,163.11);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (464.21,225.63) {Initial forward};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (464.21,166.05) {Intrinsic};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] (278.22, 44.63) rectangle (468.98,244.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (271.02, 51.35) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (271.02,162.64) {2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (274.22, 55.76) --
	(278.22, 55.76);

\path[draw=drawColor,line width= 0.8pt,line join=round] (274.22,167.05) --
	(278.22,167.05);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (278.22, 40.63) --
	(278.22, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (325.91, 40.63) --
	(325.91, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (373.60, 40.63) --
	(373.60, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (421.29, 40.63) --
	(421.29, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (468.98, 40.63) --
	(468.98, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (278.22, 28.62) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (325.91, 28.62) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (373.60, 28.62) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (421.29, 28.62) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (468.98, 28.62) {80};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (373.60, 11.11) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (257.51,144.79) {Basic reproduction number $\mathcal R_0$};
\end{scope}
\end{tikzpicture}

\end{document}