### 2020-04-30 observedrRb.tex

% Created by tikzDevice version 0.12.3 on 2020-04-30 02:30:28
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (476.98,252.94);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (238.49,252.94);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (238.49,252.94);
\end{scope}
\begin{scope}
\path[clip] ( 49.68, 44.63) rectangle (230.49,244.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 49.68, 44.63) rectangle (230.49,244.94);
\definecolor{drawColor}{RGB}{204,121,167}
\definecolor{fillColor}{RGB}{204,121,167}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] ( 72.28,220.79) --
	( 76.80,173.63) --
	( 81.32,177.48) --
	( 85.84,167.11) --
	( 90.36,170.24) --
	( 94.88,164.09) --
	( 99.40,164.49) --
	(103.92,161.24) --
	(108.44,155.52) --
	(112.96,151.79) --
	(117.49,149.31) --
	(122.01,142.65) --
	(126.53,136.63) --
	(131.05,134.42) --
	(135.57,127.76) --
	(140.09,120.12) --
	(144.61,112.80) --
	(149.13,105.10) --
	(153.65, 96.62) --
	(158.17, 88.12) --
	(162.69, 81.85) --
	(167.21, 75.98) --
	(171.73, 72.26) --
	(176.25, 69.75) --
	(180.77, 68.15) --
	(185.29, 66.82) --
	(189.81, 66.25) --
	(194.33, 65.92) --
	(198.85, 65.50) --
	(203.37, 65.12) --
	(207.89, 64.85) --
	(212.41, 64.77) --
	(216.93, 64.76) --
	(221.45, 64.72) --
	(225.97, 64.70) --
	(230.49, 64.68) --
	(230.49, 61.17) --
	(225.97, 61.18) --
	(221.45, 61.18) --
	(216.93, 61.18) --
	(212.41, 61.21) --
	(207.89, 61.26) --
	(203.37, 61.49) --
	(198.85, 61.59) --
	(194.33, 61.86) --
	(189.81, 62.24) --
	(185.29, 62.51) --
	(180.77, 63.29) --
	(176.25, 64.73) --
	(171.73, 66.54) --
	(167.21, 69.08) --
	(162.69, 72.77) --
	(158.17, 77.59) --
	(153.65, 83.55) --
	(149.13, 90.03) --
	(144.61, 97.66) --
	(140.09,105.50) --
	(135.57,112.90) --
	(131.05,120.30) --
	(126.53,125.69) --
	(122.01,127.79) --
	(117.49,129.17) --
	(112.96,134.25) --
	(108.44,131.79) --
	(103.92,134.11) --
	( 99.40,124.53) --
	( 94.88,126.59) --
	( 90.36,117.34) --
	( 85.84,117.17) --
	( 81.32,114.10) --
	( 76.80,133.97) --
	( 72.28,131.66) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 72.28,176.39) --
	( 76.80,152.48) --
	( 81.32,146.27) --
	( 85.84,146.59) --
	( 90.36,146.79) --
	( 94.88,145.82) --
	( 99.40,145.88) --
	(103.92,144.07) --
	(108.44,142.38) --
	(112.96,141.20) --
	(117.49,137.50) --
	(122.01,135.97) --
	(126.53,131.50) --
	(131.05,126.47) --
	(135.57,120.20) --
	(140.09,112.76) --
	(144.61,105.14) --
	(149.13, 97.01) --
	(153.65, 89.53) --
	(158.17, 82.81) --
	(162.69, 77.55) --
	(167.21, 73.12) --
	(171.73, 69.77) --
	(176.25, 67.31) --
	(180.77, 65.78) --
	(185.29, 64.72) --
	(189.81, 64.05) --
	(194.33, 63.59) --
	(198.85, 63.28) --
	(203.37, 63.09) --
	(207.89, 62.95) --
	(212.41, 62.87) --
	(216.93, 62.83) --
	(221.45, 62.80) --
	(225.97, 62.79) --
	(230.49, 62.78);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 49.68,148.19) -- (230.49,148.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.68, 82.86) -- (230.49, 82.86);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (225.97,151.13) {Initial forward};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (225.97, 85.80) {Intrinsic};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] ( 49.68, 44.63) rectangle (230.49,244.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 40.23) {4.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48, 77.32) {5.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,114.42) {5.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,151.51) {6.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,188.60) {6.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 42.48,225.70) {7.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 44.63) --
	( 49.68, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68, 81.73) --
	( 49.68, 81.73);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,118.82) --
	( 49.68,118.82);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,155.92) --
	( 49.68,155.92);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,193.01) --
	( 49.68,193.01);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 45.68,230.11) --
	( 49.68,230.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 49.68, 40.63) --
	( 49.68, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] ( 94.88, 40.63) --
	( 94.88, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (140.09, 40.63) --
	(140.09, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (185.29, 40.63) --
	(185.29, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (230.49, 40.63) --
	(230.49, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 49.68, 28.62) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at ( 94.88, 28.62) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (140.09, 28.62) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (185.29, 28.62) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (230.49, 28.62) {80};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (140.09, 11.11) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at ( 19.02,144.79) {Mean serial interval (days)};
\end{scope}
\begin{scope}
\path[clip] (238.49,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round,fill=fillColor] (238.49,  0.00) rectangle (476.98,252.94);
\end{scope}
\begin{scope}
\path[clip] (278.22, 44.63) rectangle (468.98,244.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (278.22, 44.63) rectangle (468.98,244.94);
\definecolor{drawColor}{RGB}{204,121,167}
\definecolor{fillColor}{RGB}{204,121,167}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (302.06,234.62) --
	(306.83,197.06) --
	(311.60,195.15) --
	(316.37,191.13) --
	(321.14,181.66) --
	(325.91,181.35) --
	(330.68,178.29) --
	(335.45,176.09) --
	(340.22,169.20) --
	(344.99,167.73) --
	(349.76,168.07) --
	(354.52,163.34) --
	(359.29,159.24) --
	(364.06,157.86) --
	(368.83,154.36) --
	(373.60,148.61) --
	(378.37,142.57) --
	(383.14,138.04) --
	(387.91,132.32) --
	(392.68,125.43) --
	(397.45,121.09) --
	(402.22,116.67) --
	(406.98,112.55) --
	(411.75,110.87) --
	(416.52,109.50) --
	(421.29,107.86) --
	(426.06,107.31) --
	(430.83,106.72) --
	(435.60,106.37) --
	(440.37,105.43) --
	(445.14,105.10) --
	(449.91,105.07) --
	(454.67,105.07) --
	(459.44,105.07) --
	(464.21,105.07) --
	(468.98,105.06) --
	(468.98,101.07) --
	(464.21,101.08) --
	(459.44,101.08) --
	(454.67,101.08) --
	(449.91,101.09) --
	(445.14,101.13) --
	(440.37,101.79) --
	(435.60,101.93) --
	(430.83,102.21) --
	(426.06,102.51) --
	(421.29,102.81) --
	(416.52,103.59) --
	(411.75,105.30) --
	(406.98,107.46) --
	(402.22,110.09) --
	(397.45,113.28) --
	(392.68,117.56) --
	(387.91,122.30) --
	(383.14,128.42) --
	(378.37,133.81) --
	(373.60,138.60) --
	(368.83,144.24) --
	(364.06,148.30) --
	(359.29,151.04) --
	(354.52,151.25) --
	(349.76,150.12) --
	(344.99,152.69) --
	(340.22,148.77) --
	(335.45,153.72) --
	(330.68,147.75) --
	(325.91,149.79) --
	(321.14,142.05) --
	(316.37,149.57) --
	(311.60,135.43) --
	(306.83,152.82) --
	(302.06,179.40) --
	cycle;

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.06,199.67) --
	(306.83,177.79) --
	(311.60,167.17) --
	(316.37,168.67) --
	(321.14,165.81) --
	(325.91,165.25) --
	(330.68,165.39) --
	(335.45,163.75) --
	(340.22,161.34) --
	(344.99,161.47) --
	(349.76,159.03) --
	(354.52,158.56) --
	(359.29,155.58) --
	(364.06,152.47) --
	(368.83,148.44) --
	(373.60,142.95) --
	(378.37,137.94) --
	(383.14,132.30) --
	(387.91,126.84) --
	(392.68,121.60) --
	(397.45,117.50) --
	(402.22,113.71) --
	(406.98,110.74) --
	(411.75,108.55) --
	(416.52,107.00) --
	(421.29,105.86) --
	(426.06,105.08) --
	(430.83,104.48) --
	(435.60,103.98) --
	(440.37,103.71) --
	(445.14,103.46) --
	(449.91,103.23) --
	(454.67,102.94) --
	(459.44,102.88) --
	(464.21,102.87) --
	(468.98,102.85);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (278.22,164.82) -- (468.98,164.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (278.22,111.20) -- (468.98,111.20);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (464.21,167.76) {Initial forward};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.42] at (464.21,114.14) {Intrinsic};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] (278.22, 44.63) rectangle (468.98,244.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (271.02,110.33) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  1.28] at (271.02,210.49) {3};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (274.22,114.74) --
	(278.22,114.74);

\path[draw=drawColor,line width= 0.8pt,line join=round] (274.22,214.90) --
	(278.22,214.90);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.8pt,line join=round] (278.22, 40.63) --
	(278.22, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (325.91, 40.63) --
	(325.91, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (373.60, 40.63) --
	(373.60, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (421.29, 40.63) --
	(421.29, 44.63);

\path[draw=drawColor,line width= 0.8pt,line join=round] (468.98, 40.63) --
	(468.98, 44.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (278.22, 28.62) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (325.91, 28.62) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (373.60, 28.62) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (421.29, 28.62) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.28] at (468.98, 28.62) {80};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (373.60, 11.11) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (476.98,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.60] at (257.51,144.79) {Basic reproduction number $\mathcal R_0$};
\end{scope}
\end{tikzpicture}

\end{document}