### 2020-04-30 serial_analysis.tex

% Created by tikzDevice version 0.12.3 on 2020-04-30 02:40:57
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (650.43,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (216.81,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (216.81,216.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 34.64, 30.69) rectangle (211.31,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (100.89,194.15) --
	(106.42,189.20) --
	(111.94,184.25) --
	(117.46,179.29) --
	(122.98,174.34) --
	(128.50,169.39) --
	(134.02,164.43) --
	(139.54,159.48) --
	(145.06,154.52) --
	(150.58,149.57) --
	(156.10,144.62) --
	(161.62,139.66) --
	(167.14,134.71) --
	(172.66,129.76) --
	(178.19,124.80) --
	(183.71,119.85) --
	(189.23,114.90) --
	(194.75,109.94) --
	(200.27,104.99) --
	(205.79,100.04) --
	(211.31, 95.08);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 34.64,104.99) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 45.69,161.34) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 51.21,136.36) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 56.73,144.62) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 62.25,144.62) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 67.77,164.43) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.29,164.43) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 78.81,144.62) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 84.33,159.48) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 89.85,137.19) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 95.37,120.06) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (100.89,120.16) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (106.42,124.02) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (111.94,113.45) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (117.46,121.02) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (122.98,112.94) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (128.50,109.78) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (134.02,113.37) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (139.54,101.81) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (145.06,105.30) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (150.58, 99.53) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (156.10, 97.67) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (161.62, 97.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (167.14, 92.46) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (172.66, 89.55) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (178.19, 95.98) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (183.71, 94.09) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (189.23, 85.17) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (200.27, 70.31) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (211.31, 60.41) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.64,104.99) --
	( 34.64,104.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.64,104.99) --
	( 34.64,104.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.64,104.99) --
	( 34.64,104.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 45.69,173.47) --
	( 45.69,173.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 45.69,173.47) --
	( 45.69,142.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 45.69,142.26) --
	( 45.69,142.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 51.21,167.16) --
	( 51.21,167.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 51.21,167.16) --
	( 51.21,115.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 51.21,115.39) --
	( 51.21,115.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 56.73,154.03) --
	( 56.73,154.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 56.73,154.03) --
	( 56.73,135.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 56.73,135.21) --
	( 56.73,135.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.25,173.35) --
	( 62.25,173.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.25,173.35) --
	( 62.25,107.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.25,107.47) --
	( 62.25,107.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.77,164.43) --
	( 67.77,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.77,164.43) --
	( 67.77,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 67.77,164.43) --
	( 67.77,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.29,187.96) --
	( 73.29,187.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.29,187.96) --
	( 73.29,140.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.29,140.90) --
	( 73.29,140.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.81,154.03) --
	( 78.81,154.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.81,154.03) --
	( 78.81,135.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.81,135.21) --
	( 78.81,135.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.33,159.48) --
	( 84.33,159.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.33,159.48) --
	( 84.33,159.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.33,159.48) --
	( 84.33,159.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.85,158.36) --
	( 89.85,158.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.85,158.36) --
	( 89.85,120.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.85,120.22) --
	( 89.85,120.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.37,168.89) --
	( 95.37,168.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.37,168.89) --
	( 95.37, 85.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.37, 85.67) --
	( 95.37, 85.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.89,157.00) --
	(100.89,157.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.89,157.00) --
	(100.89, 96.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.89, 96.94) --
	(100.89, 96.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.42,147.34) --
	(106.42,147.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.42,147.34) --
	(106.42, 94.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.42, 94.59) --
	(106.42, 94.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.94,156.63) --
	(111.94,156.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.94,156.63) --
	(111.94, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.94, 85.17) --
	(111.94, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (117.46,164.43) --
	(117.46,164.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (117.46,164.43) --
	(117.46, 94.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (117.46, 94.22) --
	(117.46, 94.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (122.98,163.56) --
	(122.98,163.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (122.98,163.56) --
	(122.98, 72.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (122.98, 72.91) --
	(122.98, 72.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (128.50,142.39) --
	(128.50,142.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (128.50,142.39) --
	(128.50, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (128.50, 85.17) --
	(128.50, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.02,141.52) --
	(134.02,141.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.02,141.52) --
	(134.02, 90.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.02, 90.13) --
	(134.02, 90.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.54,125.30) --
	(139.54,125.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.54,125.30) --
	(139.54, 70.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.54, 70.07) --
	(139.54, 70.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.06,143.75) --
	(145.06,143.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.06,143.75) --
	(145.06, 65.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.06, 65.36) --
	(145.06, 65.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.58,135.45) --
	(150.58,135.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.58,135.45) --
	(150.58, 70.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.58, 70.07) --
	(150.58, 70.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.10,129.26) --
	(156.10,129.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.10,129.26) --
	(156.10, 40.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.10, 40.59) --
	(156.10, 40.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.62,125.79) --
	(161.62,125.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.62,125.79) --
	(161.62, 58.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.62, 58.43) --
	(161.62, 58.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.14,109.94) --
	(167.14,109.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.14,109.94) --
	(167.14, 64.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.14, 64.37) --
	(167.14, 64.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.66,107.96) --
	(172.66,107.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.66,107.96) --
	(172.66, 73.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.66, 73.29) --
	(172.66, 73.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.19,109.94) --
	(178.19,109.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.19,109.94) --
	(178.19, 77.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (178.19, 77.74) --
	(178.19, 77.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.71,100.04) --
	(183.71,100.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.71,100.04) --
	(183.71, 85.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.71, 85.67) --
	(183.71, 85.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (189.23, 85.17) --
	(189.23, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (189.23, 85.17) --
	(189.23, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (189.23, 85.17) --
	(189.23, 85.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.27, 70.31) --
	(200.27, 70.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.27, 70.31) --
	(200.27, 70.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.27, 70.31) --
	(200.27, 70.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.31, 83.94) --
	(211.31, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.31, 83.94) --
	(211.31, 36.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.31, 36.88) --
	(211.31, 36.88);

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 34.64,156.47) --
	( 36.88,155.53) --
	( 39.12,154.58) --
	( 41.35,153.61) --
	( 43.59,152.64) --
	( 45.83,151.65) --
	( 48.06,150.65) --
	( 50.30,149.64) --
	( 52.54,148.61) --
	( 54.77,147.57) --
	( 57.01,146.52) --
	( 59.24,145.46) --
	( 61.48,144.38) --
	( 63.72,143.29) --
	( 65.95,142.18) --
	( 68.19,141.06) --
	( 70.43,139.93) --
	( 72.66,138.78) --
	( 74.90,137.62) --
	( 77.13,136.44) --
	( 79.37,135.25) --
	( 81.61,134.04) --
	( 83.84,132.82) --
	( 86.08,131.59) --
	( 88.32,130.33) --
	( 90.55,129.06) --
	( 92.79,127.78) --
	( 95.02,126.48) --
	( 97.26,125.19) --
	( 99.50,123.96) --
	(101.73,122.80) --
	(103.97,121.71) --
	(106.21,120.70) --
	(108.44,119.75) --
	(110.68,118.89) --
	(112.91,118.11) --
	(115.15,117.44) --
	(117.39,116.71) --
	(119.62,115.77) --
	(121.86,114.72) --
	(124.10,113.71) --
	(126.33,112.61) --
	(128.57,111.44) --
	(130.80,110.24) --
	(133.04,109.07) --
	(135.28,107.93) --
	(137.51,106.71) --
	(139.75,105.56) --
	(141.99,104.51) --
	(144.22,103.49) --
	(146.46,102.45) --
	(148.69,101.42) --
	(150.93,100.41) --
	(153.17, 99.39) --
	(155.40, 98.40) --
	(157.64, 97.46) --
	(159.88, 96.58) --
	(162.11, 95.79) --
	(164.35, 95.01) --
	(166.58, 94.17) --
	(168.82, 93.29) --
	(171.06, 92.35) --
	(173.29, 91.38) --
	(175.53, 90.35) --
	(177.77, 89.28) --
	(180.00, 88.16) --
	(182.24, 86.99) --
	(184.47, 85.78) --
	(186.71, 84.52) --
	(188.95, 83.22) --
	(191.18, 81.87) --
	(193.42, 80.48) --
	(195.66, 79.04) --
	(197.89, 77.57) --
	(200.13, 76.04) --
	(202.36, 74.48) --
	(204.60, 72.87) --
	(206.84, 71.22) --
	(209.07, 69.53) --
	(211.31, 67.79);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.64, 30.69) rectangle (211.31,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69, 37.56) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69, 87.10) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69,136.63) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.69,186.17) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89, 40.59) --
	( 34.64, 40.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89, 90.13) --
	( 34.64, 90.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89,139.66) --
	( 34.64,139.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.89,189.20) --
	( 34.64,189.20);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 51.21, 27.94) --
	( 51.21, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.42, 27.94) --
	(106.42, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.62, 27.94) --
	(161.62, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 51.21, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (106.42, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (161.62, 19.68) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (122.98,  7.64) {Symptom onset day (infector)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,112.42) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 34.64,202.22) {A. Forward serial interval};
\end{scope}
\begin{scope}
\path[clip] (216.81,  0.00) rectangle (433.62,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (216.81,  0.00) rectangle (433.62,216.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (251.45, 30.69) rectangle (428.12,194.15);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (251.45,108.69) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (275.01, 43.54) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (280.90,114.12) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (286.79, 81.54) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (292.68,130.40) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (298.57, 92.40) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (304.45, 83.87) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (310.34,101.09) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (316.23, 88.33) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (322.12, 97.47) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (328.01,105.63) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (333.90,109.04) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (339.79,110.37) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (345.68,109.44) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (351.57,103.13) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (357.45,111.73) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (363.34,116.53) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (369.23,117.73) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (375.12,115.83) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (381.01,117.47) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (386.90,120.45) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (392.79,126.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (398.68,130.12) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (404.56,128.31) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (410.45,147.37) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (416.34,140.35) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (422.23,179.26) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (428.12,157.54) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.45,108.69) --
	(251.45,108.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.45,108.69) --
	(251.45,108.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.45,108.69) --
	(251.45,108.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.01, 43.54) --
	(275.01, 43.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.01, 43.54) --
	(275.01, 43.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.01, 43.54) --
	(275.01, 43.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.90,119.27) --
	(280.90,119.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.90,119.27) --
	(280.90,108.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.90,108.96) --
	(280.90,108.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (286.79, 81.54) --
	(286.79, 81.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (286.79, 81.54) --
	(286.79, 81.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (286.79, 81.54) --
	(286.79, 81.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.68,130.40) --
	(292.68,130.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.68,130.40) --
	(292.68,130.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.68,130.40) --
	(292.68,130.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (298.57, 92.40) --
	(298.57, 92.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (298.57, 92.40) --
	(298.57, 92.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (298.57, 92.40) --
	(298.57, 92.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (304.45,139.36) --
	(304.45,139.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (304.45,139.36) --
	(304.45, 38.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (304.45, 38.12) --
	(304.45, 38.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (310.34,136.92) --
	(310.34,136.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (310.34,136.92) --
	(310.34, 86.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (310.34, 86.97) --
	(310.34, 86.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.23,107.74) --
	(316.23,107.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.23,107.74) --
	(316.23, 60.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.23, 60.78) --
	(316.23, 60.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (322.12,145.87) --
	(322.12,145.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] (322.12,145.87) --
	(322.12, 54.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (322.12, 54.67) --
	(322.12, 54.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (328.01,168.40) --
	(328.01,168.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (328.01,168.40) --
	(328.01, 69.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (328.01, 69.33) --
	(328.01, 69.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90,165.69) --
	(333.90,165.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90,165.69) --
	(333.90, 70.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90, 70.69) --
	(333.90, 70.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.79,179.26) --
	(339.79,179.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.79,179.26) --
	(339.79, 71.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.79, 71.23) --
	(339.79, 71.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.68,180.89) --
	(345.68,180.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.68,180.89) --
	(345.68, 78.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.68, 78.29) --
	(345.68, 78.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.57,141.26) --
	(351.57,141.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.57,141.26) --
	(351.57, 70.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.57, 70.69) --
	(351.57, 70.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (357.45,162.97) --
	(357.45,162.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (357.45,162.97) --
	(357.45, 76.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (357.45, 76.12) --
	(357.45, 76.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (363.34,156.19) --
	(363.34,156.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (363.34,156.19) --
	(363.34, 92.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (363.34, 92.40) --
	(363.34, 92.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.23,154.15) --
	(369.23,154.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.23,154.15) --
	(369.23, 85.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.23, 85.62) --
	(369.23, 85.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.12,153.34) --
	(375.12,153.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.12,153.34) --
	(375.12, 86.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.12, 86.97) --
	(375.12, 86.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (381.01,148.59) --
	(381.01,148.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (381.01,148.59) --
	(381.01, 91.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (381.01, 91.45) --
	(381.01, 91.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (386.90,164.47) --
	(386.90,164.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] (386.90,164.47) --
	(386.90, 80.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (386.90, 80.05) --
	(386.90, 80.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (392.79,186.72) --
	(392.79,186.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (392.79,186.72) --
	(392.79, 86.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (392.79, 86.97) --
	(392.79, 86.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.68,177.36) --
	(398.68,177.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.68,177.36) --
	(398.68,100.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.68,100.27) --
	(398.68,100.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.56,155.92) --
	(404.56,155.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.56,155.92) --
	(404.56,103.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.56,103.26) --
	(404.56,103.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.45,173.83) --
	(410.45,173.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.45,173.83) --
	(410.45,114.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.45,114.12) --
	(410.45,114.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.34,171.12) --
	(416.34,171.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.34,171.12) --
	(416.34, 93.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.34, 93.08) --
	(416.34, 93.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (422.23,179.26) --
	(422.23,179.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (422.23,179.26) --
	(422.23,179.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (422.23,179.26) --
	(422.23,179.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (428.12,157.54) --
	(428.12,157.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (428.12,157.54) --
	(428.12,157.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (428.12,157.54) --
	(428.12,157.54);

\path[draw=drawColor,line width= 1.1pt,line join=round] (251.45, 93.58) --
	(253.69, 93.29) --
	(255.93, 93.04) --
	(258.16, 92.81) --
	(260.40, 92.62) --
	(262.64, 92.45) --
	(264.87, 92.32) --
	(267.11, 92.23) --
	(269.35, 92.16) --
	(271.58, 92.13) --
	(273.82, 92.13) --
	(276.05, 92.16) --
	(278.29, 92.23) --
	(280.53, 92.33) --
	(282.76, 92.46) --
	(285.00, 92.63) --
	(287.24, 92.83) --
	(289.47, 93.07) --
	(291.71, 93.35) --
	(293.94, 93.66) --
	(296.18, 94.00) --
	(298.42, 94.38) --
	(300.65, 94.80) --
	(302.89, 95.25) --
	(305.13, 95.74) --
	(307.36, 96.27) --
	(309.60, 96.84) --
	(311.83, 97.44) --
	(314.07, 98.08) --
	(316.31, 98.76) --
	(318.54, 99.48) --
	(320.78,100.24) --
	(323.02,101.03) --
	(325.25,101.87) --
	(327.49,102.75) --
	(329.72,103.64) --
	(331.96,104.48) --
	(334.20,105.29) --
	(336.43,106.05) --
	(338.67,106.77) --
	(340.91,107.42) --
	(343.14,107.90) --
	(345.38,108.26) --
	(347.61,108.57) --
	(349.85,108.93) --
	(352.09,109.40) --
	(354.32,109.92) --
	(356.56,110.52) --
	(358.80,111.39) --
	(361.03,112.74) --
	(363.27,113.91) --
	(365.50,114.59) --
	(367.74,115.15) --
	(369.98,115.88) --
	(372.21,116.66) --
	(374.45,117.45) --
	(376.69,118.26) --
	(378.92,119.02) --
	(381.16,119.78) --
	(383.39,120.62) --
	(385.63,121.59) --
	(387.87,122.75) --
	(390.10,124.03) --
	(392.34,125.40) --
	(394.58,126.87) --
	(396.81,128.44) --
	(399.05,130.11) --
	(401.28,131.88) --
	(403.52,133.75) --
	(405.76,135.71) --
	(407.99,137.77) --
	(410.23,139.94) --
	(412.47,142.20) --
	(414.70,144.56) --
	(416.94,147.03) --
	(419.17,149.59) --
	(421.41,152.26) --
	(423.65,155.02) --
	(425.88,157.89) --
	(428.12,160.86);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.45, 30.69) rectangle (428.12,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (246.50, 35.09) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (246.50, 89.37) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (246.50,143.66) {10};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (248.70, 38.12) --
	(251.45, 38.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (248.70, 92.40) --
	(251.45, 92.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (248.70,146.69) --
	(251.45,146.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.45, 27.94) --
	(251.45, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (310.34, 27.94) --
	(310.34, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.23, 27.94) --
	(369.23, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (428.12, 27.94) --
	(428.12, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (251.45, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (310.34, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (369.23, 19.68) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (428.12, 19.68) {30};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (339.79,  7.64) {Symptom onset day (infectee)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (229.89,112.42) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (251.45,202.22) {B. Backward serial interval};
\end{scope}
\begin{scope}
\path[clip] (433.62,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (433.62,  0.00) rectangle (650.43,216.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (460.93, 30.69) rectangle (644.93,194.15);
\definecolor{drawColor}{RGB}{240,228,66}
\definecolor{fillColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (478.18,186.72) --
	(483.93,172.02) --
	(489.68,167.41) --
	(495.43,164.97) --
	(501.18,170.94) --
	(506.93,165.55) --
	(512.68,162.02) --
	(518.43,156.83) --
	(524.18,104.98) --
	(529.93, 92.34) --
	(535.68, 89.99) --
	(541.43, 81.76) --
	(547.18, 79.80) --
	(552.93, 73.13) --
	(558.68, 70.34) --
	(564.43, 68.88) --
	(570.18, 64.62) --
	(575.93, 62.47) --
	(581.68, 59.85) --
	(587.43, 58.86) --
	(593.18, 57.46) --
	(598.93, 56.56) --
	(604.68, 55.29) --
	(610.43, 55.07) --
	(616.18, 54.52) --
	(621.93, 54.37) --
	(627.68, 54.37) --
	(633.43, 54.35) --
	(639.18, 54.35) --
	(644.93, 53.84) --
	(644.93, 48.78) --
	(639.18, 49.43) --
	(633.43, 49.43) --
	(627.68, 49.75) --
	(621.93, 49.75) --
	(616.18, 49.90) --
	(610.43, 49.74) --
	(604.68, 50.17) --
	(598.93, 51.03) --
	(593.18, 51.66) --
	(587.43, 52.72) --
	(581.68, 54.51) --
	(575.93, 55.72) --
	(570.18, 58.33) --
	(564.43, 60.67) --
	(558.68, 61.38) --
	(552.93, 63.32) --
	(547.18, 68.18) --
	(541.43, 67.77) --
	(535.68, 72.19) --
	(529.93, 71.54) --
	(524.18, 74.09) --
	(518.43, 99.35) --
	(512.68,104.50) --
	(506.93,103.95) --
	(501.18,102.98) --
	(495.43,103.05) --
	(489.68, 92.92) --
	(483.93, 95.55) --
	(478.18, 92.27) --
	cycle;
\definecolor{drawColor}{RGB}{0,114,178}
\definecolor{fillColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (478.18,125.90) --
	(483.93,119.23) --
	(489.68,117.31) --
	(495.43,115.79) --
	(501.18,119.31) --
	(506.93,116.22) --
	(512.68,114.76) --
	(518.43,112.24) --
	(524.18, 84.09) --
	(529.93, 76.40) --
	(535.68, 74.61) --
	(541.43, 69.22) --
	(547.18, 68.02) --
	(552.93, 63.34) --
	(558.68, 61.72) --
	(564.43, 60.71) --
	(570.18, 57.76) --
	(575.93, 56.38) --
	(581.68, 54.72) --
	(587.43, 54.07) --
	(593.18, 52.92) --
	(598.93, 52.36) --
	(604.68, 51.40) --
	(610.43, 51.36) --
	(616.18, 51.02) --
	(621.93, 50.93) --
	(627.68, 50.93) --
	(633.43, 50.85) --
	(639.18, 50.85) --
	(644.93, 50.60) --
	(644.93, 47.19) --
	(639.18, 47.47) --
	(633.43, 47.47) --
	(627.68, 47.71) --
	(621.93, 47.71) --
	(616.18, 47.77) --
	(610.43, 47.85) --
	(604.68, 48.08) --
	(598.93, 48.58) --
	(593.18, 49.16) --
	(587.43, 49.86) --
	(581.68, 50.91) --
	(575.93, 51.95) --
	(570.18, 53.53) --
	(564.43, 55.54) --
	(558.68, 55.84) --
	(552.93, 57.08) --
	(547.18, 60.14) --
	(541.43, 60.06) --
	(535.68, 63.08) --
	(529.93, 62.94) --
	(524.18, 64.75) --
	(518.43, 79.69) --
	(512.68, 84.36) --
	(506.93, 83.64) --
	(501.18, 82.64) --
	(495.43, 82.42) --
	(489.68, 76.41) --
	(483.93, 77.92) --
	(478.18, 75.82) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (460.93, 38.12) -- (644.93, 38.12);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.18,127.94) --
	(483.93,127.84) --
	(489.68,122.64) --
	(495.43,125.45) --
	(501.18,129.15) --
	(506.93,128.97) --
	(512.68,130.66) --
	(518.43,125.72) --
	(524.18, 87.52) --
	(529.93, 82.25) --
	(535.68, 80.72) --
	(541.43, 74.16) --
	(547.18, 73.09) --
	(552.93, 67.65) --
	(558.68, 65.54) --
	(564.43, 64.88) --
	(570.18, 61.54) --
	(575.93, 59.17) --
	(581.68, 56.95) --
	(587.43, 55.38) --
	(593.18, 54.56) --
	(598.93, 53.59) --
	(604.68, 52.65) --
	(610.43, 52.34) --
	(616.18, 52.19) --
	(621.93, 52.07) --
	(627.68, 52.07) --
	(633.43, 51.94) --
	(639.18, 51.94) --
	(644.93, 51.46);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (478.18, 97.05) --
	(483.93, 96.63) --
	(489.68, 94.07) --
	(495.43, 95.65) --
	(501.18, 97.81) --
	(506.93, 97.47) --
	(512.68, 98.36) --
	(518.43, 95.36) --
	(524.18, 73.48) --
	(529.93, 69.78) --
	(535.68, 68.56) --
	(541.43, 64.35) --
	(547.18, 63.49) --
	(552.93, 60.05) --
	(558.68, 58.57) --
	(564.43, 58.06) --
	(570.18, 55.78) --
	(575.93, 54.18) --
	(581.68, 52.66) --
	(587.43, 51.66) --
	(593.18, 51.10) --
	(598.93, 50.42) --
	(604.68, 49.74) --
	(610.43, 49.50) --
	(616.18, 49.38) --
	(621.93, 49.29) --
	(627.68, 49.29) --
	(633.43, 49.20) --
	(639.18, 49.20) --
	(644.93, 48.92);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (460.93, 30.69) rectangle (644.93,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.98, 35.09) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.98, 71.15) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.98,107.22) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.98,143.29) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.98,179.35) {5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.18, 38.12) --
	(460.93, 38.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.18, 74.18) --
	(460.93, 74.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.18,110.25) --
	(460.93,110.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.18,146.32) --
	(460.93,146.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.18,182.38) --
	(460.93,182.38);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.18, 27.94) --
	(478.18, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (535.68, 27.94) --
	(535.68, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.18, 27.94) --
	(593.18, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (478.18, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (535.68, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (593.18, 19.68) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (552.93,  7.64) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (446.70,112.42) {Basic reproduction number $\mathcal R_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (528.28,146.93) rectangle (643.82,192.34);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (533.78,166.88) rectangle (548.24,181.34);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{240,228,66}
\definecolor{fillColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor,fill opacity=0.50] (534.49,167.59) rectangle (547.53,180.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (535.23,174.11) -- (546.79,174.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (533.78,152.43) rectangle (548.24,166.88);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,114,178}
\definecolor{fillColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line cap=rect,fill=fillColor,fill opacity=0.50] (534.49,153.14) rectangle (547.53,166.17);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (535.23,159.66) -- (546.79,159.66);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (553.74,171.08) {6 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (553.74,156.63) {8 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (460.93,202.22) {C. Cohort-averaged estimates};
\end{scope}
\end{tikzpicture}

\end{document}