### 2020-05-15 serial_analysis.tex
done

% Created by tikzDevice version 0.12.3 on 2020-05-14 21:20:58
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (867.24,289.08);
\begin{scope}
\path[clip] ( 10.17,  0.00) rectangle (278.91,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 10.17,  0.00) rectangle (278.91,289.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 44.82, 30.69) rectangle (273.41,266.42);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate=-45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (206.43,192.30) {maximum observable};

\node[text=drawColor,rotate=-45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 92.13, 70.86) {minimum observable};

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (130.54,266.42) --
	(137.68,259.28) --
	(144.83,252.14) --
	(151.97,244.99) --
	(159.11,237.85) --
	(166.26,230.71) --
	(173.40,223.56) --
	(180.54,216.42) --
	(187.69,209.27) --
	(194.83,202.13) --
	(201.97,194.99) --
	(209.12,187.84) --
	(216.26,180.70) --
	(223.40,173.56) --
	(230.55,166.41) --
	(237.69,159.27) --
	(244.83,152.13) --
	(251.98,144.98) --
	(259.12,137.84) --
	(266.27,130.70) --
	(273.41,123.55);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 44.82,116.41) --
	( 51.96,109.26) --
	( 59.10,102.12) --
	( 66.25, 94.98) --
	( 73.39, 87.83) --
	( 80.53, 80.69) --
	( 87.68, 73.55) --
	( 94.82, 66.40) --
	(101.96, 59.26) --
	(109.11, 52.12) --
	(116.25, 44.97) --
	(123.39, 37.83) --
	(130.54, 30.69);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 44.82,137.84) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 59.10,219.10) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 66.25,183.08) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 73.39,194.99) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 80.53,194.99) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 87.68,223.56) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 94.82,223.56) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (101.96,194.99) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (109.11,216.42) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (116.25,184.27) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (123.39,159.58) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (130.54,159.72) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (137.68,165.29) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (144.83,150.04) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (151.97,160.95) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (159.11,149.30) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (166.26,144.74) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (173.40,149.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (180.54,133.26) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (187.69,138.29) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (194.83,129.96) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (201.97,127.28) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (209.12,127.33) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (216.26,119.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (223.40,115.57) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (230.55,124.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (237.69,122.12) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (244.83,109.26) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (259.12, 87.83) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (273.41, 73.55) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.82,137.84) --
	( 44.82,137.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.82,137.84) --
	( 44.82,137.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 44.82,137.84) --
	( 44.82,137.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 59.10,236.60) --
	( 59.10,236.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 59.10,236.60) --
	( 59.10,191.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 59.10,191.59) --
	( 59.10,191.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 66.25,227.49) --
	( 66.25,227.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 66.25,227.49) --
	( 66.25,152.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 66.25,152.84) --
	( 66.25,152.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.39,208.56) --
	( 73.39,208.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.39,208.56) --
	( 73.39,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 73.39,181.41) --
	( 73.39,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.53,236.42) --
	( 80.53,236.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.53,236.42) --
	( 80.53,141.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.53,141.41) --
	( 80.53,141.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 87.68,223.56) --
	( 87.68,223.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 87.68,223.56) --
	( 87.68,223.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 87.68,223.56) --
	( 87.68,223.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.82,257.49) --
	( 94.82,257.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.82,257.49) --
	( 94.82,189.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.82,189.63) --
	( 94.82,189.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (101.96,208.56) --
	(101.96,208.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (101.96,208.56) --
	(101.96,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (101.96,181.41) --
	(101.96,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (109.11,216.42) --
	(109.11,216.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (109.11,216.42) --
	(109.11,216.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (109.11,216.42) --
	(109.11,216.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.25,214.81) --
	(116.25,214.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.25,214.81) --
	(116.25,159.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.25,159.81) --
	(116.25,159.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.39,229.99) --
	(123.39,229.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.39,229.99) --
	(123.39,109.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.39,109.98) --
	(123.39,109.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.54,212.85) --
	(130.54,212.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.54,212.85) --
	(130.54,126.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.54,126.23) --
	(130.54,126.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.68,198.92) --
	(137.68,198.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.68,198.92) --
	(137.68,122.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.68,122.84) --
	(137.68,122.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (144.83,212.31) --
	(144.83,212.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (144.83,212.31) --
	(144.83,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (144.83,109.26) --
	(144.83,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.97,223.56) --
	(151.97,223.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.97,223.56) --
	(151.97,122.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.97,122.30) --
	(151.97,122.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.11,222.31) --
	(159.11,222.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.11,222.31) --
	(159.11, 91.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.11, 91.58) --
	(159.11, 91.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.26,191.77) --
	(166.26,191.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.26,191.77) --
	(166.26,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.26,109.26) --
	(166.26,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (173.40,190.52) --
	(173.40,190.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (173.40,190.52) --
	(173.40,116.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (173.40,116.41) --
	(173.40,116.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (180.54,167.13) --
	(180.54,167.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (180.54,167.13) --
	(180.54, 87.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (180.54, 87.48) --
	(180.54, 87.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (187.69,193.74) --
	(187.69,193.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (187.69,193.74) --
	(187.69, 80.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (187.69, 80.69) --
	(187.69, 80.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (194.83,181.77) --
	(194.83,181.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (194.83,181.77) --
	(194.83, 87.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (194.83, 87.48) --
	(194.83, 87.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.97,172.84) --
	(201.97,172.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.97,172.84) --
	(201.97, 44.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.97, 44.97) --
	(201.97, 44.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.12,167.84) --
	(209.12,167.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.12,167.84) --
	(209.12, 70.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.12, 70.69) --
	(209.12, 70.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.26,144.98) --
	(216.26,144.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.26,144.98) --
	(216.26, 79.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.26, 79.26) --
	(216.26, 79.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.40,142.13) --
	(223.40,142.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.40,142.13) --
	(223.40, 92.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.40, 92.12) --
	(223.40, 92.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (230.55,144.98) --
	(230.55,144.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (230.55,144.98) --
	(230.55, 98.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (230.55, 98.55) --
	(230.55, 98.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (237.69,130.70) --
	(237.69,130.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (237.69,130.70) --
	(237.69,109.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (237.69,109.98) --
	(237.69,109.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (244.83,109.26) --
	(244.83,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (244.83,109.26) --
	(244.83,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (244.83,109.26) --
	(244.83,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (259.12, 87.83) --
	(259.12, 87.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (259.12, 87.83) --
	(259.12, 87.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (259.12, 87.83) --
	(259.12, 87.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (273.41,107.48) --
	(273.41,107.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (273.41,107.48) --
	(273.41, 39.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (273.41, 39.62) --
	(273.41, 39.62);

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 44.82,212.08) --
	( 47.71,210.73) --
	( 50.60,209.35) --
	( 53.50,207.96) --
	( 56.39,206.55) --
	( 59.28,205.13) --
	( 62.18,203.68) --
	( 65.07,202.22) --
	( 67.96,200.75) --
	( 70.86,199.25) --
	( 73.75,197.73) --
	( 76.65,196.20) --
	( 79.54,194.64) --
	( 82.43,193.07) --
	( 85.33,191.48) --
	( 88.22,189.86) --
	( 91.11,188.23) --
	( 94.01,186.57) --
	( 96.90,184.90) --
	( 99.79,183.20) --
	(102.69,181.48) --
	(105.58,179.74) --
	(108.47,177.98) --
	(111.37,176.19) --
	(114.26,174.39) --
	(117.16,172.56) --
	(120.05,170.70) --
	(122.94,168.82) --
	(125.84,166.96) --
	(128.73,165.20) --
	(131.62,163.53) --
	(134.52,161.96) --
	(137.41,160.49) --
	(140.30,159.13) --
	(143.20,157.88) --
	(146.09,156.76) --
	(148.98,155.79) --
	(151.88,154.74) --
	(154.77,153.38) --
	(157.67,151.87) --
	(160.56,150.42) --
	(163.45,148.83) --
	(166.35,147.14) --
	(169.24,145.41) --
	(172.13,143.73) --
	(175.03,142.08) --
	(177.92,140.32) --
	(180.81,138.66) --
	(183.71,137.15) --
	(186.60,135.67) --
	(189.50,134.18) --
	(192.39,132.70) --
	(195.28,131.23) --
	(198.18,129.77) --
	(201.07,128.34) --
	(203.96,126.97) --
	(206.86,125.71) --
	(209.75,124.58) --
	(212.64,123.44) --
	(215.54,122.24) --
	(218.43,120.96) --
	(221.32,119.62) --
	(224.22,118.21) --
	(227.11,116.73) --
	(230.01,115.18) --
	(232.90,113.56) --
	(235.79,111.88) --
	(238.69,110.13) --
	(241.58,108.32) --
	(244.47,106.44) --
	(247.37,104.50) --
	(250.26,102.49) --
	(253.15,100.42) --
	(256.05, 98.29) --
	(258.94, 96.10) --
	(261.83, 93.84) --
	(264.73, 91.52) --
	(267.62, 89.14) --
	(270.52, 86.70) --
	(273.41, 84.20);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 44.82, 30.69) rectangle (273.41,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 39.87, 41.94) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 39.87,113.38) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 39.87,184.81) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 39.87,256.25) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 42.07, 44.97) --
	( 44.82, 44.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 42.07,116.41) --
	( 44.82,116.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 42.07,187.84) --
	( 44.82,187.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 42.07,259.28) --
	( 44.82,259.28);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 66.25, 27.94) --
	( 66.25, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.68, 27.94) --
	(137.68, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.12, 27.94) --
	(209.12, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 66.25, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (137.68, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (209.12, 19.68) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (159.11,  7.64) {Symptom onset day (infector)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 23.25,148.55) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 44.82,274.49) {A. Forward serial interval};
\end{scope}
\begin{scope}
\path[clip] (295.68,  0.00) rectangle (571.56,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (295.68,  0.00) rectangle (571.56,289.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (330.32, 30.69) rectangle (566.06,266.42);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (388.02,200.87) {maximum observable};

\node[text=drawColor,rotate= 45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (521.60, 75.14) {minimum observable};

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (330.32,140.70) --
	(338.18,148.55) --
	(346.04,156.41) --
	(353.90,164.27) --
	(361.76,172.13) --
	(369.61,179.99) --
	(377.47,187.84) --
	(385.33,195.70) --
	(393.19,203.56) --
	(401.05,211.42) --
	(408.90,219.28) --
	(416.76,227.13) --
	(424.62,234.99) --
	(432.48,242.85) --
	(440.33,250.71) --
	(448.19,258.56) --
	(456.05,266.42);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (479.62, 30.69) --
	(487.48, 38.54) --
	(495.34, 46.40) --
	(503.20, 54.26) --
	(511.06, 62.12) --
	(518.91, 69.98) --
	(526.77, 77.83) --
	(534.63, 85.69) --
	(542.49, 93.55) --
	(550.35,101.41) --
	(558.20,109.26) --
	(566.06,117.12);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (330.32,140.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (361.76, 46.40) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (369.61,148.55) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (377.47,101.41) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (385.33,172.13) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (393.19,117.12) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (401.05,104.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (408.90,129.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (416.76,111.23) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (424.62,124.46) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (432.48,136.28) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (440.33,141.20) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (448.19,143.14) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (456.05,141.78) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (463.91,132.65) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (471.77,145.10) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (479.62,152.05) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.48,153.79) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (495.34,151.04) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (503.20,153.41) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (511.06,157.72) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (518.91,166.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (526.77,171.71) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (534.63,169.11) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (542.49,196.68) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (550.35,186.53) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (558.20,242.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (566.06,211.42) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (330.32,140.70) --
	(330.32,140.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (330.32,140.70) --
	(330.32,140.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (330.32,140.70) --
	(330.32,140.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.76, 46.40) --
	(361.76, 46.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.76, 46.40) --
	(361.76, 46.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.76, 46.40) --
	(361.76, 46.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.61,156.02) --
	(369.61,156.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.61,156.02) --
	(369.61,141.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.61,141.09) --
	(369.61,141.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (377.47,101.41) --
	(377.47,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (377.47,101.41) --
	(377.47,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (377.47,101.41) --
	(377.47,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (385.33,172.13) --
	(385.33,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (385.33,172.13) --
	(385.33,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (385.33,172.13) --
	(385.33,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.19,117.12) --
	(393.19,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.19,117.12) --
	(393.19,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.19,117.12) --
	(393.19,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.05,185.09) --
	(401.05,185.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.05,185.09) --
	(401.05, 38.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.05, 38.54) --
	(401.05, 38.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (408.90,181.56) --
	(408.90,181.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (408.90,181.56) --
	(408.90,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (408.90,109.26) --
	(408.90,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.76,139.32) --
	(416.76,139.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.76,139.32) --
	(416.76, 71.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.76, 71.35) --
	(416.76, 71.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (424.62,194.52) --
	(424.62,194.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (424.62,194.52) --
	(424.62, 62.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (424.62, 62.51) --
	(424.62, 62.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (432.48,227.13) --
	(432.48,227.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (432.48,227.13) --
	(432.48, 83.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (432.48, 83.73) --
	(432.48, 83.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (440.33,223.20) --
	(440.33,223.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (440.33,223.20) --
	(440.33, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (440.33, 85.69) --
	(440.33, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (448.19,242.85) --
	(448.19,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (448.19,242.85) --
	(448.19, 86.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (448.19, 86.48) --
	(448.19, 86.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (456.05,245.21) --
	(456.05,245.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (456.05,245.21) --
	(456.05, 96.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (456.05, 96.69) --
	(456.05, 96.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.91,187.84) --
	(463.91,187.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.91,187.84) --
	(463.91, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.91, 85.69) --
	(463.91, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (471.77,219.28) --
	(471.77,219.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (471.77,219.28) --
	(471.77, 93.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (471.77, 93.55) --
	(471.77, 93.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.62,209.45) --
	(479.62,209.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.62,209.45) --
	(479.62,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.62,117.12) --
	(479.62,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.48,206.51) --
	(487.48,206.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.48,206.51) --
	(487.48,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.48,107.30) --
	(487.48,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (495.34,205.33) --
	(495.34,205.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (495.34,205.33) --
	(495.34,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (495.34,109.26) --
	(495.34,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (503.20,198.45) --
	(503.20,198.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (503.20,198.45) --
	(503.20,115.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (503.20,115.75) --
	(503.20,115.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (511.06,221.44) --
	(511.06,221.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (511.06,221.44) --
	(511.06, 99.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (511.06, 99.25) --
	(511.06, 99.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (518.91,253.65) --
	(518.91,253.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (518.91,253.65) --
	(518.91,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (518.91,109.26) --
	(518.91,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77,240.10) --
	(526.77,240.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77,240.10) --
	(526.77,128.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77,128.52) --
	(526.77,128.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (534.63,209.06) --
	(534.63,209.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (534.63,209.06) --
	(534.63,132.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (534.63,132.84) --
	(534.63,132.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (542.49,234.99) --
	(542.49,234.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (542.49,234.99) --
	(542.49,148.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (542.49,148.55) --
	(542.49,148.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.35,231.06) --
	(550.35,231.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.35,231.06) --
	(550.35,118.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (550.35,118.10) --
	(550.35,118.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (558.20,242.85) --
	(558.20,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (558.20,242.85) --
	(558.20,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (558.20,242.85) --
	(558.20,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (566.06,211.42) --
	(566.06,211.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (566.06,211.42) --
	(566.06,211.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (566.06,211.42) --
	(566.06,211.42);

\path[draw=drawColor,line width= 1.1pt,line join=round] (330.32,118.83) --
	(333.31,118.41) --
	(336.29,118.04) --
	(339.28,117.71) --
	(342.26,117.43) --
	(345.24,117.20) --
	(348.23,117.01) --
	(351.21,116.87) --
	(354.20,116.77) --
	(357.18,116.72) --
	(360.16,116.72) --
	(363.15,116.77) --
	(366.13,116.87) --
	(369.12,117.01) --
	(372.10,117.21) --
	(375.08,117.45) --
	(378.07,117.75) --
	(381.05,118.09) --
	(384.04,118.49) --
	(387.02,118.94) --
	(390.00,119.44) --
	(392.99,119.99) --
	(395.97,120.59) --
	(398.96,121.25) --
	(401.94,121.96) --
	(404.92,122.73) --
	(407.91,123.54) --
	(410.89,124.42) --
	(413.88,125.35) --
	(416.86,126.33) --
	(419.84,127.37) --
	(422.83,128.47) --
	(425.81,129.62) --
	(428.80,130.83) --
	(431.78,132.10) --
	(434.76,133.38) --
	(437.75,134.61) --
	(440.73,135.77) --
	(443.72,136.88) --
	(446.70,137.92) --
	(449.68,138.87) --
	(452.67,139.56) --
	(455.65,140.07) --
	(458.64,140.53) --
	(461.62,141.04) --
	(464.60,141.73) --
	(467.59,142.49) --
	(470.57,143.35) --
	(473.56,144.61) --
	(476.54,146.56) --
	(479.52,148.26) --
	(482.51,149.31) --
	(485.49,150.22) --
	(488.48,151.24) --
	(491.46,152.28) --
	(494.44,153.32) --
	(497.43,154.37) --
	(500.41,155.31) --
	(503.40,156.26) --
	(506.38,157.32) --
	(509.36,158.61) --
	(512.35,160.23) --
	(515.33,162.05) --
	(518.32,164.02) --
	(521.30,166.15) --
	(524.28,168.43) --
	(527.27,170.87) --
	(530.25,173.46) --
	(533.24,176.21) --
	(536.22,179.13) --
	(539.20,182.20) --
	(542.19,185.42) --
	(545.17,188.81) --
	(548.16,192.36) --
	(551.14,196.07) --
	(554.12,199.95) --
	(557.11,203.98) --
	(560.09,208.18) --
	(563.08,212.54) --
	(566.06,217.06);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (330.32, 30.69) rectangle (566.06,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (325.37, 35.51) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (325.37,114.09) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (325.37,192.67) {10};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.57, 38.54) --
	(330.32, 38.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.57,117.12) --
	(330.32,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.57,195.70) --
	(330.32,195.70);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (330.32, 27.94) --
	(330.32, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (408.90, 27.94) --
	(408.90, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.48, 27.94) --
	(487.48, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (566.06, 27.94) --
	(566.06, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (330.32, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (408.90, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (487.48, 19.68) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (566.06, 19.68) {30};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (448.19,  7.64) {Symptom onset day (infectee)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (308.76,148.55) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (330.32,274.49) {B. Backward serial interval};
\end{scope}
\begin{scope}
\path[clip] (578.16,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (578.16,  0.00) rectangle (867.24,289.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (605.47, 30.69) rectangle (861.74,266.42);
\definecolor{drawColor}{RGB}{240,228,66}
\definecolor{fillColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (629.50,255.71) --
	(637.51,234.51) --
	(645.51,227.85) --
	(653.52,224.34) --
	(661.53,232.95) --
	(669.54,225.18) --
	(677.55,220.08) --
	(685.56,212.60) --
	(693.56,137.83) --
	(701.57,119.60) --
	(709.58,116.21) --
	(717.59,104.34) --
	(725.60,101.52) --
	(733.61, 91.89) --
	(741.61, 87.88) --
	(749.62, 85.77) --
	(757.63, 79.62) --
	(765.64, 76.53) --
	(773.65, 72.74) --
	(781.66, 71.32) --
	(789.66, 69.30) --
	(797.67, 68.00) --
	(805.68, 66.17) --
	(813.69, 65.86) --
	(821.70, 65.06) --
	(829.71, 64.84) --
	(837.71, 64.84) --
	(845.72, 64.82) --
	(853.73, 64.82) --
	(861.74, 64.08) --
	(861.74, 56.77) --
	(853.73, 57.72) --
	(845.72, 57.72) --
	(837.71, 58.18) --
	(829.71, 58.18) --
	(821.70, 58.40) --
	(813.69, 58.16) --
	(805.68, 58.79) --
	(797.67, 60.03) --
	(789.66, 60.94) --
	(781.66, 62.46) --
	(773.65, 65.05) --
	(765.64, 66.78) --
	(757.63, 70.56) --
	(749.62, 73.93) --
	(741.61, 74.95) --
	(733.61, 77.75) --
	(725.60, 84.76) --
	(717.59, 84.17) --
	(709.58, 90.54) --
	(701.57, 89.60) --
	(693.56, 93.27) --
	(685.56,129.71) --
	(677.55,137.14) --
	(669.54,136.34) --
	(661.53,134.94) --
	(653.52,135.04) --
	(645.51,120.43) --
	(637.51,124.23) --
	(629.50,119.49) --
	cycle;
\definecolor{drawColor}{RGB}{0,114,178}
\definecolor{fillColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round,line cap=round,fill=fillColor,fill opacity=0.50] (629.50,167.99) --
	(637.51,158.37) --
	(645.51,155.61) --
	(653.52,153.41) --
	(661.53,158.49) --
	(669.54,154.03) --
	(677.55,151.93) --
	(685.56,148.30) --
	(693.56,107.71) --
	(701.57, 96.60) --
	(709.58, 94.03) --
	(717.59, 86.26) --
	(725.60, 84.52) --
	(733.61, 77.77) --
	(741.61, 75.45) --
	(749.62, 73.98) --
	(757.63, 69.74) --
	(765.64, 67.73) --
	(773.65, 65.35) --
	(781.66, 64.41) --
	(789.66, 62.76) --
	(797.67, 61.94) --
	(805.68, 60.56) --
	(813.69, 60.51) --
	(821.70, 60.01) --
	(829.71, 59.88) --
	(837.71, 59.88) --
	(845.72, 59.77) --
	(853.73, 59.77) --
	(861.74, 59.40) --
	(861.74, 54.49) --
	(853.73, 54.89) --
	(845.72, 54.89) --
	(837.71, 55.23) --
	(829.71, 55.23) --
	(821.70, 55.32) --
	(813.69, 55.43) --
	(805.68, 55.76) --
	(797.67, 56.48) --
	(789.66, 57.32) --
	(781.66, 58.34) --
	(773.65, 59.85) --
	(765.64, 61.35) --
	(757.63, 63.63) --
	(749.62, 66.52) --
	(741.61, 66.97) --
	(733.61, 68.74) --
	(725.60, 73.15) --
	(717.59, 73.04) --
	(709.58, 77.40) --
	(701.57, 77.19) --
	(693.56, 79.82) --
	(685.56,101.35) --
	(677.55,108.09) --
	(669.54,107.05) --
	(661.53,105.61) --
	(653.52,105.29) --
	(645.51, 96.62) --
	(637.51, 98.80) --
	(629.50, 95.78) --
	cycle;
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (605.47, 41.40) -- (861.74, 41.40);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (629.50,170.94) --
	(637.51,170.79) --
	(645.51,163.30) --
	(653.52,167.34) --
	(661.53,172.68) --
	(669.54,172.43) --
	(677.55,174.86) --
	(685.56,167.73) --
	(693.56,112.65) --
	(701.57,105.04) --
	(709.58,102.84) --
	(717.59, 93.38) --
	(725.60, 91.84) --
	(733.61, 83.99) --
	(741.61, 80.95) --
	(749.62, 80.00) --
	(757.63, 75.18) --
	(765.64, 71.77) --
	(773.65, 68.56) --
	(781.66, 66.30) --
	(789.66, 65.11) --
	(797.67, 63.71) --
	(805.68, 62.37) --
	(813.69, 61.91) --
	(821.70, 61.69) --
	(829.71, 61.52) --
	(837.71, 61.52) --
	(845.72, 61.33) --
	(853.73, 61.33) --
	(861.74, 60.65);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (629.50,126.39) --
	(637.51,125.79) --
	(645.51,122.09) --
	(653.52,124.37) --
	(661.53,127.48) --
	(669.54,126.99) --
	(677.55,128.28) --
	(685.56,123.96) --
	(693.56, 92.40) --
	(701.57, 87.06) --
	(709.58, 85.30) --
	(717.59, 79.23) --
	(725.60, 77.99) --
	(733.61, 73.04) --
	(741.61, 70.90) --
	(749.62, 70.16) --
	(757.63, 66.88) --
	(765.64, 64.57) --
	(773.65, 62.37) --
	(781.66, 60.94) --
	(789.66, 60.13) --
	(797.67, 59.14) --
	(805.68, 58.16) --
	(813.69, 57.82) --
	(821.70, 57.64) --
	(829.71, 57.52) --
	(837.71, 57.52) --
	(845.72, 57.39) --
	(853.73, 57.39) --
	(861.74, 56.99);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (605.47, 30.69) rectangle (861.74,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (600.52, 38.37) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (600.52, 90.38) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (600.52,142.40) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (600.52,194.41) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (600.52,246.42) {5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.72, 41.40) --
	(605.47, 41.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.72, 93.41) --
	(605.47, 93.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.72,145.43) --
	(605.47,145.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.72,197.44) --
	(605.47,197.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.72,249.45) --
	(605.47,249.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (629.50, 27.94) --
	(629.50, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (709.58, 27.94) --
	(709.58, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (789.66, 27.94) --
	(789.66, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (629.50, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (709.58, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (789.66, 19.68) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (733.61,  7.64) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (591.24,148.55) {Basic reproduction number $\mathcal R_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (721.97,208.36) rectangle (837.50,253.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (727.47,228.31) rectangle (741.92,242.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{240,228,66}
\definecolor{fillColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor,fill opacity=0.50] (728.18,229.02) rectangle (741.21,242.05);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (728.91,235.54) -- (740.48,235.54);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (727.47,213.86) rectangle (741.92,228.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,114,178}
\definecolor{fillColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line cap=rect,fill=fillColor,fill opacity=0.50] (728.18,214.57) rectangle (741.21,227.60);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (728.91,221.09) -- (740.48,221.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (747.42,232.51) {6 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (747.42,218.05) {8 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (867.24,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (605.47,274.49) {C. Cohort-averaged estimates};
\end{scope}
\end{tikzpicture}

\end{document}