### 2020-05-28 serial_guide.tex
done

% Created by tikzDevice version 0.12.3 on 2020-05-27 19:48:26
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,578.16);
\begin{scope}
\path[clip] (  0.00,385.44) rectangle (722.70,578.16);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,385.44) rectangle (722.70,578.16);
\end{scope}
\begin{scope}
\path[clip] (  8.25,393.69) rectangle (717.20,555.50);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (  8.25,393.69) rectangle (717.20,555.50);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04,437.82) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.16,437.82) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (443.29,437.82) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (604.41,437.82) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04,486.85) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.16,535.89) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (443.29,511.37) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04,462.34) circle (  3.57);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (282.16,421.76) {$p+\tau_{\rm i1}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (121.04,421.76) {$p$};
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (121.04,409.50) {Infector (reference cohort)};
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,409.50) {Infectee};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,421.76) {$p+\tau_{\rm g}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (604.41,421.76) {$p+\tau_{\rm g}+\tau_{\rm i2}$};

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 40.48,437.82) -- (684.98,437.82);

\path[draw=drawColor,line width= 1.1pt,line join=round] (672.46,430.59) --
	(684.98,437.82) --
	(672.46,445.05);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (121.04,437.82) -- (121.04,486.85);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (282.16,437.82) -- (282.16,535.89);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (443.29,437.82) -- (443.29,511.37);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (604.41,437.82) -- (604.41,535.89);

\path[draw=drawColor,line width= 1.1pt,line join=round] (121.04,462.34) -- (282.16,462.34);

\path[draw=drawColor,line width= 1.1pt,line join=round] (275.90,458.72) --
	(282.16,462.34) --
	(275.90,465.95);

\path[draw=drawColor,line width= 1.1pt,line join=round] (121.04,486.85) -- (443.29,486.85);

\path[draw=drawColor,line width= 1.1pt,line join=round] (437.03,483.24) --
	(443.29,486.85) --
	(437.03,490.47);

\path[draw=drawColor,line width= 1.1pt,line join=round] (443.29,511.37) -- (604.41,511.37);

\path[draw=drawColor,line width= 1.1pt,line join=round] (598.15,507.76) --
	(604.41,511.37) --
	(598.15,514.99);

\path[draw=drawColor,line width= 1.7pt,line join=round] (282.16,535.89) -- (604.41,535.89);

\path[draw=drawColor,line width= 1.7pt,line join=round] (598.15,532.28) --
	(604.41,535.89) --
	(598.15,539.50);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,544.35) {Intrinsic serial interval, $\tau_{\rm s} = (\tau_{\rm g} + \tau_{\rm i2}) - \tau_{\rm i1}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (201.60,470.79) {Intrinsic incubation period, $\tau_{\rm i1}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (201.60,495.31) {Intrinsic generation interval, $\tau_{\rm g}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (523.85,519.83) {Intrinsic incubation period, $\tau_{\rm i2}$};
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round,fill=fillColor] (121.04,437.82) circle (  5.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  8.25,563.57) {A. Intrinsic serial interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,192.72) rectangle (722.70,385.44);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,192.72) rectangle (722.70,385.44);
\end{scope}
\begin{scope}
\path[clip] (  8.25,200.97) rectangle (717.20,362.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (  8.25,200.97) rectangle (717.20,362.78);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04,245.10) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.16,245.10) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (443.29,245.10) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (604.41,245.10) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04,294.13) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.16,343.17) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (443.29,318.65) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.16,269.62) circle (  3.57);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (282.16,229.04) {$p$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (121.04,229.04) {$p-\tau_{\rm i1}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (121.04,216.78) {Infector};
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (282.16,216.78) {Reference cohort};
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,216.78) {Infectee};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,229.04) {$p-\tau_{\rm i1}+\tau_{\rm g}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (604.41,229.04) {$p-\tau_{\rm i1}+\tau_{\rm g}+\tau_{\rm i2}$};

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 40.48,245.10) -- (684.98,245.10);

\path[draw=drawColor,line width= 1.1pt,line join=round] (672.46,237.87) --
	(684.98,245.10) --
	(672.46,252.33);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (121.04,245.10) -- (121.04,294.13);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (282.16,245.10) -- (282.16,343.17);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (443.29,245.10) -- (443.29,318.65);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (604.41,245.10) -- (604.41,343.17);

\path[draw=drawColor,line width= 1.1pt,line join=round] (282.16,269.62) -- (121.04,269.62);

\path[draw=drawColor,line width= 1.1pt,line join=round] (127.30,273.23) --
	(121.04,269.62) --
	(127.30,266.00);

\path[draw=drawColor,line width= 1.1pt,line join=round] (121.04,294.13) -- (443.29,294.13);

\path[draw=drawColor,line width= 1.1pt,line join=round] (437.03,290.52) --
	(443.29,294.13) --
	(437.03,297.75);

\path[draw=drawColor,line width= 1.1pt,line join=round] (443.29,318.65) -- (604.41,318.65);

\path[draw=drawColor,line width= 1.1pt,line join=round] (598.15,315.04) --
	(604.41,318.65) --
	(598.15,322.27);

\path[draw=drawColor,line width= 1.7pt,line join=round] (282.16,343.17) -- (604.41,343.17);

\path[draw=drawColor,line width= 1.7pt,line join=round] (598.15,339.56) --
	(604.41,343.17) --
	(598.15,346.78);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,351.63) {Forward serial interval, $\tau_{\rm s} = (\tau_{\rm g} + \tau_{\rm i2}) - \tau_{\rm i1}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (201.60,278.07) {Backward incubation period, $\tau_{\rm i1}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (201.60,302.59) {Forward generation interval, $\tau_{\rm g}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (523.85,327.11) {Forward incubation period, $\tau_{\rm i2}$};
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round,fill=fillColor] (282.16,245.10) circle (  5.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  8.25,370.85) {B. Forward serial interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,192.72);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,192.72);
\end{scope}
\begin{scope}
\path[clip] (  8.25,  8.25) rectangle (717.20,170.06);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (  8.25,  8.25) rectangle (717.20,170.06);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04, 52.38) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.16, 52.38) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (443.29, 52.38) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (604.41, 52.38) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (604.41, 76.90) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.04,125.93) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (604.41,150.45) circle (  3.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (443.29,101.41) circle (  3.57);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (604.41, 36.32) {$s$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29, 36.32) {$s-\tau_{\rm i2}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (121.04, 36.32) {$s-\tau_{\rm i2}-\tau_{\rm g}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (282.16, 36.32) {$s-\tau_{\rm i2}-\tau_{\rm g}+\tau_{\rm i1}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (121.04, 24.06) {Infector};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29, 24.06) {Infectee};
\definecolor{drawColor}{RGB}{255,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (604.41, 24.06) {Reference cohort};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 40.48, 52.38) -- (684.98, 52.38);

\path[draw=drawColor,line width= 1.1pt,line join=round] (672.46, 45.15) --
	(684.98, 52.38) --
	(672.46, 59.61);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (121.04, 52.38) -- (121.04,125.93);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (282.16, 52.38) -- (282.16,150.45);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (443.29, 52.38) -- (443.29,101.41);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (604.41, 52.38) -- (604.41,150.45);

\path[draw=drawColor,line width= 1.1pt,line join=round] (604.41, 76.90) -- (443.29, 76.90);

\path[draw=drawColor,line width= 1.1pt,line join=round] (449.55, 80.51) --
	(443.29, 76.90) --
	(449.55, 73.28);

\path[draw=drawColor,line width= 1.1pt,line join=round] (443.29,101.41) -- (121.04,101.41);

\path[draw=drawColor,line width= 1.1pt,line join=round] (127.30,105.03) --
	(121.04,101.41) --
	(127.30, 97.80);

\path[draw=drawColor,line width= 1.1pt,line join=round] (121.04,125.93) -- (282.16,125.93);

\path[draw=drawColor,line width= 1.1pt,line join=round] (275.90,122.32) --
	(282.16,125.93) --
	(275.90,129.55);

\path[draw=drawColor,line width= 1.7pt,line join=round] (604.41,150.45) -- (282.16,150.45);

\path[draw=drawColor,line width= 1.7pt,line join=round] (288.42,154.06) --
	(282.16,150.45) --
	(288.42,146.84);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (443.29,158.91) {Backward serial interval, $\tau_{\rm s} = (\tau_{\rm g} + \tau_{\rm i2}) - \tau_{\rm i1}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (523.85, 85.35) {Backward incubation period, $\tau_{\rm i2}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (362.72,109.87) {Backward generation interval, $\tau_{\rm g}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (201.60,134.39) {Forward incubation period, $\tau_{\rm i1}$};
\definecolor{fillColor}{RGB}{255,0,0}

\path[draw=drawColor,line width= 1.5pt,line join=round,line cap=round,fill=fillColor] (604.41, 52.38) circle (  5.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  8.25,178.13) {C. Backward serial interval};
\end{scope}
\end{tikzpicture}

\end{document}