### 2020-08-28 rR.tex
update

% Created by tikzDevice version 0.12.3 on 2020-08-27 23:38:11
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,433.62);
\begin{scope}
\path[clip] (  0.00,216.81) rectangle (289.08,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,216.81) rectangle (289.08,433.62);
\end{scope}
\begin{scope}
\path[clip] ( 27.31,247.50) rectangle (283.58,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31,247.50) rectangle (283.58,410.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31,247.50) --
	( 30.02,248.23) --
	( 32.72,248.97) --
	( 35.42,249.73) --
	( 38.13,250.50) --
	( 40.83,251.28) --
	( 43.53,252.08) --
	( 46.24,252.88) --
	( 48.94,253.70) --
	( 51.64,254.53) --
	( 54.35,255.37) --
	( 57.05,256.23) --
	( 59.75,257.09) --
	( 62.46,257.97) --
	( 65.16,258.86) --
	( 67.86,259.76) --
	( 70.57,260.67) --
	( 73.27,261.60) --
	( 75.97,262.53) --
	( 78.67,263.48) --
	( 81.38,264.43) --
	( 84.08,265.40) --
	( 86.78,266.38) --
	( 89.49,267.37) --
	( 92.19,268.38) --
	( 94.89,269.40) --
	( 97.60,270.43) --
	(100.30,271.47) --
	(103.00,272.54) --
	(105.71,273.61) --
	(108.41,274.70) --
	(111.11,275.81) --
	(113.82,276.93) --
	(116.52,278.07) --
	(119.22,279.22) --
	(121.93,280.38) --
	(124.63,281.55) --
	(127.33,282.74) --
	(130.04,283.95) --
	(132.74,285.16) --
	(135.44,286.40) --
	(138.15,287.65) --
	(140.85,288.91) --
	(143.55,290.20) --
	(146.26,291.50) --
	(148.96,292.81) --
	(151.66,294.15) --
	(154.37,295.51) --
	(157.07,296.88) --
	(159.77,298.27) --
	(162.48,299.67) --
	(165.18,301.09) --
	(167.88,302.53) --
	(170.58,303.98) --
	(173.29,305.45) --
	(175.99,306.94) --
	(178.69,308.45) --
	(181.40,309.97) --
	(184.10,311.51) --
	(186.80,313.08) --
	(189.51,314.67) --
	(192.21,316.27) --
	(194.91,317.90) --
	(197.62,319.55) --
	(200.32,321.23) --
	(203.02,322.92) --
	(205.73,324.64) --
	(208.43,326.37) --
	(211.13,328.13) --
	(213.84,329.90) --
	(216.54,331.70) --
	(219.24,333.51) --
	(221.95,335.35) --
	(224.65,337.21) --
	(227.35,339.08) --
	(230.06,340.98) --
	(232.76,342.90) --
	(235.46,344.83) --
	(238.17,346.79) --
	(240.87,348.77);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 61.48,257.69) --
	( 64.19,258.56) --
	( 66.89,259.45) --
	( 69.59,260.35) --
	( 72.30,261.26) --
	( 75.00,262.19) --
	( 77.70,263.13) --
	( 80.40,264.08) --
	( 83.11,265.05) --
	( 85.81,266.03) --
	( 88.51,267.03) --
	( 91.22,268.04) --
	( 93.92,269.06) --
	( 96.62,270.09) --
	( 99.33,271.14) --
	(102.03,272.21) --
	(104.73,273.28) --
	(107.44,274.37) --
	(110.14,275.47) --
	(112.84,276.58) --
	(115.55,277.71) --
	(118.25,278.84) --
	(120.95,279.99) --
	(123.66,281.16) --
	(126.36,282.34) --
	(129.06,283.53) --
	(131.77,284.74) --
	(134.47,285.97) --
	(137.17,287.21) --
	(139.88,288.47) --
	(142.58,289.75) --
	(145.28,291.04) --
	(147.99,292.36) --
	(150.69,293.69) --
	(153.39,295.03) --
	(156.10,296.39) --
	(158.80,297.76) --
	(161.50,299.15) --
	(164.21,300.55) --
	(166.91,301.97) --
	(169.61,303.41) --
	(172.31,304.87) --
	(175.02,306.34) --
	(177.72,307.84) --
	(180.42,309.35) --
	(183.13,310.89) --
	(185.83,312.45) --
	(188.53,314.03) --
	(191.24,315.62) --
	(193.94,317.24) --
	(196.64,318.87) --
	(199.35,320.53) --
	(202.05,322.19) --
	(204.75,323.88) --
	(207.46,325.59) --
	(210.16,327.32) --
	(212.86,329.07) --
	(215.57,330.84) --
	(218.27,332.63) --
	(220.97,334.45) --
	(223.68,336.29) --
	(226.38,338.15) --
	(229.08,340.04) --
	(231.79,341.96) --
	(234.49,343.90) --
	(237.19,345.87) --
	(239.90,347.86) --
	(242.60,349.87) --
	(245.30,351.90) --
	(248.01,353.96) --
	(250.71,356.04) --
	(253.41,358.15) --
	(256.12,360.28) --
	(258.82,362.43) --
	(261.52,364.60) --
	(264.22,366.80) --
	(266.93,369.02) --
	(269.63,371.27) --
	(272.33,373.53) --
	(275.04,375.82);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 52.94,254.98) --
	( 55.64,255.81) --
	( 58.35,256.66) --
	( 61.05,257.53) --
	( 63.75,258.41) --
	( 66.46,259.30) --
	( 69.16,260.20) --
	( 71.86,261.12) --
	( 74.57,262.04) --
	( 77.27,262.99) --
	( 79.97,263.94) --
	( 82.68,264.91) --
	( 85.38,265.89) --
	( 88.08,266.89) --
	( 90.79,267.90) --
	( 93.49,268.92) --
	( 96.19,269.95) --
	( 98.89,271.00) --
	(101.60,272.05) --
	(104.30,273.12) --
	(107.00,274.20) --
	(109.71,275.30) --
	(112.41,276.40) --
	(115.11,277.52) --
	(117.82,278.66) --
	(120.52,279.81) --
	(123.22,280.97) --
	(125.93,282.15) --
	(128.63,283.34) --
	(131.33,284.55) --
	(134.04,285.78) --
	(136.74,287.03) --
	(139.44,288.29) --
	(142.15,289.56) --
	(144.85,290.85) --
	(147.55,292.16) --
	(150.26,293.48) --
	(152.96,294.81) --
	(155.66,296.16) --
	(158.37,297.52) --
	(161.07,298.90) --
	(163.77,300.30) --
	(166.48,301.72) --
	(169.18,303.16) --
	(171.88,304.61) --
	(174.59,306.09) --
	(177.29,307.58) --
	(179.99,309.10) --
	(182.70,310.64) --
	(185.40,312.20) --
	(188.10,313.77) --
	(190.80,315.36) --
	(193.51,316.96) --
	(196.21,318.59) --
	(198.91,320.23) --
	(201.62,321.89) --
	(204.32,323.58) --
	(207.02,325.28) --
	(209.73,327.01) --
	(212.43,328.76) --
	(215.13,330.54) --
	(217.84,332.33) --
	(220.54,334.16) --
	(223.24,336.01) --
	(225.95,337.88) --
	(228.65,339.78) --
	(231.35,341.70) --
	(234.06,343.64) --
	(236.76,345.60) --
	(239.46,347.59) --
	(242.17,349.60) --
	(244.87,351.63) --
	(247.57,353.69) --
	(250.28,355.76) --
	(252.98,357.86) --
	(255.68,359.99) --
	(258.39,362.13) --
	(261.09,364.30) --
	(263.79,366.49) --
	(266.50,368.71);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 44.40,252.37) --
	( 47.10,253.17) --
	( 49.80,253.99) --
	( 52.51,254.82) --
	( 55.21,255.66) --
	( 57.91,256.51) --
	( 60.62,257.37) --
	( 63.32,258.25) --
	( 66.02,259.14) --
	( 68.73,260.05) --
	( 71.43,260.97) --
	( 74.13,261.90) --
	( 76.84,262.84) --
	( 79.54,263.79) --
	( 82.24,264.76) --
	( 84.95,265.74) --
	( 87.65,266.74) --
	( 90.35,267.74) --
	( 93.06,268.76) --
	( 95.76,269.79) --
	( 98.46,270.83) --
	(101.17,271.88) --
	(103.87,272.95) --
	(106.57,274.03) --
	(109.28,275.12) --
	(111.98,276.23) --
	(114.68,277.35) --
	(117.39,278.49) --
	(120.09,279.64) --
	(122.79,280.80) --
	(125.49,281.98) --
	(128.20,283.18) --
	(130.90,284.39) --
	(133.60,285.61) --
	(136.31,286.85) --
	(139.01,288.09) --
	(141.71,289.35) --
	(144.42,290.63) --
	(147.12,291.92) --
	(149.82,293.22) --
	(152.53,294.54) --
	(155.23,295.88) --
	(157.93,297.23) --
	(160.64,298.61) --
	(163.34,300.00) --
	(166.04,301.41) --
	(168.75,302.85) --
	(171.45,304.30) --
	(174.15,305.78) --
	(176.86,307.28) --
	(179.56,308.79) --
	(182.26,310.32) --
	(184.97,311.86) --
	(187.67,313.42) --
	(190.37,315.01) --
	(193.08,316.61) --
	(195.78,318.23) --
	(198.48,319.88) --
	(201.19,321.54) --
	(203.89,323.23) --
	(206.59,324.94) --
	(209.30,326.68) --
	(212.00,328.43) --
	(214.70,330.22) --
	(217.40,332.02) --
	(220.11,333.85) --
	(222.81,335.71) --
	(225.51,337.58) --
	(228.22,339.48) --
	(230.92,341.40) --
	(233.62,343.34) --
	(236.33,345.30) --
	(239.03,347.28) --
	(241.73,349.29) --
	(244.44,351.32) --
	(247.14,353.37) --
	(249.84,355.45) --
	(252.55,357.54) --
	(255.25,359.66) --
	(257.95,361.80);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 35.86,249.88) --
	( 38.56,250.65) --
	( 41.26,251.43) --
	( 43.97,252.22) --
	( 46.67,253.03) --
	( 49.37,253.84) --
	( 52.07,254.67) --
	( 54.78,255.51) --
	( 57.48,256.37) --
	( 60.18,257.24) --
	( 62.89,258.11) --
	( 65.59,259.01) --
	( 68.29,259.91) --
	( 71.00,260.83) --
	( 73.70,261.76) --
	( 76.40,262.70) --
	( 79.11,263.65) --
	( 81.81,264.62) --
	( 84.51,265.60) --
	( 87.22,266.59) --
	( 89.92,267.59) --
	( 92.62,268.60) --
	( 95.33,269.63) --
	( 98.03,270.67) --
	(100.73,271.72) --
	(103.44,272.79) --
	(106.14,273.87) --
	(108.84,274.96) --
	(111.55,276.07) --
	(114.25,277.19) --
	(116.95,278.32) --
	(119.66,279.47) --
	(122.36,280.64) --
	(125.06,281.81) --
	(127.77,283.00) --
	(130.47,284.20) --
	(133.17,285.41) --
	(135.88,286.63) --
	(138.58,287.87) --
	(141.28,289.12) --
	(143.98,290.39) --
	(146.69,291.67) --
	(149.39,292.97) --
	(152.09,294.29) --
	(154.80,295.63) --
	(157.50,296.98) --
	(160.20,298.36) --
	(162.91,299.76) --
	(165.61,301.18) --
	(168.31,302.62) --
	(171.02,304.07) --
	(173.72,305.54) --
	(176.42,307.02) --
	(179.13,308.53) --
	(181.83,310.05) --
	(184.53,311.59) --
	(187.24,313.15) --
	(189.94,314.73) --
	(192.64,316.33) --
	(195.35,317.95) --
	(198.05,319.60) --
	(200.75,321.27) --
	(203.46,322.96) --
	(206.16,324.67) --
	(208.86,326.41) --
	(211.57,328.16) --
	(214.27,329.94) --
	(216.97,331.74) --
	(219.68,333.57) --
	(222.38,335.41) --
	(225.08,337.27) --
	(227.79,339.16) --
	(230.49,341.07) --
	(233.19,342.99) --
	(235.89,344.94) --
	(238.60,346.91) --
	(241.30,348.91) --
	(244.00,350.92) --
	(246.71,352.95) --
	(249.41,355.01);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 61.48,261.57) --
	( 64.85,255.75) --
	( 58.12,255.75) --
	cycle;

\path[fill=fillColor] (104.19,276.95) --
	(107.56,271.12) --
	(100.83,271.12) --
	cycle;

\path[fill=fillColor] (146.90,295.71) --
	(150.27,289.89) --
	(143.54,289.89) --
	cycle;

\path[fill=fillColor] (189.62,318.55) --
	(192.98,312.72) --
	(186.25,312.72) --
	cycle;

\path[fill=fillColor] (232.33,346.23) --
	(235.69,340.41) --
	(228.96,340.41) --
	cycle;

\path[fill=fillColor] (275.04,379.71) --
	(278.40,373.88) --
	(271.67,373.88) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 50.44,252.48) --
	( 55.44,252.48) --
	( 55.44,257.47) --
	( 50.44,257.47) --
	cycle;

\path[fill=fillColor] ( 93.15,267.25) --
	( 98.15,267.25) --
	( 98.15,272.24) --
	( 93.15,272.24) --
	cycle;

\path[fill=fillColor] (135.86,285.28) --
	(140.86,285.28) --
	(140.86,290.28) --
	(135.86,290.28) --
	cycle;

\path[fill=fillColor] (178.58,307.22) --
	(183.57,307.22) --
	(183.57,312.21) --
	(178.58,312.21) --
	cycle;

\path[fill=fillColor] (221.29,333.88) --
	(226.28,333.88) --
	(226.28,338.88) --
	(221.29,338.88) --
	cycle;

\path[fill=fillColor] (264.00,366.21) --
	(268.99,366.21) --
	(268.99,371.20) --
	(264.00,371.20) --
	cycle;
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87,252.37) -- ( 47.93,252.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.40,248.84) -- ( 44.40,255.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.58,266.54) -- ( 90.64,266.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.11,263.01) -- ( 87.11,270.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.29,283.90) -- (133.35,283.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.82,280.37) -- (129.82,287.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.00,304.89) -- (176.06,304.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.53,301.36) -- (172.53,308.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.71,330.58) -- (218.77,330.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.24,327.04) -- (215.24,334.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.42,361.80) -- (261.49,361.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.95,358.27) -- (257.95,365.33);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.36,247.39) rectangle ( 38.35,252.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.36,247.39) -- ( 38.35,252.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.36,252.38) -- ( 38.35,247.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.07,260.97) rectangle ( 81.06,265.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.07,260.97) -- ( 81.06,265.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.07,265.96) -- ( 81.06,260.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (118.78,277.67) rectangle (123.78,282.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (118.78,277.67) -- (123.78,282.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (118.78,282.67) -- (123.78,277.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.49,297.83) rectangle (166.49,302.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.49,297.83) -- (166.49,302.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.49,302.82) -- (166.49,297.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.20,322.52) rectangle (209.20,327.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.20,322.52) -- (209.20,327.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.20,327.51) -- (209.20,322.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.91,352.51) rectangle (251.91,357.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.91,352.51) -- (251.91,357.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.91,357.51) -- (251.91,352.51);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 27.31,247.50) circle (  2.50);

\path[fill=fillColor] ( 70.02,260.49) circle (  2.50);

\path[fill=fillColor] (112.74,276.48) circle (  2.50);

\path[fill=fillColor] (155.45,296.05) circle (  2.50);

\path[fill=fillColor] (198.16,319.89) circle (  2.50);

\path[fill=fillColor] (240.87,348.77) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31,247.50) rectangle (283.58,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,247.50) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,294.20) --
	( 27.31,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,340.91) --
	( 27.31,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,387.61) --
	( 27.31,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,244.75) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.74,244.75) --
	(112.74,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.16,244.75) --
	(198.16,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.74,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.16,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (155.45,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,329.23) {Basic reproduction number $\mathcal R_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 37.78,389.03) rectangle ( 52.24,403.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 39.23,396.25) -- ( 50.79,396.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 45.01,396.25) circle (  2.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 37.78,374.57) rectangle ( 52.24,389.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 39.23,381.80) -- ( 50.79,381.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 45.01,385.68) --
	( 48.37,379.86) --
	( 41.65,379.86) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 37.78,360.12) rectangle ( 52.24,374.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 39.23,367.35) -- ( 50.79,367.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 42.51,364.85) --
	( 47.51,364.85) --
	( 47.51,369.84) --
	( 42.51,369.84) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 37.78,345.67) rectangle ( 52.24,360.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 39.23,352.89) -- ( 50.79,352.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.48,352.89) -- ( 48.54,352.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.01,349.36) -- ( 45.01,356.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 37.78,331.21) rectangle ( 52.24,345.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 39.23,338.44) -- ( 50.79,338.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.51,335.94) rectangle ( 47.51,340.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.51,335.94) -- ( 47.51,340.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.51,340.94) -- ( 47.51,335.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 57.74,393.22) {Intrinsic generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 57.74,378.77) {Forward serial interval ($\rho=0.75$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 57.74,364.32) {Forward serial interval ($\rho=0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 57.74,349.86) {Forward serial interval ($\rho=0.25$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 57.74,335.41) {Forward serial interval ($\rho=0$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,419.03) {A};
\end{scope}
\begin{scope}
\path[clip] (289.08,216.81) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,216.81) rectangle (578.16,433.62);
\end{scope}
\begin{scope}
\path[clip] (316.39,247.50) rectangle (572.66,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (316.39,247.50) rectangle (572.66,410.96);
\definecolor{drawColor}{RGB}{13,8,135}

\path[draw=drawColor,line width= 1.1pt,line join=round] (316.39,247.64) --
	(319.64,248.49) --
	(322.88,249.36) --
	(326.12,250.25) --
	(329.37,251.16) --
	(332.61,252.09) --
	(335.86,253.03) --
	(339.10,254.00) --
	(342.34,254.99) --
	(345.59,255.99) --
	(348.83,257.01) --
	(352.08,258.06) --
	(355.32,259.12) --
	(358.56,260.20) --
	(361.81,261.30) --
	(365.05,262.42) --
	(368.30,263.56) --
	(371.54,264.72) --
	(374.78,265.90) --
	(378.03,267.10) --
	(381.27,268.32) --
	(384.51,269.55) --
	(387.76,270.81) --
	(391.00,272.08) --
	(394.25,273.37) --
	(397.49,274.68) --
	(400.73,276.01) --
	(403.98,277.36) --
	(407.22,278.74) --
	(410.47,280.13) --
	(413.71,281.55) --
	(416.95,282.99) --
	(420.20,284.44) --
	(423.44,285.91) --
	(426.69,287.40) --
	(429.93,288.91) --
	(433.17,290.45) --
	(436.42,292.02) --
	(439.66,293.61) --
	(442.90,295.23) --
	(446.15,296.88) --
	(449.39,298.55) --
	(452.64,300.24) --
	(455.88,301.95) --
	(459.12,303.68) --
	(462.37,305.44) --
	(465.61,307.23) --
	(468.86,309.04) --
	(472.10,310.89) --
	(475.34,312.76) --
	(478.59,314.68) --
	(481.83,316.62) --
	(485.08,318.59) --
	(488.32,320.58) --
	(491.56,322.60) --
	(494.81,324.66) --
	(498.05,326.74) --
	(501.29,328.85) --
	(504.54,330.99) --
	(507.78,333.17) --
	(511.03,335.39) --
	(514.27,337.63) --
	(517.51,339.91) --
	(520.76,342.22) --
	(524.00,344.56) --
	(527.25,346.93) --
	(530.49,349.33) --
	(533.73,351.76) --
	(536.98,354.22) --
	(540.22,356.72) --
	(543.47,359.24) --
	(546.71,361.80) --
	(549.95,364.38) --
	(553.20,367.00) --
	(556.44,369.66) --
	(559.68,372.34) --
	(562.93,375.05) --
	(566.17,377.80) --
	(569.42,380.57) --
	(572.66,383.38);
\definecolor{drawColor}{RGB}{126,3,168}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (319.64,248.37) --
	(322.88,249.29) --
	(326.12,250.21) --
	(329.37,251.14) --
	(332.61,252.09) --
	(335.86,253.04) --
	(339.10,254.01) --
	(342.34,254.98) --
	(345.59,255.97) --
	(348.83,256.97) --
	(352.08,257.98) --
	(355.32,259.00) --
	(358.56,260.04) --
	(361.81,261.08) --
	(365.05,262.13) --
	(368.30,263.20) --
	(371.54,264.27) --
	(374.78,265.36) --
	(378.03,266.46) --
	(381.27,267.57) --
	(384.51,268.69) --
	(387.76,269.82) --
	(391.00,270.96) --
	(394.25,272.12) --
	(397.49,273.28) --
	(400.73,274.45) --
	(403.98,275.64) --
	(407.22,276.84) --
	(410.47,278.04) --
	(413.71,279.26) --
	(416.95,280.49) --
	(420.20,281.73) --
	(423.44,282.98) --
	(426.69,284.25) --
	(429.93,285.52) --
	(433.17,286.81) --
	(436.42,288.10) --
	(439.66,289.40) --
	(442.90,290.71) --
	(446.15,292.02) --
	(449.39,293.35) --
	(452.64,294.69) --
	(455.88,296.04) --
	(459.12,297.40) --
	(462.37,298.77) --
	(465.61,300.15) --
	(468.86,301.54) --
	(472.10,302.93) --
	(475.34,304.32) --
	(478.59,305.72) --
	(481.83,307.12) --
	(485.08,308.54) --
	(488.32,309.96) --
	(491.56,311.39) --
	(494.81,312.82) --
	(498.05,314.26) --
	(501.29,315.71) --
	(504.54,317.16) --
	(507.78,318.61) --
	(511.03,320.07) --
	(514.27,321.53) --
	(517.51,322.99) --
	(520.76,324.45) --
	(524.00,325.93) --
	(527.25,327.40) --
	(530.49,328.89) --
	(533.73,330.37) --
	(536.98,331.86) --
	(540.22,333.36) --
	(543.47,334.86) --
	(546.71,336.37) --
	(549.95,337.88) --
	(553.20,339.39) --
	(556.44,340.91) --
	(559.68,342.43) --
	(562.93,343.96) --
	(566.17,345.49) --
	(569.42,347.03) --
	(572.66,348.57);
\definecolor{drawColor}{RGB}{204,70,120}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (319.64,248.30) --
	(322.88,249.24) --
	(326.12,250.18) --
	(329.37,251.12) --
	(332.61,252.08) --
	(335.86,253.04) --
	(339.10,254.00) --
	(342.34,254.98) --
	(345.59,255.96) --
	(348.83,256.94) --
	(352.08,257.93) --
	(355.32,258.93) --
	(358.56,259.94) --
	(361.81,260.95) --
	(365.05,261.97) --
	(368.30,262.99) --
	(371.54,264.02) --
	(374.78,265.05) --
	(378.03,266.09) --
	(381.27,267.14) --
	(384.51,268.20) --
	(387.76,269.26) --
	(391.00,270.34) --
	(394.25,271.42) --
	(397.49,272.50) --
	(400.73,273.59) --
	(403.98,274.68) --
	(407.22,275.78) --
	(410.47,276.87) --
	(413.71,277.98) --
	(416.95,279.10) --
	(420.20,280.22) --
	(423.44,281.36) --
	(426.69,282.50) --
	(429.93,283.64) --
	(433.17,284.78) --
	(436.42,285.92) --
	(439.66,287.06) --
	(442.90,288.19) --
	(446.15,289.31) --
	(449.39,290.45) --
	(452.64,291.59) --
	(455.88,292.74) --
	(459.12,293.89) --
	(462.37,295.05) --
	(465.61,296.19) --
	(468.86,297.34) --
	(472.10,298.47) --
	(475.34,299.59) --
	(478.59,300.69) --
	(481.83,301.79) --
	(485.08,302.89) --
	(488.32,303.98) --
	(491.56,305.07) --
	(494.81,306.15) --
	(498.05,307.23) --
	(501.29,308.29) --
	(504.54,309.34) --
	(507.78,310.37) --
	(511.03,311.39) --
	(514.27,312.39) --
	(517.51,313.37) --
	(520.76,314.35) --
	(524.00,315.32) --
	(527.25,316.28) --
	(530.49,317.22) --
	(533.73,318.16) --
	(536.98,319.09) --
	(540.22,320.00) --
	(543.47,320.91) --
	(546.71,321.80) --
	(549.95,322.67) --
	(553.20,323.54) --
	(556.44,324.40) --
	(559.68,325.24) --
	(562.93,326.07) --
	(566.17,326.89) --
	(569.42,327.70) --
	(572.66,328.50);
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (319.64,248.24) --
	(322.88,249.19) --
	(326.12,250.15) --
	(329.37,251.11) --
	(332.61,252.07) --
	(335.86,253.03) --
	(339.10,254.00) --
	(342.34,254.96) --
	(345.59,255.93) --
	(348.83,256.91) --
	(352.08,257.88) --
	(355.32,258.86) --
	(358.56,259.84) --
	(361.81,260.82) --
	(365.05,261.80) --
	(368.30,262.78) --
	(371.54,263.77) --
	(374.78,264.75) --
	(378.03,265.74) --
	(381.27,266.73) --
	(384.51,267.73) --
	(387.76,268.73) --
	(391.00,269.73) --
	(394.25,270.74) --
	(397.49,271.75) --
	(400.73,272.76) --
	(403.98,273.76) --
	(407.22,274.76) --
	(410.47,275.76) --
	(413.71,276.76) --
	(416.95,277.77) --
	(420.20,278.79) --
	(423.44,279.81) --
	(426.69,280.83) --
	(429.93,281.84) --
	(433.17,282.85) --
	(436.42,283.85) --
	(439.66,284.83) --
	(442.90,285.80) --
	(446.15,286.75) --
	(449.39,287.70) --
	(452.64,288.67) --
	(455.88,289.63) --
	(459.12,290.59) --
	(462.37,291.54) --
	(465.61,292.47) --
	(468.86,293.39) --
	(472.10,294.29) --
	(475.34,295.16) --
	(478.59,295.99) --
	(481.83,296.81) --
	(485.08,297.61) --
	(488.32,298.41) --
	(491.56,299.19) --
	(494.81,299.95) --
	(498.05,300.69) --
	(501.29,301.40) --
	(504.54,302.09) --
	(507.78,302.75) --
	(511.03,303.37) --
	(514.27,303.96) --
	(517.51,304.52) --
	(520.76,305.06) --
	(524.00,305.58) --
	(527.25,306.08) --
	(530.49,306.56) --
	(533.73,307.01) --
	(536.98,307.43) --
	(540.22,307.83) --
	(543.47,308.21) --
	(546.71,308.56) --
	(549.95,308.88) --
	(553.20,309.18) --
	(556.44,309.45) --
	(559.68,309.69) --
	(562.93,309.91) --
	(566.17,310.11) --
	(569.42,310.28) --
	(572.66,310.43);
\definecolor{drawColor}{RGB}{240,249,33}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] (319.64,248.18) --
	(322.88,249.15) --
	(326.12,250.12) --
	(329.37,251.09) --
	(332.61,252.06) --
	(335.86,253.02) --
	(339.10,253.99) --
	(342.34,254.95) --
	(345.59,255.91) --
	(348.83,256.87) --
	(352.08,257.82) --
	(355.32,258.78) --
	(358.56,259.73) --
	(361.81,260.68) --
	(365.05,261.63) --
	(368.30,262.58) --
	(371.54,263.52) --
	(374.78,264.45) --
	(378.03,265.39) --
	(381.27,266.33) --
	(384.51,267.27) --
	(387.76,268.21) --
	(391.00,269.15) --
	(394.25,270.09) --
	(397.49,271.03) --
	(400.73,271.96) --
	(403.98,272.88) --
	(407.22,273.79) --
	(410.47,274.70) --
	(413.71,275.60) --
	(416.95,276.51) --
	(420.20,277.43) --
	(423.44,278.34) --
	(426.69,279.25) --
	(429.93,280.15) --
	(433.17,281.03) --
	(436.42,281.90) --
	(439.66,282.74) --
	(442.90,283.56) --
	(446.15,284.35) --
	(449.39,285.14) --
	(452.64,285.93) --
	(455.88,286.72) --
	(459.12,287.50) --
	(462.37,288.26) --
	(465.61,289.01) --
	(468.86,289.72) --
	(472.10,290.40) --
	(475.34,291.04) --
	(478.59,291.64) --
	(481.83,292.21) --
	(485.08,292.76) --
	(488.32,293.29) --
	(491.56,293.79) --
	(494.81,294.27) --
	(498.05,294.72) --
	(501.29,295.13) --
	(504.54,295.51) --
	(507.78,295.84) --
	(511.03,296.12) --
	(514.27,296.36) --
	(517.51,296.57) --
	(520.76,296.74) --
	(524.00,296.88) --
	(527.25,296.99) --
	(530.49,297.06) --
	(533.73,297.10) --
	(536.98,297.11) --
	(540.22,297.08) --
	(543.47,297.02) --
	(546.71,296.91) --
	(549.95,296.77) --
	(553.20,296.60) --
	(556.44,296.39) --
	(559.68,296.14) --
	(562.93,295.86) --
	(566.17,295.54) --
	(569.42,295.19) --
	(572.66,294.80);
\definecolor{fillColor}{RGB}{126,3,168}

\path[fill=fillColor] (316.39,251.38) --
	(319.76,245.55) --
	(313.03,245.55) --
	cycle;

\path[fill=fillColor] (324.94,253.76) --
	(328.30,247.94) --
	(321.57,247.94) --
	cycle;

\path[fill=fillColor] (333.48,256.22) --
	(336.84,250.40) --
	(330.11,250.40) --
	cycle;

\path[fill=fillColor] (342.02,258.76) --
	(345.38,252.93) --
	(338.66,252.93) --
	cycle;

\path[fill=fillColor] (350.56,261.38) --
	(353.93,255.55) --
	(347.20,255.55) --
	cycle;

\path[fill=fillColor] (359.10,264.07) --
	(362.47,258.25) --
	(355.74,258.25) --
	cycle;

\path[fill=fillColor] (367.65,266.85) --
	(371.01,261.02) --
	(364.28,261.02) --
	cycle;

\path[fill=fillColor] (376.19,269.70) --
	(379.55,263.87) --
	(372.83,263.87) --
	cycle;

\path[fill=fillColor] (384.73,272.63) --
	(388.09,266.80) --
	(381.37,266.80) --
	cycle;

\path[fill=fillColor] (393.27,275.64) --
	(396.64,269.81) --
	(389.91,269.81) --
	cycle;

\path[fill=fillColor] (401.82,278.72) --
	(405.18,272.89) --
	(398.45,272.89) --
	cycle;

\path[fill=fillColor] (410.36,281.88) --
	(413.72,276.05) --
	(406.99,276.05) --
	cycle;

\path[fill=fillColor] (418.90,285.11) --
	(422.26,279.29) --
	(415.54,279.29) --
	cycle;

\path[fill=fillColor] (427.44,288.42) --
	(430.81,282.59) --
	(424.08,282.59) --
	cycle;

\path[fill=fillColor] (435.98,291.80) --
	(439.35,285.97) --
	(432.62,285.97) --
	cycle;

\path[fill=fillColor] (444.53,295.24) --
	(447.89,289.41) --
	(441.16,289.41) --
	cycle;

\path[fill=fillColor] (453.07,298.75) --
	(456.43,292.92) --
	(449.71,292.92) --
	cycle;

\path[fill=fillColor] (461.61,302.32) --
	(464.97,296.50) --
	(458.25,296.50) --
	cycle;

\path[fill=fillColor] (470.15,305.95) --
	(473.52,300.13) --
	(466.79,300.13) --
	cycle;

\path[fill=fillColor] (478.70,309.64) --
	(482.06,303.81) --
	(475.33,303.81) --
	cycle;

\path[fill=fillColor] (487.24,313.38) --
	(490.60,307.55) --
	(483.87,307.55) --
	cycle;

\path[fill=fillColor] (495.78,317.16) --
	(499.14,311.34) --
	(492.42,311.34) --
	cycle;

\path[fill=fillColor] (504.32,320.99) --
	(507.69,315.16) --
	(500.96,315.16) --
	cycle;

\path[fill=fillColor] (512.86,324.85) --
	(516.23,319.03) --
	(509.50,319.03) --
	cycle;

\path[fill=fillColor] (521.41,328.74) --
	(524.77,322.92) --
	(518.04,322.92) --
	cycle;

\path[fill=fillColor] (529.95,332.66) --
	(533.31,326.83) --
	(526.59,326.83) --
	cycle;

\path[fill=fillColor] (538.49,336.58) --
	(541.85,330.76) --
	(535.13,330.76) --
	cycle;

\path[fill=fillColor] (547.03,340.52) --
	(550.40,334.69) --
	(543.67,334.69) --
	cycle;

\path[fill=fillColor] (555.58,344.44) --
	(558.94,338.62) --
	(552.21,338.62) --
	cycle;

\path[fill=fillColor] (564.12,348.36) --
	(567.48,342.53) --
	(560.75,342.53) --
	cycle;

\path[fill=fillColor] (572.66,352.25) --
	(576.02,346.42) --
	(569.30,346.42) --
	cycle;
\definecolor{fillColor}{RGB}{204,70,120}

\path[fill=fillColor] (313.90,245.00) --
	(318.89,245.00) --
	(318.89,249.99) --
	(313.90,249.99) --
	cycle;

\path[fill=fillColor] (322.44,247.37) --
	(327.43,247.37) --
	(327.43,252.37) --
	(322.44,252.37) --
	cycle;

\path[fill=fillColor] (330.98,249.82) --
	(335.98,249.82) --
	(335.98,254.81) --
	(330.98,254.81) --
	cycle;

\path[fill=fillColor] (339.52,252.32) --
	(344.52,252.32) --
	(344.52,257.32) --
	(339.52,257.32) --
	cycle;

\path[fill=fillColor] (348.06,254.89) --
	(353.06,254.89) --
	(353.06,259.88) --
	(348.06,259.88) --
	cycle;

\path[fill=fillColor] (356.61,257.51) --
	(361.60,257.51) --
	(361.60,262.51) --
	(356.61,262.51) --
	cycle;

\path[fill=fillColor] (365.15,260.19) --
	(370.14,260.19) --
	(370.14,265.19) --
	(365.15,265.19) --
	cycle;

\path[fill=fillColor] (373.69,262.93) --
	(378.69,262.93) --
	(378.69,267.92) --
	(373.69,267.92) --
	cycle;

\path[fill=fillColor] (382.23,265.71) --
	(387.23,265.71) --
	(387.23,270.71) --
	(382.23,270.71) --
	cycle;

\path[fill=fillColor] (390.78,268.54) --
	(395.77,268.54) --
	(395.77,273.54) --
	(390.78,273.54) --
	cycle;

\path[fill=fillColor] (399.32,271.42) --
	(404.31,271.42) --
	(404.31,276.41) --
	(399.32,276.41) --
	cycle;

\path[fill=fillColor] (407.86,274.33) --
	(412.86,274.33) --
	(412.86,279.32) --
	(407.86,279.32) --
	cycle;

\path[fill=fillColor] (416.40,277.27) --
	(421.40,277.27) --
	(421.40,282.27) --
	(416.40,282.27) --
	cycle;

\path[fill=fillColor] (424.94,280.24) --
	(429.94,280.24) --
	(429.94,285.24) --
	(424.94,285.24) --
	cycle;

\path[fill=fillColor] (433.49,283.23) --
	(438.48,283.23) --
	(438.48,288.23) --
	(433.49,288.23) --
	cycle;

\path[fill=fillColor] (442.03,286.24) --
	(447.02,286.24) --
	(447.02,291.23) --
	(442.03,291.23) --
	cycle;

\path[fill=fillColor] (450.57,289.25) --
	(455.57,289.25) --
	(455.57,294.24) --
	(450.57,294.24) --
	cycle;

\path[fill=fillColor] (459.11,292.25) --
	(464.11,292.25) --
	(464.11,297.25) --
	(459.11,297.25) --
	cycle;

\path[fill=fillColor] (467.66,295.25) --
	(472.65,295.25) --
	(472.65,300.24) --
	(467.66,300.24) --
	cycle;

\path[fill=fillColor] (476.20,298.22) --
	(481.19,298.22) --
	(481.19,303.21) --
	(476.20,303.21) --
	cycle;

\path[fill=fillColor] (484.74,301.15) --
	(489.74,301.15) --
	(489.74,306.15) --
	(484.74,306.15) --
	cycle;

\path[fill=fillColor] (493.28,304.05) --
	(498.28,304.05) --
	(498.28,309.04) --
	(493.28,309.04) --
	cycle;

\path[fill=fillColor] (501.82,306.88) --
	(506.82,306.88) --
	(506.82,311.88) --
	(501.82,311.88) --
	cycle;

\path[fill=fillColor] (510.37,309.64) --
	(515.36,309.64) --
	(515.36,314.64) --
	(510.37,314.64) --
	cycle;

\path[fill=fillColor] (518.91,312.32) --
	(523.90,312.32) --
	(523.90,317.31) --
	(518.91,317.31) --
	cycle;

\path[fill=fillColor] (527.45,314.89) --
	(532.45,314.89) --
	(532.45,319.88) --
	(527.45,319.88) --
	cycle;

\path[fill=fillColor] (535.99,317.34) --
	(540.99,317.34) --
	(540.99,322.33) --
	(535.99,322.33) --
	cycle;

\path[fill=fillColor] (544.54,319.65) --
	(549.53,319.65) --
	(549.53,324.64) --
	(544.54,324.64) --
	cycle;

\path[fill=fillColor] (553.08,321.80) --
	(558.07,321.80) --
	(558.07,326.79) --
	(553.08,326.79) --
	cycle;

\path[fill=fillColor] (561.62,323.76) --
	(566.62,323.76) --
	(566.62,328.76) --
	(561.62,328.76) --
	cycle;

\path[fill=fillColor] (570.16,325.52) --
	(575.16,325.52) --
	(575.16,330.52) --
	(570.16,330.52) --
	cycle;
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.86,247.50) -- (319.93,247.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.39,243.96) -- (316.39,251.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.40,249.87) -- (328.47,249.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (324.94,246.33) -- (324.94,253.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.95,252.29) -- (337.01,252.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.48,248.76) -- (333.48,255.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (338.49,254.76) -- (345.55,254.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.02,251.23) -- (342.02,258.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (347.03,257.28) -- (354.09,257.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.56,253.74) -- (350.56,260.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (355.57,259.83) -- (362.64,259.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.10,256.30) -- (359.10,263.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.11,262.43) -- (371.18,262.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.65,258.89) -- (367.65,265.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.66,265.05) -- (379.72,265.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.19,261.52) -- (376.19,268.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.20,267.70) -- (388.26,267.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.73,264.16) -- (384.73,271.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.74,270.36) -- (396.81,270.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.27,266.83) -- (393.27,273.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.28,273.03) -- (405.35,273.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.82,269.50) -- (401.82,276.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.83,275.71) -- (413.89,275.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.36,272.18) -- (410.36,279.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (415.37,278.38) -- (422.43,278.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.90,274.85) -- (418.90,281.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.91,281.04) -- (430.97,281.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.44,277.51) -- (427.44,284.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (432.45,283.67) -- (439.52,283.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.98,280.13) -- (435.98,287.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (440.99,286.26) -- (448.06,286.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.53,282.73) -- (444.53,289.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (449.54,288.80) -- (456.60,288.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.07,285.27) -- (453.07,292.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.08,291.28) -- (465.14,291.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.61,287.75) -- (461.61,294.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (466.62,293.69) -- (473.69,293.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.15,290.15) -- (470.15,297.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (475.16,296.00) -- (482.23,296.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.70,292.47) -- (478.70,299.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.71,298.20) -- (490.77,298.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.24,294.67) -- (487.24,301.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (492.25,300.28) -- (499.31,300.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.78,296.75) -- (495.78,303.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (500.79,302.21) -- (507.85,302.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (504.32,298.68) -- (504.32,305.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (509.33,303.98) -- (516.40,303.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (512.86,300.45) -- (512.86,307.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (517.87,305.56) -- (524.94,305.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (521.41,302.03) -- (521.41,309.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.42,306.93) -- (533.48,306.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (529.95,303.40) -- (529.95,310.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (534.96,308.07) -- (542.02,308.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (538.49,304.54) -- (538.49,311.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (543.50,308.95) -- (550.57,308.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.03,305.42) -- (547.03,312.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.04,309.54) -- (559.11,309.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (555.58,306.01) -- (555.58,313.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.59,309.82) -- (567.65,309.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.12,306.29) -- (564.12,313.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13,309.77) -- (576.19,309.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66,306.24) -- (572.66,313.30);
\definecolor{drawColor}{RGB}{240,249,33}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.90,245.00) rectangle (318.89,249.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.90,245.00) -- (318.89,249.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.90,249.99) -- (318.89,245.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.44,247.36) rectangle (327.43,252.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.44,247.36) -- (327.43,252.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.44,252.36) -- (327.43,247.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (330.98,249.77) rectangle (335.98,254.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (330.98,249.77) -- (335.98,254.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (330.98,254.76) -- (335.98,249.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (339.52,252.21) rectangle (344.52,257.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (339.52,252.21) -- (344.52,257.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (339.52,257.20) -- (344.52,252.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (348.06,254.68) rectangle (353.06,259.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (348.06,254.68) -- (353.06,259.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (348.06,259.67) -- (353.06,254.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.61,257.17) rectangle (361.60,262.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.61,257.17) -- (361.60,262.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.61,262.16) -- (361.60,257.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.15,259.67) rectangle (370.14,264.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.15,259.67) -- (370.14,264.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.15,264.67) -- (370.14,259.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (373.69,262.19) rectangle (378.69,267.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (373.69,262.19) -- (378.69,267.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (373.69,267.19) -- (378.69,262.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.23,264.71) rectangle (387.23,269.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.23,264.71) -- (387.23,269.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.23,269.70) -- (387.23,264.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (390.78,267.21) rectangle (395.77,272.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (390.78,267.21) -- (395.77,272.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (390.78,272.21) -- (395.77,267.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (399.32,269.70) rectangle (404.31,274.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (399.32,269.70) -- (404.31,274.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (399.32,274.69) -- (404.31,269.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (407.86,272.15) rectangle (412.86,277.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (407.86,272.15) -- (412.86,277.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (407.86,277.15) -- (412.86,272.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (416.40,274.57) rectangle (421.40,279.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (416.40,274.57) -- (421.40,279.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (416.40,279.56) -- (421.40,274.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.94,276.93) rectangle (429.94,281.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.94,276.93) -- (429.94,281.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.94,281.93) -- (429.94,276.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.49,279.22) rectangle (438.48,284.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.49,279.22) -- (438.48,284.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.49,284.22) -- (438.48,279.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (442.03,281.44) rectangle (447.02,286.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (442.03,281.44) -- (447.02,286.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (442.03,286.43) -- (447.02,281.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (450.57,283.55) rectangle (455.57,288.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (450.57,283.55) -- (455.57,288.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (450.57,288.55) -- (455.57,283.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.11,285.55) rectangle (464.11,290.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.11,285.55) -- (464.11,290.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.11,290.55) -- (464.11,285.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.66,287.42) rectangle (472.65,292.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.66,287.42) -- (472.65,292.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.66,292.42) -- (472.65,287.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.20,289.14) rectangle (481.19,294.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.20,289.14) -- (481.19,294.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.20,294.14) -- (481.19,289.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (484.74,290.69) rectangle (489.74,295.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (484.74,290.69) -- (489.74,295.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (484.74,295.69) -- (489.74,290.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (493.28,292.05) rectangle (498.28,297.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (493.28,292.05) -- (498.28,297.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (493.28,297.04) -- (498.28,292.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.82,293.19) rectangle (506.82,298.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.82,293.19) -- (506.82,298.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.82,298.19) -- (506.82,293.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.37,294.09) rectangle (515.36,299.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.37,294.09) -- (515.36,299.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.37,299.09) -- (515.36,294.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (518.91,294.73) rectangle (523.90,299.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (518.91,294.73) -- (523.90,299.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (518.91,299.73) -- (523.90,294.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (527.45,295.08) rectangle (532.45,300.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (527.45,295.08) -- (532.45,300.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (527.45,300.07) -- (532.45,295.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.99,295.11) rectangle (540.99,300.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.99,295.11) -- (540.99,300.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.99,300.11) -- (540.99,295.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.54,294.80) rectangle (549.53,299.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.54,294.80) -- (549.53,299.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.54,299.79) -- (549.53,294.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.08,294.12) rectangle (558.07,299.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.08,294.12) -- (558.07,299.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.08,299.12) -- (558.07,294.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.62,293.05) rectangle (566.62,298.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.62,293.05) -- (566.62,298.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.62,298.05) -- (566.62,293.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16,291.57) rectangle (575.16,296.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16,291.57) -- (575.16,296.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16,296.57) -- (575.16,291.57);
\definecolor{fillColor}{RGB}{13,8,135}

\path[fill=fillColor] (316.39,247.50) circle (  2.50);

\path[fill=fillColor] (324.94,249.88) circle (  2.50);

\path[fill=fillColor] (333.48,252.37) circle (  2.50);

\path[fill=fillColor] (342.02,254.96) circle (  2.50);

\path[fill=fillColor] (350.56,257.67) circle (  2.50);

\path[fill=fillColor] (359.10,260.49) circle (  2.50);

\path[fill=fillColor] (367.65,263.43) circle (  2.50);

\path[fill=fillColor] (376.19,266.50) circle (  2.50);

\path[fill=fillColor] (384.73,269.69) circle (  2.50);

\path[fill=fillColor] (393.27,273.02) circle (  2.50);

\path[fill=fillColor] (401.82,276.48) circle (  2.50);

\path[fill=fillColor] (410.36,280.09) circle (  2.50);

\path[fill=fillColor] (418.90,283.85) circle (  2.50);

\path[fill=fillColor] (427.44,287.76) circle (  2.50);

\path[fill=fillColor] (435.98,291.82) circle (  2.50);

\path[fill=fillColor] (444.53,296.05) circle (  2.50);

\path[fill=fillColor] (453.07,300.46) circle (  2.50);

\path[fill=fillColor] (461.61,305.04) circle (  2.50);

\path[fill=fillColor] (470.15,309.80) circle (  2.50);

\path[fill=fillColor] (478.70,314.74) circle (  2.50);

\path[fill=fillColor] (487.24,319.89) circle (  2.50);

\path[fill=fillColor] (495.78,325.23) circle (  2.50);

\path[fill=fillColor] (504.32,330.79) circle (  2.50);

\path[fill=fillColor] (512.86,336.55) circle (  2.50);

\path[fill=fillColor] (521.41,342.55) circle (  2.50);

\path[fill=fillColor] (529.95,348.77) circle (  2.50);

\path[fill=fillColor] (538.49,355.23) circle (  2.50);

\path[fill=fillColor] (547.03,361.93) circle (  2.50);

\path[fill=fillColor] (555.58,368.89) circle (  2.50);

\path[fill=fillColor] (564.12,376.12) circle (  2.50);

\path[fill=fillColor] (572.66,383.61) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.39,247.50) rectangle (572.66,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,247.50) --
	(316.39,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,294.20) --
	(316.39,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,340.91) --
	(316.39,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,387.61) --
	(316.39,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (316.39,244.75) --
	(316.39,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.82,244.75) --
	(401.82,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.24,244.75) --
	(487.24,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (316.39,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (401.82,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (487.24,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (444.53,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,329.23) {Reproduction number ${\mathcal R}_{\textrm{\tiny intrinsic}}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (326.79,389.03) rectangle (341.24,403.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{13,8,135}

\path[draw=drawColor,line width= 1.1pt,line join=round] (328.24,396.25) -- (339.80,396.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{13,8,135}

\path[fill=fillColor] (334.02,396.25) circle (  2.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (326.79,374.57) rectangle (341.24,389.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{126,3,168}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (328.24,381.80) -- (339.80,381.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{126,3,168}

\path[fill=fillColor] (334.02,385.68) --
	(337.38,379.86) --
	(330.65,379.86) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (326.79,360.12) rectangle (341.24,374.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{204,70,120}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (328.24,367.35) -- (339.80,367.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{204,70,120}

\path[fill=fillColor] (331.52,364.85) --
	(336.52,364.85) --
	(336.52,369.84) --
	(331.52,369.84) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (326.79,345.67) rectangle (341.24,360.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (328.24,352.89) -- (339.80,352.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (330.49,352.89) -- (337.55,352.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.02,349.36) -- (334.02,356.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (326.79,331.21) rectangle (341.24,345.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{240,249,33}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] (328.24,338.44) -- (339.80,338.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{240,249,33}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (331.52,335.94) rectangle (336.52,340.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (331.52,335.94) -- (336.52,340.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (331.52,340.94) -- (336.52,335.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (346.74,393.22) {Intrinsic generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (346.74,378.77) {Intrinsic serial interval ($\rho=0.75$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (346.74,364.32) {Intrinsic serial interval ($\rho=0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (346.74,349.86) {Intrinsic serial interval ($\rho=0.25$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (346.74,335.41) {Intrinsic serial interval ($\rho=0$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (316.39,419.03) {B};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 27.31, 30.69) rectangle (283.58,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31, 30.69) rectangle (283.58,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31, 63.20) --
	( 30.56, 63.20) --
	( 33.80, 63.20) --
	( 37.04, 63.20) --
	( 40.29, 63.20) --
	( 43.53, 63.20) --
	( 46.78, 63.20) --
	( 50.02, 63.20) --
	( 53.26, 63.20) --
	( 56.51, 63.20) --
	( 59.75, 63.20) --
	( 63.00, 63.20) --
	( 66.24, 63.20) --
	( 69.48, 63.20) --
	( 72.73, 63.20) --
	( 75.97, 63.20) --
	( 79.22, 63.20) --
	( 82.46, 63.20) --
	( 85.70, 63.20) --
	( 88.95, 63.20) --
	( 92.19, 63.20) --
	( 95.43, 63.20) --
	( 98.68, 63.20) --
	(101.92, 63.20) --
	(105.17, 63.20) --
	(108.41, 63.20) --
	(111.65, 63.20) --
	(114.90, 63.20) --
	(118.14, 63.20) --
	(121.39, 63.20) --
	(124.63, 63.20) --
	(127.87, 63.20) --
	(131.12, 63.20) --
	(134.36, 63.20) --
	(137.61, 63.20) --
	(140.85, 63.20) --
	(144.09, 63.20) --
	(147.34, 63.20) --
	(150.58, 63.20) --
	(153.82, 63.20) --
	(157.07, 63.20) --
	(160.31, 63.20) --
	(163.56, 63.20) --
	(166.80, 63.20) --
	(170.04, 63.20) --
	(173.29, 63.20) --
	(176.53, 63.20) --
	(179.78, 63.20) --
	(183.02, 63.20) --
	(186.26, 63.20) --
	(189.51, 63.20) --
	(192.75, 63.20) --
	(196.00, 63.20) --
	(199.24, 63.20) --
	(202.48, 63.20) --
	(205.73, 63.20) --
	(208.97, 63.20) --
	(212.21, 63.20) --
	(215.46, 63.20) --
	(218.70, 63.20) --
	(221.95, 63.20) --
	(225.19, 63.20) --
	(228.43, 63.20) --
	(231.68, 63.20) --
	(234.92, 63.20) --
	(238.17, 63.20) --
	(241.41, 63.20) --
	(244.65, 63.20) --
	(247.90, 63.20) --
	(251.14, 63.20) --
	(254.39, 63.20) --
	(257.63, 63.20) --
	(260.87, 63.20) --
	(264.12, 63.20) --
	(267.36, 63.20) --
	(270.60, 63.20) --
	(273.85, 63.20) --
	(277.09, 63.20) --
	(280.34, 63.20) --
	(283.58, 63.20);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 27.31, 64.16) --
	( 30.56, 65.16) --
	( 33.80, 66.15) --
	( 37.04, 67.13) --
	( 40.29, 68.09) --
	( 43.53, 69.05) --
	( 46.78, 69.99) --
	( 50.02, 70.92) --
	( 53.26, 71.84) --
	( 56.51, 72.75) --
	( 59.75, 73.64) --
	( 63.00, 74.53) --
	( 66.24, 75.40) --
	( 69.48, 76.26) --
	( 72.73, 77.11) --
	( 75.97, 77.95) --
	( 79.22, 78.78) --
	( 82.46, 79.59) --
	( 85.70, 80.40) --
	( 88.95, 81.20) --
	( 92.19, 81.98) --
	( 95.43, 82.75) --
	( 98.68, 83.50) --
	(101.92, 84.24) --
	(105.17, 84.98) --
	(108.41, 85.70) --
	(111.65, 86.41) --
	(114.90, 87.11) --
	(118.14, 87.80) --
	(121.39, 88.49) --
	(124.63, 89.16) --
	(127.87, 89.82) --
	(131.12, 90.46) --
	(134.36, 91.09) --
	(137.61, 91.71) --
	(140.85, 92.33) --
	(144.09, 92.93) --
	(147.34, 93.54) --
	(150.58, 94.13) --
	(153.82, 94.73) --
	(157.07, 95.33) --
	(160.31, 95.92) --
	(163.56, 96.50) --
	(166.80, 97.08) --
	(170.04, 97.64) --
	(173.29, 98.20) --
	(176.53, 98.76) --
	(179.78, 99.31) --
	(183.02, 99.86) --
	(186.26,100.41) --
	(189.51,100.95) --
	(192.75,101.49) --
	(196.00,102.03) --
	(199.24,102.56) --
	(202.48,103.08) --
	(205.73,103.60) --
	(208.97,104.12) --
	(212.21,104.63) --
	(215.46,105.14) --
	(218.70,105.65) --
	(221.95,106.15) --
	(225.19,106.65) --
	(228.43,107.14) --
	(231.68,107.63) --
	(234.92,108.12) --
	(238.17,108.60) --
	(241.41,109.07) --
	(244.65,109.54) --
	(247.90,110.01) --
	(251.14,110.48) --
	(254.39,110.94) --
	(257.63,111.39) --
	(260.87,111.84) --
	(264.12,112.29) --
	(267.36,112.73) --
	(270.60,113.17) --
	(273.85,113.61) --
	(277.09,114.04) --
	(280.34,114.46) --
	(283.58,114.89);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 27.31, 64.16) --
	( 30.56, 65.64) --
	( 33.80, 67.11) --
	( 37.04, 68.56) --
	( 40.29, 69.99) --
	( 43.53, 71.41) --
	( 46.78, 72.81) --
	( 50.02, 74.19) --
	( 53.26, 75.56) --
	( 56.51, 76.91) --
	( 59.75, 78.25) --
	( 63.00, 79.56) --
	( 66.24, 80.86) --
	( 69.48, 82.15) --
	( 72.73, 83.42) --
	( 75.97, 84.67) --
	( 79.22, 85.91) --
	( 82.46, 87.13) --
	( 85.70, 88.34) --
	( 88.95, 89.53) --
	( 92.19, 90.70) --
	( 95.43, 91.85) --
	( 98.68, 92.98) --
	(101.92, 94.10) --
	(105.17, 95.20) --
	(108.41, 96.29) --
	(111.65, 97.36) --
	(114.90, 98.41) --
	(118.14, 99.46) --
	(121.39,100.49) --
	(124.63,101.51) --
	(127.87,102.50) --
	(131.12,103.47) --
	(134.36,104.43) --
	(137.61,105.37) --
	(140.85,106.29) --
	(144.09,107.21) --
	(147.34,108.12) --
	(150.58,109.03) --
	(153.82,109.94) --
	(157.07,110.85) --
	(160.31,111.75) --
	(163.56,112.63) --
	(166.80,113.50) --
	(170.04,114.37) --
	(173.29,115.22) --
	(176.53,116.07) --
	(179.78,116.92) --
	(183.02,117.75) --
	(186.26,118.59) --
	(189.51,119.42) --
	(192.75,120.24) --
	(196.00,121.06) --
	(199.24,121.87) --
	(202.48,122.67) --
	(205.73,123.47) --
	(208.97,124.26) --
	(212.21,125.04) --
	(215.46,125.82) --
	(218.70,126.59) --
	(221.95,127.36) --
	(225.19,128.12) --
	(228.43,128.87) --
	(231.68,129.62) --
	(234.92,130.36) --
	(238.17,131.10) --
	(241.41,131.83) --
	(244.65,132.55) --
	(247.90,133.26) --
	(251.14,133.97) --
	(254.39,134.68) --
	(257.63,135.37) --
	(260.87,136.07) --
	(264.12,136.75) --
	(267.36,137.43) --
	(270.60,138.10) --
	(273.85,138.77) --
	(277.09,139.43) --
	(280.34,140.08) --
	(283.58,140.73);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 64.08) --
	( 30.56, 66.04) --
	( 33.80, 67.97) --
	( 37.04, 69.88) --
	( 40.29, 71.77) --
	( 43.53, 73.64) --
	( 46.78, 75.49) --
	( 50.02, 77.31) --
	( 53.26, 79.12) --
	( 56.51, 80.91) --
	( 59.75, 82.67) --
	( 63.00, 84.42) --
	( 66.24, 86.14) --
	( 69.48, 87.84) --
	( 72.73, 89.53) --
	( 75.97, 91.19) --
	( 79.22, 92.83) --
	( 82.46, 94.46) --
	( 85.70, 96.06) --
	( 88.95, 97.65) --
	( 92.19, 99.21) --
	( 95.43,100.75) --
	( 98.68,102.26) --
	(101.92,103.75) --
	(105.17,105.23) --
	(108.41,106.68) --
	(111.65,108.11) --
	(114.90,109.53) --
	(118.14,110.93) --
	(121.39,112.32) --
	(124.63,113.68) --
	(127.87,115.02) --
	(131.12,116.33) --
	(134.36,117.61) --
	(137.61,118.88) --
	(140.85,120.13) --
	(144.09,121.37) --
	(147.34,122.61) --
	(150.58,123.84) --
	(153.82,125.07) --
	(157.07,126.29) --
	(160.31,127.51) --
	(163.56,128.71) --
	(166.80,129.90) --
	(170.04,131.07) --
	(173.29,132.24) --
	(176.53,133.40) --
	(179.78,134.54) --
	(183.02,135.68) --
	(186.26,136.82) --
	(189.51,137.95) --
	(192.75,139.07) --
	(196.00,140.19) --
	(199.24,141.29) --
	(202.48,142.39) --
	(205.73,143.48) --
	(208.97,144.56) --
	(212.21,145.63) --
	(215.46,146.69) --
	(218.70,147.75) --
	(221.95,148.80) --
	(225.19,149.84) --
	(228.43,150.87) --
	(231.68,151.89) --
	(234.92,152.91) --
	(238.17,153.91) --
	(241.41,154.91) --
	(244.65,155.90) --
	(247.90,156.88) --
	(251.14,157.85) --
	(254.39,158.82) --
	(257.63,159.78) --
	(260.87,160.73) --
	(264.12,161.67) --
	(267.36,162.60) --
	(270.60,163.52) --
	(273.85,164.44) --
	(277.09,165.35) --
	(280.34,166.24) --
	(283.58,167.14);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 27.31, 63.95) --
	( 30.56, 66.36) --
	( 33.80, 68.74) --
	( 37.04, 71.10) --
	( 40.29, 73.43) --
	( 43.53, 75.74) --
	( 46.78, 78.03) --
	( 50.02, 80.29) --
	( 53.26, 82.53) --
	( 56.51, 84.74) --
	( 59.75, 86.93) --
	( 63.00, 89.09) --
	( 66.24, 91.23) --
	( 69.48, 93.35) --
	( 72.73, 95.44) --
	( 75.97, 97.51) --
	( 79.22, 99.55) --
	( 82.46,101.58) --
	( 85.70,103.58) --
	( 88.95,105.56) --
	( 92.19,107.51) --
	( 95.43,109.43) --
	( 98.68,111.32) --
	(101.92,113.19) --
	(105.17,115.04) --
	(108.41,116.86) --
	(111.65,118.67) --
	(114.90,120.45) --
	(118.14,122.21) --
	(121.39,123.96) --
	(124.63,125.67) --
	(127.87,127.35) --
	(131.12,129.01) --
	(134.36,130.63) --
	(137.61,132.24) --
	(140.85,133.82) --
	(144.09,135.39) --
	(147.34,136.96) --
	(150.58,138.52) --
	(153.82,140.07) --
	(157.07,141.63) --
	(160.31,143.18) --
	(163.56,144.71) --
	(166.80,146.22) --
	(170.04,147.72) --
	(173.29,149.21) --
	(176.53,150.69) --
	(179.78,152.15) --
	(183.02,153.61) --
	(186.26,155.06) --
	(189.51,156.51) --
	(192.75,157.94) --
	(196.00,159.37) --
	(199.24,160.79) --
	(202.48,162.19) --
	(205.73,163.58) --
	(208.97,164.97) --
	(212.21,166.34) --
	(215.46,167.71) --
	(218.70,169.06) --
	(221.95,170.41) --
	(225.19,171.74) --
	(228.43,173.07) --
	(231.68,174.38) --
	(234.92,175.69) --
	(238.17,176.98) --
	(241.41,178.27) --
	(244.65,179.54) --
	(247.90,180.81) --
	(251.14,182.06) --
	(254.39,183.31) --
	(257.63,184.54) --
	(260.87,185.77) --
	(264.12,186.98) --
	(267.36,188.19) --
	(270.60,189.38) --
	(273.85,190.57) --
	(277.09,191.74) --
	(280.34,192.91) --
	(283.58,194.06);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] ( 27.31, 68.04) --
	( 30.68, 62.21) --
	( 23.95, 62.21) --
	cycle;

\path[fill=fillColor] ( 35.86, 70.65) --
	( 39.22, 64.82) --
	( 32.49, 64.82) --
	cycle;

\path[fill=fillColor] ( 44.40, 73.00) --
	( 47.76, 67.17) --
	( 41.03, 67.17) --
	cycle;

\path[fill=fillColor] ( 52.94, 75.66) --
	( 56.30, 69.83) --
	( 49.58, 69.83) --
	cycle;

\path[fill=fillColor] ( 61.48, 78.15) --
	( 64.85, 72.32) --
	( 58.12, 72.32) --
	cycle;

\path[fill=fillColor] ( 70.02, 80.38) --
	( 73.39, 74.55) --
	( 66.66, 74.55) --
	cycle;

\path[fill=fillColor] ( 78.57, 82.50) --
	( 81.93, 76.67) --
	( 75.20, 76.67) --
	cycle;

\path[fill=fillColor] ( 87.11, 84.65) --
	( 90.47, 78.82) --
	( 83.75, 78.82) --
	cycle;

\path[fill=fillColor] ( 95.65, 86.82) --
	( 99.01, 80.99) --
	( 92.29, 80.99) --
	cycle;

\path[fill=fillColor] (104.19, 88.79) --
	(107.56, 82.96) --
	(100.83, 82.96) --
	cycle;

\path[fill=fillColor] (112.74, 90.70) --
	(116.10, 84.87) --
	(109.37, 84.87) --
	cycle;

\path[fill=fillColor] (121.28, 92.30) --
	(124.64, 86.48) --
	(117.91, 86.48) --
	cycle;

\path[fill=fillColor] (129.82, 94.04) --
	(133.18, 88.21) --
	(126.46, 88.21) --
	cycle;

\path[fill=fillColor] (138.36, 95.68) --
	(141.73, 89.85) --
	(135.00, 89.85) --
	cycle;

\path[fill=fillColor] (146.90, 97.35) --
	(150.27, 91.53) --
	(143.54, 91.53) --
	cycle;

\path[fill=fillColor] (155.45, 98.90) --
	(158.81, 93.07) --
	(152.08, 93.07) --
	cycle;

\path[fill=fillColor] (163.99,100.37) --
	(167.35, 94.54) --
	(160.63, 94.54) --
	cycle;

\path[fill=fillColor] (172.53,101.89) --
	(175.89, 96.06) --
	(169.17, 96.06) --
	cycle;

\path[fill=fillColor] (181.07,103.46) --
	(184.44, 97.64) --
	(177.71, 97.64) --
	cycle;

\path[fill=fillColor] (189.62,104.87) --
	(192.98, 99.04) --
	(186.25, 99.04) --
	cycle;

\path[fill=fillColor] (198.16,106.29) --
	(201.52,100.47) --
	(194.79,100.47) --
	cycle;

\path[fill=fillColor] (206.70,107.65) --
	(210.06,101.82) --
	(203.34,101.82) --
	cycle;

\path[fill=fillColor] (215.24,108.93) --
	(218.61,103.11) --
	(211.88,103.11) --
	cycle;

\path[fill=fillColor] (223.78,110.30) --
	(227.15,104.48) --
	(220.42,104.48) --
	cycle;

\path[fill=fillColor] (232.33,111.62) --
	(235.69,105.79) --
	(228.96,105.79) --
	cycle;

\path[fill=fillColor] (240.87,112.82) --
	(244.23,106.99) --
	(237.51,106.99) --
	cycle;

\path[fill=fillColor] (249.41,114.10) --
	(252.77,108.28) --
	(246.05,108.28) --
	cycle;

\path[fill=fillColor] (257.95,115.26) --
	(261.32,109.43) --
	(254.59,109.43) --
	cycle;

\path[fill=fillColor] (266.50,116.55) --
	(269.86,110.72) --
	(263.13,110.72) --
	cycle;

\path[fill=fillColor] (275.04,117.65) --
	(278.40,111.83) --
	(271.67,111.83) --
	cycle;

\path[fill=fillColor] (283.58,118.80) --
	(286.94,112.97) --
	(280.22,112.97) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] ( 24.82, 61.68) --
	( 29.81, 61.68) --
	( 29.81, 66.68) --
	( 24.82, 66.68) --
	cycle;

\path[fill=fillColor] ( 33.36, 65.53) --
	( 38.35, 65.53) --
	( 38.35, 70.52) --
	( 33.36, 70.52) --
	cycle;

\path[fill=fillColor] ( 41.90, 69.01) --
	( 46.90, 69.01) --
	( 46.90, 74.00) --
	( 41.90, 74.00) --
	cycle;

\path[fill=fillColor] ( 50.44, 72.95) --
	( 55.44, 72.95) --
	( 55.44, 77.95) --
	( 50.44, 77.95) --
	cycle;

\path[fill=fillColor] ( 58.98, 76.64) --
	( 63.98, 76.64) --
	( 63.98, 81.63) --
	( 58.98, 81.63) --
	cycle;

\path[fill=fillColor] ( 67.53, 79.97) --
	( 72.52, 79.97) --
	( 72.52, 84.96) --
	( 67.53, 84.96) --
	cycle;

\path[fill=fillColor] ( 76.07, 83.16) --
	( 81.06, 83.16) --
	( 81.06, 88.16) --
	( 76.07, 88.16) --
	cycle;

\path[fill=fillColor] ( 84.61, 86.41) --
	( 89.61, 86.41) --
	( 89.61, 91.41) --
	( 84.61, 91.41) --
	cycle;

\path[fill=fillColor] ( 93.15, 89.65) --
	( 98.15, 89.65) --
	( 98.15, 94.65) --
	( 93.15, 94.65) --
	cycle;

\path[fill=fillColor] (101.70, 92.60) --
	(106.69, 92.60) --
	(106.69, 97.60) --
	(101.70, 97.60) --
	cycle;

\path[fill=fillColor] (110.24, 95.42) --
	(115.23, 95.42) --
	(115.23,100.42) --
	(110.24,100.42) --
	cycle;

\path[fill=fillColor] (118.78, 97.90) --
	(123.78, 97.90) --
	(123.78,102.89) --
	(118.78,102.89) --
	cycle;

\path[fill=fillColor] (127.32,100.49) --
	(132.32,100.49) --
	(132.32,105.49) --
	(127.32,105.49) --
	cycle;

\path[fill=fillColor] (135.86,103.01) --
	(140.86,103.01) --
	(140.86,108.00) --
	(135.86,108.00) --
	cycle;

\path[fill=fillColor] (144.41,105.51) --
	(149.40,105.51) --
	(149.40,110.51) --
	(144.41,110.51) --
	cycle;

\path[fill=fillColor] (152.95,107.88) --
	(157.94,107.88) --
	(157.94,112.88) --
	(152.95,112.88) --
	cycle;

\path[fill=fillColor] (161.49,110.12) --
	(166.49,110.12) --
	(166.49,115.11) --
	(161.49,115.11) --
	cycle;

\path[fill=fillColor] (170.03,112.40) --
	(175.03,112.40) --
	(175.03,117.39) --
	(170.03,117.39) --
	cycle;

\path[fill=fillColor] (178.58,114.82) --
	(183.57,114.82) --
	(183.57,119.82) --
	(178.58,119.82) --
	cycle;

\path[fill=fillColor] (187.12,116.97) --
	(192.11,116.97) --
	(192.11,121.96) --
	(187.12,121.96) --
	cycle;

\path[fill=fillColor] (195.66,119.16) --
	(200.66,119.16) --
	(200.66,124.15) --
	(195.66,124.15) --
	cycle;

\path[fill=fillColor] (204.20,121.21) --
	(209.20,121.21) --
	(209.20,126.20) --
	(204.20,126.20) --
	cycle;

\path[fill=fillColor] (212.74,123.21) --
	(217.74,123.21) --
	(217.74,128.21) --
	(212.74,128.21) --
	cycle;

\path[fill=fillColor] (221.29,125.29) --
	(226.28,125.29) --
	(226.28,130.28) --
	(221.29,130.28) --
	cycle;

\path[fill=fillColor] (229.83,127.27) --
	(234.82,127.27) --
	(234.82,132.26) --
	(229.83,132.26) --
	cycle;

\path[fill=fillColor] (238.37,129.12) --
	(243.37,129.12) --
	(243.37,134.11) --
	(238.37,134.11) --
	cycle;

\path[fill=fillColor] (246.91,131.10) --
	(251.91,131.10) --
	(251.91,136.09) --
	(246.91,136.09) --
	cycle;

\path[fill=fillColor] (255.46,132.88) --
	(260.45,132.88) --
	(260.45,137.87) --
	(255.46,137.87) --
	cycle;

\path[fill=fillColor] (264.00,134.82) --
	(268.99,134.82) --
	(268.99,139.81) --
	(264.00,139.81) --
	cycle;

\path[fill=fillColor] (272.54,136.50) --
	(277.54,136.50) --
	(277.54,141.49) --
	(272.54,141.49) --
	cycle;

\path[fill=fillColor] (281.08,138.27) --
	(286.08,138.27) --
	(286.08,143.27) --
	(281.08,143.27) --
	cycle;
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 23.78, 64.15) -- ( 30.85, 64.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.31, 60.62) -- ( 27.31, 67.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.32, 69.18) -- ( 39.39, 69.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.86, 65.65) -- ( 35.86, 72.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.87, 73.76) -- ( 47.93, 73.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.40, 70.23) -- ( 44.40, 77.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.41, 78.96) -- ( 56.47, 78.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.94, 75.42) -- ( 52.94, 82.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.95, 83.79) -- ( 65.01, 83.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.48, 80.26) -- ( 61.48, 87.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.49, 88.22) -- ( 73.56, 88.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 70.02, 84.69) -- ( 70.02, 91.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.03, 92.50) -- ( 82.10, 92.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.57, 88.97) -- ( 78.57, 96.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.58, 96.86) -- ( 90.64, 96.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 87.11, 93.33) -- ( 87.11,100.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.12,101.17) -- ( 99.18,101.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65, 97.63) -- ( 95.65,104.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.66,105.11) -- (107.73,105.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.19,101.57) -- (104.19,108.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.20,108.80) -- (116.27,108.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (112.74,105.26) -- (112.74,112.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.75,112.21) -- (124.81,112.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.28,108.68) -- (121.28,115.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (126.29,115.67) -- (133.35,115.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.82,112.14) -- (129.82,119.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.83,119.09) -- (141.89,119.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (138.36,115.56) -- (138.36,122.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (143.37,122.43) -- (150.44,122.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (146.90,118.90) -- (146.90,125.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.91,125.69) -- (158.98,125.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (155.45,122.16) -- (155.45,129.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (160.46,128.71) -- (167.52,128.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.99,125.18) -- (163.99,132.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.00,131.76) -- (176.06,131.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (172.53,128.22) -- (172.53,135.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.54,135.09) -- (184.61,135.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.07,131.55) -- (181.07,138.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.08,138.02) -- (193.15,138.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.62,134.49) -- (189.62,141.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.63,141.02) -- (201.69,141.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (198.16,137.49) -- (198.16,144.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.17,143.80) -- (210.23,143.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.70,140.26) -- (206.70,147.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.71,146.57) -- (218.77,146.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (215.24,143.04) -- (215.24,150.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.25,149.39) -- (227.32,149.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.78,145.86) -- (223.78,152.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.79,152.08) -- (235.86,152.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (232.33,148.55) -- (232.33,155.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (237.34,154.61) -- (244.40,154.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.87,151.07) -- (240.87,158.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.88,157.36) -- (252.94,157.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (249.41,153.83) -- (249.41,160.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.42,159.82) -- (261.49,159.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.95,156.28) -- (257.95,163.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.96,162.43) -- (270.03,162.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (266.50,158.90) -- (266.50,165.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.51,164.72) -- (278.57,164.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.04,161.19) -- (275.04,168.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.05,167.20) -- (287.11,167.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.58,163.67) -- (283.58,170.73);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 24.82, 61.58) rectangle ( 29.81, 66.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 24.82, 61.58) -- ( 29.81, 66.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 24.82, 66.58) -- ( 29.81, 61.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.36, 67.75) rectangle ( 38.35, 72.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.36, 67.75) -- ( 38.35, 72.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.36, 72.74) -- ( 38.35, 67.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.90, 73.39) rectangle ( 46.90, 78.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.90, 73.39) -- ( 46.90, 78.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.90, 78.39) -- ( 46.90, 73.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.44, 79.80) rectangle ( 55.44, 84.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.44, 79.80) -- ( 55.44, 84.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.44, 84.80) -- ( 55.44, 79.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.98, 85.74) rectangle ( 63.98, 90.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.98, 85.74) -- ( 63.98, 90.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.98, 90.74) -- ( 63.98, 85.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 67.53, 91.27) rectangle ( 72.52, 96.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 67.53, 91.27) -- ( 72.52, 96.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 67.53, 96.26) -- ( 72.52, 91.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.07, 96.65) rectangle ( 81.06,101.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.07, 96.65) -- ( 81.06,101.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 76.07,101.64) -- ( 81.06, 96.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.61,102.12) rectangle ( 89.61,107.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.61,102.12) -- ( 89.61,107.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 84.61,107.12) -- ( 89.61,102.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.15,107.48) rectangle ( 98.15,112.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.15,107.48) -- ( 98.15,112.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 93.15,112.48) -- ( 98.15,107.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.70,112.41) rectangle (106.69,117.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.70,112.41) -- (106.69,117.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.70,117.40) -- (106.69,112.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.24,116.94) rectangle (115.23,121.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.24,116.94) -- (115.23,121.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (110.24,121.93) -- (115.23,116.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (118.78,121.34) rectangle (123.78,126.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (118.78,121.34) -- (123.78,126.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (118.78,126.34) -- (123.78,121.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.32,125.66) rectangle (132.32,130.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.32,125.66) -- (132.32,130.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.32,130.66) -- (132.32,125.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (135.86,130.03) rectangle (140.86,135.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (135.86,130.03) -- (140.86,135.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (135.86,135.02) -- (140.86,130.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.41,134.21) rectangle (149.40,139.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.41,134.21) -- (149.40,139.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.41,139.21) -- (149.40,134.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.95,138.41) rectangle (157.94,143.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.95,138.41) -- (157.94,143.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.95,143.40) -- (157.94,138.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.49,142.23) rectangle (166.49,147.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.49,142.23) -- (166.49,147.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.49,147.22) -- (166.49,142.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.03,146.06) rectangle (175.03,151.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.03,146.06) -- (175.03,151.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.03,151.05) -- (175.03,146.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (178.58,150.34) rectangle (183.57,155.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (178.58,150.34) -- (183.57,155.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (178.58,155.34) -- (183.57,150.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.12,154.10) rectangle (192.11,159.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.12,154.10) -- (192.11,159.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.12,159.10) -- (192.11,154.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.66,157.97) rectangle (200.66,162.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.66,157.97) -- (200.66,162.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.66,162.96) -- (200.66,157.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.20,161.49) rectangle (209.20,166.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.20,161.49) -- (209.20,166.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.20,166.49) -- (209.20,161.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.74,165.09) rectangle (217.74,170.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.74,165.09) -- (217.74,170.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.74,170.08) -- (217.74,165.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.29,168.70) rectangle (226.28,173.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.29,168.70) -- (226.28,173.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.29,173.69) -- (226.28,168.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.83,172.11) rectangle (234.82,177.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.83,172.11) -- (234.82,177.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.83,177.11) -- (234.82,172.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.37,175.36) rectangle (243.37,180.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.37,175.36) -- (243.37,180.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.37,180.35) -- (243.37,175.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.91,178.96) rectangle (251.91,183.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.91,178.96) -- (251.91,183.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.91,183.95) -- (251.91,178.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.46,182.15) rectangle (260.45,187.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.46,182.15) -- (260.45,187.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (255.46,187.14) -- (260.45,182.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.00,185.46) rectangle (268.99,190.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.00,185.46) -- (268.99,190.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.00,190.45) -- (268.99,185.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (272.54,188.38) rectangle (277.54,193.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (272.54,188.38) -- (277.54,193.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (272.54,193.38) -- (277.54,188.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (281.08,191.66) rectangle (286.08,196.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (281.08,191.66) -- (286.08,196.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (281.08,196.65) -- (286.08,191.66);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] ( 27.31, 63.20) circle (  2.50);

\path[fill=fillColor] ( 35.86, 63.20) circle (  2.50);

\path[fill=fillColor] ( 44.40, 63.20) circle (  2.50);

\path[fill=fillColor] ( 52.94, 63.20) circle (  2.50);

\path[fill=fillColor] ( 61.48, 63.20) circle (  2.50);

\path[fill=fillColor] ( 70.02, 63.20) circle (  2.50);

\path[fill=fillColor] ( 78.57, 63.20) circle (  2.50);

\path[fill=fillColor] ( 87.11, 63.20) circle (  2.50);

\path[fill=fillColor] ( 95.65, 63.20) circle (  2.50);

\path[fill=fillColor] (104.19, 63.20) circle (  2.50);

\path[fill=fillColor] (112.74, 63.20) circle (  2.50);

\path[fill=fillColor] (121.28, 63.20) circle (  2.50);

\path[fill=fillColor] (129.82, 63.20) circle (  2.50);

\path[fill=fillColor] (138.36, 63.20) circle (  2.50);

\path[fill=fillColor] (146.90, 63.20) circle (  2.50);

\path[fill=fillColor] (155.45, 63.20) circle (  2.50);

\path[fill=fillColor] (163.99, 63.20) circle (  2.50);

\path[fill=fillColor] (172.53, 63.20) circle (  2.50);

\path[fill=fillColor] (181.07, 63.20) circle (  2.50);

\path[fill=fillColor] (189.62, 63.20) circle (  2.50);

\path[fill=fillColor] (198.16, 63.20) circle (  2.50);

\path[fill=fillColor] (206.70, 63.20) circle (  2.50);

\path[fill=fillColor] (215.24, 63.20) circle (  2.50);

\path[fill=fillColor] (223.78, 63.20) circle (  2.50);

\path[fill=fillColor] (232.33, 63.20) circle (  2.50);

\path[fill=fillColor] (240.87, 63.20) circle (  2.50);

\path[fill=fillColor] (249.41, 63.20) circle (  2.50);

\path[fill=fillColor] (257.95, 63.20) circle (  2.50);

\path[fill=fillColor] (266.50, 63.20) circle (  2.50);

\path[fill=fillColor] (275.04, 63.20) circle (  2.50);

\path[fill=fillColor] (283.58, 63.20) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 30.69) rectangle (283.58,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 60.52) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,170.08) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 63.55) --
	( 27.31, 63.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,173.11) --
	( 27.31,173.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 27.94) --
	( 27.31, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (112.74, 27.94) --
	(112.74, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.16, 27.94) --
	(198.16, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (112.74, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.16, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (155.45,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,112.42) {Mean interval (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,202.22) {C};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,216.81);
\end{scope}
\begin{scope}
\path[clip] (323.24, 30.69) rectangle (572.66,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (323.24, 30.69) rectangle (572.66,194.15);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (323.24, 58.94) --
	(326.39, 58.94) --
	(329.55, 58.94) --
	(332.71, 58.94) --
	(335.87, 58.94) --
	(339.02, 58.94) --
	(342.18, 58.94) --
	(345.34, 58.94) --
	(348.49, 58.94) --
	(351.65, 58.94) --
	(354.81, 58.94) --
	(357.97, 58.94) --
	(361.12, 58.94) --
	(364.28, 58.94) --
	(367.44, 58.94) --
	(370.60, 58.94) --
	(373.75, 58.94) --
	(376.91, 58.94) --
	(380.07, 58.94) --
	(383.22, 58.94) --
	(386.38, 58.94) --
	(389.54, 58.94) --
	(392.70, 58.94) --
	(395.85, 58.94) --
	(399.01, 58.94) --
	(402.17, 58.94) --
	(405.32, 58.94) --
	(408.48, 58.94) --
	(411.64, 58.94) --
	(414.80, 58.94) --
	(417.95, 58.94) --
	(421.11, 58.94) --
	(424.27, 58.94) --
	(427.43, 58.94) --
	(430.58, 58.94) --
	(433.74, 58.94) --
	(436.90, 58.94) --
	(440.05, 58.94) --
	(443.21, 58.94) --
	(446.37, 58.94) --
	(449.53, 58.94) --
	(452.68, 58.94) --
	(455.84, 58.94) --
	(459.00, 58.94) --
	(462.16, 58.94) --
	(465.31, 58.94) --
	(468.47, 58.94) --
	(471.63, 58.94) --
	(474.78, 58.94) --
	(477.94, 58.94) --
	(481.10, 58.94) --
	(484.26, 58.94) --
	(487.41, 58.94) --
	(490.57, 58.94) --
	(493.73, 58.94) --
	(496.89, 58.94) --
	(500.04, 58.94) --
	(503.20, 58.94) --
	(506.36, 58.94) --
	(509.51, 58.94) --
	(512.67, 58.94) --
	(515.83, 58.94) --
	(518.99, 58.94) --
	(522.14, 58.94) --
	(525.30, 58.94) --
	(528.46, 58.94) --
	(531.62, 58.94) --
	(534.77, 58.94) --
	(537.93, 58.94) --
	(541.09, 58.94) --
	(544.24, 58.94) --
	(547.40, 58.94) --
	(550.56, 58.94) --
	(553.72, 58.94) --
	(556.87, 58.94) --
	(560.03, 58.94) --
	(563.19, 58.94) --
	(566.35, 58.94) --
	(569.50, 58.94) --
	(572.66, 58.94);
\definecolor{drawColor}{RGB}{230,159,0}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (323.24, 97.64) --
	(326.39, 97.20) --
	(329.55, 96.77) --
	(332.71, 96.35) --
	(335.87, 95.93) --
	(339.02, 95.52) --
	(342.18, 95.12) --
	(345.34, 94.73) --
	(348.49, 94.34) --
	(351.65, 93.96) --
	(354.81, 93.58) --
	(357.97, 93.22) --
	(361.12, 92.86) --
	(364.28, 92.51) --
	(367.44, 92.16) --
	(370.60, 91.82) --
	(373.75, 91.49) --
	(376.91, 91.17) --
	(380.07, 90.85) --
	(383.22, 90.54) --
	(386.38, 90.23) --
	(389.54, 89.94) --
	(392.70, 89.65) --
	(395.85, 89.37) --
	(399.01, 89.10) --
	(402.17, 88.84) --
	(405.32, 88.58) --
	(408.48, 88.32) --
	(411.64, 88.07) --
	(414.80, 87.82) --
	(417.95, 87.59) --
	(421.11, 87.36) --
	(424.27, 87.14) --
	(427.43, 86.92) --
	(430.58, 86.72) --
	(433.74, 86.51) --
	(436.90, 86.31) --
	(440.05, 86.12) --
	(443.21, 85.92) --
	(446.37, 85.72) --
	(449.53, 85.52) --
	(452.68, 85.33) --
	(455.84, 85.14) --
	(459.00, 84.95) --
	(462.16, 84.77) --
	(465.31, 84.59) --
	(468.47, 84.41) --
	(471.63, 84.24) --
	(474.78, 84.07) --
	(477.94, 83.89) --
	(481.10, 83.72) --
	(484.26, 83.55) --
	(487.41, 83.38) --
	(490.57, 83.21) --
	(493.73, 83.04) --
	(496.89, 82.88) --
	(500.04, 82.72) --
	(503.20, 82.56) --
	(506.36, 82.40) --
	(509.51, 82.25) --
	(512.67, 82.09) --
	(515.83, 81.94) --
	(518.99, 81.79) --
	(522.14, 81.64) --
	(525.30, 81.49) --
	(528.46, 81.35) --
	(531.62, 81.20) --
	(534.77, 81.06) --
	(537.93, 80.92) --
	(541.09, 80.79) --
	(544.24, 80.65) --
	(547.40, 80.52) --
	(550.56, 80.39) --
	(553.72, 80.26) --
	(556.87, 80.13) --
	(560.03, 80.00) --
	(563.19, 79.88) --
	(566.35, 79.76) --
	(569.50, 79.64) --
	(572.66, 79.52);
\definecolor{drawColor}{RGB}{86,180,233}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (323.24,115.70) --
	(326.39,114.93) --
	(329.55,114.17) --
	(332.71,113.43) --
	(335.87,112.70) --
	(339.02,111.98) --
	(342.18,111.27) --
	(345.34,110.58) --
	(348.49,109.90) --
	(351.65,109.23) --
	(354.81,108.58) --
	(357.97,107.94) --
	(361.12,107.32) --
	(364.28,106.70) --
	(367.44,106.10) --
	(370.60,105.51) --
	(373.75,104.94) --
	(376.91,104.37) --
	(380.07,103.82) --
	(383.22,103.28) --
	(386.38,102.76) --
	(389.54,102.25) --
	(392.70,101.76) --
	(395.85,101.28) --
	(399.01,100.81) --
	(402.17,100.36) --
	(405.32, 99.92) --
	(408.48, 99.48) --
	(411.64, 99.06) --
	(414.80, 98.64) --
	(417.95, 98.23) --
	(421.11, 97.85) --
	(424.27, 97.47) --
	(427.43, 97.11) --
	(430.58, 96.76) --
	(433.74, 96.42) --
	(436.90, 96.09) --
	(440.05, 95.75) --
	(443.21, 95.42) --
	(446.37, 95.09) --
	(449.53, 94.76) --
	(452.68, 94.43) --
	(455.84, 94.12) --
	(459.00, 93.81) --
	(462.16, 93.51) --
	(465.31, 93.21) --
	(468.47, 92.92) --
	(471.63, 92.63) --
	(474.78, 92.34) --
	(477.94, 92.06) --
	(481.10, 91.77) --
	(484.26, 91.49) --
	(487.41, 91.21) --
	(490.57, 90.94) --
	(493.73, 90.67) --
	(496.89, 90.40) --
	(500.04, 90.13) --
	(503.20, 89.87) --
	(506.36, 89.61) --
	(509.51, 89.36) --
	(512.67, 89.11) --
	(515.83, 88.86) --
	(518.99, 88.61) --
	(522.14, 88.37) --
	(525.30, 88.13) --
	(528.46, 87.90) --
	(531.62, 87.67) --
	(534.77, 87.44) --
	(537.93, 87.22) --
	(541.09, 87.00) --
	(544.24, 86.78) --
	(547.40, 86.56) --
	(550.56, 86.35) --
	(553.72, 86.14) --
	(556.87, 85.94) --
	(560.03, 85.74) --
	(563.19, 85.54) --
	(566.35, 85.35) --
	(569.50, 85.16) --
	(572.66, 84.97);
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (323.24,133.03) --
	(326.39,131.91) --
	(329.55,130.81) --
	(332.71,129.73) --
	(335.87,128.67) --
	(339.02,127.63) --
	(342.18,126.61) --
	(345.34,125.61) --
	(348.49,124.63) --
	(351.65,123.67) --
	(354.81,122.72) --
	(357.97,121.80) --
	(361.12,120.90) --
	(364.28,120.02) --
	(367.44,119.15) --
	(370.60,118.31) --
	(373.75,117.48) --
	(376.91,116.67) --
	(380.07,115.88) --
	(383.22,115.11) --
	(386.38,114.36) --
	(389.54,113.63) --
	(392.70,112.93) --
	(395.85,112.25) --
	(399.01,111.59) --
	(402.17,110.95) --
	(405.32,110.32) --
	(408.48,109.71) --
	(411.64,109.11) --
	(414.80,108.52) --
	(417.95,107.94) --
	(421.11,107.40) --
	(424.27,106.87) --
	(427.43,106.37) --
	(430.58,105.88) --
	(433.74,105.40) --
	(436.90,104.94) --
	(440.05,104.47) --
	(443.21,104.01) --
	(446.37,103.55) --
	(449.53,103.09) --
	(452.68,102.64) --
	(455.84,102.20) --
	(459.00,101.77) --
	(462.16,101.35) --
	(465.31,100.94) --
	(468.47,100.54) --
	(471.63,100.14) --
	(474.78, 99.74) --
	(477.94, 99.35) --
	(481.10, 98.96) --
	(484.26, 98.57) --
	(487.41, 98.18) --
	(490.57, 97.81) --
	(493.73, 97.43) --
	(496.89, 97.07) --
	(500.04, 96.71) --
	(503.20, 96.35) --
	(506.36, 96.00) --
	(509.51, 95.65) --
	(512.67, 95.31) --
	(515.83, 94.97) --
	(518.99, 94.63) --
	(522.14, 94.31) --
	(525.30, 93.98) --
	(528.46, 93.66) --
	(531.62, 93.35) --
	(534.77, 93.04) --
	(537.93, 92.74) --
	(541.09, 92.44) --
	(544.24, 92.14) --
	(547.40, 91.86) --
	(550.56, 91.57) --
	(553.72, 91.29) --
	(556.87, 91.02) --
	(560.03, 90.75) --
	(563.19, 90.48) --
	(566.35, 90.23) --
	(569.50, 89.97) --
	(572.66, 89.72);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] (323.24,149.67) --
	(326.39,148.19) --
	(329.55,146.73) --
	(332.71,145.31) --
	(335.87,143.91) --
	(339.02,142.53) --
	(342.18,141.18) --
	(345.34,139.86) --
	(348.49,138.57) --
	(351.65,137.30) --
	(354.81,136.06) --
	(357.97,134.85) --
	(361.12,133.66) --
	(364.28,132.50) --
	(367.44,131.37) --
	(370.60,130.26) --
	(373.75,129.17) --
	(376.91,128.11) --
	(380.07,127.07) --
	(383.22,126.06) --
	(386.38,125.09) --
	(389.54,124.14) --
	(392.70,123.23) --
	(395.85,122.35) --
	(399.01,121.49) --
	(402.17,120.66) --
	(405.32,119.84) --
	(408.48,119.05) --
	(411.64,118.27) --
	(414.80,117.51) --
	(417.95,116.77) --
	(421.11,116.06) --
	(424.27,115.39) --
	(427.43,114.74) --
	(430.58,114.12) --
	(433.74,113.51) --
	(436.90,112.91) --
	(440.05,112.32) --
	(443.21,111.73) --
	(446.37,111.14) --
	(449.53,110.55) --
	(452.68,109.97) --
	(455.84,109.41) --
	(459.00,108.87) --
	(462.16,108.33) --
	(465.31,107.81) --
	(468.47,107.30) --
	(471.63,106.80) --
	(474.78,106.29) --
	(477.94,105.80) --
	(481.10,105.30) --
	(484.26,104.81) --
	(487.41,104.32) --
	(490.57,103.84) --
	(493.73,103.37) --
	(496.89,102.91) --
	(500.04,102.46) --
	(503.20,102.01) --
	(506.36,101.57) --
	(509.51,101.13) --
	(512.67,100.70) --
	(515.83,100.27) --
	(518.99, 99.86) --
	(522.14, 99.44) --
	(525.30, 99.04) --
	(528.46, 98.64) --
	(531.62, 98.25) --
	(534.77, 97.86) --
	(537.93, 97.48) --
	(541.09, 97.11) --
	(544.24, 96.75) --
	(547.40, 96.39) --
	(550.56, 96.04) --
	(553.72, 95.69) --
	(556.87, 95.35) --
	(560.03, 95.02) --
	(563.19, 94.69) --
	(566.35, 94.37) --
	(569.50, 94.06) --
	(572.66, 93.75);
\definecolor{fillColor}{RGB}{230,159,0}

\path[fill=fillColor] (323.24,101.62) --
	(326.60, 95.79) --
	(319.87, 95.79) --
	cycle;

\path[fill=fillColor] (331.55,100.41) --
	(334.91, 94.58) --
	(328.19, 94.58) --
	cycle;

\path[fill=fillColor] (339.86, 99.32) --
	(343.23, 93.49) --
	(336.50, 93.49) --
	cycle;

\path[fill=fillColor] (348.18, 98.18) --
	(351.54, 92.35) --
	(344.81, 92.35) --
	cycle;

\path[fill=fillColor] (356.49, 97.16) --
	(359.86, 91.33) --
	(353.13, 91.33) --
	cycle;

\path[fill=fillColor] (364.81, 96.27) --
	(368.17, 90.44) --
	(361.44, 90.44) --
	cycle;

\path[fill=fillColor] (373.12, 95.43) --
	(376.48, 89.61) --
	(369.76, 89.61) --
	cycle;

\path[fill=fillColor] (381.44, 94.63) --
	(384.80, 88.80) --
	(378.07, 88.80) --
	cycle;

\path[fill=fillColor] (389.75, 93.76) --
	(393.11, 87.94) --
	(386.39, 87.94) --
	cycle;

\path[fill=fillColor] (398.06, 92.98) --
	(401.43, 87.16) --
	(394.70, 87.16) --
	cycle;

\path[fill=fillColor] (406.38, 92.29) --
	(409.74, 86.46) --
	(403.01, 86.46) --
	cycle;

\path[fill=fillColor] (414.69, 91.68) --
	(418.06, 85.86) --
	(411.33, 85.86) --
	cycle;

\path[fill=fillColor] (423.01, 91.09) --
	(426.37, 85.26) --
	(419.64, 85.26) --
	cycle;

\path[fill=fillColor] (431.32, 90.57) --
	(434.68, 84.74) --
	(427.96, 84.74) --
	cycle;

\path[fill=fillColor] (439.63, 90.05) --
	(443.00, 84.22) --
	(436.27, 84.22) --
	cycle;

\path[fill=fillColor] (447.95, 89.54) --
	(451.31, 83.72) --
	(444.58, 83.72) --
	cycle;

\path[fill=fillColor] (456.26, 89.09) --
	(459.63, 83.26) --
	(452.90, 83.26) --
	cycle;

\path[fill=fillColor] (464.58, 88.59) --
	(467.94, 82.77) --
	(461.21, 82.77) --
	cycle;

\path[fill=fillColor] (472.89, 88.04) --
	(476.25, 82.21) --
	(469.53, 82.21) --
	cycle;

\path[fill=fillColor] (481.20, 87.56) --
	(484.57, 81.73) --
	(477.84, 81.73) --
	cycle;

\path[fill=fillColor] (489.52, 87.14) --
	(492.88, 81.31) --
	(486.15, 81.31) --
	cycle;

\path[fill=fillColor] (497.83, 86.70) --
	(501.20, 80.87) --
	(494.47, 80.87) --
	cycle;

\path[fill=fillColor] (506.15, 86.30) --
	(509.51, 80.48) --
	(502.78, 80.48) --
	cycle;

\path[fill=fillColor] (514.46, 85.86) --
	(517.82, 80.04) --
	(511.10, 80.04) --
	cycle;

\path[fill=fillColor] (522.78, 85.49) --
	(526.14, 79.67) --
	(519.41, 79.67) --
	cycle;

\path[fill=fillColor] (531.09, 85.15) --
	(534.45, 79.33) --
	(527.73, 79.33) --
	cycle;

\path[fill=fillColor] (539.40, 84.78) --
	(542.77, 78.96) --
	(536.04, 78.96) --
	cycle;

\path[fill=fillColor] (547.72, 84.39) --
	(551.08, 78.57) --
	(544.35, 78.57) --
	cycle;

\path[fill=fillColor] (556.03, 84.03) --
	(559.40, 78.21) --
	(552.67, 78.21) --
	cycle;

\path[fill=fillColor] (564.35, 83.73) --
	(567.71, 77.90) --
	(560.98, 77.90) --
	cycle;

\path[fill=fillColor] (572.66, 83.38) --
	(576.02, 77.56) --
	(569.30, 77.56) --
	cycle;
\definecolor{fillColor}{RGB}{86,180,233}

\path[fill=fillColor] (320.74,113.39) --
	(325.73,113.39) --
	(325.73,118.38) --
	(320.74,118.38) --
	cycle;

\path[fill=fillColor] (329.05,111.23) --
	(334.05,111.23) --
	(334.05,116.23) --
	(329.05,116.23) --
	cycle;

\path[fill=fillColor] (337.37,109.33) --
	(342.36,109.33) --
	(342.36,114.33) --
	(337.37,114.33) --
	cycle;

\path[fill=fillColor] (345.68,107.33) --
	(350.68,107.33) --
	(350.68,112.32) --
	(345.68,112.32) --
	cycle;

\path[fill=fillColor] (353.99,105.51) --
	(358.99,105.51) --
	(358.99,110.51) --
	(353.99,110.51) --
	cycle;

\path[fill=fillColor] (362.31,103.96) --
	(367.30,103.96) --
	(367.30,108.96) --
	(362.31,108.96) --
	cycle;

\path[fill=fillColor] (370.62,102.51) --
	(375.62,102.51) --
	(375.62,107.50) --
	(370.62,107.50) --
	cycle;

\path[fill=fillColor] (378.94,101.12) --
	(383.93,101.12) --
	(383.93,106.11) --
	(378.94,106.11) --
	cycle;

\path[fill=fillColor] (387.25, 99.64) --
	(392.25, 99.64) --
	(392.25,104.64) --
	(387.25,104.64) --
	cycle;

\path[fill=fillColor] (395.57, 98.32) --
	(400.56, 98.32) --
	(400.56,103.31) --
	(395.57,103.31) --
	cycle;

\path[fill=fillColor] (403.88, 97.13) --
	(408.88, 97.13) --
	(408.88,102.13) --
	(403.88,102.13) --
	cycle;

\path[fill=fillColor] (412.19, 96.11) --
	(417.19, 96.11) --
	(417.19,101.11) --
	(412.19,101.11) --
	cycle;

\path[fill=fillColor] (420.51, 95.11) --
	(425.50, 95.11) --
	(425.50,100.10) --
	(420.51,100.10) --
	cycle;

\path[fill=fillColor] (428.82, 94.21) --
	(433.82, 94.21) --
	(433.82, 99.20) --
	(428.82, 99.20) --
	cycle;

\path[fill=fillColor] (437.14, 93.32) --
	(442.13, 93.32) --
	(442.13, 98.32) --
	(437.14, 98.32) --
	cycle;

\path[fill=fillColor] (445.45, 92.49) --
	(450.45, 92.49) --
	(450.45, 97.48) --
	(445.45, 97.48) --
	cycle;

\path[fill=fillColor] (453.76, 91.72) --
	(458.76, 91.72) --
	(458.76, 96.71) --
	(453.76, 96.71) --
	cycle;

\path[fill=fillColor] (462.08, 90.90) --
	(467.07, 90.90) --
	(467.07, 95.89) --
	(462.08, 95.89) --
	cycle;

\path[fill=fillColor] (470.39, 90.00) --
	(475.39, 90.00) --
	(475.39, 94.99) --
	(470.39, 94.99) --
	cycle;

\path[fill=fillColor] (478.71, 89.22) --
	(483.70, 89.22) --
	(483.70, 94.21) --
	(478.71, 94.21) --
	cycle;

\path[fill=fillColor] (487.02, 88.51) --
	(492.02, 88.51) --
	(492.02, 93.51) --
	(487.02, 93.51) --
	cycle;

\path[fill=fillColor] (495.34, 87.80) --
	(500.33, 87.80) --
	(500.33, 92.79) --
	(495.34, 92.79) --
	cycle;

\path[fill=fillColor] (503.65, 87.15) --
	(508.64, 87.15) --
	(508.64, 92.14) --
	(503.65, 92.14) --
	cycle;

\path[fill=fillColor] (511.96, 86.43) --
	(516.96, 86.43) --
	(516.96, 91.43) --
	(511.96, 91.43) --
	cycle;

\path[fill=fillColor] (520.28, 85.84) --
	(525.27, 85.84) --
	(525.27, 90.83) --
	(520.28, 90.83) --
	cycle;

\path[fill=fillColor] (528.59, 85.28) --
	(533.59, 85.28) --
	(533.59, 90.27) --
	(528.59, 90.27) --
	cycle;

\path[fill=fillColor] (536.91, 84.68) --
	(541.90, 84.68) --
	(541.90, 89.67) --
	(536.91, 89.67) --
	cycle;

\path[fill=fillColor] (545.22, 84.07) --
	(550.22, 84.07) --
	(550.22, 89.07) --
	(545.22, 89.07) --
	cycle;

\path[fill=fillColor] (553.53, 83.47) --
	(558.53, 83.47) --
	(558.53, 88.47) --
	(553.53, 88.47) --
	cycle;

\path[fill=fillColor] (561.85, 82.98) --
	(566.84, 82.98) --
	(566.84, 87.97) --
	(561.85, 87.97) --
	cycle;

\path[fill=fillColor] (570.16, 82.43) --
	(575.16, 82.43) --
	(575.16, 87.42) --
	(570.16, 87.42) --
	cycle;
\definecolor{drawColor}{RGB}{0,158,115}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (319.70,133.33) -- (326.77,133.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.24,129.79) -- (323.24,136.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.02,130.17) -- (335.08,130.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (331.55,126.64) -- (331.55,133.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (336.33,127.42) -- (343.40,127.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (339.86,123.89) -- (339.86,130.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (344.65,124.53) -- (351.71,124.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (348.18,121.00) -- (348.18,128.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (352.96,121.87) -- (360.02,121.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (356.49,118.34) -- (356.49,125.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.27,119.65) -- (368.34,119.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.81,116.11) -- (364.81,123.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.59,117.55) -- (376.65,117.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (373.12,114.02) -- (373.12,121.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.90,115.54) -- (384.97,115.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (381.44,112.01) -- (381.44,119.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (386.22,113.47) -- (393.28,113.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.75,109.94) -- (389.75,117.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (394.53,111.60) -- (401.60,111.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (398.06,108.06) -- (398.06,115.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (402.85,109.91) -- (409.91,109.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (406.38,106.37) -- (406.38,113.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (411.16,108.50) -- (418.22,108.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (414.69,104.97) -- (414.69,112.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.47,107.09) -- (426.54,107.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.01,103.56) -- (423.01,110.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.79,105.80) -- (434.85,105.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.32,102.27) -- (431.32,109.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.10,104.55) -- (443.17,104.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (439.63,101.02) -- (439.63,108.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.42,103.41) -- (451.48,103.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (447.95, 99.87) -- (447.95,106.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.73,102.33) -- (459.79,102.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.26, 98.80) -- (456.26,105.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.04,101.17) -- (468.11,101.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.58, 97.64) -- (464.58,104.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.36, 99.96) -- (476.42, 99.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (472.89, 96.42) -- (472.89,103.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.67, 98.89) -- (484.74, 98.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (481.20, 95.35) -- (481.20,102.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.99, 97.92) -- (493.05, 97.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (489.52, 94.38) -- (489.52,101.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.30, 96.93) -- (501.36, 96.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (497.83, 93.40) -- (497.83,100.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.61, 96.04) -- (509.68, 96.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (506.15, 92.51) -- (506.15, 99.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.93, 95.07) -- (517.99, 95.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (514.46, 91.53) -- (514.46, 98.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.24, 94.27) -- (526.31, 94.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.78, 90.73) -- (522.78, 97.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (527.56, 93.49) -- (534.62, 93.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (531.09, 89.96) -- (531.09, 97.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.87, 92.69) -- (542.94, 92.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (539.40, 89.16) -- (539.40, 96.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.19, 91.90) -- (551.25, 91.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.72, 88.37) -- (547.72, 95.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.50, 91.05) -- (559.56, 91.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (556.03, 87.52) -- (556.03, 94.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.81, 90.38) -- (567.88, 90.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.35, 86.85) -- (564.35, 93.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13, 89.65) -- (576.19, 89.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66, 86.12) -- (572.66, 93.18);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (320.74,147.59) rectangle (325.73,152.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (320.74,147.59) -- (325.73,152.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (320.74,152.58) -- (325.73,147.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.05,143.39) rectangle (334.05,148.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.05,143.39) -- (334.05,148.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.05,148.38) -- (334.05,143.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (337.37,139.76) rectangle (342.36,144.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (337.37,139.76) -- (342.36,144.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (337.37,144.75) -- (342.36,139.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.68,135.95) rectangle (350.68,140.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.68,135.95) -- (350.68,140.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.68,140.94) -- (350.68,135.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (353.99,132.41) rectangle (358.99,137.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (353.99,132.41) -- (358.99,137.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (353.99,137.40) -- (358.99,132.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (362.31,129.50) rectangle (367.30,134.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (362.31,129.50) -- (367.30,134.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (362.31,134.50) -- (367.30,129.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (370.62,126.75) rectangle (375.62,131.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (370.62,126.75) -- (375.62,131.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (370.62,131.74) -- (375.62,126.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (378.94,124.08) rectangle (383.93,129.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (378.94,124.08) -- (383.93,129.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (378.94,129.08) -- (383.93,124.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.25,121.44) rectangle (392.25,126.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.25,121.44) -- (392.25,126.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.25,126.43) -- (392.25,121.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.57,118.99) rectangle (400.56,123.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.57,118.99) -- (400.56,123.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.57,123.99) -- (400.56,118.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (403.88,116.78) rectangle (408.88,121.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (403.88,116.78) -- (408.88,121.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (403.88,121.77) -- (408.88,116.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.19,115.02) rectangle (417.19,120.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.19,115.02) -- (417.19,120.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.19,120.02) -- (417.19,115.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.51,113.20) rectangle (425.50,118.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.51,113.20) -- (425.50,118.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.51,118.19) -- (425.50,113.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.82,111.53) rectangle (433.82,116.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.82,111.53) -- (433.82,116.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (428.82,116.52) -- (433.82,111.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.14,109.90) rectangle (442.13,114.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.14,109.90) -- (442.13,114.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.14,114.89) -- (442.13,109.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (445.45,108.47) rectangle (450.45,113.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (445.45,108.47) -- (450.45,113.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (445.45,113.47) -- (450.45,108.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.76,107.09) rectangle (458.76,112.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.76,107.09) -- (458.76,112.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (453.76,112.08) -- (458.76,107.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.08,105.59) rectangle (467.07,110.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.08,105.59) -- (467.07,110.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (462.08,110.58) -- (467.07,105.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.39,104.07) rectangle (475.39,109.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.39,104.07) -- (475.39,109.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (470.39,109.07) -- (475.39,104.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.71,102.72) rectangle (483.70,107.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.71,102.72) -- (483.70,107.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (478.71,107.72) -- (483.70,102.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.02,101.49) rectangle (492.02,106.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.02,101.49) -- (492.02,106.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (487.02,106.49) -- (492.02,101.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.34,100.25) rectangle (500.33,105.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.34,100.25) -- (500.33,105.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.34,105.25) -- (500.33,100.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (503.65, 99.11) rectangle (508.64,104.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (503.65, 99.11) -- (508.64,104.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (503.65,104.10) -- (508.64, 99.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.96, 97.90) rectangle (516.96,102.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.96, 97.90) -- (516.96,102.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.96,102.90) -- (516.96, 97.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (520.28, 96.90) rectangle (525.27,101.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (520.28, 96.90) -- (525.27,101.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (520.28,101.90) -- (525.27, 96.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.59, 95.91) rectangle (533.59,100.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.59, 95.91) -- (533.59,100.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.59,100.91) -- (533.59, 95.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.91, 94.94) rectangle (541.90, 99.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.91, 94.94) -- (541.90, 99.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.91, 99.93) -- (541.90, 94.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (545.22, 94.00) rectangle (550.22, 98.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (545.22, 94.00) -- (550.22, 98.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (545.22, 98.99) -- (550.22, 94.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.53, 92.89) rectangle (558.53, 97.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.53, 92.89) -- (558.53, 97.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.53, 97.89) -- (558.53, 92.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.85, 92.04) rectangle (566.84, 97.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.85, 92.04) -- (566.84, 97.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.85, 97.04) -- (566.84, 92.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16, 91.16) rectangle (575.16, 96.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16, 91.16) -- (575.16, 96.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16, 96.15) -- (575.16, 91.16);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (323.24, 58.94) circle (  2.50);

\path[fill=fillColor] (331.55, 58.94) circle (  2.50);

\path[fill=fillColor] (339.86, 58.94) circle (  2.50);

\path[fill=fillColor] (348.18, 58.94) circle (  2.50);

\path[fill=fillColor] (356.49, 58.94) circle (  2.50);

\path[fill=fillColor] (364.81, 58.94) circle (  2.50);

\path[fill=fillColor] (373.12, 58.94) circle (  2.50);

\path[fill=fillColor] (381.44, 58.94) circle (  2.50);

\path[fill=fillColor] (389.75, 58.94) circle (  2.50);

\path[fill=fillColor] (398.06, 58.94) circle (  2.50);

\path[fill=fillColor] (406.38, 58.94) circle (  2.50);

\path[fill=fillColor] (414.69, 58.94) circle (  2.50);

\path[fill=fillColor] (423.01, 58.94) circle (  2.50);

\path[fill=fillColor] (431.32, 58.94) circle (  2.50);

\path[fill=fillColor] (439.63, 58.94) circle (  2.50);

\path[fill=fillColor] (447.95, 58.94) circle (  2.50);

\path[fill=fillColor] (456.26, 58.94) circle (  2.50);

\path[fill=fillColor] (464.58, 58.94) circle (  2.50);

\path[fill=fillColor] (472.89, 58.94) circle (  2.50);

\path[fill=fillColor] (481.20, 58.94) circle (  2.50);

\path[fill=fillColor] (489.52, 58.94) circle (  2.50);

\path[fill=fillColor] (497.83, 58.94) circle (  2.50);

\path[fill=fillColor] (506.15, 58.94) circle (  2.50);

\path[fill=fillColor] (514.46, 58.94) circle (  2.50);

\path[fill=fillColor] (522.78, 58.94) circle (  2.50);

\path[fill=fillColor] (531.09, 58.94) circle (  2.50);

\path[fill=fillColor] (539.40, 58.94) circle (  2.50);

\path[fill=fillColor] (547.72, 58.94) circle (  2.50);

\path[fill=fillColor] (556.03, 58.94) circle (  2.50);

\path[fill=fillColor] (564.35, 58.94) circle (  2.50);

\path[fill=fillColor] (572.66, 58.94) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (323.24, 30.69) rectangle (572.66,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29, 27.66) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29, 66.58) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,105.50) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,144.42) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (318.29,183.34) {0.8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49, 30.69) --
	(323.24, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49, 69.61) --
	(323.24, 69.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,108.53) --
	(323.24,108.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,147.45) --
	(323.24,147.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.49,186.37) --
	(323.24,186.37);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.24, 27.94) --
	(323.24, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.38, 27.94) --
	(406.38, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (489.52, 27.94) --
	(489.52, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (323.24, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (406.38, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.52, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (447.95,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,112.42) {Squared coefficient of variation};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (323.24,202.22) {D};
\end{scope}
\end{tikzpicture}

\end{document}