### 2020-09-15 serial_analysis.tex

% Created by tikzDevice version 0.12.3 on 2020-09-14 16:14:30
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (1156.32,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (289.08,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (289.08,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 31.71, 30.69) rectangle (283.58,266.42);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.71, 30.69) rectangle (283.58,266.42);
\definecolor{fillColor}{gray}{0.35}

\path[fill=fillColor] ( 31.71, 30.69) rectangle ( 38.40, 34.10);

\path[fill=fillColor] ( 46.57, 30.69) rectangle ( 53.26, 37.52);

\path[fill=fillColor] ( 54.00, 30.69) rectangle ( 60.69, 40.94);

\path[fill=fillColor] ( 61.43, 30.69) rectangle ( 68.12, 37.52);

\path[fill=fillColor] ( 68.86, 30.69) rectangle ( 75.55, 40.94);

\path[fill=fillColor] ( 76.29, 30.69) rectangle ( 82.98, 34.10);

\path[fill=fillColor] ( 83.72, 30.69) rectangle ( 90.41, 37.52);

\path[fill=fillColor] ( 91.15, 30.69) rectangle ( 97.84, 40.94);

\path[fill=fillColor] ( 98.58, 30.69) rectangle (105.27, 37.52);

\path[fill=fillColor] (106.01, 30.69) rectangle (112.70, 44.35);

\path[fill=fillColor] (113.44, 30.69) rectangle (120.13, 85.35);

\path[fill=fillColor] (120.87, 30.69) rectangle (127.56, 85.35);

\path[fill=fillColor] (128.30, 30.69) rectangle (134.99, 92.18);

\path[fill=fillColor] (135.73, 30.69) rectangle (142.42, 95.60);

\path[fill=fillColor] (143.16, 30.69) rectangle (149.84,150.26);

\path[fill=fillColor] (150.59, 30.69) rectangle (157.27,194.68);

\path[fill=fillColor] (158.02, 30.69) rectangle (164.70,204.93);

\path[fill=fillColor] (165.45, 30.69) rectangle (172.13,198.09);

\path[fill=fillColor] (172.88, 30.69) rectangle (179.56,218.59);

\path[fill=fillColor] (180.31, 30.69) rectangle (186.99,245.92);

\path[fill=fillColor] (187.74, 30.69) rectangle (194.42,256.17);

\path[fill=fillColor] (195.17, 30.69) rectangle (201.85,198.09);

\path[fill=fillColor] (202.60, 30.69) rectangle (209.28,187.84);

\path[fill=fillColor] (210.03, 30.69) rectangle (216.71,191.26);

\path[fill=fillColor] (217.46, 30.69) rectangle (224.14,177.59);

\path[fill=fillColor] (224.89, 30.69) rectangle (231.57,136.60);

\path[fill=fillColor] (232.31, 30.69) rectangle (239.00,116.10);

\path[fill=fillColor] (239.74, 30.69) rectangle (246.43,102.43);

\path[fill=fillColor] (247.17, 30.69) rectangle (253.86, 75.10);

\path[fill=fillColor] (254.60, 30.69) rectangle (261.29, 61.43);

\path[fill=fillColor] (262.03, 30.69) rectangle (268.72, 51.18);

\path[fill=fillColor] (269.46, 30.69) rectangle (276.15, 37.52);

\path[fill=fillColor] (276.89, 30.69) rectangle (283.58, 34.10);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.71, 30.69) rectangle (283.58,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76, 27.66) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76, 95.98) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,164.31) {40};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,232.64) {60};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96, 30.69) --
	( 31.71, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96, 99.02) --
	( 31.71, 99.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,167.34) --
	( 31.71,167.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,235.67) --
	( 31.71,235.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.63, 27.94) --
	( 79.63, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (131.64, 27.94) --
	(131.64, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.65, 27.94) --
	(183.65, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (235.66, 27.94) --
	(235.66, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 79.63, 19.68) {Jan 13};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (131.64, 19.68) {Jan 20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (183.65, 19.68) {Jan 27};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (235.66, 19.68) {Feb 03};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (157.65,  7.64) {Symptom onset date};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,148.55) {Number of cases with a known infector/infectee pair};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 31.71,274.49) {A. Epidemic curve};
\end{scope}
\begin{scope}
\path[clip] (299.25,  0.00) rectangle (567.99,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (299.25,  0.00) rectangle (567.99,289.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (333.90, 30.69) rectangle (562.49,266.42);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate=-45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (499.08,195.87) {maximum observable};

\node[text=drawColor,rotate=-45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (384.78, 74.43) {minimum observable};

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (419.62,266.42) --
	(426.76,259.28) --
	(433.91,252.14) --
	(441.05,244.99) --
	(448.19,237.85) --
	(455.34,230.71) --
	(462.48,223.56) --
	(469.62,216.42) --
	(476.77,209.27) --
	(483.91,202.13) --
	(491.05,194.99) --
	(498.20,187.84) --
	(505.34,180.70) --
	(512.48,173.56) --
	(519.63,166.41) --
	(526.77,159.27) --
	(533.91,152.13) --
	(541.06,144.98) --
	(548.20,137.84) --
	(555.35,130.70) --
	(562.49,123.55);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (333.90,116.41) --
	(341.04,109.26) --
	(348.18,102.12) --
	(355.33, 94.98) --
	(362.47, 87.83) --
	(369.61, 80.69) --
	(376.76, 73.55) --
	(383.90, 66.40) --
	(391.04, 59.26) --
	(398.19, 52.12) --
	(405.33, 44.97) --
	(412.47, 37.83) --
	(419.62, 30.69);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (333.90,137.84) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (348.18,219.10) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (355.33,183.08) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (362.47,194.99) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (369.61,194.99) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (376.76,223.56) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (383.90,223.56) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (391.04,194.99) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (398.19,216.42) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (405.33,184.27) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (412.47,159.58) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (419.62,159.72) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (426.76,165.29) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (433.91,150.04) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (441.05,160.95) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (448.19,149.30) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (455.34,144.74) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (462.48,149.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (469.62,133.26) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (476.77,138.29) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (483.91,129.96) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (491.05,127.28) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (498.20,127.33) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (505.34,119.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (512.48,115.57) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (519.63,124.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (526.77,122.12) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (533.91,109.26) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (548.20, 87.83) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (562.49, 73.55) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90,137.84) --
	(333.90,137.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90,137.84) --
	(333.90,137.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90,137.84) --
	(333.90,137.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.18,236.60) --
	(348.18,236.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.18,236.60) --
	(348.18,191.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.18,191.59) --
	(348.18,191.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.33,227.49) --
	(355.33,227.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.33,227.49) --
	(355.33,152.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.33,152.84) --
	(355.33,152.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.47,208.56) --
	(362.47,208.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.47,208.56) --
	(362.47,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.47,181.41) --
	(362.47,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.61,236.42) --
	(369.61,236.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.61,236.42) --
	(369.61,141.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.61,141.41) --
	(369.61,141.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.76,223.56) --
	(376.76,223.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.76,223.56) --
	(376.76,223.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.76,223.56) --
	(376.76,223.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (383.90,257.49) --
	(383.90,257.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (383.90,257.49) --
	(383.90,189.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (383.90,189.63) --
	(383.90,189.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (391.04,208.56) --
	(391.04,208.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (391.04,208.56) --
	(391.04,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (391.04,181.41) --
	(391.04,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.19,216.42) --
	(398.19,216.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.19,216.42) --
	(398.19,216.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.19,216.42) --
	(398.19,216.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.33,214.81) --
	(405.33,214.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.33,214.81) --
	(405.33,159.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.33,159.81) --
	(405.33,159.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.47,229.99) --
	(412.47,229.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.47,229.99) --
	(412.47,109.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.47,109.98) --
	(412.47,109.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (419.62,212.85) --
	(419.62,212.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (419.62,212.85) --
	(419.62,126.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (419.62,126.23) --
	(419.62,126.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (426.76,198.92) --
	(426.76,198.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (426.76,198.92) --
	(426.76,122.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (426.76,122.84) --
	(426.76,122.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (433.91,212.31) --
	(433.91,212.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (433.91,212.31) --
	(433.91,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (433.91,109.26) --
	(433.91,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (441.05,223.56) --
	(441.05,223.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (441.05,223.56) --
	(441.05,122.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (441.05,122.30) --
	(441.05,122.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (448.19,222.31) --
	(448.19,222.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (448.19,222.31) --
	(448.19, 91.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (448.19, 91.58) --
	(448.19, 91.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (455.34,191.77) --
	(455.34,191.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (455.34,191.77) --
	(455.34,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (455.34,109.26) --
	(455.34,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.48,190.52) --
	(462.48,190.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.48,190.52) --
	(462.48,116.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.48,116.41) --
	(462.48,116.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.62,167.13) --
	(469.62,167.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.62,167.13) --
	(469.62, 87.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.62, 87.48) --
	(469.62, 87.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.77,193.74) --
	(476.77,193.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.77,193.74) --
	(476.77, 80.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.77, 80.69) --
	(476.77, 80.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (483.91,181.77) --
	(483.91,181.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (483.91,181.77) --
	(483.91, 87.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (483.91, 87.48) --
	(483.91, 87.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.05,172.84) --
	(491.05,172.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.05,172.84) --
	(491.05, 44.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.05, 44.97) --
	(491.05, 44.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (498.20,167.84) --
	(498.20,167.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (498.20,167.84) --
	(498.20, 70.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (498.20, 70.69) --
	(498.20, 70.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (505.34,144.98) --
	(505.34,144.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (505.34,144.98) --
	(505.34, 79.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (505.34, 79.26) --
	(505.34, 79.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (512.48,142.13) --
	(512.48,142.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (512.48,142.13) --
	(512.48, 92.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (512.48, 92.12) --
	(512.48, 92.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (519.63,144.98) --
	(519.63,144.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (519.63,144.98) --
	(519.63, 98.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (519.63, 98.55) --
	(519.63, 98.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77,130.70) --
	(526.77,130.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77,130.70) --
	(526.77,109.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77,109.98) --
	(526.77,109.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.91,109.26) --
	(533.91,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.91,109.26) --
	(533.91,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.91,109.26) --
	(533.91,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (548.20, 87.83) --
	(548.20, 87.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (548.20, 87.83) --
	(548.20, 87.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (548.20, 87.83) --
	(548.20, 87.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (562.49,107.48) --
	(562.49,107.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (562.49,107.48) --
	(562.49, 39.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (562.49, 39.62) --
	(562.49, 39.62);

\path[draw=drawColor,line width= 1.1pt,line join=round] (333.90,212.08) --
	(336.79,210.73) --
	(339.68,209.35) --
	(342.58,207.96) --
	(345.47,206.55) --
	(348.36,205.13) --
	(351.26,203.68) --
	(354.15,202.22) --
	(357.04,200.75) --
	(359.94,199.25) --
	(362.83,197.73) --
	(365.73,196.20) --
	(368.62,194.64) --
	(371.51,193.07) --
	(374.41,191.48) --
	(377.30,189.86) --
	(380.19,188.23) --
	(383.09,186.57) --
	(385.98,184.90) --
	(388.87,183.20) --
	(391.77,181.48) --
	(394.66,179.74) --
	(397.55,177.98) --
	(400.45,176.19) --
	(403.34,174.39) --
	(406.24,172.56) --
	(409.13,170.70) --
	(412.02,168.82) --
	(414.92,166.96) --
	(417.81,165.20) --
	(420.70,163.53) --
	(423.60,161.96) --
	(426.49,160.49) --
	(429.38,159.13) --
	(432.28,157.88) --
	(435.17,156.76) --
	(438.06,155.79) --
	(440.96,154.74) --
	(443.85,153.38) --
	(446.75,151.87) --
	(449.64,150.42) --
	(452.53,148.83) --
	(455.43,147.14) --
	(458.32,145.41) --
	(461.21,143.73) --
	(464.11,142.08) --
	(467.00,140.32) --
	(469.89,138.66) --
	(472.79,137.15) --
	(475.68,135.67) --
	(478.58,134.18) --
	(481.47,132.70) --
	(484.36,131.23) --
	(487.26,129.77) --
	(490.15,128.34) --
	(493.04,126.97) --
	(495.94,125.71) --
	(498.83,124.58) --
	(501.72,123.44) --
	(504.62,122.24) --
	(507.51,120.96) --
	(510.40,119.62) --
	(513.30,118.21) --
	(516.19,116.73) --
	(519.09,115.18) --
	(521.98,113.56) --
	(524.87,111.88) --
	(527.77,110.13) --
	(530.66,108.32) --
	(533.55,106.44) --
	(536.45,104.50) --
	(539.34,102.49) --
	(542.23,100.42) --
	(545.13, 98.29) --
	(548.02, 96.10) --
	(550.91, 93.84) --
	(553.81, 91.52) --
	(556.70, 89.14) --
	(559.60, 86.70) --
	(562.49, 84.20);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.90, 30.69) rectangle (562.49,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (328.95, 41.94) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (328.95,113.38) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (328.95,184.81) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (328.95,256.25) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.15, 44.97) --
	(333.90, 44.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.15,116.41) --
	(333.90,116.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.15,187.84) --
	(333.90,187.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.15,259.28) --
	(333.90,259.28);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.76, 27.94) --
	(376.76, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (426.76, 27.94) --
	(426.76, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.77, 27.94) --
	(476.77, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77, 27.94) --
	(526.77, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (376.76, 19.68) {Jan 13};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (426.76, 19.68) {Jan 20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (476.77, 19.68) {Jan 27};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (526.77, 19.68) {Feb 03};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (448.19,  7.64) {Symptom onset date (infector)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (312.33,148.55) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (333.90,274.49) {B. Forward serial interval};
\end{scope}
\begin{scope}
\path[clip] (584.76,  0.00) rectangle (860.64,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (584.76,  0.00) rectangle (860.64,289.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (619.40, 30.69) rectangle (855.14,266.42);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (673.17,204.80) {maximum observable};

\node[text=drawColor,rotate= 45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (806.75, 79.07) {minimum observable};

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (619.40,140.70) --
	(627.26,148.55) --
	(635.12,156.41) --
	(642.98,164.27) --
	(650.84,172.13) --
	(658.69,179.99) --
	(666.55,187.84) --
	(674.41,195.70) --
	(682.27,203.56) --
	(690.13,211.42) --
	(697.98,219.28) --
	(705.84,227.13) --
	(713.70,234.99) --
	(721.56,242.85) --
	(729.41,250.71) --
	(737.27,258.56) --
	(745.13,266.42);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (768.70, 30.69) --
	(776.56, 38.54) --
	(784.42, 46.40) --
	(792.28, 54.26) --
	(800.14, 62.12) --
	(807.99, 69.98) --
	(815.85, 77.83) --
	(823.71, 85.69) --
	(831.57, 93.55) --
	(839.43,101.41) --
	(847.28,109.26) --
	(855.14,117.12);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (619.40,140.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (650.84, 46.40) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (658.69,148.55) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (666.55,101.41) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (674.41,172.13) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (682.27,117.12) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (690.13,104.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (697.98,129.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (705.84,111.23) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (713.70,124.46) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (721.56,136.28) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (729.41,141.20) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (737.27,143.14) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (745.13,141.78) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (752.99,132.65) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (760.85,145.10) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (768.70,152.05) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (776.56,153.79) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (784.42,151.04) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (792.28,153.41) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (800.14,157.72) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (807.99,166.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (815.85,171.71) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (823.71,169.11) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (831.57,196.68) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (839.43,186.53) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (847.28,242.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (855.14,211.42) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.40,140.70) --
	(619.40,140.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.40,140.70) --
	(619.40,140.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.40,140.70) --
	(619.40,140.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (650.84, 46.40) --
	(650.84, 46.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (650.84, 46.40) --
	(650.84, 46.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (650.84, 46.40) --
	(650.84, 46.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.69,156.02) --
	(658.69,156.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.69,156.02) --
	(658.69,141.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.69,141.09) --
	(658.69,141.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.55,101.41) --
	(666.55,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.55,101.41) --
	(666.55,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.55,101.41) --
	(666.55,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.41,172.13) --
	(674.41,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.41,172.13) --
	(674.41,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.41,172.13) --
	(674.41,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (682.27,117.12) --
	(682.27,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (682.27,117.12) --
	(682.27,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (682.27,117.12) --
	(682.27,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.13,185.09) --
	(690.13,185.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.13,185.09) --
	(690.13, 38.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.13, 38.54) --
	(690.13, 38.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (697.98,181.56) --
	(697.98,181.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (697.98,181.56) --
	(697.98,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (697.98,109.26) --
	(697.98,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (705.84,139.32) --
	(705.84,139.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (705.84,139.32) --
	(705.84, 71.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (705.84, 71.35) --
	(705.84, 71.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (713.70,194.52) --
	(713.70,194.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (713.70,194.52) --
	(713.70, 62.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (713.70, 62.51) --
	(713.70, 62.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (721.56,227.13) --
	(721.56,227.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (721.56,227.13) --
	(721.56, 83.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (721.56, 83.73) --
	(721.56, 83.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (729.41,223.20) --
	(729.41,223.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (729.41,223.20) --
	(729.41, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (729.41, 85.69) --
	(729.41, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (737.27,242.85) --
	(737.27,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (737.27,242.85) --
	(737.27, 86.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (737.27, 86.48) --
	(737.27, 86.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (745.13,245.21) --
	(745.13,245.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (745.13,245.21) --
	(745.13, 96.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (745.13, 96.69) --
	(745.13, 96.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (752.99,187.84) --
	(752.99,187.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (752.99,187.84) --
	(752.99, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (752.99, 85.69) --
	(752.99, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (760.85,219.28) --
	(760.85,219.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (760.85,219.28) --
	(760.85, 93.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (760.85, 93.55) --
	(760.85, 93.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (768.70,209.45) --
	(768.70,209.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (768.70,209.45) --
	(768.70,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (768.70,117.12) --
	(768.70,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (776.56,206.51) --
	(776.56,206.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (776.56,206.51) --
	(776.56,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (776.56,107.30) --
	(776.56,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (784.42,205.33) --
	(784.42,205.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (784.42,205.33) --
	(784.42,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (784.42,109.26) --
	(784.42,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (792.28,198.45) --
	(792.28,198.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (792.28,198.45) --
	(792.28,115.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (792.28,115.75) --
	(792.28,115.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (800.14,221.44) --
	(800.14,221.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (800.14,221.44) --
	(800.14, 99.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (800.14, 99.25) --
	(800.14, 99.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (807.99,253.65) --
	(807.99,253.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (807.99,253.65) --
	(807.99,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (807.99,109.26) --
	(807.99,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (815.85,240.10) --
	(815.85,240.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (815.85,240.10) --
	(815.85,128.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (815.85,128.52) --
	(815.85,128.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (823.71,209.06) --
	(823.71,209.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (823.71,209.06) --
	(823.71,132.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (823.71,132.84) --
	(823.71,132.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (831.57,234.99) --
	(831.57,234.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (831.57,234.99) --
	(831.57,148.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (831.57,148.55) --
	(831.57,148.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (839.43,231.06) --
	(839.43,231.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (839.43,231.06) --
	(839.43,118.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (839.43,118.10) --
	(839.43,118.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (847.28,242.85) --
	(847.28,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (847.28,242.85) --
	(847.28,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (847.28,242.85) --
	(847.28,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (855.14,211.42) --
	(855.14,211.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (855.14,211.42) --
	(855.14,211.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (855.14,211.42) --
	(855.14,211.42);

\path[draw=drawColor,line width= 1.1pt,line join=round] (619.40,118.83) --
	(622.39,118.41) --
	(625.37,118.04) --
	(628.36,117.71) --
	(631.34,117.43) --
	(634.32,117.20) --
	(637.31,117.01) --
	(640.29,116.87) --
	(643.28,116.77) --
	(646.26,116.72) --
	(649.24,116.72) --
	(652.23,116.77) --
	(655.21,116.87) --
	(658.20,117.01) --
	(661.18,117.21) --
	(664.16,117.45) --
	(667.15,117.75) --
	(670.13,118.09) --
	(673.12,118.49) --
	(676.10,118.94) --
	(679.08,119.44) --
	(682.07,119.99) --
	(685.05,120.59) --
	(688.04,121.25) --
	(691.02,121.96) --
	(694.00,122.73) --
	(696.99,123.54) --
	(699.97,124.42) --
	(702.96,125.35) --
	(705.94,126.33) --
	(708.92,127.37) --
	(711.91,128.47) --
	(714.89,129.62) --
	(717.88,130.83) --
	(720.86,132.10) --
	(723.84,133.38) --
	(726.83,134.61) --
	(729.81,135.77) --
	(732.80,136.88) --
	(735.78,137.92) --
	(738.76,138.87) --
	(741.75,139.56) --
	(744.73,140.07) --
	(747.72,140.53) --
	(750.70,141.04) --
	(753.68,141.73) --
	(756.67,142.49) --
	(759.65,143.35) --
	(762.64,144.61) --
	(765.62,146.56) --
	(768.60,148.26) --
	(771.59,149.31) --
	(774.57,150.22) --
	(777.56,151.24) --
	(780.54,152.28) --
	(783.52,153.32) --
	(786.51,154.37) --
	(789.49,155.31) --
	(792.48,156.26) --
	(795.46,157.32) --
	(798.44,158.61) --
	(801.43,160.23) --
	(804.41,162.05) --
	(807.40,164.02) --
	(810.38,166.15) --
	(813.36,168.43) --
	(816.35,170.87) --
	(819.33,173.46) --
	(822.32,176.21) --
	(825.30,179.13) --
	(828.28,182.20) --
	(831.27,185.42) --
	(834.25,188.81) --
	(837.24,192.36) --
	(840.22,196.07) --
	(843.20,199.95) --
	(846.19,203.98) --
	(849.17,208.18) --
	(852.16,212.54) --
	(855.14,217.06);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (619.40, 30.69) rectangle (855.14,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (614.45, 35.51) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (614.45,114.09) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (614.45,192.67) {10};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.65, 38.54) --
	(619.40, 38.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.65,117.12) --
	(619.40,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.65,195.70) --
	(619.40,195.70);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (642.98, 27.94) --
	(642.98, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (697.98, 27.94) --
	(697.98, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (752.99, 27.94) --
	(752.99, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (807.99, 27.94) --
	(807.99, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (642.98, 19.68) {Jan 13};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (697.98, 19.68) {Jan 20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (752.99, 19.68) {Jan 27};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (807.99, 19.68) {Feb 03};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (737.27,  7.64) {Symptom onset date (infectee)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (597.84,148.55) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (619.40,274.49) {C. Backward serial interval};
\end{scope}
\begin{scope}
\path[clip] (867.24,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (867.24,  0.00) rectangle (1156.32,289.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (894.55, 30.69) rectangle (1150.82,266.42);
\definecolor{fillColor}{RGB}{240,228,66}

\path[fill=fillColor,fill opacity=0.50] (918.58,255.71) --
	(926.59,234.51) --
	(934.59,227.85) --
	(942.60,224.34) --
	(950.61,232.95) --
	(958.62,225.18) --
	(966.63,220.08) --
	(974.64,212.60) --
	(982.64,137.83) --
	(990.65,119.60) --
	(998.66,116.21) --
	(1006.67,104.34) --
	(1014.68,101.52) --
	(1022.69, 91.89) --
	(1030.69, 87.88) --
	(1038.70, 85.77) --
	(1046.71, 79.62) --
	(1054.72, 76.53) --
	(1062.73, 72.74) --
	(1070.74, 71.32) --
	(1078.74, 69.30) --
	(1086.75, 68.00) --
	(1094.76, 66.17) --
	(1102.77, 65.86) --
	(1110.78, 65.06) --
	(1118.79, 64.84) --
	(1126.79, 64.84) --
	(1134.80, 64.82) --
	(1142.81, 64.82) --
	(1150.82, 64.08) --
	(1150.82, 56.77) --
	(1142.81, 57.72) --
	(1134.80, 57.72) --
	(1126.79, 58.18) --
	(1118.79, 58.18) --
	(1110.78, 58.40) --
	(1102.77, 58.16) --
	(1094.76, 58.79) --
	(1086.75, 60.03) --
	(1078.74, 60.94) --
	(1070.74, 62.46) --
	(1062.73, 65.05) --
	(1054.72, 66.78) --
	(1046.71, 70.56) --
	(1038.70, 73.93) --
	(1030.69, 74.95) --
	(1022.69, 77.75) --
	(1014.68, 84.76) --
	(1006.67, 84.17) --
	(998.66, 90.54) --
	(990.65, 89.60) --
	(982.64, 93.27) --
	(974.64,129.71) --
	(966.63,137.14) --
	(958.62,136.34) --
	(950.61,134.94) --
	(942.60,135.04) --
	(934.59,120.43) --
	(926.59,124.23) --
	(918.58,119.49) --
	cycle;
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (918.58,255.71) --
	(926.59,234.51) --
	(934.59,227.85) --
	(942.60,224.34) --
	(950.61,232.95) --
	(958.62,225.18) --
	(966.63,220.08) --
	(974.64,212.60) --
	(982.64,137.83) --
	(990.65,119.60) --
	(998.66,116.21) --
	(1006.67,104.34) --
	(1014.68,101.52) --
	(1022.69, 91.89) --
	(1030.69, 87.88) --
	(1038.70, 85.77) --
	(1046.71, 79.62) --
	(1054.72, 76.53) --
	(1062.73, 72.74) --
	(1070.74, 71.32) --
	(1078.74, 69.30) --
	(1086.75, 68.00) --
	(1094.76, 66.17) --
	(1102.77, 65.86) --
	(1110.78, 65.06) --
	(1118.79, 64.84) --
	(1126.79, 64.84) --
	(1134.80, 64.82) --
	(1142.81, 64.82) --
	(1150.82, 64.08);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (1150.82, 56.77) --
	(1142.81, 57.72) --
	(1134.80, 57.72) --
	(1126.79, 58.18) --
	(1118.79, 58.18) --
	(1110.78, 58.40) --
	(1102.77, 58.16) --
	(1094.76, 58.79) --
	(1086.75, 60.03) --
	(1078.74, 60.94) --
	(1070.74, 62.46) --
	(1062.73, 65.05) --
	(1054.72, 66.78) --
	(1046.71, 70.56) --
	(1038.70, 73.93) --
	(1030.69, 74.95) --
	(1022.69, 77.75) --
	(1014.68, 84.76) --
	(1006.67, 84.17) --
	(998.66, 90.54) --
	(990.65, 89.60) --
	(982.64, 93.27) --
	(974.64,129.71) --
	(966.63,137.14) --
	(958.62,136.34) --
	(950.61,134.94) --
	(942.60,135.04) --
	(934.59,120.43) --
	(926.59,124.23) --
	(918.58,119.49);
\definecolor{fillColor}{RGB}{0,114,178}

\path[fill=fillColor,fill opacity=0.50] (918.58,167.99) --
	(926.59,158.37) --
	(934.59,155.61) --
	(942.60,153.41) --
	(950.61,158.49) --
	(958.62,154.03) --
	(966.63,151.93) --
	(974.64,148.30) --
	(982.64,107.71) --
	(990.65, 96.60) --
	(998.66, 94.03) --
	(1006.67, 86.26) --
	(1014.68, 84.52) --
	(1022.69, 77.77) --
	(1030.69, 75.45) --
	(1038.70, 73.98) --
	(1046.71, 69.74) --
	(1054.72, 67.73) --
	(1062.73, 65.35) --
	(1070.74, 64.41) --
	(1078.74, 62.76) --
	(1086.75, 61.94) --
	(1094.76, 60.56) --
	(1102.77, 60.51) --
	(1110.78, 60.01) --
	(1118.79, 59.88) --
	(1126.79, 59.88) --
	(1134.80, 59.77) --
	(1142.81, 59.77) --
	(1150.82, 59.40) --
	(1150.82, 54.49) --
	(1142.81, 54.89) --
	(1134.80, 54.89) --
	(1126.79, 55.23) --
	(1118.79, 55.23) --
	(1110.78, 55.32) --
	(1102.77, 55.43) --
	(1094.76, 55.76) --
	(1086.75, 56.48) --
	(1078.74, 57.32) --
	(1070.74, 58.34) --
	(1062.73, 59.85) --
	(1054.72, 61.35) --
	(1046.71, 63.63) --
	(1038.70, 66.52) --
	(1030.69, 66.97) --
	(1022.69, 68.74) --
	(1014.68, 73.15) --
	(1006.67, 73.04) --
	(998.66, 77.40) --
	(990.65, 77.19) --
	(982.64, 79.82) --
	(974.64,101.35) --
	(966.63,108.09) --
	(958.62,107.05) --
	(950.61,105.61) --
	(942.60,105.29) --
	(934.59, 96.62) --
	(926.59, 98.80) --
	(918.58, 95.78) --
	cycle;
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round,line cap=round] (918.58,167.99) --
	(926.59,158.37) --
	(934.59,155.61) --
	(942.60,153.41) --
	(950.61,158.49) --
	(958.62,154.03) --
	(966.63,151.93) --
	(974.64,148.30) --
	(982.64,107.71) --
	(990.65, 96.60) --
	(998.66, 94.03) --
	(1006.67, 86.26) --
	(1014.68, 84.52) --
	(1022.69, 77.77) --
	(1030.69, 75.45) --
	(1038.70, 73.98) --
	(1046.71, 69.74) --
	(1054.72, 67.73) --
	(1062.73, 65.35) --
	(1070.74, 64.41) --
	(1078.74, 62.76) --
	(1086.75, 61.94) --
	(1094.76, 60.56) --
	(1102.77, 60.51) --
	(1110.78, 60.01) --
	(1118.79, 59.88) --
	(1126.79, 59.88) --
	(1134.80, 59.77) --
	(1142.81, 59.77) --
	(1150.82, 59.40);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round,line cap=round] (1150.82, 54.49) --
	(1142.81, 54.89) --
	(1134.80, 54.89) --
	(1126.79, 55.23) --
	(1118.79, 55.23) --
	(1110.78, 55.32) --
	(1102.77, 55.43) --
	(1094.76, 55.76) --
	(1086.75, 56.48) --
	(1078.74, 57.32) --
	(1070.74, 58.34) --
	(1062.73, 59.85) --
	(1054.72, 61.35) --
	(1046.71, 63.63) --
	(1038.70, 66.52) --
	(1030.69, 66.97) --
	(1022.69, 68.74) --
	(1014.68, 73.15) --
	(1006.67, 73.04) --
	(998.66, 77.40) --
	(990.65, 77.19) --
	(982.64, 79.82) --
	(974.64,101.35) --
	(966.63,108.09) --
	(958.62,107.05) --
	(950.61,105.61) --
	(942.60,105.29) --
	(934.59, 96.62) --
	(926.59, 98.80) --
	(918.58, 95.78);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (894.55, 41.40) -- (1150.82, 41.40);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (918.58,170.94) --
	(926.59,170.79) --
	(934.59,163.30) --
	(942.60,167.34) --
	(950.61,172.68) --
	(958.62,172.43) --
	(966.63,174.86) --
	(974.64,167.73) --
	(982.64,112.65) --
	(990.65,105.04) --
	(998.66,102.84) --
	(1006.67, 93.38) --
	(1014.68, 91.84) --
	(1022.69, 83.99) --
	(1030.69, 80.95) --
	(1038.70, 80.00) --
	(1046.71, 75.18) --
	(1054.72, 71.77) --
	(1062.73, 68.56) --
	(1070.74, 66.30) --
	(1078.74, 65.11) --
	(1086.75, 63.71) --
	(1094.76, 62.37) --
	(1102.77, 61.91) --
	(1110.78, 61.69) --
	(1118.79, 61.52) --
	(1126.79, 61.52) --
	(1134.80, 61.33) --
	(1142.81, 61.33) --
	(1150.82, 60.65);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (918.58,126.39) --
	(926.59,125.79) --
	(934.59,122.09) --
	(942.60,124.37) --
	(950.61,127.48) --
	(958.62,126.99) --
	(966.63,128.28) --
	(974.64,123.96) --
	(982.64, 92.40) --
	(990.65, 87.06) --
	(998.66, 85.30) --
	(1006.67, 79.23) --
	(1014.68, 77.99) --
	(1022.69, 73.04) --
	(1030.69, 70.90) --
	(1038.70, 70.16) --
	(1046.71, 66.88) --
	(1054.72, 64.57) --
	(1062.73, 62.37) --
	(1070.74, 60.94) --
	(1078.74, 60.13) --
	(1086.75, 59.14) --
	(1094.76, 58.16) --
	(1102.77, 57.82) --
	(1110.78, 57.64) --
	(1118.79, 57.52) --
	(1126.79, 57.52) --
	(1134.80, 57.39) --
	(1142.81, 57.39) --
	(1150.82, 56.99);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (894.55, 30.69) rectangle (1150.82,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (889.60, 38.37) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (889.60, 90.38) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (889.60,142.40) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (889.60,194.41) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (889.60,246.42) {5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (891.80, 41.40) --
	(894.55, 41.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (891.80, 93.41) --
	(894.55, 93.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (891.80,145.43) --
	(894.55,145.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (891.80,197.44) --
	(894.55,197.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (891.80,249.45) --
	(894.55,249.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (918.58, 27.94) --
	(918.58, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (998.66, 27.94) --
	(998.66, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (1078.74, 27.94) --
	(1078.74, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (918.58, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (998.66, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (1078.74, 19.68) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (1022.69,  7.64) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (880.32,148.55) {Basic reproduction number $\mathcal R_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (1011.05,208.36) rectangle (1126.58,253.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (1016.55,228.31) rectangle (1031.00,242.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{240,228,66}
\definecolor{fillColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor,fill opacity=0.50] (1017.26,229.02) rectangle (1030.29,242.05);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (1017.99,235.54) -- (1029.56,235.54);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (1016.55,213.86) rectangle (1031.00,228.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,114,178}
\definecolor{fillColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line cap=rect,fill=fillColor,fill opacity=0.50] (1017.26,214.57) rectangle (1030.29,227.60);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (1017.99,221.09) -- (1029.56,221.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (1036.50,232.51) {6 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (1036.50,218.05) {8 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (1156.32,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (894.55,274.49) {D. Cohort-averaged $\mathcal R_0$ estimates};
\end{scope}
\end{tikzpicture}

\end{document}