### 2020-09-20 serial_analysis.tex
forgot to commit

% Created by tikzDevice version 0.12.3 on 2020-09-20 18:46:20
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,289.08) rectangle (289.08,578.16);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,289.08) rectangle (289.08,578.16);
\end{scope}
\begin{scope}
\path[clip] ( 31.71,319.77) rectangle (283.58,555.50);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.71,319.77) rectangle (283.58,555.50);
\definecolor{fillColor}{gray}{0.35}

\path[fill=fillColor] ( 31.71,319.77) rectangle ( 38.40,323.18);

\path[fill=fillColor] ( 46.57,319.77) rectangle ( 53.26,326.60);

\path[fill=fillColor] ( 54.00,319.77) rectangle ( 60.69,330.02);

\path[fill=fillColor] ( 61.43,319.77) rectangle ( 68.12,326.60);

\path[fill=fillColor] ( 68.86,319.77) rectangle ( 75.55,330.02);

\path[fill=fillColor] ( 76.29,319.77) rectangle ( 82.98,323.18);

\path[fill=fillColor] ( 83.72,319.77) rectangle ( 90.41,326.60);

\path[fill=fillColor] ( 91.15,319.77) rectangle ( 97.84,330.02);

\path[fill=fillColor] ( 98.58,319.77) rectangle (105.27,326.60);

\path[fill=fillColor] (106.01,319.77) rectangle (112.70,333.43);

\path[fill=fillColor] (113.44,319.77) rectangle (120.13,374.43);

\path[fill=fillColor] (120.87,319.77) rectangle (127.56,374.43);

\path[fill=fillColor] (128.30,319.77) rectangle (134.99,381.26);

\path[fill=fillColor] (135.73,319.77) rectangle (142.42,384.68);

\path[fill=fillColor] (143.16,319.77) rectangle (149.84,439.34);

\path[fill=fillColor] (150.59,319.77) rectangle (157.27,483.76);

\path[fill=fillColor] (158.02,319.77) rectangle (164.70,494.01);

\path[fill=fillColor] (165.45,319.77) rectangle (172.13,487.17);

\path[fill=fillColor] (172.88,319.77) rectangle (179.56,507.67);

\path[fill=fillColor] (180.31,319.77) rectangle (186.99,535.00);

\path[fill=fillColor] (187.74,319.77) rectangle (194.42,545.25);

\path[fill=fillColor] (195.17,319.77) rectangle (201.85,487.17);

\path[fill=fillColor] (202.60,319.77) rectangle (209.28,476.92);

\path[fill=fillColor] (210.03,319.77) rectangle (216.71,480.34);

\path[fill=fillColor] (217.46,319.77) rectangle (224.14,466.67);

\path[fill=fillColor] (224.89,319.77) rectangle (231.57,425.68);

\path[fill=fillColor] (232.31,319.77) rectangle (239.00,405.18);

\path[fill=fillColor] (239.74,319.77) rectangle (246.43,391.51);

\path[fill=fillColor] (247.17,319.77) rectangle (253.86,364.18);

\path[fill=fillColor] (254.60,319.77) rectangle (261.29,350.51);

\path[fill=fillColor] (262.03,319.77) rectangle (268.72,340.26);

\path[fill=fillColor] (269.46,319.77) rectangle (276.15,326.60);

\path[fill=fillColor] (276.89,319.77) rectangle (283.58,323.18);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.71,319.77) rectangle (283.58,555.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,316.74) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,385.06) {20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,453.39) {40};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,521.72) {60};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,319.77) --
	( 31.71,319.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,388.10) --
	( 31.71,388.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,456.42) --
	( 31.71,456.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,524.75) --
	( 31.71,524.75);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.63,317.02) --
	( 79.63,319.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (131.64,317.02) --
	(131.64,319.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (183.65,317.02) --
	(183.65,319.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (235.66,317.02) --
	(235.66,319.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 79.63,308.76) {Jan 13};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (131.64,308.76) {Jan 20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (183.65,308.76) {Jan 27};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (235.66,308.76) {Feb 03};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (157.65,296.72) {Symptom onset date};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,437.63) {Number of cases with a known infector/infectee pair};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 31.71,563.57) {A. Epidemic curve};
\end{scope}
\begin{scope}
\path[clip] (299.25,289.08) rectangle (567.99,578.16);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (299.25,289.08) rectangle (567.99,578.16);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (333.90,319.77) rectangle (562.49,555.50);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate=-45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (499.08,484.95) {maximum observable};

\node[text=drawColor,rotate=-45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (384.78,363.51) {minimum observable};

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (419.62,555.50) --
	(426.76,548.36) --
	(433.91,541.22) --
	(441.05,534.07) --
	(448.19,526.93) --
	(455.34,519.79) --
	(462.48,512.64) --
	(469.62,505.50) --
	(476.77,498.35) --
	(483.91,491.21) --
	(491.05,484.07) --
	(498.20,476.92) --
	(505.34,469.78) --
	(512.48,462.64) --
	(519.63,455.49) --
	(526.77,448.35) --
	(533.91,441.21) --
	(541.06,434.06) --
	(548.20,426.92) --
	(555.35,419.78) --
	(562.49,412.63);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (333.90,405.49) --
	(341.04,398.34) --
	(348.18,391.20) --
	(355.33,384.06) --
	(362.47,376.91) --
	(369.61,369.77) --
	(376.76,362.63) --
	(383.90,355.48) --
	(391.04,348.34) --
	(398.19,341.20) --
	(405.33,334.05) --
	(412.47,326.91) --
	(419.62,319.77);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (333.90,426.92) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (348.18,508.18) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (355.33,472.16) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (362.47,484.07) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (369.61,484.07) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (376.76,512.64) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (383.90,512.64) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (391.04,484.07) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (398.19,505.50) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (405.33,473.35) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (412.47,448.66) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (419.62,448.80) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (426.76,454.37) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (433.91,439.12) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (441.05,450.03) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (448.19,438.38) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (455.34,433.82) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (462.48,439.01) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (469.62,422.34) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (476.77,427.37) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (483.91,419.04) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (491.05,416.36) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (498.20,416.41) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (505.34,408.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (512.48,404.65) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (519.63,413.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (526.77,411.20) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (533.91,398.34) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (548.20,376.91) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (562.49,362.63) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90,426.92) --
	(333.90,426.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90,426.92) --
	(333.90,426.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.90,426.92) --
	(333.90,426.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.18,525.68) --
	(348.18,525.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.18,525.68) --
	(348.18,480.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.18,480.67) --
	(348.18,480.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.33,516.57) --
	(355.33,516.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.33,516.57) --
	(355.33,441.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.33,441.92) --
	(355.33,441.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.47,497.64) --
	(362.47,497.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.47,497.64) --
	(362.47,470.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.47,470.49) --
	(362.47,470.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.61,525.50) --
	(369.61,525.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.61,525.50) --
	(369.61,430.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.61,430.49) --
	(369.61,430.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.76,512.64) --
	(376.76,512.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.76,512.64) --
	(376.76,512.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.76,512.64) --
	(376.76,512.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (383.90,546.57) --
	(383.90,546.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (383.90,546.57) --
	(383.90,478.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (383.90,478.71) --
	(383.90,478.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (391.04,497.64) --
	(391.04,497.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (391.04,497.64) --
	(391.04,470.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (391.04,470.49) --
	(391.04,470.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.19,505.50) --
	(398.19,505.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.19,505.50) --
	(398.19,505.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.19,505.50) --
	(398.19,505.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.33,503.89) --
	(405.33,503.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.33,503.89) --
	(405.33,448.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.33,448.89) --
	(405.33,448.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.47,519.07) --
	(412.47,519.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.47,519.07) --
	(412.47,399.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.47,399.06) --
	(412.47,399.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (419.62,501.93) --
	(419.62,501.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (419.62,501.93) --
	(419.62,415.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (419.62,415.31) --
	(419.62,415.31);

\path[draw=drawColor,line width= 0.6pt,line join=round] (426.76,488.00) --
	(426.76,488.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (426.76,488.00) --
	(426.76,411.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (426.76,411.92) --
	(426.76,411.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (433.91,501.39) --
	(433.91,501.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (433.91,501.39) --
	(433.91,398.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (433.91,398.34) --
	(433.91,398.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (441.05,512.64) --
	(441.05,512.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (441.05,512.64) --
	(441.05,411.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (441.05,411.38) --
	(441.05,411.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (448.19,511.39) --
	(448.19,511.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (448.19,511.39) --
	(448.19,380.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (448.19,380.66) --
	(448.19,380.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (455.34,480.85) --
	(455.34,480.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (455.34,480.85) --
	(455.34,398.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (455.34,398.34) --
	(455.34,398.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.48,479.60) --
	(462.48,479.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.48,479.60) --
	(462.48,405.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.48,405.49) --
	(462.48,405.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.62,456.21) --
	(469.62,456.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.62,456.21) --
	(469.62,376.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.62,376.56) --
	(469.62,376.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.77,482.82) --
	(476.77,482.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.77,482.82) --
	(476.77,369.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.77,369.77) --
	(476.77,369.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (483.91,470.85) --
	(483.91,470.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (483.91,470.85) --
	(483.91,376.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (483.91,376.56) --
	(483.91,376.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.05,461.92) --
	(491.05,461.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.05,461.92) --
	(491.05,334.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.05,334.05) --
	(491.05,334.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (498.20,456.92) --
	(498.20,456.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (498.20,456.92) --
	(498.20,359.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (498.20,359.77) --
	(498.20,359.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (505.34,434.06) --
	(505.34,434.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (505.34,434.06) --
	(505.34,368.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (505.34,368.34) --
	(505.34,368.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (512.48,431.21) --
	(512.48,431.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (512.48,431.21) --
	(512.48,381.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (512.48,381.20) --
	(512.48,381.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (519.63,434.06) --
	(519.63,434.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (519.63,434.06) --
	(519.63,387.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (519.63,387.63) --
	(519.63,387.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77,419.78) --
	(526.77,419.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77,419.78) --
	(526.77,399.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77,399.06) --
	(526.77,399.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.91,398.34) --
	(533.91,398.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.91,398.34) --
	(533.91,398.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.91,398.34) --
	(533.91,398.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (548.20,376.91) --
	(548.20,376.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (548.20,376.91) --
	(548.20,376.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (548.20,376.91) --
	(548.20,376.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (562.49,396.56) --
	(562.49,396.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (562.49,396.56) --
	(562.49,328.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (562.49,328.70) --
	(562.49,328.70);

\path[draw=drawColor,line width= 1.1pt,line join=round] (333.90,501.16) --
	(336.79,499.81) --
	(339.68,498.43) --
	(342.58,497.04) --
	(345.47,495.63) --
	(348.36,494.21) --
	(351.26,492.76) --
	(354.15,491.30) --
	(357.04,489.83) --
	(359.94,488.33) --
	(362.83,486.81) --
	(365.73,485.28) --
	(368.62,483.72) --
	(371.51,482.15) --
	(374.41,480.56) --
	(377.30,478.94) --
	(380.19,477.31) --
	(383.09,475.65) --
	(385.98,473.98) --
	(388.87,472.28) --
	(391.77,470.56) --
	(394.66,468.82) --
	(397.55,467.06) --
	(400.45,465.27) --
	(403.34,463.47) --
	(406.24,461.64) --
	(409.13,459.78) --
	(412.02,457.90) --
	(414.92,456.04) --
	(417.81,454.28) --
	(420.70,452.61) --
	(423.60,451.04) --
	(426.49,449.57) --
	(429.38,448.21) --
	(432.28,446.96) --
	(435.17,445.84) --
	(438.06,444.87) --
	(440.96,443.82) --
	(443.85,442.46) --
	(446.75,440.95) --
	(449.64,439.50) --
	(452.53,437.91) --
	(455.43,436.22) --
	(458.32,434.49) --
	(461.21,432.81) --
	(464.11,431.16) --
	(467.00,429.40) --
	(469.89,427.74) --
	(472.79,426.23) --
	(475.68,424.75) --
	(478.58,423.26) --
	(481.47,421.78) --
	(484.36,420.31) --
	(487.26,418.85) --
	(490.15,417.42) --
	(493.04,416.05) --
	(495.94,414.79) --
	(498.83,413.66) --
	(501.72,412.52) --
	(504.62,411.32) --
	(507.51,410.04) --
	(510.40,408.70) --
	(513.30,407.29) --
	(516.19,405.81) --
	(519.09,404.26) --
	(521.98,402.64) --
	(524.87,400.96) --
	(527.77,399.21) --
	(530.66,397.40) --
	(533.55,395.52) --
	(536.45,393.58) --
	(539.34,391.57) --
	(542.23,389.50) --
	(545.13,387.37) --
	(548.02,385.18) --
	(550.91,382.92) --
	(553.81,380.60) --
	(556.70,378.22) --
	(559.60,375.78) --
	(562.49,373.28);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.90,319.77) rectangle (562.49,555.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (328.95,331.02) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (328.95,402.46) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (328.95,473.89) {10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (328.95,545.33) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.15,334.05) --
	(333.90,334.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.15,405.49) --
	(333.90,405.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.15,476.92) --
	(333.90,476.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.15,548.36) --
	(333.90,548.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.76,317.02) --
	(376.76,319.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (426.76,317.02) --
	(426.76,319.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.77,317.02) --
	(476.77,319.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (526.77,317.02) --
	(526.77,319.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (376.76,308.76) {Jan 13};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (426.76,308.76) {Jan 20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (476.77,308.76) {Jan 27};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (526.77,308.76) {Feb 03};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (448.19,296.72) {Symptom onset date (infector)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (312.33,437.63) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (333.90,563.57) {B. Forward serial interval};
\end{scope}
\begin{scope}
\path[clip] (  6.60,  0.00) rectangle (282.48,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  6.60,  0.00) rectangle (282.48,289.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.24, 30.69) rectangle (276.98,266.42);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 95.01,204.80) {maximum observable};

\node[text=drawColor,rotate= 45.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (228.59, 79.07) {minimum observable};

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 41.24,140.70) --
	( 49.10,148.55) --
	( 56.96,156.41) --
	( 64.82,164.27) --
	( 72.68,172.13) --
	( 80.53,179.99) --
	( 88.39,187.84) --
	( 96.25,195.70) --
	(104.11,203.56) --
	(111.97,211.42) --
	(119.82,219.28) --
	(127.68,227.13) --
	(135.54,234.99) --
	(143.40,242.85) --
	(151.25,250.71) --
	(159.11,258.56) --
	(166.97,266.42);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (190.54, 30.69) --
	(198.40, 38.54) --
	(206.26, 46.40) --
	(214.12, 54.26) --
	(221.98, 62.12) --
	(229.83, 69.98) --
	(237.69, 77.83) --
	(245.55, 85.69) --
	(253.41, 93.55) --
	(261.27,101.41) --
	(269.12,109.26) --
	(276.98,117.12);
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 41.24,140.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 72.68, 46.40) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 80.53,148.55) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 88.39,101.41) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 96.25,172.13) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (104.11,117.12) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (111.97,104.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (119.82,129.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (127.68,111.23) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (135.54,124.46) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (143.40,136.28) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (151.25,141.20) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (159.11,143.14) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (166.97,141.78) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (174.83,132.65) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (182.69,145.10) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (190.54,152.05) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (198.40,153.79) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (206.26,151.04) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (214.12,153.41) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (221.98,157.72) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (229.83,166.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (237.69,171.71) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (245.55,169.11) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (253.41,196.68) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (261.27,186.53) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (269.12,242.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (276.98,211.42) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.24,140.70) --
	( 41.24,140.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.24,140.70) --
	( 41.24,140.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.24,140.70) --
	( 41.24,140.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 72.68, 46.40) --
	( 72.68, 46.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 72.68, 46.40) --
	( 72.68, 46.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 72.68, 46.40) --
	( 72.68, 46.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.53,156.02) --
	( 80.53,156.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.53,156.02) --
	( 80.53,141.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.53,141.09) --
	( 80.53,141.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.39,101.41) --
	( 88.39,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.39,101.41) --
	( 88.39,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.39,101.41) --
	( 88.39,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.25,172.13) --
	( 96.25,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.25,172.13) --
	( 96.25,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.25,172.13) --
	( 96.25,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (104.11,117.12) --
	(104.11,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (104.11,117.12) --
	(104.11,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (104.11,117.12) --
	(104.11,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.97,185.09) --
	(111.97,185.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.97,185.09) --
	(111.97, 38.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.97, 38.54) --
	(111.97, 38.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (119.82,181.56) --
	(119.82,181.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (119.82,181.56) --
	(119.82,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (119.82,109.26) --
	(119.82,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (127.68,139.32) --
	(127.68,139.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (127.68,139.32) --
	(127.68, 71.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (127.68, 71.35) --
	(127.68, 71.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.54,194.52) --
	(135.54,194.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.54,194.52) --
	(135.54, 62.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.54, 62.51) --
	(135.54, 62.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.40,227.13) --
	(143.40,227.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.40,227.13) --
	(143.40, 83.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.40, 83.73) --
	(143.40, 83.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.25,223.20) --
	(151.25,223.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.25,223.20) --
	(151.25, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.25, 85.69) --
	(151.25, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.11,242.85) --
	(159.11,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.11,242.85) --
	(159.11, 86.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.11, 86.48) --
	(159.11, 86.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.97,245.21) --
	(166.97,245.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.97,245.21) --
	(166.97, 96.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.97, 96.69) --
	(166.97, 96.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (174.83,187.84) --
	(174.83,187.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (174.83,187.84) --
	(174.83, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (174.83, 85.69) --
	(174.83, 85.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (182.69,219.28) --
	(182.69,219.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (182.69,219.28) --
	(182.69, 93.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (182.69, 93.55) --
	(182.69, 93.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (190.54,209.45) --
	(190.54,209.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (190.54,209.45) --
	(190.54,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (190.54,117.12) --
	(190.54,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.40,206.51) --
	(198.40,206.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.40,206.51) --
	(198.40,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (198.40,107.30) --
	(198.40,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (206.26,205.33) --
	(206.26,205.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (206.26,205.33) --
	(206.26,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (206.26,109.26) --
	(206.26,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.12,198.45) --
	(214.12,198.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.12,198.45) --
	(214.12,115.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.12,115.75) --
	(214.12,115.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.98,221.44) --
	(221.98,221.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.98,221.44) --
	(221.98, 99.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.98, 99.25) --
	(221.98, 99.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.83,253.65) --
	(229.83,253.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.83,253.65) --
	(229.83,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.83,109.26) --
	(229.83,109.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (237.69,240.10) --
	(237.69,240.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (237.69,240.10) --
	(237.69,128.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (237.69,128.52) --
	(237.69,128.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (245.55,209.06) --
	(245.55,209.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (245.55,209.06) --
	(245.55,132.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (245.55,132.84) --
	(245.55,132.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (253.41,234.99) --
	(253.41,234.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (253.41,234.99) --
	(253.41,148.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (253.41,148.55) --
	(253.41,148.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.27,231.06) --
	(261.27,231.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.27,231.06) --
	(261.27,118.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.27,118.10) --
	(261.27,118.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (269.12,242.85) --
	(269.12,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (269.12,242.85) --
	(269.12,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (269.12,242.85) --
	(269.12,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.98,211.42) --
	(276.98,211.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.98,211.42) --
	(276.98,211.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.98,211.42) --
	(276.98,211.42);

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 41.24,118.83) --
	( 44.23,118.41) --
	( 47.21,118.04) --
	( 50.20,117.71) --
	( 53.18,117.43) --
	( 56.16,117.20) --
	( 59.15,117.01) --
	( 62.13,116.87) --
	( 65.12,116.77) --
	( 68.10,116.72) --
	( 71.08,116.72) --
	( 74.07,116.77) --
	( 77.05,116.87) --
	( 80.04,117.01) --
	( 83.02,117.21) --
	( 86.00,117.45) --
	( 88.99,117.75) --
	( 91.97,118.09) --
	( 94.96,118.49) --
	( 97.94,118.94) --
	(100.92,119.44) --
	(103.91,119.99) --
	(106.89,120.59) --
	(109.88,121.25) --
	(112.86,121.96) --
	(115.84,122.73) --
	(118.83,123.54) --
	(121.81,124.42) --
	(124.80,125.35) --
	(127.78,126.33) --
	(130.76,127.37) --
	(133.75,128.47) --
	(136.73,129.62) --
	(139.72,130.83) --
	(142.70,132.10) --
	(145.68,133.38) --
	(148.67,134.61) --
	(151.65,135.77) --
	(154.64,136.88) --
	(157.62,137.92) --
	(160.60,138.87) --
	(163.59,139.56) --
	(166.57,140.07) --
	(169.56,140.53) --
	(172.54,141.04) --
	(175.52,141.73) --
	(178.51,142.49) --
	(181.49,143.35) --
	(184.48,144.61) --
	(187.46,146.56) --
	(190.44,148.26) --
	(193.43,149.31) --
	(196.41,150.22) --
	(199.40,151.24) --
	(202.38,152.28) --
	(205.36,153.32) --
	(208.35,154.37) --
	(211.33,155.31) --
	(214.32,156.26) --
	(217.30,157.32) --
	(220.28,158.61) --
	(223.27,160.23) --
	(226.25,162.05) --
	(229.24,164.02) --
	(232.22,166.15) --
	(235.20,168.43) --
	(238.19,170.87) --
	(241.17,173.46) --
	(244.16,176.21) --
	(247.14,179.13) --
	(250.12,182.20) --
	(253.11,185.42) --
	(256.09,188.81) --
	(259.08,192.36) --
	(262.06,196.07) --
	(265.04,199.95) --
	(268.03,203.98) --
	(271.01,208.18) --
	(274.00,212.54) --
	(276.98,217.06);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.24, 30.69) rectangle (276.98,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.29, 35.51) {-10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.29,114.09) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.29,192.67) {10};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.49, 38.54) --
	( 41.24, 38.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.49,117.12) --
	( 41.24,117.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 38.49,195.70) --
	( 41.24,195.70);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.82, 27.94) --
	( 64.82, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (119.82, 27.94) --
	(119.82, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (174.83, 27.94) --
	(174.83, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.83, 27.94) --
	(229.83, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 64.82, 19.68) {Jan 13};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (119.82, 19.68) {Jan 20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (174.83, 19.68) {Jan 27};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (229.83, 19.68) {Feb 03};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (159.11,  7.64) {Symptom onset date (infectee)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 19.68,148.55) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 41.24,274.49) {C. Backward serial interval};
\end{scope}
\begin{scope}
\path[clip] (289.08,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (289.08,  0.00) rectangle (578.16,289.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (316.39, 30.69) rectangle (572.66,266.42);
\definecolor{fillColor}{RGB}{240,228,66}

\path[fill=fillColor,fill opacity=0.50] (340.42,255.71) --
	(348.43,234.51) --
	(356.43,227.85) --
	(364.44,224.34) --
	(372.45,232.95) --
	(380.46,225.18) --
	(388.47,220.08) --
	(396.48,212.60) --
	(404.48,137.83) --
	(412.49,119.60) --
	(420.50,116.21) --
	(428.51,104.34) --
	(436.52,101.52) --
	(444.53, 91.89) --
	(452.53, 87.88) --
	(460.54, 85.77) --
	(468.55, 79.62) --
	(476.56, 76.53) --
	(484.57, 72.74) --
	(492.58, 71.32) --
	(500.58, 69.30) --
	(508.59, 68.00) --
	(516.60, 66.17) --
	(524.61, 65.86) --
	(532.62, 65.06) --
	(540.63, 64.84) --
	(548.63, 64.84) --
	(556.64, 64.82) --
	(564.65, 64.82) --
	(572.66, 64.08) --
	(572.66, 56.77) --
	(564.65, 57.72) --
	(556.64, 57.72) --
	(548.63, 58.18) --
	(540.63, 58.18) --
	(532.62, 58.40) --
	(524.61, 58.16) --
	(516.60, 58.79) --
	(508.59, 60.03) --
	(500.58, 60.94) --
	(492.58, 62.46) --
	(484.57, 65.05) --
	(476.56, 66.78) --
	(468.55, 70.56) --
	(460.54, 73.93) --
	(452.53, 74.95) --
	(444.53, 77.75) --
	(436.52, 84.76) --
	(428.51, 84.17) --
	(420.50, 90.54) --
	(412.49, 89.60) --
	(404.48, 93.27) --
	(396.48,129.71) --
	(388.47,137.14) --
	(380.46,136.34) --
	(372.45,134.94) --
	(364.44,135.04) --
	(356.43,120.43) --
	(348.43,124.23) --
	(340.42,119.49) --
	cycle;
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.42,255.71) --
	(348.43,234.51) --
	(356.43,227.85) --
	(364.44,224.34) --
	(372.45,232.95) --
	(380.46,225.18) --
	(388.47,220.08) --
	(396.48,212.60) --
	(404.48,137.83) --
	(412.49,119.60) --
	(420.50,116.21) --
	(428.51,104.34) --
	(436.52,101.52) --
	(444.53, 91.89) --
	(452.53, 87.88) --
	(460.54, 85.77) --
	(468.55, 79.62) --
	(476.56, 76.53) --
	(484.57, 72.74) --
	(492.58, 71.32) --
	(500.58, 69.30) --
	(508.59, 68.00) --
	(516.60, 66.17) --
	(524.61, 65.86) --
	(532.62, 65.06) --
	(540.63, 64.84) --
	(548.63, 64.84) --
	(556.64, 64.82) --
	(564.65, 64.82) --
	(572.66, 64.08);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (572.66, 56.77) --
	(564.65, 57.72) --
	(556.64, 57.72) --
	(548.63, 58.18) --
	(540.63, 58.18) --
	(532.62, 58.40) --
	(524.61, 58.16) --
	(516.60, 58.79) --
	(508.59, 60.03) --
	(500.58, 60.94) --
	(492.58, 62.46) --
	(484.57, 65.05) --
	(476.56, 66.78) --
	(468.55, 70.56) --
	(460.54, 73.93) --
	(452.53, 74.95) --
	(444.53, 77.75) --
	(436.52, 84.76) --
	(428.51, 84.17) --
	(420.50, 90.54) --
	(412.49, 89.60) --
	(404.48, 93.27) --
	(396.48,129.71) --
	(388.47,137.14) --
	(380.46,136.34) --
	(372.45,134.94) --
	(364.44,135.04) --
	(356.43,120.43) --
	(348.43,124.23) --
	(340.42,119.49);
\definecolor{fillColor}{RGB}{0,114,178}

\path[fill=fillColor,fill opacity=0.50] (340.42,167.99) --
	(348.43,158.37) --
	(356.43,155.61) --
	(364.44,153.41) --
	(372.45,158.49) --
	(380.46,154.03) --
	(388.47,151.93) --
	(396.48,148.30) --
	(404.48,107.71) --
	(412.49, 96.60) --
	(420.50, 94.03) --
	(428.51, 86.26) --
	(436.52, 84.52) --
	(444.53, 77.77) --
	(452.53, 75.45) --
	(460.54, 73.98) --
	(468.55, 69.74) --
	(476.56, 67.73) --
	(484.57, 65.35) --
	(492.58, 64.41) --
	(500.58, 62.76) --
	(508.59, 61.94) --
	(516.60, 60.56) --
	(524.61, 60.51) --
	(532.62, 60.01) --
	(540.63, 59.88) --
	(548.63, 59.88) --
	(556.64, 59.77) --
	(564.65, 59.77) --
	(572.66, 59.40) --
	(572.66, 54.49) --
	(564.65, 54.89) --
	(556.64, 54.89) --
	(548.63, 55.23) --
	(540.63, 55.23) --
	(532.62, 55.32) --
	(524.61, 55.43) --
	(516.60, 55.76) --
	(508.59, 56.48) --
	(500.58, 57.32) --
	(492.58, 58.34) --
	(484.57, 59.85) --
	(476.56, 61.35) --
	(468.55, 63.63) --
	(460.54, 66.52) --
	(452.53, 66.97) --
	(444.53, 68.74) --
	(436.52, 73.15) --
	(428.51, 73.04) --
	(420.50, 77.40) --
	(412.49, 77.19) --
	(404.48, 79.82) --
	(396.48,101.35) --
	(388.47,108.09) --
	(380.46,107.05) --
	(372.45,105.61) --
	(364.44,105.29) --
	(356.43, 96.62) --
	(348.43, 98.80) --
	(340.42, 95.78) --
	cycle;
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round,line cap=round] (340.42,167.99) --
	(348.43,158.37) --
	(356.43,155.61) --
	(364.44,153.41) --
	(372.45,158.49) --
	(380.46,154.03) --
	(388.47,151.93) --
	(396.48,148.30) --
	(404.48,107.71) --
	(412.49, 96.60) --
	(420.50, 94.03) --
	(428.51, 86.26) --
	(436.52, 84.52) --
	(444.53, 77.77) --
	(452.53, 75.45) --
	(460.54, 73.98) --
	(468.55, 69.74) --
	(476.56, 67.73) --
	(484.57, 65.35) --
	(492.58, 64.41) --
	(500.58, 62.76) --
	(508.59, 61.94) --
	(516.60, 60.56) --
	(524.61, 60.51) --
	(532.62, 60.01) --
	(540.63, 59.88) --
	(548.63, 59.88) --
	(556.64, 59.77) --
	(564.65, 59.77) --
	(572.66, 59.40);

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round,line cap=round] (572.66, 54.49) --
	(564.65, 54.89) --
	(556.64, 54.89) --
	(548.63, 55.23) --
	(540.63, 55.23) --
	(532.62, 55.32) --
	(524.61, 55.43) --
	(516.60, 55.76) --
	(508.59, 56.48) --
	(500.58, 57.32) --
	(492.58, 58.34) --
	(484.57, 59.85) --
	(476.56, 61.35) --
	(468.55, 63.63) --
	(460.54, 66.52) --
	(452.53, 66.97) --
	(444.53, 68.74) --
	(436.52, 73.15) --
	(428.51, 73.04) --
	(420.50, 77.40) --
	(412.49, 77.19) --
	(404.48, 79.82) --
	(396.48,101.35) --
	(388.47,108.09) --
	(380.46,107.05) --
	(372.45,105.61) --
	(364.44,105.29) --
	(356.43, 96.62) --
	(348.43, 98.80) --
	(340.42, 95.78);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (316.39, 41.40) -- (572.66, 41.40);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.42,170.94) --
	(348.43,170.79) --
	(356.43,163.30) --
	(364.44,167.34) --
	(372.45,172.68) --
	(380.46,172.43) --
	(388.47,174.86) --
	(396.48,167.73) --
	(404.48,112.65) --
	(412.49,105.04) --
	(420.50,102.84) --
	(428.51, 93.38) --
	(436.52, 91.84) --
	(444.53, 83.99) --
	(452.53, 80.95) --
	(460.54, 80.00) --
	(468.55, 75.18) --
	(476.56, 71.77) --
	(484.57, 68.56) --
	(492.58, 66.30) --
	(500.58, 65.11) --
	(508.59, 63.71) --
	(516.60, 62.37) --
	(524.61, 61.91) --
	(532.62, 61.69) --
	(540.63, 61.52) --
	(548.63, 61.52) --
	(556.64, 61.33) --
	(564.65, 61.33) --
	(572.66, 60.65);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (340.42,126.39) --
	(348.43,125.79) --
	(356.43,122.09) --
	(364.44,124.37) --
	(372.45,127.48) --
	(380.46,126.99) --
	(388.47,128.28) --
	(396.48,123.96) --
	(404.48, 92.40) --
	(412.49, 87.06) --
	(420.50, 85.30) --
	(428.51, 79.23) --
	(436.52, 77.99) --
	(444.53, 73.04) --
	(452.53, 70.90) --
	(460.54, 70.16) --
	(468.55, 66.88) --
	(476.56, 64.57) --
	(484.57, 62.37) --
	(492.58, 60.94) --
	(500.58, 60.13) --
	(508.59, 59.14) --
	(516.60, 58.16) --
	(524.61, 57.82) --
	(532.62, 57.64) --
	(540.63, 57.52) --
	(548.63, 57.52) --
	(556.64, 57.39) --
	(564.65, 57.39) --
	(572.66, 56.99);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (316.39, 30.69) rectangle (572.66,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44, 38.37) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44, 90.38) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,142.40) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,194.41) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (311.44,246.42) {5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64, 41.40) --
	(316.39, 41.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64, 93.41) --
	(316.39, 93.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,145.43) --
	(316.39,145.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,197.44) --
	(316.39,197.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (313.64,249.45) --
	(316.39,249.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (364.44, 27.94) --
	(364.44, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.50, 27.94) --
	(420.50, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.56, 27.94) --
	(476.56, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.62, 27.94) --
	(532.62, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (364.44, 19.68) {Jan 13};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (420.50, 19.68) {Jan 20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (476.56, 19.68) {Jan 27};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (532.62, 19.68) {Feb 03};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (444.53,  7.64) {Symptom onset date (infector)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (302.16,148.55) {Basic reproduction number $\mathcal R_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (432.89,208.36) rectangle (548.42,253.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (438.39,228.31) rectangle (452.84,242.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{240,228,66}
\definecolor{fillColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line cap=rect,fill=fillColor,fill opacity=0.50] (439.10,229.02) rectangle (452.13,242.05);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{240,228,66}

\path[draw=drawColor,line width= 0.6pt,line join=round] (439.83,235.54) -- (451.40,235.54);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (438.39,213.86) rectangle (452.84,228.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,114,178}
\definecolor{fillColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line cap=rect,fill=fillColor,fill opacity=0.50] (439.10,214.57) rectangle (452.13,227.60);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,114,178}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (439.83,221.09) -- (451.40,221.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (458.34,232.51) {6 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (458.34,218.05) {8 day doubling period};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (316.39,274.49) {D. Cohort-averaged $\mathcal R_0$ estimates};
\end{scope}
\end{tikzpicture}

\end{document}