### 2020-09-28 forward_corr.tex
update

% Created by tikzDevice version 0.12.3 on 2020-09-28 00:14:35
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{gray}{0.95}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,  0.00) -- (  0.00,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 30.43,  0.00) -- ( 30.43,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.86,  0.00) -- ( 60.86,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.29,  0.00) -- ( 91.29,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.72,  0.00) -- (121.72,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.15,  0.00) -- (152.15,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.58,  0.00) -- (182.58,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (213.01,  0.00) -- (213.01,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.44,  0.00) -- (243.44,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.87,  0.00) -- (273.87,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.29,  0.00) -- (304.29,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.72,  0.00) -- (334.72,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.15,  0.00) -- (365.15,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.58,  0.00) -- (395.58,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.01,  0.00) -- (426.01,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.44,  0.00) -- (456.44,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (486.87,  0.00) -- (486.87,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (517.30,  0.00) -- (517.30,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.73,  0.00) -- (547.73,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (578.16,  0.00) -- (578.16,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,  0.00) -- (578.16,  0.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 30.43) -- (578.16, 30.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 60.86) -- (578.16, 60.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 91.29) -- (578.16, 91.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,121.72) -- (578.16,121.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,152.15) -- (578.16,152.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,182.58) -- (578.16,182.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,213.01) -- (578.16,213.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,243.44) -- (578.16,243.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,273.87) -- (578.16,273.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,304.29) -- (578.16,304.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,334.72) -- (578.16,334.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,365.15) -- (578.16,365.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,395.58) -- (578.16,395.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,426.01) -- (578.16,426.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,456.44) -- (578.16,456.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,486.87) -- (578.16,486.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,517.30) -- (578.16,517.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,547.73) -- (578.16,547.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,578.16) -- (578.16,578.16);
\end{scope}
\begin{scope}
\path[clip] (  0.00,412.97) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,412.97) rectangle (578.16,578.16);
\end{scope}
\begin{scope}
\path[clip] ( 40.70,443.99) rectangle (572.66,555.28);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 40.70,443.99) rectangle (572.66,555.28);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.87) --
	( 49.28,444.83) --
	( 57.86,444.83) --
	( 66.44,445.43) --
	( 75.02,445.66) --
	( 83.60,446.22) --
	( 92.18,446.31) --
	(100.76,447.61) --
	(109.34,447.61) --
	(117.92,448.49) --
	(126.50,449.19) --
	(135.08,450.49) --
	(143.66,451.74) --
	(152.24,454.24) --
	(160.82,455.12) --
	(169.40,458.60) --
	(177.98,460.51) --
	(186.56,463.75) --
	(195.14,469.23) --
	(203.72,472.61) --
	(212.30,477.86) --
	(220.88,481.80) --
	(229.46,489.82) --
	(238.04,497.94) --
	(246.62,502.53) --
	(255.20,508.10) --
	(263.78,517.98) --
	(272.36,523.55) --
	(280.94,528.79) --
	(289.52,534.22) --
	(298.10,533.76) --
	(306.68,535.43) --
	(315.26,531.67) --
	(323.84,527.31) --
	(332.42,522.81) --
	(341.00,515.11) --
	(349.58,506.34) --
	(358.16,499.06) --
	(366.74,490.57) --
	(375.32,481.01) --
	(383.90,474.65) --
	(392.48,471.50) --
	(401.06,463.52) --
	(409.64,462.18) --
	(418.22,456.66) --
	(426.80,454.47) --
	(435.38,452.02) --
	(443.96,450.86) --
	(452.54,449.28) --
	(461.12,448.63) --
	(469.70,447.61) --
	(478.28,446.68) --
	(486.86,445.61) --
	(495.44,445.89) --
	(504.02,445.15) --
	(512.60,444.87) --
	(521.18,444.55) --
	(529.76,444.50) --
	(538.34,444.36) --
	(546.92,444.13) --
	(555.50,444.22) --
	(564.08,444.08) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.41) --
	( 49.28,444.73) --
	( 57.86,444.92) --
	( 66.44,444.92) --
	( 75.02,444.96) --
	( 83.60,445.71) --
	( 92.18,446.68) --
	(100.76,445.85) --
	(109.34,447.19) --
	(117.92,447.52) --
	(126.50,448.72) --
	(135.08,449.14) --
	(143.66,449.79) --
	(152.24,451.41) --
	(160.82,453.64) --
	(169.40,455.17) --
	(177.98,458.28) --
	(186.56,459.44) --
	(195.14,462.45) --
	(203.72,467.93) --
	(212.30,471.41) --
	(220.88,478.46) --
	(229.46,479.53) --
	(238.04,488.11) --
	(246.62,497.48) --
	(255.20,501.28) --
	(263.78,507.22) --
	(272.36,516.82) --
	(280.94,521.69) --
	(289.52,526.84) --
	(298.10,534.27) --
	(306.68,533.80) --
	(315.26,534.22) --
	(323.84,531.53) --
	(332.42,529.67) --
	(341.00,522.16) --
	(349.58,514.60) --
	(358.16,508.15) --
	(366.74,500.45) --
	(375.32,493.03) --
	(383.90,479.80) --
	(392.48,475.16) --
	(401.06,472.94) --
	(409.64,463.98) --
	(418.22,462.55) --
	(426.80,458.60) --
	(435.38,455.54) --
	(443.96,452.99) --
	(452.54,451.51) --
	(461.12,449.65) --
	(469.70,448.30) --
	(478.28,448.03) --
	(486.86,447.47) --
	(495.44,446.36) --
	(504.02,445.20) --
	(512.60,445.61) --
	(521.18,444.92) --
	(529.76,444.69) --
	(538.34,444.87) --
	(546.92,444.22) --
	(555.50,444.22) --
	(564.08,444.22) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.59) --
	( 49.28,444.59) --
	( 57.86,444.87) --
	( 66.44,444.87) --
	( 75.02,444.92) --
	( 83.60,445.48) --
	( 92.18,446.26) --
	(100.76,445.85) --
	(109.34,446.77) --
	(117.92,447.38) --
	(126.50,448.91) --
	(135.08,448.86) --
	(143.66,449.74) --
	(152.24,450.67) --
	(160.82,451.97) --
	(169.40,455.68) --
	(177.98,456.98) --
	(186.56,459.21) --
	(195.14,461.57) --
	(203.72,466.81) --
	(212.30,469.97) --
	(220.88,475.30) --
	(229.46,480.13) --
	(238.04,486.11) --
	(246.62,494.09) --
	(255.20,499.75) --
	(263.78,504.90) --
	(272.36,514.92) --
	(280.94,519.79) --
	(289.52,525.64) --
	(298.10,533.43) --
	(306.68,530.37) --
	(315.26,535.47) --
	(323.84,535.52) --
	(332.42,528.79) --
	(341.00,524.29) --
	(349.58,518.26) --
	(358.16,510.98) --
	(366.74,502.95) --
	(375.32,492.24) --
	(383.90,485.97) --
	(392.48,475.16) --
	(401.06,471.04) --
	(409.64,468.58) --
	(418.22,464.03) --
	(426.80,458.19) --
	(435.38,456.47) --
	(443.96,453.59) --
	(452.54,451.69) --
	(461.12,449.93) --
	(469.70,448.86) --
	(478.28,447.38) --
	(486.86,447.38) --
	(495.44,446.31) --
	(504.02,445.52) --
	(512.60,445.89) --
	(521.18,444.78) --
	(529.76,444.92) --
	(538.34,444.73) --
	(546.92,444.45) --
	(555.50,444.55) --
	(564.08,444.22) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.50) --
	( 49.28,445.29) --
	( 57.86,444.73) --
	( 66.44,445.52) --
	( 75.02,445.71) --
	( 83.60,445.94) --
	( 92.18,446.54) --
	(100.76,447.19) --
	(109.34,448.44) --
	(117.92,448.49) --
	(126.50,449.60) --
	(135.08,450.39) --
	(143.66,451.32) --
	(152.24,453.78) --
	(160.82,456.89) --
	(169.40,458.56) --
	(177.98,461.06) --
	(186.56,464.08) --
	(195.14,467.79) --
	(203.72,472.85) --
	(212.30,479.62) --
	(220.88,482.40) --
	(229.46,491.59) --
	(238.04,496.88) --
	(246.62,500.63) --
	(255.20,511.67) --
	(263.78,517.61) --
	(272.36,521.51) --
	(280.94,527.91) --
	(289.52,532.83) --
	(298.10,533.94) --
	(306.68,534.73) --
	(315.26,530.42) --
	(323.84,527.91) --
	(332.42,524.11) --
	(341.00,513.39) --
	(349.58,508.89) --
	(358.16,496.69) --
	(366.74,489.92) --
	(375.32,482.45) --
	(383.90,471.78) --
	(392.48,470.99) --
	(401.06,464.40) --
	(409.64,460.04) --
	(418.22,456.10) --
	(426.80,454.66) --
	(435.38,452.62) --
	(443.96,450.35) --
	(452.54,449.19) --
	(461.12,448.21) --
	(469.70,446.73) --
	(478.28,446.82) --
	(486.86,445.99) --
	(495.44,445.61) --
	(504.02,445.01) --
	(512.60,445.01) --
	(521.18,444.87) --
	(529.76,444.64) --
	(538.34,444.50) --
	(546.92,444.27) --
	(555.50,444.04) --
	(564.08,444.18) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.64) --
	( 49.28,444.64) --
	( 57.86,444.83) --
	( 66.44,444.96) --
	( 75.02,445.24) --
	( 83.60,445.52) --
	( 92.18,446.03) --
	(100.76,445.85) --
	(109.34,447.52) --
	(117.92,447.70) --
	(126.50,448.72) --
	(135.08,448.72) --
	(143.66,450.95) --
	(152.24,450.72) --
	(160.82,453.41) --
	(169.40,455.12) --
	(177.98,457.86) --
	(186.56,461.20) --
	(195.14,463.24) --
	(203.72,465.70) --
	(212.30,470.25) --
	(220.88,478.55) --
	(229.46,483.05) --
	(238.04,487.83) --
	(246.62,496.09) --
	(255.20,498.78) --
	(263.78,507.68) --
	(272.36,517.98) --
	(280.94,520.07) --
	(289.52,524.38) --
	(298.10,531.39) --
	(306.68,534.13) --
	(315.26,534.27) --
	(323.84,531.67) --
	(332.42,528.65) --
	(341.00,526.29) --
	(349.58,513.62) --
	(358.16,507.31) --
	(366.74,502.02) --
	(375.32,490.57) --
	(383.90,484.35) --
	(392.48,477.02) --
	(401.06,473.22) --
	(409.64,465.52) --
	(418.22,460.97) --
	(426.80,458.37) --
	(435.38,454.89) --
	(443.96,452.85) --
	(452.54,451.32) --
	(461.12,448.63) --
	(469.70,448.58) --
	(478.28,447.42) --
	(486.86,446.68) --
	(495.44,446.31) --
	(504.02,445.75) --
	(512.60,445.29) --
	(521.18,445.15) --
	(529.76,444.64) --
	(538.34,444.32) --
	(546.92,444.50) --
	(555.50,444.41) --
	(564.08,444.36) --
	(572.66,444.22);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.32) --
	( 49.28,444.73) --
	( 57.86,444.83) --
	( 66.44,445.29) --
	( 75.02,445.10) --
	( 83.60,445.43) --
	( 92.18,446.08) --
	(100.76,446.54) --
	(109.34,446.82) --
	(117.92,448.26) --
	(126.50,448.63) --
	(135.08,449.42) --
	(143.66,450.44) --
	(152.24,451.92) --
	(160.82,454.38) --
	(169.40,456.28) --
	(177.98,457.21) --
	(186.56,461.43) --
	(195.14,465.10) --
	(203.72,468.62) --
	(212.30,472.06) --
	(220.88,476.60) --
	(229.46,482.68) --
	(238.04,491.35) --
	(246.62,496.64) --
	(255.20,505.78) --
	(263.78,511.49) --
	(272.36,516.17) --
	(280.94,525.50) --
	(289.52,528.14) --
	(298.10,530.69) --
	(306.68,535.15) --
	(315.26,535.66) --
	(323.84,530.42) --
	(332.42,526.05) --
	(341.00,520.02) --
	(349.58,514.46) --
	(358.16,504.16) --
	(366.74,498.50) --
	(375.32,490.84) --
	(383.90,481.06) --
	(392.48,474.24) --
	(401.06,468.35) --
	(409.64,466.49) --
	(418.22,461.11) --
	(426.80,456.70) --
	(435.38,454.34) --
	(443.96,452.67) --
	(452.54,451.92) --
	(461.12,449.74) --
	(469.70,448.21) --
	(478.28,447.28) --
	(486.86,447.01) --
	(495.44,445.71) --
	(504.02,445.52) --
	(512.60,445.20) --
	(521.18,445.34) --
	(529.76,444.55) --
	(538.34,444.41) --
	(546.92,444.18) --
	(555.50,444.13) --
	(564.08,444.27) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.59) --
	( 49.28,444.83) --
	( 57.86,445.20) --
	( 66.44,444.96) --
	( 75.02,445.61) --
	( 83.60,445.75) --
	( 92.18,446.59) --
	(100.76,446.91) --
	(109.34,447.79) --
	(117.92,447.93) --
	(126.50,449.65) --
	(135.08,450.53) --
	(143.66,451.13) --
	(152.24,454.38) --
	(160.82,454.80) --
	(169.40,458.70) --
	(177.98,460.32) --
	(186.56,462.87) --
	(195.14,467.37) --
	(203.72,472.06) --
	(212.30,476.60) --
	(220.88,480.13) --
	(229.46,489.04) --
	(238.04,495.25) --
	(246.62,502.53) --
	(255.20,509.35) --
	(263.78,515.52) --
	(272.36,523.60) --
	(280.94,528.37) --
	(289.52,530.23) --
	(298.10,534.64) --
	(306.68,537.42) --
	(315.26,530.55) --
	(323.84,529.86) --
	(332.42,520.72) --
	(341.00,515.20) --
	(349.58,504.21) --
	(358.16,501.47) --
	(366.74,492.33) --
	(375.32,480.92) --
	(383.90,476.88) --
	(392.48,471.45) --
	(401.06,463.98) --
	(409.64,462.41) --
	(418.22,457.68) --
	(426.80,454.52) --
	(435.38,453.59) --
	(443.96,452.02) --
	(452.54,451.13) --
	(461.12,447.79) --
	(469.70,447.33) --
	(478.28,446.91) --
	(486.86,446.22) --
	(495.44,445.48) --
	(504.02,445.34) --
	(512.60,445.15) --
	(521.18,444.55) --
	(529.76,444.69) --
	(538.34,444.45) --
	(546.92,444.22) --
	(555.50,444.13) --
	(564.08,444.13) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.45) --
	( 49.28,444.64) --
	( 57.86,445.15) --
	( 66.44,445.15) --
	( 75.02,445.29) --
	( 83.60,446.03) --
	( 92.18,446.12) --
	(100.76,447.05) --
	(109.34,447.79) --
	(117.92,447.61) --
	(126.50,449.05) --
	(135.08,450.07) --
	(143.66,450.76) --
	(152.24,453.32) --
	(160.82,455.03) --
	(169.40,457.95) --
	(177.98,459.07) --
	(186.56,462.04) --
	(195.14,466.58) --
	(203.72,469.83) --
	(212.30,475.16) --
	(220.88,479.90) --
	(229.46,485.28) --
	(238.04,494.51) --
	(246.62,499.52) --
	(255.20,510.70) --
	(263.78,510.89) --
	(272.36,521.93) --
	(280.94,525.73) --
	(289.52,530.42) --
	(298.10,534.08) --
	(306.68,534.96) --
	(315.26,533.43) --
	(323.84,529.77) --
	(332.42,525.87) --
	(341.00,516.45) --
	(349.58,507.31) --
	(358.16,501.98) --
	(366.74,494.79) --
	(375.32,484.58) --
	(383.90,480.27) --
	(392.48,470.80) --
	(401.06,468.30) --
	(409.64,462.08) --
	(418.22,458.28) --
	(426.80,456.79) --
	(435.38,452.94) --
	(443.96,452.34) --
	(452.54,450.81) --
	(461.12,448.86) --
	(469.70,447.79) --
	(478.28,447.01) --
	(486.86,446.73) --
	(495.44,445.38) --
	(504.02,446.08) --
	(512.60,445.15) --
	(521.18,444.92) --
	(529.76,444.41) --
	(538.34,444.45) --
	(546.92,444.45) --
	(555.50,444.45) --
	(564.08,444.55) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.27) --
	( 49.28,444.73) --
	( 57.86,444.96) --
	( 66.44,445.34) --
	( 75.02,445.06) --
	( 83.60,445.85) --
	( 92.18,446.12) --
	(100.76,447.01) --
	(109.34,447.28) --
	(117.92,447.61) --
	(126.50,449.05) --
	(135.08,449.46) --
	(143.66,450.81) --
	(152.24,452.29) --
	(160.82,455.31) --
	(169.40,455.87) --
	(177.98,459.67) --
	(186.56,461.20) --
	(195.14,466.17) --
	(203.72,468.44) --
	(212.30,473.36) --
	(220.88,478.74) --
	(229.46,485.65) --
	(238.04,492.05) --
	(246.62,498.08) --
	(255.20,507.96) --
	(263.78,512.32) --
	(272.36,521.74) --
	(280.94,525.87) --
	(289.52,529.21) --
	(298.10,535.19) --
	(306.68,534.17) --
	(315.26,536.63) --
	(323.84,529.39) --
	(332.42,526.24) --
	(341.00,518.03) --
	(349.58,511.63) --
	(358.16,502.77) --
	(366.74,497.76) --
	(375.32,484.40) --
	(383.90,479.57) --
	(392.48,472.47) --
	(401.06,469.51) --
	(409.64,463.20) --
	(418.22,458.05) --
	(426.80,456.24) --
	(435.38,453.22) --
	(443.96,451.97) --
	(452.54,449.70) --
	(461.12,449.51) --
	(469.70,448.03) --
	(478.28,447.28) --
	(486.86,446.91) --
	(495.44,445.43) --
	(504.02,445.94) --
	(512.60,445.24) --
	(521.18,444.92) --
	(529.76,444.22) --
	(538.34,444.50) --
	(546.92,444.55) --
	(555.50,444.36) --
	(564.08,444.36) --
	(572.66,444.27);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.45) --
	( 49.28,445.01) --
	( 57.86,445.34) --
	( 66.44,444.83) --
	( 75.02,445.66) --
	( 83.60,446.63) --
	( 92.18,446.40) --
	(100.76,447.28) --
	(109.34,447.33) --
	(117.92,448.30) --
	(126.50,449.84) --
	(135.08,450.30) --
	(143.66,451.92) --
	(152.24,454.10) --
	(160.82,457.17) --
	(169.40,457.30) --
	(177.98,461.11) --
	(186.56,465.65) --
	(195.14,468.02) --
	(203.72,472.06) --
	(212.30,478.41) --
	(220.88,482.77) --
	(229.46,491.26) --
	(238.04,497.71) --
	(246.62,503.14) --
	(255.20,512.88) --
	(263.78,520.12) --
	(272.36,523.32) --
	(280.94,528.14) --
	(289.52,530.09) --
	(298.10,536.77) --
	(306.68,538.21) --
	(315.26,526.57) --
	(323.84,528.75) --
	(332.42,518.96) --
	(341.00,515.99) --
	(349.58,503.51) --
	(358.16,497.20) --
	(366.74,488.11) --
	(375.32,480.31) --
	(383.90,476.00) --
	(392.48,470.15) --
	(401.06,464.59) --
	(409.64,460.37) --
	(418.22,456.19) --
	(426.80,453.83) --
	(435.38,453.96) --
	(443.96,450.72) --
	(452.54,449.93) --
	(461.12,448.44) --
	(469.70,447.56) --
	(478.28,446.59) --
	(486.86,445.71) --
	(495.44,445.61) --
	(504.02,444.96) --
	(512.60,444.83) --
	(521.18,444.78) --
	(529.76,444.55) --
	(538.34,444.36) --
	(546.92,444.36) --
	(555.50,444.13) --
	(564.08,444.32);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.18) --
	( 49.28,444.96) --
	( 57.86,445.01) --
	( 66.44,444.96) --
	( 75.02,445.80) --
	( 83.60,445.57) --
	( 92.18,445.80) --
	(100.76,446.59) --
	(109.34,447.38) --
	(117.92,448.35) --
	(126.50,448.63) --
	(135.08,448.82) --
	(143.66,451.55) --
	(152.24,451.88) --
	(160.82,453.64) --
	(169.40,455.54) --
	(177.98,459.11) --
	(186.56,460.92) --
	(195.14,465.42) --
	(203.72,470.99) --
	(212.30,473.54) --
	(220.88,477.44) --
	(229.46,483.65) --
	(238.04,494.00) --
	(246.62,500.17) --
	(255.20,503.23) --
	(263.78,512.09) --
	(272.36,523.46) --
	(280.94,526.15) --
	(289.52,535.01) --
	(298.10,532.64) --
	(306.68,532.73) --
	(315.26,540.06) --
	(323.84,533.85) --
	(332.42,522.81) --
	(341.00,518.35) --
	(349.58,512.00) --
	(358.16,502.16) --
	(366.74,495.99) --
	(375.32,486.62) --
	(383.90,477.76) --
	(392.48,472.98) --
	(401.06,470.11) --
	(409.64,463.15) --
	(418.22,457.17) --
	(426.80,457.54) --
	(435.38,454.01) --
	(443.96,451.13) --
	(452.54,449.88) --
	(461.12,448.49) --
	(469.70,447.42) --
	(478.28,446.96) --
	(486.86,446.59) --
	(495.44,445.61) --
	(504.02,445.52) --
	(512.60,444.87) --
	(521.18,444.55) --
	(529.76,444.64) --
	(538.34,444.59) --
	(546.92,444.41) --
	(555.50,444.50) --
	(564.08,444.22) --
	(572.66,444.32);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.36) --
	( 49.28,444.69) --
	( 57.86,444.64) --
	( 66.44,445.24) --
	( 75.02,445.43) --
	( 83.60,445.66) --
	( 92.18,445.24) --
	(100.76,446.17) --
	(109.34,446.73) --
	(117.92,447.84) --
	(126.50,448.72) --
	(135.08,448.26) --
	(143.66,449.56) --
	(152.24,451.78) --
	(160.82,452.29) --
	(169.40,453.73) --
	(177.98,457.58) --
	(186.56,458.79) --
	(195.14,463.10) --
	(203.72,466.30) --
	(212.30,471.27) --
	(220.88,473.82) --
	(229.46,479.48) --
	(238.04,487.09) --
	(246.62,495.90) --
	(255.20,500.35) --
	(263.78,505.78) --
	(272.36,514.18) --
	(280.94,523.78) --
	(289.52,526.43) --
	(298.10,537.23) --
	(306.68,532.55) --
	(315.26,533.20) --
	(323.84,538.63) --
	(332.42,529.67) --
	(341.00,520.63) --
	(349.58,521.28) --
	(358.16,506.57) --
	(366.74,500.68) --
	(375.32,494.14) --
	(383.90,483.89) --
	(392.48,475.68) --
	(401.06,470.99) --
	(409.64,467.00) --
	(418.22,460.37) --
	(426.80,459.02) --
	(435.38,455.68) --
	(443.96,453.59) --
	(452.54,451.37) --
	(461.12,449.51) --
	(469.70,448.54) --
	(478.28,447.93) --
	(486.86,447.28) --
	(495.44,445.85) --
	(504.02,445.89) --
	(512.60,445.66) --
	(521.18,444.73) --
	(529.76,444.96) --
	(538.34,444.73) --
	(546.92,444.36) --
	(555.50,444.08) --
	(564.08,444.18) --
	(572.66,444.27);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.45) --
	( 49.28,444.45) --
	( 57.86,444.96) --
	( 66.44,445.38) --
	( 75.02,445.38) --
	( 83.60,445.66) --
	( 92.18,445.52) --
	(100.76,446.08) --
	(109.34,447.15) --
	(117.92,448.03) --
	(126.50,448.91) --
	(135.08,448.72) --
	(143.66,450.39) --
	(152.24,451.78) --
	(160.82,453.32) --
	(169.40,455.12) --
	(177.98,457.77) --
	(186.56,459.86) --
	(195.14,463.80) --
	(203.72,468.53) --
	(212.30,472.80) --
	(220.88,475.58) --
	(229.46,481.24) --
	(238.04,487.88) --
	(246.62,500.08) --
	(255.20,500.82) --
	(263.78,508.61) --
	(272.36,516.31) --
	(280.94,525.64) --
	(289.52,527.63) --
	(298.10,534.40) --
	(306.68,534.31) --
	(315.26,534.87) --
	(323.84,538.77) --
	(332.42,527.54) --
	(341.00,518.91) --
	(349.58,515.29) --
	(358.16,509.03) --
	(366.74,501.10) --
	(375.32,488.01) --
	(383.90,480.59) --
	(392.48,475.95) --
	(401.06,470.99) --
	(409.64,465.89) --
	(418.22,460.55) --
	(426.80,458.14) --
	(435.38,454.99) --
	(443.96,453.50) --
	(452.54,451.69) --
	(461.12,449.93) --
	(469.70,447.70) --
	(478.28,447.52) --
	(486.86,446.63) --
	(495.44,446.08) --
	(504.02,445.34) --
	(512.60,445.24) --
	(521.18,445.10) --
	(529.76,445.01) --
	(538.34,444.87) --
	(546.92,444.27) --
	(555.50,444.36) --
	(564.08,444.22) --
	(572.66,444.27);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.64) --
	( 49.28,444.36) --
	( 57.86,445.06) --
	( 66.44,445.57) --
	( 75.02,445.43) --
	( 83.60,445.48) --
	( 92.18,446.22) --
	(100.76,445.71) --
	(109.34,447.93) --
	(117.92,448.03) --
	(126.50,448.58) --
	(135.08,449.88) --
	(143.66,450.44) --
	(152.24,452.16) --
	(160.82,453.55) --
	(169.40,456.79) --
	(177.98,458.97) --
	(186.56,461.29) --
	(195.14,464.63) --
	(203.72,468.62) --
	(212.30,473.77) --
	(220.88,477.44) --
	(229.46,482.08) --
	(238.04,493.30) --
	(246.62,499.71) --
	(255.20,505.23) --
	(263.78,512.97) --
	(272.36,518.77) --
	(280.94,526.70) --
	(289.52,531.76) --
	(298.10,532.69) --
	(306.68,536.12) --
	(315.26,534.59) --
	(323.84,536.12) --
	(332.42,524.38) --
	(341.00,520.77) --
	(349.58,510.98) --
	(358.16,503.00) --
	(366.74,497.71) --
	(375.32,486.67) --
	(383.90,479.06) --
	(392.48,473.31) --
	(401.06,468.72) --
	(409.64,464.17) --
	(418.22,459.44) --
	(426.80,456.38) --
	(435.38,453.92) --
	(443.96,452.11) --
	(452.54,450.21) --
	(461.12,449.74) --
	(469.70,447.70) --
	(478.28,447.24) --
	(486.86,446.17) --
	(495.44,445.89) --
	(504.02,445.24) --
	(512.60,445.34) --
	(521.18,444.73) --
	(529.76,444.92) --
	(538.34,444.87) --
	(546.92,444.41) --
	(555.50,444.13) --
	(564.08,444.18) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.22) --
	( 49.28,444.83) --
	( 57.86,445.34) --
	( 66.44,445.34) --
	( 75.02,445.24) --
	( 83.60,445.89) --
	( 92.18,445.85) --
	(100.76,446.87) --
	(109.34,447.24) --
	(117.92,448.49) --
	(126.50,449.51) --
	(135.08,450.44) --
	(143.66,450.30) --
	(152.24,452.57) --
	(160.82,454.43) --
	(169.40,458.09) --
	(177.98,459.58) --
	(186.56,461.20) --
	(195.14,467.97) --
	(203.72,468.11) --
	(212.30,476.09) --
	(220.88,479.48) --
	(229.46,484.12) --
	(238.04,495.67) --
	(246.62,500.68) --
	(255.20,507.36) --
	(263.78,513.34) --
	(272.36,520.26) --
	(280.94,529.81) --
	(289.52,533.25) --
	(298.10,532.36) --
	(306.68,537.10) --
	(315.26,533.29) --
	(323.84,531.99) --
	(332.42,524.66) --
	(341.00,517.75) --
	(349.58,508.01) --
	(358.16,500.35) --
	(366.74,493.67) --
	(375.32,485.37) --
	(383.90,478.41) --
	(392.48,470.53) --
	(401.06,466.86) --
	(409.64,461.71) --
	(418.22,458.56) --
	(426.80,455.40) --
	(435.38,452.94) --
	(443.96,451.23) --
	(452.54,449.46) --
	(461.12,449.05) --
	(469.70,447.47) --
	(478.28,446.26) --
	(486.86,446.40) --
	(495.44,445.29) --
	(504.02,445.48) --
	(512.60,444.73) --
	(521.18,444.55) --
	(529.76,444.55) --
	(538.34,444.45) --
	(546.92,444.22) --
	(555.50,444.18) --
	(564.08,444.18) --
	(572.66,444.18);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.36) --
	( 49.28,444.50) --
	( 57.86,444.45) --
	( 66.44,444.83) --
	( 75.02,445.34) --
	( 83.60,445.29) --
	( 92.18,445.66) --
	(100.76,445.75) --
	(109.34,446.36) --
	(117.92,447.05) --
	(126.50,448.72) --
	(135.08,448.58) --
	(143.66,449.46) --
	(152.24,450.53) --
	(160.82,453.50) --
	(169.40,454.24) --
	(177.98,456.24) --
	(186.56,459.58) --
	(195.14,461.85) --
	(203.72,463.80) --
	(212.30,469.55) --
	(220.88,474.24) --
	(229.46,479.25) --
	(238.04,482.96) --
	(246.62,491.87) --
	(255.20,496.74) --
	(263.78,504.02) --
	(272.36,511.26) --
	(280.94,519.00) --
	(289.52,527.63) --
	(298.10,526.75) --
	(306.68,532.04) --
	(315.26,536.03) --
	(323.84,530.83) --
	(332.42,531.34) --
	(341.00,525.73) --
	(349.58,518.68) --
	(358.16,510.98) --
	(366.74,503.18) --
	(375.32,495.85) --
	(383.90,487.18) --
	(392.48,479.57) --
	(401.06,476.14) --
	(409.64,470.48) --
	(418.22,463.61) --
	(426.80,459.35) --
	(435.38,457.44) --
	(443.96,455.50) --
	(452.54,451.69) --
	(461.12,451.00) --
	(469.70,449.19) --
	(478.28,447.47) --
	(486.86,447.10) --
	(495.44,445.99) --
	(504.02,446.03) --
	(512.60,445.43) --
	(521.18,445.10) --
	(529.76,444.64) --
	(538.34,444.64) --
	(546.92,444.36) --
	(555.50,444.32) --
	(564.08,444.22) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.41) --
	( 49.28,444.78) --
	( 57.86,445.38) --
	( 66.44,445.43) --
	( 75.02,445.48) --
	( 83.60,445.85) --
	( 92.18,446.31) --
	(100.76,447.70) --
	(109.34,448.54) --
	(117.92,448.17) --
	(126.50,449.88) --
	(135.08,451.32) --
	(143.66,452.76) --
	(152.24,455.40) --
	(160.82,456.33) --
	(169.40,459.02) --
	(177.98,462.87) --
	(186.56,464.08) --
	(195.14,469.04) --
	(203.72,476.74) --
	(212.30,480.69) --
	(220.88,482.73) --
	(229.46,491.96) --
	(238.04,497.90) --
	(246.62,505.13) --
	(255.20,512.42) --
	(263.78,520.02) --
	(272.36,528.33) --
	(280.94,527.12) --
	(289.52,532.92) --
	(298.10,535.52) --
	(306.68,533.06) --
	(315.26,533.66) --
	(323.84,524.20) --
	(332.42,516.08) --
	(341.00,512.51) --
	(349.58,500.59) --
	(358.16,496.88) --
	(366.74,487.18) --
	(375.32,478.60) --
	(383.90,472.80) --
	(392.48,467.51) --
	(401.06,462.87) --
	(409.64,459.95) --
	(418.22,457.54) --
	(426.80,453.36) --
	(435.38,451.60) --
	(443.96,450.39) --
	(452.54,448.40) --
	(461.12,447.70) --
	(469.70,446.59) --
	(478.28,445.80) --
	(486.86,445.80) --
	(495.44,445.38) --
	(504.02,445.38) --
	(512.60,444.59) --
	(521.18,444.45) --
	(529.76,444.32) --
	(538.34,444.41) --
	(546.92,444.27) --
	(555.50,444.08) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.55) --
	( 49.28,445.01) --
	( 57.86,445.20) --
	( 66.44,445.48) --
	( 75.02,445.38) --
	( 83.60,445.94) --
	( 92.18,446.08) --
	(100.76,447.56) --
	(109.34,448.49) --
	(117.92,448.54) --
	(126.50,449.88) --
	(135.08,450.90) --
	(143.66,453.92) --
	(152.24,455.08) --
	(160.82,456.38) --
	(169.40,459.53) --
	(177.98,462.59) --
	(186.56,464.82) --
	(195.14,468.76) --
	(203.72,476.93) --
	(212.30,481.89) --
	(220.88,483.79) --
	(229.46,490.33) --
	(238.04,500.22) --
	(246.62,504.30) --
	(255.20,514.23) --
	(263.78,520.07) --
	(272.36,525.27) --
	(280.94,527.08) --
	(289.52,532.97) --
	(298.10,531.48) --
	(306.68,535.89) --
	(315.26,529.12) --
	(323.84,523.09) --
	(332.42,517.06) --
	(341.00,509.86) --
	(349.58,500.73) --
	(358.16,495.44) --
	(366.74,489.31) --
	(375.32,477.44) --
	(383.90,472.61) --
	(392.48,467.79) --
	(401.06,462.08) --
	(409.64,459.81) --
	(418.22,456.84) --
	(426.80,453.83) --
	(435.38,452.11) --
	(443.96,449.98) --
	(452.54,449.23) --
	(461.12,448.07) --
	(469.70,447.28) --
	(478.28,446.12) --
	(486.86,445.57) --
	(495.44,445.71) --
	(504.02,444.96) --
	(512.60,444.87) --
	(521.18,444.45) --
	(529.76,444.50) --
	(538.34,444.36) --
	(546.92,444.08) --
	(555.50,444.08) --
	(564.08,444.08);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.78) --
	( 49.28,444.73) --
	( 57.86,445.29) --
	( 66.44,445.34) --
	( 75.02,445.24) --
	( 83.60,446.12) --
	( 92.18,446.03) --
	(100.76,447.56) --
	(109.34,448.17) --
	(117.92,448.03) --
	(126.50,450.25) --
	(135.08,451.32) --
	(143.66,453.13) --
	(152.24,455.22) --
	(160.82,456.42) --
	(169.40,457.95) --
	(177.98,464.40) --
	(186.56,463.43) --
	(195.14,468.76) --
	(203.72,476.05) --
	(212.30,483.47) --
	(220.88,482.96) --
	(229.46,488.94) --
	(238.04,499.66) --
	(246.62,504.76) --
	(255.20,512.69) --
	(263.78,522.11) --
	(272.36,523.64) --
	(280.94,528.98) --
	(289.52,533.29) --
	(298.10,533.25) --
	(306.68,535.66) --
	(315.26,531.58) --
	(323.84,523.69) --
	(332.42,517.29) --
	(341.00,510.24) --
	(349.58,503.83) --
	(358.16,496.83) --
	(366.74,490.01) --
	(375.32,478.50) --
	(383.90,474.84) --
	(392.48,468.86) --
	(401.06,462.69) --
	(409.64,460.00) --
	(418.22,456.05) --
	(426.80,454.94) --
	(435.38,451.37) --
	(443.96,450.02) --
	(452.54,449.00) --
	(461.12,447.38) --
	(469.70,446.82) --
	(478.28,446.03) --
	(486.86,445.99) --
	(495.44,445.06) --
	(504.02,445.20) --
	(512.60,445.01) --
	(521.18,444.27) --
	(529.76,444.27) --
	(538.34,444.32) --
	(546.92,444.18) --
	(555.50,444.04) --
	(564.08,444.18) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.64) --
	( 49.28,444.83) --
	( 57.86,445.43) --
	( 66.44,444.96) --
	( 75.02,445.20) --
	( 83.60,446.31) --
	( 92.18,446.08) --
	(100.76,447.56) --
	(109.34,448.12) --
	(117.92,448.30) --
	(126.50,449.84) --
	(135.08,451.41) --
	(143.66,452.67) --
	(152.24,455.36) --
	(160.82,456.24) --
	(169.40,459.62) --
	(177.98,462.92) --
	(186.56,463.47) --
	(195.14,469.13) --
	(203.72,477.58) --
	(212.30,482.26) --
	(220.88,484.16) --
	(229.46,489.08) --
	(238.04,502.44) --
	(246.62,502.07) --
	(255.20,514.27) --
	(263.78,521.32) --
	(272.36,523.74) --
	(280.94,530.37) --
	(289.52,536.45) --
	(298.10,531.16) --
	(306.68,537.14) --
	(315.26,532.73) --
	(323.84,520.86) --
	(332.42,519.98) --
	(341.00,511.30) --
	(349.58,503.51) --
	(358.16,494.97) --
	(366.74,485.74) --
	(375.32,480.22) --
	(383.90,472.24) --
	(392.48,467.09) --
	(401.06,462.18) --
	(409.64,460.09) --
	(418.22,456.33) --
	(426.80,452.85) --
	(435.38,452.20) --
	(443.96,450.72) --
	(452.54,449.05) --
	(461.12,446.96) --
	(469.70,446.82) --
	(478.28,446.40) --
	(486.86,445.80) --
	(495.44,445.66) --
	(504.02,445.10) --
	(512.60,444.78) --
	(521.18,444.69) --
	(529.76,444.50) --
	(538.34,444.41) --
	(546.92,444.22) --
	(555.50,444.27);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.87) --
	( 49.28,445.01) --
	( 57.86,445.10) --
	( 66.44,445.38) --
	( 75.02,445.66) --
	( 83.60,446.45) --
	( 92.18,446.17) --
	(100.76,447.28) --
	(109.34,448.03) --
	(117.92,448.30) --
	(126.50,450.25) --
	(135.08,450.76) --
	(143.66,452.34) --
	(152.24,453.04) --
	(160.82,455.45) --
	(169.40,458.28) --
	(177.98,462.13) --
	(186.56,465.14) --
	(195.14,469.04) --
	(203.72,471.18) --
	(212.30,478.69) --
	(220.88,484.35) --
	(229.46,489.36) --
	(238.04,496.27) --
	(246.62,502.95) --
	(255.20,510.79) --
	(263.78,514.60) --
	(272.36,524.43) --
	(280.94,525.08) --
	(289.52,530.51) --
	(298.10,535.56) --
	(306.68,534.87) --
	(315.26,531.02) --
	(323.84,526.01) --
	(332.42,522.81) --
	(341.00,513.07) --
	(349.58,505.32) --
	(358.16,499.19) --
	(366.74,489.45) --
	(375.32,481.84) --
	(383.90,478.60) --
	(392.48,469.32) --
	(401.06,465.47) --
	(409.64,460.60) --
	(418.22,457.63) --
	(426.80,455.17) --
	(435.38,453.41) --
	(443.96,451.09) --
	(452.54,449.74) --
	(461.12,448.26) --
	(469.70,446.87) --
	(478.28,445.75) --
	(486.86,446.22) --
	(495.44,445.75) --
	(504.02,445.24) --
	(512.60,444.92) --
	(521.18,444.83) --
	(529.76,444.64) --
	(538.34,444.41) --
	(546.92,444.64) --
	(555.50,444.13) --
	(564.08,444.04) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.59) --
	( 49.28,445.34) --
	( 57.86,444.96) --
	( 66.44,445.20) --
	( 75.02,445.61) --
	( 83.60,445.89) --
	( 92.18,446.40) --
	(100.76,447.47) --
	(109.34,447.61) --
	(117.92,448.21) --
	(126.50,449.33) --
	(135.08,450.25) --
	(143.66,452.80) --
	(152.24,452.43) --
	(160.82,453.87) --
	(169.40,458.00) --
	(177.98,460.88) --
	(186.56,465.33) --
	(195.14,466.72) --
	(203.72,471.45) --
	(212.30,475.86) --
	(220.88,483.52) --
	(229.46,489.36) --
	(238.04,493.12) --
	(246.62,501.14) --
	(255.20,509.54) --
	(263.78,513.16) --
	(272.36,522.30) --
	(280.94,527.63) --
	(289.52,529.81) --
	(298.10,532.60) --
	(306.68,535.98) --
	(315.26,530.23) --
	(323.84,528.47) --
	(332.42,522.90) --
	(341.00,517.33) --
	(349.58,509.17) --
	(358.16,498.82) --
	(366.74,490.94) --
	(375.32,483.38) --
	(383.90,479.90) --
	(392.48,470.34) --
	(401.06,466.17) --
	(409.64,462.27) --
	(418.22,457.81) --
	(426.80,456.33) --
	(435.38,452.25) --
	(443.96,451.51) --
	(452.54,450.21) --
	(461.12,448.44) --
	(469.70,447.38) --
	(478.28,446.77) --
	(486.86,445.89) --
	(495.44,445.71) --
	(504.02,445.57) --
	(512.60,444.73) --
	(521.18,444.83) --
	(529.76,444.45) --
	(538.34,444.59) --
	(546.92,444.18) --
	(555.50,444.18) --
	(564.08,444.18) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.87) --
	( 49.28,445.10) --
	( 57.86,445.20) --
	( 66.44,445.06) --
	( 75.02,445.57) --
	( 83.60,445.57) --
	( 92.18,446.50) --
	(100.76,446.87) --
	(109.34,448.03) --
	(117.92,447.79) --
	(126.50,448.77) --
	(135.08,449.98) --
	(143.66,451.60) --
	(152.24,452.99) --
	(160.82,454.10) --
	(169.40,455.91) --
	(177.98,460.51) --
	(186.56,464.63) --
	(195.14,465.84) --
	(203.72,470.11) --
	(212.30,473.73) --
	(220.88,482.08) --
	(229.46,487.69) --
	(238.04,491.26) --
	(246.62,500.03) --
	(255.20,507.87) --
	(263.78,511.67) --
	(272.36,518.59) --
	(280.94,527.17) --
	(289.52,525.92) --
	(298.10,535.98) --
	(306.68,533.29) --
	(315.26,536.07) --
	(323.84,527.49) --
	(332.42,523.55) --
	(341.00,521.14) --
	(349.58,511.67) --
	(358.16,500.73) --
	(366.74,492.42) --
	(375.32,486.11) --
	(383.90,480.45) --
	(392.48,473.59) --
	(401.06,467.93) --
	(409.64,463.52) --
	(418.22,460.09) --
	(426.80,456.28) --
	(435.38,453.50) --
	(443.96,451.83) --
	(452.54,450.53) --
	(461.12,449.65) --
	(469.70,447.89) --
	(478.28,447.19) --
	(486.86,446.36) --
	(495.44,445.38) --
	(504.02,445.38) --
	(512.60,445.34) --
	(521.18,444.78) --
	(529.76,444.73) --
	(538.34,444.41) --
	(546.92,444.45) --
	(555.50,444.04) --
	(564.08,444.22) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.64) --
	( 49.28,445.24) --
	( 57.86,444.92) --
	( 66.44,445.24) --
	( 75.02,445.48) --
	( 83.60,445.48) --
	( 92.18,446.08) --
	(100.76,446.87) --
	(109.34,447.84) --
	(117.92,447.61) --
	(126.50,449.14) --
	(135.08,449.23) --
	(143.66,451.09) --
	(152.24,452.85) --
	(160.82,453.83) --
	(169.40,454.99) --
	(177.98,459.21) --
	(186.56,463.15) --
	(195.14,466.26) --
	(203.72,469.27) --
	(212.30,473.68) --
	(220.88,479.99) --
	(229.46,484.63) --
	(238.04,492.00) --
	(246.62,499.57) --
	(255.20,505.09) --
	(263.78,510.70) --
	(272.36,517.38) --
	(280.94,524.66) --
	(289.52,528.88) --
	(298.10,531.20) --
	(306.68,533.57) --
	(315.26,536.26) --
	(323.84,528.61) --
	(332.42,524.94) --
	(341.00,521.93) --
	(349.58,513.67) --
	(358.16,504.34) --
	(366.74,493.72) --
	(375.32,485.60) --
	(383.90,481.47) --
	(392.48,475.49) --
	(401.06,467.93) --
	(409.64,464.45) --
	(418.22,461.29) --
	(426.80,457.91) --
	(435.38,453.92) --
	(443.96,453.55) --
	(452.54,450.76) --
	(461.12,449.42) --
	(469.70,448.30) --
	(478.28,447.05) --
	(486.86,446.73) --
	(495.44,445.99) --
	(504.02,445.48) --
	(512.60,445.01) --
	(521.18,445.01) --
	(529.76,444.69) --
	(538.34,444.64) --
	(546.92,444.32) --
	(555.50,444.36) --
	(564.08,444.22) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.73) --
	( 49.28,445.20) --
	( 57.86,444.96) --
	( 66.44,445.52) --
	( 75.02,445.43) --
	( 83.60,445.94) --
	( 92.18,446.54) --
	(100.76,447.89) --
	(109.34,447.70) --
	(117.92,448.54) --
	(126.50,449.46) --
	(135.08,450.11) --
	(143.66,452.06) --
	(152.24,453.04) --
	(160.82,455.22) --
	(169.40,458.14) --
	(177.98,461.20) --
	(186.56,464.96) --
	(195.14,468.16) --
	(203.72,472.75) --
	(212.30,478.46) --
	(220.88,483.14) --
	(229.46,489.87) --
	(238.04,495.99) --
	(246.62,504.16) --
	(255.20,507.73) --
	(263.78,519.42) --
	(272.36,523.32) --
	(280.94,524.71) --
	(289.52,532.64) --
	(298.10,534.54) --
	(306.68,534.82) --
	(315.26,531.06) --
	(323.84,527.63) --
	(332.42,521.69) --
	(341.00,515.52) --
	(349.58,506.80) --
	(358.16,498.04) --
	(366.74,488.94) --
	(375.32,481.01) --
	(383.90,478.27) --
	(392.48,469.18) --
	(401.06,463.80) --
	(409.64,460.55) --
	(418.22,458.74) --
	(426.80,456.14) --
	(435.38,452.71) --
	(443.96,450.90) --
	(452.54,449.19) --
	(461.12,449.19) --
	(469.70,448.49) --
	(478.28,446.96) --
	(486.86,446.26) --
	(495.44,446.22) --
	(504.02,445.24) --
	(512.60,445.24) --
	(521.18,445.01) --
	(529.76,444.69) --
	(538.34,444.13) --
	(546.92,444.55) --
	(555.50,444.08) --
	(564.08,444.08) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.45) --
	( 49.28,444.73) --
	( 57.86,445.06) --
	( 66.44,444.92) --
	( 75.02,445.52) --
	( 83.60,445.15) --
	( 92.18,446.22) --
	(100.76,446.45) --
	(109.34,446.26) --
	(117.92,447.70) --
	(126.50,449.33) --
	(135.08,449.93) --
	(143.66,450.67) --
	(152.24,451.04) --
	(160.82,452.76) --
	(169.40,456.38) --
	(177.98,457.12) --
	(186.56,459.48) --
	(195.14,462.83) --
	(203.72,466.81) --
	(212.30,471.78) --
	(220.88,474.42) --
	(229.46,480.92) --
	(238.04,487.74) --
	(246.62,493.81) --
	(255.20,498.41) --
	(263.78,504.67) --
	(272.36,515.20) --
	(280.94,519.79) --
	(289.52,525.64) --
	(298.10,529.58) --
	(306.68,529.44) --
	(315.26,537.65) --
	(323.84,533.85) --
	(332.42,526.98) --
	(341.00,524.89) --
	(349.58,515.43) --
	(358.16,510.47) --
	(366.74,501.14) --
	(375.32,494.18) --
	(383.90,486.30) --
	(392.48,477.30) --
	(401.06,473.59) --
	(409.64,466.35) --
	(418.22,464.54) --
	(426.80,458.42) --
	(435.38,456.05) --
	(443.96,453.08) --
	(452.54,451.51) --
	(461.12,449.28) --
	(469.70,449.14) --
	(478.28,447.89) --
	(486.86,447.28) --
	(495.44,446.12) --
	(504.02,445.61) --
	(512.60,445.24) --
	(521.18,444.87) --
	(529.76,444.83) --
	(538.34,444.41) --
	(546.92,444.45) --
	(555.50,444.45) --
	(564.08,444.08) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.55) --
	( 49.28,444.69) --
	( 57.86,445.10) --
	( 66.44,444.96) --
	( 75.02,445.24) --
	( 83.60,445.52) --
	( 92.18,446.17) --
	(100.76,446.54) --
	(109.34,446.77) --
	(117.92,447.89) --
	(126.50,448.72) --
	(135.08,450.58) --
	(143.66,450.95) --
	(152.24,450.53) --
	(160.82,454.38) --
	(169.40,456.79) --
	(177.98,456.75) --
	(186.56,460.55) --
	(195.14,463.61) --
	(203.72,468.39) --
	(212.30,473.36) --
	(220.88,475.30) --
	(229.46,483.61) --
	(238.04,488.06) --
	(246.62,496.64) --
	(255.20,499.89) --
	(263.78,507.13) --
	(272.36,518.73) --
	(280.94,520.44) --
	(289.52,526.61) --
	(298.10,529.07) --
	(306.68,532.41) --
	(315.26,533.85) --
	(323.84,531.44) --
	(332.42,525.68) --
	(341.00,521.88) --
	(349.58,512.74) --
	(358.16,507.55) --
	(366.74,498.55) --
	(375.32,491.45) --
	(383.90,485.70) --
	(392.48,475.16) --
	(401.06,471.27) --
	(409.64,467.14) --
	(418.22,463.52) --
	(426.80,458.23) --
	(435.38,454.61) --
	(443.96,453.92) --
	(452.54,451.83) --
	(461.12,450.25) --
	(469.70,449.00) --
	(478.28,448.40) --
	(486.86,447.05) --
	(495.44,446.22) --
	(504.02,445.57) --
	(512.60,445.57) --
	(521.18,445.34) --
	(529.76,445.43) --
	(538.34,444.64) --
	(546.92,444.55) --
	(555.50,444.41) --
	(564.08,444.50) --
	(572.66,444.22);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.69) --
	( 49.28,445.06) --
	( 57.86,444.92) --
	( 66.44,445.61) --
	( 75.02,445.48) --
	( 83.60,446.22) --
	( 92.18,446.45) --
	(100.76,447.01) --
	(109.34,448.17) --
	(117.92,448.54) --
	(126.50,450.86) --
	(135.08,451.00) --
	(143.66,450.90) --
	(152.24,455.22) --
	(160.82,456.56) --
	(169.40,457.44) --
	(177.98,461.06) --
	(186.56,464.36) --
	(195.14,470.02) --
	(203.72,473.40) --
	(212.30,478.32) --
	(220.88,482.59) --
	(229.46,489.13) --
	(238.04,496.92) --
	(246.62,501.00) --
	(255.20,508.47) --
	(263.78,518.03) --
	(272.36,522.07) --
	(280.94,527.31) --
	(289.52,530.37) --
	(298.10,531.39) --
	(306.68,535.24) --
	(315.26,532.13) --
	(323.84,526.47) --
	(332.42,519.42) --
	(341.00,512.74) --
	(349.58,507.03) --
	(358.16,497.94) --
	(366.74,490.24) --
	(375.32,482.68) --
	(383.90,475.54) --
	(392.48,469.92) --
	(401.06,467.28) --
	(409.64,460.78) --
	(418.22,457.21) --
	(426.80,455.36) --
	(435.38,451.41) --
	(443.96,450.95) --
	(452.54,449.79) --
	(461.12,448.54) --
	(469.70,447.61) --
	(478.28,447.05) --
	(486.86,446.36) --
	(495.44,445.66) --
	(504.02,445.48) --
	(512.60,444.92) --
	(521.18,444.83) --
	(529.76,444.50) --
	(538.34,444.04) --
	(546.92,444.45) --
	(555.50,444.50) --
	(564.08,444.08) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.83) --
	( 49.28,444.92) --
	( 57.86,445.06) --
	( 66.44,445.66) --
	( 75.02,445.38) --
	( 83.60,446.22) --
	( 92.18,446.59) --
	(100.76,447.42) --
	(109.34,448.17) --
	(117.92,448.54) --
	(126.50,450.72) --
	(135.08,451.55) --
	(143.66,452.16) --
	(152.24,455.17) --
	(160.82,456.38) --
	(169.40,458.70) --
	(177.98,461.20) --
	(186.56,466.40) --
	(195.14,469.88) --
	(203.72,473.82) --
	(212.30,479.94) --
	(220.88,486.72) --
	(229.46,487.41) --
	(238.04,498.55) --
	(246.62,503.88) --
	(255.20,509.77) --
	(263.78,516.45) --
	(272.36,521.60) --
	(280.94,529.16) --
	(289.52,533.52) --
	(298.10,531.06) --
	(306.68,531.99) --
	(315.26,529.67) --
	(323.84,527.08) --
	(332.42,517.66) --
	(341.00,511.44) --
	(349.58,503.14) --
	(358.16,498.36) --
	(366.74,486.34) --
	(375.32,481.10) --
	(383.90,475.21) --
	(392.48,468.99) --
	(401.06,464.68) --
	(409.64,461.39) --
	(418.22,457.30) --
	(426.80,454.99) --
	(435.38,452.90) --
	(443.96,452.02) --
	(452.54,449.23) --
	(461.12,449.00) --
	(469.70,446.68) --
	(478.28,447.24) --
	(486.86,446.77) --
	(495.44,445.61) --
	(504.02,444.96) --
	(512.60,444.96) --
	(521.18,445.10) --
	(529.76,444.59) --
	(538.34,444.55) --
	(546.92,444.18) --
	(555.50,444.27) --
	(564.08,444.13) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.50) --
	( 49.28,444.83) --
	( 57.86,445.10) --
	( 66.44,445.24) --
	( 75.02,445.57) --
	( 83.60,445.15) --
	( 92.18,446.45) --
	(100.76,446.82) --
	(109.34,447.52) --
	(117.92,448.49) --
	(126.50,448.63) --
	(135.08,450.16) --
	(143.66,452.16) --
	(152.24,452.43) --
	(160.82,455.40) --
	(169.40,457.86) --
	(177.98,458.33) --
	(186.56,462.36) --
	(195.14,467.74) --
	(203.72,469.55) --
	(212.30,474.28) --
	(220.88,482.03) --
	(229.46,485.32) --
	(238.04,491.35) --
	(246.62,499.43) --
	(255.20,506.57) --
	(263.78,512.79) --
	(272.36,515.66) --
	(280.94,525.92) --
	(289.52,529.58) --
	(298.10,533.52) --
	(306.68,528.84) --
	(315.26,533.15) --
	(323.84,530.23) --
	(332.42,520.86) --
	(341.00,518.35) --
	(349.58,509.45) --
	(358.16,501.93) --
	(366.74,493.21) --
	(375.32,486.99) --
	(383.90,479.71) --
	(392.48,473.31) --
	(401.06,468.90) --
	(409.64,462.83) --
	(418.22,460.13) --
	(426.80,456.38) --
	(435.38,453.78) --
	(443.96,452.16) --
	(452.54,450.25) --
	(461.12,449.19) --
	(469.70,447.98) --
	(478.28,447.66) --
	(486.86,446.54) --
	(495.44,446.96) --
	(504.02,445.71) --
	(512.60,445.71) --
	(521.18,445.20) --
	(529.76,444.59) --
	(538.34,444.45) --
	(546.92,444.96) --
	(555.50,444.18) --
	(564.08,444.04) --
	(572.66,444.22);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.87) --
	( 49.28,444.69) --
	( 57.86,445.24) --
	( 66.44,445.29) --
	( 75.02,445.66) --
	( 83.60,446.26) --
	( 92.18,446.45) --
	(100.76,446.59) --
	(109.34,447.89) --
	(117.92,448.54) --
	(126.50,449.23) --
	(135.08,451.04) --
	(143.66,451.65) --
	(152.24,453.69) --
	(160.82,455.68) --
	(169.40,457.21) --
	(177.98,460.64) --
	(186.56,463.98) --
	(195.14,467.79) --
	(203.72,473.31) --
	(212.30,477.16) --
	(220.88,480.82) --
	(229.46,489.08) --
	(238.04,496.32) --
	(246.62,502.21) --
	(255.20,508.38) --
	(263.78,515.71) --
	(272.36,520.77) --
	(280.94,529.02) --
	(289.52,531.30) --
	(298.10,531.30) --
	(306.68,531.16) --
	(315.26,530.32) --
	(323.84,530.04) --
	(332.42,520.72) --
	(341.00,513.25) --
	(349.58,506.99) --
	(358.16,498.92) --
	(366.74,488.53) --
	(375.32,481.15) --
	(383.90,476.42) --
	(392.48,473.31) --
	(401.06,468.39) --
	(409.64,461.94) --
	(418.22,457.63) --
	(426.80,454.61) --
	(435.38,454.52) --
	(443.96,452.06) --
	(452.54,450.44) --
	(461.12,448.40) --
	(469.70,447.42) --
	(478.28,447.05) --
	(486.86,446.36) --
	(495.44,445.34) --
	(504.02,445.57) --
	(512.60,445.20) --
	(521.18,444.78) --
	(529.76,444.87) --
	(538.34,444.27) --
	(546.92,444.50) --
	(555.50,444.18) --
	(564.08,444.04) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.73) --
	( 49.28,444.96) --
	( 57.86,444.69) --
	( 66.44,445.29) --
	( 75.02,445.15) --
	( 83.60,445.94) --
	( 92.18,446.77) --
	(100.76,446.36) --
	(109.34,446.45) --
	(117.92,448.07) --
	(126.50,449.14) --
	(135.08,449.98) --
	(143.66,451.27) --
	(152.24,452.16) --
	(160.82,454.38) --
	(169.40,455.40) --
	(177.98,458.56) --
	(186.56,460.69) --
	(195.14,465.38) --
	(203.72,468.48) --
	(212.30,474.79) --
	(220.88,477.86) --
	(229.46,482.59) --
	(238.04,490.89) --
	(246.62,495.81) --
	(255.20,503.05) --
	(263.78,510.65) --
	(272.36,515.94) --
	(280.94,523.04) --
	(289.52,528.14) --
	(298.10,529.12) --
	(306.68,532.73) --
	(315.26,531.48) --
	(323.84,527.45) --
	(332.42,524.89) --
	(341.00,516.03) --
	(349.58,512.97) --
	(358.16,506.43) --
	(366.74,495.81) --
	(375.32,488.39) --
	(383.90,480.41) --
	(392.48,475.91) --
	(401.06,473.26) --
	(409.64,467.05) --
	(418.22,460.27) --
	(426.80,457.44) --
	(435.38,456.33) --
	(443.96,453.22) --
	(452.54,451.13) --
	(461.12,449.60) --
	(469.70,448.21) --
	(478.28,447.42) --
	(486.86,446.63) --
	(495.44,446.31) --
	(504.02,446.03) --
	(512.60,445.34) --
	(521.18,444.87) --
	(529.76,445.06) --
	(538.34,444.59) --
	(546.92,444.36) --
	(555.50,444.32) --
	(564.08,444.36) --
	(572.66,444.18);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.83) --
	( 49.28,444.83) --
	( 57.86,444.92) --
	( 66.44,445.06) --
	( 75.02,445.29) --
	( 83.60,445.94) --
	( 92.18,447.05) --
	(100.76,446.45) --
	(109.34,446.59) --
	(117.92,448.03) --
	(126.50,449.56) --
	(135.08,450.58) --
	(143.66,450.86) --
	(152.24,452.94) --
	(160.82,454.15) --
	(169.40,456.33) --
	(177.98,459.48) --
	(186.56,461.29) --
	(195.14,465.75) --
	(203.72,470.53) --
	(212.30,474.28) --
	(220.88,478.78) --
	(229.46,485.56) --
	(238.04,491.59) --
	(246.62,496.41) --
	(255.20,506.34) --
	(263.78,511.53) --
	(272.36,517.47) --
	(280.94,525.08) --
	(289.52,529.16) --
	(298.10,528.47) --
	(306.68,534.27) --
	(315.26,531.95) --
	(323.84,527.12) --
	(332.42,522.48) --
	(341.00,518.77) --
	(349.58,513.48) --
	(358.16,503.83) --
	(366.74,492.75) --
	(375.32,487.46) --
	(383.90,479.02) --
	(392.48,476.00) --
	(401.06,470.15) --
	(409.64,464.63) --
	(418.22,459.76) --
	(426.80,456.66) --
	(435.38,455.36) --
	(443.96,452.53) --
	(452.54,450.39) --
	(461.12,448.82) --
	(469.70,448.77) --
	(478.28,447.01) --
	(486.86,445.94) --
	(495.44,445.71) --
	(504.02,445.85) --
	(512.60,445.38) --
	(521.18,444.87) --
	(529.76,444.55) --
	(538.34,444.41) --
	(546.92,444.27) --
	(555.50,444.27) --
	(564.08,444.32) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.36) --
	( 49.28,444.96) --
	( 57.86,444.64) --
	( 66.44,445.10) --
	( 75.02,444.78) --
	( 83.60,445.57) --
	( 92.18,446.08) --
	(100.76,447.24) --
	(109.34,446.54) --
	(117.92,446.54) --
	(126.50,448.77) --
	(135.08,449.74) --
	(143.66,450.16) --
	(152.24,451.65) --
	(160.82,453.73) --
	(169.40,454.61) --
	(177.98,457.26) --
	(186.56,459.62) --
	(195.14,462.64) --
	(203.72,466.54) --
	(212.30,472.34) --
	(220.88,474.93) --
	(229.46,482.45) --
	(238.04,487.18) --
	(246.62,491.45) --
	(255.20,499.01) --
	(263.78,509.26) --
	(272.36,514.41) --
	(280.94,519.19) --
	(289.52,526.52) --
	(298.10,529.35) --
	(306.68,531.58) --
	(315.26,535.47) --
	(323.84,529.44) --
	(332.42,524.99) --
	(341.00,520.86) --
	(349.58,520.49) --
	(358.16,506.85) --
	(366.74,501.28) --
	(375.32,494.18) --
	(383.90,483.70) --
	(392.48,479.99) --
	(401.06,473.36) --
	(409.64,468.11) --
	(418.22,463.24) --
	(426.80,459.16) --
	(435.38,455.91) --
	(443.96,453.64) --
	(452.54,450.44) --
	(461.12,449.88) --
	(469.70,448.86) --
	(478.28,448.12) --
	(486.86,446.68) --
	(495.44,446.50) --
	(504.02,445.52) --
	(512.60,445.89) --
	(521.18,444.69) --
	(529.76,445.06) --
	(538.34,444.83) --
	(546.92,444.27) --
	(555.50,444.41) --
	(564.08,444.55) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.50) --
	( 49.28,444.92) --
	( 57.86,444.69) --
	( 66.44,444.87) --
	( 75.02,445.24) --
	( 83.60,445.66) --
	( 92.18,446.03) --
	(100.76,446.87) --
	(109.34,447.28) --
	(117.92,446.82) --
	(126.50,448.63) --
	(135.08,449.33) --
	(143.66,451.41) --
	(152.24,451.46) --
	(160.82,453.92) --
	(169.40,454.85) --
	(177.98,457.58) --
	(186.56,459.95) --
	(195.14,463.20) --
	(203.72,468.07) --
	(212.30,471.87) --
	(220.88,476.37) --
	(229.46,482.26) --
	(238.04,486.07) --
	(246.62,492.51) --
	(255.20,501.61) --
	(263.78,510.47) --
	(272.36,516.03) --
	(280.94,522.25) --
	(289.52,524.89) --
	(298.10,531.44) --
	(306.68,534.45) --
	(315.26,534.36) --
	(323.84,528.14) --
	(332.42,522.71) --
	(341.00,519.47) --
	(349.58,515.34) --
	(358.16,509.03) --
	(366.74,499.75) --
	(375.32,493.03) --
	(383.90,485.51) --
	(392.48,474.98) --
	(401.06,472.66) --
	(409.64,468.25) --
	(418.22,463.01) --
	(426.80,458.14) --
	(435.38,455.91) --
	(443.96,454.52) --
	(452.54,450.39) --
	(461.12,449.51) --
	(469.70,449.28) --
	(478.28,447.19) --
	(486.86,446.82) --
	(495.44,446.08) --
	(504.02,445.66) --
	(512.60,445.71) --
	(521.18,444.64) --
	(529.76,444.87) --
	(538.34,444.83) --
	(546.92,444.36) --
	(555.50,444.27) --
	(564.08,444.27) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.96) --
	( 49.28,444.36) --
	( 57.86,444.78) --
	( 66.44,445.20) --
	( 75.02,445.66) --
	( 83.60,446.17) --
	( 92.18,446.26) --
	(100.76,447.10) --
	(109.34,447.93) --
	(117.92,448.03) --
	(126.50,448.86) --
	(135.08,450.16) --
	(143.66,451.88) --
	(152.24,452.80) --
	(160.82,455.36) --
	(169.40,457.40) --
	(177.98,460.04) --
	(186.56,462.83) --
	(195.14,466.72) --
	(203.72,472.06) --
	(212.30,478.09) --
	(220.88,482.87) --
	(229.46,485.74) --
	(238.04,496.64) --
	(246.62,501.65) --
	(255.20,511.44) --
	(263.78,515.57) --
	(272.36,524.48) --
	(280.94,530.28) --
	(289.52,533.11) --
	(298.10,537.51) --
	(306.68,535.94) --
	(315.26,535.10) --
	(323.84,528.93) --
	(332.42,522.90) --
	(341.00,515.57) --
	(349.58,508.61) --
	(358.16,500.22) --
	(366.74,490.84) --
	(375.32,482.49) --
	(383.90,476.05) --
	(392.48,469.41) --
	(401.06,464.68) --
	(409.64,460.23) --
	(418.22,457.58) --
	(426.80,454.75) --
	(435.38,452.06) --
	(443.96,450.44) --
	(452.54,449.56) --
	(461.12,448.63) --
	(469.70,446.91) --
	(478.28,446.03) --
	(486.86,446.03) --
	(495.44,445.80) --
	(504.02,444.92) --
	(512.60,445.01) --
	(521.18,445.01) --
	(529.76,444.45) --
	(538.34,444.59) --
	(546.92,444.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (564.08,444.13) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,445.01) --
	( 49.28,444.59) --
	( 57.86,444.50) --
	( 66.44,445.29) --
	( 75.02,445.20) --
	( 83.60,446.17) --
	( 92.18,446.26) --
	(100.76,446.87) --
	(109.34,447.05) --
	(117.92,448.21) --
	(126.50,448.63) --
	(135.08,449.93) --
	(143.66,450.95) --
	(152.24,452.71) --
	(160.82,454.71) --
	(169.40,455.82) --
	(177.98,460.37) --
	(186.56,461.62) --
	(195.14,464.63) --
	(203.72,471.27) --
	(212.30,475.63) --
	(220.88,480.13) --
	(229.46,486.02) --
	(238.04,494.37) --
	(246.62,499.66) --
	(255.20,508.06) --
	(263.78,514.69) --
	(272.36,521.83) --
	(280.94,526.19) --
	(289.52,532.97) --
	(298.10,534.36) --
	(306.68,536.17) --
	(315.26,532.60) --
	(323.84,532.55) --
	(332.42,526.29) --
	(341.00,519.84) --
	(349.58,509.82) --
	(358.16,498.59) --
	(366.74,494.51) --
	(375.32,486.02) --
	(383.90,478.55) --
	(392.48,471.59) --
	(401.06,467.79) --
	(409.64,463.38) --
	(418.22,459.07) --
	(426.80,456.47) --
	(435.38,452.76) --
	(443.96,451.27) --
	(452.54,449.56) --
	(461.12,449.70) --
	(469.70,448.17) --
	(478.28,446.68) --
	(486.86,446.12) --
	(495.44,445.80) --
	(504.02,445.52) --
	(512.60,445.15) --
	(521.18,444.83) --
	(529.76,444.55) --
	(538.34,444.59) --
	(546.92,444.41) --
	(555.50,444.36) --
	(564.08,444.27) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.78) --
	( 49.28,444.78) --
	( 57.86,444.36) --
	( 66.44,445.24) --
	( 75.02,445.06) --
	( 83.60,445.57) --
	( 92.18,446.59) --
	(100.76,446.87) --
	(109.34,446.40) --
	(117.92,447.89) --
	(126.50,448.58) --
	(135.08,449.65) --
	(143.66,450.53) --
	(152.24,451.78) --
	(160.82,453.83) --
	(169.40,455.17) --
	(177.98,458.33) --
	(186.56,461.48) --
	(195.14,462.96) --
	(203.72,468.53) --
	(212.30,472.20) --
	(220.88,478.88) --
	(229.46,484.44) --
	(238.04,488.94) --
	(246.62,496.18) --
	(255.20,504.53) --
	(263.78,512.83) --
	(272.36,518.31) --
	(280.94,524.80) --
	(289.52,528.00) --
	(298.10,536.31) --
	(306.68,534.45) --
	(315.26,535.61) --
	(323.84,532.73) --
	(332.42,528.10) --
	(341.00,519.88) --
	(349.58,514.64) --
	(358.16,502.86) --
	(366.74,498.87) --
	(375.32,490.38) --
	(383.90,481.38) --
	(392.48,475.21) --
	(401.06,467.23) --
	(409.64,466.35) --
	(418.22,461.20) --
	(426.80,456.61) --
	(435.38,453.59) --
	(443.96,452.94) --
	(452.54,450.86) --
	(461.12,449.56) --
	(469.70,447.42) --
	(478.28,447.93) --
	(486.86,446.31) --
	(495.44,446.12) --
	(504.02,445.06) --
	(512.60,445.57) --
	(521.18,444.78) --
	(529.76,444.59) --
	(538.34,444.41) --
	(546.92,444.18) --
	(555.50,444.32) --
	(564.08,444.32) --
	(572.66,444.18);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.92) --
	( 49.28,444.55) --
	( 57.86,444.78) --
	( 66.44,445.38) --
	( 75.02,445.24) --
	( 83.60,446.17) --
	( 92.18,446.96) --
	(100.76,446.45) --
	(109.34,447.15) --
	(117.92,448.54) --
	(126.50,449.88) --
	(135.08,450.16) --
	(143.66,451.13) --
	(152.24,452.94) --
	(160.82,454.75) --
	(169.40,457.72) --
	(177.98,461.16) --
	(186.56,463.01) --
	(195.14,466.07) --
	(203.72,472.38) --
	(212.30,477.07) --
	(220.88,482.08) --
	(229.46,488.90) --
	(238.04,495.48) --
	(246.62,504.76) --
	(255.20,508.01) --
	(263.78,514.78) --
	(272.36,526.84) --
	(280.94,526.29) --
	(289.52,536.82) --
	(298.10,534.78) --
	(306.68,535.24) --
	(315.26,536.03) --
	(323.84,528.05) --
	(332.42,521.23) --
	(341.00,515.71) --
	(349.58,507.64) --
	(358.16,502.58) --
	(366.74,488.48) --
	(375.32,483.75) --
	(383.90,474.47) --
	(392.48,469.23) --
	(401.06,467.32) --
	(409.64,461.80) --
	(418.22,456.70) --
	(426.80,456.33) --
	(435.38,451.88) --
	(443.96,451.83) --
	(452.54,449.79) --
	(461.12,447.93) --
	(469.70,447.24) --
	(478.28,446.08) --
	(486.86,445.99) --
	(495.44,445.75) --
	(504.02,445.06) --
	(512.60,444.73) --
	(521.18,444.83) --
	(529.76,444.50) --
	(538.34,444.18) --
	(546.92,444.18) --
	(555.50,444.27) --
	(564.08,444.04) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.83) --
	( 49.28,444.59) --
	( 57.86,444.83) --
	( 66.44,445.38) --
	( 75.02,445.24) --
	( 83.60,445.89) --
	( 92.18,446.50) --
	(100.76,446.91) --
	(109.34,446.54) --
	(117.92,448.03) --
	(126.50,450.07) --
	(135.08,449.74) --
	(143.66,450.72) --
	(152.24,452.94) --
	(160.82,454.34) --
	(169.40,456.66) --
	(177.98,459.76) --
	(186.56,461.48) --
	(195.14,465.33) --
	(203.72,471.18) --
	(212.30,473.87) --
	(220.88,479.29) --
	(229.46,487.60) --
	(238.04,493.95) --
	(246.62,499.89) --
	(255.20,505.60) --
	(263.78,512.97) --
	(272.36,522.62) --
	(280.94,524.15) --
	(289.52,531.58) --
	(298.10,534.31) --
	(306.68,537.61) --
	(315.26,533.66) --
	(323.84,533.85) --
	(332.42,525.08) --
	(341.00,519.19) --
	(349.58,509.77) --
	(358.16,505.36) --
	(366.74,492.38) --
	(375.32,484.67) --
	(383.90,478.88) --
	(392.48,472.01) --
	(401.06,467.37) --
	(409.64,464.08) --
	(418.22,457.63) --
	(426.80,455.96) --
	(435.38,451.78) --
	(443.96,451.41) --
	(452.54,449.84) --
	(461.12,448.17) --
	(469.70,447.52) --
	(478.28,446.59) --
	(486.86,446.17) --
	(495.44,445.61) --
	(504.02,445.24) --
	(512.60,444.96) --
	(521.18,444.36) --
	(529.76,444.69) --
	(538.34,444.45) --
	(546.92,444.45) --
	(555.50,443.99) --
	(564.08,444.18) --
	(572.66,444.18);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.73) --
	( 49.28,444.73) --
	( 57.86,445.15) --
	( 66.44,445.24) --
	( 75.02,445.52) --
	( 83.60,446.08) --
	( 92.18,446.31) --
	(100.76,447.61) --
	(109.34,447.42) --
	(117.92,448.03) --
	(126.50,449.79) --
	(135.08,450.44) --
	(143.66,451.97) --
	(152.24,453.22) --
	(160.82,454.15) --
	(169.40,458.33) --
	(177.98,460.32) --
	(186.56,463.80) --
	(195.14,467.05) --
	(203.72,472.06) --
	(212.30,477.67) --
	(220.88,480.50) --
	(229.46,488.48) --
	(238.04,493.44) --
	(246.62,503.18) --
	(255.20,513.30) --
	(263.78,514.13) --
	(272.36,520.44) --
	(280.94,527.40) --
	(289.52,533.66) --
	(298.10,532.73) --
	(306.68,535.61) --
	(315.26,532.78) --
	(323.84,527.96) --
	(332.42,517.80) --
	(341.00,518.73) --
	(349.58,504.95) --
	(358.16,499.66) --
	(366.74,491.54) --
	(375.32,482.96) --
	(383.90,475.26) --
	(392.48,470.48) --
	(401.06,467.05) --
	(409.64,460.00) --
	(418.22,457.35) --
	(426.80,454.75) --
	(435.38,452.71) --
	(443.96,450.86) --
	(452.54,449.93) --
	(461.12,448.21) --
	(469.70,447.10) --
	(478.28,447.10) --
	(486.86,446.31) --
	(495.44,445.34) --
	(504.02,445.29) --
	(512.60,444.83) --
	(521.18,444.59) --
	(529.76,444.36) --
	(538.34,444.59) --
	(546.92,444.08) --
	(555.50,444.13);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.69) --
	( 49.28,444.64) --
	( 57.86,444.92) --
	( 66.44,444.83) --
	( 75.02,445.34) --
	( 83.60,445.48) --
	( 92.18,446.26) --
	(100.76,446.22) --
	(109.34,447.84) --
	(117.92,447.42) --
	(126.50,449.00) --
	(135.08,448.95) --
	(143.66,450.86) --
	(152.24,452.94) --
	(160.82,452.94) --
	(169.40,454.06) --
	(177.98,458.33) --
	(186.56,461.29) --
	(195.14,464.87) --
	(203.72,466.91) --
	(212.30,472.80) --
	(220.88,476.79) --
	(229.46,482.96) --
	(238.04,489.64) --
	(246.62,495.07) --
	(255.20,504.39) --
	(263.78,511.72) --
	(272.36,515.01) --
	(280.94,521.97) --
	(289.52,531.48) --
	(298.10,531.81) --
	(306.68,536.54) --
	(315.26,533.85) --
	(323.84,530.28) --
	(332.42,528.42) --
	(341.00,520.07) --
	(349.58,512.56) --
	(358.16,503.56) --
	(366.74,500.03) --
	(375.32,490.94) --
	(383.90,482.73) --
	(392.48,477.53) --
	(401.06,469.64) --
	(409.64,465.75) --
	(418.22,459.76) --
	(426.80,457.07) --
	(435.38,453.50) --
	(443.96,452.16) --
	(452.54,450.81) --
	(461.12,449.09) --
	(469.70,447.98) --
	(478.28,447.05) --
	(486.86,446.36) --
	(495.44,446.31) --
	(504.02,445.24) --
	(512.60,445.29) --
	(521.18,444.59) --
	(529.76,444.73) --
	(538.34,444.55) --
	(546.92,444.45) --
	(555.50,444.04) --
	(564.08,444.13) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.55) --
	( 49.28,444.96) --
	( 57.86,444.64) --
	( 66.44,445.01) --
	( 75.02,445.15) --
	( 83.60,445.75) --
	( 92.18,445.94) --
	(100.76,446.54) --
	(109.34,447.75) --
	(117.92,447.15) --
	(126.50,449.51) --
	(135.08,448.95) --
	(143.66,451.32) --
	(152.24,452.76) --
	(160.82,452.71) --
	(169.40,454.66) --
	(177.98,457.95) --
	(186.56,462.69) --
	(195.14,464.03) --
	(203.72,467.74) --
	(212.30,473.54) --
	(220.88,476.97) --
	(229.46,482.40) --
	(238.04,491.08) --
	(246.62,497.39) --
	(255.20,503.28) --
	(263.78,510.51) --
	(272.36,517.29) --
	(280.94,521.74) --
	(289.52,529.77) --
	(298.10,532.64) --
	(306.68,537.61) --
	(315.26,534.82) --
	(323.84,530.97) --
	(332.42,526.70) --
	(341.00,521.93) --
	(349.58,509.17) --
	(358.16,504.25) --
	(366.74,495.81) --
	(375.32,487.83) --
	(383.90,479.29) --
	(392.48,476.37) --
	(401.06,469.64) --
	(409.64,465.79) --
	(418.22,460.37) --
	(426.80,457.12) --
	(435.38,455.36) --
	(443.96,451.78) --
	(452.54,450.81) --
	(461.12,448.68) --
	(469.70,448.17) --
	(478.28,446.91) --
	(486.86,446.91) --
	(495.44,445.94) --
	(504.02,445.80) --
	(512.60,445.06) --
	(521.18,444.59) --
	(529.76,444.32) --
	(538.34,444.36) --
	(546.92,444.50) --
	(555.50,444.18) --
	(564.08,444.27) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.41) --
	( 49.28,444.96) --
	( 57.86,444.83) --
	( 66.44,445.01) --
	( 75.02,445.29) --
	( 83.60,445.85) --
	( 92.18,446.08) --
	(100.76,446.59) --
	(109.34,447.42) --
	(117.92,448.35) --
	(126.50,448.77) --
	(135.08,449.05) --
	(143.66,451.41) --
	(152.24,453.27) --
	(160.82,453.04) --
	(169.40,455.36) --
	(177.98,459.21) --
	(186.56,462.13) --
	(195.14,465.19) --
	(203.72,467.93) --
	(212.30,475.35) --
	(220.88,478.18) --
	(229.46,482.63) --
	(238.04,492.56) --
	(246.62,500.40) --
	(255.20,503.09) --
	(263.78,510.37) --
	(272.36,517.52) --
	(280.94,526.43) --
	(289.52,529.49) --
	(298.10,535.56) --
	(306.68,534.45) --
	(315.26,533.57) --
	(323.84,531.06) --
	(332.42,524.25) --
	(341.00,518.31) --
	(349.58,511.30) --
	(358.16,502.86) --
	(366.74,494.23) --
	(375.32,487.92) --
	(383.90,480.96) --
	(392.48,474.84) --
	(401.06,470.80) --
	(409.64,463.01) --
	(418.22,459.21) --
	(426.80,457.77) --
	(435.38,453.59) --
	(443.96,451.69) --
	(452.54,451.88) --
	(461.12,448.26) --
	(469.70,447.70) --
	(478.28,447.98) --
	(486.86,446.17) --
	(495.44,446.17) --
	(504.02,445.94) --
	(512.60,444.73) --
	(521.18,445.10) --
	(529.76,444.55) --
	(538.34,444.83) --
	(546.92,444.08) --
	(555.50,444.04) --
	(564.08,444.22) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.45) --
	( 49.28,444.64) --
	( 57.86,444.96) --
	( 66.44,445.24) --
	( 75.02,445.24) --
	( 83.60,445.66) --
	( 92.18,446.17) --
	(100.76,446.82) --
	(109.34,447.52) --
	(117.92,448.03) --
	(126.50,448.72) --
	(135.08,449.42) --
	(143.66,451.92) --
	(152.24,452.20) --
	(160.82,453.83) --
	(169.40,456.10) --
	(177.98,458.74) --
	(186.56,461.25) --
	(195.14,465.65) --
	(203.72,468.81) --
	(212.30,473.87) --
	(220.88,478.18) --
	(229.46,485.97) --
	(238.04,492.05) --
	(246.62,498.96) --
	(255.20,504.99) --
	(263.78,510.42) --
	(272.36,518.82) --
	(280.94,527.68) --
	(289.52,527.35) --
	(298.10,536.86) --
	(306.68,535.19) --
	(315.26,535.84) --
	(323.84,528.93) --
	(332.42,525.96) --
	(341.00,518.45) --
	(349.58,511.67) --
	(358.16,501.05) --
	(366.74,495.11) --
	(375.32,485.46) --
	(383.90,479.90) --
	(392.48,473.12) --
	(401.06,469.51) --
	(409.64,463.75) --
	(418.22,460.00) --
	(426.80,455.82) --
	(435.38,452.94) --
	(443.96,451.74) --
	(452.54,450.72) --
	(461.12,448.21) --
	(469.70,447.28) --
	(478.28,447.01) --
	(486.86,446.22) --
	(495.44,446.12) --
	(504.02,445.94) --
	(512.60,445.06) --
	(521.18,444.55) --
	(529.76,444.64) --
	(538.34,444.45) --
	(546.92,444.64) --
	(555.50,444.13) --
	(564.08,444.13) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.96) --
	( 49.28,444.22) --
	( 57.86,445.29) --
	( 66.44,445.34) --
	( 75.02,445.71) --
	( 83.60,445.75) --
	( 92.18,446.91) --
	(100.76,446.17) --
	(109.34,447.47) --
	(117.92,448.35) --
	(126.50,449.42) --
	(135.08,449.79) --
	(143.66,450.49) --
	(152.24,453.64) --
	(160.82,455.50) --
	(169.40,456.75) --
	(177.98,460.18) --
	(186.56,462.50) --
	(195.14,467.28) --
	(203.72,471.36) --
	(212.30,478.13) --
	(220.88,480.87) --
	(229.46,489.59) --
	(238.04,494.42) --
	(246.62,502.21) --
	(255.20,512.97) --
	(263.78,513.85) --
	(272.36,525.64) --
	(280.94,527.26) --
	(289.52,534.31) --
	(298.10,539.65) --
	(306.68,533.43) --
	(315.26,532.41) --
	(323.84,530.69) --
	(332.42,526.61) --
	(341.00,513.16) --
	(349.58,507.73) --
	(358.16,497.48) --
	(366.74,492.42) --
	(375.32,481.01) --
	(383.90,476.88) --
	(392.48,469.46) --
	(401.06,465.10) --
	(409.64,461.99) --
	(418.22,458.14) --
	(426.80,454.71) --
	(435.38,452.71) --
	(443.96,450.86) --
	(452.54,449.93) --
	(461.12,448.86) --
	(469.70,447.10) --
	(478.28,447.05) --
	(486.86,445.29) --
	(495.44,445.38) --
	(504.02,445.24) --
	(512.60,444.92) --
	(521.18,445.01) --
	(529.76,444.32) --
	(538.34,444.55) --
	(546.92,444.08) --
	(555.50,444.13) --
	(564.08,444.13);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.73) --
	( 49.28,444.36) --
	( 57.86,445.20) --
	( 66.44,445.52) --
	( 75.02,445.29) --
	( 83.60,446.40) --
	( 92.18,446.45) --
	(100.76,446.73) --
	(109.34,447.42) --
	(117.92,448.26) --
	(126.50,449.05) --
	(135.08,450.16) --
	(143.66,451.18) --
	(152.24,453.04) --
	(160.82,455.82) --
	(169.40,457.35) --
	(177.98,460.09) --
	(186.56,462.50) --
	(195.14,466.72) --
	(203.72,473.36) --
	(212.30,477.76) --
	(220.88,480.96) --
	(229.46,487.69) --
	(238.04,494.93) --
	(246.62,500.08) --
	(255.20,512.04) --
	(263.78,514.78) --
	(272.36,523.27) --
	(280.94,527.49) --
	(289.52,534.03) --
	(298.10,538.58) --
	(306.68,535.15) --
	(315.26,529.67) --
	(323.84,530.55) --
	(332.42,527.35) --
	(341.00,514.74) --
	(349.58,507.64) --
	(358.16,497.25) --
	(366.74,493.30) --
	(375.32,482.45) --
	(383.90,472.89) --
	(392.48,470.80) --
	(401.06,466.03) --
	(409.64,461.71) --
	(418.22,455.77) --
	(426.80,456.38) --
	(435.38,452.29) --
	(443.96,450.67) --
	(452.54,449.56) --
	(461.12,448.82) --
	(469.70,447.52) --
	(478.28,447.10) --
	(486.86,445.85) --
	(495.44,445.24) --
	(504.02,445.43) --
	(512.60,445.48) --
	(521.18,444.73) --
	(529.76,444.13) --
	(538.34,444.73) --
	(546.92,444.32) --
	(555.50,444.08) --
	(564.08,444.22);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.64) --
	( 49.28,444.55) --
	( 57.86,445.29) --
	( 66.44,445.57) --
	( 75.02,445.48) --
	( 83.60,446.12) --
	( 92.18,446.26) --
	(100.76,447.24) --
	(109.34,447.52) --
	(117.92,448.12) --
	(126.50,449.42) --
	(135.08,449.93) --
	(143.66,451.13) --
	(152.24,453.41) --
	(160.82,455.22) --
	(169.40,457.35) --
	(177.98,460.37) --
	(186.56,463.85) --
	(195.14,466.17) --
	(203.72,474.05) --
	(212.30,476.46) --
	(220.88,481.75) --
	(229.46,488.94) --
	(238.04,495.58) --
	(246.62,501.33) --
	(255.20,511.67) --
	(263.78,519.24) --
	(272.36,521.93) --
	(280.94,528.61) --
	(289.52,534.27) --
	(298.10,536.77) --
	(306.68,534.96) --
	(315.26,533.06) --
	(323.84,528.61) --
	(332.42,522.95) --
	(341.00,514.36) --
	(349.58,509.26) --
	(358.16,497.71) --
	(366.74,490.47) --
	(375.32,481.33) --
	(383.90,474.98) --
	(392.48,467.42) --
	(401.06,464.96) --
	(409.64,460.69) --
	(418.22,457.91) --
	(426.80,454.47) --
	(435.38,452.29) --
	(443.96,450.62) --
	(452.54,449.46) --
	(461.12,448.12) --
	(469.70,447.70) --
	(478.28,446.73) --
	(486.86,446.50) --
	(495.44,445.01) --
	(504.02,445.24) --
	(512.60,444.78) --
	(521.18,444.45) --
	(529.76,444.55) --
	(538.34,444.27) --
	(546.92,444.45) --
	(555.50,444.32) --
	(564.08,444.08) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.41) --
	( 49.28,444.87) --
	( 57.86,445.01) --
	( 66.44,445.43) --
	( 75.02,445.38) --
	( 83.60,446.40) --
	( 92.18,446.63) --
	(100.76,446.96) --
	(109.34,447.42) --
	(117.92,447.70) --
	(126.50,449.56) --
	(135.08,450.58) --
	(143.66,450.86) --
	(152.24,453.69) --
	(160.82,454.29) --
	(169.40,459.02) --
	(177.98,459.62) --
	(186.56,463.34) --
	(195.14,466.77) --
	(203.72,472.66) --
	(212.30,475.63) --
	(220.88,482.77) --
	(229.46,489.08) --
	(238.04,493.54) --
	(246.62,503.28) --
	(255.20,506.20) --
	(263.78,518.86) --
	(272.36,521.74) --
	(280.94,527.86) --
	(289.52,533.01) --
	(298.10,537.93) --
	(306.68,536.03) --
	(315.26,529.44) --
	(323.84,530.46) --
	(332.42,525.50) --
	(341.00,515.01) --
	(349.58,506.80) --
	(358.16,500.45) --
	(366.74,491.87) --
	(375.32,482.31) --
	(383.90,476.37) --
	(392.48,469.64) --
	(401.06,464.96) --
	(409.64,460.55) --
	(418.22,457.49) --
	(426.80,454.89) --
	(435.38,454.38) --
	(443.96,450.72) --
	(452.54,448.91) --
	(461.12,447.79) --
	(469.70,448.07) --
	(478.28,446.87) --
	(486.86,445.99) --
	(495.44,445.24) --
	(504.02,445.34) --
	(512.60,444.73) --
	(521.18,444.69) --
	(529.76,444.41) --
	(538.34,444.41) --
	(546.92,444.27) --
	(555.50,444.18) --
	(564.08,444.32) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.27) --
	( 49.28,444.55) --
	( 57.86,445.38) --
	( 66.44,445.24) --
	( 75.02,445.61) --
	( 83.60,446.31) --
	( 92.18,446.40) --
	(100.76,446.73) --
	(109.34,447.61) --
	(117.92,448.40) --
	(126.50,449.37) --
	(135.08,450.58) --
	(143.66,451.18) --
	(152.24,452.90) --
	(160.82,454.06) --
	(169.40,458.65) --
	(177.98,459.72) --
	(186.56,463.52) --
	(195.14,466.26) --
	(203.72,472.15) --
	(212.30,476.83) --
	(220.88,479.94) --
	(229.46,488.71) --
	(238.04,494.46) --
	(246.62,501.93) --
	(255.20,506.01) --
	(263.78,517.43) --
	(272.36,523.50) --
	(280.94,527.35) --
	(289.52,533.38) --
	(298.10,535.66) --
	(306.68,535.52) --
	(315.26,533.99) --
	(323.84,530.69) --
	(332.42,524.94) --
	(341.00,516.50) --
	(349.58,506.43) --
	(358.16,501.70) --
	(366.74,493.72) --
	(375.32,481.15) --
	(383.90,474.14) --
	(392.48,470.76) --
	(401.06,465.28) --
	(409.64,461.76) --
	(418.22,457.40) --
	(426.80,455.82) --
	(435.38,452.71) --
	(443.96,450.02) --
	(452.54,448.54) --
	(461.12,448.86) --
	(469.70,448.21) --
	(478.28,446.54) --
	(486.86,446.08) --
	(495.44,445.29) --
	(504.02,445.10) --
	(512.60,445.06) --
	(521.18,444.78) --
	(529.76,444.50) --
	(538.34,444.50) --
	(546.92,444.27) --
	(555.50,444.41) --
	(564.08,444.08) --
	(572.66,444.18);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.70,443.99) rectangle (572.66,555.28);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.75,464.11) {500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.75,487.30) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.75,510.50) {1500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.75,533.69) {2000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.95,467.14) --
	( 40.70,467.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.95,490.33) --
	( 40.70,490.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.95,513.53) --
	( 40.70,513.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.95,536.72) --
	( 40.70,536.72);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,441.24) --
	( 40.70,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (126.50,441.24) --
	(126.50,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.30,441.24) --
	(212.30,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (298.10,441.24) --
	(298.10,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (383.90,441.24) --
	(383.90,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.70,441.24) --
	(469.70,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.50,441.24) --
	(555.50,443.99);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 40.70,432.98) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (126.50,432.98) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (212.30,432.98) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (298.10,432.98) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (383.90,432.98) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (469.70,432.98) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (555.50,432.98) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (306.68,420.80) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,499.64) {Daily incidence};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);

\path[] ( 45.09,453.59) rectangle (121.42,556.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);

\path[] ( 45.09,453.59) rectangle (121.42,556.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 50.59,528.47) rectangle ( 67.93,545.81);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.32,537.14) -- ( 66.20,537.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 50.59,511.12) rectangle ( 67.93,528.47);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.32,519.79) -- ( 66.20,519.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 50.59,493.78) rectangle ( 67.93,511.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.32,502.45) -- ( 66.20,502.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 50.59,476.43) rectangle ( 67.93,493.78);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.32,485.11) -- ( 66.20,485.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 50.59,459.09) rectangle ( 67.93,476.43);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 52.32,467.76) -- ( 66.20,467.76);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 73.43,534.11) {$\rho=0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 73.43,516.76) {$\rho=0.2$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 73.43,499.42) {$\rho=0.4$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 73.43,482.07) {$\rho=0.6$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 73.43,464.73) {$\rho=0.8$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 40.70,563.57) {A};
\end{scope}
\begin{scope}
\path[clip] (  0.00,206.49) rectangle (192.72,412.97);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,206.49) rectangle (192.72,412.97);
\end{scope}
\begin{scope}
\path[clip] ( 27.50,237.50) rectangle (187.22,390.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.50,237.50) rectangle (187.22,390.09);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,356.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,356.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,358.01) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,357.20) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,357.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,357.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,358.23) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,357.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,357.73) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,357.72) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,357.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,358.48) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,357.72) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,358.49) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,354.92) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,356.93) circle (  2.50);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,356.64) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,357.22) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,357.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,357.23) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,357.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,357.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,358.20) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,357.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,357.88) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,357.66) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,357.56) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,358.51) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,357.78) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,358.07) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,357.24) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,355.42) circle (  2.50);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,357.11) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,357.46) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,357.42) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,357.37) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,357.66) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,357.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,358.20) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,357.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,357.88) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,357.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,357.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,358.38) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,357.37) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,359.12) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,355.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,356.17) circle (  2.50);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,357.29) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,357.65) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,357.48) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,357.69) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,357.34) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,358.05) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,358.17) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,357.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,357.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,357.55) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,357.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,358.36) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,357.69) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,358.30) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,357.14) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,357.74) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,358.36) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,357.14) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,357.19) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,357.54) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,357.47) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,358.07) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,358.17) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,357.84) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,357.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,357.55) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,357.66) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,358.50) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,357.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,357.56) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,358.55) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,356.45) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,356.77) --
	( 29.46,356.84) --
	( 31.41,356.90) --
	( 33.37,356.97) --
	( 35.32,357.03) --
	( 37.28,357.09) --
	( 39.24,357.15) --
	( 41.19,357.21) --
	( 43.15,357.27) --
	( 45.11,357.32) --
	( 47.06,357.37) --
	( 49.02,357.42) --
	( 50.98,357.47) --
	( 52.93,357.51) --
	( 54.89,357.56) --
	( 56.85,357.60) --
	( 58.80,357.64) --
	( 60.76,357.67) --
	( 62.72,357.71) --
	( 64.67,357.74) --
	( 66.63,357.77) --
	( 68.59,357.79) --
	( 70.54,357.82) --
	( 72.50,357.84) --
	( 74.46,357.86) --
	( 76.41,357.88) --
	( 78.37,357.89) --
	( 80.33,357.91) --
	( 82.28,357.91) --
	( 84.24,357.92) --
	( 86.20,357.92) --
	( 88.15,357.92) --
	( 90.11,357.91) --
	( 92.07,357.91) --
	( 94.02,357.91) --
	( 95.98,357.90) --
	( 97.94,357.90) --
	( 99.89,357.91) --
	(101.85,357.91) --
	(103.80,357.91) --
	(105.76,357.91) --
	(107.72,357.91) --
	(109.67,357.91) --
	(111.63,357.91) --
	(113.59,357.91) --
	(115.54,357.91) --
	(117.50,357.91) --
	(119.46,357.90) --
	(121.41,357.90) --
	(123.37,357.91) --
	(125.33,357.92) --
	(127.28,357.93) --
	(129.24,357.94) --
	(131.20,357.94) --
	(133.15,357.95) --
	(135.11,357.95) --
	(137.07,357.94) --
	(139.02,357.92) --
	(140.98,357.88) --
	(142.94,357.84) --
	(144.89,357.80) --
	(146.85,357.75) --
	(148.81,357.70) --
	(150.76,357.65) --
	(152.72,357.59) --
	(154.68,357.53) --
	(156.63,357.46) --
	(158.59,357.39) --
	(160.55,357.32) --
	(162.50,357.24) --
	(164.46,357.16) --
	(166.42,357.07) --
	(168.37,356.98) --
	(170.33,356.88) --
	(172.28,356.78) --
	(174.24,356.68) --
	(176.20,356.57) --
	(178.15,356.46) --
	(180.11,356.34) --
	(182.07,356.22);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,356.74) --
	( 29.46,356.81) --
	( 31.41,356.88) --
	( 33.37,356.94) --
	( 35.32,357.00) --
	( 37.28,357.06) --
	( 39.24,357.12) --
	( 41.19,357.18) --
	( 43.15,357.24) --
	( 45.11,357.29) --
	( 47.06,357.34) --
	( 49.02,357.39) --
	( 50.98,357.44) --
	( 52.93,357.48) --
	( 54.89,357.53) --
	( 56.85,357.57) --
	( 58.80,357.61) --
	( 60.76,357.64) --
	( 62.72,357.68) --
	( 64.67,357.71) --
	( 66.63,357.74) --
	( 68.59,357.77) --
	( 70.54,357.79) --
	( 72.50,357.82) --
	( 74.46,357.84) --
	( 76.41,357.86) --
	( 78.37,357.88) --
	( 80.33,357.89) --
	( 82.28,357.91) --
	( 84.24,357.91) --
	( 86.20,357.92) --
	( 88.15,357.92) --
	( 90.11,357.92) --
	( 92.07,357.92) --
	( 94.02,357.92) --
	( 95.98,357.91) --
	( 97.94,357.92) --
	( 99.89,357.92) --
	(101.85,357.92) --
	(103.80,357.92) --
	(105.76,357.91) --
	(107.72,357.91) --
	(109.67,357.90) --
	(111.63,357.89) --
	(113.59,357.88) --
	(115.54,357.87) --
	(117.50,357.87) --
	(119.46,357.86) --
	(121.41,357.87) --
	(123.37,357.88) --
	(125.33,357.90) --
	(127.28,357.93) --
	(129.24,357.96) --
	(131.20,357.99) --
	(133.15,358.01) --
	(135.11,358.03) --
	(137.07,358.03) --
	(139.02,358.03) --
	(140.98,358.00) --
	(142.94,357.97) --
	(144.89,357.93) --
	(146.85,357.89) --
	(148.81,357.85) --
	(150.76,357.80) --
	(152.72,357.75) --
	(154.68,357.68) --
	(156.63,357.62) --
	(158.59,357.54) --
	(160.55,357.45) --
	(162.50,357.36) --
	(164.46,357.26) --
	(166.42,357.15) --
	(168.37,357.03) --
	(170.33,356.91) --
	(172.28,356.78) --
	(174.24,356.64) --
	(176.20,356.49) --
	(178.15,356.34) --
	(180.11,356.18) --
	(182.07,356.02);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,357.12) --
	( 29.46,357.15) --
	( 31.41,357.19) --
	( 33.37,357.22) --
	( 35.32,357.26) --
	( 37.28,357.29) --
	( 39.24,357.33) --
	( 41.19,357.36) --
	( 43.15,357.40) --
	( 45.11,357.43) --
	( 47.06,357.46) --
	( 49.02,357.50) --
	( 50.98,357.53) --
	( 52.93,357.56) --
	( 54.89,357.59) --
	( 56.85,357.61) --
	( 58.80,357.64) --
	( 60.76,357.67) --
	( 62.72,357.69) --
	( 64.67,357.72) --
	( 66.63,357.75) --
	( 68.59,357.77) --
	( 70.54,357.80) --
	( 72.50,357.82) --
	( 74.46,357.84) --
	( 76.41,357.86) --
	( 78.37,357.88) --
	( 80.33,357.89) --
	( 82.28,357.90) --
	( 84.24,357.91) --
	( 86.20,357.92) --
	( 88.15,357.92) --
	( 90.11,357.93) --
	( 92.07,357.93) --
	( 94.02,357.93) --
	( 95.98,357.93) --
	( 97.94,357.93) --
	( 99.89,357.94) --
	(101.85,357.93) --
	(103.80,357.93) --
	(105.76,357.92) --
	(107.72,357.91) --
	(109.67,357.89) --
	(111.63,357.88) --
	(113.59,357.86) --
	(115.54,357.85) --
	(117.50,357.84) --
	(119.46,357.84) --
	(121.41,357.84) --
	(123.37,357.85) --
	(125.33,357.87) --
	(127.28,357.89) --
	(129.24,357.92) --
	(131.20,357.95) --
	(133.15,357.97) --
	(135.11,357.98) --
	(137.07,357.98) --
	(139.02,357.97) --
	(140.98,357.95) --
	(142.94,357.91) --
	(144.89,357.87) --
	(146.85,357.83) --
	(148.81,357.78) --
	(150.76,357.73) --
	(152.72,357.68) --
	(154.68,357.62) --
	(156.63,357.55) --
	(158.59,357.47) --
	(160.55,357.39) --
	(162.50,357.30) --
	(164.46,357.21) --
	(166.42,357.11) --
	(168.37,357.00) --
	(170.33,356.89) --
	(172.28,356.77) --
	(174.24,356.65) --
	(176.20,356.52) --
	(178.15,356.38) --
	(180.11,356.24) --
	(182.07,356.10);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,357.34) --
	( 29.46,357.36) --
	( 31.41,357.39) --
	( 33.37,357.41) --
	( 35.32,357.43) --
	( 37.28,357.45) --
	( 39.24,357.48) --
	( 41.19,357.50) --
	( 43.15,357.52) --
	( 45.11,357.55) --
	( 47.06,357.57) --
	( 49.02,357.59) --
	( 50.98,357.61) --
	( 52.93,357.63) --
	( 54.89,357.65) --
	( 56.85,357.67) --
	( 58.80,357.69) --
	( 60.76,357.71) --
	( 62.72,357.73) --
	( 64.67,357.75) --
	( 66.63,357.76) --
	( 68.59,357.78) --
	( 70.54,357.80) --
	( 72.50,357.82) --
	( 74.46,357.83) --
	( 76.41,357.85) --
	( 78.37,357.86) --
	( 80.33,357.87) --
	( 82.28,357.88) --
	( 84.24,357.89) --
	( 86.20,357.89) --
	( 88.15,357.90) --
	( 90.11,357.90) --
	( 92.07,357.90) --
	( 94.02,357.90) --
	( 95.98,357.91) --
	( 97.94,357.91) --
	( 99.89,357.91) --
	(101.85,357.91) --
	(103.80,357.91) --
	(105.76,357.90) --
	(107.72,357.90) --
	(109.67,357.89) --
	(111.63,357.88) --
	(113.59,357.87) --
	(115.54,357.86) --
	(117.50,357.85) --
	(119.46,357.85) --
	(121.41,357.85) --
	(123.37,357.85) --
	(125.33,357.86) --
	(127.28,357.87) --
	(129.24,357.87) --
	(131.20,357.88) --
	(133.15,357.88) --
	(135.11,357.89) --
	(137.07,357.89) --
	(139.02,357.89) --
	(140.98,357.89) --
	(142.94,357.88) --
	(144.89,357.87) --
	(146.85,357.87) --
	(148.81,357.86) --
	(150.76,357.85) --
	(152.72,357.85) --
	(154.68,357.84) --
	(156.63,357.83) --
	(158.59,357.82) --
	(160.55,357.80) --
	(162.50,357.79) --
	(164.46,357.77) --
	(166.42,357.75) --
	(168.37,357.74) --
	(170.33,357.72) --
	(172.28,357.70) --
	(174.24,357.67) --
	(176.20,357.65) --
	(178.15,357.63) --
	(180.11,357.60) --
	(182.07,357.58);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,357.81) --
	( 29.46,357.77) --
	( 31.41,357.74) --
	( 33.37,357.71) --
	( 35.32,357.69) --
	( 37.28,357.66) --
	( 39.24,357.64) --
	( 41.19,357.63) --
	( 43.15,357.62) --
	( 45.11,357.61) --
	( 47.06,357.60) --
	( 49.02,357.59) --
	( 50.98,357.59) --
	( 52.93,357.59) --
	( 54.89,357.59) --
	( 56.85,357.59) --
	( 58.80,357.59) --
	( 60.76,357.60) --
	( 62.72,357.62) --
	( 64.67,357.64) --
	( 66.63,357.67) --
	( 68.59,357.70) --
	( 70.54,357.74) --
	( 72.50,357.77) --
	( 74.46,357.80) --
	( 76.41,357.83) --
	( 78.37,357.85) --
	( 80.33,357.86) --
	( 82.28,357.88) --
	( 84.24,357.89) --
	( 86.20,357.89) --
	( 88.15,357.90) --
	( 90.11,357.90) --
	( 92.07,357.91) --
	( 94.02,357.91) --
	( 95.98,357.91) --
	( 97.94,357.92) --
	( 99.89,357.92) --
	(101.85,357.92) --
	(103.80,357.92) --
	(105.76,357.91) --
	(107.72,357.90) --
	(109.67,357.89) --
	(111.63,357.88) --
	(113.59,357.87) --
	(115.54,357.87) --
	(117.50,357.86) --
	(119.46,357.85) --
	(121.41,357.85) --
	(123.37,357.86) --
	(125.33,357.87) --
	(127.28,357.88) --
	(129.24,357.90) --
	(131.20,357.92) --
	(133.15,357.93) --
	(135.11,357.94) --
	(137.07,357.95) --
	(139.02,357.95) --
	(140.98,357.94) --
	(142.94,357.93) --
	(144.89,357.92) --
	(146.85,357.91) --
	(148.81,357.90) --
	(150.76,357.88) --
	(152.72,357.87) --
	(154.68,357.85) --
	(156.63,357.82) --
	(158.59,357.79) --
	(160.55,357.76) --
	(162.50,357.72) --
	(164.46,357.68) --
	(166.42,357.63) --
	(168.37,357.58) --
	(170.33,357.52) --
	(172.28,357.46) --
	(174.24,357.40) --
	(176.20,357.34) --
	(178.15,357.27) --
	(180.11,357.20) --
	(182.07,357.12);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.50,237.50) rectangle (187.22,390.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,234.47) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,278.07) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,321.67) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,365.26) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,237.50) --
	( 27.50,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,281.10) --
	( 27.50,281.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,324.70) --
	( 27.50,324.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,368.29) --
	( 27.50,368.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.50,234.75) --
	( 27.50,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 53.26,234.75) --
	( 53.26,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.02,234.75) --
	( 79.02,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (104.78,234.75) --
	(104.78,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.54,234.75) --
	(130.54,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.31,234.75) --
	(156.31,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (182.07,234.75) --
	(182.07,237.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.50,226.49) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 53.26,226.49) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 79.02,226.49) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (104.78,226.49) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (130.54,226.49) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (156.31,226.49) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (182.07,226.49) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (107.36,214.31) {Primary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,313.80) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.50,398.38) {B. Incubation period};
\end{scope}
\begin{scope}
\path[clip] (192.72,206.49) rectangle (385.44,412.97);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (192.72,206.49) rectangle (385.44,412.97);
\end{scope}
\begin{scope}
\path[clip] (220.22,237.50) rectangle (379.94,390.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.22,237.50) rectangle (379.94,390.09);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,345.69) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,345.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,346.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,345.70) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,343.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,342.54) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,340.69) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,338.91) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,338.45) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,340.23) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,342.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,344.54) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,343.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,343.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,344.37) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,346.69) circle (  2.50);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,345.65) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,345.38) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,346.55) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,345.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,343.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,342.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,340.83) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,338.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,338.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,339.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,342.27) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,344.42) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,344.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,342.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,343.07) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,342.06) circle (  2.50);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,346.03) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,345.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,346.12) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,345.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,343.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,342.66) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,340.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,339.00) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,338.83) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,339.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,342.34) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,344.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,343.44) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,343.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,338.93) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,340.60) circle (  2.50);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,346.05) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,345.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,346.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,345.73) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,343.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,342.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,341.03) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,339.16) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,338.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,339.83) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,342.50) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,344.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,344.19) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,344.00) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,340.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,350.98) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,347.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,345.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,345.51) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,345.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,343.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,342.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,341.14) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,339.03) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,338.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,339.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,341.84) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,344.33) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,346.47) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,344.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,343.58) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,347.56) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,346.12) --
	(222.18,346.10) --
	(224.13,346.08) --
	(226.09,346.04) --
	(228.04,346.00) --
	(230.00,345.94) --
	(231.96,345.87) --
	(233.91,345.78) --
	(235.87,345.68) --
	(237.83,345.57) --
	(239.78,345.45) --
	(241.74,345.33) --
	(243.70,345.19) --
	(245.65,345.05) --
	(247.61,344.91) --
	(249.57,344.76) --
	(251.52,344.60) --
	(253.48,344.43) --
	(255.44,344.23) --
	(257.39,344.01) --
	(259.35,343.78) --
	(261.31,343.53) --
	(263.26,343.28) --
	(265.22,343.03) --
	(267.18,342.78) --
	(269.13,342.54) --
	(271.09,342.32) --
	(273.05,342.09) --
	(275.00,341.83) --
	(276.96,341.53) --
	(278.92,341.22) --
	(280.87,340.89) --
	(282.83,340.57) --
	(284.79,340.28) --
	(286.74,340.01) --
	(288.70,339.79) --
	(290.66,339.63) --
	(292.61,339.54) --
	(294.57,339.51) --
	(296.52,339.53) --
	(298.48,339.59) --
	(300.44,339.69) --
	(302.39,339.80) --
	(304.35,339.94) --
	(306.31,340.09) --
	(308.26,340.24) --
	(310.22,340.38) --
	(312.18,340.52) --
	(314.13,340.65) --
	(316.09,340.81) --
	(318.05,341.00) --
	(320.00,341.22) --
	(321.96,341.45) --
	(323.92,341.70) --
	(325.87,341.94) --
	(327.83,342.18) --
	(329.79,342.41) --
	(331.74,342.61) --
	(333.70,342.79) --
	(335.66,342.95) --
	(337.61,343.12) --
	(339.57,343.29) --
	(341.53,343.46) --
	(343.48,343.63) --
	(345.44,343.80) --
	(347.40,343.97) --
	(349.35,344.13) --
	(351.31,344.30) --
	(353.27,344.45) --
	(355.22,344.61) --
	(357.18,344.76) --
	(359.14,344.90) --
	(361.09,345.05) --
	(363.05,345.19) --
	(365.00,345.34) --
	(366.96,345.48) --
	(368.92,345.61) --
	(370.87,345.75) --
	(372.83,345.88) --
	(374.79,346.02);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,346.05) --
	(222.18,346.03) --
	(224.13,346.02) --
	(226.09,345.99) --
	(228.04,345.95) --
	(230.00,345.90) --
	(231.96,345.83) --
	(233.91,345.76) --
	(235.87,345.67) --
	(237.83,345.57) --
	(239.78,345.46) --
	(241.74,345.34) --
	(243.70,345.21) --
	(245.65,345.08) --
	(247.61,344.94) --
	(249.57,344.80) --
	(251.52,344.65) --
	(253.48,344.48) --
	(255.44,344.29) --
	(257.39,344.08) --
	(259.35,343.86) --
	(261.31,343.62) --
	(263.26,343.38) --
	(265.22,343.14) --
	(267.18,342.90) --
	(269.13,342.67) --
	(271.09,342.45) --
	(273.05,342.23) --
	(275.00,341.96) --
	(276.96,341.66) --
	(278.92,341.34) --
	(280.87,341.01) --
	(282.83,340.69) --
	(284.79,340.38) --
	(286.74,340.11) --
	(288.70,339.88) --
	(290.66,339.71) --
	(292.61,339.62) --
	(294.57,339.58) --
	(296.52,339.60) --
	(298.48,339.65) --
	(300.44,339.74) --
	(302.39,339.85) --
	(304.35,339.98) --
	(306.31,340.12) --
	(308.26,340.27) --
	(310.22,340.40) --
	(312.18,340.53) --
	(314.13,340.65) --
	(316.09,340.82) --
	(318.05,341.04) --
	(320.00,341.28) --
	(321.96,341.54) --
	(323.92,341.82) --
	(325.87,342.08) --
	(327.83,342.33) --
	(329.79,342.54) --
	(331.74,342.71) --
	(333.70,342.83) --
	(335.66,342.91) --
	(337.61,343.00) --
	(339.57,343.08) --
	(341.53,343.16) --
	(343.48,343.22) --
	(345.44,343.28) --
	(347.40,343.33) --
	(349.35,343.36) --
	(351.31,343.37) --
	(353.27,343.36) --
	(355.22,343.33) --
	(357.18,343.29) --
	(359.14,343.23) --
	(361.09,343.16) --
	(363.05,343.08) --
	(365.00,342.98) --
	(366.96,342.87) --
	(368.92,342.75) --
	(370.87,342.61) --
	(372.83,342.46) --
	(374.79,342.30);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,346.34) --
	(222.18,346.29) --
	(224.13,346.24) --
	(226.09,346.18) --
	(228.04,346.11) --
	(230.00,346.03) --
	(231.96,345.94) --
	(233.91,345.84) --
	(235.87,345.72) --
	(237.83,345.60) --
	(239.78,345.47) --
	(241.74,345.34) --
	(243.70,345.20) --
	(245.65,345.05) --
	(247.61,344.90) --
	(249.57,344.74) --
	(251.52,344.58) --
	(253.48,344.41) --
	(255.44,344.21) --
	(257.39,344.00) --
	(259.35,343.78) --
	(261.31,343.55) --
	(263.26,343.31) --
	(265.22,343.07) --
	(267.18,342.83) --
	(269.13,342.61) --
	(271.09,342.39) --
	(273.05,342.17) --
	(275.00,341.92) --
	(276.96,341.63) --
	(278.92,341.32) --
	(280.87,341.00) --
	(282.83,340.69) --
	(284.79,340.40) --
	(286.74,340.13) --
	(288.70,339.92) --
	(290.66,339.76) --
	(292.61,339.66) --
	(294.57,339.64) --
	(296.52,339.67) --
	(298.48,339.74) --
	(300.44,339.84) --
	(302.39,339.97) --
	(304.35,340.12) --
	(306.31,340.27) --
	(308.26,340.42) --
	(310.22,340.57) --
	(312.18,340.69) --
	(314.13,340.80) --
	(316.09,340.96) --
	(318.05,341.16) --
	(320.00,341.39) --
	(321.96,341.64) --
	(323.92,341.89) --
	(325.87,342.13) --
	(327.83,342.35) --
	(329.79,342.52) --
	(331.74,342.65) --
	(333.70,342.70) --
	(335.66,342.72) --
	(337.61,342.73) --
	(339.57,342.74) --
	(341.53,342.73) --
	(343.48,342.71) --
	(345.44,342.68) --
	(347.40,342.63) --
	(349.35,342.56) --
	(351.31,342.47) --
	(353.27,342.35) --
	(355.22,342.22) --
	(357.18,342.06) --
	(359.14,341.88) --
	(361.09,341.69) --
	(363.05,341.48) --
	(365.00,341.25) --
	(366.96,341.00) --
	(368.92,340.74) --
	(370.87,340.45) --
	(372.83,340.15) --
	(374.79,339.83);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,346.38) --
	(222.18,346.33) --
	(224.13,346.28) --
	(226.09,346.22) --
	(228.04,346.14) --
	(230.00,346.06) --
	(231.96,345.97) --
	(233.91,345.86) --
	(235.87,345.75) --
	(237.83,345.63) --
	(239.78,345.50) --
	(241.74,345.37) --
	(243.70,345.22) --
	(245.65,345.08) --
	(247.61,344.93) --
	(249.57,344.77) --
	(251.52,344.62) --
	(253.48,344.44) --
	(255.44,344.25) --
	(257.39,344.04) --
	(259.35,343.82) --
	(261.31,343.59) --
	(263.26,343.36) --
	(265.22,343.13) --
	(267.18,342.89) --
	(269.13,342.67) --
	(271.09,342.46) --
	(273.05,342.24) --
	(275.00,341.99) --
	(276.96,341.70) --
	(278.92,341.40) --
	(280.87,341.08) --
	(282.83,340.78) --
	(284.79,340.49) --
	(286.74,340.23) --
	(288.70,340.02) --
	(290.66,339.86) --
	(292.61,339.76) --
	(294.57,339.73) --
	(296.52,339.75) --
	(298.48,339.81) --
	(300.44,339.89) --
	(302.39,340.01) --
	(304.35,340.13) --
	(306.31,340.27) --
	(308.26,340.41) --
	(310.22,340.55) --
	(312.18,340.67) --
	(314.13,340.78) --
	(316.09,340.91) --
	(318.05,341.05) --
	(320.00,341.21) --
	(321.96,341.38) --
	(323.92,341.56) --
	(325.87,341.74) --
	(327.83,341.93) --
	(329.79,342.11) --
	(331.74,342.28) --
	(333.70,342.45) --
	(335.66,342.61) --
	(337.61,342.78) --
	(339.57,342.95) --
	(341.53,343.12) --
	(343.48,343.31) --
	(345.44,343.50) --
	(347.40,343.70) --
	(349.35,343.90) --
	(351.31,344.12) --
	(353.27,344.34) --
	(355.22,344.58) --
	(357.18,344.82) --
	(359.14,345.07) --
	(361.09,345.33) --
	(363.05,345.59) --
	(365.00,345.86) --
	(366.96,346.14) --
	(368.92,346.43) --
	(370.87,346.72) --
	(372.83,347.02) --
	(374.79,347.33);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,347.01) --
	(222.18,346.89) --
	(224.13,346.78) --
	(226.09,346.65) --
	(228.04,346.52) --
	(230.00,346.39) --
	(231.96,346.25) --
	(233.91,346.10) --
	(235.87,345.94) --
	(237.83,345.78) --
	(239.78,345.62) --
	(241.74,345.45) --
	(243.70,345.28) --
	(245.65,345.10) --
	(247.61,344.92) --
	(249.57,344.74) --
	(251.52,344.56) --
	(253.48,344.37) --
	(255.44,344.17) --
	(257.39,343.96) --
	(259.35,343.74) --
	(261.31,343.52) --
	(263.26,343.29) --
	(265.22,343.07) --
	(267.18,342.85) --
	(269.13,342.63) --
	(271.09,342.42) --
	(273.05,342.20) --
	(275.00,341.94) --
	(276.96,341.65) --
	(278.92,341.33) --
	(280.87,341.01) --
	(282.83,340.69) --
	(284.79,340.38) --
	(286.74,340.11) --
	(288.70,339.89) --
	(290.66,339.72) --
	(292.61,339.62) --
	(294.57,339.58) --
	(296.52,339.57) --
	(298.48,339.59) --
	(300.44,339.65) --
	(302.39,339.73) --
	(304.35,339.82) --
	(306.31,339.94) --
	(308.26,340.06) --
	(310.22,340.20) --
	(312.18,340.33) --
	(314.13,340.48) --
	(316.09,340.68) --
	(318.05,340.92) --
	(320.00,341.19) --
	(321.96,341.48) --
	(323.92,341.78) --
	(325.87,342.09) --
	(327.83,342.38) --
	(329.79,342.65) --
	(331.74,342.90) --
	(333.70,343.10) --
	(335.66,343.28) --
	(337.61,343.47) --
	(339.57,343.66) --
	(341.53,343.86) --
	(343.48,344.05) --
	(345.44,344.24) --
	(347.40,344.43) --
	(349.35,344.62) --
	(351.31,344.80) --
	(353.27,344.97) --
	(355.22,345.14) --
	(357.18,345.30) --
	(359.14,345.45) --
	(361.09,345.60) --
	(363.05,345.75) --
	(365.00,345.89) --
	(366.96,346.03) --
	(368.92,346.17) --
	(370.87,346.30) --
	(372.83,346.43) --
	(374.79,346.55);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.22,237.50) rectangle (379.94,390.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,234.47) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,278.07) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,321.67) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,365.26) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,237.50) --
	(220.22,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,281.10) --
	(220.22,281.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,324.70) --
	(220.22,324.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,368.29) --
	(220.22,368.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.22,234.75) --
	(220.22,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (245.98,234.75) --
	(245.98,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.74,234.75) --
	(271.74,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (297.50,234.75) --
	(297.50,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.26,234.75) --
	(323.26,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (349.03,234.75) --
	(349.03,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (374.79,234.75) --
	(374.79,237.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (220.22,226.49) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.98,226.49) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.74,226.49) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (297.50,226.49) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (323.26,226.49) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (349.03,226.49) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (374.79,226.49) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (300.08,214.31) {Primary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (205.80,313.80) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (220.22,398.38) {C. Generation interval};
\end{scope}
\begin{scope}
\path[clip] (385.44,206.49) rectangle (578.16,412.97);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (385.44,206.49) rectangle (578.16,412.97);
\end{scope}
\begin{scope}
\path[clip] (412.94,237.50) rectangle (572.66,390.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (412.94,237.50) rectangle (572.66,390.09);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94,369.28) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24,362.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55,365.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85,365.00) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16,362.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46,361.48) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77,357.71) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07,351.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,342.52) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,332.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,321.34) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,310.19) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,301.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,297.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,287.48) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,279.29) circle (  2.50);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94,365.64) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24,358.37) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55,361.32) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85,361.46) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16,361.50) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46,359.26) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77,356.35) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07,351.11) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,343.69) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,334.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,325.46) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,316.50) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,309.10) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,302.82) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,299.01) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,295.21) circle (  2.50);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94,363.32) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24,355.97) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55,357.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85,359.54) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16,358.46) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46,357.46) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77,354.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07,350.94) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,345.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,337.45) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,329.93) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,322.82) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,316.14) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,314.22) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,301.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,299.44) circle (  2.50);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94,357.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24,354.25) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55,355.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85,356.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16,355.19) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46,355.24) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77,352.83) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07,350.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,345.94) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,340.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,334.78) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,329.05) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,324.01) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,320.28) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,315.23) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,312.23) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94,354.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24,351.30) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55,352.97) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85,353.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16,352.63) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46,352.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77,351.69) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07,350.19) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,346.97) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,343.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,339.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,336.17) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,332.07) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,329.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,326.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,318.93) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94,365.83) --
	(414.90,366.04) --
	(416.85,366.21) --
	(418.81,366.36) --
	(420.76,366.47) --
	(422.72,366.56) --
	(424.68,366.62) --
	(426.63,366.66) --
	(428.59,366.66) --
	(430.55,366.64) --
	(432.50,366.59) --
	(434.46,366.50) --
	(436.42,366.37) --
	(438.37,366.21) --
	(440.33,366.01) --
	(442.29,365.77) --
	(444.24,365.49) --
	(446.20,365.20) --
	(448.16,364.91) --
	(450.11,364.61) --
	(452.07,364.30) --
	(454.03,363.95) --
	(455.98,363.57) --
	(457.94,363.13) --
	(459.90,362.64) --
	(461.85,362.08) --
	(463.81,361.44) --
	(465.77,360.72) --
	(467.72,359.95) --
	(469.68,359.12) --
	(471.64,358.25) --
	(473.59,357.31) --
	(475.55,356.32) --
	(477.51,355.27) --
	(479.46,354.16) --
	(481.42,352.99) --
	(483.38,351.76) --
	(485.33,350.46) --
	(487.29,349.03) --
	(489.24,347.46) --
	(491.20,345.76) --
	(493.16,343.97) --
	(495.11,342.11) --
	(497.07,340.21) --
	(499.03,338.30) --
	(500.98,336.40) --
	(502.94,334.54) --
	(504.90,332.75) --
	(506.85,331.01) --
	(508.81,329.22) --
	(510.77,327.36) --
	(512.72,325.46) --
	(514.68,323.53) --
	(516.64,321.60) --
	(518.59,319.67) --
	(520.55,317.77) --
	(522.51,315.90) --
	(524.46,314.10) --
	(526.42,312.37) --
	(528.38,310.69) --
	(530.33,309.00) --
	(532.29,307.33) --
	(534.25,305.66) --
	(536.20,304.00) --
	(538.16,302.35) --
	(540.12,300.72) --
	(542.07,299.11) --
	(544.03,297.51) --
	(545.99,295.94) --
	(547.94,294.38) --
	(549.90,292.85) --
	(551.86,291.33) --
	(553.81,289.83) --
	(555.77,288.34) --
	(557.72,286.86) --
	(559.68,285.41) --
	(561.64,283.97) --
	(563.59,282.54) --
	(565.55,281.13) --
	(567.51,279.73);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94,361.86) --
	(414.90,362.08) --
	(416.85,362.26) --
	(418.81,362.42) --
	(420.76,362.55) --
	(422.72,362.67) --
	(424.68,362.76) --
	(426.63,362.83) --
	(428.59,362.89) --
	(430.55,362.91) --
	(432.50,362.91) --
	(434.46,362.88) --
	(436.42,362.82) --
	(438.37,362.73) --
	(440.33,362.61) --
	(442.29,362.45) --
	(444.24,362.25) --
	(446.20,362.06) --
	(448.16,361.88) --
	(450.11,361.69) --
	(452.07,361.49) --
	(454.03,361.27) --
	(455.98,361.02) --
	(457.94,360.72) --
	(459.90,360.37) --
	(461.85,359.95) --
	(463.81,359.46) --
	(465.77,358.89) --
	(467.72,358.26) --
	(469.68,357.59) --
	(471.64,356.86) --
	(473.59,356.07) --
	(475.55,355.24) --
	(477.51,354.35) --
	(479.46,353.41) --
	(481.42,352.42) --
	(483.38,351.38) --
	(485.33,350.29) --
	(487.29,349.09) --
	(489.24,347.77) --
	(491.20,346.34) --
	(493.16,344.83) --
	(495.11,343.27) --
	(497.07,341.67) --
	(499.03,340.05) --
	(500.98,338.44) --
	(502.94,336.87) --
	(504.90,335.34) --
	(506.85,333.86) --
	(508.81,332.29) --
	(510.77,330.66) --
	(512.72,328.97) --
	(514.68,327.25) --
	(516.64,325.52) --
	(518.59,323.81) --
	(520.55,322.14) --
	(522.51,320.53) --
	(524.46,319.00) --
	(526.42,317.58) --
	(528.38,316.22) --
	(530.33,314.87) --
	(532.29,313.53) --
	(534.25,312.20) --
	(536.20,310.90) --
	(538.16,309.62) --
	(540.12,308.37) --
	(542.07,307.15) --
	(544.03,305.96) --
	(545.99,304.82) --
	(547.94,303.71) --
	(549.90,302.63) --
	(551.86,301.58) --
	(553.81,300.56) --
	(555.77,299.57) --
	(557.72,298.61) --
	(559.68,297.68) --
	(561.64,296.77) --
	(563.59,295.90) --
	(565.55,295.05) --
	(567.51,294.23);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94,359.64) --
	(414.90,359.75) --
	(416.85,359.84) --
	(418.81,359.91) --
	(420.76,359.96) --
	(422.72,360.01) --
	(424.68,360.04) --
	(426.63,360.06) --
	(428.59,360.07) --
	(430.55,360.05) --
	(432.50,360.03) --
	(434.46,359.98) --
	(436.42,359.91) --
	(438.37,359.82) --
	(440.33,359.70) --
	(442.29,359.56) --
	(444.24,359.39) --
	(446.20,359.24) --
	(448.16,359.09) --
	(450.11,358.96) --
	(452.07,358.82) --
	(454.03,358.67) --
	(455.98,358.49) --
	(457.94,358.28) --
	(459.90,358.03) --
	(461.85,357.72) --
	(463.81,357.34) --
	(465.77,356.90) --
	(467.72,356.41) --
	(469.68,355.89) --
	(471.64,355.32) --
	(473.59,354.70) --
	(475.55,354.05) --
	(477.51,353.35) --
	(479.46,352.61) --
	(481.42,351.82) --
	(483.38,350.99) --
	(485.33,350.11) --
	(487.29,349.15) --
	(489.24,348.09) --
	(491.20,346.94) --
	(493.16,345.73) --
	(495.11,344.48) --
	(497.07,343.19) --
	(499.03,341.89) --
	(500.98,340.59) --
	(502.94,339.32) --
	(504.90,338.09) --
	(506.85,336.90) --
	(508.81,335.67) --
	(510.77,334.39) --
	(512.72,333.08) --
	(514.68,331.75) --
	(516.64,330.41) --
	(518.59,329.07) --
	(520.55,327.73) --
	(522.51,326.42) --
	(524.46,325.12) --
	(526.42,323.87) --
	(528.38,322.63) --
	(530.33,321.38) --
	(532.29,320.14) --
	(534.25,318.89) --
	(536.20,317.64) --
	(538.16,316.39) --
	(540.12,315.15) --
	(542.07,313.92) --
	(544.03,312.69) --
	(545.99,311.47) --
	(547.94,310.26) --
	(549.90,309.06) --
	(551.86,307.86) --
	(553.81,306.66) --
	(555.77,305.47) --
	(557.72,304.28) --
	(559.68,303.10) --
	(561.64,301.92) --
	(563.59,300.75) --
	(565.55,299.58) --
	(567.51,298.41);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94,355.71) --
	(414.90,355.88) --
	(416.85,356.03) --
	(418.81,356.17) --
	(420.76,356.28) --
	(422.72,356.39) --
	(424.68,356.49) --
	(426.63,356.57) --
	(428.59,356.63) --
	(430.55,356.68) --
	(432.50,356.71) --
	(434.46,356.72) --
	(436.42,356.72) --
	(438.37,356.70) --
	(440.33,356.65) --
	(442.29,356.59) --
	(444.24,356.50) --
	(446.20,356.41) --
	(448.16,356.32) --
	(450.11,356.23) --
	(452.07,356.13) --
	(454.03,356.01) --
	(455.98,355.88) --
	(457.94,355.71) --
	(459.90,355.52) --
	(461.85,355.29) --
	(463.81,355.01) --
	(465.77,354.69) --
	(467.72,354.35) --
	(469.68,353.98) --
	(471.64,353.58) --
	(473.59,353.16) --
	(475.55,352.70) --
	(477.51,352.21) --
	(479.46,351.68) --
	(481.42,351.12) --
	(483.38,350.52) --
	(485.33,349.88) --
	(487.29,349.17) --
	(489.24,348.40) --
	(491.20,347.56) --
	(493.16,346.67) --
	(495.11,345.75) --
	(497.07,344.80) --
	(499.03,343.84) --
	(500.98,342.88) --
	(502.94,341.92) --
	(504.90,340.99) --
	(506.85,340.07) --
	(508.81,339.10) --
	(510.77,338.09) --
	(512.72,337.04) --
	(514.68,335.97) --
	(516.64,334.89) --
	(518.59,333.82) --
	(520.55,332.75) --
	(522.51,331.71) --
	(524.46,330.71) --
	(526.42,329.76) --
	(528.38,328.83) --
	(530.33,327.90) --
	(532.29,326.97) --
	(534.25,326.04) --
	(536.20,325.11) --
	(538.16,324.19) --
	(540.12,323.28) --
	(542.07,322.38) --
	(544.03,321.49) --
	(545.99,320.61) --
	(547.94,319.75) --
	(549.90,318.90) --
	(551.86,318.05) --
	(553.81,317.22) --
	(555.77,316.39) --
	(557.72,315.57) --
	(559.68,314.76) --
	(561.64,313.96) --
	(563.59,313.16) --
	(565.55,312.38) --
	(567.51,311.60);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94,352.93) --
	(414.90,353.03) --
	(416.85,353.12) --
	(418.81,353.20) --
	(420.76,353.27) --
	(422.72,353.33) --
	(424.68,353.39) --
	(426.63,353.44) --
	(428.59,353.48) --
	(430.55,353.52) --
	(432.50,353.54) --
	(434.46,353.55) --
	(436.42,353.56) --
	(438.37,353.55) --
	(440.33,353.53) --
	(442.29,353.50) --
	(444.24,353.45) --
	(446.20,353.41) --
	(448.16,353.37) --
	(450.11,353.33) --
	(452.07,353.29) --
	(454.03,353.24) --
	(455.98,353.18) --
	(457.94,353.10) --
	(459.90,353.00) --
	(461.85,352.88) --
	(463.81,352.72) --
	(465.77,352.54) --
	(467.72,352.33) --
	(469.68,352.12) --
	(471.64,351.88) --
	(473.59,351.63) --
	(475.55,351.35) --
	(477.51,351.05) --
	(479.46,350.73) --
	(481.42,350.39) --
	(483.38,350.01) --
	(485.33,349.61) --
	(487.29,349.16) --
	(489.24,348.67) --
	(491.20,348.14) --
	(493.16,347.57) --
	(495.11,346.98) --
	(497.07,346.37) --
	(499.03,345.76) --
	(500.98,345.13) --
	(502.94,344.51) --
	(504.90,343.90) --
	(506.85,343.30) --
	(508.81,342.68) --
	(510.77,342.05) --
	(512.72,341.39) --
	(514.68,340.73) --
	(516.64,340.06) --
	(518.59,339.38) --
	(520.55,338.69) --
	(522.51,338.00) --
	(524.46,337.32) --
	(526.42,336.63) --
	(528.38,335.94) --
	(530.33,335.25) --
	(532.29,334.55) --
	(534.25,333.84) --
	(536.20,333.12) --
	(538.16,332.39) --
	(540.12,331.66) --
	(542.07,330.91) --
	(544.03,330.16) --
	(545.99,329.40) --
	(547.94,328.62) --
	(549.90,327.84) --
	(551.86,327.05) --
	(553.81,326.25) --
	(555.77,325.44) --
	(557.72,324.63) --
	(559.68,323.80) --
	(561.64,322.97) --
	(563.59,322.13) --
	(565.55,321.28) --
	(567.51,320.42);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (412.94,237.50) rectangle (572.66,390.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,234.47) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,278.07) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,321.67) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,365.26) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,237.50) --
	(412.94,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,281.10) --
	(412.94,281.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,324.70) --
	(412.94,324.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,368.29) --
	(412.94,368.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.94,234.75) --
	(412.94,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (438.70,234.75) --
	(438.70,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.46,234.75) --
	(464.46,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.22,234.75) --
	(490.22,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (515.98,234.75) --
	(515.98,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (541.75,234.75) --
	(541.75,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (567.51,234.75) --
	(567.51,237.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (412.94,226.49) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (438.70,226.49) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (464.46,226.49) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (490.22,226.49) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (515.98,226.49) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (541.75,226.49) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (567.51,226.49) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (492.80,214.31) {Primary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (398.52,313.80) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (412.94,398.38) {D. Serial interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (192.72,206.49);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (192.72,206.49);
\end{scope}
\begin{scope}
\path[clip] ( 27.50, 31.02) rectangle (187.22,183.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.50, 31.02) rectangle (187.22,183.61);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,101.82) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,106.84) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,104.72) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,104.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,105.08) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,105.28) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,106.06) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,107.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,110.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,115.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,122.02) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,130.26) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,138.54) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,144.77) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,150.65) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,156.29) circle (  2.50);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,102.25) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,106.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,104.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,104.51) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,104.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,105.33) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,106.11) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,107.94) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,110.88) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,115.51) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,122.04) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,130.29) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,138.47) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,144.64) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,150.10) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,155.45) circle (  2.50);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,102.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,106.44) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,104.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,104.28) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,105.03) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,105.20) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,106.16) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,107.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,110.88) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,115.42) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,121.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,130.30) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,138.44) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,144.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,150.12) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,156.03) circle (  2.50);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,101.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,106.84) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,105.21) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,103.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,105.04) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,105.24) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,106.15) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,107.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,110.83) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,115.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,121.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,130.18) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,138.51) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,144.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,150.23) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,156.30) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,101.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,107.17) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,105.21) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,104.04) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,105.02) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,105.37) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,106.15) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,107.82) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,110.81) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,115.37) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,121.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,130.04) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,138.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,144.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,150.25) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,155.48) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,104.12) --
	( 29.46,104.08) --
	( 31.41,104.06) --
	( 33.37,104.04) --
	( 35.32,104.03) --
	( 37.28,104.02) --
	( 39.24,104.01) --
	( 41.19,104.00) --
	( 43.15,104.00) --
	( 45.11,104.01) --
	( 47.06,104.02) --
	( 49.02,104.05) --
	( 50.98,104.08) --
	( 52.93,104.12) --
	( 54.89,104.17) --
	( 56.85,104.23) --
	( 58.80,104.31) --
	( 60.76,104.37) --
	( 62.72,104.42) --
	( 64.67,104.46) --
	( 66.63,104.50) --
	( 68.59,104.54) --
	( 70.54,104.60) --
	( 72.50,104.67) --
	( 74.46,104.77) --
	( 76.41,104.90) --
	( 78.37,105.07) --
	( 80.33,105.27) --
	( 82.28,105.45) --
	( 84.24,105.62) --
	( 86.20,105.80) --
	( 88.15,105.99) --
	( 90.11,106.21) --
	( 92.07,106.47) --
	( 94.02,106.77) --
	( 95.98,107.14) --
	( 97.94,107.58) --
	( 99.89,108.10) --
	(101.85,108.68) --
	(103.80,109.32) --
	(105.76,109.99) --
	(107.72,110.71) --
	(109.67,111.47) --
	(111.63,112.28) --
	(113.59,113.12) --
	(115.54,114.00) --
	(117.50,114.93) --
	(119.46,115.88) --
	(121.41,116.89) --
	(123.37,117.99) --
	(125.33,119.17) --
	(127.28,120.41) --
	(129.24,121.70) --
	(131.20,123.03) --
	(133.15,124.37) --
	(135.11,125.71) --
	(137.07,127.04) --
	(139.02,128.34) --
	(140.98,129.58) --
	(142.94,130.82) --
	(144.89,132.09) --
	(146.85,133.37) --
	(148.81,134.67) --
	(150.76,135.99) --
	(152.72,137.32) --
	(154.68,138.65) --
	(156.63,139.99) --
	(158.59,141.33) --
	(160.55,142.67) --
	(162.50,144.00) --
	(164.46,145.34) --
	(166.42,146.68) --
	(168.37,148.03) --
	(170.33,149.39) --
	(172.28,150.75) --
	(174.24,152.12) --
	(176.20,153.50) --
	(178.15,154.88) --
	(180.11,156.26) --
	(182.07,157.66);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,104.38) --
	( 29.46,104.31) --
	( 31.41,104.26) --
	( 33.37,104.22) --
	( 35.32,104.18) --
	( 37.28,104.14) --
	( 39.24,104.11) --
	( 41.19,104.09) --
	( 43.15,104.07) --
	( 45.11,104.06) --
	( 47.06,104.05) --
	( 49.02,104.06) --
	( 50.98,104.08) --
	( 52.93,104.10) --
	( 54.89,104.14) --
	( 56.85,104.20) --
	( 58.80,104.26) --
	( 60.76,104.32) --
	( 62.72,104.37) --
	( 64.67,104.41) --
	( 66.63,104.44) --
	( 68.59,104.49) --
	( 70.54,104.55) --
	( 72.50,104.62) --
	( 74.46,104.73) --
	( 76.41,104.86) --
	( 78.37,105.03) --
	( 80.33,105.23) --
	( 82.28,105.42) --
	( 84.24,105.59) --
	( 86.20,105.78) --
	( 88.15,105.97) --
	( 90.11,106.20) --
	( 92.07,106.46) --
	( 94.02,106.77) --
	( 95.98,107.14) --
	( 97.94,107.58) --
	( 99.89,108.10) --
	(101.85,108.69) --
	(103.80,109.33) --
	(105.76,110.01) --
	(107.72,110.73) --
	(109.67,111.50) --
	(111.63,112.30) --
	(113.59,113.15) --
	(115.54,114.03) --
	(117.50,114.95) --
	(119.46,115.91) --
	(121.41,116.92) --
	(123.37,118.01) --
	(125.33,119.19) --
	(127.28,120.44) --
	(129.24,121.73) --
	(131.20,123.06) --
	(133.15,124.40) --
	(135.11,125.74) --
	(137.07,127.06) --
	(139.02,128.34) --
	(140.98,129.58) --
	(142.94,130.80) --
	(144.89,132.04) --
	(146.85,133.31) --
	(148.81,134.59) --
	(150.76,135.88) --
	(152.72,137.18) --
	(154.68,138.48) --
	(156.63,139.79) --
	(158.59,141.10) --
	(160.55,142.40) --
	(162.50,143.70) --
	(164.46,145.00) --
	(166.42,146.30) --
	(168.37,147.61) --
	(170.33,148.92) --
	(172.28,150.23) --
	(174.24,151.55) --
	(176.20,152.87) --
	(178.15,154.20) --
	(180.11,155.53) --
	(182.07,156.86);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,104.38) --
	( 29.46,104.32) --
	( 31.41,104.27) --
	( 33.37,104.22) --
	( 35.32,104.18) --
	( 37.28,104.15) --
	( 39.24,104.12) --
	( 41.19,104.09) --
	( 43.15,104.08) --
	( 45.11,104.06) --
	( 47.06,104.06) --
	( 49.02,104.07) --
	( 50.98,104.08) --
	( 52.93,104.11) --
	( 54.89,104.15) --
	( 56.85,104.20) --
	( 58.80,104.27) --
	( 60.76,104.33) --
	( 62.72,104.37) --
	( 64.67,104.41) --
	( 66.63,104.45) --
	( 68.59,104.50) --
	( 70.54,104.56) --
	( 72.50,104.63) --
	( 74.46,104.74) --
	( 76.41,104.87) --
	( 78.37,105.04) --
	( 80.33,105.24) --
	( 82.28,105.42) --
	( 84.24,105.60) --
	( 86.20,105.77) --
	( 88.15,105.96) --
	( 90.11,106.18) --
	( 92.07,106.44) --
	( 94.02,106.74) --
	( 95.98,107.11) --
	( 97.94,107.55) --
	( 99.89,108.07) --
	(101.85,108.65) --
	(103.80,109.29) --
	(105.76,109.96) --
	(107.72,110.68) --
	(109.67,111.44) --
	(111.63,112.25) --
	(113.59,113.09) --
	(115.54,113.97) --
	(117.50,114.89) --
	(119.46,115.85) --
	(121.41,116.85) --
	(123.37,117.94) --
	(125.33,119.12) --
	(127.28,120.36) --
	(129.24,121.65) --
	(131.20,122.97) --
	(133.15,124.31) --
	(135.11,125.64) --
	(137.07,126.97) --
	(139.02,128.25) --
	(140.98,129.49) --
	(142.94,130.72) --
	(144.89,131.98) --
	(146.85,133.25) --
	(148.81,134.54) --
	(150.76,135.85) --
	(152.72,137.16) --
	(154.68,138.49) --
	(156.63,139.81) --
	(158.59,141.14) --
	(160.55,142.46) --
	(162.50,143.78) --
	(164.46,145.11) --
	(166.42,146.44) --
	(168.37,147.77) --
	(170.33,149.11) --
	(172.28,150.46) --
	(174.24,151.81) --
	(176.20,153.17) --
	(178.15,154.53) --
	(180.11,155.90) --
	(182.07,157.27);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,104.14) --
	( 29.46,104.10) --
	( 31.41,104.08) --
	( 33.37,104.06) --
	( 35.32,104.04) --
	( 37.28,104.03) --
	( 39.24,104.02) --
	( 41.19,104.01) --
	( 43.15,104.01) --
	( 45.11,104.01) --
	( 47.06,104.02) --
	( 49.02,104.04) --
	( 50.98,104.07) --
	( 52.93,104.11) --
	( 54.89,104.16) --
	( 56.85,104.22) --
	( 58.80,104.29) --
	( 60.76,104.35) --
	( 62.72,104.40) --
	( 64.67,104.44) --
	( 66.63,104.47) --
	( 68.59,104.51) --
	( 70.54,104.56) --
	( 72.50,104.63) --
	( 74.46,104.72) --
	( 76.41,104.85) --
	( 78.37,105.01) --
	( 80.33,105.21) --
	( 82.28,105.39) --
	( 84.24,105.57) --
	( 86.20,105.75) --
	( 88.15,105.94) --
	( 90.11,106.16) --
	( 92.07,106.42) --
	( 94.02,106.73) --
	( 95.98,107.10) --
	( 97.94,107.54) --
	( 99.89,108.06) --
	(101.85,108.65) --
	(103.80,109.28) --
	(105.76,109.95) --
	(107.72,110.66) --
	(109.67,111.41) --
	(111.63,112.21) --
	(113.59,113.05) --
	(115.54,113.92) --
	(117.50,114.84) --
	(119.46,115.79) --
	(121.41,116.80) --
	(123.37,117.89) --
	(125.33,119.07) --
	(127.28,120.31) --
	(129.24,121.60) --
	(131.20,122.92) --
	(133.15,124.26) --
	(135.11,125.61) --
	(137.07,126.93) --
	(139.02,128.23) --
	(140.98,129.47) --
	(142.94,130.71) --
	(144.89,131.97) --
	(146.85,133.25) --
	(148.81,134.55) --
	(150.76,135.87) --
	(152.72,137.19) --
	(154.68,138.52) --
	(156.63,139.86) --
	(158.59,141.20) --
	(160.55,142.54) --
	(162.50,143.87) --
	(164.46,145.21) --
	(166.42,146.56) --
	(168.37,147.91) --
	(170.33,149.26) --
	(172.28,150.63) --
	(174.24,152.00) --
	(176.20,153.37) --
	(178.15,154.76) --
	(180.11,156.14) --
	(182.07,157.54);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,104.16) --
	( 29.46,104.13) --
	( 31.41,104.11) --
	( 33.37,104.10) --
	( 35.32,104.09) --
	( 37.28,104.08) --
	( 39.24,104.08) --
	( 41.19,104.07) --
	( 43.15,104.07) --
	( 45.11,104.08) --
	( 47.06,104.09) --
	( 49.02,104.11) --
	( 50.98,104.14) --
	( 52.93,104.18) --
	( 54.89,104.23) --
	( 56.85,104.29) --
	( 58.80,104.37) --
	( 60.76,104.43) --
	( 62.72,104.47) --
	( 64.67,104.50) --
	( 66.63,104.53) --
	( 68.59,104.56) --
	( 70.54,104.61) --
	( 72.50,104.67) --
	( 74.46,104.76) --
	( 76.41,104.88) --
	( 78.37,105.04) --
	( 80.33,105.23) --
	( 82.28,105.41) --
	( 84.24,105.58) --
	( 86.20,105.76) --
	( 88.15,105.95) --
	( 90.11,106.17) --
	( 92.07,106.42) --
	( 94.02,106.73) --
	( 95.98,107.10) --
	( 97.94,107.53) --
	( 99.89,108.05) --
	(101.85,108.63) --
	(103.80,109.25) --
	(105.76,109.91) --
	(107.72,110.62) --
	(109.67,111.36) --
	(111.63,112.15) --
	(113.59,112.98) --
	(115.54,113.85) --
	(117.50,114.76) --
	(119.46,115.71) --
	(121.41,116.72) --
	(123.37,117.82) --
	(125.33,119.01) --
	(127.28,120.26) --
	(129.24,121.57) --
	(131.20,122.91) --
	(133.15,124.26) --
	(135.11,125.62) --
	(137.07,126.95) --
	(139.02,128.25) --
	(140.98,129.49) --
	(142.94,130.72) --
	(144.89,131.98) --
	(146.85,133.25) --
	(148.81,134.54) --
	(150.76,135.85) --
	(152.72,137.16) --
	(154.68,138.48) --
	(156.63,139.81) --
	(158.59,141.13) --
	(160.55,142.44) --
	(162.50,143.75) --
	(164.46,145.06) --
	(166.42,146.38) --
	(168.37,147.70) --
	(170.33,149.02) --
	(172.28,150.35) --
	(174.24,151.68) --
	(176.20,153.01) --
	(178.15,154.35) --
	(180.11,155.70) --
	(182.07,157.04);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.50, 31.02) rectangle (187.22,183.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55, 27.99) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55, 59.78) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55, 91.57) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,123.36) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,155.14) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75, 31.02) --
	( 27.50, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75, 62.81) --
	( 27.50, 62.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75, 94.60) --
	( 27.50, 94.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,126.39) --
	( 27.50,126.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,158.18) --
	( 27.50,158.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.50, 28.27) --
	( 27.50, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 53.26, 28.27) --
	( 53.26, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.02, 28.27) --
	( 79.02, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (104.78, 28.27) --
	(104.78, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.54, 28.27) --
	(130.54, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.31, 28.27) --
	(156.31, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (182.07, 28.27) --
	(182.07, 31.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.50, 20.01) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 53.26, 20.01) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 79.02, 20.01) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (104.78, 20.01) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (130.54, 20.01) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (156.31, 20.01) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (182.07, 20.01) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (107.36,  7.82) {Secondary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,107.31) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.50,191.89) {E. Incubation period};
\end{scope}
\begin{scope}
\path[clip] (192.72,  0.00) rectangle (385.44,206.49);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (192.72,  0.00) rectangle (385.44,206.49);
\end{scope}
\begin{scope}
\path[clip] (220.22, 31.02) rectangle (379.94,183.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.22, 31.02) rectangle (379.94,183.61);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,101.30) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,101.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,100.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,101.12) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,101.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,101.58) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,101.97) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,103.26) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,105.38) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,108.65) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,113.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,118.88) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,123.70) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,126.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,127.48) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,125.68) circle (  2.50);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,101.73) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,101.93) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,100.46) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,101.05) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,101.63) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,101.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,101.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,103.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,105.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,108.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,113.65) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,118.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,123.55) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,126.05) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,128.36) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,123.16) circle (  2.50);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,101.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,102.22) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,100.56) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,100.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,101.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,101.45) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,101.92) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,103.46) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,105.36) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,108.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,113.49) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,119.03) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,123.70) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,126.40) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,127.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,127.28) circle (  2.50);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,101.08) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,102.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,100.83) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,100.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,101.71) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,101.37) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,101.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,103.47) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,105.55) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,108.81) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,113.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,118.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,123.22) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,127.21) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,127.27) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,128.95) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,101.01) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,102.91) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,101.10) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,100.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,101.45) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,101.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,101.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,103.49) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,105.54) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,108.52) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,113.38) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,118.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,123.35) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,127.24) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,126.77) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,128.39) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,101.69) --
	(222.18,101.56) --
	(224.13,101.44) --
	(226.09,101.33) --
	(228.04,101.23) --
	(230.00,101.13) --
	(231.96,101.05) --
	(233.91,100.97) --
	(235.87,100.90) --
	(237.83,100.84) --
	(239.78,100.79) --
	(241.74,100.75) --
	(243.70,100.72) --
	(245.65,100.71) --
	(247.61,100.70) --
	(249.57,100.71) --
	(251.52,100.73) --
	(253.48,100.75) --
	(255.44,100.78) --
	(257.39,100.81) --
	(259.35,100.85) --
	(261.31,100.90) --
	(263.26,100.97) --
	(265.22,101.04) --
	(267.18,101.13) --
	(269.13,101.24) --
	(271.09,101.38) --
	(273.05,101.52) --
	(275.00,101.65) --
	(276.96,101.77) --
	(278.92,101.90) --
	(280.87,102.03) --
	(282.83,102.18) --
	(284.79,102.36) --
	(286.74,102.57) --
	(288.70,102.82) --
	(290.66,103.12) --
	(292.61,103.49) --
	(294.57,103.91) --
	(296.52,104.37) --
	(298.48,104.88) --
	(300.44,105.42) --
	(302.39,106.00) --
	(304.35,106.60) --
	(306.31,107.22) --
	(308.26,107.86) --
	(310.22,108.50) --
	(312.18,109.16) --
	(314.13,109.83) --
	(316.09,110.59) --
	(318.05,111.43) --
	(320.00,112.32) --
	(321.96,113.24) --
	(323.92,114.18) --
	(325.87,115.11) --
	(327.83,116.01) --
	(329.79,116.85) --
	(331.74,117.63) --
	(333.70,118.31) --
	(335.66,118.95) --
	(337.61,119.59) --
	(339.57,120.22) --
	(341.53,120.85) --
	(343.48,121.47) --
	(345.44,122.06) --
	(347.40,122.64) --
	(349.35,123.19) --
	(351.31,123.71) --
	(353.27,124.20) --
	(355.22,124.65) --
	(357.18,125.07) --
	(359.14,125.48) --
	(361.09,125.86) --
	(363.05,126.22) --
	(365.00,126.55) --
	(366.96,126.86) --
	(368.92,127.15) --
	(370.87,127.41) --
	(372.83,127.65) --
	(374.79,127.87);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,102.09) --
	(222.18,101.92) --
	(224.13,101.77) --
	(226.09,101.62) --
	(228.04,101.49) --
	(230.00,101.37) --
	(231.96,101.26) --
	(233.91,101.16) --
	(235.87,101.07) --
	(237.83,100.99) --
	(239.78,100.92) --
	(241.74,100.86) --
	(243.70,100.81) --
	(245.65,100.78) --
	(247.61,100.76) --
	(249.57,100.76) --
	(251.52,100.76) --
	(253.48,100.78) --
	(255.44,100.80) --
	(257.39,100.82) --
	(259.35,100.86) --
	(261.31,100.91) --
	(263.26,100.97) --
	(265.22,101.04) --
	(267.18,101.13) --
	(269.13,101.24) --
	(271.09,101.37) --
	(273.05,101.52) --
	(275.00,101.66) --
	(276.96,101.79) --
	(278.92,101.92) --
	(280.87,102.06) --
	(282.83,102.22) --
	(284.79,102.41) --
	(286.74,102.63) --
	(288.70,102.89) --
	(290.66,103.20) --
	(292.61,103.57) --
	(294.57,103.99) --
	(296.52,104.46) --
	(298.48,104.97) --
	(300.44,105.51) --
	(302.39,106.08) --
	(304.35,106.68) --
	(306.31,107.30) --
	(308.26,107.94) --
	(310.22,108.58) --
	(312.18,109.23) --
	(314.13,109.90) --
	(316.09,110.66) --
	(318.05,111.50) --
	(320.00,112.40) --
	(321.96,113.33) --
	(323.92,114.27) --
	(325.87,115.20) --
	(327.83,116.10) --
	(329.79,116.94) --
	(331.74,117.70) --
	(333.70,118.36) --
	(335.66,118.97) --
	(337.61,119.58) --
	(339.57,120.18) --
	(341.53,120.77) --
	(343.48,121.35) --
	(345.44,121.90) --
	(347.40,122.43) --
	(349.35,122.93) --
	(351.31,123.39) --
	(353.27,123.81) --
	(355.22,124.19) --
	(357.18,124.54) --
	(359.14,124.86) --
	(361.09,125.15) --
	(363.05,125.41) --
	(365.00,125.65) --
	(366.96,125.86) --
	(368.92,126.03) --
	(370.87,126.19) --
	(372.83,126.31) --
	(374.79,126.41);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,102.24) --
	(222.18,102.06) --
	(224.13,101.90) --
	(226.09,101.75) --
	(228.04,101.61) --
	(230.00,101.48) --
	(231.96,101.36) --
	(233.91,101.25) --
	(235.87,101.15) --
	(237.83,101.06) --
	(239.78,100.99) --
	(241.74,100.92) --
	(243.70,100.87) --
	(245.65,100.83) --
	(247.61,100.80) --
	(249.57,100.79) --
	(251.52,100.79) --
	(253.48,100.80) --
	(255.44,100.81) --
	(257.39,100.83) --
	(259.35,100.86) --
	(261.31,100.90) --
	(263.26,100.95) --
	(265.22,101.02) --
	(267.18,101.10) --
	(269.13,101.21) --
	(271.09,101.34) --
	(273.05,101.48) --
	(275.00,101.61) --
	(276.96,101.74) --
	(278.92,101.87) --
	(280.87,102.01) --
	(282.83,102.17) --
	(284.79,102.36) --
	(286.74,102.58) --
	(288.70,102.84) --
	(290.66,103.15) --
	(292.61,103.52) --
	(294.57,103.94) --
	(296.52,104.41) --
	(298.48,104.92) --
	(300.44,105.46) --
	(302.39,106.04) --
	(304.35,106.64) --
	(306.31,107.27) --
	(308.26,107.91) --
	(310.22,108.56) --
	(312.18,109.21) --
	(314.13,109.88) --
	(316.09,110.63) --
	(318.05,111.45) --
	(320.00,112.32) --
	(321.96,113.22) --
	(323.92,114.13) --
	(325.87,115.04) --
	(327.83,115.92) --
	(329.79,116.76) --
	(331.74,117.53) --
	(333.70,118.23) --
	(335.66,118.89) --
	(337.61,119.55) --
	(339.57,120.20) --
	(341.53,120.86) --
	(343.48,121.50) --
	(345.44,122.13) --
	(347.40,122.75) --
	(349.35,123.34) --
	(351.31,123.91) --
	(353.27,124.45) --
	(355.22,124.96) --
	(357.18,125.45) --
	(359.14,125.92) --
	(361.09,126.38) --
	(363.05,126.81) --
	(365.00,127.23) --
	(366.96,127.63) --
	(368.92,128.00) --
	(370.87,128.37) --
	(372.83,128.71) --
	(374.79,129.03);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,102.10) --
	(222.18,101.94) --
	(224.13,101.80) --
	(226.09,101.67) --
	(228.04,101.55) --
	(230.00,101.44) --
	(231.96,101.34) --
	(233.91,101.24) --
	(235.87,101.15) --
	(237.83,101.08) --
	(239.78,101.01) --
	(241.74,100.95) --
	(243.70,100.91) --
	(245.65,100.88) --
	(247.61,100.86) --
	(249.57,100.85) --
	(251.52,100.85) --
	(253.48,100.86) --
	(255.44,100.87) --
	(257.39,100.88) --
	(259.35,100.90) --
	(261.31,100.92) --
	(263.26,100.96) --
	(265.22,101.02) --
	(267.18,101.09) --
	(269.13,101.19) --
	(271.09,101.31) --
	(273.05,101.45) --
	(275.00,101.58) --
	(276.96,101.72) --
	(278.92,101.86) --
	(280.87,102.02) --
	(282.83,102.19) --
	(284.79,102.39) --
	(286.74,102.62) --
	(288.70,102.90) --
	(290.66,103.21) --
	(292.61,103.58) --
	(294.57,104.00) --
	(296.52,104.47) --
	(298.48,104.98) --
	(300.44,105.52) --
	(302.39,106.09) --
	(304.35,106.69) --
	(306.31,107.30) --
	(308.26,107.94) --
	(310.22,108.58) --
	(312.18,109.23) --
	(314.13,109.90) --
	(316.09,110.64) --
	(318.05,111.44) --
	(320.00,112.29) --
	(321.96,113.17) --
	(323.92,114.06) --
	(325.87,114.95) --
	(327.83,115.82) --
	(329.79,116.65) --
	(331.74,117.43) --
	(333.70,118.14) --
	(335.66,118.81) --
	(337.61,119.49) --
	(339.57,120.16) --
	(341.53,120.84) --
	(343.48,121.51) --
	(345.44,122.16) --
	(347.40,122.81) --
	(349.35,123.44) --
	(351.31,124.05) --
	(353.27,124.64) --
	(355.22,125.20) --
	(357.18,125.75) --
	(359.14,126.29) --
	(361.09,126.81) --
	(363.05,127.32) --
	(365.00,127.81) --
	(366.96,128.29) --
	(368.92,128.75) --
	(370.87,129.20) --
	(372.83,129.64) --
	(374.79,130.06);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,102.10) --
	(222.18,101.95) --
	(224.13,101.82) --
	(226.09,101.69) --
	(228.04,101.57) --
	(230.00,101.46) --
	(231.96,101.36) --
	(233.91,101.27) --
	(235.87,101.18) --
	(237.83,101.10) --
	(239.78,101.04) --
	(241.74,100.98) --
	(243.70,100.94) --
	(245.65,100.91) --
	(247.61,100.89) --
	(249.57,100.88) --
	(251.52,100.88) --
	(253.48,100.89) --
	(255.44,100.89) --
	(257.39,100.90) --
	(259.35,100.91) --
	(261.31,100.94) --
	(263.26,100.97) --
	(265.22,101.02) --
	(267.18,101.09) --
	(269.13,101.18) --
	(271.09,101.30) --
	(273.05,101.43) --
	(275.00,101.57) --
	(276.96,101.70) --
	(278.92,101.84) --
	(280.87,101.99) --
	(282.83,102.16) --
	(284.79,102.36) --
	(286.74,102.59) --
	(288.70,102.86) --
	(290.66,103.18) --
	(292.61,103.54) --
	(294.57,103.95) --
	(296.52,104.41) --
	(298.48,104.90) --
	(300.44,105.43) --
	(302.39,105.98) --
	(304.35,106.57) --
	(306.31,107.17) --
	(308.26,107.79) --
	(310.22,108.42) --
	(312.18,109.07) --
	(314.13,109.73) --
	(316.09,110.47) --
	(318.05,111.27) --
	(320.00,112.13) --
	(321.96,113.01) --
	(323.92,113.91) --
	(325.87,114.80) --
	(327.83,115.67) --
	(329.79,116.50) --
	(331.74,117.28) --
	(333.70,117.98) --
	(335.66,118.64) --
	(337.61,119.31) --
	(339.57,119.98) --
	(341.53,120.65) --
	(343.48,121.31) --
	(345.44,121.96) --
	(347.40,122.59) --
	(349.35,123.21) --
	(351.31,123.81) --
	(353.27,124.38) --
	(355.22,124.93) --
	(357.18,125.47) --
	(359.14,125.98) --
	(361.09,126.49) --
	(363.05,126.97) --
	(365.00,127.45) --
	(366.96,127.90) --
	(368.92,128.34) --
	(370.87,128.77) --
	(372.83,129.18) --
	(374.79,129.58);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.22, 31.02) rectangle (379.94,183.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27, 27.99) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27, 59.78) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27, 91.57) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,123.36) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,155.14) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47, 31.02) --
	(220.22, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47, 62.81) --
	(220.22, 62.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47, 94.60) --
	(220.22, 94.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,126.39) --
	(220.22,126.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,158.18) --
	(220.22,158.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.22, 28.27) --
	(220.22, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (245.98, 28.27) --
	(245.98, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.74, 28.27) --
	(271.74, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (297.50, 28.27) --
	(297.50, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.26, 28.27) --
	(323.26, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (349.03, 28.27) --
	(349.03, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (374.79, 28.27) --
	(374.79, 31.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (220.22, 20.01) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.98, 20.01) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.74, 20.01) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (297.50, 20.01) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (323.26, 20.01) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (349.03, 20.01) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (374.79, 20.01) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (300.08,  7.82) {Secondary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (205.80,107.31) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (220.22,191.89) {F. Generation interval};
\end{scope}
\begin{scope}
\path[clip] (385.44,  0.00) rectangle (578.16,206.49);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (385.44,  0.00) rectangle (578.16,206.49);
\end{scope}
\begin{scope}
\path[clip] (412.94, 31.02) rectangle (572.66,183.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (412.94, 31.02) rectangle (572.66,183.61);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94, 76.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24, 86.83) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55, 85.55) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85, 86.31) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16, 86.44) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46, 88.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77, 88.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07, 91.38) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38, 94.91) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,101.20) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,110.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,122.63) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,135.35) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,145.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,153.70) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,162.89) circle (  2.50);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94, 76.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24, 87.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55, 88.68) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85, 89.22) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16, 89.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46, 90.63) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77, 91.04) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07, 93.34) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38, 96.81) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,102.70) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,111.21) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,122.45) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,133.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,143.31) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,151.06) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,157.71) circle (  2.50);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94, 80.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24, 92.10) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55, 91.50) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85, 90.92) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16, 92.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46, 92.32) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77, 93.28) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07, 95.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38, 98.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,103.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,111.72) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,122.19) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,133.10) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,141.14) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,148.25) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,156.21) circle (  2.50);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94, 85.03) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24, 94.32) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55, 94.35) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85, 92.82) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16, 94.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46, 94.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77, 95.66) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07, 97.29) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,100.66) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,105.31) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,112.94) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,121.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,131.34) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,138.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,145.17) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,151.07) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94, 88.35) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24, 97.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55, 96.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85, 95.29) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16, 96.55) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46, 97.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77, 97.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07, 99.36) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,102.45) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,106.99) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,113.52) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,121.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,130.25) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,136.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,141.69) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,147.05) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94, 80.73) --
	(414.90, 81.08) --
	(416.85, 81.42) --
	(418.81, 81.77) --
	(420.76, 82.11) --
	(422.72, 82.43) --
	(424.68, 82.74) --
	(426.63, 83.04) --
	(428.59, 83.33) --
	(430.55, 83.62) --
	(432.50, 83.90) --
	(434.46, 84.18) --
	(436.42, 84.45) --
	(438.37, 84.73) --
	(440.33, 85.01) --
	(442.29, 85.29) --
	(444.24, 85.58) --
	(446.20, 85.84) --
	(448.16, 86.05) --
	(450.11, 86.23) --
	(452.07, 86.38) --
	(454.03, 86.53) --
	(455.98, 86.69) --
	(457.94, 86.85) --
	(459.90, 87.05) --
	(461.85, 87.28) --
	(463.81, 87.56) --
	(465.77, 87.87) --
	(467.72, 88.13) --
	(469.68, 88.36) --
	(471.64, 88.56) --
	(473.59, 88.78) --
	(475.55, 89.03) --
	(477.51, 89.32) --
	(479.46, 89.69) --
	(481.42, 90.15) --
	(483.38, 90.73) --
	(485.33, 91.44) --
	(487.29, 92.24) --
	(489.24, 93.08) --
	(491.20, 93.96) --
	(493.16, 94.90) --
	(495.11, 95.90) --
	(497.07, 96.95) --
	(499.03, 98.08) --
	(500.98, 99.27) --
	(502.94,100.53) --
	(504.90,101.88) --
	(506.85,103.32) --
	(508.81,104.91) --
	(510.77,106.62) --
	(512.72,108.43) --
	(514.68,110.31) --
	(516.64,112.26) --
	(518.59,114.23) --
	(520.55,116.21) --
	(522.51,118.17) --
	(524.46,120.09) --
	(526.42,121.95) --
	(528.38,123.80) --
	(530.33,125.70) --
	(532.29,127.63) --
	(534.25,129.59) --
	(536.20,131.58) --
	(538.16,133.59) --
	(540.12,135.62) --
	(542.07,137.66) --
	(544.03,139.71) --
	(545.99,141.76) --
	(547.94,143.81) --
	(549.90,145.87) --
	(551.86,147.95) --
	(553.81,150.04) --
	(555.77,152.14) --
	(557.72,154.26) --
	(559.68,156.39) --
	(561.64,158.54) --
	(563.59,160.70) --
	(565.55,162.87) --
	(567.51,165.06);
\definecolor{drawColor}{RGB}{59,82,139}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94, 80.69) --
	(414.90, 81.29) --
	(416.85, 81.89) --
	(418.81, 82.47) --
	(420.76, 83.03) --
	(422.72, 83.56) --
	(424.68, 84.07) --
	(426.63, 84.55) --
	(428.59, 85.01) --
	(430.55, 85.45) --
	(432.50, 85.87) --
	(434.46, 86.29) --
	(436.42, 86.68) --
	(438.37, 87.07) --
	(440.33, 87.46) --
	(442.29, 87.83) --
	(444.24, 88.21) --
	(446.20, 88.53) --
	(448.16, 88.79) --
	(450.11, 89.00) --
	(452.07, 89.17) --
	(454.03, 89.32) --
	(455.98, 89.47) --
	(457.94, 89.62) --
	(459.90, 89.79) --
	(461.85, 90.00) --
	(463.81, 90.26) --
	(465.77, 90.55) --
	(467.72, 90.77) --
	(469.68, 90.95) --
	(471.64, 91.10) --
	(473.59, 91.25) --
	(475.55, 91.43) --
	(477.51, 91.65) --
	(479.46, 91.94) --
	(481.42, 92.33) --
	(483.38, 92.84) --
	(485.33, 93.48) --
	(487.29, 94.22) --
	(489.24, 95.00) --
	(491.20, 95.82) --
	(493.16, 96.71) --
	(495.11, 97.64) --
	(497.07, 98.63) --
	(499.03, 99.69) --
	(500.98,100.80) --
	(502.94,101.98) --
	(504.90,103.23) --
	(506.85,104.56) --
	(508.81,106.02) --
	(510.77,107.61) --
	(512.72,109.29) --
	(514.68,111.04) --
	(516.64,112.84) --
	(518.59,114.66) --
	(520.55,116.49) --
	(522.51,118.29) --
	(524.46,120.05) --
	(526.42,121.75) --
	(528.38,123.43) --
	(530.33,125.15) --
	(532.29,126.90) --
	(534.25,128.68) --
	(536.20,130.48) --
	(538.16,132.30) --
	(540.12,134.12) --
	(542.07,135.96) --
	(544.03,137.79) --
	(545.99,139.63) --
	(547.94,141.45) --
	(549.90,143.29) --
	(551.86,145.13) --
	(553.81,146.98) --
	(555.77,148.84) --
	(557.72,150.71) --
	(559.68,152.59) --
	(561.64,154.48) --
	(563.59,156.38) --
	(565.55,158.29) --
	(567.51,160.20);
\definecolor{drawColor}{RGB}{33,144,140}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94, 85.11) --
	(414.90, 85.56) --
	(416.85, 86.01) --
	(418.81, 86.45) --
	(420.76, 86.87) --
	(422.72, 87.27) --
	(424.68, 87.65) --
	(426.63, 88.01) --
	(428.59, 88.36) --
	(430.55, 88.69) --
	(432.50, 89.01) --
	(434.46, 89.32) --
	(436.42, 89.62) --
	(438.37, 89.92) --
	(440.33, 90.22) --
	(442.29, 90.51) --
	(444.24, 90.80) --
	(446.20, 91.05) --
	(448.16, 91.24) --
	(450.11, 91.39) --
	(452.07, 91.51) --
	(454.03, 91.61) --
	(455.98, 91.71) --
	(457.94, 91.82) --
	(459.90, 91.96) --
	(461.85, 92.13) --
	(463.81, 92.35) --
	(465.77, 92.60) --
	(467.72, 92.81) --
	(469.68, 92.99) --
	(471.64, 93.14) --
	(473.59, 93.31) --
	(475.55, 93.50) --
	(477.51, 93.73) --
	(479.46, 94.03) --
	(481.42, 94.41) --
	(483.38, 94.89) --
	(485.33, 95.49) --
	(487.29, 96.17) --
	(489.24, 96.90) --
	(491.20, 97.67) --
	(493.16, 98.49) --
	(495.11, 99.36) --
	(497.07,100.28) --
	(499.03,101.25) --
	(500.98,102.29) --
	(502.94,103.38) --
	(504.90,104.53) --
	(506.85,105.76) --
	(508.81,107.10) --
	(510.77,108.55) --
	(512.72,110.08) --
	(514.68,111.67) --
	(516.64,113.31) --
	(518.59,114.97) --
	(520.55,116.64) --
	(522.51,118.29) --
	(524.46,119.92) --
	(526.42,121.49) --
	(528.38,123.05) --
	(530.33,124.65) --
	(532.29,126.27) --
	(534.25,127.93) --
	(536.20,129.61) --
	(538.16,131.30) --
	(540.12,133.01) --
	(542.07,134.73) --
	(544.03,136.46) --
	(545.99,138.19) --
	(547.94,139.92) --
	(549.90,141.66) --
	(551.86,143.41) --
	(553.81,145.17) --
	(555.77,146.95) --
	(557.72,148.74) --
	(559.68,150.54) --
	(561.64,152.35) --
	(563.59,154.17) --
	(565.55,156.01) --
	(567.51,157.86);
\definecolor{drawColor}{RGB}{93,200,99}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94, 88.68) --
	(414.90, 89.03) --
	(416.85, 89.38) --
	(418.81, 89.72) --
	(420.76, 90.06) --
	(422.72, 90.37) --
	(424.68, 90.67) --
	(426.63, 90.96) --
	(428.59, 91.23) --
	(430.55, 91.50) --
	(432.50, 91.76) --
	(434.46, 92.01) --
	(436.42, 92.26) --
	(438.37, 92.50) --
	(440.33, 92.75) --
	(442.29, 92.99) --
	(444.24, 93.24) --
	(446.20, 93.45) --
	(448.16, 93.61) --
	(450.11, 93.74) --
	(452.07, 93.85) --
	(454.03, 93.94) --
	(455.98, 94.03) --
	(457.94, 94.14) --
	(459.90, 94.26) --
	(461.85, 94.42) --
	(463.81, 94.63) --
	(465.77, 94.86) --
	(467.72, 95.06) --
	(469.68, 95.23) --
	(471.64, 95.40) --
	(473.59, 95.57) --
	(475.55, 95.76) --
	(477.51, 95.99) --
	(479.46, 96.28) --
	(481.42, 96.64) --
	(483.38, 97.10) --
	(485.33, 97.66) --
	(487.29, 98.29) --
	(489.24, 98.96) --
	(491.20, 99.67) --
	(493.16,100.43) --
	(495.11,101.23) --
	(497.07,102.08) --
	(499.03,102.98) --
	(500.98,103.92) --
	(502.94,104.92) --
	(504.90,105.96) --
	(506.85,107.07) --
	(508.81,108.29) --
	(510.77,109.61) --
	(512.72,111.00) --
	(514.68,112.45) --
	(516.64,113.94) --
	(518.59,115.45) --
	(520.55,116.96) --
	(522.51,118.45) --
	(524.46,119.90) --
	(526.42,121.30) --
	(528.38,122.69) --
	(530.33,124.11) --
	(532.29,125.55) --
	(534.25,127.01) --
	(536.20,128.49) --
	(538.16,129.99) --
	(540.12,131.49) --
	(542.07,132.99) --
	(544.03,134.50) --
	(545.99,136.01) --
	(547.94,137.51) --
	(549.90,139.01) --
	(551.86,140.52) --
	(553.81,142.04) --
	(555.77,143.57) --
	(557.72,145.10) --
	(559.68,146.64) --
	(561.64,148.19) --
	(563.59,149.74) --
	(565.55,151.30) --
	(567.51,152.87);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94, 91.91) --
	(414.90, 92.20) --
	(416.85, 92.48) --
	(418.81, 92.77) --
	(420.76, 93.04) --
	(422.72, 93.30) --
	(424.68, 93.55) --
	(426.63, 93.78) --
	(428.59, 94.01) --
	(430.55, 94.23) --
	(432.50, 94.44) --
	(434.46, 94.65) --
	(436.42, 94.86) --
	(438.37, 95.06) --
	(440.33, 95.27) --
	(442.29, 95.48) --
	(444.24, 95.69) --
	(446.20, 95.86) --
	(448.16, 96.00) --
	(450.11, 96.10) --
	(452.07, 96.18) --
	(454.03, 96.26) --
	(455.98, 96.33) --
	(457.94, 96.42) --
	(459.90, 96.52) --
	(461.85, 96.66) --
	(463.81, 96.84) --
	(465.77, 97.06) --
	(467.72, 97.25) --
	(469.68, 97.41) --
	(471.64, 97.58) --
	(473.59, 97.75) --
	(475.55, 97.94) --
	(477.51, 98.17) --
	(479.46, 98.45) --
	(481.42, 98.80) --
	(483.38, 99.22) --
	(485.33, 99.74) --
	(487.29,100.32) --
	(489.24,100.94) --
	(491.20,101.59) --
	(493.16,102.29) --
	(495.11,103.03) --
	(497.07,103.81) --
	(499.03,104.63) --
	(500.98,105.50) --
	(502.94,106.40) --
	(504.90,107.36) --
	(506.85,108.36) --
	(508.81,109.47) --
	(510.77,110.66) --
	(512.72,111.92) --
	(514.68,113.24) --
	(516.64,114.59) --
	(518.59,115.95) --
	(520.55,117.31) --
	(522.51,118.65) --
	(524.46,119.95) --
	(526.42,121.20) --
	(528.38,122.43) --
	(530.33,123.68) --
	(532.29,124.95) --
	(534.25,126.24) --
	(536.20,127.54) --
	(538.16,128.85) --
	(540.12,130.16) --
	(542.07,131.48) --
	(544.03,132.79) --
	(545.99,134.10) --
	(547.94,135.40) --
	(549.90,136.70) --
	(551.86,138.00) --
	(553.81,139.31) --
	(555.77,140.62) --
	(557.72,141.94) --
	(559.68,143.26) --
	(561.64,144.58) --
	(563.59,145.90) --
	(565.55,147.23) --
	(567.51,148.56);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (412.94, 31.02) rectangle (572.66,183.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99, 27.99) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99, 59.78) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99, 91.57) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,123.36) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,155.14) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19, 31.02) --
	(412.94, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19, 62.81) --
	(412.94, 62.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19, 94.60) --
	(412.94, 94.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,126.39) --
	(412.94,126.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,158.18) --
	(412.94,158.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.94, 28.27) --
	(412.94, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (438.70, 28.27) --
	(438.70, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.46, 28.27) --
	(464.46, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.22, 28.27) --
	(490.22, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (515.98, 28.27) --
	(515.98, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (541.75, 28.27) --
	(541.75, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (567.51, 28.27) --
	(567.51, 31.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (412.94, 20.01) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (438.70, 20.01) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (464.46, 20.01) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (490.22, 20.01) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (515.98, 20.01) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (541.75, 20.01) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (567.51, 20.01) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (492.80,  7.82) {Secondary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (398.52,107.31) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (412.94,191.89) {G. Serial interval};
\end{scope}
\end{tikzpicture}

\end{document}