### 2020-10-05 rR.tex
so close

% Created by tikzDevice version 0.12.3 on 2020-10-04 14:51:35
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,433.62);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,247.50) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,294.20) --
	( 27.31,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,340.91) --
	( 27.31,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,387.61) --
	( 27.31,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,329.23) {Basic reproduction number $\mathcal R_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.86,244.47) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.86,291.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.86,337.88) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.86,384.58) {4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (317.06,247.50) --
	(319.81,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (317.06,294.20) --
	(319.81,294.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (317.06,340.91) --
	(319.81,340.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (317.06,387.61) --
	(319.81,387.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (305.58,329.23) {Reproduction number ${\mathcal R}_{\textrm{\tiny intrinsic}}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36, 60.52) {5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.36,170.08) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56, 63.55) --
	( 27.31, 63.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.56,173.11) --
	( 27.31,173.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,112.42) {Mean interval (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.86, 27.66) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.86, 66.58) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.86,105.50) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.86,144.42) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.86,183.34) {0.8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (317.06, 30.69) --
	(319.81, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (317.06, 69.61) --
	(319.81, 69.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (317.06,108.53) --
	(319.81,108.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (317.06,147.45) --
	(319.81,147.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (317.06,186.37) --
	(319.81,186.37);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (298.73,112.42) {Squared coefficient of variation};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,419.03) {A};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (319.81,419.03) {B};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.31,202.22) {C};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (319.81,202.22) {D};
\end{scope}
\begin{scope}
\path[clip] ( 27.31,247.50) rectangle (280.16,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31,247.50) rectangle (280.16,410.96);
\definecolor{drawColor}{RGB}{13,8,135}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31,247.50) --
	( 29.98,248.23) --
	( 32.65,248.97) --
	( 35.31,249.73) --
	( 37.98,250.50) --
	( 40.65,251.28) --
	( 43.32,252.08) --
	( 45.98,252.88) --
	( 48.65,253.70) --
	( 51.32,254.53) --
	( 53.98,255.37) --
	( 56.65,256.23) --
	( 59.32,257.09) --
	( 61.99,257.97) --
	( 64.65,258.86) --
	( 67.32,259.76) --
	( 69.99,260.67) --
	( 72.65,261.60) --
	( 75.32,262.53) --
	( 77.99,263.48) --
	( 80.66,264.43) --
	( 83.32,265.40) --
	( 85.99,266.38) --
	( 88.66,267.37) --
	( 91.32,268.38) --
	( 93.99,269.40) --
	( 96.66,270.43) --
	( 99.33,271.47) --
	(101.99,272.54) --
	(104.66,273.61) --
	(107.33,274.70) --
	(109.99,275.81) --
	(112.66,276.93) --
	(115.33,278.07) --
	(118.00,279.22) --
	(120.66,280.38) --
	(123.33,281.55) --
	(126.00,282.74) --
	(128.66,283.95) --
	(131.33,285.16) --
	(134.00,286.40) --
	(136.67,287.65) --
	(139.33,288.91) --
	(142.00,290.20) --
	(144.67,291.50) --
	(147.33,292.81) --
	(150.00,294.15) --
	(152.67,295.51) --
	(155.34,296.88) --
	(158.00,298.27) --
	(160.67,299.67) --
	(163.34,301.09) --
	(166.00,302.53) --
	(168.67,303.98) --
	(171.34,305.45) --
	(174.01,306.94) --
	(176.67,308.45) --
	(179.34,309.97) --
	(182.01,311.51) --
	(184.67,313.08) --
	(187.34,314.67) --
	(190.01,316.27) --
	(192.68,317.90) --
	(195.34,319.55) --
	(198.01,321.23) --
	(200.68,322.92) --
	(203.34,324.64) --
	(206.01,326.37) --
	(208.68,328.13) --
	(211.35,329.90) --
	(214.01,331.70) --
	(216.68,333.51) --
	(219.35,335.35) --
	(222.01,337.21) --
	(224.68,339.08) --
	(227.35,340.98) --
	(230.02,342.90) --
	(232.68,344.83) --
	(235.35,346.79) --
	(238.02,348.77);
\definecolor{drawColor}{RGB}{126,3,168}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 61.03,257.69) --
	( 63.69,258.56) --
	( 66.36,259.45) --
	( 69.03,260.35) --
	( 71.69,261.26) --
	( 74.36,262.19) --
	( 77.03,263.13) --
	( 79.70,264.08) --
	( 82.36,265.05) --
	( 85.03,266.03) --
	( 87.70,267.03) --
	( 90.36,268.04) --
	( 93.03,269.06) --
	( 95.70,270.09) --
	( 98.37,271.14) --
	(101.03,272.21) --
	(103.70,273.28) --
	(106.37,274.37) --
	(109.03,275.47) --
	(111.70,276.58) --
	(114.37,277.71) --
	(117.04,278.84) --
	(119.70,279.99) --
	(122.37,281.16) --
	(125.04,282.34) --
	(127.70,283.53) --
	(130.37,284.74) --
	(133.04,285.97) --
	(135.71,287.21) --
	(138.37,288.47) --
	(141.04,289.75) --
	(143.71,291.04) --
	(146.37,292.36) --
	(149.04,293.69) --
	(151.71,295.03) --
	(154.38,296.39) --
	(157.04,297.76) --
	(159.71,299.15) --
	(162.38,300.55) --
	(165.04,301.97) --
	(167.71,303.41) --
	(170.38,304.87) --
	(173.05,306.34) --
	(175.71,307.84) --
	(178.38,309.35) --
	(181.05,310.89) --
	(183.71,312.45) --
	(186.38,314.03) --
	(189.05,315.62) --
	(191.72,317.24) --
	(194.38,318.87) --
	(197.05,320.53) --
	(199.72,322.19) --
	(202.38,323.88) --
	(205.05,325.59) --
	(207.72,327.32) --
	(210.39,329.07) --
	(213.05,330.84) --
	(215.72,332.63) --
	(218.39,334.45) --
	(221.05,336.29) --
	(223.72,338.15) --
	(226.39,340.04) --
	(229.06,341.96) --
	(231.72,343.90) --
	(234.39,345.87) --
	(237.06,347.86) --
	(239.72,349.87) --
	(242.39,351.90) --
	(245.06,353.96) --
	(247.73,356.04) --
	(250.39,358.15) --
	(253.06,360.28) --
	(255.73,362.43) --
	(258.39,364.60) --
	(261.06,366.80) --
	(263.73,369.02) --
	(266.40,371.27) --
	(269.06,373.53) --
	(271.73,375.82);
\definecolor{drawColor}{RGB}{204,70,120}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 52.60,254.98) --
	( 55.26,255.81) --
	( 57.93,256.66) --
	( 60.60,257.53) --
	( 63.27,258.41) --
	( 65.93,259.30) --
	( 68.60,260.20) --
	( 71.27,261.12) --
	( 73.93,262.04) --
	( 76.60,262.99) --
	( 79.27,263.94) --
	( 81.94,264.91) --
	( 84.60,265.89) --
	( 87.27,266.89) --
	( 89.94,267.90) --
	( 92.60,268.92) --
	( 95.27,269.95) --
	( 97.94,271.00) --
	(100.61,272.05) --
	(103.27,273.12) --
	(105.94,274.20) --
	(108.61,275.30) --
	(111.27,276.40) --
	(113.94,277.52) --
	(116.61,278.66) --
	(119.28,279.81) --
	(121.94,280.97) --
	(124.61,282.15) --
	(127.28,283.34) --
	(129.94,284.55) --
	(132.61,285.78) --
	(135.28,287.03) --
	(137.95,288.29) --
	(140.61,289.56) --
	(143.28,290.85) --
	(145.95,292.16) --
	(148.62,293.48) --
	(151.28,294.81) --
	(153.95,296.16) --
	(156.62,297.52) --
	(159.28,298.90) --
	(161.95,300.30) --
	(164.62,301.72) --
	(167.29,303.16) --
	(169.95,304.61) --
	(172.62,306.09) --
	(175.29,307.58) --
	(177.95,309.10) --
	(180.62,310.64) --
	(183.29,312.20) --
	(185.96,313.77) --
	(188.62,315.36) --
	(191.29,316.96) --
	(193.96,318.59) --
	(196.62,320.23) --
	(199.29,321.89) --
	(201.96,323.58) --
	(204.63,325.28) --
	(207.29,327.01) --
	(209.96,328.76) --
	(212.63,330.54) --
	(215.29,332.33) --
	(217.96,334.16) --
	(220.63,336.01) --
	(223.30,337.88) --
	(225.96,339.78) --
	(228.63,341.70) --
	(231.30,343.64) --
	(233.96,345.60) --
	(236.63,347.59) --
	(239.30,349.60) --
	(241.97,351.63) --
	(244.63,353.69) --
	(247.30,355.76) --
	(249.97,357.86) --
	(252.63,359.99) --
	(255.30,362.13) --
	(257.97,364.30) --
	(260.64,366.49) --
	(263.30,368.71);
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 44.17,252.37) --
	( 46.84,253.17) --
	( 49.50,253.99) --
	( 52.17,254.82) --
	( 54.84,255.66) --
	( 57.51,256.51) --
	( 60.17,257.37) --
	( 62.84,258.25) --
	( 65.51,259.14) --
	( 68.17,260.05) --
	( 70.84,260.97) --
	( 73.51,261.90) --
	( 76.18,262.84) --
	( 78.84,263.79) --
	( 81.51,264.76) --
	( 84.18,265.74) --
	( 86.84,266.74) --
	( 89.51,267.74) --
	( 92.18,268.76) --
	( 94.85,269.79) --
	( 97.51,270.83) --
	(100.18,271.88) --
	(102.85,272.95) --
	(105.51,274.03) --
	(108.18,275.12) --
	(110.85,276.23) --
	(113.52,277.35) --
	(116.18,278.49) --
	(118.85,279.64) --
	(121.52,280.80) --
	(124.18,281.98) --
	(126.85,283.18) --
	(129.52,284.39) --
	(132.19,285.61) --
	(134.85,286.85) --
	(137.52,288.09) --
	(140.19,289.35) --
	(142.85,290.63) --
	(145.52,291.92) --
	(148.19,293.22) --
	(150.86,294.54) --
	(153.52,295.88) --
	(156.19,297.23) --
	(158.86,298.61) --
	(161.52,300.00) --
	(164.19,301.41) --
	(166.86,302.85) --
	(169.53,304.30) --
	(172.19,305.78) --
	(174.86,307.28) --
	(177.53,308.79) --
	(180.19,310.32) --
	(182.86,311.86) --
	(185.53,313.42) --
	(188.20,315.01) --
	(190.86,316.61) --
	(193.53,318.23) --
	(196.20,319.88) --
	(198.86,321.54) --
	(201.53,323.23) --
	(204.20,324.94) --
	(206.87,326.68) --
	(209.53,328.43) --
	(212.20,330.22) --
	(214.87,332.02) --
	(217.53,333.85) --
	(220.20,335.71) --
	(222.87,337.58) --
	(225.54,339.48) --
	(228.20,341.40) --
	(230.87,343.34) --
	(233.54,345.30) --
	(236.20,347.28) --
	(238.87,349.29) --
	(241.54,351.32) --
	(244.21,353.37) --
	(246.87,355.45) --
	(249.54,357.54) --
	(252.21,359.66) --
	(254.87,361.80);
\definecolor{drawColor}{RGB}{216,224,30}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 35.74,249.88) --
	( 38.41,250.65) --
	( 41.08,251.43) --
	( 43.74,252.22) --
	( 46.41,253.03) --
	( 49.08,253.84) --
	( 51.74,254.67) --
	( 54.41,255.51) --
	( 57.08,256.37) --
	( 59.75,257.24) --
	( 62.41,258.11) --
	( 65.08,259.01) --
	( 67.75,259.91) --
	( 70.41,260.83) --
	( 73.08,261.76) --
	( 75.75,262.70) --
	( 78.42,263.65) --
	( 81.08,264.62) --
	( 83.75,265.60) --
	( 86.42,266.59) --
	( 89.08,267.59) --
	( 91.75,268.60) --
	( 94.42,269.63) --
	( 97.09,270.67) --
	( 99.75,271.72) --
	(102.42,272.79) --
	(105.09,273.87) --
	(107.75,274.96) --
	(110.42,276.07) --
	(113.09,277.19) --
	(115.76,278.32) --
	(118.42,279.47) --
	(121.09,280.64) --
	(123.76,281.81) --
	(126.42,283.00) --
	(129.09,284.20) --
	(131.76,285.41) --
	(134.43,286.63) --
	(137.09,287.87) --
	(139.76,289.12) --
	(142.43,290.39) --
	(145.09,291.67) --
	(147.76,292.97) --
	(150.43,294.29) --
	(153.10,295.63) --
	(155.76,296.98) --
	(158.43,298.36) --
	(161.10,299.76) --
	(163.76,301.18) --
	(166.43,302.62) --
	(169.10,304.07) --
	(171.77,305.54) --
	(174.43,307.02) --
	(177.10,308.53) --
	(179.77,310.05) --
	(182.43,311.59) --
	(185.10,313.15) --
	(187.77,314.73) --
	(190.44,316.33) --
	(193.10,317.95) --
	(195.77,319.60) --
	(198.44,321.27) --
	(201.10,322.96) --
	(203.77,324.67) --
	(206.44,326.41) --
	(209.11,328.16) --
	(211.77,329.94) --
	(214.44,331.74) --
	(217.11,333.57) --
	(219.77,335.41) --
	(222.44,337.27) --
	(225.11,339.16) --
	(227.78,341.07) --
	(230.44,342.99) --
	(233.11,344.94) --
	(235.78,346.91) --
	(238.44,348.91) --
	(241.11,350.92) --
	(243.78,352.95) --
	(246.45,355.01);
\definecolor{fillColor}{RGB}{126,3,168}

\path[fill=fillColor] ( 61.03,261.57) --
	( 64.39,255.75) --
	( 57.66,255.75) --
	cycle;

\path[fill=fillColor] (103.17,276.95) --
	(106.53,271.12) --
	( 99.80,271.12) --
	cycle;

\path[fill=fillColor] (145.31,295.71) --
	(148.67,289.89) --
	(141.94,289.89) --
	cycle;

\path[fill=fillColor] (187.45,318.55) --
	(190.81,312.72) --
	(184.08,312.72) --
	cycle;

\path[fill=fillColor] (229.59,346.23) --
	(232.95,340.41) --
	(226.23,340.41) --
	cycle;

\path[fill=fillColor] (271.73,379.71) --
	(275.09,373.88) --
	(268.37,373.88) --
	cycle;
\definecolor{fillColor}{RGB}{204,70,120}

\path[fill=fillColor] ( 50.10,252.48) --
	( 55.10,252.48) --
	( 55.10,257.47) --
	( 50.10,257.47) --
	cycle;

\path[fill=fillColor] ( 92.24,267.25) --
	( 97.24,267.25) --
	( 97.24,272.24) --
	( 92.24,272.24) --
	cycle;

\path[fill=fillColor] (134.38,285.28) --
	(139.38,285.28) --
	(139.38,290.28) --
	(134.38,290.28) --
	cycle;

\path[fill=fillColor] (176.52,307.22) --
	(181.52,307.22) --
	(181.52,312.21) --
	(176.52,312.21) --
	cycle;

\path[fill=fillColor] (218.66,333.88) --
	(223.66,333.88) --
	(223.66,338.88) --
	(218.66,338.88) --
	cycle;

\path[fill=fillColor] (260.80,366.21) --
	(265.80,366.21) --
	(265.80,371.20) --
	(260.80,371.20) --
	cycle;
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.64,252.37) -- ( 47.70,252.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.17,248.84) -- ( 44.17,255.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 82.78,266.54) -- ( 89.84,266.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.31,263.01) -- ( 86.31,270.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (124.92,283.90) -- (131.98,283.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.45,280.37) -- (128.45,287.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (167.06,304.89) -- (174.12,304.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.59,301.36) -- (170.59,308.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.20,330.58) -- (216.27,330.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.73,327.04) -- (212.73,334.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.34,361.80) -- (258.41,361.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.87,358.27) -- (254.87,365.33);
\definecolor{drawColor}{RGB}{216,224,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.24,247.39) rectangle ( 38.24,252.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.24,247.39) -- ( 38.24,252.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.24,252.38) -- ( 38.24,247.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.38,260.97) rectangle ( 80.38,265.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.38,260.97) -- ( 80.38,265.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.38,265.96) -- ( 80.38,260.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.53,277.67) rectangle (122.52,282.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.53,277.67) -- (122.52,282.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.53,282.67) -- (122.52,277.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.67,297.83) rectangle (164.66,302.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.67,297.83) -- (164.66,302.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.67,302.82) -- (164.66,297.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (201.81,322.52) rectangle (206.80,327.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (201.81,322.52) -- (206.80,327.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (201.81,327.51) -- (206.80,322.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.95,352.51) rectangle (248.94,357.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.95,352.51) -- (248.94,357.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.95,357.51) -- (248.94,352.51);
\definecolor{fillColor}{RGB}{13,8,135}

\path[fill=fillColor] ( 27.31,247.50) circle (  2.50);

\path[fill=fillColor] ( 69.45,260.49) circle (  2.50);

\path[fill=fillColor] (111.60,276.48) circle (  2.50);

\path[fill=fillColor] (153.74,296.05) circle (  2.50);

\path[fill=fillColor] (195.88,319.89) circle (  2.50);

\path[fill=fillColor] (238.02,348.77) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31,247.50) rectangle (280.16,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 36.66,389.03) rectangle ( 51.11,403.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{13,8,135}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 38.10,396.25) -- ( 49.66,396.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{13,8,135}

\path[fill=fillColor] ( 43.88,396.25) circle (  2.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 36.66,374.57) rectangle ( 51.11,389.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{126,3,168}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 38.10,381.80) -- ( 49.66,381.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{126,3,168}

\path[fill=fillColor] ( 43.88,385.68) --
	( 47.25,379.86) --
	( 40.52,379.86) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 36.66,360.12) rectangle ( 51.11,374.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{204,70,120}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 38.10,367.35) -- ( 49.66,367.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{204,70,120}

\path[fill=fillColor] ( 41.38,364.85) --
	( 46.38,364.85) --
	( 46.38,369.84) --
	( 41.38,369.84) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 36.66,345.67) rectangle ( 51.11,360.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 38.10,352.89) -- ( 49.66,352.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.35,352.89) -- ( 47.41,352.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 43.88,349.36) -- ( 43.88,356.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 36.66,331.21) rectangle ( 51.11,345.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{216,224,30}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 38.10,338.44) -- ( 49.66,338.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{216,224,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.38,335.94) rectangle ( 46.38,340.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.38,335.94) -- ( 46.38,340.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.38,340.94) -- ( 46.38,335.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 56.61,393.22) {Intrinsic generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 56.61,378.77) {Forward serial interval ($\rho=0.75$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 56.61,364.32) {Forward serial interval ($\rho=0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 56.61,349.86) {Forward serial interval ($\rho=0.25$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 56.61,335.41) {Forward serial interval ($\rho=0$)};
\end{scope}
\begin{scope}
\path[clip] (319.81,247.50) rectangle (572.66,410.96);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (319.81,247.50) rectangle (572.66,410.96);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (319.81,247.64) --
	(323.02,248.49) --
	(326.22,249.36) --
	(329.42,250.25) --
	(332.62,251.16) --
	(335.82,252.09) --
	(339.02,253.03) --
	(342.22,254.00) --
	(345.42,254.99) --
	(348.62,255.99) --
	(351.82,257.01) --
	(355.02,258.06) --
	(358.22,259.12) --
	(361.42,260.20) --
	(364.62,261.30) --
	(367.82,262.42) --
	(371.02,263.56) --
	(374.22,264.72) --
	(377.43,265.90) --
	(380.63,267.10) --
	(383.83,268.32) --
	(387.03,269.55) --
	(390.23,270.81) --
	(393.43,272.08) --
	(396.63,273.37) --
	(399.83,274.68) --
	(403.03,276.01) --
	(406.23,277.36) --
	(409.43,278.74) --
	(412.63,280.13) --
	(415.83,281.55) --
	(419.03,282.99) --
	(422.23,284.44) --
	(425.43,285.91) --
	(428.63,287.40) --
	(431.83,288.91) --
	(435.04,290.45) --
	(438.24,292.02) --
	(441.44,293.61) --
	(444.64,295.23) --
	(447.84,296.88) --
	(451.04,298.55) --
	(454.24,300.24) --
	(457.44,301.95) --
	(460.64,303.68) --
	(463.84,305.44) --
	(467.04,307.23) --
	(470.24,309.04) --
	(473.44,310.89) --
	(476.64,312.76) --
	(479.84,314.68) --
	(483.04,316.62) --
	(486.24,318.59) --
	(489.45,320.58) --
	(492.65,322.60) --
	(495.85,324.66) --
	(499.05,326.74) --
	(502.25,328.85) --
	(505.45,330.99) --
	(508.65,333.17) --
	(511.85,335.39) --
	(515.05,337.63) --
	(518.25,339.91) --
	(521.45,342.22) --
	(524.65,344.56) --
	(527.85,346.93) --
	(531.05,349.33) --
	(534.25,351.76) --
	(537.45,354.22) --
	(540.65,356.72) --
	(543.85,359.24) --
	(547.06,361.80) --
	(550.26,364.38) --
	(553.46,367.00) --
	(556.66,369.66) --
	(559.86,372.34) --
	(563.06,375.05) --
	(566.26,377.80) --
	(569.46,380.57) --
	(572.66,383.38);
\definecolor{drawColor}{RGB}{109,109,109}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (323.02,248.37) --
	(326.22,249.29) --
	(329.42,250.21) --
	(332.62,251.14) --
	(335.82,252.09) --
	(339.02,253.04) --
	(342.22,254.01) --
	(345.42,254.98) --
	(348.62,255.97) --
	(351.82,256.97) --
	(355.02,257.98) --
	(358.22,259.00) --
	(361.42,260.04) --
	(364.62,261.08) --
	(367.82,262.13) --
	(371.02,263.20) --
	(374.22,264.27) --
	(377.43,265.36) --
	(380.63,266.46) --
	(383.83,267.57) --
	(387.03,268.69) --
	(390.23,269.82) --
	(393.43,270.96) --
	(396.63,272.12) --
	(399.83,273.28) --
	(403.03,274.45) --
	(406.23,275.64) --
	(409.43,276.84) --
	(412.63,278.04) --
	(415.83,279.26) --
	(419.03,280.49) --
	(422.23,281.73) --
	(425.43,282.98) --
	(428.63,284.25) --
	(431.83,285.52) --
	(435.04,286.81) --
	(438.24,288.10) --
	(441.44,289.40) --
	(444.64,290.71) --
	(447.84,292.02) --
	(451.04,293.35) --
	(454.24,294.69) --
	(457.44,296.04) --
	(460.64,297.40) --
	(463.84,298.77) --
	(467.04,300.15) --
	(470.24,301.54) --
	(473.44,302.93) --
	(476.64,304.32) --
	(479.84,305.72) --
	(483.04,307.12) --
	(486.24,308.54) --
	(489.45,309.96) --
	(492.65,311.39) --
	(495.85,312.82) --
	(499.05,314.26) --
	(502.25,315.71) --
	(505.45,317.16) --
	(508.65,318.61) --
	(511.85,320.07) --
	(515.05,321.53) --
	(518.25,322.99) --
	(521.45,324.45) --
	(524.65,325.93) --
	(527.85,327.40) --
	(531.05,328.89) --
	(534.25,330.37) --
	(537.45,331.86) --
	(540.65,333.36) --
	(543.85,334.86) --
	(547.06,336.37) --
	(550.26,337.88) --
	(553.46,339.39) --
	(556.66,340.91) --
	(559.86,342.43) --
	(563.06,343.96) --
	(566.26,345.49) --
	(569.46,347.03) --
	(572.66,348.57);
\definecolor{drawColor}{RGB}{149,149,149}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (323.02,248.30) --
	(326.22,249.24) --
	(329.42,250.18) --
	(332.62,251.12) --
	(335.82,252.08) --
	(339.02,253.04) --
	(342.22,254.00) --
	(345.42,254.98) --
	(348.62,255.96) --
	(351.82,256.94) --
	(355.02,257.93) --
	(358.22,258.93) --
	(361.42,259.94) --
	(364.62,260.95) --
	(367.82,261.97) --
	(371.02,262.99) --
	(374.22,264.02) --
	(377.43,265.05) --
	(380.63,266.09) --
	(383.83,267.14) --
	(387.03,268.20) --
	(390.23,269.26) --
	(393.43,270.34) --
	(396.63,271.42) --
	(399.83,272.50) --
	(403.03,273.59) --
	(406.23,274.68) --
	(409.43,275.78) --
	(412.63,276.87) --
	(415.83,277.98) --
	(419.03,279.10) --
	(422.23,280.22) --
	(425.43,281.36) --
	(428.63,282.50) --
	(431.83,283.64) --
	(435.04,284.78) --
	(438.24,285.92) --
	(441.44,287.06) --
	(444.64,288.19) --
	(447.84,289.31) --
	(451.04,290.45) --
	(454.24,291.59) --
	(457.44,292.74) --
	(460.64,293.89) --
	(463.84,295.05) --
	(467.04,296.19) --
	(470.24,297.34) --
	(473.44,298.47) --
	(476.64,299.59) --
	(479.84,300.69) --
	(483.04,301.79) --
	(486.24,302.89) --
	(489.45,303.98) --
	(492.65,305.07) --
	(495.85,306.15) --
	(499.05,307.23) --
	(502.25,308.29) --
	(505.45,309.34) --
	(508.65,310.37) --
	(511.85,311.39) --
	(515.05,312.39) --
	(518.25,313.37) --
	(521.45,314.35) --
	(524.65,315.32) --
	(527.85,316.28) --
	(531.05,317.22) --
	(534.25,318.16) --
	(537.45,319.09) --
	(540.65,320.00) --
	(543.85,320.91) --
	(547.06,321.80) --
	(550.26,322.67) --
	(553.46,323.54) --
	(556.66,324.40) --
	(559.86,325.24) --
	(563.06,326.07) --
	(566.26,326.89) --
	(569.46,327.70) --
	(572.66,328.50);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (323.02,248.24) --
	(326.22,249.19) --
	(329.42,250.15) --
	(332.62,251.11) --
	(335.82,252.07) --
	(339.02,253.03) --
	(342.22,254.00) --
	(345.42,254.96) --
	(348.62,255.93) --
	(351.82,256.91) --
	(355.02,257.88) --
	(358.22,258.86) --
	(361.42,259.84) --
	(364.62,260.82) --
	(367.82,261.80) --
	(371.02,262.78) --
	(374.22,263.77) --
	(377.43,264.75) --
	(380.63,265.74) --
	(383.83,266.73) --
	(387.03,267.73) --
	(390.23,268.73) --
	(393.43,269.73) --
	(396.63,270.74) --
	(399.83,271.75) --
	(403.03,272.76) --
	(406.23,273.76) --
	(409.43,274.76) --
	(412.63,275.76) --
	(415.83,276.76) --
	(419.03,277.77) --
	(422.23,278.79) --
	(425.43,279.81) --
	(428.63,280.83) --
	(431.83,281.84) --
	(435.04,282.85) --
	(438.24,283.85) --
	(441.44,284.83) --
	(444.64,285.80) --
	(447.84,286.75) --
	(451.04,287.70) --
	(454.24,288.67) --
	(457.44,289.63) --
	(460.64,290.59) --
	(463.84,291.54) --
	(467.04,292.47) --
	(470.24,293.39) --
	(473.44,294.29) --
	(476.64,295.16) --
	(479.84,295.99) --
	(483.04,296.81) --
	(486.24,297.61) --
	(489.45,298.41) --
	(492.65,299.19) --
	(495.85,299.95) --
	(499.05,300.69) --
	(502.25,301.40) --
	(505.45,302.09) --
	(508.65,302.75) --
	(511.85,303.37) --
	(515.05,303.96) --
	(518.25,304.52) --
	(521.45,305.06) --
	(524.65,305.58) --
	(527.85,306.08) --
	(531.05,306.56) --
	(534.25,307.01) --
	(537.45,307.43) --
	(540.65,307.83) --
	(543.85,308.21) --
	(547.06,308.56) --
	(550.26,308.88) --
	(553.46,309.18) --
	(556.66,309.45) --
	(559.86,309.69) --
	(563.06,309.91) --
	(566.26,310.11) --
	(569.46,310.28) --
	(572.66,310.43);
\definecolor{drawColor}{gray}{0.80}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] (323.02,248.18) --
	(326.22,249.15) --
	(329.42,250.12) --
	(332.62,251.09) --
	(335.82,252.06) --
	(339.02,253.02) --
	(342.22,253.99) --
	(345.42,254.95) --
	(348.62,255.91) --
	(351.82,256.87) --
	(355.02,257.82) --
	(358.22,258.78) --
	(361.42,259.73) --
	(364.62,260.68) --
	(367.82,261.63) --
	(371.02,262.58) --
	(374.22,263.52) --
	(377.43,264.45) --
	(380.63,265.39) --
	(383.83,266.33) --
	(387.03,267.27) --
	(390.23,268.21) --
	(393.43,269.15) --
	(396.63,270.09) --
	(399.83,271.03) --
	(403.03,271.96) --
	(406.23,272.88) --
	(409.43,273.79) --
	(412.63,274.70) --
	(415.83,275.60) --
	(419.03,276.51) --
	(422.23,277.43) --
	(425.43,278.34) --
	(428.63,279.25) --
	(431.83,280.15) --
	(435.04,281.03) --
	(438.24,281.90) --
	(441.44,282.74) --
	(444.64,283.56) --
	(447.84,284.35) --
	(451.04,285.14) --
	(454.24,285.93) --
	(457.44,286.72) --
	(460.64,287.50) --
	(463.84,288.26) --
	(467.04,289.01) --
	(470.24,289.72) --
	(473.44,290.40) --
	(476.64,291.04) --
	(479.84,291.64) --
	(483.04,292.21) --
	(486.24,292.76) --
	(489.45,293.29) --
	(492.65,293.79) --
	(495.85,294.27) --
	(499.05,294.72) --
	(502.25,295.13) --
	(505.45,295.51) --
	(508.65,295.84) --
	(511.85,296.12) --
	(515.05,296.36) --
	(518.25,296.57) --
	(521.45,296.74) --
	(524.65,296.88) --
	(527.85,296.99) --
	(531.05,297.06) --
	(534.25,297.10) --
	(537.45,297.11) --
	(540.65,297.08) --
	(543.85,297.02) --
	(547.06,296.91) --
	(550.26,296.77) --
	(553.46,296.60) --
	(556.66,296.39) --
	(559.86,296.14) --
	(563.06,295.86) --
	(566.26,295.54) --
	(569.46,295.19) --
	(572.66,294.80);
\definecolor{fillColor}{RGB}{109,109,109}

\path[fill=fillColor] (319.81,251.38) --
	(323.18,245.55) --
	(316.45,245.55) --
	cycle;

\path[fill=fillColor] (328.24,253.76) --
	(331.61,247.94) --
	(324.88,247.94) --
	cycle;

\path[fill=fillColor] (336.67,256.22) --
	(340.03,250.40) --
	(333.31,250.40) --
	cycle;

\path[fill=fillColor] (345.10,258.76) --
	(348.46,252.93) --
	(341.74,252.93) --
	cycle;

\path[fill=fillColor] (353.53,261.38) --
	(356.89,255.55) --
	(350.16,255.55) --
	cycle;

\path[fill=fillColor] (361.96,264.07) --
	(365.32,258.25) --
	(358.59,258.25) --
	cycle;

\path[fill=fillColor] (370.38,266.85) --
	(373.75,261.02) --
	(367.02,261.02) --
	cycle;

\path[fill=fillColor] (378.81,269.70) --
	(382.18,263.87) --
	(375.45,263.87) --
	cycle;

\path[fill=fillColor] (387.24,272.63) --
	(390.60,266.80) --
	(383.88,266.80) --
	cycle;

\path[fill=fillColor] (395.67,275.64) --
	(399.03,269.81) --
	(392.30,269.81) --
	cycle;

\path[fill=fillColor] (404.10,278.72) --
	(407.46,272.89) --
	(400.73,272.89) --
	cycle;

\path[fill=fillColor] (412.52,281.88) --
	(415.89,276.05) --
	(409.16,276.05) --
	cycle;

\path[fill=fillColor] (420.95,285.11) --
	(424.32,279.29) --
	(417.59,279.29) --
	cycle;

\path[fill=fillColor] (429.38,288.42) --
	(432.74,282.59) --
	(426.02,282.59) --
	cycle;

\path[fill=fillColor] (437.81,291.80) --
	(441.17,285.97) --
	(434.45,285.97) --
	cycle;

\path[fill=fillColor] (446.24,295.24) --
	(449.60,289.41) --
	(442.87,289.41) --
	cycle;

\path[fill=fillColor] (454.67,298.75) --
	(458.03,292.92) --
	(451.30,292.92) --
	cycle;

\path[fill=fillColor] (463.09,302.32) --
	(466.46,296.50) --
	(459.73,296.50) --
	cycle;

\path[fill=fillColor] (471.52,305.95) --
	(474.89,300.13) --
	(468.16,300.13) --
	cycle;

\path[fill=fillColor] (479.95,309.64) --
	(483.31,303.81) --
	(476.59,303.81) --
	cycle;

\path[fill=fillColor] (488.38,313.38) --
	(491.74,307.55) --
	(485.01,307.55) --
	cycle;

\path[fill=fillColor] (496.81,317.16) --
	(500.17,311.34) --
	(493.44,311.34) --
	cycle;

\path[fill=fillColor] (505.23,320.99) --
	(508.60,315.16) --
	(501.87,315.16) --
	cycle;

\path[fill=fillColor] (513.66,324.85) --
	(517.03,319.03) --
	(510.30,319.03) --
	cycle;

\path[fill=fillColor] (522.09,328.74) --
	(525.45,322.92) --
	(518.73,322.92) --
	cycle;

\path[fill=fillColor] (530.52,332.66) --
	(533.88,326.83) --
	(527.16,326.83) --
	cycle;

\path[fill=fillColor] (538.95,336.58) --
	(542.31,330.76) --
	(535.58,330.76) --
	cycle;

\path[fill=fillColor] (547.38,340.52) --
	(550.74,334.69) --
	(544.01,334.69) --
	cycle;

\path[fill=fillColor] (555.80,344.44) --
	(559.17,338.62) --
	(552.44,338.62) --
	cycle;

\path[fill=fillColor] (564.23,348.36) --
	(567.60,342.53) --
	(560.87,342.53) --
	cycle;

\path[fill=fillColor] (572.66,352.25) --
	(576.02,346.42) --
	(569.30,346.42) --
	cycle;
\definecolor{fillColor}{RGB}{149,149,149}

\path[fill=fillColor] (317.32,245.00) --
	(322.31,245.00) --
	(322.31,249.99) --
	(317.32,249.99) --
	cycle;

\path[fill=fillColor] (325.75,247.37) --
	(330.74,247.37) --
	(330.74,252.37) --
	(325.75,252.37) --
	cycle;

\path[fill=fillColor] (334.17,249.82) --
	(339.17,249.82) --
	(339.17,254.81) --
	(334.17,254.81) --
	cycle;

\path[fill=fillColor] (342.60,252.32) --
	(347.60,252.32) --
	(347.60,257.32) --
	(342.60,257.32) --
	cycle;

\path[fill=fillColor] (351.03,254.89) --
	(356.03,254.89) --
	(356.03,259.88) --
	(351.03,259.88) --
	cycle;

\path[fill=fillColor] (359.46,257.51) --
	(364.45,257.51) --
	(364.45,262.51) --
	(359.46,262.51) --
	cycle;

\path[fill=fillColor] (367.89,260.19) --
	(372.88,260.19) --
	(372.88,265.19) --
	(367.89,265.19) --
	cycle;

\path[fill=fillColor] (376.31,262.93) --
	(381.31,262.93) --
	(381.31,267.92) --
	(376.31,267.92) --
	cycle;

\path[fill=fillColor] (384.74,265.71) --
	(389.74,265.71) --
	(389.74,270.71) --
	(384.74,270.71) --
	cycle;

\path[fill=fillColor] (393.17,268.54) --
	(398.17,268.54) --
	(398.17,273.54) --
	(393.17,273.54) --
	cycle;

\path[fill=fillColor] (401.60,271.42) --
	(406.59,271.42) --
	(406.59,276.41) --
	(401.60,276.41) --
	cycle;

\path[fill=fillColor] (410.03,274.33) --
	(415.02,274.33) --
	(415.02,279.32) --
	(410.03,279.32) --
	cycle;

\path[fill=fillColor] (418.46,277.27) --
	(423.45,277.27) --
	(423.45,282.27) --
	(418.46,282.27) --
	cycle;

\path[fill=fillColor] (426.88,280.24) --
	(431.88,280.24) --
	(431.88,285.24) --
	(426.88,285.24) --
	cycle;

\path[fill=fillColor] (435.31,283.23) --
	(440.31,283.23) --
	(440.31,288.23) --
	(435.31,288.23) --
	cycle;

\path[fill=fillColor] (443.74,286.24) --
	(448.73,286.24) --
	(448.73,291.23) --
	(443.74,291.23) --
	cycle;

\path[fill=fillColor] (452.17,289.25) --
	(457.16,289.25) --
	(457.16,294.24) --
	(452.17,294.24) --
	cycle;

\path[fill=fillColor] (460.60,292.25) --
	(465.59,292.25) --
	(465.59,297.25) --
	(460.60,297.25) --
	cycle;

\path[fill=fillColor] (469.02,295.25) --
	(474.02,295.25) --
	(474.02,300.24) --
	(469.02,300.24) --
	cycle;

\path[fill=fillColor] (477.45,298.22) --
	(482.45,298.22) --
	(482.45,303.21) --
	(477.45,303.21) --
	cycle;

\path[fill=fillColor] (485.88,301.15) --
	(490.88,301.15) --
	(490.88,306.15) --
	(485.88,306.15) --
	cycle;

\path[fill=fillColor] (494.31,304.05) --
	(499.30,304.05) --
	(499.30,309.04) --
	(494.31,309.04) --
	cycle;

\path[fill=fillColor] (502.74,306.88) --
	(507.73,306.88) --
	(507.73,311.88) --
	(502.74,311.88) --
	cycle;

\path[fill=fillColor] (511.17,309.64) --
	(516.16,309.64) --
	(516.16,314.64) --
	(511.17,314.64) --
	cycle;

\path[fill=fillColor] (519.59,312.32) --
	(524.59,312.32) --
	(524.59,317.31) --
	(519.59,317.31) --
	cycle;

\path[fill=fillColor] (528.02,314.89) --
	(533.02,314.89) --
	(533.02,319.88) --
	(528.02,319.88) --
	cycle;

\path[fill=fillColor] (536.45,317.34) --
	(541.44,317.34) --
	(541.44,322.33) --
	(536.45,322.33) --
	cycle;

\path[fill=fillColor] (544.88,319.65) --
	(549.87,319.65) --
	(549.87,324.64) --
	(544.88,324.64) --
	cycle;

\path[fill=fillColor] (553.31,321.80) --
	(558.30,321.80) --
	(558.30,326.79) --
	(553.31,326.79) --
	cycle;

\path[fill=fillColor] (561.73,323.76) --
	(566.73,323.76) --
	(566.73,328.76) --
	(561.73,328.76) --
	cycle;

\path[fill=fillColor] (570.16,325.52) --
	(575.16,325.52) --
	(575.16,330.52) --
	(570.16,330.52) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.28,247.50) -- (323.35,247.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (319.81,243.96) -- (319.81,251.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (324.71,249.87) -- (331.78,249.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.24,246.33) -- (328.24,253.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.14,252.29) -- (340.20,252.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (336.67,248.76) -- (336.67,255.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (341.57,254.76) -- (348.63,254.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.10,251.23) -- (345.10,258.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.00,257.28) -- (357.06,257.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (353.53,253.74) -- (353.53,260.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (358.42,259.83) -- (365.49,259.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.96,256.30) -- (361.96,263.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.85,262.43) -- (373.92,262.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (370.38,258.89) -- (370.38,265.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (375.28,265.05) -- (382.34,265.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (378.81,261.52) -- (378.81,268.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.71,267.70) -- (390.77,267.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.24,264.16) -- (387.24,271.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (392.14,270.36) -- (399.20,270.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.67,266.83) -- (395.67,273.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (400.56,273.03) -- (407.63,273.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (404.10,269.50) -- (404.10,276.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (408.99,275.71) -- (416.06,275.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.52,272.18) -- (412.52,279.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.42,278.38) -- (424.48,278.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.95,274.85) -- (420.95,281.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.85,281.04) -- (432.91,281.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.38,277.51) -- (429.38,284.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (434.28,283.67) -- (441.34,283.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.81,280.13) -- (437.81,287.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (442.71,286.26) -- (449.77,286.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (446.24,282.73) -- (446.24,289.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (451.13,288.80) -- (458.20,288.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.67,285.27) -- (454.67,292.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.56,291.28) -- (466.63,291.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.09,287.75) -- (463.09,294.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.99,293.69) -- (475.05,293.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.52,290.15) -- (471.52,297.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.42,296.00) -- (483.48,296.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.95,292.47) -- (479.95,299.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (484.85,298.20) -- (491.91,298.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (488.38,294.67) -- (488.38,301.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (493.27,300.28) -- (500.34,300.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (496.81,296.75) -- (496.81,303.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.70,302.21) -- (508.77,302.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.23,298.68) -- (505.23,305.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.13,303.98) -- (517.19,303.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (513.66,300.45) -- (513.66,307.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (518.56,305.56) -- (525.62,305.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.09,302.03) -- (522.09,309.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.99,306.93) -- (534.05,306.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (530.52,303.40) -- (530.52,310.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.42,308.07) -- (542.48,308.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (538.95,304.54) -- (538.95,311.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (543.84,308.95) -- (550.91,308.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.38,305.42) -- (547.38,312.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.27,309.54) -- (559.34,309.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (555.80,306.01) -- (555.80,313.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.70,309.82) -- (567.76,309.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.23,306.29) -- (564.23,313.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13,309.77) -- (576.19,309.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66,306.24) -- (572.66,313.30);
\definecolor{drawColor}{gray}{0.80}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (317.32,245.00) rectangle (322.31,249.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (317.32,245.00) -- (322.31,249.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (317.32,249.99) -- (322.31,245.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (325.75,247.36) rectangle (330.74,252.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (325.75,247.36) -- (330.74,252.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (325.75,252.36) -- (330.74,247.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.17,249.77) rectangle (339.17,254.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.17,249.77) -- (339.17,254.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.17,254.76) -- (339.17,249.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.60,252.21) rectangle (347.60,257.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.60,252.21) -- (347.60,257.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.60,257.20) -- (347.60,252.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.03,254.68) rectangle (356.03,259.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.03,254.68) -- (356.03,259.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.03,259.67) -- (356.03,254.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.46,257.17) rectangle (364.45,262.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.46,257.17) -- (364.45,262.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.46,262.16) -- (364.45,257.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.89,259.67) rectangle (372.88,264.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.89,259.67) -- (372.88,264.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.89,264.67) -- (372.88,259.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.31,262.19) rectangle (381.31,267.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.31,262.19) -- (381.31,267.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.31,267.19) -- (381.31,262.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.74,264.71) rectangle (389.74,269.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.74,264.71) -- (389.74,269.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.74,269.70) -- (389.74,264.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.17,267.21) rectangle (398.17,272.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.17,267.21) -- (398.17,272.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.17,272.21) -- (398.17,267.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.60,269.70) rectangle (406.59,274.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.60,269.70) -- (406.59,274.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.60,274.69) -- (406.59,269.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.03,272.15) rectangle (415.02,277.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.03,272.15) -- (415.02,277.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.03,277.15) -- (415.02,272.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.46,274.57) rectangle (423.45,279.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.46,274.57) -- (423.45,279.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.46,279.56) -- (423.45,274.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.88,276.93) rectangle (431.88,281.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.88,276.93) -- (431.88,281.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.88,281.93) -- (431.88,276.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.31,279.22) rectangle (440.31,284.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.31,279.22) -- (440.31,284.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.31,284.22) -- (440.31,279.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.74,281.44) rectangle (448.73,286.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.74,281.44) -- (448.73,286.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.74,286.43) -- (448.73,281.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.17,283.55) rectangle (457.16,288.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.17,283.55) -- (457.16,288.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.17,288.55) -- (457.16,283.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (460.60,285.55) rectangle (465.59,290.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (460.60,285.55) -- (465.59,290.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (460.60,290.55) -- (465.59,285.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.02,287.42) rectangle (474.02,292.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.02,287.42) -- (474.02,292.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.02,292.42) -- (474.02,287.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.45,289.14) rectangle (482.45,294.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.45,289.14) -- (482.45,294.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.45,294.14) -- (482.45,289.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.88,290.69) rectangle (490.88,295.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.88,290.69) -- (490.88,295.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.88,295.69) -- (490.88,290.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.31,292.05) rectangle (499.30,297.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.31,292.05) -- (499.30,297.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.31,297.04) -- (499.30,292.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.74,293.19) rectangle (507.73,298.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.74,293.19) -- (507.73,298.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.74,298.19) -- (507.73,293.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.17,294.09) rectangle (516.16,299.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.17,294.09) -- (516.16,299.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.17,299.09) -- (516.16,294.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.59,294.73) rectangle (524.59,299.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.59,294.73) -- (524.59,299.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.59,299.73) -- (524.59,294.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.02,295.08) rectangle (533.02,300.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.02,295.08) -- (533.02,300.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.02,300.07) -- (533.02,295.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.45,295.11) rectangle (541.44,300.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.45,295.11) -- (541.44,300.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.45,300.11) -- (541.44,295.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.88,294.80) rectangle (549.87,299.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.88,294.80) -- (549.87,299.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.88,299.79) -- (549.87,294.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.31,294.12) rectangle (558.30,299.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.31,294.12) -- (558.30,299.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.31,299.12) -- (558.30,294.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.73,293.05) rectangle (566.73,298.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.73,293.05) -- (566.73,298.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.73,298.05) -- (566.73,293.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16,291.57) rectangle (575.16,296.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16,291.57) -- (575.16,296.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16,296.57) -- (575.16,291.57);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (319.81,247.50) circle (  2.50);

\path[fill=fillColor] (328.24,249.88) circle (  2.50);

\path[fill=fillColor] (336.67,252.37) circle (  2.50);

\path[fill=fillColor] (345.10,254.96) circle (  2.50);

\path[fill=fillColor] (353.53,257.67) circle (  2.50);

\path[fill=fillColor] (361.96,260.49) circle (  2.50);

\path[fill=fillColor] (370.38,263.43) circle (  2.50);

\path[fill=fillColor] (378.81,266.50) circle (  2.50);

\path[fill=fillColor] (387.24,269.69) circle (  2.50);

\path[fill=fillColor] (395.67,273.02) circle (  2.50);

\path[fill=fillColor] (404.10,276.48) circle (  2.50);

\path[fill=fillColor] (412.52,280.09) circle (  2.50);

\path[fill=fillColor] (420.95,283.85) circle (  2.50);

\path[fill=fillColor] (429.38,287.76) circle (  2.50);

\path[fill=fillColor] (437.81,291.82) circle (  2.50);

\path[fill=fillColor] (446.24,296.05) circle (  2.50);

\path[fill=fillColor] (454.67,300.46) circle (  2.50);

\path[fill=fillColor] (463.09,305.04) circle (  2.50);

\path[fill=fillColor] (471.52,309.80) circle (  2.50);

\path[fill=fillColor] (479.95,314.74) circle (  2.50);

\path[fill=fillColor] (488.38,319.89) circle (  2.50);

\path[fill=fillColor] (496.81,325.23) circle (  2.50);

\path[fill=fillColor] (505.23,330.79) circle (  2.50);

\path[fill=fillColor] (513.66,336.55) circle (  2.50);

\path[fill=fillColor] (522.09,342.55) circle (  2.50);

\path[fill=fillColor] (530.52,348.77) circle (  2.50);

\path[fill=fillColor] (538.95,355.23) circle (  2.50);

\path[fill=fillColor] (547.38,361.93) circle (  2.50);

\path[fill=fillColor] (555.80,368.89) circle (  2.50);

\path[fill=fillColor] (564.23,376.12) circle (  2.50);

\path[fill=fillColor] (572.66,383.61) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.81,247.50) rectangle (572.66,410.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (329.08,389.03) rectangle (343.54,403.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (330.53,396.25) -- (342.09,396.25);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (336.31,396.25) circle (  2.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (329.08,374.57) rectangle (343.54,389.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{109,109,109}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (330.53,381.80) -- (342.09,381.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{109,109,109}

\path[fill=fillColor] (336.31,385.68) --
	(339.67,379.86) --
	(332.95,379.86) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (329.08,360.12) rectangle (343.54,374.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{149,149,149}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (330.53,367.35) -- (342.09,367.35);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{149,149,149}

\path[fill=fillColor] (333.81,364.85) --
	(338.81,364.85) --
	(338.81,369.84) --
	(333.81,369.84) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (329.08,345.67) rectangle (343.54,360.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (330.53,352.89) -- (342.09,352.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (332.78,352.89) -- (339.84,352.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (336.31,349.36) -- (336.31,356.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (329.08,331.21) rectangle (343.54,345.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.80}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] (330.53,338.44) -- (342.09,338.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.80}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.81,335.94) rectangle (338.81,340.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.81,335.94) -- (338.81,340.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.81,340.94) -- (338.81,335.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (349.04,393.22) {Intrinsic generation interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (349.04,378.77) {Intrinsic serial interval ($\rho=0.75$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (349.04,364.32) {Intrinsic serial interval ($\rho=0.5$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (349.04,349.86) {Intrinsic serial interval ($\rho=0.25$)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (349.04,335.41) {Intrinsic serial interval ($\rho=0$)};
\end{scope}
\begin{scope}
\path[clip] ( 27.31, 30.69) rectangle (280.16,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.31, 30.69) rectangle (280.16,194.15);
\definecolor{drawColor}{RGB}{13,8,135}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.31, 63.20) --
	( 30.51, 63.20) --
	( 33.71, 63.20) --
	( 36.91, 63.20) --
	( 40.12, 63.20) --
	( 43.32, 63.20) --
	( 46.52, 63.20) --
	( 49.72, 63.20) --
	( 52.92, 63.20) --
	( 56.12, 63.20) --
	( 59.32, 63.20) --
	( 62.52, 63.20) --
	( 65.72, 63.20) --
	( 68.92, 63.20) --
	( 72.12, 63.20) --
	( 75.32, 63.20) --
	( 78.52, 63.20) --
	( 81.72, 63.20) --
	( 84.92, 63.20) --
	( 88.12, 63.20) --
	( 91.32, 63.20) --
	( 94.53, 63.20) --
	( 97.73, 63.20) --
	(100.93, 63.20) --
	(104.13, 63.20) --
	(107.33, 63.20) --
	(110.53, 63.20) --
	(113.73, 63.20) --
	(116.93, 63.20) --
	(120.13, 63.20) --
	(123.33, 63.20) --
	(126.53, 63.20) --
	(129.73, 63.20) --
	(132.93, 63.20) --
	(136.13, 63.20) --
	(139.33, 63.20) --
	(142.53, 63.20) --
	(145.73, 63.20) --
	(148.94, 63.20) --
	(152.14, 63.20) --
	(155.34, 63.20) --
	(158.54, 63.20) --
	(161.74, 63.20) --
	(164.94, 63.20) --
	(168.14, 63.20) --
	(171.34, 63.20) --
	(174.54, 63.20) --
	(177.74, 63.20) --
	(180.94, 63.20) --
	(184.14, 63.20) --
	(187.34, 63.20) --
	(190.54, 63.20) --
	(193.74, 63.20) --
	(196.94, 63.20) --
	(200.14, 63.20) --
	(203.34, 63.20) --
	(206.55, 63.20) --
	(209.75, 63.20) --
	(212.95, 63.20) --
	(216.15, 63.20) --
	(219.35, 63.20) --
	(222.55, 63.20) --
	(225.75, 63.20) --
	(228.95, 63.20) --
	(232.15, 63.20) --
	(235.35, 63.20) --
	(238.55, 63.20) --
	(241.75, 63.20) --
	(244.95, 63.20) --
	(248.15, 63.20) --
	(251.35, 63.20) --
	(254.55, 63.20) --
	(257.75, 63.20) --
	(260.96, 63.20) --
	(264.16, 63.20) --
	(267.36, 63.20) --
	(270.56, 63.20) --
	(273.76, 63.20) --
	(276.96, 63.20) --
	(280.16, 63.20);
\definecolor{drawColor}{RGB}{126,3,168}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] ( 27.31, 64.16) --
	( 30.51, 65.16) --
	( 33.71, 66.15) --
	( 36.91, 67.13) --
	( 40.12, 68.09) --
	( 43.32, 69.05) --
	( 46.52, 69.99) --
	( 49.72, 70.92) --
	( 52.92, 71.84) --
	( 56.12, 72.75) --
	( 59.32, 73.64) --
	( 62.52, 74.53) --
	( 65.72, 75.40) --
	( 68.92, 76.26) --
	( 72.12, 77.11) --
	( 75.32, 77.95) --
	( 78.52, 78.78) --
	( 81.72, 79.59) --
	( 84.92, 80.40) --
	( 88.12, 81.20) --
	( 91.32, 81.98) --
	( 94.53, 82.75) --
	( 97.73, 83.50) --
	(100.93, 84.24) --
	(104.13, 84.98) --
	(107.33, 85.70) --
	(110.53, 86.41) --
	(113.73, 87.11) --
	(116.93, 87.80) --
	(120.13, 88.49) --
	(123.33, 89.16) --
	(126.53, 89.82) --
	(129.73, 90.46) --
	(132.93, 91.09) --
	(136.13, 91.71) --
	(139.33, 92.33) --
	(142.53, 92.93) --
	(145.73, 93.54) --
	(148.94, 94.13) --
	(152.14, 94.73) --
	(155.34, 95.33) --
	(158.54, 95.92) --
	(161.74, 96.50) --
	(164.94, 97.08) --
	(168.14, 97.64) --
	(171.34, 98.20) --
	(174.54, 98.76) --
	(177.74, 99.31) --
	(180.94, 99.86) --
	(184.14,100.41) --
	(187.34,100.95) --
	(190.54,101.49) --
	(193.74,102.03) --
	(196.94,102.56) --
	(200.14,103.08) --
	(203.34,103.60) --
	(206.55,104.12) --
	(209.75,104.63) --
	(212.95,105.14) --
	(216.15,105.65) --
	(219.35,106.15) --
	(222.55,106.65) --
	(225.75,107.14) --
	(228.95,107.63) --
	(232.15,108.12) --
	(235.35,108.60) --
	(238.55,109.07) --
	(241.75,109.54) --
	(244.95,110.01) --
	(248.15,110.48) --
	(251.35,110.94) --
	(254.55,111.39) --
	(257.75,111.84) --
	(260.96,112.29) --
	(264.16,112.73) --
	(267.36,113.17) --
	(270.56,113.61) --
	(273.76,114.04) --
	(276.96,114.46) --
	(280.16,114.89);
\definecolor{drawColor}{RGB}{204,70,120}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] ( 27.31, 64.16) --
	( 30.51, 65.64) --
	( 33.71, 67.11) --
	( 36.91, 68.56) --
	( 40.12, 69.99) --
	( 43.32, 71.41) --
	( 46.52, 72.81) --
	( 49.72, 74.19) --
	( 52.92, 75.56) --
	( 56.12, 76.91) --
	( 59.32, 78.25) --
	( 62.52, 79.56) --
	( 65.72, 80.86) --
	( 68.92, 82.15) --
	( 72.12, 83.42) --
	( 75.32, 84.67) --
	( 78.52, 85.91) --
	( 81.72, 87.13) --
	( 84.92, 88.34) --
	( 88.12, 89.53) --
	( 91.32, 90.70) --
	( 94.53, 91.85) --
	( 97.73, 92.98) --
	(100.93, 94.10) --
	(104.13, 95.20) --
	(107.33, 96.29) --
	(110.53, 97.36) --
	(113.73, 98.41) --
	(116.93, 99.46) --
	(120.13,100.49) --
	(123.33,101.51) --
	(126.53,102.50) --
	(129.73,103.47) --
	(132.93,104.43) --
	(136.13,105.37) --
	(139.33,106.29) --
	(142.53,107.21) --
	(145.73,108.12) --
	(148.94,109.03) --
	(152.14,109.94) --
	(155.34,110.85) --
	(158.54,111.75) --
	(161.74,112.63) --
	(164.94,113.50) --
	(168.14,114.37) --
	(171.34,115.22) --
	(174.54,116.07) --
	(177.74,116.92) --
	(180.94,117.75) --
	(184.14,118.59) --
	(187.34,119.42) --
	(190.54,120.24) --
	(193.74,121.06) --
	(196.94,121.87) --
	(200.14,122.67) --
	(203.34,123.47) --
	(206.55,124.26) --
	(209.75,125.04) --
	(212.95,125.82) --
	(216.15,126.59) --
	(219.35,127.36) --
	(222.55,128.12) --
	(225.75,128.87) --
	(228.95,129.62) --
	(232.15,130.36) --
	(235.35,131.10) --
	(238.55,131.83) --
	(241.75,132.55) --
	(244.95,133.26) --
	(248.15,133.97) --
	(251.35,134.68) --
	(254.55,135.37) --
	(257.75,136.07) --
	(260.96,136.75) --
	(264.16,137.43) --
	(267.36,138.10) --
	(270.56,138.77) --
	(273.76,139.43) --
	(276.96,140.08) --
	(280.16,140.73);
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] ( 27.31, 64.08) --
	( 30.51, 66.04) --
	( 33.71, 67.97) --
	( 36.91, 69.88) --
	( 40.12, 71.77) --
	( 43.32, 73.64) --
	( 46.52, 75.49) --
	( 49.72, 77.31) --
	( 52.92, 79.12) --
	( 56.12, 80.91) --
	( 59.32, 82.67) --
	( 62.52, 84.42) --
	( 65.72, 86.14) --
	( 68.92, 87.84) --
	( 72.12, 89.53) --
	( 75.32, 91.19) --
	( 78.52, 92.83) --
	( 81.72, 94.46) --
	( 84.92, 96.06) --
	( 88.12, 97.65) --
	( 91.32, 99.21) --
	( 94.53,100.75) --
	( 97.73,102.26) --
	(100.93,103.75) --
	(104.13,105.23) --
	(107.33,106.68) --
	(110.53,108.11) --
	(113.73,109.53) --
	(116.93,110.93) --
	(120.13,112.32) --
	(123.33,113.68) --
	(126.53,115.02) --
	(129.73,116.33) --
	(132.93,117.61) --
	(136.13,118.88) --
	(139.33,120.13) --
	(142.53,121.37) --
	(145.73,122.61) --
	(148.94,123.84) --
	(152.14,125.07) --
	(155.34,126.29) --
	(158.54,127.51) --
	(161.74,128.71) --
	(164.94,129.90) --
	(168.14,131.07) --
	(171.34,132.24) --
	(174.54,133.40) --
	(177.74,134.54) --
	(180.94,135.68) --
	(184.14,136.82) --
	(187.34,137.95) --
	(190.54,139.07) --
	(193.74,140.19) --
	(196.94,141.29) --
	(200.14,142.39) --
	(203.34,143.48) --
	(206.55,144.56) --
	(209.75,145.63) --
	(212.95,146.69) --
	(216.15,147.75) --
	(219.35,148.80) --
	(222.55,149.84) --
	(225.75,150.87) --
	(228.95,151.89) --
	(232.15,152.91) --
	(235.35,153.91) --
	(238.55,154.91) --
	(241.75,155.90) --
	(244.95,156.88) --
	(248.15,157.85) --
	(251.35,158.82) --
	(254.55,159.78) --
	(257.75,160.73) --
	(260.96,161.67) --
	(264.16,162.60) --
	(267.36,163.52) --
	(270.56,164.44) --
	(273.76,165.35) --
	(276.96,166.24) --
	(280.16,167.14);
\definecolor{drawColor}{RGB}{216,224,30}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] ( 27.31, 63.95) --
	( 30.51, 66.36) --
	( 33.71, 68.74) --
	( 36.91, 71.10) --
	( 40.12, 73.43) --
	( 43.32, 75.74) --
	( 46.52, 78.03) --
	( 49.72, 80.29) --
	( 52.92, 82.53) --
	( 56.12, 84.74) --
	( 59.32, 86.93) --
	( 62.52, 89.09) --
	( 65.72, 91.23) --
	( 68.92, 93.35) --
	( 72.12, 95.44) --
	( 75.32, 97.51) --
	( 78.52, 99.55) --
	( 81.72,101.58) --
	( 84.92,103.58) --
	( 88.12,105.56) --
	( 91.32,107.51) --
	( 94.53,109.43) --
	( 97.73,111.32) --
	(100.93,113.19) --
	(104.13,115.04) --
	(107.33,116.86) --
	(110.53,118.67) --
	(113.73,120.45) --
	(116.93,122.21) --
	(120.13,123.96) --
	(123.33,125.67) --
	(126.53,127.35) --
	(129.73,129.01) --
	(132.93,130.63) --
	(136.13,132.24) --
	(139.33,133.82) --
	(142.53,135.39) --
	(145.73,136.96) --
	(148.94,138.52) --
	(152.14,140.07) --
	(155.34,141.63) --
	(158.54,143.18) --
	(161.74,144.71) --
	(164.94,146.22) --
	(168.14,147.72) --
	(171.34,149.21) --
	(174.54,150.69) --
	(177.74,152.15) --
	(180.94,153.61) --
	(184.14,155.06) --
	(187.34,156.51) --
	(190.54,157.94) --
	(193.74,159.37) --
	(196.94,160.79) --
	(200.14,162.19) --
	(203.34,163.58) --
	(206.55,164.97) --
	(209.75,166.34) --
	(212.95,167.71) --
	(216.15,169.06) --
	(219.35,170.41) --
	(222.55,171.74) --
	(225.75,173.07) --
	(228.95,174.38) --
	(232.15,175.69) --
	(235.35,176.98) --
	(238.55,178.27) --
	(241.75,179.54) --
	(244.95,180.81) --
	(248.15,182.06) --
	(251.35,183.31) --
	(254.55,184.54) --
	(257.75,185.77) --
	(260.96,186.98) --
	(264.16,188.19) --
	(267.36,189.38) --
	(270.56,190.57) --
	(273.76,191.74) --
	(276.96,192.91) --
	(280.16,194.06);
\definecolor{fillColor}{RGB}{126,3,168}

\path[fill=fillColor] ( 27.31, 68.04) --
	( 30.68, 62.21) --
	( 23.95, 62.21) --
	cycle;

\path[fill=fillColor] ( 35.74, 70.65) --
	( 39.11, 64.82) --
	( 32.38, 64.82) --
	cycle;

\path[fill=fillColor] ( 44.17, 73.00) --
	( 47.53, 67.17) --
	( 40.81, 67.17) --
	cycle;

\path[fill=fillColor] ( 52.60, 75.66) --
	( 55.96, 69.83) --
	( 49.23, 69.83) --
	cycle;

\path[fill=fillColor] ( 61.03, 78.15) --
	( 64.39, 72.32) --
	( 57.66, 72.32) --
	cycle;

\path[fill=fillColor] ( 69.45, 80.38) --
	( 72.82, 74.55) --
	( 66.09, 74.55) --
	cycle;

\path[fill=fillColor] ( 77.88, 82.50) --
	( 81.25, 76.67) --
	( 74.52, 76.67) --
	cycle;

\path[fill=fillColor] ( 86.31, 84.65) --
	( 89.67, 78.82) --
	( 82.95, 78.82) --
	cycle;

\path[fill=fillColor] ( 94.74, 86.82) --
	( 98.10, 80.99) --
	( 91.37, 80.99) --
	cycle;

\path[fill=fillColor] (103.17, 88.79) --
	(106.53, 82.96) --
	( 99.80, 82.96) --
	cycle;

\path[fill=fillColor] (111.60, 90.70) --
	(114.96, 84.87) --
	(108.23, 84.87) --
	cycle;

\path[fill=fillColor] (120.02, 92.30) --
	(123.39, 86.48) --
	(116.66, 86.48) --
	cycle;

\path[fill=fillColor] (128.45, 94.04) --
	(131.82, 88.21) --
	(125.09, 88.21) --
	cycle;

\path[fill=fillColor] (136.88, 95.68) --
	(140.24, 89.85) --
	(133.52, 89.85) --
	cycle;

\path[fill=fillColor] (145.31, 97.35) --
	(148.67, 91.53) --
	(141.94, 91.53) --
	cycle;

\path[fill=fillColor] (153.74, 98.90) --
	(157.10, 93.07) --
	(150.37, 93.07) --
	cycle;

\path[fill=fillColor] (162.16,100.37) --
	(165.53, 94.54) --
	(158.80, 94.54) --
	cycle;

\path[fill=fillColor] (170.59,101.89) --
	(173.96, 96.06) --
	(167.23, 96.06) --
	cycle;

\path[fill=fillColor] (179.02,103.46) --
	(182.38, 97.64) --
	(175.66, 97.64) --
	cycle;

\path[fill=fillColor] (187.45,104.87) --
	(190.81, 99.04) --
	(184.08, 99.04) --
	cycle;

\path[fill=fillColor] (195.88,106.29) --
	(199.24,100.47) --
	(192.51,100.47) --
	cycle;

\path[fill=fillColor] (204.30,107.65) --
	(207.67,101.82) --
	(200.94,101.82) --
	cycle;

\path[fill=fillColor] (212.73,108.93) --
	(216.10,103.11) --
	(209.37,103.11) --
	cycle;

\path[fill=fillColor] (221.16,110.30) --
	(224.53,104.48) --
	(217.80,104.48) --
	cycle;

\path[fill=fillColor] (229.59,111.62) --
	(232.95,105.79) --
	(226.23,105.79) --
	cycle;

\path[fill=fillColor] (238.02,112.82) --
	(241.38,106.99) --
	(234.65,106.99) --
	cycle;

\path[fill=fillColor] (246.45,114.10) --
	(249.81,108.28) --
	(243.08,108.28) --
	cycle;

\path[fill=fillColor] (254.87,115.26) --
	(258.24,109.43) --
	(251.51,109.43) --
	cycle;

\path[fill=fillColor] (263.30,116.55) --
	(266.67,110.72) --
	(259.94,110.72) --
	cycle;

\path[fill=fillColor] (271.73,117.65) --
	(275.09,111.83) --
	(268.37,111.83) --
	cycle;

\path[fill=fillColor] (280.16,118.80) --
	(283.52,112.97) --
	(276.79,112.97) --
	cycle;
\definecolor{fillColor}{RGB}{204,70,120}

\path[fill=fillColor] ( 24.82, 61.68) --
	( 29.81, 61.68) --
	( 29.81, 66.68) --
	( 24.82, 66.68) --
	cycle;

\path[fill=fillColor] ( 33.24, 65.53) --
	( 38.24, 65.53) --
	( 38.24, 70.52) --
	( 33.24, 70.52) --
	cycle;

\path[fill=fillColor] ( 41.67, 69.01) --
	( 46.67, 69.01) --
	( 46.67, 74.00) --
	( 41.67, 74.00) --
	cycle;

\path[fill=fillColor] ( 50.10, 72.95) --
	( 55.10, 72.95) --
	( 55.10, 77.95) --
	( 50.10, 77.95) --
	cycle;

\path[fill=fillColor] ( 58.53, 76.64) --
	( 63.52, 76.64) --
	( 63.52, 81.63) --
	( 58.53, 81.63) --
	cycle;

\path[fill=fillColor] ( 66.96, 79.97) --
	( 71.95, 79.97) --
	( 71.95, 84.96) --
	( 66.96, 84.96) --
	cycle;

\path[fill=fillColor] ( 75.38, 83.16) --
	( 80.38, 83.16) --
	( 80.38, 88.16) --
	( 75.38, 88.16) --
	cycle;

\path[fill=fillColor] ( 83.81, 86.41) --
	( 88.81, 86.41) --
	( 88.81, 91.41) --
	( 83.81, 91.41) --
	cycle;

\path[fill=fillColor] ( 92.24, 89.65) --
	( 97.24, 89.65) --
	( 97.24, 94.65) --
	( 92.24, 94.65) --
	cycle;

\path[fill=fillColor] (100.67, 92.60) --
	(105.66, 92.60) --
	(105.66, 97.60) --
	(100.67, 97.60) --
	cycle;

\path[fill=fillColor] (109.10, 95.42) --
	(114.09, 95.42) --
	(114.09,100.42) --
	(109.10,100.42) --
	cycle;

\path[fill=fillColor] (117.53, 97.90) --
	(122.52, 97.90) --
	(122.52,102.89) --
	(117.53,102.89) --
	cycle;

\path[fill=fillColor] (125.95,100.49) --
	(130.95,100.49) --
	(130.95,105.49) --
	(125.95,105.49) --
	cycle;

\path[fill=fillColor] (134.38,103.01) --
	(139.38,103.01) --
	(139.38,108.00) --
	(134.38,108.00) --
	cycle;

\path[fill=fillColor] (142.81,105.51) --
	(147.81,105.51) --
	(147.81,110.51) --
	(142.81,110.51) --
	cycle;

\path[fill=fillColor] (151.24,107.88) --
	(156.23,107.88) --
	(156.23,112.88) --
	(151.24,112.88) --
	cycle;

\path[fill=fillColor] (159.67,110.12) --
	(164.66,110.12) --
	(164.66,115.11) --
	(159.67,115.11) --
	cycle;

\path[fill=fillColor] (168.09,112.40) --
	(173.09,112.40) --
	(173.09,117.39) --
	(168.09,117.39) --
	cycle;

\path[fill=fillColor] (176.52,114.82) --
	(181.52,114.82) --
	(181.52,119.82) --
	(176.52,119.82) --
	cycle;

\path[fill=fillColor] (184.95,116.97) --
	(189.95,116.97) --
	(189.95,121.96) --
	(184.95,121.96) --
	cycle;

\path[fill=fillColor] (193.38,119.16) --
	(198.37,119.16) --
	(198.37,124.15) --
	(193.38,124.15) --
	cycle;

\path[fill=fillColor] (201.81,121.21) --
	(206.80,121.21) --
	(206.80,126.20) --
	(201.81,126.20) --
	cycle;

\path[fill=fillColor] (210.24,123.21) --
	(215.23,123.21) --
	(215.23,128.21) --
	(210.24,128.21) --
	cycle;

\path[fill=fillColor] (218.66,125.29) --
	(223.66,125.29) --
	(223.66,130.28) --
	(218.66,130.28) --
	cycle;

\path[fill=fillColor] (227.09,127.27) --
	(232.09,127.27) --
	(232.09,132.26) --
	(227.09,132.26) --
	cycle;

\path[fill=fillColor] (235.52,129.12) --
	(240.52,129.12) --
	(240.52,134.11) --
	(235.52,134.11) --
	cycle;

\path[fill=fillColor] (243.95,131.10) --
	(248.94,131.10) --
	(248.94,136.09) --
	(243.95,136.09) --
	cycle;

\path[fill=fillColor] (252.38,132.88) --
	(257.37,132.88) --
	(257.37,137.87) --
	(252.38,137.87) --
	cycle;

\path[fill=fillColor] (260.80,134.82) --
	(265.80,134.82) --
	(265.80,139.81) --
	(260.80,139.81) --
	cycle;

\path[fill=fillColor] (269.23,136.50) --
	(274.23,136.50) --
	(274.23,141.49) --
	(269.23,141.49) --
	cycle;

\path[fill=fillColor] (277.66,138.27) --
	(282.66,138.27) --
	(282.66,143.27) --
	(277.66,143.27) --
	cycle;
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 23.78, 64.15) -- ( 30.85, 64.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.31, 60.62) -- ( 27.31, 67.68);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 32.21, 69.18) -- ( 39.27, 69.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 35.74, 65.65) -- ( 35.74, 72.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 40.64, 73.76) -- ( 47.70, 73.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 44.17, 70.23) -- ( 44.17, 77.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.07, 78.96) -- ( 56.13, 78.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 52.60, 75.42) -- ( 52.60, 82.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.49, 83.79) -- ( 64.56, 83.79);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.03, 80.26) -- ( 61.03, 87.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.92, 88.22) -- ( 72.99, 88.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 69.45, 84.69) -- ( 69.45, 91.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 74.35, 92.50) -- ( 81.41, 92.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.88, 88.97) -- ( 77.88, 96.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 82.78, 96.86) -- ( 89.84, 96.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.31, 93.33) -- ( 86.31,100.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.21,101.17) -- ( 98.27,101.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 94.74, 97.63) -- ( 94.74,104.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,105.11) -- (106.70,105.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.17,101.57) -- (103.17,108.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (108.06,108.80) -- (115.13,108.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (111.60,105.26) -- (111.60,112.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (116.49,112.21) -- (123.56,112.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.02,108.68) -- (120.02,115.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (124.92,115.67) -- (131.98,115.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (128.45,112.14) -- (128.45,119.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.35,119.09) -- (140.41,119.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (136.88,115.56) -- (136.88,122.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (141.78,122.43) -- (148.84,122.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (145.31,118.90) -- (145.31,125.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (150.20,125.69) -- (157.27,125.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (153.74,122.16) -- (153.74,129.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.63,128.71) -- (165.70,128.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.16,125.18) -- (162.16,132.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (167.06,131.76) -- (174.12,131.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.59,128.22) -- (170.59,135.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (175.49,135.09) -- (182.55,135.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (179.02,131.55) -- (179.02,138.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (183.92,138.02) -- (190.98,138.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.45,134.49) -- (187.45,141.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (192.34,141.02) -- (199.41,141.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.88,137.49) -- (195.88,144.56);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.77,143.80) -- (207.84,143.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (204.30,140.26) -- (204.30,147.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (209.20,146.57) -- (216.27,146.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (212.73,143.04) -- (212.73,150.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.63,149.39) -- (224.69,149.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (221.16,145.86) -- (221.16,152.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.06,152.08) -- (233.12,152.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (229.59,148.55) -- (229.59,155.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.49,154.61) -- (241.55,154.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (238.02,151.07) -- (238.02,158.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.91,157.36) -- (249.98,157.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (246.45,153.83) -- (246.45,160.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.34,159.82) -- (258.41,159.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (254.87,156.28) -- (254.87,163.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (259.77,162.43) -- (266.83,162.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (263.30,158.90) -- (263.30,165.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.20,164.72) -- (275.26,164.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.73,161.19) -- (271.73,168.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.63,167.20) -- (283.69,167.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.16,163.67) -- (280.16,170.73);
\definecolor{drawColor}{RGB}{216,224,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 24.82, 61.58) rectangle ( 29.81, 66.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 24.82, 61.58) -- ( 29.81, 66.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 24.82, 66.58) -- ( 29.81, 61.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.24, 67.75) rectangle ( 38.24, 72.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.24, 67.75) -- ( 38.24, 72.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 33.24, 72.74) -- ( 38.24, 67.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.67, 73.39) rectangle ( 46.67, 78.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.67, 73.39) -- ( 46.67, 78.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.67, 78.39) -- ( 46.67, 73.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.10, 79.80) rectangle ( 55.10, 84.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.10, 79.80) -- ( 55.10, 84.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.10, 84.80) -- ( 55.10, 79.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.53, 85.74) rectangle ( 63.52, 90.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.53, 85.74) -- ( 63.52, 90.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.53, 90.74) -- ( 63.52, 85.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.96, 91.27) rectangle ( 71.95, 96.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.96, 91.27) -- ( 71.95, 96.26);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.96, 96.26) -- ( 71.95, 91.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.38, 96.65) rectangle ( 80.38,101.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.38, 96.65) -- ( 80.38,101.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 75.38,101.64) -- ( 80.38, 96.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.81,102.12) rectangle ( 88.81,107.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.81,102.12) -- ( 88.81,107.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.81,107.12) -- ( 88.81,102.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.24,107.48) rectangle ( 97.24,112.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.24,107.48) -- ( 97.24,112.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 92.24,112.48) -- ( 97.24,107.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.67,112.41) rectangle (105.66,117.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.67,112.41) -- (105.66,117.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.67,117.40) -- (105.66,112.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.10,116.94) rectangle (114.09,121.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.10,116.94) -- (114.09,121.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.10,121.93) -- (114.09,116.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.53,121.34) rectangle (122.52,126.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.53,121.34) -- (122.52,126.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (117.53,126.34) -- (122.52,121.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (125.95,125.66) rectangle (130.95,130.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (125.95,125.66) -- (130.95,130.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (125.95,130.66) -- (130.95,125.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.38,130.03) rectangle (139.38,135.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.38,130.03) -- (139.38,135.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (134.38,135.02) -- (139.38,130.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (142.81,134.21) rectangle (147.81,139.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (142.81,134.21) -- (147.81,139.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (142.81,139.21) -- (147.81,134.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.24,138.41) rectangle (156.23,143.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.24,138.41) -- (156.23,143.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.24,143.40) -- (156.23,138.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.67,142.23) rectangle (164.66,147.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.67,142.23) -- (164.66,147.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.67,147.22) -- (164.66,142.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (168.09,146.06) rectangle (173.09,151.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (168.09,146.06) -- (173.09,151.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (168.09,151.05) -- (173.09,146.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.52,150.34) rectangle (181.52,155.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.52,150.34) -- (181.52,155.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (176.52,155.34) -- (181.52,150.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.95,154.10) rectangle (189.95,159.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.95,154.10) -- (189.95,159.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.95,159.10) -- (189.95,154.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (193.38,157.97) rectangle (198.37,162.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (193.38,157.97) -- (198.37,162.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (193.38,162.96) -- (198.37,157.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (201.81,161.49) rectangle (206.80,166.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (201.81,161.49) -- (206.80,166.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (201.81,166.49) -- (206.80,161.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (210.24,165.09) rectangle (215.23,170.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (210.24,165.09) -- (215.23,170.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (210.24,170.08) -- (215.23,165.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.66,168.70) rectangle (223.66,173.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.66,168.70) -- (223.66,173.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (218.66,173.69) -- (223.66,168.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.09,172.11) rectangle (232.09,177.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.09,172.11) -- (232.09,177.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (227.09,177.11) -- (232.09,172.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.52,175.36) rectangle (240.52,180.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.52,175.36) -- (240.52,180.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.52,180.35) -- (240.52,175.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.95,178.96) rectangle (248.94,183.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.95,178.96) -- (248.94,183.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.95,183.95) -- (248.94,178.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.38,182.15) rectangle (257.37,187.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.38,182.15) -- (257.37,187.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (252.38,187.14) -- (257.37,182.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (260.80,185.46) rectangle (265.80,190.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (260.80,185.46) -- (265.80,190.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (260.80,190.45) -- (265.80,185.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (269.23,188.38) rectangle (274.23,193.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (269.23,188.38) -- (274.23,193.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (269.23,193.38) -- (274.23,188.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (277.66,191.66) rectangle (282.66,196.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (277.66,191.66) -- (282.66,196.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (277.66,196.65) -- (282.66,191.66);
\definecolor{fillColor}{RGB}{13,8,135}

\path[fill=fillColor] ( 27.31, 63.20) circle (  2.50);

\path[fill=fillColor] ( 35.74, 63.20) circle (  2.50);

\path[fill=fillColor] ( 44.17, 63.20) circle (  2.50);

\path[fill=fillColor] ( 52.60, 63.20) circle (  2.50);

\path[fill=fillColor] ( 61.03, 63.20) circle (  2.50);

\path[fill=fillColor] ( 69.45, 63.20) circle (  2.50);

\path[fill=fillColor] ( 77.88, 63.20) circle (  2.50);

\path[fill=fillColor] ( 86.31, 63.20) circle (  2.50);

\path[fill=fillColor] ( 94.74, 63.20) circle (  2.50);

\path[fill=fillColor] (103.17, 63.20) circle (  2.50);

\path[fill=fillColor] (111.60, 63.20) circle (  2.50);

\path[fill=fillColor] (120.02, 63.20) circle (  2.50);

\path[fill=fillColor] (128.45, 63.20) circle (  2.50);

\path[fill=fillColor] (136.88, 63.20) circle (  2.50);

\path[fill=fillColor] (145.31, 63.20) circle (  2.50);

\path[fill=fillColor] (153.74, 63.20) circle (  2.50);

\path[fill=fillColor] (162.16, 63.20) circle (  2.50);

\path[fill=fillColor] (170.59, 63.20) circle (  2.50);

\path[fill=fillColor] (179.02, 63.20) circle (  2.50);

\path[fill=fillColor] (187.45, 63.20) circle (  2.50);

\path[fill=fillColor] (195.88, 63.20) circle (  2.50);

\path[fill=fillColor] (204.30, 63.20) circle (  2.50);

\path[fill=fillColor] (212.73, 63.20) circle (  2.50);

\path[fill=fillColor] (221.16, 63.20) circle (  2.50);

\path[fill=fillColor] (229.59, 63.20) circle (  2.50);

\path[fill=fillColor] (238.02, 63.20) circle (  2.50);

\path[fill=fillColor] (246.45, 63.20) circle (  2.50);

\path[fill=fillColor] (254.87, 63.20) circle (  2.50);

\path[fill=fillColor] (263.30, 63.20) circle (  2.50);

\path[fill=fillColor] (271.73, 63.20) circle (  2.50);

\path[fill=fillColor] (280.16, 63.20) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.31, 30.69) rectangle (280.16,194.15);
\end{scope}
\begin{scope}
\path[clip] (319.81, 30.69) rectangle (572.66,194.15);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (319.81, 30.69) rectangle (572.66,194.15);
\definecolor{drawColor}{RGB}{13,8,135}

\path[draw=drawColor,line width= 1.1pt,line join=round] (319.81, 58.94) --
	(323.02, 58.94) --
	(326.22, 58.94) --
	(329.42, 58.94) --
	(332.62, 58.94) --
	(335.82, 58.94) --
	(339.02, 58.94) --
	(342.22, 58.94) --
	(345.42, 58.94) --
	(348.62, 58.94) --
	(351.82, 58.94) --
	(355.02, 58.94) --
	(358.22, 58.94) --
	(361.42, 58.94) --
	(364.62, 58.94) --
	(367.82, 58.94) --
	(371.02, 58.94) --
	(374.22, 58.94) --
	(377.43, 58.94) --
	(380.63, 58.94) --
	(383.83, 58.94) --
	(387.03, 58.94) --
	(390.23, 58.94) --
	(393.43, 58.94) --
	(396.63, 58.94) --
	(399.83, 58.94) --
	(403.03, 58.94) --
	(406.23, 58.94) --
	(409.43, 58.94) --
	(412.63, 58.94) --
	(415.83, 58.94) --
	(419.03, 58.94) --
	(422.23, 58.94) --
	(425.43, 58.94) --
	(428.63, 58.94) --
	(431.83, 58.94) --
	(435.04, 58.94) --
	(438.24, 58.94) --
	(441.44, 58.94) --
	(444.64, 58.94) --
	(447.84, 58.94) --
	(451.04, 58.94) --
	(454.24, 58.94) --
	(457.44, 58.94) --
	(460.64, 58.94) --
	(463.84, 58.94) --
	(467.04, 58.94) --
	(470.24, 58.94) --
	(473.44, 58.94) --
	(476.64, 58.94) --
	(479.84, 58.94) --
	(483.04, 58.94) --
	(486.24, 58.94) --
	(489.45, 58.94) --
	(492.65, 58.94) --
	(495.85, 58.94) --
	(499.05, 58.94) --
	(502.25, 58.94) --
	(505.45, 58.94) --
	(508.65, 58.94) --
	(511.85, 58.94) --
	(515.05, 58.94) --
	(518.25, 58.94) --
	(521.45, 58.94) --
	(524.65, 58.94) --
	(527.85, 58.94) --
	(531.05, 58.94) --
	(534.25, 58.94) --
	(537.45, 58.94) --
	(540.65, 58.94) --
	(543.85, 58.94) --
	(547.06, 58.94) --
	(550.26, 58.94) --
	(553.46, 58.94) --
	(556.66, 58.94) --
	(559.86, 58.94) --
	(563.06, 58.94) --
	(566.26, 58.94) --
	(569.46, 58.94) --
	(572.66, 58.94);
\definecolor{drawColor}{RGB}{126,3,168}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 2pt off 2pt ,line join=round] (319.81, 97.64) --
	(323.02, 97.20) --
	(326.22, 96.77) --
	(329.42, 96.35) --
	(332.62, 95.93) --
	(335.82, 95.52) --
	(339.02, 95.12) --
	(342.22, 94.73) --
	(345.42, 94.34) --
	(348.62, 93.96) --
	(351.82, 93.58) --
	(355.02, 93.22) --
	(358.22, 92.86) --
	(361.42, 92.51) --
	(364.62, 92.16) --
	(367.82, 91.82) --
	(371.02, 91.49) --
	(374.22, 91.17) --
	(377.43, 90.85) --
	(380.63, 90.54) --
	(383.83, 90.23) --
	(387.03, 89.94) --
	(390.23, 89.65) --
	(393.43, 89.37) --
	(396.63, 89.10) --
	(399.83, 88.84) --
	(403.03, 88.58) --
	(406.23, 88.32) --
	(409.43, 88.07) --
	(412.63, 87.82) --
	(415.83, 87.59) --
	(419.03, 87.36) --
	(422.23, 87.14) --
	(425.43, 86.92) --
	(428.63, 86.72) --
	(431.83, 86.51) --
	(435.04, 86.31) --
	(438.24, 86.12) --
	(441.44, 85.92) --
	(444.64, 85.72) --
	(447.84, 85.52) --
	(451.04, 85.33) --
	(454.24, 85.14) --
	(457.44, 84.95) --
	(460.64, 84.77) --
	(463.84, 84.59) --
	(467.04, 84.41) --
	(470.24, 84.24) --
	(473.44, 84.07) --
	(476.64, 83.89) --
	(479.84, 83.72) --
	(483.04, 83.55) --
	(486.24, 83.38) --
	(489.45, 83.21) --
	(492.65, 83.04) --
	(495.85, 82.88) --
	(499.05, 82.72) --
	(502.25, 82.56) --
	(505.45, 82.40) --
	(508.65, 82.25) --
	(511.85, 82.09) --
	(515.05, 81.94) --
	(518.25, 81.79) --
	(521.45, 81.64) --
	(524.65, 81.49) --
	(527.85, 81.35) --
	(531.05, 81.20) --
	(534.25, 81.06) --
	(537.45, 80.92) --
	(540.65, 80.79) --
	(543.85, 80.65) --
	(547.06, 80.52) --
	(550.26, 80.39) --
	(553.46, 80.26) --
	(556.66, 80.13) --
	(559.86, 80.00) --
	(563.06, 79.88) --
	(566.26, 79.76) --
	(569.46, 79.64) --
	(572.66, 79.52);
\definecolor{drawColor}{RGB}{204,70,120}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 2pt ,line join=round] (319.81,115.70) --
	(323.02,114.93) --
	(326.22,114.17) --
	(329.42,113.43) --
	(332.62,112.70) --
	(335.82,111.98) --
	(339.02,111.27) --
	(342.22,110.58) --
	(345.42,109.90) --
	(348.62,109.23) --
	(351.82,108.58) --
	(355.02,107.94) --
	(358.22,107.32) --
	(361.42,106.70) --
	(364.62,106.10) --
	(367.82,105.51) --
	(371.02,104.94) --
	(374.22,104.37) --
	(377.43,103.82) --
	(380.63,103.28) --
	(383.83,102.76) --
	(387.03,102.25) --
	(390.23,101.76) --
	(393.43,101.28) --
	(396.63,100.81) --
	(399.83,100.36) --
	(403.03, 99.92) --
	(406.23, 99.48) --
	(409.43, 99.06) --
	(412.63, 98.64) --
	(415.83, 98.23) --
	(419.03, 97.85) --
	(422.23, 97.47) --
	(425.43, 97.11) --
	(428.63, 96.76) --
	(431.83, 96.42) --
	(435.04, 96.09) --
	(438.24, 95.75) --
	(441.44, 95.42) --
	(444.64, 95.09) --
	(447.84, 94.76) --
	(451.04, 94.43) --
	(454.24, 94.12) --
	(457.44, 93.81) --
	(460.64, 93.51) --
	(463.84, 93.21) --
	(467.04, 92.92) --
	(470.24, 92.63) --
	(473.44, 92.34) --
	(476.64, 92.06) --
	(479.84, 91.77) --
	(483.04, 91.49) --
	(486.24, 91.21) --
	(489.45, 90.94) --
	(492.65, 90.67) --
	(495.85, 90.40) --
	(499.05, 90.13) --
	(502.25, 89.87) --
	(505.45, 89.61) --
	(508.65, 89.36) --
	(511.85, 89.11) --
	(515.05, 88.86) --
	(518.25, 88.61) --
	(521.45, 88.37) --
	(524.65, 88.13) --
	(527.85, 87.90) --
	(531.05, 87.67) --
	(534.25, 87.44) --
	(537.45, 87.22) --
	(540.65, 87.00) --
	(543.85, 86.78) --
	(547.06, 86.56) --
	(550.26, 86.35) --
	(553.46, 86.14) --
	(556.66, 85.94) --
	(559.86, 85.74) --
	(563.06, 85.54) --
	(566.26, 85.35) --
	(569.46, 85.16) --
	(572.66, 84.97);
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 4pt off 4pt ,line join=round] (319.81,133.03) --
	(323.02,131.91) --
	(326.22,130.81) --
	(329.42,129.73) --
	(332.62,128.67) --
	(335.82,127.63) --
	(339.02,126.61) --
	(342.22,125.61) --
	(345.42,124.63) --
	(348.62,123.67) --
	(351.82,122.72) --
	(355.02,121.80) --
	(358.22,120.90) --
	(361.42,120.02) --
	(364.62,119.15) --
	(367.82,118.31) --
	(371.02,117.48) --
	(374.22,116.67) --
	(377.43,115.88) --
	(380.63,115.11) --
	(383.83,114.36) --
	(387.03,113.63) --
	(390.23,112.93) --
	(393.43,112.25) --
	(396.63,111.59) --
	(399.83,110.95) --
	(403.03,110.32) --
	(406.23,109.71) --
	(409.43,109.11) --
	(412.63,108.52) --
	(415.83,107.94) --
	(419.03,107.40) --
	(422.23,106.87) --
	(425.43,106.37) --
	(428.63,105.88) --
	(431.83,105.40) --
	(435.04,104.94) --
	(438.24,104.47) --
	(441.44,104.01) --
	(444.64,103.55) --
	(447.84,103.09) --
	(451.04,102.64) --
	(454.24,102.20) --
	(457.44,101.77) --
	(460.64,101.35) --
	(463.84,100.94) --
	(467.04,100.54) --
	(470.24,100.14) --
	(473.44, 99.74) --
	(476.64, 99.35) --
	(479.84, 98.96) --
	(483.04, 98.57) --
	(486.24, 98.18) --
	(489.45, 97.81) --
	(492.65, 97.43) --
	(495.85, 97.07) --
	(499.05, 96.71) --
	(502.25, 96.35) --
	(505.45, 96.00) --
	(508.65, 95.65) --
	(511.85, 95.31) --
	(515.05, 94.97) --
	(518.25, 94.63) --
	(521.45, 94.31) --
	(524.65, 93.98) --
	(527.85, 93.66) --
	(531.05, 93.35) --
	(534.25, 93.04) --
	(537.45, 92.74) --
	(540.65, 92.44) --
	(543.85, 92.14) --
	(547.06, 91.86) --
	(550.26, 91.57) --
	(553.46, 91.29) --
	(556.66, 91.02) --
	(559.86, 90.75) --
	(563.06, 90.48) --
	(566.26, 90.23) --
	(569.46, 89.97) --
	(572.66, 89.72);
\definecolor{drawColor}{RGB}{216,224,30}

\path[draw=drawColor,line width= 1.1pt,dash pattern=on 1pt off 3pt ,line join=round] (319.81,149.67) --
	(323.02,148.19) --
	(326.22,146.73) --
	(329.42,145.31) --
	(332.62,143.91) --
	(335.82,142.53) --
	(339.02,141.18) --
	(342.22,139.86) --
	(345.42,138.57) --
	(348.62,137.30) --
	(351.82,136.06) --
	(355.02,134.85) --
	(358.22,133.66) --
	(361.42,132.50) --
	(364.62,131.37) --
	(367.82,130.26) --
	(371.02,129.17) --
	(374.22,128.11) --
	(377.43,127.07) --
	(380.63,126.06) --
	(383.83,125.09) --
	(387.03,124.14) --
	(390.23,123.23) --
	(393.43,122.35) --
	(396.63,121.49) --
	(399.83,120.66) --
	(403.03,119.84) --
	(406.23,119.05) --
	(409.43,118.27) --
	(412.63,117.51) --
	(415.83,116.77) --
	(419.03,116.06) --
	(422.23,115.39) --
	(425.43,114.74) --
	(428.63,114.12) --
	(431.83,113.51) --
	(435.04,112.91) --
	(438.24,112.32) --
	(441.44,111.73) --
	(444.64,111.14) --
	(447.84,110.55) --
	(451.04,109.97) --
	(454.24,109.41) --
	(457.44,108.87) --
	(460.64,108.33) --
	(463.84,107.81) --
	(467.04,107.30) --
	(470.24,106.80) --
	(473.44,106.29) --
	(476.64,105.80) --
	(479.84,105.30) --
	(483.04,104.81) --
	(486.24,104.32) --
	(489.45,103.84) --
	(492.65,103.37) --
	(495.85,102.91) --
	(499.05,102.46) --
	(502.25,102.01) --
	(505.45,101.57) --
	(508.65,101.13) --
	(511.85,100.70) --
	(515.05,100.27) --
	(518.25, 99.86) --
	(521.45, 99.44) --
	(524.65, 99.04) --
	(527.85, 98.64) --
	(531.05, 98.25) --
	(534.25, 97.86) --
	(537.45, 97.48) --
	(540.65, 97.11) --
	(543.85, 96.75) --
	(547.06, 96.39) --
	(550.26, 96.04) --
	(553.46, 95.69) --
	(556.66, 95.35) --
	(559.86, 95.02) --
	(563.06, 94.69) --
	(566.26, 94.37) --
	(569.46, 94.06) --
	(572.66, 93.75);
\definecolor{fillColor}{RGB}{126,3,168}

\path[fill=fillColor] (319.81,101.62) --
	(323.18, 95.79) --
	(316.45, 95.79) --
	cycle;

\path[fill=fillColor] (328.24,100.41) --
	(331.61, 94.58) --
	(324.88, 94.58) --
	cycle;

\path[fill=fillColor] (336.67, 99.32) --
	(340.03, 93.49) --
	(333.31, 93.49) --
	cycle;

\path[fill=fillColor] (345.10, 98.18) --
	(348.46, 92.35) --
	(341.74, 92.35) --
	cycle;

\path[fill=fillColor] (353.53, 97.16) --
	(356.89, 91.33) --
	(350.16, 91.33) --
	cycle;

\path[fill=fillColor] (361.96, 96.27) --
	(365.32, 90.44) --
	(358.59, 90.44) --
	cycle;

\path[fill=fillColor] (370.38, 95.43) --
	(373.75, 89.61) --
	(367.02, 89.61) --
	cycle;

\path[fill=fillColor] (378.81, 94.63) --
	(382.18, 88.80) --
	(375.45, 88.80) --
	cycle;

\path[fill=fillColor] (387.24, 93.76) --
	(390.60, 87.94) --
	(383.88, 87.94) --
	cycle;

\path[fill=fillColor] (395.67, 92.98) --
	(399.03, 87.16) --
	(392.30, 87.16) --
	cycle;

\path[fill=fillColor] (404.10, 92.29) --
	(407.46, 86.46) --
	(400.73, 86.46) --
	cycle;

\path[fill=fillColor] (412.52, 91.68) --
	(415.89, 85.86) --
	(409.16, 85.86) --
	cycle;

\path[fill=fillColor] (420.95, 91.09) --
	(424.32, 85.26) --
	(417.59, 85.26) --
	cycle;

\path[fill=fillColor] (429.38, 90.57) --
	(432.74, 84.74) --
	(426.02, 84.74) --
	cycle;

\path[fill=fillColor] (437.81, 90.05) --
	(441.17, 84.22) --
	(434.45, 84.22) --
	cycle;

\path[fill=fillColor] (446.24, 89.54) --
	(449.60, 83.72) --
	(442.87, 83.72) --
	cycle;

\path[fill=fillColor] (454.67, 89.09) --
	(458.03, 83.26) --
	(451.30, 83.26) --
	cycle;

\path[fill=fillColor] (463.09, 88.59) --
	(466.46, 82.77) --
	(459.73, 82.77) --
	cycle;

\path[fill=fillColor] (471.52, 88.04) --
	(474.89, 82.21) --
	(468.16, 82.21) --
	cycle;

\path[fill=fillColor] (479.95, 87.56) --
	(483.31, 81.73) --
	(476.59, 81.73) --
	cycle;

\path[fill=fillColor] (488.38, 87.14) --
	(491.74, 81.31) --
	(485.01, 81.31) --
	cycle;

\path[fill=fillColor] (496.81, 86.70) --
	(500.17, 80.87) --
	(493.44, 80.87) --
	cycle;

\path[fill=fillColor] (505.23, 86.30) --
	(508.60, 80.48) --
	(501.87, 80.48) --
	cycle;

\path[fill=fillColor] (513.66, 85.86) --
	(517.03, 80.04) --
	(510.30, 80.04) --
	cycle;

\path[fill=fillColor] (522.09, 85.49) --
	(525.45, 79.67) --
	(518.73, 79.67) --
	cycle;

\path[fill=fillColor] (530.52, 85.15) --
	(533.88, 79.33) --
	(527.16, 79.33) --
	cycle;

\path[fill=fillColor] (538.95, 84.78) --
	(542.31, 78.96) --
	(535.58, 78.96) --
	cycle;

\path[fill=fillColor] (547.38, 84.39) --
	(550.74, 78.57) --
	(544.01, 78.57) --
	cycle;

\path[fill=fillColor] (555.80, 84.03) --
	(559.17, 78.21) --
	(552.44, 78.21) --
	cycle;

\path[fill=fillColor] (564.23, 83.73) --
	(567.60, 77.90) --
	(560.87, 77.90) --
	cycle;

\path[fill=fillColor] (572.66, 83.38) --
	(576.02, 77.56) --
	(569.30, 77.56) --
	cycle;
\definecolor{fillColor}{RGB}{204,70,120}

\path[fill=fillColor] (317.32,113.39) --
	(322.31,113.39) --
	(322.31,118.38) --
	(317.32,118.38) --
	cycle;

\path[fill=fillColor] (325.75,111.23) --
	(330.74,111.23) --
	(330.74,116.23) --
	(325.75,116.23) --
	cycle;

\path[fill=fillColor] (334.17,109.33) --
	(339.17,109.33) --
	(339.17,114.33) --
	(334.17,114.33) --
	cycle;

\path[fill=fillColor] (342.60,107.33) --
	(347.60,107.33) --
	(347.60,112.32) --
	(342.60,112.32) --
	cycle;

\path[fill=fillColor] (351.03,105.51) --
	(356.03,105.51) --
	(356.03,110.51) --
	(351.03,110.51) --
	cycle;

\path[fill=fillColor] (359.46,103.96) --
	(364.45,103.96) --
	(364.45,108.96) --
	(359.46,108.96) --
	cycle;

\path[fill=fillColor] (367.89,102.51) --
	(372.88,102.51) --
	(372.88,107.50) --
	(367.89,107.50) --
	cycle;

\path[fill=fillColor] (376.31,101.12) --
	(381.31,101.12) --
	(381.31,106.11) --
	(376.31,106.11) --
	cycle;

\path[fill=fillColor] (384.74, 99.64) --
	(389.74, 99.64) --
	(389.74,104.64) --
	(384.74,104.64) --
	cycle;

\path[fill=fillColor] (393.17, 98.32) --
	(398.17, 98.32) --
	(398.17,103.31) --
	(393.17,103.31) --
	cycle;

\path[fill=fillColor] (401.60, 97.13) --
	(406.59, 97.13) --
	(406.59,102.13) --
	(401.60,102.13) --
	cycle;

\path[fill=fillColor] (410.03, 96.11) --
	(415.02, 96.11) --
	(415.02,101.11) --
	(410.03,101.11) --
	cycle;

\path[fill=fillColor] (418.46, 95.11) --
	(423.45, 95.11) --
	(423.45,100.10) --
	(418.46,100.10) --
	cycle;

\path[fill=fillColor] (426.88, 94.21) --
	(431.88, 94.21) --
	(431.88, 99.20) --
	(426.88, 99.20) --
	cycle;

\path[fill=fillColor] (435.31, 93.32) --
	(440.31, 93.32) --
	(440.31, 98.32) --
	(435.31, 98.32) --
	cycle;

\path[fill=fillColor] (443.74, 92.49) --
	(448.73, 92.49) --
	(448.73, 97.48) --
	(443.74, 97.48) --
	cycle;

\path[fill=fillColor] (452.17, 91.72) --
	(457.16, 91.72) --
	(457.16, 96.71) --
	(452.17, 96.71) --
	cycle;

\path[fill=fillColor] (460.60, 90.90) --
	(465.59, 90.90) --
	(465.59, 95.89) --
	(460.60, 95.89) --
	cycle;

\path[fill=fillColor] (469.02, 90.00) --
	(474.02, 90.00) --
	(474.02, 94.99) --
	(469.02, 94.99) --
	cycle;

\path[fill=fillColor] (477.45, 89.22) --
	(482.45, 89.22) --
	(482.45, 94.21) --
	(477.45, 94.21) --
	cycle;

\path[fill=fillColor] (485.88, 88.51) --
	(490.88, 88.51) --
	(490.88, 93.51) --
	(485.88, 93.51) --
	cycle;

\path[fill=fillColor] (494.31, 87.80) --
	(499.30, 87.80) --
	(499.30, 92.79) --
	(494.31, 92.79) --
	cycle;

\path[fill=fillColor] (502.74, 87.15) --
	(507.73, 87.15) --
	(507.73, 92.14) --
	(502.74, 92.14) --
	cycle;

\path[fill=fillColor] (511.17, 86.43) --
	(516.16, 86.43) --
	(516.16, 91.43) --
	(511.17, 91.43) --
	cycle;

\path[fill=fillColor] (519.59, 85.84) --
	(524.59, 85.84) --
	(524.59, 90.83) --
	(519.59, 90.83) --
	cycle;

\path[fill=fillColor] (528.02, 85.28) --
	(533.02, 85.28) --
	(533.02, 90.27) --
	(528.02, 90.27) --
	cycle;

\path[fill=fillColor] (536.45, 84.68) --
	(541.44, 84.68) --
	(541.44, 89.67) --
	(536.45, 89.67) --
	cycle;

\path[fill=fillColor] (544.88, 84.07) --
	(549.87, 84.07) --
	(549.87, 89.07) --
	(544.88, 89.07) --
	cycle;

\path[fill=fillColor] (553.31, 83.47) --
	(558.30, 83.47) --
	(558.30, 88.47) --
	(553.31, 88.47) --
	cycle;

\path[fill=fillColor] (561.73, 82.98) --
	(566.73, 82.98) --
	(566.73, 87.97) --
	(561.73, 87.97) --
	cycle;

\path[fill=fillColor] (570.16, 82.43) --
	(575.16, 82.43) --
	(575.16, 87.42) --
	(570.16, 87.42) --
	cycle;
\definecolor{drawColor}{RGB}{248,148,65}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (316.28,133.33) -- (323.35,133.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (319.81,129.79) -- (319.81,136.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (324.71,130.17) -- (331.78,130.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.24,126.64) -- (328.24,133.70);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.14,127.42) -- (340.20,127.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (336.67,123.89) -- (336.67,130.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (341.57,124.53) -- (348.63,124.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.10,121.00) -- (345.10,128.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.00,121.87) -- (357.06,121.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (353.53,118.34) -- (353.53,125.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (358.42,119.65) -- (365.49,119.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (361.96,116.11) -- (361.96,123.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.85,117.55) -- (373.92,117.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (370.38,114.02) -- (370.38,121.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (375.28,115.54) -- (382.34,115.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (378.81,112.01) -- (378.81,119.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.71,113.47) -- (390.77,113.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.24,109.94) -- (387.24,117.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (392.14,111.60) -- (399.20,111.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.67,108.06) -- (395.67,115.13);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (400.56,109.91) -- (407.63,109.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (404.10,106.37) -- (404.10,113.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (408.99,108.50) -- (416.06,108.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.52,104.97) -- (412.52,112.03);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (417.42,107.09) -- (424.48,107.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (420.95,103.56) -- (420.95,110.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (425.85,105.80) -- (432.91,105.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (429.38,102.27) -- (429.38,109.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (434.28,104.55) -- (441.34,104.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (437.81,101.02) -- (437.81,108.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (442.71,103.41) -- (449.77,103.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (446.24, 99.87) -- (446.24,106.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (451.13,102.33) -- (458.20,102.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.67, 98.80) -- (454.67,105.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (459.56,101.17) -- (466.63,101.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.09, 97.64) -- (463.09,104.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (467.99, 99.96) -- (475.05, 99.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (471.52, 96.42) -- (471.52,103.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.42, 98.89) -- (483.48, 98.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.95, 95.35) -- (479.95,102.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (484.85, 97.92) -- (491.91, 97.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (488.38, 94.38) -- (488.38,101.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (493.27, 96.93) -- (500.34, 96.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (496.81, 93.40) -- (496.81,100.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.70, 96.04) -- (508.77, 96.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.23, 92.51) -- (505.23, 99.57);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (510.13, 95.07) -- (517.19, 95.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (513.66, 91.53) -- (513.66, 98.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (518.56, 94.27) -- (525.62, 94.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (522.09, 90.73) -- (522.09, 97.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.99, 93.49) -- (534.05, 93.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (530.52, 89.96) -- (530.52, 97.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.42, 92.69) -- (542.48, 92.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (538.95, 89.16) -- (538.95, 96.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (543.84, 91.90) -- (550.91, 91.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.38, 88.37) -- (547.38, 95.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (552.27, 91.05) -- (559.34, 91.05);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (555.80, 87.52) -- (555.80, 94.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (560.70, 90.38) -- (567.76, 90.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (564.23, 86.85) -- (564.23, 93.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (569.13, 89.65) -- (576.19, 89.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (572.66, 86.12) -- (572.66, 93.18);
\definecolor{drawColor}{RGB}{216,224,30}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (317.32,147.59) rectangle (322.31,152.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (317.32,147.59) -- (322.31,152.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (317.32,152.58) -- (322.31,147.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (325.75,143.39) rectangle (330.74,148.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (325.75,143.39) -- (330.74,148.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (325.75,148.38) -- (330.74,143.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.17,139.76) rectangle (339.17,144.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.17,139.76) -- (339.17,144.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.17,144.75) -- (339.17,139.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.60,135.95) rectangle (347.60,140.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.60,135.95) -- (347.60,140.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (342.60,140.94) -- (347.60,135.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.03,132.41) rectangle (356.03,137.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.03,132.41) -- (356.03,137.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (351.03,137.40) -- (356.03,132.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.46,129.50) rectangle (364.45,134.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.46,129.50) -- (364.45,134.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (359.46,134.50) -- (364.45,129.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.89,126.75) rectangle (372.88,131.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.89,126.75) -- (372.88,131.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (367.89,131.74) -- (372.88,126.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.31,124.08) rectangle (381.31,129.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.31,124.08) -- (381.31,129.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.31,129.08) -- (381.31,124.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.74,121.44) rectangle (389.74,126.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.74,121.44) -- (389.74,126.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.74,126.43) -- (389.74,121.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.17,118.99) rectangle (398.17,123.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.17,118.99) -- (398.17,123.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (393.17,123.99) -- (398.17,118.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.60,116.78) rectangle (406.59,121.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.60,116.78) -- (406.59,121.77);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.60,121.77) -- (406.59,116.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.03,115.02) rectangle (415.02,120.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.03,115.02) -- (415.02,120.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.03,120.02) -- (415.02,115.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.46,113.20) rectangle (423.45,118.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.46,113.20) -- (423.45,118.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (418.46,118.19) -- (423.45,113.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.88,111.53) rectangle (431.88,116.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.88,111.53) -- (431.88,116.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.88,116.52) -- (431.88,111.53);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.31,109.90) rectangle (440.31,114.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.31,109.90) -- (440.31,114.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.31,114.89) -- (440.31,109.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.74,108.47) rectangle (448.73,113.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.74,108.47) -- (448.73,113.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.74,113.47) -- (448.73,108.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.17,107.09) rectangle (457.16,112.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.17,107.09) -- (457.16,112.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.17,112.08) -- (457.16,107.09);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (460.60,105.59) rectangle (465.59,110.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (460.60,105.59) -- (465.59,110.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (460.60,110.58) -- (465.59,105.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.02,104.07) rectangle (474.02,109.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.02,104.07) -- (474.02,109.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.02,109.07) -- (474.02,104.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.45,102.72) rectangle (482.45,107.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.45,102.72) -- (482.45,107.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.45,107.72) -- (482.45,102.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.88,101.49) rectangle (490.88,106.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.88,101.49) -- (490.88,106.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.88,106.49) -- (490.88,101.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.31,100.25) rectangle (499.30,105.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.31,100.25) -- (499.30,105.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.31,105.25) -- (499.30,100.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.74, 99.11) rectangle (507.73,104.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.74, 99.11) -- (507.73,104.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (502.74,104.10) -- (507.73, 99.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.17, 97.90) rectangle (516.16,102.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.17, 97.90) -- (516.16,102.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (511.17,102.90) -- (516.16, 97.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.59, 96.90) rectangle (524.59,101.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.59, 96.90) -- (524.59,101.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (519.59,101.90) -- (524.59, 96.90);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.02, 95.91) rectangle (533.02,100.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.02, 95.91) -- (533.02,100.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (528.02,100.91) -- (533.02, 95.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.45, 94.94) rectangle (541.44, 99.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.45, 94.94) -- (541.44, 99.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.45, 99.93) -- (541.44, 94.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.88, 94.00) rectangle (549.87, 98.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.88, 94.00) -- (549.87, 98.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (544.88, 98.99) -- (549.87, 94.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.31, 92.89) rectangle (558.30, 97.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.31, 92.89) -- (558.30, 97.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (553.31, 97.89) -- (558.30, 92.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.73, 92.04) rectangle (566.73, 97.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.73, 92.04) -- (566.73, 97.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (561.73, 97.04) -- (566.73, 92.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16, 91.16) rectangle (575.16, 96.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16, 91.16) -- (575.16, 96.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (570.16, 96.15) -- (575.16, 91.16);
\definecolor{fillColor}{RGB}{13,8,135}

\path[fill=fillColor] (319.81, 58.94) circle (  2.50);

\path[fill=fillColor] (328.24, 58.94) circle (  2.50);

\path[fill=fillColor] (336.67, 58.94) circle (  2.50);

\path[fill=fillColor] (345.10, 58.94) circle (  2.50);

\path[fill=fillColor] (353.53, 58.94) circle (  2.50);

\path[fill=fillColor] (361.96, 58.94) circle (  2.50);

\path[fill=fillColor] (370.38, 58.94) circle (  2.50);

\path[fill=fillColor] (378.81, 58.94) circle (  2.50);

\path[fill=fillColor] (387.24, 58.94) circle (  2.50);

\path[fill=fillColor] (395.67, 58.94) circle (  2.50);

\path[fill=fillColor] (404.10, 58.94) circle (  2.50);

\path[fill=fillColor] (412.52, 58.94) circle (  2.50);

\path[fill=fillColor] (420.95, 58.94) circle (  2.50);

\path[fill=fillColor] (429.38, 58.94) circle (  2.50);

\path[fill=fillColor] (437.81, 58.94) circle (  2.50);

\path[fill=fillColor] (446.24, 58.94) circle (  2.50);

\path[fill=fillColor] (454.67, 58.94) circle (  2.50);

\path[fill=fillColor] (463.09, 58.94) circle (  2.50);

\path[fill=fillColor] (471.52, 58.94) circle (  2.50);

\path[fill=fillColor] (479.95, 58.94) circle (  2.50);

\path[fill=fillColor] (488.38, 58.94) circle (  2.50);

\path[fill=fillColor] (496.81, 58.94) circle (  2.50);

\path[fill=fillColor] (505.23, 58.94) circle (  2.50);

\path[fill=fillColor] (513.66, 58.94) circle (  2.50);

\path[fill=fillColor] (522.09, 58.94) circle (  2.50);

\path[fill=fillColor] (530.52, 58.94) circle (  2.50);

\path[fill=fillColor] (538.95, 58.94) circle (  2.50);

\path[fill=fillColor] (547.38, 58.94) circle (  2.50);

\path[fill=fillColor] (555.80, 58.94) circle (  2.50);

\path[fill=fillColor] (564.23, 58.94) circle (  2.50);

\path[fill=fillColor] (572.66, 58.94) circle (  2.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (319.81, 30.69) rectangle (572.66,194.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31,244.75) --
	( 27.31,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.60,244.75) --
	(111.60,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.88,244.75) --
	(195.88,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (111.60,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (195.88,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (153.74,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (319.81,244.75) --
	(319.81,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.10,244.75) --
	(404.10,247.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.38,244.75) --
	(488.38,247.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (319.81,236.49) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (404.10,236.49) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (488.38,236.49) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (446.24,224.45) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.31, 27.94) --
	( 27.31, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.60, 27.94) --
	(111.60, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.88, 27.94) --
	(195.88, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.31, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (111.60, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (195.88, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (153.74,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (319.81, 27.94) --
	(319.81, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.10, 27.94) --
	(404.10, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.38, 27.94) --
	(488.38, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (319.81, 19.68) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (404.10, 19.68) {0.1};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (488.38, 19.68) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,433.62);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (446.24,  7.64) {Exponential growth rate $r$ (1/day)};
\end{scope}
\end{tikzpicture}

\end{document}