### 2020-10-17 forward_corr.tex
done

% Created by tikzDevice version 0.12.3 on 2020-10-16 22:00:47
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,578.16);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{gray}{0.95}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,  0.00) -- (  0.00,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 30.43,  0.00) -- ( 30.43,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 60.86,  0.00) -- ( 60.86,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.29,  0.00) -- ( 91.29,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (121.72,  0.00) -- (121.72,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.15,  0.00) -- (152.15,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.58,  0.00) -- (182.58,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (213.01,  0.00) -- (213.01,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (243.44,  0.00) -- (243.44,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.87,  0.00) -- (273.87,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.29,  0.00) -- (304.29,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.72,  0.00) -- (334.72,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.15,  0.00) -- (365.15,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.58,  0.00) -- (395.58,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (426.01,  0.00) -- (426.01,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (456.44,  0.00) -- (456.44,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (486.87,  0.00) -- (486.87,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (517.30,  0.00) -- (517.30,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (547.73,  0.00) -- (547.73,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (578.16,  0.00) -- (578.16,578.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,  0.00) -- (578.16,  0.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 30.43) -- (578.16, 30.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 60.86) -- (578.16, 60.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00, 91.29) -- (578.16, 91.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,121.72) -- (578.16,121.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,152.15) -- (578.16,152.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,182.58) -- (578.16,182.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,213.01) -- (578.16,213.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,243.44) -- (578.16,243.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,273.87) -- (578.16,273.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,304.29) -- (578.16,304.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,334.72) -- (578.16,334.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,365.15) -- (578.16,365.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,395.58) -- (578.16,395.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,426.01) -- (578.16,426.01);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,456.44) -- (578.16,456.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,486.87) -- (578.16,486.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,517.30) -- (578.16,517.30);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,547.73) -- (578.16,547.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (  0.00,578.16) -- (578.16,578.16);
\end{scope}
\begin{scope}
\path[clip] (  0.00,412.97) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,412.97) rectangle (578.16,578.16);
\end{scope}
\begin{scope}
\path[clip] ( 40.70,443.99) rectangle (572.66,555.28);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 40.70,443.99) rectangle (572.66,555.28);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.87) --
	( 49.28,444.83) --
	( 57.86,444.83) --
	( 66.44,445.43) --
	( 75.02,445.66) --
	( 83.60,446.22) --
	( 92.18,446.31) --
	(100.76,447.61) --
	(109.34,447.61) --
	(117.92,448.49) --
	(126.50,449.19) --
	(135.08,450.49) --
	(143.66,451.74) --
	(152.24,454.24) --
	(160.82,455.12) --
	(169.40,458.60) --
	(177.98,460.51) --
	(186.56,463.75) --
	(195.14,469.23) --
	(203.72,472.61) --
	(212.30,477.86) --
	(220.88,481.80) --
	(229.46,489.82) --
	(238.04,497.94) --
	(246.62,502.53) --
	(255.20,508.10) --
	(263.78,517.98) --
	(272.36,523.55) --
	(280.94,528.79) --
	(289.52,534.22) --
	(298.10,533.76) --
	(306.68,535.43) --
	(315.26,531.67) --
	(323.84,527.31) --
	(332.42,522.81) --
	(341.00,515.11) --
	(349.58,506.34) --
	(358.16,499.06) --
	(366.74,490.57) --
	(375.32,481.01) --
	(383.90,474.65) --
	(392.48,471.50) --
	(401.06,463.52) --
	(409.64,462.18) --
	(418.22,456.66) --
	(426.80,454.47) --
	(435.38,452.02) --
	(443.96,450.86) --
	(452.54,449.28) --
	(461.12,448.63) --
	(469.70,447.61) --
	(478.28,446.68) --
	(486.86,445.61) --
	(495.44,445.89) --
	(504.02,445.15) --
	(512.60,444.87) --
	(521.18,444.55) --
	(529.76,444.50) --
	(538.34,444.36) --
	(546.92,444.13) --
	(555.50,444.22) --
	(564.08,444.08) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.41) --
	( 49.28,444.64) --
	( 57.86,445.01) --
	( 66.44,444.87) --
	( 75.02,445.01) --
	( 83.60,445.57) --
	( 92.18,446.45) --
	(100.76,445.89) --
	(109.34,447.05) --
	(117.92,447.56) --
	(126.50,448.77) --
	(135.08,449.00) --
	(143.66,449.93) --
	(152.24,451.46) --
	(160.82,453.08) --
	(169.40,455.17) --
	(177.98,458.09) --
	(186.56,459.35) --
	(195.14,462.08) --
	(203.72,467.84) --
	(212.30,470.66) --
	(220.88,478.18) --
	(229.46,478.27) --
	(238.04,487.04) --
	(246.62,496.09) --
	(255.20,501.51) --
	(263.78,504.58) --
	(272.36,516.17) --
	(280.94,520.53) --
	(289.52,524.89) --
	(298.10,534.68) --
	(306.68,532.09) --
	(315.26,534.31) --
	(323.84,530.65) --
	(332.42,530.55) --
	(341.00,523.78) --
	(349.58,515.90) --
	(358.16,510.10) --
	(366.74,500.63) --
	(375.32,493.30) --
	(383.90,484.26) --
	(392.48,476.37) --
	(401.06,471.27) --
	(409.64,466.86) --
	(418.22,462.92) --
	(426.80,458.60) --
	(435.38,456.05) --
	(443.96,452.94) --
	(452.54,451.23) --
	(461.12,449.46) --
	(469.70,447.98) --
	(478.28,447.05) --
	(486.86,447.24) --
	(495.44,446.12) --
	(504.02,445.75) --
	(512.60,445.06) --
	(521.18,445.01) --
	(529.76,444.83) --
	(538.34,444.50) --
	(546.92,444.41) --
	(555.50,444.50) --
	(564.08,444.36) --
	(572.66,444.18);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.41) --
	( 49.28,444.59) --
	( 57.86,444.73) --
	( 66.44,445.10) --
	( 75.02,444.73) --
	( 83.60,445.52) --
	( 92.18,446.12) --
	(100.76,445.89) --
	(109.34,446.77) --
	(117.92,447.10) --
	(126.50,448.82) --
	(135.08,448.77) --
	(143.66,449.70) --
	(152.24,450.39) --
	(160.82,451.55) --
	(169.40,455.26) --
	(177.98,456.61) --
	(186.56,458.79) --
	(195.14,461.43) --
	(203.72,465.75) --
	(212.30,469.46) --
	(220.88,473.82) --
	(229.46,480.50) --
	(238.04,485.05) --
	(246.62,492.56) --
	(255.20,499.15) --
	(263.78,503.14) --
	(272.36,513.16) --
	(280.94,518.86) --
	(289.52,524.76) --
	(298.10,531.20) --
	(306.68,529.86) --
	(315.26,535.47) --
	(323.84,533.52) --
	(332.42,528.79) --
	(341.00,526.89) --
	(349.58,516.03) --
	(358.16,513.25) --
	(366.74,505.55) --
	(375.32,492.61) --
	(383.90,488.01) --
	(392.48,477.48) --
	(401.06,473.12) --
	(409.64,467.88) --
	(418.22,465.14) --
	(426.80,459.39) --
	(435.38,456.01) --
	(443.96,455.54) --
	(452.54,452.11) --
	(461.12,450.86) --
	(469.70,449.33) --
	(478.28,447.70) --
	(486.86,447.01) --
	(495.44,446.22) --
	(504.02,446.59) --
	(512.60,445.71) --
	(521.18,444.96) --
	(529.76,445.20) --
	(538.34,444.55) --
	(546.92,444.55) --
	(555.50,444.87) --
	(564.08,444.36) --
	(572.66,444.27);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.55) --
	( 49.28,444.64) --
	( 57.86,444.96) --
	( 66.44,444.87) --
	( 75.02,445.34) --
	( 83.60,445.57) --
	( 92.18,446.08) --
	(100.76,445.89) --
	(109.34,447.56) --
	(117.92,447.84) --
	(126.50,448.58) --
	(135.08,449.05) --
	(143.66,451.00) --
	(152.24,450.72) --
	(160.82,453.36) --
	(169.40,455.91) --
	(177.98,457.91) --
	(186.56,461.25) --
	(195.14,463.80) --
	(203.72,465.98) --
	(212.30,471.92) --
	(220.88,479.43) --
	(229.46,481.98) --
	(238.04,489.17) --
	(246.62,495.99) --
	(255.20,499.61) --
	(263.78,509.82) --
	(272.36,518.08) --
	(280.94,521.00) --
	(289.52,524.71) --
	(298.10,531.99) --
	(306.68,533.57) --
	(315.26,532.32) --
	(323.84,533.48) --
	(332.42,529.30) --
	(341.00,523.09) --
	(349.58,515.66) --
	(358.16,509.59) --
	(366.74,500.40) --
	(375.32,488.94) --
	(383.90,483.75) --
	(392.48,475.81) --
	(401.06,472.34) --
	(409.64,465.14) --
	(418.22,461.20) --
	(426.80,457.49) --
	(435.38,454.85) --
	(443.96,453.22) --
	(452.54,450.44) --
	(461.12,448.54) --
	(469.70,447.89) --
	(478.28,446.96) --
	(486.86,447.33) --
	(495.44,445.71) --
	(504.02,444.92) --
	(512.60,445.43) --
	(521.18,445.10) --
	(529.76,444.83) --
	(538.34,444.87) --
	(546.92,444.32) --
	(555.50,444.27) --
	(564.08,443.99) --
	(572.66,444.18);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.32) --
	( 49.28,444.73) --
	( 57.86,444.83) --
	( 66.44,445.29) --
	( 75.02,445.10) --
	( 83.60,445.43) --
	( 92.18,446.08) --
	(100.76,446.54) --
	(109.34,446.82) --
	(117.92,448.26) --
	(126.50,448.63) --
	(135.08,449.42) --
	(143.66,450.44) --
	(152.24,451.92) --
	(160.82,454.38) --
	(169.40,456.28) --
	(177.98,457.21) --
	(186.56,461.43) --
	(195.14,465.10) --
	(203.72,468.62) --
	(212.30,472.06) --
	(220.88,476.60) --
	(229.46,482.68) --
	(238.04,491.35) --
	(246.62,496.64) --
	(255.20,505.78) --
	(263.78,511.49) --
	(272.36,516.17) --
	(280.94,525.50) --
	(289.52,528.14) --
	(298.10,530.69) --
	(306.68,535.15) --
	(315.26,535.66) --
	(323.84,530.42) --
	(332.42,526.05) --
	(341.00,520.02) --
	(349.58,514.46) --
	(358.16,504.16) --
	(366.74,498.50) --
	(375.32,490.84) --
	(383.90,481.06) --
	(392.48,474.24) --
	(401.06,468.35) --
	(409.64,466.49) --
	(418.22,461.11) --
	(426.80,456.70) --
	(435.38,454.34) --
	(443.96,452.67) --
	(452.54,451.92) --
	(461.12,449.74) --
	(469.70,448.21) --
	(478.28,447.28) --
	(486.86,447.01) --
	(495.44,445.71) --
	(504.02,445.52) --
	(512.60,445.20) --
	(521.18,445.34) --
	(529.76,444.55) --
	(538.34,444.41) --
	(546.92,444.18) --
	(555.50,444.13) --
	(564.08,444.27) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.50) --
	( 49.28,444.78) --
	( 57.86,445.24) --
	( 66.44,445.10) --
	( 75.02,445.48) --
	( 83.60,445.85) --
	( 92.18,446.50) --
	(100.76,446.82) --
	(109.34,447.47) --
	(117.92,448.12) --
	(126.50,449.51) --
	(135.08,450.53) --
	(143.66,450.95) --
	(152.24,454.01) --
	(160.82,454.89) --
	(169.40,458.70) --
	(177.98,459.67) --
	(186.56,462.45) --
	(195.14,467.09) --
	(203.72,471.59) --
	(212.30,476.37) --
	(220.88,480.13) --
	(229.46,488.57) --
	(238.04,494.65) --
	(246.62,501.47) --
	(255.20,509.54) --
	(263.78,516.68) --
	(272.36,522.34) --
	(280.94,526.33) --
	(289.52,531.20) --
	(298.10,533.29) --
	(306.68,535.47) --
	(315.26,532.78) --
	(323.84,529.53) --
	(332.42,521.74) --
	(341.00,514.69) --
	(349.58,505.41) --
	(358.16,500.03) --
	(366.74,492.28) --
	(375.32,483.33) --
	(383.90,477.99) --
	(392.48,471.45) --
	(401.06,466.40) --
	(409.64,462.55) --
	(418.22,458.42) --
	(426.80,455.40) --
	(435.38,453.83) --
	(443.96,451.55) --
	(452.54,450.49) --
	(461.12,448.30) --
	(469.70,448.07) --
	(478.28,447.38) --
	(486.86,446.45) --
	(495.44,445.34) --
	(504.02,445.43) --
	(512.60,445.29) --
	(521.18,445.06) --
	(529.76,444.50) --
	(538.34,444.50) --
	(546.92,444.36) --
	(555.50,444.22) --
	(564.08,444.18) --
	(572.66,444.18);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.32) --
	( 49.28,444.64) --
	( 57.86,445.24) --
	( 66.44,445.06) --
	( 75.02,445.29) --
	( 83.60,445.80) --
	( 92.18,446.31) --
	(100.76,446.91) --
	(109.34,447.52) --
	(117.92,447.52) --
	(126.50,448.86) --
	(135.08,450.21) --
	(143.66,450.49) --
	(152.24,452.94) --
	(160.82,454.94) --
	(169.40,456.66) --
	(177.98,460.00) --
	(186.56,462.04) --
	(195.14,466.44) --
	(203.72,469.55) --
	(212.30,474.19) --
	(220.88,479.43) --
	(229.46,486.34) --
	(238.04,492.84) --
	(246.62,498.92) --
	(255.20,508.70) --
	(263.78,512.79) --
	(272.36,520.49) --
	(280.94,525.27) --
	(289.52,529.02) --
	(298.10,534.27) --
	(306.68,534.08) --
	(315.26,534.27) --
	(323.84,526.94) --
	(332.42,525.96) --
	(341.00,517.29) --
	(349.58,511.35) --
	(358.16,501.61) --
	(366.74,495.30) --
	(375.32,486.62) --
	(383.90,478.69) --
	(392.48,470.85) --
	(401.06,469.37) --
	(409.64,463.71) --
	(418.22,457.54) --
	(426.80,456.42) --
	(435.38,453.96) --
	(443.96,452.53) --
	(452.54,449.79) --
	(461.12,449.84) --
	(469.70,447.61) --
	(478.28,447.89) --
	(486.86,446.45) --
	(495.44,445.75) --
	(504.02,445.48) --
	(512.60,445.29) --
	(521.18,444.92) --
	(529.76,444.50) --
	(538.34,444.45) --
	(546.92,444.27) --
	(555.50,444.22) --
	(564.08,444.22) --
	(572.66,444.18);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.45) --
	( 49.28,444.96) --
	( 57.86,445.34) --
	( 66.44,444.78) --
	( 75.02,445.94) --
	( 83.60,446.54) --
	( 92.18,446.50) --
	(100.76,447.24) --
	(109.34,447.42) --
	(117.92,448.49) --
	(126.50,449.93) --
	(135.08,450.44) --
	(143.66,452.39) --
	(152.24,454.24) --
	(160.82,456.93) --
	(169.40,458.33) --
	(177.98,461.90) --
	(186.56,465.38) --
	(195.14,468.21) --
	(203.72,472.01) --
	(212.30,478.83) --
	(220.88,484.12) --
	(229.46,491.26) --
	(238.04,498.04) --
	(246.62,505.04) --
	(255.20,513.39) --
	(263.78,519.70) --
	(272.36,523.69) --
	(280.94,529.86) --
	(289.52,531.90) --
	(298.10,535.94) --
	(306.68,537.28) --
	(315.26,529.30) --
	(323.84,529.12) --
	(332.42,519.28) --
	(341.00,513.67) --
	(349.58,502.81) --
	(358.16,494.83) --
	(366.74,488.15) --
	(375.32,479.99) --
	(383.90,476.05) --
	(392.48,467.42) --
	(401.06,465.38) --
	(409.64,459.58) --
	(418.22,455.91) --
	(426.80,454.94) --
	(435.38,452.80) --
	(443.96,451.41) --
	(452.54,449.46) --
	(461.12,448.21) --
	(469.70,447.33) --
	(478.28,446.82) --
	(486.86,446.08) --
	(495.44,445.71) --
	(504.02,444.92) --
	(512.60,444.50) --
	(521.18,444.78) --
	(529.76,444.59) --
	(538.34,444.69) --
	(546.92,444.36) --
	(555.50,444.36) --
	(564.08,444.04) --
	(572.66,444.18);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.18) --
	( 49.28,444.96) --
	( 57.86,445.01) --
	( 66.44,444.96) --
	( 75.02,445.80) --
	( 83.60,445.57) --
	( 92.18,445.80) --
	(100.76,446.59) --
	(109.34,447.38) --
	(117.92,448.35) --
	(126.50,448.63) --
	(135.08,448.82) --
	(143.66,451.55) --
	(152.24,451.88) --
	(160.82,453.64) --
	(169.40,455.54) --
	(177.98,459.11) --
	(186.56,460.92) --
	(195.14,465.42) --
	(203.72,470.99) --
	(212.30,473.54) --
	(220.88,477.44) --
	(229.46,483.65) --
	(238.04,494.00) --
	(246.62,500.17) --
	(255.20,503.23) --
	(263.78,512.09) --
	(272.36,523.46) --
	(280.94,526.15) --
	(289.52,535.01) --
	(298.10,532.64) --
	(306.68,532.73) --
	(315.26,540.06) --
	(323.84,533.85) --
	(332.42,522.81) --
	(341.00,518.35) --
	(349.58,512.00) --
	(358.16,502.16) --
	(366.74,495.99) --
	(375.32,486.62) --
	(383.90,477.76) --
	(392.48,472.98) --
	(401.06,470.11) --
	(409.64,463.15) --
	(418.22,457.17) --
	(426.80,457.54) --
	(435.38,454.01) --
	(443.96,451.13) --
	(452.54,449.88) --
	(461.12,448.49) --
	(469.70,447.42) --
	(478.28,446.96) --
	(486.86,446.59) --
	(495.44,445.61) --
	(504.02,445.52) --
	(512.60,444.87) --
	(521.18,444.55) --
	(529.76,444.64) --
	(538.34,444.59) --
	(546.92,444.41) --
	(555.50,444.50) --
	(564.08,444.22) --
	(572.66,444.32);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.32) --
	( 49.28,444.73) --
	( 57.86,444.69) --
	( 66.44,445.20) --
	( 75.02,445.34) --
	( 83.60,445.66) --
	( 92.18,445.38) --
	(100.76,446.26) --
	(109.34,446.68) --
	(117.92,447.89) --
	(126.50,448.58) --
	(135.08,448.68) --
	(143.66,449.42) --
	(152.24,451.83) --
	(160.82,452.48) --
	(169.40,453.92) --
	(177.98,457.58) --
	(186.56,458.56) --
	(195.14,462.87) --
	(203.72,467.00) --
	(212.30,471.13) --
	(220.88,474.10) --
	(229.46,478.78) --
	(238.04,486.72) --
	(246.62,497.48) --
	(255.20,500.77) --
	(263.78,506.20) --
	(272.36,513.99) --
	(280.94,525.50) --
	(289.52,528.10) --
	(298.10,536.59) --
	(306.68,532.50) --
	(315.26,535.10) --
	(323.84,540.90) --
	(332.42,532.13) --
	(341.00,519.70) --
	(349.58,518.45) --
	(358.16,507.31) --
	(366.74,500.73) --
	(375.32,492.42) --
	(383.90,482.96) --
	(392.48,475.21) --
	(401.06,471.18) --
	(409.64,465.98) --
	(418.22,459.16) --
	(426.80,458.84) --
	(435.38,455.12) --
	(443.96,453.13) --
	(452.54,451.23) --
	(461.12,449.14) --
	(469.70,448.77) --
	(478.28,446.73) --
	(486.86,446.73) --
	(495.44,445.94) --
	(504.02,445.89) --
	(512.60,445.15) --
	(521.18,444.83) --
	(529.76,444.96) --
	(538.34,444.27) --
	(546.92,444.36) --
	(555.50,444.36) --
	(564.08,444.13) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.32) --
	( 49.28,445.06) --
	( 57.86,445.34) --
	( 66.44,445.57) --
	( 75.02,445.48) --
	( 83.60,445.80) --
	( 92.18,445.99) --
	(100.76,447.52) --
	(109.34,448.40) --
	(117.92,448.44) --
	(126.50,448.95) --
	(135.08,450.62) --
	(143.66,451.88) --
	(152.24,453.32) --
	(160.82,456.10) --
	(169.40,457.86) --
	(177.98,460.41) --
	(186.56,465.10) --
	(195.14,468.16) --
	(203.72,473.08) --
	(212.30,476.42) --
	(220.88,482.45) --
	(229.46,489.55) --
	(238.04,501.38) --
	(246.62,502.12) --
	(255.20,510.10) --
	(263.78,518.31) --
	(272.36,529.30) --
	(280.94,530.04) --
	(289.52,533.11) --
	(298.10,535.29) --
	(306.68,533.62) --
	(315.26,536.35) --
	(323.84,528.14) --
	(332.42,520.12) --
	(341.00,510.14) --
	(349.58,505.27) --
	(358.16,499.01) --
	(366.74,488.85) --
	(375.32,478.32) --
	(383.90,476.46) --
	(392.48,469.27) --
	(401.06,464.26) --
	(409.64,460.27) --
	(418.22,456.84) --
	(426.80,454.66) --
	(435.38,453.50) --
	(443.96,450.07) --
	(452.54,450.02) --
	(461.12,447.19) --
	(469.70,447.19) --
	(478.28,446.36) --
	(486.86,445.94) --
	(495.44,445.66) --
	(504.02,445.06) --
	(512.60,444.59) --
	(521.18,444.92) --
	(529.76,445.06) --
	(538.34,444.32) --
	(546.92,444.22) --
	(555.50,444.18) --
	(564.08,444.04) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.36) --
	( 49.28,444.64) --
	( 57.86,445.20) --
	( 66.44,445.52) --
	( 75.02,445.20) --
	( 83.60,445.66) --
	( 92.18,446.08) --
	(100.76,446.59) --
	(109.34,447.42) --
	(117.92,448.54) --
	(126.50,448.91) --
	(135.08,450.25) --
	(143.66,450.81) --
	(152.24,452.02) --
	(160.82,454.34) --
	(169.40,458.00) --
	(177.98,459.21) --
	(186.56,461.43) --
	(195.14,467.56) --
	(203.72,468.76) --
	(212.30,475.54) --
	(220.88,479.53) --
	(229.46,484.03) --
	(238.04,496.60) --
	(246.62,499.98) --
	(255.20,507.03) --
	(263.78,513.76) --
	(272.36,519.56) --
	(280.94,529.44) --
	(289.52,533.71) --
	(298.10,530.88) --
	(306.68,535.75) --
	(315.26,533.85) --
	(323.84,533.94) --
	(332.42,522.58) --
	(341.00,516.92) --
	(349.58,507.87) --
	(358.16,501.65) --
	(366.74,495.76) --
	(375.32,485.51) --
	(383.90,477.44) --
	(392.48,471.78) --
	(401.06,466.17) --
	(409.64,462.96) --
	(418.22,457.63) --
	(426.80,456.14) --
	(435.38,453.27) --
	(443.96,451.37) --
	(452.54,450.25) --
	(461.12,449.00) --
	(469.70,447.28) --
	(478.28,446.87) --
	(486.86,446.40) --
	(495.44,445.66) --
	(504.02,445.94) --
	(512.60,444.83) --
	(521.18,444.69) --
	(529.76,444.78) --
	(538.34,444.27) --
	(546.92,444.50) --
	(555.50,444.27) --
	(564.08,444.04) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.36) --
	( 49.28,444.50) --
	( 57.86,444.45) --
	( 66.44,444.83) --
	( 75.02,445.34) --
	( 83.60,445.29) --
	( 92.18,445.66) --
	(100.76,445.75) --
	(109.34,446.36) --
	(117.92,447.05) --
	(126.50,448.72) --
	(135.08,448.58) --
	(143.66,449.46) --
	(152.24,450.53) --
	(160.82,453.50) --
	(169.40,454.24) --
	(177.98,456.24) --
	(186.56,459.58) --
	(195.14,461.85) --
	(203.72,463.80) --
	(212.30,469.55) --
	(220.88,474.24) --
	(229.46,479.25) --
	(238.04,482.96) --
	(246.62,491.87) --
	(255.20,496.74) --
	(263.78,504.02) --
	(272.36,511.26) --
	(280.94,519.00) --
	(289.52,527.63) --
	(298.10,526.75) --
	(306.68,532.04) --
	(315.26,536.03) --
	(323.84,530.83) --
	(332.42,531.34) --
	(341.00,525.73) --
	(349.58,518.68) --
	(358.16,510.98) --
	(366.74,503.18) --
	(375.32,495.85) --
	(383.90,487.18) --
	(392.48,479.57) --
	(401.06,476.14) --
	(409.64,470.48) --
	(418.22,463.61) --
	(426.80,459.35) --
	(435.38,457.44) --
	(443.96,455.50) --
	(452.54,451.69) --
	(461.12,451.00) --
	(469.70,449.19) --
	(478.28,447.47) --
	(486.86,447.10) --
	(495.44,445.99) --
	(504.02,446.03) --
	(512.60,445.43) --
	(521.18,445.10) --
	(529.76,444.64) --
	(538.34,444.64) --
	(546.92,444.36) --
	(555.50,444.32) --
	(564.08,444.22) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.45) --
	( 49.28,444.55) --
	( 57.86,444.36) --
	( 66.44,444.83) --
	( 75.02,445.29) --
	( 83.60,445.71) --
	( 92.18,445.20) --
	(100.76,445.99) --
	(109.34,446.17) --
	(117.92,447.56) --
	(126.50,448.77) --
	(135.08,448.35) --
	(143.66,449.70) --
	(152.24,451.23) --
	(160.82,452.71) --
	(169.40,455.40) --
	(177.98,456.42) --
	(186.56,458.79) --
	(195.14,463.01) --
	(203.72,463.94) --
	(212.30,468.62) --
	(220.88,476.97) --
	(229.46,480.50) --
	(238.04,483.28) --
	(246.62,491.12) --
	(255.20,497.25) --
	(263.78,504.39) --
	(272.36,512.51) --
	(280.94,518.21) --
	(289.52,527.40) --
	(298.10,526.38) --
	(306.68,529.67) --
	(315.26,536.96) --
	(323.84,531.02) --
	(332.42,532.04) --
	(341.00,526.61) --
	(349.58,517.52) --
	(358.16,513.44) --
	(366.74,502.63) --
	(375.32,494.79) --
	(383.90,489.17) --
	(392.48,480.04) --
	(401.06,474.70) --
	(409.64,469.64) --
	(418.22,462.18) --
	(426.80,460.00) --
	(435.38,457.77) --
	(443.96,454.47) --
	(452.54,451.51) --
	(461.12,450.53) --
	(469.70,448.44) --
	(478.28,447.75) --
	(486.86,446.12) --
	(495.44,446.63) --
	(504.02,446.31) --
	(512.60,445.29) --
	(521.18,445.24) --
	(529.76,444.73) --
	(538.34,444.45) --
	(546.92,444.22) --
	(555.50,444.22) --
	(564.08,444.41) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.87) --
	( 49.28,444.73) --
	( 57.86,445.24) --
	( 66.44,445.48) --
	( 75.02,445.29) --
	( 83.60,445.99) --
	( 92.18,446.08) --
	(100.76,447.75) --
	(109.34,448.21) --
	(117.92,448.03) --
	(126.50,450.30) --
	(135.08,451.09) --
	(143.66,453.64) --
	(152.24,455.26) --
	(160.82,456.05) --
	(169.40,458.93) --
	(177.98,463.66) --
	(186.56,464.87) --
	(195.14,468.99) --
	(203.72,475.58) --
	(212.30,481.75) --
	(220.88,484.07) --
	(229.46,490.29) --
	(238.04,499.47) --
	(246.62,505.36) --
	(255.20,513.44) --
	(263.78,518.86) --
	(272.36,526.01) --
	(280.94,529.63) --
	(289.52,535.52) --
	(298.10,530.93) --
	(306.68,535.19) --
	(315.26,530.18) --
	(323.84,525.96) --
	(332.42,516.41) --
	(341.00,511.58) --
	(349.58,501.14) --
	(358.16,495.44) --
	(366.74,486.62) --
	(375.32,478.27) --
	(383.90,474.10) --
	(392.48,468.48) --
	(401.06,462.50) --
	(409.64,458.93) --
	(418.22,456.56) --
	(426.80,454.24) --
	(435.38,451.83) --
	(443.96,450.02) --
	(452.54,448.40) --
	(461.12,447.24) --
	(469.70,446.77) --
	(478.28,446.50) --
	(486.86,445.75) --
	(495.44,445.24) --
	(504.02,445.38) --
	(512.60,444.55) --
	(521.18,444.32) --
	(529.76,444.50) --
	(538.34,444.18) --
	(546.92,444.27) --
	(555.50,444.04) --
	(564.08,444.18) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.64) --
	( 49.28,444.83) --
	( 57.86,445.43) --
	( 66.44,444.92) --
	( 75.02,445.24) --
	( 83.60,446.17) --
	( 92.18,446.59) --
	(100.76,447.33) --
	(109.34,448.03) --
	(117.92,448.17) --
	(126.50,450.16) --
	(135.08,451.32) --
	(143.66,452.94) --
	(152.24,455.50) --
	(160.82,456.01) --
	(169.40,459.62) --
	(177.98,463.20) --
	(186.56,464.31) --
	(195.14,468.25) --
	(203.72,478.04) --
	(212.30,481.94) --
	(220.88,483.42) --
	(229.46,488.20) --
	(238.04,500.86) --
	(246.62,505.46) --
	(255.20,514.18) --
	(263.78,519.51) --
	(272.36,524.15) --
	(280.94,529.16) --
	(289.52,535.84) --
	(298.10,532.09) --
	(306.68,533.01) --
	(315.26,533.15) --
	(323.84,524.34) --
	(332.42,519.10) --
	(341.00,509.96) --
	(349.58,501.89) --
	(358.16,496.74) --
	(366.74,487.64) --
	(375.32,480.45) --
	(383.90,473.36) --
	(392.48,467.32) --
	(401.06,463.61) --
	(409.64,460.00) --
	(418.22,457.35) --
	(426.80,452.80) --
	(435.38,451.83) --
	(443.96,449.74) --
	(452.54,447.98) --
	(461.12,447.19) --
	(469.70,446.87) --
	(478.28,446.31) --
	(486.86,445.80) --
	(495.44,445.24) --
	(504.02,444.73) --
	(512.60,444.87) --
	(521.18,444.27) --
	(529.76,444.50) --
	(538.34,444.08) --
	(546.92,444.27) --
	(555.50,444.04) --
	(564.08,443.99) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.87) --
	( 49.28,445.01) --
	( 57.86,445.10) --
	( 66.44,445.38) --
	( 75.02,445.66) --
	( 83.60,446.45) --
	( 92.18,446.17) --
	(100.76,447.28) --
	(109.34,448.03) --
	(117.92,448.30) --
	(126.50,450.25) --
	(135.08,450.76) --
	(143.66,452.34) --
	(152.24,453.04) --
	(160.82,455.45) --
	(169.40,458.28) --
	(177.98,462.13) --
	(186.56,465.14) --
	(195.14,469.04) --
	(203.72,471.18) --
	(212.30,478.69) --
	(220.88,484.35) --
	(229.46,489.36) --
	(238.04,496.27) --
	(246.62,502.95) --
	(255.20,510.79) --
	(263.78,514.60) --
	(272.36,524.43) --
	(280.94,525.08) --
	(289.52,530.51) --
	(298.10,535.56) --
	(306.68,534.87) --
	(315.26,531.02) --
	(323.84,526.01) --
	(332.42,522.81) --
	(341.00,513.07) --
	(349.58,505.32) --
	(358.16,499.19) --
	(366.74,489.45) --
	(375.32,481.84) --
	(383.90,478.60) --
	(392.48,469.32) --
	(401.06,465.47) --
	(409.64,460.60) --
	(418.22,457.63) --
	(426.80,455.17) --
	(435.38,453.41) --
	(443.96,451.09) --
	(452.54,449.74) --
	(461.12,448.26) --
	(469.70,446.87) --
	(478.28,445.75) --
	(486.86,446.22) --
	(495.44,445.75) --
	(504.02,445.24) --
	(512.60,444.92) --
	(521.18,444.83) --
	(529.76,444.64) --
	(538.34,444.41) --
	(546.92,444.64) --
	(555.50,444.13) --
	(564.08,444.04) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.73) --
	( 49.28,445.20) --
	( 57.86,445.06) --
	( 66.44,445.15) --
	( 75.02,445.57) --
	( 83.60,445.89) --
	( 92.18,446.40) --
	(100.76,447.15) --
	(109.34,447.61) --
	(117.92,448.26) --
	(126.50,449.28) --
	(135.08,450.07) --
	(143.66,452.25) --
	(152.24,452.25) --
	(160.82,454.34) --
	(169.40,457.49) --
	(177.98,459.86) --
	(186.56,465.19) --
	(195.14,466.58) --
	(203.72,470.90) --
	(212.30,474.28) --
	(220.88,483.47) --
	(229.46,488.39) --
	(238.04,492.10) --
	(246.62,499.89) --
	(255.20,508.24) --
	(263.78,514.13) --
	(272.36,521.14) --
	(280.94,526.94) --
	(289.52,526.80) --
	(298.10,532.55) --
	(306.68,535.89) --
	(315.26,533.66) --
	(323.84,528.37) --
	(332.42,524.48) --
	(341.00,519.84) --
	(349.58,508.89) --
	(358.16,499.98) --
	(366.74,490.89) --
	(375.32,483.93) --
	(383.90,479.06) --
	(392.48,472.24) --
	(401.06,465.38) --
	(409.64,463.20) --
	(418.22,458.97) --
	(426.80,457.30) --
	(435.38,452.76) --
	(443.96,451.65) --
	(452.54,449.65) --
	(461.12,448.54) --
	(469.70,447.70) --
	(478.28,446.59) --
	(486.86,445.89) --
	(495.44,446.08) --
	(504.02,445.38) --
	(512.60,445.24) --
	(521.18,444.59) --
	(529.76,444.55) --
	(538.34,444.64) --
	(546.92,444.32) --
	(555.50,444.27) --
	(564.08,444.22) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.64) --
	( 49.28,445.29) --
	( 57.86,445.06) --
	( 66.44,445.06) --
	( 75.02,445.66) --
	( 83.60,445.52) --
	( 92.18,446.40) --
	(100.76,446.68) --
	(109.34,448.17) --
	(117.92,447.28) --
	(126.50,449.65) --
	(135.08,449.33) --
	(143.66,451.46) --
	(152.24,452.76) --
	(160.82,454.20) --
	(169.40,455.54) --
	(177.98,460.23) --
	(186.56,463.15) --
	(195.14,466.35) --
	(203.72,469.41) --
	(212.30,474.24) --
	(220.88,480.92) --
	(229.46,485.60) --
	(238.04,492.89) --
	(246.62,499.19) --
	(255.20,506.11) --
	(263.78,512.23) --
	(272.36,518.82) --
	(280.94,524.99) --
	(289.52,528.88) --
	(298.10,532.64) --
	(306.68,533.38) --
	(315.26,535.05) --
	(323.84,526.57) --
	(332.42,523.55) --
	(341.00,520.77) --
	(349.58,513.76) --
	(358.16,503.46) --
	(366.74,493.12) --
	(375.32,485.83) --
	(383.90,479.57) --
	(392.48,474.01) --
	(401.06,468.11) --
	(409.64,462.83) --
	(418.22,459.21) --
	(426.80,457.40) --
	(435.38,453.96) --
	(443.96,452.67) --
	(452.54,450.16) --
	(461.12,448.30) --
	(469.70,448.26) --
	(478.28,446.87) --
	(486.86,446.91) --
	(495.44,445.61) --
	(504.02,445.43) --
	(512.60,444.92) --
	(521.18,445.01) --
	(529.76,444.73) --
	(538.34,444.50) --
	(546.92,444.22) --
	(555.50,444.36) --
	(564.08,444.27) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.59) --
	( 49.28,444.92) --
	( 57.86,445.10) --
	( 66.44,445.01) --
	( 75.02,445.57) --
	( 83.60,445.52) --
	( 92.18,445.80) --
	(100.76,446.91) --
	(109.34,447.33) --
	(117.92,447.84) --
	(126.50,448.68) --
	(135.08,449.70) --
	(143.66,450.21) --
	(152.24,451.88) --
	(160.82,453.41) --
	(169.40,455.50) --
	(177.98,458.65) --
	(186.56,461.25) --
	(195.14,465.24) --
	(203.72,467.93) --
	(212.30,474.01) --
	(220.88,477.35) --
	(229.46,483.56) --
	(238.04,490.01) --
	(246.62,495.58) --
	(255.20,505.74) --
	(263.78,507.50) --
	(272.36,516.68) --
	(280.94,521.79) --
	(289.52,527.31) --
	(298.10,532.27) --
	(306.68,533.38) --
	(315.26,536.91) --
	(323.84,530.18) --
	(332.42,525.96) --
	(341.00,521.42) --
	(349.58,515.48) --
	(358.16,509.35) --
	(366.74,497.85) --
	(375.32,486.76) --
	(383.90,481.47) --
	(392.48,476.97) --
	(401.06,469.92) --
	(409.64,464.50) --
	(418.22,461.34) --
	(426.80,458.00) --
	(435.38,454.52) --
	(443.96,451.88) --
	(452.54,450.86) --
	(461.12,449.23) --
	(469.70,449.28) --
	(478.28,447.28) --
	(486.86,446.12) --
	(495.44,446.08) --
	(504.02,445.85) --
	(512.60,444.83) --
	(521.18,444.96) --
	(529.76,444.78) --
	(538.34,444.64) --
	(546.92,444.27) --
	(555.50,444.13) --
	(564.08,444.22) --
	(572.66,444.32);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.45) --
	( 49.28,444.73) --
	( 57.86,445.06) --
	( 66.44,444.92) --
	( 75.02,445.52) --
	( 83.60,445.15) --
	( 92.18,446.22) --
	(100.76,446.45) --
	(109.34,446.26) --
	(117.92,447.70) --
	(126.50,449.33) --
	(135.08,449.93) --
	(143.66,450.67) --
	(152.24,451.04) --
	(160.82,452.76) --
	(169.40,456.38) --
	(177.98,457.12) --
	(186.56,459.48) --
	(195.14,462.83) --
	(203.72,466.81) --
	(212.30,471.78) --
	(220.88,474.42) --
	(229.46,480.92) --
	(238.04,487.74) --
	(246.62,493.81) --
	(255.20,498.41) --
	(263.78,504.67) --
	(272.36,515.20) --
	(280.94,519.79) --
	(289.52,525.64) --
	(298.10,529.58) --
	(306.68,529.44) --
	(315.26,537.65) --
	(323.84,533.85) --
	(332.42,526.98) --
	(341.00,524.89) --
	(349.58,515.43) --
	(358.16,510.47) --
	(366.74,501.14) --
	(375.32,494.18) --
	(383.90,486.30) --
	(392.48,477.30) --
	(401.06,473.59) --
	(409.64,466.35) --
	(418.22,464.54) --
	(426.80,458.42) --
	(435.38,456.05) --
	(443.96,453.08) --
	(452.54,451.51) --
	(461.12,449.28) --
	(469.70,449.14) --
	(478.28,447.89) --
	(486.86,447.28) --
	(495.44,446.12) --
	(504.02,445.61) --
	(512.60,445.24) --
	(521.18,444.87) --
	(529.76,444.83) --
	(538.34,444.41) --
	(546.92,444.45) --
	(555.50,444.45) --
	(564.08,444.08) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.59) --
	( 49.28,444.73) --
	( 57.86,445.20) --
	( 66.44,444.83) --
	( 75.02,445.24) --
	( 83.60,445.61) --
	( 92.18,446.26) --
	(100.76,446.36) --
	(109.34,447.05) --
	(117.92,447.61) --
	(126.50,448.86) --
	(135.08,450.53) --
	(143.66,450.62) --
	(152.24,450.67) --
	(160.82,454.61) --
	(169.40,456.52) --
	(177.98,456.38) --
	(186.56,460.64) --
	(195.14,463.15) --
	(203.72,468.39) --
	(212.30,472.52) --
	(220.88,475.81) --
	(229.46,482.17) --
	(238.04,488.15) --
	(246.62,494.70) --
	(255.20,500.40) --
	(263.78,506.11) --
	(272.36,517.47) --
	(280.94,520.86) --
	(289.52,525.36) --
	(298.10,528.79) --
	(306.68,531.85) --
	(315.26,533.62) --
	(323.84,531.39) --
	(332.42,525.59) --
	(341.00,521.93) --
	(349.58,515.80) --
	(358.16,505.92) --
	(366.74,501.28) --
	(375.32,495.30) --
	(383.90,484.44) --
	(392.48,478.09) --
	(401.06,471.18) --
	(409.64,468.21) --
	(418.22,462.13) --
	(426.80,457.77) --
	(435.38,456.10) --
	(443.96,453.78) --
	(452.54,452.43) --
	(461.12,449.88) --
	(469.70,448.26) --
	(478.28,447.89) --
	(486.86,447.10) --
	(495.44,446.26) --
	(504.02,445.99) --
	(512.60,445.06) --
	(521.18,445.24) --
	(529.76,444.78) --
	(538.34,444.96) --
	(546.92,444.18) --
	(555.50,444.41) --
	(564.08,444.50) --
	(572.66,444.27);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.78) --
	( 49.28,444.87) --
	( 57.86,444.96) --
	( 66.44,444.83) --
	( 75.02,445.89) --
	( 83.60,445.24) --
	( 92.18,446.36) --
	(100.76,446.50) --
	(109.34,447.05) --
	(117.92,447.98) --
	(126.50,448.72) --
	(135.08,450.62) --
	(143.66,451.46) --
	(152.24,451.04) --
	(160.82,455.26) --
	(169.40,456.24) --
	(177.98,458.14) --
	(186.56,460.64) --
	(195.14,465.01) --
	(203.72,469.92) --
	(212.30,473.40) --
	(220.88,478.50) --
	(229.46,485.70) --
	(238.04,487.69) --
	(246.62,497.52) --
	(255.20,503.09) --
	(263.78,508.33) --
	(272.36,517.61) --
	(280.94,521.00) --
	(289.52,529.02) --
	(298.10,529.30) --
	(306.68,531.02) --
	(315.26,532.13) --
	(323.84,532.13) --
	(332.42,525.50) --
	(341.00,518.03) --
	(349.58,513.39) --
	(358.16,504.72) --
	(366.74,499.94) --
	(375.32,490.24) --
	(383.90,482.17) --
	(392.48,474.47) --
	(401.06,470.85) --
	(409.64,465.56) --
	(418.22,461.80) --
	(426.80,457.03) --
	(435.38,454.89) --
	(443.96,451.97) --
	(452.54,451.18) --
	(461.12,450.72) --
	(469.70,448.30) --
	(478.28,446.77) --
	(486.86,447.47) --
	(495.44,446.26) --
	(504.02,446.12) --
	(512.60,445.06) --
	(521.18,444.78) --
	(529.76,444.96) --
	(538.34,444.55) --
	(546.92,444.50) --
	(555.50,444.22) --
	(564.08,444.18) --
	(572.66,444.22);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.69) --
	( 49.28,444.64) --
	( 57.86,445.24) --
	( 66.44,445.06) --
	( 75.02,445.71) --
	( 83.60,445.20) --
	( 92.18,446.17) --
	(100.76,447.28) --
	(109.34,447.19) --
	(117.92,448.21) --
	(126.50,449.09) --
	(135.08,450.16) --
	(143.66,451.92) --
	(152.24,452.62) --
	(160.82,455.45) --
	(169.40,457.68) --
	(177.98,458.19) --
	(186.56,462.08) --
	(195.14,467.93) --
	(203.72,470.48) --
	(212.30,473.40) --
	(220.88,481.66) --
	(229.46,485.70) --
	(238.04,491.12) --
	(246.62,499.75) --
	(255.20,505.64) --
	(263.78,515.15) --
	(272.36,515.85) --
	(280.94,525.13) --
	(289.52,531.06) --
	(298.10,534.82) --
	(306.68,528.51) --
	(315.26,533.57) --
	(323.84,531.95) --
	(332.42,524.89) --
	(341.00,516.92) --
	(349.58,509.63) --
	(358.16,502.26) --
	(366.74,494.32) --
	(375.32,484.86) --
	(383.90,481.43) --
	(392.48,473.54) --
	(401.06,468.72) --
	(409.64,461.90) --
	(418.22,460.60) --
	(426.80,455.50) --
	(435.38,454.43) --
	(443.96,452.20) --
	(452.54,451.13) --
	(461.12,448.35) --
	(469.70,447.52) --
	(478.28,447.93) --
	(486.86,447.24) --
	(495.44,446.45) --
	(504.02,445.06) --
	(512.60,445.34) --
	(521.18,445.01) --
	(529.76,444.73) --
	(538.34,444.64) --
	(546.92,444.36) --
	(555.50,444.32) --
	(564.08,444.41) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.87) --
	( 49.28,444.69) --
	( 57.86,445.24) --
	( 66.44,445.29) --
	( 75.02,445.66) --
	( 83.60,446.26) --
	( 92.18,446.45) --
	(100.76,446.59) --
	(109.34,447.89) --
	(117.92,448.54) --
	(126.50,449.23) --
	(135.08,451.04) --
	(143.66,451.65) --
	(152.24,453.69) --
	(160.82,455.68) --
	(169.40,457.21) --
	(177.98,460.64) --
	(186.56,463.98) --
	(195.14,467.79) --
	(203.72,473.31) --
	(212.30,477.16) --
	(220.88,480.82) --
	(229.46,489.08) --
	(238.04,496.32) --
	(246.62,502.21) --
	(255.20,508.38) --
	(263.78,515.71) --
	(272.36,520.77) --
	(280.94,529.02) --
	(289.52,531.30) --
	(298.10,531.30) --
	(306.68,531.16) --
	(315.26,530.32) --
	(323.84,530.04) --
	(332.42,520.72) --
	(341.00,513.25) --
	(349.58,506.99) --
	(358.16,498.92) --
	(366.74,488.53) --
	(375.32,481.15) --
	(383.90,476.42) --
	(392.48,473.31) --
	(401.06,468.39) --
	(409.64,461.94) --
	(418.22,457.63) --
	(426.80,454.61) --
	(435.38,454.52) --
	(443.96,452.06) --
	(452.54,450.44) --
	(461.12,448.40) --
	(469.70,447.42) --
	(478.28,447.05) --
	(486.86,446.36) --
	(495.44,445.34) --
	(504.02,445.57) --
	(512.60,445.20) --
	(521.18,444.78) --
	(529.76,444.87) --
	(538.34,444.27) --
	(546.92,444.50) --
	(555.50,444.18) --
	(564.08,444.04) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.78) --
	( 49.28,445.01) --
	( 57.86,444.64) --
	( 66.44,445.24) --
	( 75.02,445.15) --
	( 83.60,445.99) --
	( 92.18,446.87) --
	(100.76,446.26) --
	(109.34,446.50) --
	(117.92,448.07) --
	(126.50,449.19) --
	(135.08,450.58) --
	(143.66,451.00) --
	(152.24,452.48) --
	(160.82,454.38) --
	(169.40,455.91) --
	(177.98,458.79) --
	(186.56,461.57) --
	(195.14,465.24) --
	(203.72,469.51) --
	(212.30,474.89) --
	(220.88,477.90) --
	(229.46,484.81) --
	(238.04,491.96) --
	(246.62,496.18) --
	(255.20,502.95) --
	(263.78,512.32) --
	(272.36,517.33) --
	(280.94,523.50) --
	(289.52,527.26) --
	(298.10,527.72) --
	(306.68,533.52) --
	(315.26,530.93) --
	(323.84,528.47) --
	(332.42,523.87) --
	(341.00,515.15) --
	(349.58,511.49) --
	(358.16,504.81) --
	(366.74,496.78) --
	(375.32,487.60) --
	(383.90,480.78) --
	(392.48,475.03) --
	(401.06,471.87) --
	(409.64,466.49) --
	(418.22,460.74) --
	(426.80,457.26) --
	(435.38,455.73) --
	(443.96,453.59) --
	(452.54,450.30) --
	(461.12,449.51) --
	(469.70,448.03) --
	(478.28,447.61) --
	(486.86,446.59) --
	(495.44,445.75) --
	(504.02,445.94) --
	(512.60,445.85) --
	(521.18,444.78) --
	(529.76,445.15) --
	(538.34,444.45) --
	(546.92,444.32) --
	(555.50,444.36) --
	(564.08,444.22) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.92) --
	( 49.28,444.55) --
	( 57.86,445.10) --
	( 66.44,444.92) --
	( 75.02,445.57) --
	( 83.60,445.99) --
	( 92.18,446.91) --
	(100.76,446.54) --
	(109.34,446.54) --
	(117.92,448.44) --
	(126.50,449.60) --
	(135.08,450.49) --
	(143.66,451.13) --
	(152.24,453.36) --
	(160.82,453.64) --
	(169.40,456.56) --
	(177.98,459.30) --
	(186.56,461.67) --
	(195.14,466.12) --
	(203.72,470.94) --
	(212.30,474.24) --
	(220.88,479.11) --
	(229.46,485.51) --
	(238.04,491.22) --
	(246.62,497.94) --
	(255.20,506.29) --
	(263.78,512.56) --
	(272.36,518.73) --
	(280.94,524.76) --
	(289.52,529.02) --
	(298.10,527.21) --
	(306.68,535.52) --
	(315.26,529.63) --
	(323.84,529.35) --
	(332.42,522.02) --
	(341.00,515.66) --
	(349.58,509.17) --
	(358.16,503.46) --
	(366.74,496.92) --
	(375.32,486.02) --
	(383.90,479.66) --
	(392.48,475.21) --
	(401.06,471.31) --
	(409.64,465.28) --
	(418.22,459.67) --
	(426.80,456.28) --
	(435.38,453.36) --
	(443.96,452.57) --
	(452.54,450.72) --
	(461.12,448.49) --
	(469.70,448.30) --
	(478.28,446.59) --
	(486.86,446.17) --
	(495.44,446.31) --
	(504.02,445.57) --
	(512.60,445.66) --
	(521.18,444.50) --
	(529.76,444.55) --
	(538.34,444.45) --
	(546.92,444.59) --
	(555.50,444.32) --
	(564.08,444.36) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.92) --
	( 49.28,444.59) --
	( 57.86,445.01) --
	( 66.44,445.10) --
	( 75.02,445.66) --
	( 83.60,446.08) --
	( 92.18,447.05) --
	(100.76,446.73) --
	(109.34,447.10) --
	(117.92,448.49) --
	(126.50,449.46) --
	(135.08,450.72) --
	(143.66,451.97) --
	(152.24,453.64) --
	(160.82,454.71) --
	(169.40,457.49) --
	(177.98,459.90) --
	(186.56,463.47) --
	(195.14,466.86) --
	(203.72,472.43) --
	(212.30,475.58) --
	(220.88,483.10) --
	(229.46,486.02) --
	(238.04,492.24) --
	(246.62,500.86) --
	(255.20,511.07) --
	(263.78,514.69) --
	(272.36,518.63) --
	(280.94,525.31) --
	(289.52,530.65) --
	(298.10,532.18) --
	(306.68,534.36) --
	(315.26,530.04) --
	(323.84,526.24) --
	(332.42,520.35) --
	(341.00,517.06) --
	(349.58,507.59) --
	(358.16,500.35) --
	(366.74,493.63) --
	(375.32,483.52) --
	(383.90,478.50) --
	(392.48,472.52) --
	(401.06,467.65) --
	(409.64,462.73) --
	(418.22,457.91) --
	(426.80,456.01) --
	(435.38,453.83) --
	(443.96,450.86) --
	(452.54,449.46) --
	(461.12,449.51) --
	(469.70,448.07) --
	(478.28,446.63) --
	(486.86,446.68) --
	(495.44,445.38) --
	(504.02,445.66) --
	(512.60,445.20) --
	(521.18,444.96) --
	(529.76,444.69) --
	(538.34,444.45) --
	(546.92,444.27) --
	(555.50,444.32) --
	(564.08,444.13);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.96) --
	( 49.28,444.36) --
	( 57.86,444.78) --
	( 66.44,445.20) --
	( 75.02,445.66) --
	( 83.60,446.17) --
	( 92.18,446.26) --
	(100.76,447.10) --
	(109.34,447.93) --
	(117.92,448.03) --
	(126.50,448.86) --
	(135.08,450.16) --
	(143.66,451.88) --
	(152.24,452.80) --
	(160.82,455.36) --
	(169.40,457.40) --
	(177.98,460.04) --
	(186.56,462.83) --
	(195.14,466.72) --
	(203.72,472.06) --
	(212.30,478.09) --
	(220.88,482.87) --
	(229.46,485.74) --
	(238.04,496.64) --
	(246.62,501.65) --
	(255.20,511.44) --
	(263.78,515.57) --
	(272.36,524.48) --
	(280.94,530.28) --
	(289.52,533.11) --
	(298.10,537.51) --
	(306.68,535.94) --
	(315.26,535.10) --
	(323.84,528.93) --
	(332.42,522.90) --
	(341.00,515.57) --
	(349.58,508.61) --
	(358.16,500.22) --
	(366.74,490.84) --
	(375.32,482.49) --
	(383.90,476.05) --
	(392.48,469.41) --
	(401.06,464.68) --
	(409.64,460.23) --
	(418.22,457.58) --
	(426.80,454.75) --
	(435.38,452.06) --
	(443.96,450.44) --
	(452.54,449.56) --
	(461.12,448.63) --
	(469.70,446.91) --
	(478.28,446.03) --
	(486.86,446.03) --
	(495.44,445.80) --
	(504.02,444.92) --
	(512.60,445.01) --
	(521.18,445.01) --
	(529.76,444.45) --
	(538.34,444.59) --
	(546.92,444.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (564.08,444.13) --
	(572.66,444.04);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.55) --
	( 49.28,444.45) --
	( 57.86,445.29) --
	( 66.44,445.01) --
	( 75.02,446.08) --
	( 83.60,446.45) --
	( 92.18,446.73) --
	(100.76,446.87) --
	(109.34,448.30) --
	(117.92,448.44) --
	(126.50,449.74) --
	(135.08,450.76) --
	(143.66,452.39) --
	(152.24,454.47) --
	(160.82,455.68) --
	(169.40,459.25) --
	(177.98,461.76) --
	(186.56,463.06) --
	(195.14,468.44) --
	(203.72,475.12) --
	(212.30,480.04) --
	(220.88,483.10) --
	(229.46,492.89) --
	(238.04,497.43) --
	(246.62,507.13) --
	(255.20,512.00) --
	(263.78,520.77) --
	(272.36,526.05) --
	(280.94,532.18) --
	(289.52,533.85) --
	(298.10,538.26) --
	(306.68,534.13) --
	(315.26,535.19) --
	(323.84,526.43) --
	(332.42,521.23) --
	(341.00,512.00) --
	(349.58,500.86) --
	(358.16,496.97) --
	(366.74,486.25) --
	(375.32,480.41) --
	(383.90,472.71) --
	(392.48,465.38) --
	(401.06,464.91) --
	(409.64,458.74) --
	(418.22,456.70) --
	(426.80,452.62) --
	(435.38,451.41) --
	(443.96,449.84) --
	(452.54,449.51) --
	(461.12,448.26) --
	(469.70,446.36) --
	(478.28,446.40) --
	(486.86,445.80) --
	(495.44,445.20) --
	(504.02,444.73) --
	(512.60,444.92) --
	(521.18,444.73) --
	(529.76,444.41) --
	(538.34,444.32) --
	(546.92,444.41) --
	(555.50,444.04) --
	(564.08,443.99) --
	(572.66,443.99);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.73) --
	( 49.28,444.87) --
	( 57.86,444.50) --
	( 66.44,445.01) --
	( 75.02,445.10) --
	( 83.60,445.52) --
	( 92.18,446.31) --
	(100.76,447.01) --
	(109.34,446.36) --
	(117.92,447.47) --
	(126.50,448.82) --
	(135.08,449.37) --
	(143.66,450.35) --
	(152.24,451.23) --
	(160.82,453.36) --
	(169.40,455.31) --
	(177.98,457.86) --
	(186.56,461.06) --
	(195.14,462.69) --
	(203.72,466.07) --
	(212.30,473.03) --
	(220.88,478.04) --
	(229.46,482.49) --
	(238.04,489.73) --
	(246.62,495.76) --
	(255.20,503.83) --
	(263.78,508.61) --
	(272.36,517.75) --
	(280.94,525.31) --
	(289.52,529.63) --
	(298.10,536.49) --
	(306.68,533.34) --
	(315.26,536.96) --
	(323.84,535.29) --
	(332.42,528.88) --
	(341.00,522.62) --
	(349.58,513.25) --
	(358.16,506.20) --
	(366.74,498.45) --
	(375.32,487.92) --
	(383.90,481.71) --
	(392.48,475.07) --
	(401.06,468.02) --
	(409.64,465.65) --
	(418.22,461.48) --
	(426.80,457.63) --
	(435.38,455.36) --
	(443.96,452.34) --
	(452.54,450.58) --
	(461.12,449.84) --
	(469.70,447.79) --
	(478.28,447.56) --
	(486.86,446.31) --
	(495.44,445.61) --
	(504.02,445.71) --
	(512.60,444.96) --
	(521.18,444.87) --
	(529.76,444.78) --
	(538.34,444.59) --
	(546.92,444.50) --
	(555.50,444.04) --
	(564.08,444.22) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.92) --
	( 49.28,444.69) --
	( 57.86,444.64) --
	( 66.44,445.43) --
	( 75.02,445.15) --
	( 83.60,446.03) --
	( 92.18,446.59) --
	(100.76,446.87) --
	(109.34,446.59) --
	(117.92,448.26) --
	(126.50,449.98) --
	(135.08,449.70) --
	(143.66,450.76) --
	(152.24,453.27) --
	(160.82,454.52) --
	(169.40,456.75) --
	(177.98,459.30) --
	(186.56,462.69) --
	(195.14,465.47) --
	(203.72,471.13) --
	(212.30,475.03) --
	(220.88,480.04) --
	(229.46,489.08) --
	(238.04,494.37) --
	(246.62,499.84) --
	(255.20,507.59) --
	(263.78,514.87) --
	(272.36,524.80) --
	(280.94,524.66) --
	(289.52,535.70) --
	(298.10,536.59) --
	(306.68,537.00) --
	(315.26,536.77) --
	(323.84,532.13) --
	(332.42,524.57) --
	(341.00,517.33) --
	(349.58,508.47) --
	(358.16,503.60) --
	(366.74,491.40) --
	(375.32,482.49) --
	(383.90,477.16) --
	(392.48,470.90) --
	(401.06,466.07) --
	(409.64,462.78) --
	(418.22,457.77) --
	(426.80,454.99) --
	(435.38,452.53) --
	(443.96,451.65) --
	(452.54,450.02) --
	(461.12,448.17) --
	(469.70,447.79) --
	(478.28,446.73) --
	(486.86,446.17) --
	(495.44,445.80) --
	(504.02,444.87) --
	(512.60,444.92) --
	(521.18,444.78) --
	(529.76,444.45) --
	(538.34,444.78) --
	(546.92,444.08) --
	(555.50,444.45) --
	(564.08,444.08) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.73) --
	( 49.28,444.73) --
	( 57.86,445.15) --
	( 66.44,445.24) --
	( 75.02,445.52) --
	( 83.60,446.08) --
	( 92.18,446.31) --
	(100.76,447.61) --
	(109.34,447.42) --
	(117.92,448.03) --
	(126.50,449.79) --
	(135.08,450.44) --
	(143.66,451.97) --
	(152.24,453.22) --
	(160.82,454.15) --
	(169.40,458.33) --
	(177.98,460.32) --
	(186.56,463.80) --
	(195.14,467.05) --
	(203.72,472.06) --
	(212.30,477.67) --
	(220.88,480.50) --
	(229.46,488.48) --
	(238.04,493.44) --
	(246.62,503.18) --
	(255.20,513.30) --
	(263.78,514.13) --
	(272.36,520.44) --
	(280.94,527.40) --
	(289.52,533.66) --
	(298.10,532.73) --
	(306.68,535.61) --
	(315.26,532.78) --
	(323.84,527.96) --
	(332.42,517.80) --
	(341.00,518.73) --
	(349.58,504.95) --
	(358.16,499.66) --
	(366.74,491.54) --
	(375.32,482.96) --
	(383.90,475.26) --
	(392.48,470.48) --
	(401.06,467.05) --
	(409.64,460.00) --
	(418.22,457.35) --
	(426.80,454.75) --
	(435.38,452.71) --
	(443.96,450.86) --
	(452.54,449.93) --
	(461.12,448.21) --
	(469.70,447.10) --
	(478.28,447.10) --
	(486.86,446.31) --
	(495.44,445.34) --
	(504.02,445.29) --
	(512.60,444.83) --
	(521.18,444.59) --
	(529.76,444.36) --
	(538.34,444.59) --
	(546.92,444.08) --
	(555.50,444.13);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.69) --
	( 49.28,444.73) --
	( 57.86,444.87) --
	( 66.44,444.83) --
	( 75.02,445.29) --
	( 83.60,445.75) --
	( 92.18,446.12) --
	(100.76,446.26) --
	(109.34,447.93) --
	(117.92,447.24) --
	(126.50,449.14) --
	(135.08,448.91) --
	(143.66,451.00) --
	(152.24,452.76) --
	(160.82,452.71) --
	(169.40,454.75) --
	(177.98,458.28) --
	(186.56,461.90) --
	(195.14,464.22) --
	(203.72,467.65) --
	(212.30,473.08) --
	(220.88,477.02) --
	(229.46,483.28) --
	(238.04,490.38) --
	(246.62,495.62) --
	(255.20,504.39) --
	(263.78,512.46) --
	(272.36,516.36) --
	(280.94,522.95) --
	(289.52,530.00) --
	(298.10,531.71) --
	(306.68,535.84) --
	(315.26,533.99) --
	(323.84,531.99) --
	(332.42,528.19) --
	(341.00,521.32) --
	(349.58,510.19) --
	(358.16,503.00) --
	(366.74,499.19) --
	(375.32,490.52) --
	(383.90,479.66) --
	(392.48,478.60) --
	(401.06,469.74) --
	(409.64,466.21) --
	(418.22,458.74) --
	(426.80,456.19) --
	(435.38,453.92) --
	(443.96,452.48) --
	(452.54,451.27) --
	(461.12,448.72) --
	(469.70,447.56) --
	(478.28,447.38) --
	(486.86,446.08) --
	(495.44,446.68) --
	(504.02,445.24) --
	(512.60,445.15) --
	(521.18,444.83) --
	(529.76,444.41) --
	(538.34,444.83) --
	(546.92,444.41) --
	(555.50,444.18) --
	(564.08,444.13) --
	(572.66,444.13);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.50) --
	( 49.28,445.01) --
	( 57.86,444.73) --
	( 66.44,444.87) --
	( 75.02,445.34) --
	( 83.60,445.75) --
	( 92.18,446.03) --
	(100.76,446.59) --
	(109.34,447.24) --
	(117.92,448.03) --
	(126.50,448.82) --
	(135.08,448.86) --
	(143.66,451.74) --
	(152.24,452.48) --
	(160.82,452.99) --
	(169.40,455.12) --
	(177.98,458.93) --
	(186.56,461.76) --
	(195.14,463.98) --
	(203.72,467.65) --
	(212.30,474.65) --
	(220.88,476.28) --
	(229.46,483.52) --
	(238.04,490.57) --
	(246.62,498.73) --
	(255.20,501.56) --
	(263.78,511.07) --
	(272.36,515.80) --
	(280.94,523.87) --
	(289.52,529.21) --
	(298.10,531.90) --
	(306.68,538.63) --
	(315.26,533.62) --
	(323.84,529.86) --
	(332.42,527.96) --
	(341.00,521.18) --
	(349.58,513.71) --
	(358.16,504.25) --
	(366.74,497.57) --
	(375.32,487.37) --
	(383.90,479.99) --
	(392.48,475.77) --
	(401.06,470.25) --
	(409.64,466.07) --
	(418.22,459.48) --
	(426.80,456.42) --
	(435.38,454.94) --
	(443.96,453.13) --
	(452.54,451.00) --
	(461.12,448.58) --
	(469.70,447.38) --
	(478.28,447.33) --
	(486.86,446.59) --
	(495.44,446.63) --
	(504.02,445.89) --
	(512.60,445.10) --
	(521.18,444.83) --
	(529.76,445.10) --
	(538.34,444.41) --
	(546.92,444.41) --
	(555.50,444.27) --
	(564.08,444.18) --
	(572.66,444.18);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.59) --
	( 49.28,444.64) --
	( 57.86,444.96) --
	( 66.44,445.06) --
	( 75.02,445.43) --
	( 83.60,445.66) --
	( 92.18,445.94) --
	(100.76,446.63) --
	(109.34,447.52) --
	(117.92,448.07) --
	(126.50,448.72) --
	(135.08,448.86) --
	(143.66,451.37) --
	(152.24,452.67) --
	(160.82,453.36) --
	(169.40,456.19) --
	(177.98,458.42) --
	(186.56,461.06) --
	(195.14,464.63) --
	(203.72,468.86) --
	(212.30,473.91) --
	(220.88,477.02) --
	(229.46,484.58) --
	(238.04,490.71) --
	(246.62,499.10) --
	(255.20,502.72) --
	(263.78,510.10) --
	(272.36,515.99) --
	(280.94,525.82) --
	(289.52,528.05) --
	(298.10,535.05) --
	(306.68,536.49) --
	(315.26,534.03) --
	(323.84,531.90) --
	(332.42,526.43) --
	(341.00,519.75) --
	(349.58,513.62) --
	(358.16,503.79) --
	(366.74,497.11) --
	(375.32,485.09) --
	(383.90,481.80) --
	(392.48,476.19) --
	(401.06,467.00) --
	(409.64,465.65) --
	(418.22,459.25) --
	(426.80,457.12) --
	(435.38,454.20) --
	(443.96,452.16) --
	(452.54,451.13) --
	(461.12,448.91) --
	(469.70,447.79) --
	(478.28,447.24) --
	(486.86,446.54) --
	(495.44,446.54) --
	(504.02,445.85) --
	(512.60,444.83) --
	(521.18,444.78) --
	(529.76,444.59) --
	(538.34,444.73) --
	(546.92,444.59) --
	(555.50,444.27) --
	(564.08,444.13) --
	(572.66,444.27);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.96) --
	( 49.28,444.22) --
	( 57.86,445.29) --
	( 66.44,445.34) --
	( 75.02,445.71) --
	( 83.60,445.75) --
	( 92.18,446.91) --
	(100.76,446.17) --
	(109.34,447.47) --
	(117.92,448.35) --
	(126.50,449.42) --
	(135.08,449.79) --
	(143.66,450.49) --
	(152.24,453.64) --
	(160.82,455.50) --
	(169.40,456.75) --
	(177.98,460.18) --
	(186.56,462.50) --
	(195.14,467.28) --
	(203.72,471.36) --
	(212.30,478.13) --
	(220.88,480.87) --
	(229.46,489.59) --
	(238.04,494.42) --
	(246.62,502.21) --
	(255.20,512.97) --
	(263.78,513.85) --
	(272.36,525.64) --
	(280.94,527.26) --
	(289.52,534.31) --
	(298.10,539.65) --
	(306.68,533.43) --
	(315.26,532.41) --
	(323.84,530.69) --
	(332.42,526.61) --
	(341.00,513.16) --
	(349.58,507.73) --
	(358.16,497.48) --
	(366.74,492.42) --
	(375.32,481.01) --
	(383.90,476.88) --
	(392.48,469.46) --
	(401.06,465.10) --
	(409.64,461.99) --
	(418.22,458.14) --
	(426.80,454.71) --
	(435.38,452.71) --
	(443.96,450.86) --
	(452.54,449.93) --
	(461.12,448.86) --
	(469.70,447.10) --
	(478.28,447.05) --
	(486.86,445.29) --
	(495.44,445.38) --
	(504.02,445.24) --
	(512.60,444.92) --
	(521.18,445.01) --
	(529.76,444.32) --
	(538.34,444.55) --
	(546.92,444.08) --
	(555.50,444.13) --
	(564.08,444.13);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.73) --
	( 49.28,444.32) --
	( 57.86,445.24) --
	( 66.44,445.43) --
	( 75.02,445.34) --
	( 83.60,446.40) --
	( 92.18,446.36) --
	(100.76,446.96) --
	(109.34,447.47) --
	(117.92,447.98) --
	(126.50,449.33) --
	(135.08,450.11) --
	(143.66,450.62) --
	(152.24,453.22) --
	(160.82,455.68) --
	(169.40,457.26) --
	(177.98,460.23) --
	(186.56,461.71) --
	(195.14,466.91) --
	(203.72,473.36) --
	(212.30,477.16) --
	(220.88,481.06) --
	(229.46,489.08) --
	(238.04,495.85) --
	(246.62,499.43) --
	(255.20,513.81) --
	(263.78,516.45) --
	(272.36,522.48) --
	(280.94,528.37) --
	(289.52,533.06) --
	(298.10,538.81) --
	(306.68,536.49) --
	(315.26,530.69) --
	(323.84,531.06) --
	(332.42,525.59) --
	(341.00,515.76) --
	(349.58,506.52) --
	(358.16,496.69) --
	(366.74,490.98) --
	(375.32,481.52) --
	(383.90,475.16) --
	(392.48,469.92) --
	(401.06,465.84) --
	(409.64,461.57) --
	(418.22,457.07) --
	(426.80,455.50) --
	(435.38,452.34) --
	(443.96,450.35) --
	(452.54,449.88) --
	(461.12,448.03) --
	(469.70,446.96) --
	(478.28,446.82) --
	(486.86,445.61) --
	(495.44,445.06) --
	(504.02,445.75) --
	(512.60,445.01) --
	(521.18,444.64) --
	(529.76,444.45) --
	(538.34,444.59) --
	(546.92,444.32) --
	(555.50,444.32) --
	(564.08,443.99) --
	(572.66,444.08);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.55) --
	( 49.28,444.73) --
	( 57.86,445.06) --
	( 66.44,445.38) --
	( 75.02,445.48) --
	( 83.60,446.50) --
	( 92.18,446.50) --
	(100.76,446.73) --
	(109.34,447.75) --
	(117.92,448.21) --
	(126.50,449.23) --
	(135.08,449.98) --
	(143.66,451.00) --
	(152.24,453.55) --
	(160.82,455.17) --
	(169.40,457.35) --
	(177.98,460.46) --
	(186.56,462.92) --
	(195.14,466.81) --
	(203.72,473.22) --
	(212.30,476.23) --
	(220.88,482.45) --
	(229.46,489.31) --
	(238.04,494.65) --
	(246.62,502.58) --
	(255.20,510.47) --
	(263.78,519.00) --
	(272.36,520.44) --
	(280.94,531.44) --
	(289.52,531.44) --
	(298.10,539.83) --
	(306.68,535.10) --
	(315.26,530.93) --
	(323.84,530.55) --
	(332.42,523.50) --
	(341.00,514.50) --
	(349.58,508.47) --
	(358.16,499.66) --
	(366.74,488.57) --
	(375.32,481.47) --
	(383.90,475.44) --
	(392.48,468.35) --
	(401.06,464.82) --
	(409.64,459.90) --
	(418.22,457.68) --
	(426.80,454.99) --
	(435.38,451.92) --
	(443.96,450.62) --
	(452.54,448.95) --
	(461.12,448.03) --
	(469.70,447.89) --
	(478.28,447.15) --
	(486.86,445.80) --
	(495.44,445.20) --
	(504.02,445.48) --
	(512.60,444.96) --
	(521.18,444.36) --
	(529.76,444.69) --
	(538.34,444.27) --
	(546.92,444.41) --
	(555.50,444.41) --
	(564.08,444.04) --
	(572.66,444.22);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,444.18) --
	( 49.28,444.78) --
	( 57.86,445.38) --
	( 66.44,445.20) --
	( 75.02,445.52) --
	( 83.60,446.36) --
	( 92.18,446.36) --
	(100.76,446.96) --
	(109.34,447.56) --
	(117.92,448.21) --
	(126.50,449.60) --
	(135.08,450.21) --
	(143.66,451.51) --
	(152.24,452.71) --
	(160.82,454.43) --
	(169.40,459.16) --
	(177.98,459.39) --
	(186.56,463.94) --
	(195.14,466.91) --
	(203.72,472.71) --
	(212.30,477.30) --
	(220.88,480.41) --
	(229.46,489.17) --
	(238.04,496.55) --
	(246.62,503.74) --
	(255.20,506.06) --
	(263.78,519.37) --
	(272.36,523.87) --
	(280.94,528.24) --
	(289.52,531.81) --
	(298.10,535.33) --
	(306.68,538.58) --
	(315.26,531.99) --
	(323.84,531.85) --
	(332.42,520.63) --
	(341.00,515.52) --
	(349.58,507.92) --
	(358.16,500.73) --
	(366.74,489.73) --
	(375.32,482.49) --
	(383.90,473.63) --
	(392.48,471.96) --
	(401.06,465.56) --
	(409.64,460.51) --
	(418.22,456.98) --
	(426.80,454.47) --
	(435.38,451.83) --
	(443.96,450.76) --
	(452.54,448.72) --
	(461.12,448.12) --
	(469.70,447.52) --
	(478.28,446.82) --
	(486.86,446.03) --
	(495.44,445.43) --
	(504.02,445.24) --
	(512.60,445.10) --
	(521.18,444.78) --
	(529.76,444.50) --
	(538.34,444.45) --
	(546.92,444.32) --
	(555.50,444.08) --
	(564.08,444.13) --
	(572.66,444.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.70,443.99) rectangle (572.66,555.28);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.75,464.11) {500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.75,487.30) {1000};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.75,510.50) {1500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 35.75,533.69) {2000};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.95,467.14) --
	( 40.70,467.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.95,490.33) --
	( 40.70,490.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.95,513.53) --
	( 40.70,513.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.95,536.72) --
	( 40.70,536.72);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.70,441.24) --
	( 40.70,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (126.50,441.24) --
	(126.50,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.30,441.24) --
	(212.30,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (298.10,441.24) --
	(298.10,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (383.90,441.24) --
	(383.90,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.70,441.24) --
	(469.70,443.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.50,441.24) --
	(555.50,443.99);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 40.70,432.98) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (126.50,432.98) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (212.30,432.98) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (298.10,432.98) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (383.90,432.98) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (469.70,432.98) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (555.50,432.98) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (306.68,420.80) {Time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,499.64) {Daily incidence};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);

\path[] ( 42.63,462.26) rectangle (123.87,548.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);

\path[] ( 42.63,462.26) rectangle (123.87,548.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.13,519.79) rectangle ( 65.48,537.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.87,528.47) -- ( 63.74,528.47);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.13,502.45) rectangle ( 65.48,519.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.87,511.12) -- ( 63.74,511.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.13,485.11) rectangle ( 65.48,502.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.87,493.78) -- ( 63.74,493.78);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.13,467.76) rectangle ( 65.48,485.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 49.87,476.43) -- ( 63.74,476.43);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 70.98,525.44) {$\rho=0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 70.98,508.09) {$\rho=0.25$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 70.98,490.75) {$\rho=0.5$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 70.98,473.40) {$\rho=0.75$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 40.70,563.57) {A};
\end{scope}
\begin{scope}
\path[clip] (  0.00,206.49) rectangle (192.72,412.97);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,206.49) rectangle (192.72,412.97);
\end{scope}
\begin{scope}
\path[clip] ( 27.50,237.50) rectangle (187.22,390.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.50,237.50) rectangle (187.22,390.09);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,356.51) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,356.56) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,358.10) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,357.21) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,357.77) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,357.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,358.16) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,358.00) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,357.73) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,357.71) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,357.65) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,358.32) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,357.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,358.54) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,355.14) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,356.97) circle (  2.50);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,356.32) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,357.12) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,357.42) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,357.42) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,357.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,357.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,358.17) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,357.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,357.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,357.65) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,357.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,358.40) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,357.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,358.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,357.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,355.80) circle (  2.50);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,356.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,357.51) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,357.35) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,357.48) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,357.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,358.14) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,358.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,357.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,357.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,357.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,357.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,358.25) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,358.04) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,358.31) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,357.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,354.33) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,357.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,357.12) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,357.17) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,357.73) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,357.54) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,358.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,358.01) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,357.91) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,357.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,357.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,357.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,358.42) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,357.88) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,358.05) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,356.27) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,358.69) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,356.48) --
	( 29.46,356.58) --
	( 31.41,356.67) --
	( 33.37,356.76) --
	( 35.32,356.84) --
	( 37.28,356.92) --
	( 39.24,357.00) --
	( 41.19,357.08) --
	( 43.15,357.15) --
	( 45.11,357.22) --
	( 47.06,357.28) --
	( 49.02,357.35) --
	( 50.98,357.41) --
	( 52.93,357.47) --
	( 54.89,357.52) --
	( 56.85,357.57) --
	( 58.80,357.62) --
	( 60.76,357.67) --
	( 62.72,357.71) --
	( 64.67,357.74) --
	( 66.63,357.78) --
	( 68.59,357.81) --
	( 70.54,357.84) --
	( 72.50,357.86) --
	( 74.46,357.88) --
	( 76.41,357.90) --
	( 78.37,357.92) --
	( 80.33,357.94) --
	( 82.28,357.94) --
	( 84.24,357.95) --
	( 86.20,357.94) --
	( 88.15,357.94) --
	( 90.11,357.93) --
	( 92.07,357.92) --
	( 94.02,357.91) --
	( 95.98,357.91) --
	( 97.94,357.90) --
	( 99.89,357.91) --
	(101.85,357.91) --
	(103.80,357.91) --
	(105.76,357.91) --
	(107.72,357.90) --
	(109.67,357.90) --
	(111.63,357.90) --
	(113.59,357.89) --
	(115.54,357.89) --
	(117.50,357.89) --
	(119.46,357.89) --
	(121.41,357.89) --
	(123.37,357.89) --
	(125.33,357.91) --
	(127.28,357.92) --
	(129.24,357.94) --
	(131.20,357.95) --
	(133.15,357.96) --
	(135.11,357.96) --
	(137.07,357.95) --
	(139.02,357.94) --
	(140.98,357.91) --
	(142.94,357.87) --
	(144.89,357.83) --
	(146.85,357.79) --
	(148.81,357.75) --
	(150.76,357.70) --
	(152.72,357.64) --
	(154.68,357.59) --
	(156.63,357.53) --
	(158.59,357.46) --
	(160.55,357.39) --
	(162.50,357.32) --
	(164.46,357.24) --
	(166.42,357.16) --
	(168.37,357.07) --
	(170.33,356.98) --
	(172.28,356.88) --
	(174.24,356.78) --
	(176.20,356.68) --
	(178.15,356.57) --
	(180.11,356.46) --
	(182.07,356.34);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,356.47) --
	( 29.46,356.56) --
	( 31.41,356.65) --
	( 33.37,356.73) --
	( 35.32,356.81) --
	( 37.28,356.89) --
	( 39.24,356.97) --
	( 41.19,357.04) --
	( 43.15,357.12) --
	( 45.11,357.18) --
	( 47.06,357.25) --
	( 49.02,357.31) --
	( 50.98,357.37) --
	( 52.93,357.43) --
	( 54.89,357.48) --
	( 56.85,357.54) --
	( 58.80,357.58) --
	( 60.76,357.63) --
	( 62.72,357.67) --
	( 64.67,357.71) --
	( 66.63,357.74) --
	( 68.59,357.78) --
	( 70.54,357.81) --
	( 72.50,357.83) --
	( 74.46,357.86) --
	( 76.41,357.88) --
	( 78.37,357.90) --
	( 80.33,357.92) --
	( 82.28,357.93) --
	( 84.24,357.93) --
	( 86.20,357.94) --
	( 88.15,357.94) --
	( 90.11,357.93) --
	( 92.07,357.93) --
	( 94.02,357.92) --
	( 95.98,357.92) --
	( 97.94,357.92) --
	( 99.89,357.92) --
	(101.85,357.92) --
	(103.80,357.92) --
	(105.76,357.91) --
	(107.72,357.90) --
	(109.67,357.89) --
	(111.63,357.88) --
	(113.59,357.87) --
	(115.54,357.86) --
	(117.50,357.85) --
	(119.46,357.85) --
	(121.41,357.85) --
	(123.37,357.86) --
	(125.33,357.88) --
	(127.28,357.91) --
	(129.24,357.94) --
	(131.20,357.97) --
	(133.15,357.99) --
	(135.11,358.01) --
	(137.07,358.02) --
	(139.02,358.01) --
	(140.98,357.99) --
	(142.94,357.97) --
	(144.89,357.94) --
	(146.85,357.91) --
	(148.81,357.87) --
	(150.76,357.84) --
	(152.72,357.79) --
	(154.68,357.74) --
	(156.63,357.69) --
	(158.59,357.63) --
	(160.55,357.56) --
	(162.50,357.48) --
	(164.46,357.39) --
	(166.42,357.30) --
	(168.37,357.20) --
	(170.33,357.10) --
	(172.28,356.99) --
	(174.24,356.87) --
	(176.20,356.75) --
	(178.15,356.63) --
	(180.11,356.49) --
	(182.07,356.35);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,357.02) --
	( 29.46,357.07) --
	( 31.41,357.12) --
	( 33.37,357.16) --
	( 35.32,357.21) --
	( 37.28,357.25) --
	( 39.24,357.29) --
	( 41.19,357.34) --
	( 43.15,357.38) --
	( 45.11,357.42) --
	( 47.06,357.46) --
	( 49.02,357.49) --
	( 50.98,357.53) --
	( 52.93,357.56) --
	( 54.89,357.60) --
	( 56.85,357.63) --
	( 58.80,357.66) --
	( 60.76,357.69) --
	( 62.72,357.72) --
	( 64.67,357.74) --
	( 66.63,357.77) --
	( 68.59,357.80) --
	( 70.54,357.82) --
	( 72.50,357.84) --
	( 74.46,357.86) --
	( 76.41,357.88) --
	( 78.37,357.90) --
	( 80.33,357.91) --
	( 82.28,357.92) --
	( 84.24,357.92) --
	( 86.20,357.93) --
	( 88.15,357.93) --
	( 90.11,357.93) --
	( 92.07,357.93) --
	( 94.02,357.92) --
	( 95.98,357.92) --
	( 97.94,357.92) --
	( 99.89,357.92) --
	(101.85,357.92) --
	(103.80,357.91) --
	(105.76,357.90) --
	(107.72,357.89) --
	(109.67,357.87) --
	(111.63,357.85) --
	(113.59,357.84) --
	(115.54,357.83) --
	(117.50,357.82) --
	(119.46,357.82) --
	(121.41,357.82) --
	(123.37,357.85) --
	(125.33,357.89) --
	(127.28,357.93) --
	(129.24,357.99) --
	(131.20,358.04) --
	(133.15,358.08) --
	(135.11,358.12) --
	(137.07,358.14) --
	(139.02,358.14) --
	(140.98,358.12) --
	(142.94,358.08) --
	(144.89,358.04) --
	(146.85,357.99) --
	(148.81,357.94) --
	(150.76,357.89) --
	(152.72,357.82) --
	(154.68,357.75) --
	(156.63,357.66) --
	(158.59,357.56) --
	(160.55,357.45) --
	(162.50,357.33) --
	(164.46,357.19) --
	(166.42,357.05) --
	(168.37,356.89) --
	(170.33,356.73) --
	(172.28,356.55) --
	(174.24,356.37) --
	(176.20,356.17) --
	(178.15,355.97) --
	(180.11,355.75) --
	(182.07,355.53);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,357.53) --
	( 29.46,357.52) --
	( 31.41,357.52) --
	( 33.37,357.52) --
	( 35.32,357.52) --
	( 37.28,357.52) --
	( 39.24,357.52) --
	( 41.19,357.53) --
	( 43.15,357.54) --
	( 45.11,357.54) --
	( 47.06,357.55) --
	( 49.02,357.56) --
	( 50.98,357.58) --
	( 52.93,357.59) --
	( 54.89,357.60) --
	( 56.85,357.61) --
	( 58.80,357.62) --
	( 60.76,357.64) --
	( 62.72,357.66) --
	( 64.67,357.69) --
	( 66.63,357.72) --
	( 68.59,357.75) --
	( 70.54,357.78) --
	( 72.50,357.81) --
	( 74.46,357.84) --
	( 76.41,357.86) --
	( 78.37,357.88) --
	( 80.33,357.90) --
	( 82.28,357.91) --
	( 84.24,357.91) --
	( 86.20,357.92) --
	( 88.15,357.91) --
	( 90.11,357.91) --
	( 92.07,357.91) --
	( 94.02,357.91) --
	( 95.98,357.90) --
	( 97.94,357.90) --
	( 99.89,357.90) --
	(101.85,357.91) --
	(103.80,357.91) --
	(105.76,357.90) --
	(107.72,357.90) --
	(109.67,357.90) --
	(111.63,357.90) --
	(113.59,357.89) --
	(115.54,357.89) --
	(117.50,357.89) --
	(119.46,357.89) --
	(121.41,357.88) --
	(123.37,357.88) --
	(125.33,357.88) --
	(127.28,357.87) --
	(129.24,357.86) --
	(131.20,357.86) --
	(133.15,357.85) --
	(135.11,357.84) --
	(137.07,357.83) --
	(139.02,357.82) --
	(140.98,357.82) --
	(142.94,357.81) --
	(144.89,357.81) --
	(146.85,357.80) --
	(148.81,357.79) --
	(150.76,357.79) --
	(152.72,357.78) --
	(154.68,357.78) --
	(156.63,357.77) --
	(158.59,357.77) --
	(160.55,357.77) --
	(162.50,357.77) --
	(164.46,357.78) --
	(166.42,357.78) --
	(168.37,357.79) --
	(170.33,357.79) --
	(172.28,357.80) --
	(174.24,357.81) --
	(176.20,357.81) --
	(178.15,357.82) --
	(180.11,357.83) --
	(182.07,357.84);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.50,237.50) rectangle (187.22,390.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,234.47) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,278.07) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,321.67) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,365.26) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,237.50) --
	( 27.50,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,281.10) --
	( 27.50,281.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,324.70) --
	( 27.50,324.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,368.29) --
	( 27.50,368.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.50,234.75) --
	( 27.50,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 53.26,234.75) --
	( 53.26,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.02,234.75) --
	( 79.02,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (104.78,234.75) --
	(104.78,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.54,234.75) --
	(130.54,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.31,234.75) --
	(156.31,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (182.07,234.75) --
	(182.07,237.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.50,226.49) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 53.26,226.49) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 79.02,226.49) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (104.78,226.49) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (130.54,226.49) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (156.31,226.49) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (182.07,226.49) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (107.36,214.31) {Primary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,313.80) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.50,398.38) {B. Incubation period};
\end{scope}
\begin{scope}
\path[clip] (192.72,206.49) rectangle (385.44,412.97);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (192.72,206.49) rectangle (385.44,412.97);
\end{scope}
\begin{scope}
\path[clip] (220.22,237.50) rectangle (379.94,390.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.22,237.50) rectangle (379.94,390.09);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,345.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,345.38) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,346.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,345.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,343.92) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,342.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,340.70) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,338.91) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,338.47) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,340.18) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,342.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,344.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,343.66) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,344.17) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,343.66) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,346.81) circle (  2.50);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,346.15) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,345.49) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,346.69) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,345.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,343.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,342.69) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,340.82) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,338.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,338.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,339.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,342.27) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,344.27) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,345.29) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,346.28) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,345.26) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,347.46) circle (  2.50);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,346.10) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,345.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,346.38) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,345.65) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,343.83) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,342.84) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,340.72) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,339.02) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,338.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,339.63) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,342.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,344.24) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,344.52) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,346.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,340.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,342.09) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,346.83) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,346.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,345.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,345.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,343.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,342.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,340.97) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,339.20) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,338.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,339.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,341.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,345.11) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,345.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,345.16) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,341.99) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,345.00) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,346.20) --
	(222.18,346.18) --
	(224.13,346.15) --
	(226.09,346.11) --
	(228.04,346.06) --
	(230.00,346.00) --
	(231.96,345.92) --
	(233.91,345.83) --
	(235.87,345.73) --
	(237.83,345.61) --
	(239.78,345.49) --
	(241.74,345.36) --
	(243.70,345.23) --
	(245.65,345.08) --
	(247.61,344.94) --
	(249.57,344.78) --
	(251.52,344.63) --
	(253.48,344.45) --
	(255.44,344.25) --
	(257.39,344.03) --
	(259.35,343.80) --
	(261.31,343.55) --
	(263.26,343.30) --
	(265.22,343.05) --
	(267.18,342.80) --
	(269.13,342.56) --
	(271.09,342.33) --
	(273.05,342.11) --
	(275.00,341.84) --
	(276.96,341.55) --
	(278.92,341.23) --
	(280.87,340.90) --
	(282.83,340.59) --
	(284.79,340.29) --
	(286.74,340.02) --
	(288.70,339.80) --
	(290.66,339.64) --
	(292.61,339.55) --
	(294.57,339.52) --
	(296.52,339.54) --
	(298.48,339.60) --
	(300.44,339.70) --
	(302.39,339.81) --
	(304.35,339.95) --
	(306.31,340.10) --
	(308.26,340.25) --
	(310.22,340.39) --
	(312.18,340.53) --
	(314.13,340.66) --
	(316.09,340.82) --
	(318.05,341.01) --
	(320.00,341.23) --
	(321.96,341.46) --
	(323.92,341.70) --
	(325.87,341.95) --
	(327.83,342.18) --
	(329.79,342.40) --
	(331.74,342.60) --
	(333.70,342.77) --
	(335.66,342.93) --
	(337.61,343.09) --
	(339.57,343.26) --
	(341.53,343.42) --
	(343.48,343.58) --
	(345.44,343.75) --
	(347.40,343.91) --
	(349.35,344.07) --
	(351.31,344.22) --
	(353.27,344.38) --
	(355.22,344.52) --
	(357.18,344.67) --
	(359.14,344.81) --
	(361.09,344.95) --
	(363.05,345.08) --
	(365.00,345.22) --
	(366.96,345.35) --
	(368.92,345.48) --
	(370.87,345.61) --
	(372.83,345.74) --
	(374.79,345.86);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,346.45) --
	(222.18,346.40) --
	(224.13,346.36) --
	(226.09,346.30) --
	(228.04,346.23) --
	(230.00,346.16) --
	(231.96,346.07) --
	(233.91,345.96) --
	(235.87,345.85) --
	(237.83,345.73) --
	(239.78,345.60) --
	(241.74,345.46) --
	(243.70,345.32) --
	(245.65,345.17) --
	(247.61,345.01) --
	(249.57,344.85) --
	(251.52,344.69) --
	(253.48,344.51) --
	(255.44,344.31) --
	(257.39,344.09) --
	(259.35,343.86) --
	(261.31,343.62) --
	(263.26,343.37) --
	(265.22,343.13) --
	(267.18,342.88) --
	(269.13,342.65) --
	(271.09,342.42) --
	(273.05,342.20) --
	(275.00,341.93) --
	(276.96,341.63) --
	(278.92,341.31) --
	(280.87,340.98) --
	(282.83,340.65) --
	(284.79,340.35) --
	(286.74,340.08) --
	(288.70,339.85) --
	(290.66,339.68) --
	(292.61,339.58) --
	(294.57,339.54) --
	(296.52,339.53) --
	(298.48,339.56) --
	(300.44,339.61) --
	(302.39,339.69) --
	(304.35,339.79) --
	(306.31,339.91) --
	(308.26,340.04) --
	(310.22,340.17) --
	(312.18,340.32) --
	(314.13,340.47) --
	(316.09,340.68) --
	(318.05,340.93) --
	(320.00,341.21) --
	(321.96,341.52) --
	(323.92,341.84) --
	(325.87,342.16) --
	(327.83,342.47) --
	(329.79,342.77) --
	(331.74,343.03) --
	(333.70,343.26) --
	(335.66,343.47) --
	(337.61,343.68) --
	(339.57,343.90) --
	(341.53,344.12) --
	(343.48,344.34) --
	(345.44,344.56) --
	(347.40,344.78) --
	(349.35,344.99) --
	(351.31,345.20) --
	(353.27,345.40) --
	(355.22,345.59) --
	(357.18,345.78) --
	(359.14,345.96) --
	(361.09,346.14) --
	(363.05,346.31) --
	(365.00,346.49) --
	(366.96,346.66) --
	(368.92,346.82) --
	(370.87,346.98) --
	(372.83,347.14) --
	(374.79,347.29);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,346.49) --
	(222.18,346.44) --
	(224.13,346.39) --
	(226.09,346.33) --
	(228.04,346.26) --
	(230.00,346.18) --
	(231.96,346.08) --
	(233.91,345.98) --
	(235.87,345.86) --
	(237.83,345.74) --
	(239.78,345.60) --
	(241.74,345.46) --
	(243.70,345.32) --
	(245.65,345.16) --
	(247.61,345.01) --
	(249.57,344.85) --
	(251.52,344.68) --
	(253.48,344.50) --
	(255.44,344.30) --
	(257.39,344.08) --
	(259.35,343.85) --
	(261.31,343.60) --
	(263.26,343.36) --
	(265.22,343.11) --
	(267.18,342.87) --
	(269.13,342.63) --
	(271.09,342.41) --
	(273.05,342.18) --
	(275.00,341.92) --
	(276.96,341.62) --
	(278.92,341.30) --
	(280.87,340.98) --
	(282.83,340.66) --
	(284.79,340.36) --
	(286.74,340.09) --
	(288.70,339.86) --
	(290.66,339.70) --
	(292.61,339.60) --
	(294.57,339.56) --
	(296.52,339.56) --
	(298.48,339.59) --
	(300.44,339.66) --
	(302.39,339.75) --
	(304.35,339.85) --
	(306.31,339.98) --
	(308.26,340.11) --
	(310.22,340.25) --
	(312.18,340.39) --
	(314.13,340.53) --
	(316.09,340.75) --
	(318.05,341.02) --
	(320.00,341.34) --
	(321.96,341.68) --
	(323.92,342.03) --
	(325.87,342.36) --
	(327.83,342.68) --
	(329.79,342.94) --
	(331.74,343.15) --
	(333.70,343.29) --
	(335.66,343.38) --
	(337.61,343.47) --
	(339.57,343.56) --
	(341.53,343.64) --
	(343.48,343.71) --
	(345.44,343.76) --
	(347.40,343.80) --
	(349.35,343.81) --
	(351.31,343.81) --
	(353.27,343.78) --
	(355.22,343.72) --
	(357.18,343.64) --
	(359.14,343.55) --
	(361.09,343.44) --
	(363.05,343.31) --
	(365.00,343.16) --
	(366.96,342.99) --
	(368.92,342.80) --
	(370.87,342.60) --
	(372.83,342.38) --
	(374.79,342.14);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,346.92) --
	(222.18,346.83) --
	(224.13,346.74) --
	(226.09,346.64) --
	(228.04,346.53) --
	(230.00,346.41) --
	(231.96,346.29) --
	(233.91,346.15) --
	(235.87,346.01) --
	(237.83,345.87) --
	(239.78,345.71) --
	(241.74,345.55) --
	(243.70,345.39) --
	(245.65,345.22) --
	(247.61,345.05) --
	(249.57,344.88) --
	(251.52,344.71) --
	(253.48,344.52) --
	(255.44,344.32) --
	(257.39,344.11) --
	(259.35,343.88) --
	(261.31,343.65) --
	(263.26,343.42) --
	(265.22,343.19) --
	(267.18,342.96) --
	(269.13,342.74) --
	(271.09,342.52) --
	(273.05,342.30) --
	(275.00,342.03) --
	(276.96,341.73) --
	(278.92,341.40) --
	(280.87,341.07) --
	(282.83,340.74) --
	(284.79,340.43) --
	(286.74,340.15) --
	(288.70,339.92) --
	(290.66,339.74) --
	(292.61,339.65) --
	(294.57,339.61) --
	(296.52,339.61) --
	(298.48,339.65) --
	(300.44,339.72) --
	(302.39,339.81) --
	(304.35,339.93) --
	(306.31,340.06) --
	(308.26,340.20) --
	(310.22,340.34) --
	(312.18,340.48) --
	(314.13,340.63) --
	(316.09,340.83) --
	(318.05,341.08) --
	(320.00,341.36) --
	(321.96,341.67) --
	(323.92,341.98) --
	(325.87,342.29) --
	(327.83,342.59) --
	(329.79,342.85) --
	(331.74,343.06) --
	(333.70,343.22) --
	(335.66,343.35) --
	(337.61,343.49) --
	(339.57,343.62) --
	(341.53,343.75) --
	(343.48,343.87) --
	(345.44,343.98) --
	(347.40,344.09) --
	(349.35,344.18) --
	(351.31,344.26) --
	(353.27,344.33) --
	(355.22,344.37) --
	(357.18,344.41) --
	(359.14,344.43) --
	(361.09,344.45) --
	(363.05,344.45) --
	(365.00,344.44) --
	(366.96,344.42) --
	(368.92,344.39) --
	(370.87,344.35) --
	(372.83,344.30) --
	(374.79,344.23);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.22,237.50) rectangle (379.94,390.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,234.47) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,278.07) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,321.67) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,365.26) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,237.50) --
	(220.22,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,281.10) --
	(220.22,281.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,324.70) --
	(220.22,324.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,368.29) --
	(220.22,368.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.22,234.75) --
	(220.22,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (245.98,234.75) --
	(245.98,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.74,234.75) --
	(271.74,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (297.50,234.75) --
	(297.50,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.26,234.75) --
	(323.26,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (349.03,234.75) --
	(349.03,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (374.79,234.75) --
	(374.79,237.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (220.22,226.49) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.98,226.49) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.74,226.49) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (297.50,226.49) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (323.26,226.49) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (349.03,226.49) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (374.79,226.49) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (300.08,214.31) {Primary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (205.80,313.80) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (220.22,398.38) {C. Generation interval};
\end{scope}
\begin{scope}
\path[clip] (385.44,206.49) rectangle (578.16,412.97);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (385.44,206.49) rectangle (578.16,412.97);
\end{scope}
\begin{scope}
\path[clip] (412.94,237.50) rectangle (572.66,390.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (412.94,237.50) rectangle (572.66,390.09);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94,370.30) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24,362.63) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55,365.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85,364.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16,363.04) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46,361.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77,357.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07,351.58) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,342.72) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,332.52) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,321.56) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,310.21) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,301.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,297.38) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,287.12) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,277.64) circle (  2.50);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94,366.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24,358.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55,362.30) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85,361.25) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16,360.11) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46,358.56) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77,356.17) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07,351.21) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,343.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,336.10) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,326.81) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,318.72) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,310.21) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,308.72) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,299.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,293.53) circle (  2.50);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94,358.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24,355.20) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55,358.93) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85,357.01) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16,356.84) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46,356.12) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77,354.21) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07,350.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,345.40) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,339.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,332.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,326.12) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,320.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,316.84) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,311.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,304.21) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94,355.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24,352.16) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55,352.55) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85,354.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16,353.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46,353.31) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77,352.06) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07,350.45) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,346.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,342.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,338.49) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,334.64) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,330.13) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,328.08) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,323.89) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,320.20) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94,366.60) --
	(414.90,366.74) --
	(416.85,366.85) --
	(418.81,366.93) --
	(420.76,366.99) --
	(422.72,367.02) --
	(424.68,367.03) --
	(426.63,367.02) --
	(428.59,366.98) --
	(430.55,366.91) --
	(432.50,366.82) --
	(434.46,366.69) --
	(436.42,366.54) --
	(438.37,366.34) --
	(440.33,366.11) --
	(442.29,365.85) --
	(444.24,365.54) --
	(446.20,365.23) --
	(448.16,364.93) --
	(450.11,364.62) --
	(452.07,364.30) --
	(454.03,363.95) --
	(455.98,363.57) --
	(457.94,363.14) --
	(459.90,362.65) --
	(461.85,362.09) --
	(463.81,361.45) --
	(465.77,360.74) --
	(467.72,359.97) --
	(469.68,359.16) --
	(471.64,358.29) --
	(473.59,357.37) --
	(475.55,356.39) --
	(477.51,355.35) --
	(479.46,354.26) --
	(481.42,353.10) --
	(483.38,351.87) --
	(485.33,350.58) --
	(487.29,349.16) --
	(489.24,347.59) --
	(491.20,345.90) --
	(493.16,344.11) --
	(495.11,342.26) --
	(497.07,340.37) --
	(499.03,338.46) --
	(500.98,336.56) --
	(502.94,334.71) --
	(504.90,332.91) --
	(506.85,331.18) --
	(508.81,329.39) --
	(510.77,327.54) --
	(512.72,325.65) --
	(514.68,323.73) --
	(516.64,321.80) --
	(518.59,319.87) --
	(520.55,317.97) --
	(522.51,316.10) --
	(524.46,314.28) --
	(526.42,312.53) --
	(528.38,310.82) --
	(530.33,309.10) --
	(532.29,307.39) --
	(534.25,305.68) --
	(536.20,303.98) --
	(538.16,302.29) --
	(540.12,300.61) --
	(542.07,298.94) --
	(544.03,297.29) --
	(545.99,295.65) --
	(547.94,294.03) --
	(549.90,292.42) --
	(551.86,290.82) --
	(553.81,289.23) --
	(555.77,287.66) --
	(557.72,286.09) --
	(559.68,284.54) --
	(561.64,283.00) --
	(563.59,281.47) --
	(565.55,279.95) --
	(567.51,278.44);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94,363.23) --
	(414.90,363.29) --
	(416.85,363.32) --
	(418.81,363.33) --
	(420.76,363.33) --
	(422.72,363.31) --
	(424.68,363.29) --
	(426.63,363.24) --
	(428.59,363.18) --
	(430.55,363.10) --
	(432.50,363.00) --
	(434.46,362.88) --
	(436.42,362.73) --
	(438.37,362.56) --
	(440.33,362.37) --
	(442.29,362.14) --
	(444.24,361.89) --
	(446.20,361.64) --
	(448.16,361.41) --
	(450.11,361.18) --
	(452.07,360.94) --
	(454.03,360.69) --
	(455.98,360.40) --
	(457.94,360.09) --
	(459.90,359.72) --
	(461.85,359.30) --
	(463.81,358.81) --
	(465.77,358.25) --
	(467.72,357.65) --
	(469.68,357.02) --
	(471.64,356.34) --
	(473.59,355.62) --
	(475.55,354.85) --
	(477.51,354.04) --
	(479.46,353.18) --
	(481.42,352.26) --
	(483.38,351.30) --
	(485.33,350.27) --
	(487.29,349.15) --
	(489.24,347.90) --
	(491.20,346.55) --
	(493.16,345.13) --
	(495.11,343.65) --
	(497.07,342.14) --
	(499.03,340.62) --
	(500.98,339.10) --
	(502.94,337.62) --
	(504.90,336.19) --
	(506.85,334.81) --
	(508.81,333.37) --
	(510.77,331.88) --
	(512.72,330.36) --
	(514.68,328.81) --
	(516.64,327.26) --
	(518.59,325.71) --
	(520.55,324.18) --
	(522.51,322.69) --
	(524.46,321.25) --
	(526.42,319.87) --
	(528.38,318.52) --
	(530.33,317.18) --
	(532.29,315.84) --
	(534.25,314.51) --
	(536.20,313.18) --
	(538.16,311.87) --
	(540.12,310.57) --
	(542.07,309.28) --
	(544.03,308.01) --
	(545.99,306.76) --
	(547.94,305.52) --
	(549.90,304.30) --
	(551.86,303.09) --
	(553.81,301.89) --
	(555.77,300.71) --
	(557.72,299.54) --
	(559.68,298.38) --
	(561.64,297.24) --
	(563.59,296.11) --
	(565.55,294.99) --
	(567.51,293.88);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94,356.81) --
	(414.90,357.05) --
	(416.85,357.26) --
	(418.81,357.46) --
	(420.76,357.63) --
	(422.72,357.79) --
	(424.68,357.92) --
	(426.63,358.04) --
	(428.59,358.14) --
	(430.55,358.22) --
	(432.50,358.27) --
	(434.46,358.30) --
	(436.42,358.31) --
	(438.37,358.29) --
	(440.33,358.25) --
	(442.29,358.18) --
	(444.24,358.08) --
	(446.20,357.98) --
	(448.16,357.86) --
	(450.11,357.74) --
	(452.07,357.60) --
	(454.03,357.45) --
	(455.98,357.26) --
	(457.94,357.05) --
	(459.90,356.79) --
	(461.85,356.50) --
	(463.81,356.16) --
	(465.77,355.77) --
	(467.72,355.35) --
	(469.68,354.90) --
	(471.64,354.42) --
	(473.59,353.91) --
	(475.55,353.35) --
	(477.51,352.77) --
	(479.46,352.14) --
	(481.42,351.48) --
	(483.38,350.77) --
	(485.33,350.02) --
	(487.29,349.20) --
	(489.24,348.28) --
	(491.20,347.29) --
	(493.16,346.25) --
	(495.11,345.16) --
	(497.07,344.05) --
	(499.03,342.92) --
	(500.98,341.79) --
	(502.94,340.69) --
	(504.90,339.61) --
	(506.85,338.57) --
	(508.81,337.49) --
	(510.77,336.37) --
	(512.72,335.22) --
	(514.68,334.05) --
	(516.64,332.88) --
	(518.59,331.70) --
	(520.55,330.54) --
	(522.51,329.39) --
	(524.46,328.26) --
	(526.42,327.18) --
	(528.38,326.11) --
	(530.33,325.03) --
	(532.29,323.96) --
	(534.25,322.88) --
	(536.20,321.81) --
	(538.16,320.74) --
	(540.12,319.66) --
	(542.07,318.60) --
	(544.03,317.53) --
	(545.99,316.47) --
	(547.94,315.42) --
	(549.90,314.37) --
	(551.86,313.32) --
	(553.81,312.27) --
	(555.77,311.22) --
	(557.72,310.18) --
	(559.68,309.14) --
	(561.64,308.11) --
	(563.59,307.07) --
	(565.55,306.04) --
	(567.51,305.01);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94,353.87) --
	(414.90,353.94) --
	(416.85,354.00) --
	(418.81,354.05) --
	(420.76,354.09) --
	(422.72,354.13) --
	(424.68,354.16) --
	(426.63,354.19) --
	(428.59,354.21) --
	(430.55,354.22) --
	(432.50,354.23) --
	(434.46,354.22) --
	(436.42,354.21) --
	(438.37,354.18) --
	(440.33,354.14) --
	(442.29,354.09) --
	(444.24,354.02) --
	(446.20,353.96) --
	(448.16,353.92) --
	(450.11,353.88) --
	(452.07,353.83) --
	(454.03,353.79) --
	(455.98,353.73) --
	(457.94,353.65) --
	(459.90,353.54) --
	(461.85,353.41) --
	(463.81,353.24) --
	(465.77,353.03) --
	(467.72,352.80) --
	(469.68,352.56) --
	(471.64,352.29) --
	(473.59,352.00) --
	(475.55,351.69) --
	(477.51,351.35) --
	(479.46,350.99) --
	(481.42,350.59) --
	(483.38,350.16) --
	(485.33,349.71) --
	(487.29,349.20) --
	(489.24,348.63) --
	(491.20,348.02) --
	(493.16,347.37) --
	(495.11,346.69) --
	(497.07,345.99) --
	(499.03,345.28) --
	(500.98,344.57) --
	(502.94,343.86) --
	(504.90,343.18) --
	(506.85,342.50) --
	(508.81,341.79) --
	(510.77,341.05) --
	(512.72,340.29) --
	(514.68,339.51) --
	(516.64,338.72) --
	(518.59,337.93) --
	(520.55,337.15) --
	(522.51,336.37) --
	(524.46,335.63) --
	(526.42,334.90) --
	(528.38,334.19) --
	(530.33,333.48) --
	(532.29,332.77) --
	(534.25,332.05) --
	(536.20,331.34) --
	(538.16,330.62) --
	(540.12,329.91) --
	(542.07,329.20) --
	(544.03,328.49) --
	(545.99,327.79) --
	(547.94,327.09) --
	(549.90,326.39) --
	(551.86,325.70) --
	(553.81,325.01) --
	(555.77,324.32) --
	(557.72,323.63) --
	(559.68,322.95) --
	(561.64,322.27) --
	(563.59,321.59) --
	(565.55,320.91) --
	(567.51,320.23);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (412.94,237.50) rectangle (572.66,390.09);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,234.47) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,278.07) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,321.67) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,365.26) {6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,237.50) --
	(412.94,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,281.10) --
	(412.94,281.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,324.70) --
	(412.94,324.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,368.29) --
	(412.94,368.29);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.94,234.75) --
	(412.94,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (438.70,234.75) --
	(438.70,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.46,234.75) --
	(464.46,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.22,234.75) --
	(490.22,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (515.98,234.75) --
	(515.98,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (541.75,234.75) --
	(541.75,237.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (567.51,234.75) --
	(567.51,237.50);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (412.94,226.49) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (438.70,226.49) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (464.46,226.49) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (490.22,226.49) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (515.98,226.49) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (541.75,226.49) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (567.51,226.49) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (492.80,214.31) {Primary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (398.52,313.80) {Forward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (412.94,398.38) {D. Serial interval};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (192.72,206.49);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (192.72,206.49);
\end{scope}
\begin{scope}
\path[clip] ( 27.50, 31.02) rectangle (187.22,183.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 27.50, 31.02) rectangle (187.22,183.61);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,101.49) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,106.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,104.49) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,104.63) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,105.02) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,105.31) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,106.02) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,107.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,110.77) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,115.44) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,121.91) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,130.13) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,138.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,144.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,150.48) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,156.80) circle (  2.50);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,101.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,106.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,104.40) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,104.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,104.88) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,105.41) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,106.06) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,107.83) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,110.81) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,115.33) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,121.82) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,129.99) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,138.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,144.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,150.64) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,155.88) circle (  2.50);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,101.97) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,106.47) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,104.81) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,104.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,105.01) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,105.31) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,106.08) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,107.82) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,110.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,115.34) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,121.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,130.05) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,138.48) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,144.63) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,149.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,156.36) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 27.50,101.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 37.80,106.78) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.11,105.16) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.41,104.02) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.72,105.06) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 79.02,105.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 89.33,106.14) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 99.63,107.79) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.94,110.78) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (120.24,115.27) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (130.54,121.64) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.85,130.00) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.15,138.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (161.46,144.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.76,149.99) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.07,157.07) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,103.84) --
	( 29.46,103.83) --
	( 31.41,103.83) --
	( 33.37,103.83) --
	( 35.32,103.84) --
	( 37.28,103.85) --
	( 39.24,103.86) --
	( 41.19,103.87) --
	( 43.15,103.89) --
	( 45.11,103.91) --
	( 47.06,103.94) --
	( 49.02,103.97) --
	( 50.98,104.02) --
	( 52.93,104.07) --
	( 54.89,104.13) --
	( 56.85,104.20) --
	( 58.80,104.29) --
	( 60.76,104.36) --
	( 62.72,104.42) --
	( 64.67,104.46) --
	( 66.63,104.51) --
	( 68.59,104.55) --
	( 70.54,104.61) --
	( 72.50,104.68) --
	( 74.46,104.78) --
	( 76.41,104.91) --
	( 78.37,105.08) --
	( 80.33,105.28) --
	( 82.28,105.46) --
	( 84.24,105.63) --
	( 86.20,105.80) --
	( 88.15,105.98) --
	( 90.11,106.19) --
	( 92.07,106.44) --
	( 94.02,106.74) --
	( 95.98,107.10) --
	( 97.94,107.53) --
	( 99.89,108.05) --
	(101.85,108.63) --
	(103.80,109.26) --
	(105.76,109.93) --
	(107.72,110.64) --
	(109.67,111.40) --
	(111.63,112.20) --
	(113.59,113.04) --
	(115.54,113.92) --
	(117.50,114.83) --
	(119.46,115.79) --
	(121.41,116.80) --
	(123.37,117.89) --
	(125.33,119.07) --
	(127.28,120.31) --
	(129.24,121.59) --
	(131.20,122.92) --
	(133.15,124.26) --
	(135.11,125.60) --
	(137.07,126.92) --
	(139.02,128.22) --
	(140.98,129.48) --
	(142.94,130.72) --
	(144.89,131.99) --
	(146.85,133.29) --
	(148.81,134.60) --
	(150.76,135.92) --
	(152.72,137.26) --
	(154.68,138.61) --
	(156.63,139.96) --
	(158.59,141.32) --
	(160.55,142.67) --
	(162.50,144.03) --
	(164.46,145.39) --
	(166.42,146.76) --
	(168.37,148.13) --
	(170.33,149.51) --
	(172.28,150.90) --
	(174.24,152.30) --
	(176.20,153.70) --
	(178.15,155.12) --
	(180.11,156.54) --
	(182.07,157.96);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,104.00) --
	( 29.46,103.98) --
	( 31.41,103.96) --
	( 33.37,103.95) --
	( 35.32,103.94) --
	( 37.28,103.94) --
	( 39.24,103.94) --
	( 41.19,103.94) --
	( 43.15,103.95) --
	( 45.11,103.96) --
	( 47.06,103.98) --
	( 49.02,104.01) --
	( 50.98,104.04) --
	( 52.93,104.09) --
	( 54.89,104.14) --
	( 56.85,104.21) --
	( 58.80,104.29) --
	( 60.76,104.35) --
	( 62.72,104.41) --
	( 64.67,104.45) --
	( 66.63,104.49) --
	( 68.59,104.53) --
	( 70.54,104.59) --
	( 72.50,104.66) --
	( 74.46,104.76) --
	( 76.41,104.89) --
	( 78.37,105.06) --
	( 80.33,105.26) --
	( 82.28,105.43) --
	( 84.24,105.60) --
	( 86.20,105.78) --
	( 88.15,105.96) --
	( 90.11,106.17) --
	( 92.07,106.42) --
	( 94.02,106.72) --
	( 95.98,107.08) --
	( 97.94,107.51) --
	( 99.89,108.03) --
	(101.85,108.61) --
	(103.80,109.23) --
	(105.76,109.89) --
	(107.72,110.60) --
	(109.67,111.35) --
	(111.63,112.14) --
	(113.59,112.97) --
	(115.54,113.84) --
	(117.50,114.76) --
	(119.46,115.71) --
	(121.41,116.71) --
	(123.37,117.81) --
	(125.33,119.00) --
	(127.28,120.24) --
	(129.24,121.54) --
	(131.20,122.88) --
	(133.15,124.22) --
	(135.11,125.57) --
	(137.07,126.91) --
	(139.02,128.21) --
	(140.98,129.45) --
	(142.94,130.69) --
	(144.89,131.96) --
	(146.85,133.24) --
	(148.81,134.54) --
	(150.76,135.86) --
	(152.72,137.18) --
	(154.68,138.52) --
	(156.63,139.86) --
	(158.59,141.19) --
	(160.55,142.53) --
	(162.50,143.86) --
	(164.46,145.19) --
	(166.42,146.53) --
	(168.37,147.87) --
	(170.33,149.22) --
	(172.28,150.58) --
	(174.24,151.94) --
	(176.20,153.30) --
	(178.15,154.67) --
	(180.11,156.05) --
	(182.07,157.43);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,104.08) --
	( 29.46,104.04) --
	( 31.41,104.02) --
	( 33.37,104.01) --
	( 35.32,103.99) --
	( 37.28,103.99) --
	( 39.24,103.98) --
	( 41.19,103.98) --
	( 43.15,103.98) --
	( 45.11,103.98) --
	( 47.06,104.00) --
	( 49.02,104.02) --
	( 50.98,104.05) --
	( 52.93,104.09) --
	( 54.89,104.15) --
	( 56.85,104.21) --
	( 58.80,104.29) --
	( 60.76,104.35) --
	( 62.72,104.40) --
	( 64.67,104.45) --
	( 66.63,104.49) --
	( 68.59,104.53) --
	( 70.54,104.59) --
	( 72.50,104.67) --
	( 74.46,104.77) --
	( 76.41,104.90) --
	( 78.37,105.07) --
	( 80.33,105.26) --
	( 82.28,105.44) --
	( 84.24,105.61) --
	( 86.20,105.78) --
	( 88.15,105.96) --
	( 90.11,106.17) --
	( 92.07,106.41) --
	( 94.02,106.71) --
	( 95.98,107.07) --
	( 97.94,107.50) --
	( 99.89,108.01) --
	(101.85,108.59) --
	(103.80,109.21) --
	(105.76,109.88) --
	(107.72,110.59) --
	(109.67,111.35) --
	(111.63,112.14) --
	(113.59,112.98) --
	(115.54,113.85) --
	(117.50,114.77) --
	(119.46,115.72) --
	(121.41,116.72) --
	(123.37,117.81) --
	(125.33,118.98) --
	(127.28,120.22) --
	(129.24,121.51) --
	(131.20,122.83) --
	(133.15,124.17) --
	(135.11,125.50) --
	(137.07,126.83) --
	(139.02,128.12) --
	(140.98,129.36) --
	(142.94,130.60) --
	(144.89,131.85) --
	(146.85,133.13) --
	(148.81,134.43) --
	(150.76,135.74) --
	(152.72,137.07) --
	(154.68,138.40) --
	(156.63,139.73) --
	(158.59,141.07) --
	(160.55,142.41) --
	(162.50,143.74) --
	(164.46,145.08) --
	(166.42,146.43) --
	(168.37,147.78) --
	(170.33,149.14) --
	(172.28,150.51) --
	(174.24,151.88) --
	(176.20,153.26) --
	(178.15,154.64) --
	(180.11,156.03) --
	(182.07,157.43);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 27.50,104.10) --
	( 29.46,104.07) --
	( 31.41,104.05) --
	( 33.37,104.04) --
	( 35.32,104.03) --
	( 37.28,104.02) --
	( 39.24,104.02) --
	( 41.19,104.01) --
	( 43.15,104.02) --
	( 45.11,104.03) --
	( 47.06,104.04) --
	( 49.02,104.06) --
	( 50.98,104.10) --
	( 52.93,104.14) --
	( 54.89,104.19) --
	( 56.85,104.25) --
	( 58.80,104.33) --
	( 60.76,104.39) --
	( 62.72,104.44) --
	( 64.67,104.48) --
	( 66.63,104.51) --
	( 68.59,104.55) --
	( 70.54,104.60) --
	( 72.50,104.67) --
	( 74.46,104.77) --
	( 76.41,104.89) --
	( 78.37,105.06) --
	( 80.33,105.25) --
	( 82.28,105.43) --
	( 84.24,105.60) --
	( 86.20,105.77) --
	( 88.15,105.95) --
	( 90.11,106.17) --
	( 92.07,106.42) --
	( 94.02,106.71) --
	( 95.98,107.07) --
	( 97.94,107.51) --
	( 99.89,108.02) --
	(101.85,108.60) --
	(103.80,109.21) --
	(105.76,109.87) --
	(107.72,110.57) --
	(109.67,111.32) --
	(111.63,112.10) --
	(113.59,112.93) --
	(115.54,113.79) --
	(117.50,114.70) --
	(119.46,115.65) --
	(121.41,116.66) --
	(123.37,117.75) --
	(125.33,118.92) --
	(127.28,120.16) --
	(129.24,121.45) --
	(131.20,122.78) --
	(133.15,124.12) --
	(135.11,125.46) --
	(137.07,126.79) --
	(139.02,128.09) --
	(140.98,129.35) --
	(142.94,130.60) --
	(144.89,131.87) --
	(146.85,133.17) --
	(148.81,134.48) --
	(150.76,135.81) --
	(152.72,137.16) --
	(154.68,138.51) --
	(156.63,139.87) --
	(158.59,141.24) --
	(160.55,142.60) --
	(162.50,143.96) --
	(164.46,145.33) --
	(166.42,146.71) --
	(168.37,148.10) --
	(170.33,149.49) --
	(172.28,150.89) --
	(174.24,152.30) --
	(176.20,153.72) --
	(178.15,155.15) --
	(180.11,156.58) --
	(182.07,158.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 27.50, 31.02) rectangle (187.22,183.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55, 27.99) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55, 59.78) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55, 91.57) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,123.36) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 22.55,155.14) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75, 31.02) --
	( 27.50, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75, 62.81) --
	( 27.50, 62.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75, 94.60) --
	( 27.50, 94.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,126.39) --
	( 27.50,126.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.75,158.18) --
	( 27.50,158.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 27.50, 28.27) --
	( 27.50, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 53.26, 28.27) --
	( 53.26, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 79.02, 28.27) --
	( 79.02, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (104.78, 28.27) --
	(104.78, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.54, 28.27) --
	(130.54, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (156.31, 28.27) --
	(156.31, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (182.07, 28.27) --
	(182.07, 31.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 27.50, 20.01) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 53.26, 20.01) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 79.02, 20.01) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (104.78, 20.01) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (130.54, 20.01) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (156.31, 20.01) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (182.07, 20.01) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (107.36,  7.82) {Secondary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,107.31) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 27.50,191.89) {E. Incubation period};
\end{scope}
\begin{scope}
\path[clip] (192.72,  0.00) rectangle (385.44,206.49);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (192.72,  0.00) rectangle (385.44,206.49);
\end{scope}
\begin{scope}
\path[clip] (220.22, 31.02) rectangle (379.94,183.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (220.22, 31.02) rectangle (379.94,183.61);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,101.22) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,101.82) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,100.26) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,101.18) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,101.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,101.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,101.99) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,103.24) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,105.32) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,108.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,113.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,118.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,123.49) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,126.94) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,127.52) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,125.22) circle (  2.50);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,101.78) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,102.13) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,100.33) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,101.19) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,101.78) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,101.65) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,101.92) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,103.33) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,105.34) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,108.63) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,113.57) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,118.50) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,123.11) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,126.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,129.23) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,128.52) circle (  2.50);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,101.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,102.44) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,100.72) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,100.99) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,101.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,101.46) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,101.91) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,103.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,105.25) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,108.68) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,113.26) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,118.87) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,122.84) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,127.31) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,128.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,128.33) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.22,101.29) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (230.52,102.77) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (240.83,100.97) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.13,100.91) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.44,101.51) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.74,101.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.05,101.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.35,103.48) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.66,105.47) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (312.96,108.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.26,113.34) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.57,118.62) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (343.87,123.31) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.18,127.28) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (364.48,128.53) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.79,130.13) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,101.67) --
	(222.18,101.54) --
	(224.13,101.42) --
	(226.09,101.31) --
	(228.04,101.21) --
	(230.00,101.12) --
	(231.96,101.04) --
	(233.91,100.96) --
	(235.87,100.90) --
	(237.83,100.84) --
	(239.78,100.79) --
	(241.74,100.75) --
	(243.70,100.73) --
	(245.65,100.71) --
	(247.61,100.71) --
	(249.57,100.72) --
	(251.52,100.74) --
	(253.48,100.76) --
	(255.44,100.79) --
	(257.39,100.82) --
	(259.35,100.86) --
	(261.31,100.91) --
	(263.26,100.97) --
	(265.22,101.05) --
	(267.18,101.14) --
	(269.13,101.25) --
	(271.09,101.38) --
	(273.05,101.52) --
	(275.00,101.65) --
	(276.96,101.77) --
	(278.92,101.89) --
	(280.87,102.02) --
	(282.83,102.17) --
	(284.79,102.34) --
	(286.74,102.55) --
	(288.70,102.80) --
	(290.66,103.10) --
	(292.61,103.46) --
	(294.57,103.88) --
	(296.52,104.34) --
	(298.48,104.84) --
	(300.44,105.38) --
	(302.39,105.95) --
	(304.35,106.55) --
	(306.31,107.16) --
	(308.26,107.80) --
	(310.22,108.44) --
	(312.18,109.10) --
	(314.13,109.77) --
	(316.09,110.53) --
	(318.05,111.37) --
	(320.00,112.27) --
	(321.96,113.20) --
	(323.92,114.14) --
	(325.87,115.08) --
	(327.83,115.98) --
	(329.79,116.83) --
	(331.74,117.61) --
	(333.70,118.29) --
	(335.66,118.92) --
	(337.61,119.56) --
	(339.57,120.19) --
	(341.53,120.82) --
	(343.48,121.43) --
	(345.44,122.02) --
	(347.40,122.59) --
	(349.35,123.13) --
	(351.31,123.65) --
	(353.27,124.12) --
	(355.22,124.56) --
	(357.18,124.98) --
	(359.14,125.36) --
	(361.09,125.73) --
	(363.05,126.07) --
	(365.00,126.39) --
	(366.96,126.68) --
	(368.92,126.95) --
	(370.87,127.20) --
	(372.83,127.42) --
	(374.79,127.62);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,102.15) --
	(222.18,101.99) --
	(224.13,101.84) --
	(226.09,101.70) --
	(228.04,101.57) --
	(230.00,101.45) --
	(231.96,101.34) --
	(233.91,101.24) --
	(235.87,101.15) --
	(237.83,101.07) --
	(239.78,101.00) --
	(241.74,100.94) --
	(243.70,100.89) --
	(245.65,100.86) --
	(247.61,100.84) --
	(249.57,100.83) --
	(251.52,100.83) --
	(253.48,100.85) --
	(255.44,100.86) --
	(257.39,100.89) --
	(259.35,100.92) --
	(261.31,100.96) --
	(263.26,101.02) --
	(265.22,101.09) --
	(267.18,101.18) --
	(269.13,101.28) --
	(271.09,101.41) --
	(273.05,101.55) --
	(275.00,101.68) --
	(276.96,101.80) --
	(278.92,101.93) --
	(280.87,102.06) --
	(282.83,102.22) --
	(284.79,102.40) --
	(286.74,102.61) --
	(288.70,102.86) --
	(290.66,103.16) --
	(292.61,103.52) --
	(294.57,103.93) --
	(296.52,104.38) --
	(298.48,104.87) --
	(300.44,105.39) --
	(302.39,105.94) --
	(304.35,106.52) --
	(306.31,107.12) --
	(308.26,107.74) --
	(310.22,108.37) --
	(312.18,109.01) --
	(314.13,109.68) --
	(316.09,110.42) --
	(318.05,111.22) --
	(320.00,112.08) --
	(321.96,112.96) --
	(323.92,113.87) --
	(325.87,114.77) --
	(327.83,115.65) --
	(329.79,116.49) --
	(331.74,117.29) --
	(333.70,118.01) --
	(335.66,118.70) --
	(337.61,119.40) --
	(339.57,120.10) --
	(341.53,120.80) --
	(343.48,121.49) --
	(345.44,122.18) --
	(347.40,122.85) --
	(349.35,123.51) --
	(351.31,124.14) --
	(353.27,124.76) --
	(355.22,125.35) --
	(357.18,125.92) --
	(359.14,126.49) --
	(361.09,127.03) --
	(363.05,127.57) --
	(365.00,128.09) --
	(366.96,128.59) --
	(368.92,129.09) --
	(370.87,129.57) --
	(372.83,130.03) --
	(374.79,130.48);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,102.09) --
	(222.18,101.94) --
	(224.13,101.80) --
	(226.09,101.67) --
	(228.04,101.55) --
	(230.00,101.44) --
	(231.96,101.34) --
	(233.91,101.24) --
	(235.87,101.16) --
	(237.83,101.08) --
	(239.78,101.02) --
	(241.74,100.96) --
	(243.70,100.92) --
	(245.65,100.89) --
	(247.61,100.87) --
	(249.57,100.86) --
	(251.52,100.87) --
	(253.48,100.88) --
	(255.44,100.89) --
	(257.39,100.91) --
	(259.35,100.93) --
	(261.31,100.96) --
	(263.26,101.01) --
	(265.22,101.07) --
	(267.18,101.15) --
	(269.13,101.24) --
	(271.09,101.36) --
	(273.05,101.50) --
	(275.00,101.63) --
	(276.96,101.75) --
	(278.92,101.88) --
	(280.87,102.01) --
	(282.83,102.17) --
	(284.79,102.35) --
	(286.74,102.56) --
	(288.70,102.82) --
	(290.66,103.12) --
	(292.61,103.48) --
	(294.57,103.89) --
	(296.52,104.34) --
	(298.48,104.83) --
	(300.44,105.36) --
	(302.39,105.91) --
	(304.35,106.49) --
	(306.31,107.09) --
	(308.26,107.71) --
	(310.22,108.34) --
	(312.18,108.99) --
	(314.13,109.66) --
	(316.09,110.40) --
	(318.05,111.21) --
	(320.00,112.08) --
	(321.96,112.97) --
	(323.92,113.88) --
	(325.87,114.79) --
	(327.83,115.68) --
	(329.79,116.53) --
	(331.74,117.32) --
	(333.70,118.05) --
	(335.66,118.74) --
	(337.61,119.43) --
	(339.57,120.13) --
	(341.53,120.82) --
	(343.48,121.51) --
	(345.44,122.19) --
	(347.40,122.86) --
	(349.35,123.51) --
	(351.31,124.13) --
	(353.27,124.74) --
	(355.22,125.32) --
	(357.18,125.88) --
	(359.14,126.43) --
	(361.09,126.96) --
	(363.05,127.48) --
	(365.00,127.98) --
	(366.96,128.47) --
	(368.92,128.94) --
	(370.87,129.40) --
	(372.83,129.85) --
	(374.79,130.28);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] (220.22,102.20) --
	(222.18,102.04) --
	(224.13,101.89) --
	(226.09,101.76) --
	(228.04,101.63) --
	(230.00,101.51) --
	(231.96,101.40) --
	(233.91,101.30) --
	(235.87,101.21) --
	(237.83,101.13) --
	(239.78,101.06) --
	(241.74,101.00) --
	(243.70,100.95) --
	(245.65,100.92) --
	(247.61,100.89) --
	(249.57,100.88) --
	(251.52,100.89) --
	(253.48,100.89) --
	(255.44,100.90) --
	(257.39,100.91) --
	(259.35,100.93) --
	(261.31,100.95) --
	(263.26,100.99) --
	(265.22,101.05) --
	(267.18,101.12) --
	(269.13,101.22) --
	(271.09,101.33) --
	(273.05,101.47) --
	(275.00,101.60) --
	(276.96,101.73) --
	(278.92,101.87) --
	(280.87,102.02) --
	(282.83,102.19) --
	(284.79,102.39) --
	(286.74,102.61) --
	(288.70,102.88) --
	(290.66,103.19) --
	(292.61,103.55) --
	(294.57,103.96) --
	(296.52,104.42) --
	(298.48,104.90) --
	(300.44,105.42) --
	(302.39,105.97) --
	(304.35,106.55) --
	(306.31,107.15) --
	(308.26,107.77) --
	(310.22,108.40) --
	(312.18,109.04) --
	(314.13,109.71) --
	(316.09,110.44) --
	(318.05,111.24) --
	(320.00,112.09) --
	(321.96,112.97) --
	(323.92,113.86) --
	(325.87,114.75) --
	(327.83,115.63) --
	(329.79,116.47) --
	(331.74,117.27) --
	(333.70,118.01) --
	(335.66,118.71) --
	(337.61,119.43) --
	(339.57,120.14) --
	(341.53,120.86) --
	(343.48,121.58) --
	(345.44,122.29) --
	(347.40,122.99) --
	(349.35,123.68) --
	(351.31,124.35) --
	(353.27,125.00) --
	(355.22,125.64) --
	(357.18,126.26) --
	(359.14,126.87) --
	(361.09,127.47) --
	(363.05,128.07) --
	(365.00,128.65) --
	(366.96,129.22) --
	(368.92,129.78) --
	(370.87,130.33) --
	(372.83,130.87) --
	(374.79,131.39);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.22, 31.02) rectangle (379.94,183.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27, 27.99) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27, 59.78) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27, 91.57) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,123.36) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (215.27,155.14) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47, 31.02) --
	(220.22, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47, 62.81) --
	(220.22, 62.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47, 94.60) --
	(220.22, 94.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,126.39) --
	(220.22,126.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.47,158.18) --
	(220.22,158.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.22, 28.27) --
	(220.22, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (245.98, 28.27) --
	(245.98, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.74, 28.27) --
	(271.74, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (297.50, 28.27) --
	(297.50, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (323.26, 28.27) --
	(323.26, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (349.03, 28.27) --
	(349.03, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (374.79, 28.27) --
	(374.79, 31.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (220.22, 20.01) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (245.98, 20.01) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (271.74, 20.01) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (297.50, 20.01) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (323.26, 20.01) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (349.03, 20.01) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (374.79, 20.01) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (300.08,  7.82) {Secondary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (205.80,107.31) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (220.22,191.89) {F. Generation interval};
\end{scope}
\begin{scope}
\path[clip] (385.44,  0.00) rectangle (578.16,206.49);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (385.44,  0.00) rectangle (578.16,206.49);
\end{scope}
\begin{scope}
\path[clip] (412.94, 31.02) rectangle (572.66,183.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (412.94, 31.02) rectangle (572.66,183.61);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94, 76.85) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24, 86.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55, 85.75) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85, 86.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16, 86.32) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46, 88.36) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77, 88.81) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07, 91.36) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38, 94.82) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,101.07) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,110.42) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,122.41) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,135.30) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,145.90) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,153.64) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,163.17) circle (  2.50);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94, 75.76) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24, 88.86) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55, 89.39) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85, 88.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16, 90.29) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46, 90.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77, 91.99) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07, 93.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38, 97.21) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,102.97) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,111.07) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,122.03) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,133.67) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,142.63) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,151.12) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,157.98) circle (  2.50);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94, 86.18) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24, 92.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55, 92.73) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85, 91.68) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16, 92.95) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46, 94.30) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77, 94.60) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07, 96.08) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38, 99.66) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,104.59) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,112.09) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,121.78) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,132.20) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,139.98) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,145.80) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,154.16) circle (  2.50);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (412.94, 89.43) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.24, 94.38) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.55, 96.42) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (443.85, 95.08) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (454.16, 96.19) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.46, 96.84) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.77, 97.29) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (485.07, 98.74) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (495.38,101.96) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (505.68,106.49) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (515.98,113.00) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (526.29,121.70) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (536.59,130.84) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (546.90,137.37) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (557.20,142.61) circle (  2.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (567.51,149.06) circle (  2.50);
\definecolor{drawColor}{RGB}{68,1,84}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94, 80.61) --
	(414.90, 80.97) --
	(416.85, 81.34) --
	(418.81, 81.70) --
	(420.76, 82.05) --
	(422.72, 82.39) --
	(424.68, 82.71) --
	(426.63, 83.02) --
	(428.59, 83.32) --
	(430.55, 83.62) --
	(432.50, 83.91) --
	(434.46, 84.19) --
	(436.42, 84.47) --
	(438.37, 84.75) --
	(440.33, 85.04) --
	(442.29, 85.33) --
	(444.24, 85.62) --
	(446.20, 85.87) --
	(448.16, 86.08) --
	(450.11, 86.26) --
	(452.07, 86.42) --
	(454.03, 86.56) --
	(455.98, 86.71) --
	(457.94, 86.87) --
	(459.90, 87.05) --
	(461.85, 87.28) --
	(463.81, 87.55) --
	(465.77, 87.86) --
	(467.72, 88.12) --
	(469.68, 88.33) --
	(471.64, 88.53) --
	(473.59, 88.74) --
	(475.55, 88.97) --
	(477.51, 89.26) --
	(479.46, 89.61) --
	(481.42, 90.07) --
	(483.38, 90.64) --
	(485.33, 91.35) --
	(487.29, 92.14) --
	(489.24, 92.97) --
	(491.20, 93.86) --
	(493.16, 94.79) --
	(495.11, 95.79) --
	(497.07, 96.84) --
	(499.03, 97.96) --
	(500.98, 99.15) --
	(502.94,100.42) --
	(504.90,101.76) --
	(506.85,103.20) --
	(508.81,104.78) --
	(510.77,106.49) --
	(512.72,108.29) --
	(514.68,110.18) --
	(516.64,112.11) --
	(518.59,114.08) --
	(520.55,116.06) --
	(522.51,118.02) --
	(524.46,119.95) --
	(526.42,121.81) --
	(528.38,123.67) --
	(530.33,125.57) --
	(532.29,127.50) --
	(534.25,129.47) --
	(536.20,131.47) --
	(538.16,133.50) --
	(540.12,135.54) --
	(542.07,137.59) --
	(544.03,139.65) --
	(545.99,141.71) --
	(547.94,143.78) --
	(549.90,145.86) --
	(551.86,147.95) --
	(553.81,150.06) --
	(555.77,152.18) --
	(557.72,154.32) --
	(559.68,156.48) --
	(561.64,158.65) --
	(563.59,160.84) --
	(565.55,163.04) --
	(567.51,165.25);
\definecolor{drawColor}{RGB}{49,104,142}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94, 80.55) --
	(414.90, 81.20) --
	(416.85, 81.85) --
	(418.81, 82.47) --
	(420.76, 83.07) --
	(422.72, 83.64) --
	(424.68, 84.19) --
	(426.63, 84.70) --
	(428.59, 85.20) --
	(430.55, 85.67) --
	(432.50, 86.13) --
	(434.46, 86.57) --
	(436.42, 87.00) --
	(438.37, 87.41) --
	(440.33, 87.82) --
	(442.29, 88.22) --
	(444.24, 88.61) --
	(446.20, 88.94) --
	(448.16, 89.21) --
	(450.11, 89.43) --
	(452.07, 89.60) --
	(454.03, 89.75) --
	(455.98, 89.90) --
	(457.94, 90.05) --
	(459.90, 90.22) --
	(461.85, 90.42) --
	(463.81, 90.68) --
	(465.77, 90.97) --
	(467.72, 91.21) --
	(469.68, 91.40) --
	(471.64, 91.57) --
	(473.59, 91.74) --
	(475.55, 91.93) --
	(477.51, 92.17) --
	(479.46, 92.47) --
	(481.42, 92.87) --
	(483.38, 93.37) --
	(485.33, 94.01) --
	(487.29, 94.72) --
	(489.24, 95.47) --
	(491.20, 96.27) --
	(493.16, 97.11) --
	(495.11, 98.00) --
	(497.07, 98.95) --
	(499.03, 99.96) --
	(500.98,101.03) --
	(502.94,102.16) --
	(504.90,103.37) --
	(506.85,104.67) --
	(508.81,106.10) --
	(510.77,107.64) --
	(512.72,109.27) --
	(514.68,110.98) --
	(516.64,112.73) --
	(518.59,114.51) --
	(520.55,116.30) --
	(522.51,118.07) --
	(524.46,119.81) --
	(526.42,121.49) --
	(528.38,123.16) --
	(530.33,124.87) --
	(532.29,126.61) --
	(534.25,128.39) --
	(536.20,130.19) --
	(538.16,132.01) --
	(540.12,133.84) --
	(542.07,135.68) --
	(544.03,137.53) --
	(545.99,139.38) --
	(547.94,141.23) --
	(549.90,143.09) --
	(551.86,144.96) --
	(553.81,146.84) --
	(555.77,148.74) --
	(557.72,150.65) --
	(559.68,152.57) --
	(561.64,154.50) --
	(563.59,156.45) --
	(565.55,158.41) --
	(567.51,160.38);
\definecolor{drawColor}{RGB}{53,183,121}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94, 88.80) --
	(414.90, 89.01) --
	(416.85, 89.22) --
	(418.81, 89.44) --
	(420.76, 89.65) --
	(422.72, 89.85) --
	(424.68, 90.05) --
	(426.63, 90.24) --
	(428.59, 90.43) --
	(430.55, 90.61) --
	(432.50, 90.79) --
	(434.46, 90.98) --
	(436.42, 91.16) --
	(438.37, 91.35) --
	(440.33, 91.54) --
	(442.29, 91.74) --
	(444.24, 91.94) --
	(446.20, 92.12) --
	(448.16, 92.27) --
	(450.11, 92.40) --
	(452.07, 92.52) --
	(454.03, 92.63) --
	(455.98, 92.74) --
	(457.94, 92.87) --
	(459.90, 93.02) --
	(461.85, 93.21) --
	(463.81, 93.43) --
	(465.77, 93.68) --
	(467.72, 93.89) --
	(469.68, 94.07) --
	(471.64, 94.24) --
	(473.59, 94.42) --
	(475.55, 94.63) --
	(477.51, 94.87) --
	(479.46, 95.18) --
	(481.42, 95.56) --
	(483.38, 96.03) --
	(485.33, 96.61) --
	(487.29, 97.25) --
	(489.24, 97.94) --
	(491.20, 98.66) --
	(493.16, 99.43) --
	(495.11,100.24) --
	(497.07,101.11) --
	(499.03,102.02) --
	(500.98,102.99) --
	(502.94,104.02) --
	(504.90,105.10) --
	(506.85,106.26) --
	(508.81,107.54) --
	(510.77,108.91) --
	(512.72,110.36) --
	(514.68,111.88) --
	(516.64,113.44) --
	(518.59,115.02) --
	(520.55,116.60) --
	(522.51,118.17) --
	(524.46,119.71) --
	(526.42,121.19) --
	(528.38,122.67) --
	(530.33,124.17) --
	(532.29,125.71) --
	(534.25,127.27) --
	(536.20,128.85) --
	(538.16,130.45) --
	(540.12,132.06) --
	(542.07,133.68) --
	(544.03,135.31) --
	(545.99,136.94) --
	(547.94,138.56) --
	(549.90,140.20) --
	(551.86,141.85) --
	(553.81,143.51) --
	(555.77,145.18) --
	(557.72,146.86) --
	(559.68,148.55) --
	(561.64,150.25) --
	(563.59,151.97) --
	(565.55,153.69) --
	(567.51,155.43);
\definecolor{drawColor}{RGB}{253,231,37}

\path[draw=drawColor,line width= 1.1pt,line join=round] (412.94, 91.57) --
	(414.90, 91.82) --
	(416.85, 92.08) --
	(418.81, 92.33) --
	(420.76, 92.57) --
	(422.72, 92.81) --
	(424.68, 93.03) --
	(426.63, 93.25) --
	(428.59, 93.45) --
	(430.55, 93.66) --
	(432.50, 93.86) --
	(434.46, 94.05) --
	(436.42, 94.25) --
	(438.37, 94.45) --
	(440.33, 94.64) --
	(442.29, 94.85) --
	(444.24, 95.05) --
	(446.20, 95.23) --
	(448.16, 95.38) --
	(450.11, 95.51) --
	(452.07, 95.62) --
	(454.03, 95.72) --
	(455.98, 95.83) --
	(457.94, 95.94) --
	(459.90, 96.08) --
	(461.85, 96.24) --
	(463.81, 96.44) --
	(465.77, 96.66) --
	(467.72, 96.84) --
	(469.68, 96.99) --
	(471.64, 97.13) --
	(473.59, 97.27) --
	(475.55, 97.44) --
	(477.51, 97.64) --
	(479.46, 97.90) --
	(481.42, 98.22) --
	(483.38, 98.63) --
	(485.33, 99.15) --
	(487.29, 99.73) --
	(489.24,100.35) --
	(491.20,101.01) --
	(493.16,101.71) --
	(495.11,102.46) --
	(497.07,103.25) --
	(499.03,104.09) --
	(500.98,104.97) --
	(502.94,105.89) --
	(504.90,106.87) --
	(506.85,107.90) --
	(508.81,109.04) --
	(510.77,110.26) --
	(512.72,111.56) --
	(514.68,112.91) --
	(516.64,114.29) --
	(518.59,115.70) --
	(520.55,117.10) --
	(522.51,118.49) --
	(524.46,119.84) --
	(526.42,121.14) --
	(528.38,122.42) --
	(530.33,123.74) --
	(532.29,125.07) --
	(534.25,126.43) --
	(536.20,127.80) --
	(538.16,129.18) --
	(540.12,130.57) --
	(542.07,131.97) --
	(544.03,133.37) --
	(545.99,134.76) --
	(547.94,136.15) --
	(549.90,137.55) --
	(551.86,138.95) --
	(553.81,140.36) --
	(555.77,141.77) --
	(557.72,143.20) --
	(559.68,144.63) --
	(561.64,146.06) --
	(563.59,147.50) --
	(565.55,148.95) --
	(567.51,150.40);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (412.94, 31.02) rectangle (572.66,183.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99, 27.99) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99, 59.78) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99, 91.57) {4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,123.36) {6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (407.99,155.14) {8};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19, 31.02) --
	(412.94, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19, 62.81) --
	(412.94, 62.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19, 94.60) --
	(412.94, 94.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,126.39) --
	(412.94,126.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.19,158.18) --
	(412.94,158.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.94, 28.27) --
	(412.94, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (438.70, 28.27) --
	(438.70, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.46, 28.27) --
	(464.46, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.22, 28.27) --
	(490.22, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (515.98, 28.27) --
	(515.98, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (541.75, 28.27) --
	(541.75, 31.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (567.51, 28.27) --
	(567.51, 31.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (412.94, 20.01) {-10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (438.70, 20.01) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (464.46, 20.01) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (490.22, 20.01) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (515.98, 20.01) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (541.75, 20.01) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (567.51, 20.01) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (492.80,  7.82) {Secondary cohort time (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (398.52,107.31) {Backward delay (days)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,578.16);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (412.94,191.89) {G. Serial interval};
\end{scope}
\end{tikzpicture}

\end{document}